### 2017-05-04 manuscript/usses_water_ms.Rmd
adds blank ms

---
layout: 12pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \doublespacing
  - \usepackage{rotating}
  - \usepackage{color, soul}
  - \usepackage[font={normalsize},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{sectsty}
  - \usepackage{bm,mathrsfs}
  - \usepackage{mathptmx} 
bibliography: ~/Dropbox/Bibliography/usses_water.bib
csl: components/ecology_letters.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: true
fontsize: 12pt
geometry: margin=1in
linkcolor: black
urlcolor: black
---

\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\definecolor{blue}{rgb}{0,0,0.7} 
\newcommand{\new}{\textcolor{blue}}

\begin{singlespace}

\begin{centering}

\Large{\textbf{Ecosystem and community resistance to six years of drought and deluge in the Sagebrush Steppe}}

\bigskip{} \bigskip{}

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

\normalsize{Andrew T. Tredennick\textsuperscript{1}, Peter B. Adler\textsuperscript{1}, and ...\textsuperscript{2}}

\bigskip{}

\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}}

\end{centering}

\vspace{3em}

Last compile: \today

\end{singlespace}

```{r caching, include=FALSE}
library("methods")
library("knitr")
basename <- "manuscript"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)

# PDF-based figures
opts_chunk$set(dev='pdf')
```

```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

opts_chunk$set(dev='pdf', dev.args=c(version="1.3"))

```

\newpage{}

\begin{abstract}
Summarize it here
\vspace{2em}
\end{abstract}

\setlength{\parindent}{5ex}

Introduction
------------
Start it here

References
----------

### 2017-05-04 manuscript/usses_water_ms.tex
adds blank ms

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            colorlinks=true,
            linkcolor=black,
            citecolor=Blue,
            urlcolor=black,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\usepackage{todonotes}
\doublespacing
\usepackage{rotating}
\usepackage{color, soul}
\usepackage[font={normalsize},labelfont={bf},labelsep=quad]{caption}
\usepackage{sectsty}
\usepackage{bm,mathrsfs}
\usepackage{mathptmx}

\begin{document}

\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}

\definecolor{blue}{rgb}{0,0,0.7} \newcommand{\new}{\textcolor{blue}}

\begin{singlespace}

\begin{centering}

\Large{\textbf{Ecosystem and community resistance to six years of drought and deluge in the Sagebrush Steppe}}

\bigskip{} \bigskip{}

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

\normalsize{Andrew T. Tredennick\textsuperscript{1}, Peter B. Adler\textsuperscript{1}, and ...\textsuperscript{2}}

\bigskip{}

\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}}

\end{centering}

\vspace{3em}

Last compile: \today

\end{singlespace}

\newpage{}

\begin{abstract}
Summarize it here
\vspace{2em}
\end{abstract}

\setlength{\parindent}{5ex}

\subsection{Introduction}\label{introduction}

Start it here

\subsection{References}\label{references}

\end{document}

### 2017-05-04 code/usses_stability.R
updates spp comp fig, adds cover data

@@ -28,6 +28,7 @@ library(RColorBrewer)
 library(lme4)
 library(codyn)
 library(vegan)
+library(gridExtra)
 
 
 
@@ -172,24 +173,54 @@ cover_community_matrix <- read.csv("../data/vital_rates/allrecords_cover_v24.csv
   ungroup()
 
 mycol <- RColorBrewer::brewer.pal(3,"Set2")
-png(paste0(figure_path,"sppcomp_nmds.png"), width=6, height = 4, units = "in", res = 120)
-par(mfrow=c(2,3))
+# 
+# par(mfrow=c(2,3))
+nmds_df <- {}
 for(doyr in unique(cover_community_matrix$year)){
   tmp <- filter(cover_community_matrix, year == doyr) %>% select(-year)
   torm <- as.numeric(which(colSums(tmp[3:ncol(tmp)], na.rm = T) == 0)) # find spp with NA in each treatment
   tmp <- tmp[,-(torm+2)] # add 2 because we lopped off two columns above
   tmp_rda <- metaMDS(tmp[3:ncol(tmp)], "bray")
-  tmp_dist <- vegdist(tmp[3:ncol(tmp)], method = "bray")
-  tmp_clust <- hclust(tmp_dist, method = "average")
-  mds.fig <- ordiplot(tmp_rda, type = "none", main = doyr)
-  # plot just the samples, colour by habitat, pch=19 means plot a circle
-  ordicluster(tmp_rda, tmp_clust, col = "gray")
-  points(mds.fig, "sites", pch = 19, col = mycol[1], select = tmp$Treatment == "Control")
-  points(mds.fig, "sites", pch = 19, col = mycol[2], select = tmp$Treatment == "Irrigation")
-  points(mds.fig, "sites", pch = 19, col = mycol[3], select = tmp$Treatment == "Drought")
-  points(mds.fig, "sites", pch = 1, col = "grey35", select = tmp$Treatment == "Control")
-  points(mds.fig, "sites", pch = 1, col = "grey35", select = tmp$Treatment == "Irrigation")
-  points(mds.fig, "sites", pch = 1, col = "grey35", select = tmp$Treatment == "Drought")
+  ## Try to plot in ggplot
+  nmds_points <- as.data.frame(tmp_rda$points)
+  nmds_points$Treatment <- tmp$Treatment
+  nmds_points$Year <- doyr
[…]

### 2017-05-05 manuscript/usses_water_ms.Rmd
adds beginning of methods section

@@
 
-\normalsize{Andrew T. Tredennick\textsuperscript{1}, Peter B. Adler\textsuperscript{1}, and ...\textsuperscript{2}}
+\normalsize{Andrew T. Tredennick\textsuperscript{1}, Peter B. Adler\textsuperscript{1}, and Andrew R. Kleinhesselink\textsuperscript{2}}
 
@@
 \textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}}
+\textit{\small{\textsuperscript{2}Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California xxxxx}}
 
@@
 
-\newpage{}
-
 \begin{abstract}
@@
 
+Methods
+-------
+### Study Area
+We conducted our precipitation manipulation experiment at the United States Sheep Experimental Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
+The vegetation is typical of high elevation sagebrush steppe.
+The plant community is dominated by the shrub *Artemesia tripartita* and three perennial bunchgrasses, *Pseudoroegneria spicata*, *Poa secunda*, and *Hesperostipa comata*.
+During the period of our experiment, average mean annual precipitation was xxx mm and mean monthly temperature ranged from -x$^{\circ}$C in January to x$^{\circ}$C in July.
+
+### Precipitation Experiment
+Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
+In 2007, we (well, one of us [P. Adler]) relocated 14 of the original quadrats, six of which were inside a large, permanent livestock exclosure.
+We use these six plots as control plots that have recieved no treatment, just ambient precipitation.
+In spring 2011, we (well, two of us [A. Kleinhesselink and P. Adler]) established 16 new 1 m$^2$ plots.
+We avoided areas on steep hill slopes, areas with greater than 20\% cover of bare rock, and areas with greater than 10\% cover of the shrubs *Purshia tridentata* and/or *Amelanchier utahensis*.
+We established the new plots in pairs and randomly assigned each plot in a pair to receive a "drought" or "irrigation" treatment.
+
+Drought and irrigation treatments were designed to decrease and increase the amount of ambient precipitation by 50\%, respectively.
+To achieve this, we used a system of rain-out shelters and automatic irrigation [@Gherardi2013].
+The rain-out shelters consisted of transparent acrylic shingles 1-1.5 m above the ground that covered an area of 2.5 by 2 m.
+The shingles intercepted approximately 50\% of incoming rainfall, which was channeled into 75 liter containers.
+Captured rainfall was then pumped out of the containers and sprayed on to the adjacent irrigation plot via two suspended sprinklers.
+Pumping was triggered by float switches once water levels reached about 20 liters.
+We disconnected the irrigation pumps each fall and reconnected them, often with difficulty, each spring.
+The rain-out shelters remained in place throughout the year.
+
+To make sure the treatments were having the desired effects, we monitored soil moisture in four of the drought-irrigation pairs using Decagon Devices (Pulman, Washington) 5TM and EC-5 soil moisture sensors.
+We installed four sensors in each plot, two at 5 cm soil depth and two at 25 cm soil depth.
+We also installed four sensors in areas nearby the four selected plot pairs to measure ambient soil moisture at the same depths.
+Soil moisture measurements were automatically logged every four hours.
+We coupled this temporally intensive soil moisture sampling with spatially extensive readings taken at six points within all 16 plots and associated ambient measurement areas.
+These snapshot data were collected on 06/06/2012, 04/29/2015, 05/07/2015, 06/09/2015, and 05/10/2016 using a handheld EC-5 sensor.
+
+### Data Collection
+#### Aboveground Net Primary Productivity
+We estimated aboveground net primary productivity (ANPP) using a radiometer to related ground reflectance to plant biomass [see @Byrne2011 for a review].
+
+Acknowledgments
+---------------
+This work was funded by the National Science Foundation through grants DEB-1353078 and DEB-1054040 to PBA, a Graduate Research Fellowship to ARK, and a Postdoctoral Resaerch Fellowship in Biology to ATT (DBI-1400370).
+Additional support came from the Utah Agricultural Experiment Station (journal paper xxxx).
+We thank the many summer research technicians who collected the data reported in this paper.
+
+\singlespace
+\setlength{\parindent}{0ex}
+
 References

### 2017-05-05 manuscript/usses_water_ms.tex
adds beginning of methods section

@@
 \end{singlespace}
-
-\newpage{}
 
@@
 
-\subsection{References}\label{references}
+\subsection{Methods}\label{methods}
+
+\subsubsection{Study Area}\label{study-area}
+
+We conducted our precipitation manipulation experiment at the United
+States Sheep Experimental Station (USSES) near Dubois, Idaho
+(44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m above sea level. The
+vegetation is typical of high elevation sagebrush steppe. The plant
+community is dominated by the shrub \emph{Artemesia tripartita} and
+three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa
+secunda}, and \emph{Hesperostipa comata}. During the period of our
+experiment, average mean annual precipitation was xxx mm and mean
+monthly temperature ranged from -x\(^{\circ}\)C in January to
+x\(^{\circ}\)C in July.
+
+\subsubsection{Precipitation Experiment}\label{precipitation-experiment}
+
+Between 1926 and 1932, range scientists at the USSES established 26
+permanent 1 m\(^2\) quadrats to track vegetation change over time. In
+2007, we (well, one of us {[}P. Adler{]}) relocated 14 of the original
+quadrats, six of which were inside a large, permanent livestock
+exclosure. We use these six plots as control plots that have recieved no
+treatment, just ambient precipitation. In spring 2011, we (well, two of
+us {[}A. Kleinhesselink and P. Adler{]}) established 16 new 1 m\(^2\)
+plots. We avoided areas on steep hill slopes, areas with greater than
+20\% cover of bare rock, and areas with greater than 10\% cover of the
+shrubs \emph{Purshia tridentata} and/or \emph{Amelanchier utahensis}. We
+established the new plots in pairs and randomly assigned each plot in a
+pair to receive a ``drought'' or ``irrigation'' treatment.
+
+Drought and irrigation treatments were designed to decrease and increase
+the amount of ambient precipitation by 50\%, respectively. To achieve
+this, we used a system of rain-out shelters and automatic irrigation
+(Gherardi \& Sala 2013). The rain-out shelters consisted of transparent
+acrylic shingles 1-1.5 m above the ground that covered an area of 2.5 by
+2 m. The shingles intercepted approximately 50\% of incoming rainfall,
+which was channeled into 75 liter containers. Captured rainfall was then
+pumped out of the containers and sprayed on to the adjacent irrigation
+plot via two suspended sprinklers. Pumping was triggered by float
+switches once water levels reached about 20 liters. We disconnected the
+irrigation pumps each fall and reconnected them, often with difficulty,
+each spring. The rain-out shelters remained in place throughout the
+year.
+
+To make sure the treatments were having the desired effects, we
+monitored soil moisture in four of the drought-irrigation pairs using
+Decagon Devices (Pulman, Washington) 5TM and EC-5 soil moisture sensors.
+We installed four sensors in each plot, two at 5 cm soil depth and two
+at 25 cm soil depth. We also installed four sensors in areas nearby the
+four selected plot pairs to measure ambient soil moisture at the same
+depths. Soil moisture measurements were automatically logged every four
+hours. We coupled this temporally intensive soil moisture sampling with
+spatially extensive readings taken at six points within all 16 plots and
+associated ambient measurement areas. These snapshot data were collected
+on 06/06/2012, 04/29/2015, 05/07/2015, 06/09/2015, and 05/10/2016 using
+a handheld EC-5 sensor.
+
+\subsubsection{Data Collection}\label{data-collection}
+
+\paragraph{Aboveground Net Primary
+Productivity}\label{aboveground-net-primary-productivity}
+
+We estimated aboveground net primary productivity (ANPP) using a
+radiometer to related ground reflectance to plant biomass (see Byrne et
+al. 2011 for a review).
+
+\subsection{Acknowledgments}\label{acknowledgments}
+
+This work was funded by the National Science Foundation through grants
+DEB-1353078 and DEB-1054040 to PBA, a Graduate Research Fellowship to
+ARK, and a Postdoctoral Resaerch Fellowship in Biology to ATT
+(DBI-1400370). Additional support came from the Utah Agricultural
+Experiment Station (journal paper xxxx). We thank the many summer
+research technicians who collected the data reported in this paper.
+
+\singlespace
+\setlength{\parindent}{0ex}
+
+\subsection*{References}\label{references}
+\addcontentsline{toc}{subsection}{References}
+
+\hypertarget{refs}{}
+\hypertarget{ref-Byrne2011}{}
+Byrne, K.M., Lauenroth, W.K., Adler, P.B. \& Byrne, C.M. (2011).
+Estimating Aboveground Net Primary Production in Grasslands: A
+Comparison of Nondestructive Methods. \emph{Rangeland Ecology \&
+Management}, 64, 498--505.
+
+\hypertarget{ref-Gherardi2013}{}
+Gherardi, L.A. \& Sala, O.E. (2013). Automated rainfall manipulation
+system: a reliable and inexpensive tool for ecologists.
+\emph{Ecosphere}, 4, 1--10.
 

### 2017-05-05 code/calibrate_radiometer_by_year.R

@@ -30,6 +30,7 @@ out_path       <- "../data/estimated_biomass/"
 library(tidyverse) # all sorts of data wrangling
 library(dplyr)     # data summarizing/wrangling tools
 library(broom)     # tidys up model output
+library(stringr)
 library(ggthemes)
 
 
@@ -303,6 +304,7 @@ ggsave("../figures/estimated_biomass.png", width = 6, height = 4, units = "in", 
 
 
 biomass_yr_trt_summ <- permanent_quad_biomass %>%
+  filter(!str_detect(quadname, 'P1|P7')) %>%
   group_by(Treatment,year) %>%
   summarise(mean_biomass = mean(biomass_grams_est),
             sd_biomass = sd(biomass_grams_est)) %>%
@@ -324,4 +326,8 @@ ggplot(biomass_yr_trt_summ, aes(x=year, y=mean_biomass, color=Treatment))+
 ggsave("../figures/anpp_trt_trend.png", width = 4, height = 3, units = "in", dpi = 120)
 
 
-
+# Get sample sizes
+samp_size <- permanent_quad_biomass %>%
+  filter(!str_detect(quadname, 'P1|P7')) %>%
+  group_by(Treatment,year) %>%
+  summarise(num_plots = length(biomass_grams_est))

### 2017-05-06 manuscript/usses_water_ms.Rmd
adds more to methods

@@
 #### Aboveground Net Primary Productivity
-We estimated aboveground net primary productivity (ANPP) using a radiometer to related ground reflectance to plant biomass [see @Byrne2011 for a review].
+We estimated aboveground net primary productivity (ANPP) using a radiometer to relate ground reflectance to plant biomass [see @Byrne2011 for a review].
+We recorded ground reflectance at four wavelengths, two associated with red reflectance (xxx and xxx) and two associated with near-infrared reflectance (xxx and xxx).
+At each plot in each year, we took four readings of ground reflectances at the above wavelengths.
+We also took readings in ten calibration plots adjacent to the experimental site, in which we harvested all aboveground biomass, dried it to a constant weight at 60$^{\circ}$C, and weighed it to estimate ANPP.
+
+For each plot and year, we averaged the four readings for each wavelength and then calculated NDVI using the MODIS and AVHRR algorithms.
+To convert NDVI to ANPP we regressed NDVI against the dry biomass weight from the ten calibration plots.
+We fit regressions to MODIS-based NDVI and AVHRR-based NDVI for each year and retained the regression with the better fit.
+Using the best regression equation for each year, we predicted ANPP.
+
+#### Species Composition
+Species composition data came from annual census maps for each plot made using a pantograph [@Hill1920].
+The maps record the spatial location and size of each individual plant.
+Using those annual maps, we aggregated over individuals to calculate total basal cover for each species in each plot.
+
+### Data Analysis
+####  ANPP and Ecosystem Stability
+
+####  Species Synchrony and Species Composition
+
+Results
+-------
+
+Discussion
+----------
 

### 2017-05-06 manuscript/usses_water_ms.tex
adds more to methods

@@
 We estimated aboveground net primary productivity (ANPP) using a
-radiometer to related ground reflectance to plant biomass (see Byrne et
-al. 2011 for a review).
+radiometer to relate ground reflectance to plant biomass (see Byrne et
+al. 2011 for a review). We recorded ground reflectance at four
+wavelengths, two associated with red reflectance (xxx and xxx) and two
+associated with near-infrared reflectance (xxx and xxx). At each plot in
+each year, we took four readings of ground reflectances at the above
+wavelengths. We also took readings in ten calibration plots adjacent to
+the experimental site, in which we harvested all aboveground biomass,
+dried it to a constant weight at 60\(^{\circ}\)C, and weighed it to
+estimate ANPP.
+
+For each plot and year, we averaged the four readings for each
+wavelength and then calculated NDVI using the MODIS and AVHRR
+algorithms. To convert NDVI to ANPP we regressed NDVI against the dry
+biomass weight from the ten calibration plots. We fit regressions to
+MODIS-based NDVI and AVHRR-based NDVI for each year and retained the
+regression with the better fit. Using the best regression equation for
+each year, we predicted ANPP.
+
+\paragraph{Species Composition}\label{species-composition}
+
+Species composition data came from annual census maps for each plot made
+using a pantograph (Hill 1920). The maps record the spatial location and
+size of each individual plant. Using those annual maps, we aggregated
+over individuals to calculate total basal cover for each species in each
+plot.
+
+\subsubsection{Data Analysis}\label{data-analysis}
+
+\paragraph{ANPP and Ecosystem
+Stability}\label{anpp-and-ecosystem-stability}
+
+\paragraph{Species Synchrony and Species
+Composition}\label{species-synchrony-and-species-composition}
+
+\subsection{Results}\label{results}
+
+\subsection{Discussion}\label{discussion}
 
@@
 
+\hypertarget{ref-Hill1920}{}
+Hill, R.R. (1920). Charting Quadrats with a Pantograph. \emph{Ecology},
+1, 270--273.
+
 \end{document}

### 2017-05-11 manuscript/usses_water_ms.Rmd
adds some analysis ANOVA details

@@
     fig_caption: true
-    keep_tex: true
+    keep_tex: false
 fontsize: 12pt
@@
 
-\Large{\textbf{Ecosystem and community resistance to six years of drought and deluge in the Sagebrush Steppe}}
+\Large{\textbf{Ecosystem and community resistance to five years of drought and deluge in the Sagebrush Steppe}}
 
@@
 To achieve this, we used a system of rain-out shelters and automatic irrigation [@Gherardi2013].
-The rain-out shelters consisted of transparent acrylic shingles 1-1.5 m above the ground that covered an area of 2.5 by 2 m.
+The rain-out shelters consisted of transparent acrylic shingles 1-1.5 m above the ground that covered an area of $2.5\times2$ m.
 The shingles intercepted approximately 50\% of incoming rainfall, which was channeled into 75 liter containers.
@@
 ####  ANPP and Ecosystem Stability
+We used ANOVA with treatment, year, and treatment$\times$year terms to test the effects of the treatments on ANPP over the course of the five year experiment.
+To test for within-year differences of ANPP among treatments, we fit ANOVAs independently for each year with treatment terms only.
+We used the \texttt{lm()} function in \texttt{R} [v3.3.2, @R2016] to fit the models, and the \texttt{anova()} function to extract sums-of-squares tables.
 

### 2017-05-11 manuscript/usses_water_ms.Rmd
adds some more methods

@@
 
+We calculated ecosystem stability as the temporal mean of ANPP divided by its temporal standarad deviation $\left( \overline{\text{ANPP}} / \sigma(\text{ANPP}) \right)$.
+Ecosystem stability was calculated for each plot.
+We compared ecosystem stability among treatments using ANOVA (same \texttt{R} functions as above).
+
 ####  Species Synchrony and Species Composition
+We calculated the synchrony of species fluctuations through time using basal area cover data and individual density data.
+For both data types we calculated the synchrony metric described by @Loreau2008a, which ranges from 0 (perfect ascynchrony) to 1 (perfect synchrony).
+Before calculating synchrony, we averaged cover and density over plots within treatments for each species.
+We used the \texttt{synchrony()} function in the \texttt{R} package \texttt{codyn} [@Hallett2016] to calculate species synchrony.
+
+To see if community composition differed among treatments through time, we used non-dimensional multivariate scaling (NMDS) based on Bray-Curtis distances.
+For each year of the experiment, we first calculated Bray-Curtis distances among all plots, and then extracted those distances for use in the NMDS.
+We plotted the first two axes of NMDS scores to see if community composition overlapped, or not, among treatments in each year.
+We used functions in the \texttt{R} package \texttt{vegan} [@Oksanen2016] to calculate Bray-Curtis distances and the run the NMDS analysis.
 

### 2017-05-11 manuscript/usses_water_ms.Rmd
adds little abstract, for context

@@
 \begin{abstract}
-Summarize it here
+Precipitation is predicted to become more variable in the western U.S., meaning years of above and below average precipitation will become more common.
+Such periods of drought and deluge could become major drivers of plant community dynamics and ecosystem functioning.
+We used drought and irrigation treatments (50\% decrease/increase) to see how a sagebrush steppe plant community in Idaho will respond to future climate changes.
+The ecosystem was suprisingly resistant to both drought and irrigation.
+Aboveground net primary productivity (ANPP) and its temporal stability did no differ among treatments.
+The similarity of ecosystem functioning was not due to shifts at the plant community level, where species compositions and species synchrony, an important driver of ecosystem stability, among treatments were similar.
+However, ANPP does appear to be diverging among treatments with time, even though the effect is not significant.
+At least in the short-term, this sagebrush steppe ecosystem is resistant to increases and decreases in annual precipitation.
+
 \vspace{2em}
+
+\emph{Key words: drought, aboveground net primary productivity, ecosystem stability, climate change, species synchrony}
 \end{abstract}

### 2017-05-11 manuscript/usses_water_ms.Rmd
udpates refs

@@
 bibliography: ~/Dropbox/Bibliography/usses_water.bib
-csl: components/ecology_letters.csl
+csl: components/ecology.csl
 

### 2017-05-12 code/anpp_anova.R
adds within year ANOVA results table

@@ -20,7 +20,8 @@ library(tidyverse) # Data munging
 library(dplyr)     # Data summarizing
 library(broom)     # Working with model output
 library(stringr)   # Working with strings
-library(car)
+library(car)       # Type II ANOVA function
+library(xtable)    # LaTeX tables
 
 
 
@@ -80,4 +81,31 @@ capture.output(
                  cat("\n\n\n\n")
                },
                file = "../results/within_year_anovas.txt")
+
+
+##  Run again for LaTeX table
+stats_out <- {}
+for(doyear in years){
+  tmpd <- filter(anpp_data, year==doyear)
+  drought_mod <- lm(log(biomass_grams_est) ~ Treatment, 
+                    data=filter(tmpd, Treatment!="Irrigation"))
+  irrigat_mod <- lm(log(biomass_grams_est) ~ Treatment, 
+                    data=filter(tmpd, Treatment!="Drought"))
+  
+  drt <- broom::tidy(car::Anova(drought_mod))
+  irr <- broom::tidy(car::Anova(irrigat_mod))
+  
+  irr$Effect <- "Irrigation"
+  drt$Effect <- "Drought"
+  tmpout <- rbind(irr,drt)
+  tmpout <- filter(tmpout, term=="Treatment")
+  tmpout$df <- paste0(as.character(tmpout$df),", 12")
+  tmpout$Year <- doyear
+  tmpout <- select(tmpout, -c(term,sumsq))
+  tmpout <- tmpout[,c("Year","Effect","df","statistic","p.value")]
+  colnames(tmpout) <- c("Year","Effect","df","F","P")
+  stats_out <- rbind(stats_out, tmpout)
+}
+
+
 

### 2017-05-12 code/anpp_anova.R
adds results output to anova script

@@ -21,7 +21,6 @@ library(dplyr)     # Data summarizing
 library(broom)     # Working with model output
 library(stringr)   # Working with strings
 library(car)       # Type II ANOVA function
-library(xtable)    # LaTeX tables
 
 
 
@@ -107,5 +106,7 @@ for(doyear in years){
   stats_out <- rbind(stats_out, tmpout)
 }
 
+# Save output dataframe for LaTeX table making in R Markdown
+saveRDS(object = stats_out, "../results/within_year_anova_table.RDS")
 
 

### 2017-05-12 manuscript/usses_water_ms.Rmd
adds inyear ANOVA table to ms

@@
   - \usepackage{color, soul}
-  - \usepackage[font={normalsize},labelfont={bf},labelsep=quad]{caption}
   - \usepackage{sectsty}
@@
 
+\newpage{}
+
+TABLES
+------
+
+```{r table, include=TRUE, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
+library(xtable)
+# Read in within-year ANOVA results table
+inyear_anova <- readRDS("../results/within_year_anova_table.RDS")
+
+inyear_caption <- "ANOVA for treatment effects on ANPP for each year of the experiment."
+print(xtable(inyear_anova, caption = inyear_caption),
+      caption.placement="top",
+      include.rownames = F, 
+      sanitize.colnames.function = identity,
+      comment=FALSE,
+      size="normalsize")
+```
+
+\newpage{}
+
 References

### 2017-05-12 manuscript/usses_water_ms.Rmd
getting some citations right

@@
 To test for within-year differences of ANPP among treatments, we fit ANOVAs independently for each year with treatment terms only.
-We used the \texttt{lm()} function in \texttt{R} [v3.3.2, @R2016] to fit the models, and the \texttt{anova()} function to extract sums-of-squares tables.
+We used the \texttt{lm()} function in \texttt{R} [v3.3.2, @R2016] to fit the models, and the \texttt{Anova()} function from the \texttt{car} package [@Fox2011] to extract sums-of-squares tables with Type II tests.
 
@@
 inyear_caption <- "ANOVA for treatment effects on ANPP for each year of the experiment."
-print(xtable(inyear_anova, caption = inyear_caption),
+print(xtable(inyear_anova, caption = inyear_caption, digits = 2),
       caption.placement="top",
@@
       sanitize.colnames.function = identity,
-      comment=FALSE,
-      size="normalsize")
+      comment = FALSE,
+      size = "normalsize")
 ```

### 2017-05-13 manuscript/usses_water_ms.Rmd
trivial update to figure 1

@@
 
-TABLES
+Tables
 ------
@@
 
+Figures
+-------
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{../figures/histppt_and_anpptrend.png}
+  \caption{(A) Histogram of historical precipitation from 1926-2016, with the years of the experiment overlaid as colored points.
+``Drought'' treatments (squares) are shown as a 50\% reduction from the observed precipitation.
+``Irrigation'' treatment precipitation levels are not shown.
+(B) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.}
+\end{figure}
+
+\newpage{}
+
 References

### 2017-05-13 code/look_at_weather.R

@@ -47,7 +47,9 @@ g1 <- ggplot(weather, aes(x=ppt1))+
   xlab("Growing Season Precipitation (mm)")+
   ggtitle("A")+
   theme_few()+
-  theme(legend.position = c(0.8, 0.6),legend.key.size = unit(1,"pt"),legend.title = element_text(size=10),
+  theme(legend.position = c(0.8, 0.6),
+        legend.key.size = unit(1,"pt"),
+        legend.title = element_text(size=10),
         legend.text = element_text(size = 8),
         legend.key.height = unit(0.8,"line"))+
   guides(fill = guide_legend(override.aes = list(color = brewer.pal(5,"Set1"),size=1)),
@@ -85,9 +87,10 @@ g2 <- ggplot(biomass_yr_trt_summ, aes(x=year, y=mean_biomass, color=Treatment))+
   geom_point(color="white", size=3)+
   geom_point()+
   geom_point(color="grey35", shape=1)+
+ # annotate("text",x=2015,y=250,label="*")+
   scale_color_brewer(palette = "Set2", name="Treatment")+
   scale_x_continuous(breaks=c(2011:2016))+
-  scale_y_continuous(breaks=seq(50,350,50))+
+  scale_y_continuous(expand=c(0,0),limits=c(10,400),breaks=seq(50,350,50))+
   ylab(expression(paste("Estimated ANPP (g ", m^-2,")")))+
   xlab("Year")+
   ggtitle("B")+

### 2017-05-14 code/usses_stability.R
adds stability ANOVA

@@ -29,6 +29,7 @@ library(lme4)
 library(codyn)
 library(vegan)
 library(gridExtra)
+library(car)
 
 
 
@@ -98,14 +99,27 @@ ggplot(stability_richness, aes(x=avg_richness, y=stability, color=Treatment))+
   theme_few()
 ggsave(paste0(figure_path,"stability_richness.png"), height = 3, width = 4.5, units = "in", dpi = 120)
 
-ggplot(stability_richness, aes(y = stability, x = Treatment, color=Treatment))+
-  geom_jitter(width=0.1)+
-  geom_boxplot(fill=NA,outlier.color = NA, width=0.25)+
+ggplot(stability_richness, aes(y = stability, x = Treatment, fill=Treatment))+
+  geom_boxplot(outlier.color = NA, width=0.2, color="grey24",alpha=0.5)+
+  geom_jitter(width=0.1,shape=21,color="grey35")+
   scale_color_brewer(palette = "Set2")+
+  scale_fill_brewer(palette = "Set2")+
+  scale_y_continuous(breaks=seq(1.4,2.6,0.2))+
   ylab(expression(paste("Stability of ANPP (", mu/sigma,")")))+
-  guides(color=F)+
+  guides(fill=F)+
   theme_few()
 ggsave(paste0(figure_path,"stability_treatment.png"), height = 3, width = 3, units = "in", dpi = 120)
+
+
+
+####
+####  ANOVA FOR STABILITY ----
+####
+stability_mod <- lm(stability ~ Treatment, data=stability_richness)
+capture.output(
+  print(car::Anova(stability_mod)),
+  file = "../results/stability_anova.txt"
+)
 
 
 

### 2017-05-14 code/look_at_weather.R
trivia

@@ -90,7 +90,7 @@ g2 <- ggplot(biomass_yr_trt_summ, aes(x=year, y=mean_biomass, color=Treatment))+
  # annotate("text",x=2015,y=250,label="*")+
   scale_color_brewer(palette = "Set2", name="Treatment")+
   scale_x_continuous(breaks=c(2011:2016))+
-  scale_y_continuous(expand=c(0,0),limits=c(10,400),breaks=seq(50,350,50))+
+  scale_y_continuous(breaks=seq(50,350,50))+
   ylab(expression(paste("Estimated ANPP (g ", m^-2,")")))+
   xlab("Year")+
   ggtitle("B")+

### 2017-05-14 code/usses_stability.R
updates nmds figure

@@ -203,7 +203,7 @@ for(doyr in unique(cover_community_matrix$year)){
   nmds_df <- rbind(nmds_df, nmds_points)
 }
 
-ggplot(nmds_df, aes(x=MDS1, y=MDS2, color=Treatment))+
+comp1 <- ggplot(nmds_df, aes(x=MDS1, y=MDS2, color=Treatment))+
   geom_point()+
   geom_point(color="grey35",shape=1)+
   scale_color_brewer(palette = "Set2")+
@@ -212,6 +212,7 @@ ggplot(nmds_df, aes(x=MDS1, y=MDS2, color=Treatment))+
   ylab("NMDS 2")+
   xlab("NMDS 1")+
   facet_wrap("Year")+
+  ggtitle("A. Cover data")+
   theme_few()
 ggsave(paste0(figure_path,"sppcomp_nmds.png"), width=6, height = 4, units = "in", dpi = 120)
 
@@ -246,7 +247,7 @@ for(doyr in unique(density_community_matrix$year)){
   nmds_df <- rbind(nmds_df, nmds_points)
 }
 
-ggplot(nmds_df, aes(x=MDS1, y=MDS2, color=Treatment))+
+comp2 <- ggplot(nmds_df, aes(x=MDS1, y=MDS2, color=Treatment))+
   geom_point()+
   geom_point(color="grey35",shape=1)+
   scale_color_brewer(palette = "Set2")+
@@ -255,8 +256,10 @@ ggplot(nmds_df, aes(x=MDS1, y=MDS2, color=Treatment))+
   ylab("NMDS 2")+
   xlab("NMDS 1")+
   facet_wrap("Year")+
+  ggtitle("B. Density data")+
   theme_few()
 ggsave(paste0(figure_path,"sppcomp_density_nmds.png"), width=6, height = 4, units = "in", dpi = 120)
 
-
+gout <- grid.arrange(comp1,comp2,nrow=2)
+ggsave("../figures/nmds_comp_combined.png", plot = gout, width = 6, height = 8, units="in", dpi=120)
 

### 2017-05-15 manuscript/usses_water_ms.Rmd
adds nmds figure
trivia
starts an intro

@@
 
-\Large{\textbf{Ecosystem and community resistance to five years of drought and deluge in the Sagebrush Steppe}}
+\Large{\textbf{Ecosystem and community resistance to five years of drought and deluge in a sagebrush tteppe}}
 
@@
 
-\emph{Key words: drought, aboveground net primary productivity, ecosystem stability, climate change, species synchrony}
+\emph{Key words: drought, aboveground net primary productivity, ecosystem stability, climate change, species synchrony, sagebrush steppe}
 \end{abstract}
@@
 ------------
-Start it here
+Chronic alterations to resource availability should cause community composition to shift over time.
+Species' relative abundances are expected to re-order and, eventually, some species will be lost from the local community and new species will be gained [@Smith2009; @Avolio2015].
+Ecosystem functioning, such as productivity and stability, may remain relatively constant (CITATION) or shift to a new mean state (CITATION) depending on the trait-space of the new community.
+
+Different ecosystems are sensitive to different global changes.
+For example, @Flombaum2017 found aboveground net primary productivity (ANPP) in a Patagonian steppe was less sensitive to drought than a North American tallgrass prairie, relative to other global changes.
+Taken together with other results, @Flombaum2017 conclude that ecosystem sensitivity to a resource alteration is inversly related to the natural variation of the resource.
+In the above example, precipitation is much more variable in the Patagonian steppe than the North American tallgrass prairie.
+
+This leaves us with the following qualitative understanding.
+Chronic resource alteration should induce shifts in community composition and, maybe, ecosystem functioning, but only if the focal ecosystem is predisposed to be sensitive to the particular resource that is altered.
+While a crude understanding, this does allow us to make testable predictions when confronted with an ecosystem--resource pairing.
 
@@
 
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=5in]{../figures/nmds_comp_combined.png}
+  \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot, colored by treatment.
+(A) NMDS results using basal cover data; (B) NMDS results using individual density data.
+2011 is a pre-treatment year.}
+\end{figure}
+
+\newpage{}
+
 References

### 2017-05-16 manuscript/usses_water_ms.Rmd
trivia
more intro writing

@@
 
-\normalsize{Andrew T. Tredennick\textsuperscript{1}, Peter B. Adler\textsuperscript{1}, and Andrew R. Kleinhesselink\textsuperscript{2}}
+\normalsize{Andrew T. Tredennick\textsuperscript{1}, Andrew R. Kleinhesselink\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}}
 
@@
 While a crude understanding, this does allow us to make testable predictions when confronted with an ecosystem--resource pairing.
+
+In many areas of the western United States, precipitation is likely to become more variable.
+This will result in swings back-and-forth between multiyear periods of above and above average precipitation.
+An obvious question is, how will semiarid ecosystems respond to periods of drought and deluge?
+A naive expectation is that alterations in precipitation, and subsequently available soil moisture, should cause changes in plant community composition and increase or decrease ANPP dependending on the direction of precipitation change.
+But our emerging qualitative understanding suggests an alternative expectation: altering soil moisture may have little to no effect on plant community composition and ecosystem functioning.
+This expectation stems from the fact that precipitation is already a variable resource in semiarid systems, meaning the plant community is not sensitive to realistic increases/decreases in precipitation.
+
+Here we test our qualitative understanding of how altering precipitation will impact a sagebrush steppe ecosystem by imposing drought and irrigation for five years.
+In particular, we test the following competing predictions:
+
+\begin{description}
+\item [P1.] Altering precipitation will favor certain species over others, resulting in a shift in the plant community composition.
+ANPP will increase (irrigation) or decrease (drought), and the treatment effects will get stronger over time (a treatment$\times$year interaction).
+
+\item[P2.] Altering precipitation will not favor certain species and plant community composition will not change.
+ANPP will increase (irrigation) or decrease (drought), and the treatment effects will get stronger over time.
+\end{description}
+
+\noindent{}Note that both predictions assume ANPP will respond to precipitation alteration.
+Under **P1**, ANPP response stems directly from plant community shifts to species able to take advantage of the new precipitation regime.
+Under **P2**, ANPP response stems from all species responding similarly to changes in precipitation.
 

### 2017-05-17 code/community_dynamics.R
adds rank clocks

@@ -1,0 +1,250 @@
+##  usses_stability.R: R script to analyze the temporal stability of plots at
+##    USSES in Idaho. Plots received irrigation/drought treatments starting
+##    in 2011.
+##
+##  Author:       Andrew Tredennick
+##  Date created: April 24, 2017
+##
+##  UPDATES  ###################################################################
+##
+##
+##  ############################################################################
+
+##  Clear the workspace
+rm(list=ls(all.names = TRUE))
+
+
+
+####
+####  LOAD PACKAGES ----
+####
+library(tidyverse)
+library(dplyr)
+library(stringr)
+library(ggthemes)
+library(ggalt)
+library(broom)
+library(RColorBrewer)
+library(lme4)
+library(codyn)
+library(vegan)
+library(gridExtra)
+library(car)
+
+
+
+####
+####  SET DIRECTORIES; READ IN DATA ----
+####
+# Set working dir to source file location
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only works in RStudio
+
+# Set paths
+biomass_path <- "../data/estimated_biomass/"
+weather_path <- "../data/weather/"
+output_path  <- "../output/"
+figure_path  <- "../figures/"
+
+# Read in data
+plot_biomass <- readRDS(paste0(biomass_path,"permanent_plots_estimated_biomass.RDS"))
+weather_data <- read.csv(paste0(weather_path,"ClimateIPM.csv"))
+plot_info    <- read.csv("../data/estimated_biomass/quad_info.csv")
+
+
+
+####
+####  CALCULATE SYNCHRONY OF SPECIES COVER ----
+####
+cover_species <- read.csv("../data/vital_rates/allrecords_cover_v24.csv") %>%
+  group_by(quad, year, species) %>%
+  summarise(total_area = sum(area)) %>%
+  left_join(plot_info) %>%
+  filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
+  filter(!str_detect(QuadName, 'P1|P7')) %>%
+  group_by(year, species, Treatment) %>%
+  summarise(avg_area = mean(total_area)) %>%
+  ungroup()
+
+ggplot(cover_species, aes(x=year, y=avg_area, col=species))+
[…]

### 2017-05-19 manuscript/usses_water_ms.Rmd
adds rank clocks
trivia

@@
 
-\Large{\textbf{Ecosystem and community resistance to five years of drought and deluge in a sagebrush tteppe}}
+\Large{\textbf{Ecosystem and community resistance to five years of drought and deluge in a sagebrush steppe}}
 
@@
 
-\normalsize{Andrew T. Tredennick\textsuperscript{1}, Andrew R. Kleinhesselink\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}}
+\normalsize{Andrew T. Tredennick\textsuperscript{1}, Andrew R. Kleinhesselink\textsuperscript{2}, Bret Taylor\textsuperscript{3}, and Peter B. Adler\textsuperscript{1}}
 
@@
 
-\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}}
-\textit{\small{\textsuperscript{2}Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California xxxxx}}
+\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}} \\
+\textit{\small{\textsuperscript{2}Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California xxxxx}} \\
+\textit{\small{\textsuperscript{3}U.S. Sheep Experiment Station, Range Sheep Production Efficiency Research Unit, USDA-ARS, Dubois, Idaho 83423}}
 
@@
 
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=5in]{../figures/rank_clocks_cover.png}
+  \caption{Rank clocks for cover by treatment. (A) All species. (B) Dominant species.}
+\end{figure}
+
+\newpage{}
+
 References

### 2017-05-19 code/anpp_model.R

@@ -18,10 +18,10 @@ setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only for RStudio
 ####
 library(tidyverse)    # Data munging
 library(dplyr)        # Data summarizing
-library(ggthemes)
+library(ggthemes)     # Pleasing ggplot themes
 library(stringr)      # Working with strings
 library(rjags)        # For MCMC
-library(coda)         # For summarizing MCMC output
+library(coda)         # For collecting/summarizing MCMC output
 
 
 
@@ -144,8 +144,10 @@ ggplot(all_params, aes(x=effect,y=median))+
   ylab("Posterior Estimate")+
   xlab("")+
   facet_wrap(~treatment)+
-  theme_few()
-ggsave("../figures/anpp_posterior_quants.png",width = 4, height = 3, units = "in", dpi = 200)
+  theme_few()+
+  theme(axis.title = element_text(size=9),
+        axis.text = element_text(size=7))
+ggsave("../figures/anpp_posterior_quants.png",width = 4, height = 2, units = "in", dpi = 120)
 
 ggplot(anpp_data, aes(x=ppt1,y=anpp,color=Treatment))+
   geom_jitter(size = 2,width = 2, alpha=0.5, shape=21, color="grey35",aes(fill=Treatment))+

### 2017-05-23 code/anpp_repmeas_model.R
adds stan repeated measures attempt and results

@@ -1,0 +1,159 @@
+##  anpp_anova.R: Script to run ANOVA analysis to test for treatment effects
+##  on ANPP. Runs ANOVA for entirety of experiment and independent tests
+##  within years.
+##
+##  Author: Andrew Tredennick
+##  Date created: May 23, 2017
+
+##  Clear everything
+rm(list=ls(all.names = T))
+
+##  Set working directory to source file location
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only for RStudio
+
+
+
+####
+####  LOAD LIBRARIES ----
+####
+library(tidyverse)    # Data munging
+library(dplyr)        # Data summarizing
+library(ggthemes)     # Pleasing ggplot themes
+library(stringr)      # Working with strings
+library(rstan)        # For MCMC
+
+
+
+
+####
+####  READ IN AND EXTRACT EXPERIMENT ANPP DATA ----
+####
+data_path <- "../data/estimated_biomass/"
+anpp_fname <- "permanent_plots_estimated_biomass.RDS"
+permanent_quad_biomass <- readRDS(paste0(data_path,anpp_fname))
+
+weather <- read.csv("../data/weather/ClimateIPM.csv") %>%
+  select(-pptLag,-ppt2,-TmeanSpr1,-TmeanSpr2)
+
+anpp_data <- permanent_quad_biomass %>% 
+  filter(Treatment %in% c("Control","Drought","Irrigation")) %>%
+  filter(!str_detect(quadname, 'P1|P7')) %>%
+  filter(year > 2011) %>%
+  rename(anpp = biomass_grams_est) %>%
+  left_join(weather, by = "year") %>%
+  select(-QuadName,-quad,-Grazing,-paddock,-ndvi) %>%
+  mutate(ppt1_scaled = as.numeric(scale(ppt1)),
+         year_id = year - 2011)
+
+drought_data <- anpp_data %>%
+  filter(Treatment != "Irrigation") %>%
+  mutate(trt_id = as.numeric(as.factor(Treatment)) - 1)
+
+irrigate_data <- anpp_data %>%
+  filter(Treatment != "Drought") %>%
+  mutate(trt_id = as.numeric(as.factor(Treatment)) - 1)
+
+
+####
+####  SET UP AND FIT MODEL IN STAN ----
+####
+fit_stan_model <- function(model_data, check_diags=FALSE, treattype){
+  lmod <- lm(log(anpp) ~ year_id*trt_id + ppt1_scaled, model_data)
[…]

### 2017-05-23 manuscript/usses_water_ms.Rmd
updates stats table 1

@@
 library("knitr")
-basename <- "manuscript"
-opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
-               cache.path = paste("components/cache/", basename, "/", sep=""))
-opts_chunk$set(cache = 2)
 opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
                comment = NA, verbose = TRUE, echo=FALSE)
-
-# PDF-based figures
-opts_chunk$set(dev='pdf')
-```
-
-```{r plot-options, message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
-cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
-               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-
-opts_chunk$set(dev='pdf', dev.args=c(version="1.3"))
-
 ```
@@
 # Read in within-year ANOVA results table
-inyear_anova <- readRDS("../results/within_year_anova_table.RDS")
+# inyear_anova <- readRDS("../results/within_year_anova_table.RDS")
+# 
+# inyear_caption <- "ANOVA for treatment effects on ANPP for each year of the experiment."
+# print(xtable(inyear_anova, caption = inyear_caption, digits = 2),
+#       caption.placement="top",
+#       include.rownames = F, 
+#       sanitize.colnames.function = identity,
+#       comment = FALSE,
+#       size = "normalsize")
 
-inyear_caption <- "ANOVA for treatment effects on ANPP for each year of the experiment."
-print(xtable(inyear_anova, caption = inyear_caption, digits = 2),
+coef_probs <- readRDS("../results/coef_probabilities.RDS")
+
+coef_caption <- "Probability that each log(ANPP) coefficient is greater or less than zero for each of the precipitation treatments.
+Probabilities denote max(Pr(coefficient) > 0, Pr(coefficient) < 0), where the sign of the coefficient was ignored.
+Thus, probabilities represent the probability of the coefficient being important in the model.
+Bold coefficients denote Pr(coefficient) $\\geq$ 0.95."
+print(xtable(coef_probs, caption = coef_caption, digits = 2),
       caption.placement="top",

### 2017-05-23 code/anpp_repmeas_model.R

@@ -138,10 +138,9 @@ for(doparam in unique(irrigate_ref$parameter)){
 
 coef_probs <- (rbind(drought_probs, irrigate_probs)) %>%
   filter(coefficient != "(Intercept)") %>%
-  filter(coefficient != "ppt1_scaled") %>%
   mutate(coefficient = as.character(coefficient)) %>%
   spread(treatment, prob)
-coef_probs$coefficient <- c("Treatment", "Year", "Treatment x Year")
+coef_probs$coefficient <- c("Precipitation","Treatment", "Year", "Treatment x Year")
 colnames(coef_probs) <- c("log(ANPP) Coefficient", "Drought", "Irrigation")
 
 ##  Save coefficient probabilities for LaTeX table in manuscript

### 2017-05-23 code/anpp_repmeas_model.R
specifies mcmc parameters

@@ -73,7 +73,7 @@ fit_stan_model <- function(model_data, check_diags=FALSE, treattype){
   rt <- stanc("anpp_longitudinal.stan")
   sm <- stan_model(stanc_ret = rt, verbose=FALSE)
   set.seed(123)
-  fit <- sampling(sm, data=anppdat)
+  fit <- sampling(sm, data=anppdat, iter=10000, chains=4, thin=10)
   
   if(check_diags){
     pnames <- c("sigmaeps", "sigmaint", "sigmaslope","beta[1]",

### 2017-05-25 manuscript/usses_water_ms.Rmd
updates authors

@@
 
-\normalsize{Andrew T. Tredennick\textsuperscript{1}, Andrew R. Kleinhesselink\textsuperscript{2}, Bret Taylor\textsuperscript{3}, and Peter B. Adler\textsuperscript{1}}
+\normalsize{Andrew T. Tredennick\textsuperscript{1}, Andrew R. Kleinhesselink\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}}
 
@@
 \textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}} \\
-\textit{\small{\textsuperscript{2}Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California xxxxx}} \\
-\textit{\small{\textsuperscript{3}U.S. Sheep Experiment Station, Range Sheep Production Efficiency Research Unit, USDA-ARS, Dubois, Idaho 83423}}
+\textit{\small{\textsuperscript{2}Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California xxxxx}}
 

### 2017-05-31 manuscript/usses_water_ms.Rmd
updates figures and methods text

@@
 ---
-layout: 12pt
+layout: 11pt
 header-includes:
@@
   - \usepackage{todonotes}
-  - \doublespacing
+  - \linespread{1.25}
   - \usepackage{rotating}
@@
   - \usepackage{bm,mathrsfs}
-  - \usepackage{mathptmx} 
+  - \usepackage[sc]{mathpazo}
 bibliography: ~/Dropbox/Bibliography/usses_water.bib
@@
     keep_tex: false
-fontsize: 12pt
+fontsize: 11pt
 geometry: margin=1in
@@
 
-\normalsize{Andrew T. Tredennick\textsuperscript{1}, Andrew R. Kleinhesselink\textsuperscript{2}, and Peter B. Adler\textsuperscript{1}}
+\normalsize{Andrew T. Tredennick$^{1,}$\footnote{Corresponding author; e-mail: atredenn@gmail.com}, Andrew R. Kleinhesselink$^{2}$, and Peter B. Adler$^{1}$}
 
@@
 
-\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}} \\
-\textit{\small{\textsuperscript{2}Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California xxxxx}}
+\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah}} \\
+\textit{\small{\textsuperscript{2}Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California}}
 
@@
 The ecosystem was suprisingly resistant to both drought and irrigation.
-Aboveground net primary productivity (ANPP) and its temporal stability did no differ among treatments.
-The similarity of ecosystem functioning was not due to shifts at the plant community level, where species compositions and species synchrony, an important driver of ecosystem stability, among treatments were similar.
-However, ANPP does appear to be diverging among treatments with time, even though the effect is not significant.
+Aboveground net primary productivity (ANPP) responded positively to ambient growing season precipitation, but the response did not vary across treatments.
+The similarity of ecosystem functioning was not due to shifts at the plant community level, where species composition among treatments was similar and remarkably stable over the five years.
 At least in the short-term, this sagebrush steppe ecosystem is resistant to increases and decreases in annual precipitation.
@@
 
-Methods
--------
+Materials and Methods
+---------------------
 ### Study Area
@@
 ### Data Collection
-#### Aboveground Net Primary Productivity
 We estimated aboveground net primary productivity (ANPP) using a radiometer to relate ground reflectance to plant biomass [see @Byrne2011 for a review].
@@
 We fit regressions to MODIS-based NDVI and AVHRR-based NDVI for each year and retained the regression with the better fit.
-Using the best regression equation for each year, we predicted ANPP.
+Using the best regression equation for each year, we predicted ANPP (Appendix 1).
 
-#### Species Composition
 Species composition data came from annual census maps for each plot made using a pantograph [@Hill1920].
@@
 ### Data Analysis
-####  ANPP and Ecosystem Stability
-We used ANOVA with treatment, year, and treatment$\times$year terms to test the effects of the treatments on ANPP over the course of the five year experiment.
-To test for within-year differences of ANPP among treatments, we fit ANOVAs independently for each year with treatment terms only.
-We used the \texttt{lm()} function in \texttt{R} [v3.3.2, @R2016] to fit the models, and the \texttt{Anova()} function from the \texttt{car} package [@Fox2011] to extract sums-of-squares tables with Type II tests.
+We fit generalized linear mixed-effects models (GLMMs) with precipitation, treatment, year, and treatment$\times$year main effects and a random effect for plot.
+To account for temporal dependency in the data (i.e., ANPP in 2012 is likely more correlated with ANPP in 2013 than in 2016), we also include an autoregressive error structure [AR(1)].
+Thus, our GLMMs account for the fact that each plot is sampled repeatedly over time.
+Our models take the form:
 
-We calculated ecosystem stability as the temporal mean of ANPP divided by its temporal standarad deviation $\left( \overline{\text{ANPP}} / \sigma(\text{ANPP}) \right)$.
-Ecosystem stability was calculated for each plot.
-We compared ecosystem stability among treatments using ANOVA (same \texttt{R} functions as above).
+\begin{align}
+\mu_{i,t} = \textbf{x}_{i,t}\boldsymbol{\beta} + \gamma_i \\
+\textbf{y}_{i} \sim \text{MVN}\left(\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}_i \right) \\
+\boldsymbol{\Sigma}_i = \begin{bmatrix}
+       \sigma_i^2 & \rho_i\sigma_i^2 & \rho_i^2\sigma_i^2 & \rho_i^3\sigma_i^2 & \rho_i^4\sigma_i^2  \\[0.2em]
+       & \sigma_i^2 & \rho_i\sigma_i^2 & \rho_i^2\sigma_i^2 & \rho_i^3\sigma_i^2 \\[0.2em]
+       & &  \sigma_i^2 & \rho_i\sigma_i^2 & \rho_i^2\sigma_i^2 \\[0.2em]
+       & & & \sigma_i^2 & \rho_i\sigma_i^2 \\[0.2em]
+       & & & & \sigma_i^2 \\[0.2em]
+     \end{bmatrix}
+\end{align}
 
-####  Species Synchrony and Species Composition
-We calculated the synchrony of species fluctuations through time using basal area cover data and individual density data.
-For both data types we calculated the synchrony metric described by @Loreau2008a, which ranges from 0 (perfect ascynchrony) to 1 (perfect synchrony).
-Before calculating synchrony, we averaged cover and density over plots within treatments for each species.
-We used the \texttt{synchrony()} function in the \texttt{R} package \texttt{codyn} [@Hallett2016] to calculate species synchrony.
+\noindent{}where $\mu_{i,t}$ is the predicted value of log(ANPP) in plot *i* and year *t*, $\textbf{x}_{i,t}$ is the vector of covariates (including the intercept) associated with plot and year, $\boldsymbol{\beta}$ is the vector of fixed effects associated with each term in $\textbf{x}_{i,t}$, and $\gamma_i$ is the random effect for plot.
+We sample the vector of yearly observed log(ANPP) values in each plot ($\textbf{y}_{i}$) from a multivariate normal likelihood with the means from the regression equation (e.g., $\boldsymbol{\mu}_{i}$) and a variance-covariance matrix ($\Sigma_i$) with an AR(1) structure that is unique for each plot.
+We fit the above model for the drought and irrigation treatments independently because we are only interested in comparing each treatment to the control, not to eachother.
+We fit the model using a Bayesian approach, obtaining posterior estimates of all unkowns via the No-U-Turn Hamiltonian Monte Carlo sampler in Stan [@stan2016].
+We used the R package 'rstan' [@rstan2016] to link R to Stan.
 
@@
 We plotted the first two axes of NMDS scores to see if community composition overlapped, or not, among treatments in each year.
-We used functions in the \texttt{R} package \texttt{vegan} [@Oksanen2016] to calculate Bray-Curtis distances and the run the NMDS analysis.
+We used functions in the R package 'vegan' [@Oksanen2016] to calculate Bray-Curtis distances and then to run the NMDS analysis.
+Lastly, we examined rank clocks of species' abundances through time to assess the stability of community composition over the course of the experiment [@Collins2008].
+Rank clocks were made using 'ggplot2' [@Wickham2009] and R code from @Hallett2016.
 
@@
   \centering
-      \includegraphics[width=6in]{../figures/histppt_and_anpptrend.png}
+      \includegraphics[height=7in]{../figures/histppt_and_anpptrend.png}
   \caption{(A) Histogram of historical precipitation from 1926-2016, with the years of the experiment overlaid as colored points.
@@
 ``Irrigation'' treatment precipitation levels are not shown.
-(B) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.}
+(B) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.
+(C) Monthly average estimated soil volumetric water content (VWC) from SOILWAT model fit using soil moisture data from experimental plots.}
+\end{figure}
+
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=6in]{../figures/data_and_posterior_estimates.png}
+  \caption{(A) Relationship between ANPP and growing season precipitation over the course of the experiment.
+Regression lines are indpendent linear fits for each treatment for visual clarity of the mean trends.
+(B) Posterior medians (points), 80\% BCIs (heavy lines), and 95\% BCIs (light lines) of effects from the fitted repeated measures generalized linear mixed-effects model for each treatment-control comparison.
+BCI refers to 'Bayesian Credible Interval', which is the upper and lower quantiles of the posterior distribution at the specified level.}
 \end{figure}
@@
       \includegraphics[width=5in]{../figures/rank_clocks_cover.png}
-  \caption{Rank clocks for cover by treatment. (A) All species. (B) Dominant species.}
+  \caption{Rank clocks of average species' basal cover by treatment. (A) All species. (B) Dominant species. Cover is log-transformed to improve visualization of species with low cover.}
 \end{figure}

### 2017-05-31 code/anpp_repmeas_model.R

@@ -144,14 +144,15 @@ fit_error_model <- function(model_data, check_diags=FALSE, treattype){
                   Nplots = Nplots,
                   Ntimes = Ntimes,
                   y = y_matrix,
-                  x = x_array)
+                  x = x_array,
+                  xyear = xyear)
   
   rstan_options(auto_write = TRUE)
   options(mc.cores = parallel::detectCores())
   rt <- stanc("anpp_longitudinal_errorcov.stan")
   sm <- stan_model(stanc_ret = rt, verbose=FALSE)
   set.seed(123)
-  fit <- sampling(sm, data=anppdat, iter=5000, chains=4, thin=2, control = list(adapt_delta = 0.99))
+  fit <- sampling(sm, data=anppdat, iter=1500, chains=2, thin=1, control = list(adapt_delta = 0.99))
   plot(fit, pars="betas")
   
   if(check_diags){

### 2017-05-31 code/look_at_weather.R

@@ -19,6 +19,7 @@ library(tidyverse)
 library(dplyr)
 library(stringr)
 library(ggthemes)
+library(ggalt)
 library(gridExtra)
 library(RColorBrewer)
 
@@ -88,10 +89,10 @@ g2 <- ggplot(biomass_yr_trt_summ, aes(x=year, y=mean_biomass, color=Treatment))+
   geom_point(color="white", size=3)+
   geom_point()+
   geom_point(color="grey35", shape=1)+
-  annotate("text",x=2015.1,y=190,label="a",size=3)+
-  annotate("text",x=2015.1,y=210,label="a",size=3)+
-  annotate("text",x=2015.1,y=148,label="b",size=3)+
-  annotate("text",x=2015.2,y=50,label="year%*%treatment~phantom(0)~italic('P')==0.67", parse=T, size=3)+
+  # annotate("text",x=2015.1,y=190,label="a",size=3)+
+  # annotate("text",x=2015.1,y=210,label="a",size=3)+
+  # annotate("text",x=2015.1,y=148,label="b",size=3)+
+  # annotate("text",x=2015.2,y=50,label="year%*%treatment~phantom(0)~italic('P')==0.67", parse=T, size=3)+
   scale_color_brewer(palette = "Set2", name="Treatment")+
   scale_x_continuous(breaks=c(2011:2016))+
   scale_y_continuous(breaks=seq(50,350,50))+
@@ -104,8 +105,39 @@ g2 <- ggplot(biomass_yr_trt_summ, aes(x=year, y=mean_biomass, color=Treatment))+
         legend.text = element_text(size = 8),
         legend.key.height = unit(0.8,"line"))
 
+
+
+####
+####  PLOT SOIL WATER BY TREATMENT ----
+####
+soil_moisture <- read.csv("../data/soil_moisture_data/average_seasonal_soil_moisture.csv") %>%
+  select(-year) %>%
+  separate(simple_date, c("year","month","day")) %>%
+  group_by(year,month,Treatment,type,year) %>%
+  summarise(avg_vwc = mean(VWC,na.rm=TRUE)) %>%
+  filter(type=="predicted") %>%
+  mutate(month_year = as.factor(paste0(year,"-",month))) %>%
+  ungroup()
+
+g3 <- ggplot(soil_moisture, aes(x=month_year, y=avg_vwc, group=Treatment, color=Treatment))+
+  #geom_line()+
[…]

### 2017-06-01 manuscript/usses_water_ms.Rmd
results writing

@@
   - \usepackage{lineno}
-  - \linenumbers
   - \usepackage{setspace}
@@
 
+\bigskip{}
+
+Last compile: \today
+
 \end{centering}
@@
 \vspace{3em}
-
-Last compile: \today
 
@@
 
-\emph{Key words: drought, aboveground net primary productivity, ecosystem stability, climate change, species synchrony, sagebrush steppe}
+\emph{Key words: drought, aboveground net primary productivity, ecosystem resistance, climate change, species composition, sagebrush steppe}
 \end{abstract}
@@
 \setlength{\parindent}{5ex}
+
+\linenumbers
 
@@
 To account for temporal dependency in the data (i.e., ANPP in 2012 is likely more correlated with ANPP in 2013 than in 2016), we also include an autoregressive error structure [AR(1)].
-Thus, our GLMMs account for the fact that each plot is sampled repeatedly over time.
+Thus, our GLMMs account for the fact that each plot is sampled repeatedly over time, and that those yearly samples are not independent.
 Our models take the form:
@@
 \begin{align}
-\mu_{i,t} = \textbf{x}_{i,t}\boldsymbol{\beta} + \gamma_i \\
-\textbf{y}_{i} \sim \text{MVN}\left(\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}_i \right) \\
-\boldsymbol{\Sigma}_i = \begin{bmatrix}
-       \sigma_i^2 & \rho_i\sigma_i^2 & \rho_i^2\sigma_i^2 & \rho_i^3\sigma_i^2 & \rho_i^4\sigma_i^2  \\[0.2em]
-       & \sigma_i^2 & \rho_i\sigma_i^2 & \rho_i^2\sigma_i^2 & \rho_i^3\sigma_i^2 \\[0.2em]
-       & &  \sigma_i^2 & \rho_i\sigma_i^2 & \rho_i^2\sigma_i^2 \\[0.2em]
-       & & & \sigma_i^2 & \rho_i\sigma_i^2 \\[0.2em]
-       & & & & \sigma_i^2 \\[0.2em]
-     \end{bmatrix}
+\mu_{i,t} &= \textbf{x}_{i,t}\boldsymbol{\beta} + \gamma_i \\
+\textbf{y}_{i} &\sim \text{MVN}\left(\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}_i \right) \\
+\boldsymbol{\Sigma}_i &= \begin{bmatrix}
+       1 & \rho_i & \rho_i^2 & \rho_i^3 & \rho_i^4  \\[0.2em]
+       & 1 & \rho_i & \rho_i^2 & \rho_i^3 \\[0.2em]
+       & &  1 & \rho_i & \rho_i^2 \\[0.2em]
+       & & & 1 & \rho_i \\[0.2em]
+       & & & & 1 \\[0.2em]
+     \end{bmatrix} \sigma_i^2
 \end{align}
@@
 We fit the model using a Bayesian approach, obtaining posterior estimates of all unkowns via the No-U-Turn Hamiltonian Monte Carlo sampler in Stan [@stan2016].
-We used the R package 'rstan' [@rstan2016] to link R to Stan.
+We used the R package 'rstan' [@rstan2016] to link R [@R2016] to Stan.
 
@@
 
+All R code and data necessary to reproduce our analysis has been archived on Figshare (\emph{link here after acceptance}) and released on GitHub (https://github.com/atredennick/usses_water/releases/v0.1).
+We also include annotated Stan code in our model description in Appendix 1.
+
 Results
 -------
+Three of our five treatment years fell in years of below average rainfall (Fig. 1A).
+Thus, those three years represent a lower magnitude of absolute change in precipitation experienced by the treatments.
+Averaged across treatments, ANPP varied from a minimum of xxx g m$^{-2}$ in 2014 to a maximum of xxx g m$^{-2}$ in 2016 (Fig. 1B).
+ANPP was slightly higher in irrigation plots and slightly lower in drought plots (Fig. 1B), corresponding to estimated soil volumetric water content (VWC) differences among treatments (Fig. 1C).
+Such differences in soil VWC indicate our treatment infrastructure was successful.
+
+Growing season precipitation had a positive effect on ANPP (Fig.
+2 and Table 1), but ANPP over the entirety of the experiment was not different from control plots for either drought (Pr(Drought < Control) $\leq$ 0.95) or irrigation (Pr(Irrigation > Control) $\leq$ 0.95) treatments (Fig.
+2 and Table 1).
+There was also no evidence that the treatment effects became more important over time (Pr(Treatment $\times$ Year) $\leq$ 0.95; Fig. 2B and Table 1).
+
+Community composition was similar among treatments, whether looking at basal cover or density (Fig. 3).
+Likewise, community composition was remarkably stable over time, with no evidence of rank shifts or species' declines or gains (Fig. 4).
 

### 2017-06-01 manuscript/usses_water_ms.Rmd
writing trivia

@@
 Such periods of drought and deluge could become major drivers of plant community dynamics and ecosystem functioning.
-We used drought and irrigation treatments (50\% decrease/increase) to see how a sagebrush steppe plant community in Idaho will respond to future climate changes.
+Here we report the results of a five-year experiment where we used drought and irrigation treatments (50\% decrease/increase) to see how a sagebrush steppe plant community in Idaho will respond to future climate changes.
 The ecosystem was suprisingly resistant to both drought and irrigation.
 Aboveground net primary productivity (ANPP) responded positively to ambient growing season precipitation, but the response did not vary across treatments.
+There was also no evidence that treatment effects grew over time.
 The similarity of ecosystem functioning was not due to shifts at the plant community level, where species composition among treatments was similar and remarkably stable over the five years.
-At least in the short-term, this sagebrush steppe ecosystem is resistant to increases and decreases in annual precipitation.
+At least in the short-term, ecosystem functioning and community composition in this sagebrush steppe system is resistant to increases and decreases in growing season precipitation.
 
@@
 
+\textcolor{red}{[ANDY: paragraph here on SOILWAT. Is this the right thing to show/do? See Fig. 1C.]}
+
 ### Data Collection
@@
 We sample the vector of yearly observed log(ANPP) values in each plot ($\textbf{y}_{i}$) from a multivariate normal likelihood with the means from the regression equation (e.g., $\boldsymbol{\mu}_{i}$) and a variance-covariance matrix ($\Sigma_i$) with an AR(1) structure that is unique for each plot.
+Our model assumes the variance term ($\sigma^2$) is unique for each plot, but consistent across years in a plot.
+The correlation term ($\rho$) in the covariance matrix determines the correlation of measurements across years.
+The superscripts (e.g., $\rho^2$) correspond to the lag time between two measurements, and ensure that the correlation decays as the lag increases.
+
 We fit the above model for the drought and irrigation treatments independently because we are only interested in comparing each treatment to the control, not to eachother.
@@
 Community composition was similar among treatments, whether looking at basal cover or density (Fig. 3).
-Likewise, community composition was remarkably stable over time, with no evidence of rank shifts or species' declines or gains (Fig. 4).
+Likewise, community composition was remarkably stable over time, with no evidence of divergence among treatments (Fig. 3).
+Species' abundances and ranks showed no deviation over the five-year experiment, regardless of treatment (Fig. 4).
 
@@
 ---------------
-This work was funded by the National Science Foundation through grants DEB-1353078 and DEB-1054040 to PBA, a Graduate Research Fellowship to ARK, and a Postdoctoral Resaerch Fellowship in Biology to ATT (DBI-1400370).
+This work was funded by the National Science Foundation through a Postdoctoral Resaerch Fellowship in Biology to ATT (DBI-1400370), a Graduate Research Fellowship to ARK, and grants DEB-1353078 and DEB-1054040 to PBA.
 Additional support came from the Utah Agricultural Experiment Station (journal paper xxxx).
-We thank the many summer research technicians who collected the data reported in this paper.
+We thank the many summer research technicians who collected the data reported in this paper and the US Experimental Sheep Station for facilitating work on their property.
+We also thank Susan Durham for clarifying our thinking on longitudinal analyses.
 

### 2017-06-03 manuscript/usses_water_ms.Rmd
methods writing, new model

@@
 There was also no evidence that treatment effects grew over time.
-The similarity of ecosystem functioning was not due to shifts at the plant community level, where species composition among treatments was similar and remarkably stable over the five years.
+The similarity of ecosystem functioning was not due to compensatory shifts at the plant community level, where species composition among treatments was similar and remarkably stable over the five years.
 At least in the short-term, ecosystem functioning and community composition in this sagebrush steppe system is resistant to increases and decreases in growing season precipitation.
@@
 ### Data Analysis
-We fit generalized linear mixed-effects models (GLMMs) with precipitation, treatment, year, and treatment$\times$year main effects and a random effect for plot.
-To account for temporal dependency in the data (i.e., ANPP in 2012 is likely more correlated with ANPP in 2013 than in 2016), we also include an autoregressive error structure [AR(1)].
-Thus, our GLMMs account for the fact that each plot is sampled repeatedly over time, and that those yearly samples are not independent.
-Our models take the form:
+Our goal was to test whether the relationship between ANPP and growing season precipitation (hereafter, precipiation) differed among the drought, control, and irrigation treatments.
+To achieve this goal, we fit a multi-level random intercept and random slope regression with log(ANPP) as the response variabile and precipitation as the sole predictor.
+We fit the model under a Bayesian framework, allowing us to test for treatment differences by comparing the posterior distributions of the treatment-level coefficients [e.g., @Tredennick2013].
+Both log(ANPP) and precipitation were standardized to have mean 0 and unit variance before fitting the model [i.e., $(x_i - \bar{x})/\sigma_x$].
+
+Our multi-level model has three grouping levels for coefficients, representin the nested structure of the data: (i) overall coefficients, (ii) treatment coefficients, and (iii) plot coefficients.
+Each subsequent level is drawn from the distribution of coefficients at the previous level.
+Formally, our model is defined as follows:
+
+\vspace{-3em}
 
 \begin{align}
-\mu_{i,t} &= \textbf{x}_{i,t}\boldsymbol{\beta} + \gamma_i \\
-\textbf{y}_{i} &\sim \text{MVN}\left(\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}_i \right) \\
-\boldsymbol{\Sigma}_i &= \begin{bmatrix}
-       1 & \rho_i & \rho_i^2 & \rho_i^3 & \rho_i^4  \\[0.2em]
-       & 1 & \rho_i & \rho_i^2 & \rho_i^3 \\[0.2em]
-       & &  1 & \rho_i & \rho_i^2 \\[0.2em]
-       & & & 1 & \rho_i \\[0.2em]
-       & & & & 1 \\[0.2em]
-     \end{bmatrix} \sigma_i^2
+\mu_{i(j(k))} &= \beta_{0,j(k)} + \beta_{1,j(k)}x_i, \\
+y_{i(j(k))} &\sim \mathcal{N} \left(\mu_{i(j(k))}, \sigma^2_{k} \right),
 \end{align}
 
-\noindent{}where $\mu_{i,t}$ is the predicted value of log(ANPP) in plot *i* and year *t*, $\textbf{x}_{i,t}$ is the vector of covariates (including the intercept) associated with plot and year, $\boldsymbol{\beta}$ is the vector of fixed effects associated with each term in $\textbf{x}_{i,t}$, and $\gamma_i$ is the random effect for plot.
-We sample the vector of yearly observed log(ANPP) values in each plot ($\textbf{y}_{i}$) from a multivariate normal likelihood with the means from the regression equation (e.g., $\boldsymbol{\mu}_{i}$) and a variance-covariance matrix ($\Sigma_i$) with an AR(1) structure that is unique for each plot.
-Our model assumes the variance term ($\sigma^2$) is unique for each plot, but consistent across years in a plot.
-The correlation term ($\rho$) in the covariance matrix determines the correlation of measurements across years.
-The superscripts (e.g., $\rho^2$) correspond to the lag time between two measurements, and ensure that the correlation decays as the lag increases.
+\noindent{} where $\mu_{i(j(k))}$ is the deterministic prediction from the regression model for observation *i* for plot *j* associated with treatment *k*, $\beta_{0,j(k)}$ is the intercept for plot *j* associated with treatment *k*, $\beta_{1,j(k)}$ is the slope term for the effect of precipitation for plot *j* associated with treatment *k*, and $\sigma^2_k$ is the process variance for treatment *k*.
+Data include the standardized log(ANPP) observations ($y_{i(j(k))}$) and precipitation ($x_i$).
+Although we include observation subscript *i* on the *x*s, observations within a year all share the same precipitation values.
+
+The intercept and slope terms are modeled hierarchichally to account for the non-independence of observations across years within plots and to allow us to test the hypothesis that our treatments alter the ANPP-precipitation relationship.
+As noted above, plot-level coefficients are drawn from treatment-level coeffecients, which are drawn from overall coefficients.
+We also include a covariance structure among the intercept and slope at each level.
+Formally, our hierarhical structure is as follows.
+Note that we drop the intercept (0) and slope (1) subscripts and instead refer to a vector of coefficients, $\boldsymbol{\beta}$.
+
+\vspace{-2em}
+
+\begin{align}
+\boldsymbol{\beta}_{j(k)} &\sim \mathcal{N} \left( \boldsymbol{\beta}_k, \Sigma(k) \right), \\
+\boldsymbol{\beta}_{k} &\sim \mathcal{N} \left( \boldsymbol{\beta}, \Sigma  \right), \\
+\boldsymbol{\beta} &\sim \mathcal{N} \left( 0, 1 \right).
+\end{align}
 
@@
 
-coef_probs <- readRDS("../results/coef_probabilities.RDS")
-
-coef_caption <- "Probability that each log(ANPP) coefficient is greater or less than zero for each of the precipitation treatments.
-Probabilities denote max(Pr(coefficient) > 0, Pr(coefficient) < 0), where the sign of the coefficient was ignored.
+# coef_probs <- readRDS("../results/coef_probabilities.RDS")
+# 
+# coef_caption <- "Probability that each log(ANPP) coefficient is greater or less than zero for each of the precipitation treatments.
+# Probabilities denote max(Pr(coefficient) > 0, Pr(coefficient) < 0), where the sign of the coefficient was ignored.
 Thus, probabilities represent the probability of the coefficient being important in the model.
 Bold coefficients denote Pr(coefficient) $\\geq$ 0.95."
-print(xtable(coef_probs, caption = coef_caption, digits = 2),
-      caption.placement="top",
-      include.rownames = F,
-      sanitize.colnames.function = identity,
-      comment = FALSE,
-      size = "normalsize")
+# print(xtable(coef_probs, caption = coef_caption, digits = 2),
+#       caption.placement="top",
+#       include.rownames = F,
+#       sanitize.colnames.function = identity,
+#       comment = FALSE,
+#       size = "normalsize")
 ```
@@
   \centering
-      \includegraphics[width=6in]{../figures/data_and_posterior_estimates.png}
+      \includegraphics[width=6in]{../figures/glmm_treatment_posteriors.png}
   \caption{(A) Relationship between ANPP and growing season precipitation over the course of the experiment.
@@
 (B) Posterior medians (points), 80\% BCIs (heavy lines), and 95\% BCIs (light lines) of effects from the fitted repeated measures generalized linear mixed-effects model for each treatment-control comparison.
+BCI refers to 'Bayesian Credible Interval', which is the upper and lower quantiles of the posterior distribution at the specified level.}
+\end{figure}
+
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=4in]{../figures/glmm_yeardiffs.png}
+  \caption{Posterior medians (points), 80\% BCIs (heavy lines), and 95\% BCIs (light lines) of the difference between log(ANPP) in control and treatment plots over time.
 BCI refers to 'Bayesian Credible Interval', which is the upper and lower quantiles of the posterior distribution at the specified level.}

### 2017-06-03 manuscript/usses_water_ms.Rmd
funding trivia

@@
 ---------------
-This work was funded by the National Science Foundation through a Postdoctoral Resaerch Fellowship in Biology to ATT (DBI-1400370), a Graduate Research Fellowship to ARK, and grants DEB-1353078 and DEB-1054040 to PBA.
-Additional support came from the Utah Agricultural Experiment Station (journal paper xxxx).
+We gratefully acknowledge the support of the Utah Agricultural Experiment Station (journal paper xxxx).
 We thank the many summer research technicians who collected the data reported in this paper and the US Experimental Sheep Station for facilitating work on their property.
 We also thank Susan Durham for clarifying our thinking on longitudinal analyses.
+
+Funding
+-------
+NSF DBI-1400370 to Andrew Tredennick.  
+NSF Graduate Research Fellowship to Andrew Kleinhesselink.  
+NSF DEB-1353078 and DEB-1054040 to Peter Adler.
 

### 2017-06-05 code/00_source_usses_water_scripts.R
adds all sourcing script

@@ -1,0 +1,51 @@
+##  00_source_usses_water_scripts.R: script to source all code to repdroduce
+##  analysis and results for Tredennick et al. This is the main code for the
+##  'usses_water' repository.
+##
+##  Author: Andrew Tredennick (atredenn@gmail.com)
+##  Date created: June 4, 2017
+
+
+##  Clear the workspace
+rm(list=ls(all.names = TRUE))
+
+##  Set working directory to */usses_water/code/
+# setwd(my/path/usses_water/code/)
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only for RStudio, sets to file location
+
+
+
+####  1. LOAD NECESSARY PACKAGES ----
+# needed_libraries <- c(
+#   "tidyverse",
+#   "dplyr",
+#   "ggthemes",
+#   "stringr",
+#   "rstan"
+# )
+# install.packages(needed_libraries)
+library(tidyverse)
+library(dplyr)
+library(ggthemes)
+library(stringr)
+library(rstan)
+
+####  2. FIT ANPP-NDVI REGRESSIONS ----
+source("calibrate_radiometer_by_year.R")
+
+####  3. MAKE DATA FIGURE (Fig. 1) ----
+source("look_at_weather.R")
+
+####  4. FIT ANPP-PRECIP MODEL ----
+source("anpp_randcoefs_model.R")
+
+####  5. PLOT MODEL RESULTS ----
+source("plot_model_results.R")
+
+####  6. RUN NMDS ANALYSIS AND PLOTS ----
+source("community_dynamics.R")
+
+
+
+
+

### 2017-06-05 manuscript/usses_water_ms.Rmd
adds author commands for edits
trivia

@@
 \definecolor{blue}{rgb}{0,0,0.7} 
-\newcommand{\new}{\textcolor{blue}}
+\definecolor{red}{rgb}{0.7,0,0}
+\newcommand{\pba}{\textcolor{blue}}
+\newcommand{\ark}{\textcolor{red}}
 
@@
 
-Our multi-level model has three grouping levels for coefficients, representin the nested structure of the data: (i) overall coefficients, (ii) treatment coefficients, and (iii) plot coefficients.
+Our multi-level model has three grouping levels for coefficients, representing the nested structure of the data: (i) overall coefficients, (ii) treatment coefficients, and (iii) plot coefficients.
 Each subsequent level is drawn from the distribution of coefficients at the previous level.

### 2017-06-05 manuscript/usses_water_ms.Rmd
some intro writing

@@
 ------------
-Chronic alterations to resource availability should cause community composition to shift over time.
+As the rate of climate change accelerates, ecologists are being pressed to forecast its impacts on ecosystem functions and services.
+Making such forecasts faces two challenges.
+First, future conditions are likely to be outside the historical range of variability [@Smith2011], meaning we cannot simply look to the past to predict the future.
+Second, ecosystems will likely exhibit unique responses to climate change induced resource alterations [e.g., @Byrne2017], meaning we cannot simply look across space to predict the future.
+These two challenges motivate the use of *in situ* experimental manipulations of resource availability [@Avolio2015].
+
+Manipulating potentially limiting resources offers a route to understanding how ecosystems will respond to resource levels that fall outside the historical range of variability [@Knapp2017].
+Chronic alterations to resource availability should cause community composition to shift over time, the focal ecosystem is sensitive to the manipulated resource.
 Species' relative abundances are expected to re-order and, eventually, some species will be lost from the local community and new species will be gained [@Smith2009; @Avolio2015].
-Ecosystem functioning, such as productivity and stability, may remain relatively constant (CITATION) or shift to a new mean state (CITATION) depending on the trait-space of the new community.
+On the heels of changes in community composition, subsequent changes in the level of ecosystem functioning may occur.
 
-Different ecosystems are sensitive to different global changes.
-For example, @Flombaum2017 found aboveground net primary productivity (ANPP) in a Patagonian steppe was less sensitive to drought than a North American tallgrass prairie, relative to other global changes.
-Taken together with other results, @Flombaum2017 conclude that ecosystem sensitivity to a resource alteration is inversly related to the natural variation of the resource.
-In the above example, precipitation is much more variable in the Patagonian steppe than the North American tallgrass prairie.
+@Smith2009 introduced the 'Hierarchical Response Framework' for understanding the interplay of community composition and ecosystem functioning in response to resource manipulations over time.
+In the near term, ecosystem functioning such as annual net primary productivity (ANPP) will reflect the physiological responses of individual species to the manipulated resource level.
+For example, ANPP may decline under simulated drought because the initial community consisted of drought-intolerant species.
+Over longer time spans, ecosystem functioning may recover as new species, better suited to take advantage of the new resource regime, colonize local communities.
+For example, ANPP may initially decline, but eventually rise back to pre-treatment levels once drought-tolerant species colonize.
+It is also possible that ecosystem functioning shifts to a new mean state, reflective of the suite of species in the new community.
 
-This leaves us with the following qualitative understanding.
-Chronic resource alteration should induce shifts in community composition and, maybe, ecosystem functioning, but only if the focal ecosystem is predisposed to be sensitive to the particular resource that is altered.
-While a crude understanding, this does allow us to make testable predictions when confronted with an ecosystem--resource pairing.
+Much of the research on ecosystem and community responses to global climate change has focused on grassland systems, where water is typically a limiting resource.
+The sensitivity of ANPP to water availability in grasslands has been characterized spatially (across sites) and temporally (within sites).
+Spatially, there is a strong positive relationship between the amount of precipitation at a given site and mean ANPP (Knapp citation).
+Temporally, however, the response of ANPP to interannual precipitation variability is much weaker (Hsu and Adler).
 

### 2017-06-05 manuscript/usses_water_ms.Rmd
updates fig 1 and adds more results

@@
 The plant community is dominated by the shrub *Artemesia tripartita* and three perennial bunchgrasses, *Pseudoroegneria spicata*, *Poa secunda*, and *Hesperostipa comata*.
-During the period of our experiment, average mean annual precipitation was xxx mm and mean monthly temperature ranged from -x$^{\circ}$C in January to x$^{\circ}$C in July.
+During the period of our experiment (2011 -- 2015), average mean annual precipitation was 265 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
 
@@
 We used the R package 'rstan' [@rstan2016] to link R [@R2016] to Stan.
+We obtained samples from the posterior distribution for all model parameters from four parallel MCMC chains run for 10,000 iterations, saving every $10^{\text{th}}$ sample.
+Traceplots of all parameters were visually inspected to ensure well-mixed chains and convergence.
+We also made sure all scale reduction factors ($\hat{R}$) were less than 1.1.
 
@@
 
-Growing season precipitation had a positive effect on ANPP (Fig.
-2 and Table 1), but ANPP over the entirety of the experiment was not different from control plots for either drought (Pr(Drought < Control) $\leq$ 0.95) or irrigation (Pr(Irrigation > Control) $\leq$ 0.95) treatments (Fig.
-2 and Table 1).
-There was also no evidence that the treatment effects became more important over time (Pr(Treatment $\times$ Year) $\leq$ 0.95; Fig. 2B and Table 1).
+Growing season precipitation had a positive effect on ANPP (mean of $\beta_{1}$ = 0.67; 80\% BCI = 0.24, 1.10; 95\% BCI = -0.11, 1.34).
+Average ANPP was similar among treatments (similar intercepts, Fig. 2A), as was the effect of precipitation (similar slopes, Fig. 2B).
+In an average precipitation year (i.e., $x$ = 0 in Eq.
+1), the probability that ANPP in a drought plot is less than ANPP in a control plot was 0.58, and the probability that ANPP in an irrigation plot is higher than in a control plot was 0.57.
+In other words, the posterior values of $\beta_0(\text{control}) - \beta_0(\text{drought})$ and $\beta_0(\text{control}) - \beta_0(\text{drought})$ broadly overlapped zero.
+There was also no evidence that the treatment effects became more important over time because differences between average treatment and control plots were centered on zero regardless of year of the experiment (Fig.
+3).
 
-Community composition was similar among treatments, whether looking at basal cover or density (Fig. 3).
-Likewise, community composition was remarkably stable over time, with no evidence of divergence among treatments (Fig. 3).
-Species' abundances and ranks showed no deviation over the five-year experiment, regardless of treatment (Fig. 4).
+Community composition was similar among treatments, whether looking at basal cover or density (Fig. 4).
+Likewise, community composition was remarkably stable over time, with no evidence of divergence among treatments (Fig. 4).
+Species' abundances and ranks showed little deviation over the five-year experiment, regardless of treatment (Fig. 5).
 
@@
 We thank the many summer research technicians who collected the data reported in this paper and the US Experimental Sheep Station for facilitating work on their property.
-We also thank Susan Durham for clarifying our thinking on longitudinal analyses.
+We also thank Susan Durham for clarifying our thinking on the statistical analyses.
 
@@
   \centering
-      \includegraphics[height=7in]{../figures/histppt_and_anpptrend.png}
+      \includegraphics[height=7in]{../figures/Figure1.png}
   \caption{(A) Histogram of historical precipitation from 1926-2016, with the years of the experiment overlaid as colored points.
@@
 ``Irrigation'' treatment precipitation levels are not shown.
-(B) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.
-(C) Monthly average estimated soil volumetric water content (VWC) from SOILWAT model fit using soil moisture data from experimental plots.}
+(B) Monthly average estimated soil volumetric water content (VWC) from SOILWAT model fit using soil moisture data from experimental plots.
+(C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.
+(D) Scatterplot of ANPP versus growing season precipitation.}
 \end{figure}
@@
   \caption{Posterior medians (points), 80\% BCIs (heavy lines), and 95\% BCIs (light lines) of the difference between log(ANPP) in control and treatment plots over time.
-BCI refers to 'Bayesian Credible Interval', which is the upper and lower quantiles of the posterior distribution at the specified level.}
+BCI refers to 'Bayesian Credible Interval', which are the upper and lower quantiles of the posterior distribution at the specified level.}
 \end{figure}

### 2017-06-05 code/look_at_weather.R

@@ -37,7 +37,7 @@ trt_data <- rbind(trt_data, data.frame(year = trt_data$year,
                                        ppt1 = trt_data$ppt1*0.5,
                                        Treatment = "Drought"))
 
-g1 <- ggplot(weather, aes(x=ppt1))+
+ppt_histogram <- ggplot(weather, aes(x=ppt1))+
   geom_histogram(bins=20,color="white",fill="grey35")+
   geom_point(data=trt_data, aes(x=ppt1, y=rep(seq(0.3,1,length.out = 5),2),shape=Treatment, fill=as.factor(year)), size=2)+
   scale_x_continuous(expand=c(0,0), limits=c(0,620), breaks=seq(0,600,100))+
@@ -48,14 +48,14 @@ g1 <- ggplot(weather, aes(x=ppt1))+
   xlab("Growing Season Precipitation (mm)")+
   ggtitle("A")+
   theme_few()+
-  theme(legend.position = c(0.8, 0.6),
+  theme(legend.position = c(0.8, 0.5),
         legend.key.size = unit(1,"pt"),
         legend.title = element_text(size=10),
         legend.text = element_text(size = 8),
         legend.key.height = unit(0.8,"line"))+
   guides(fill = guide_legend(override.aes = list(color = brewer.pal(5,"Set1"),size=1)),
          shape = guide_legend(override.aes = list(size=1)))
-ggsave(plot = g1,"../figures/historical_precip_trts.png", width = 4, height = 3.5, units="in", dpi=120)
+#ggsave(plot = g1,"../figures/historical_precip_trts.png", width = 4, height = 3.5, units="in", dpi=120)
 
 
 
@@ -83,7 +83,7 @@ biomass_yr_trt_summ <- permanent_quad_biomass %>%
   filter(year > 2011)
 
 stats_lab <- expression(paste("year*treatment  ", italic("P"), " = 0.67"))
-g2 <- ggplot(biomass_yr_trt_summ, aes(x=year, y=mean_biomass, color=Treatment))+
+anpp_means <- ggplot(biomass_yr_trt_summ, aes(x=year, y=mean_biomass, color=Treatment))+
   geom_line()+
   geom_errorbar(aes(ymin=mean_biomass-sd_biomass, ymax=mean_biomass+sd_biomass), width=0.05)+
   geom_point(color="white", size=3)+
[…]

### 2017-06-05 code/plot_model_results.R

@@ -100,3 +100,21 @@ g2 <- ggplot(ydiffs, aes(y=`50%`, x=pptyear, color=Treatment))+
 ggsave("../figures/glmm_yeardiffs.png",plot = g2, width = 4, height = 3.5, units = "in", dpi = 120)
 
 
+
+####
+####  EXTRACT AND SAVE ALL COEFFICIENT SUMMARIES ----
+####
+all_betas <- as.data.frame(summary(all_fit, pars = c("beta_treat","beta_mu"), probs = c(0.025,0.1,0.5,0.9,0.975))$summary)
+write.csv(all_betas, "../results/beta_summaries.csv")
+
+
+
+####
+####  CALCULATE PROBABILITY OF DIFFERENCE BASED ON INTERCEPTS ----
+####
+intercept_diffs <- reshape2::melt(rstan::extract(all_fit, pars="inter_diffs")) %>%
+  rename(iteration = iterations, treat_id = Var2, estimate = value, stan_name = L1)
+Pr_control_above_drought <- 1 - ecdf(filter(intercept_diffs,treat_id == 1)[,"estimate"])(0)
+Pr_control_below_irrigation <- ecdf(filter(intercept_diffs,treat_id == 2)[,"estimate"])(0)
+mean_pptyear_probs <- data.frame(Pr_control_above_drought,Pr_control_below_irrigation)
+write.csv(mean_pptyear_probs, "../results/mean_pptyear_probs.csv")

### 2017-06-05 manuscript/Appendix1.Rmd
more results and writing

---
layout: 11pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \onehalfspacing
  - \usepackage{rotating}
  - \usepackage{color, soul}
  - \usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{tikz}
  - \usepackage{bm,mathrsfs}
  - \usepackage[sc]{mathpazo}
bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
csl: components/ecology_letters.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
    number_sections: true
fontsize: 11pt
geometry: margin=1in
linkcolor: black
urlcolor: black
---

\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{A1-\arabic{figure}}  
\renewcommand\thetable{A1-\arabic{table}}  
\renewcommand\thesection{Section A1.\arabic{section}}

\begin{center}
\textbf{\Large{Appendix 1}} \\
A.T. Tredennick, A.R. Kleinhesselink, \& P.B. Adler \\
``Ecosystem and community resistance...'' \\
\emph{PeerJ}
\end{center}

\section{Estimating ANPP}

### 2017-06-05 manuscript/usses_water_ms.Rmd
more results and writing

@@
 \mu_{i(j(k))} &= \beta_{0,j(k)} + \beta_{1,j(k)}x_i, \\
-y_{i(j(k))} &\sim \mathcal{N} \left(\mu_{i(j(k))}, \sigma^2_{k} \right),
+y_{i(j(k))} &\sim \text{Normal} \left(\mu_{i(j(k))}, \sigma^2_{k} \right),
 \end{align}
@@
 We also include a covariance structure among the intercept and slope at each level.
-Formally, our hierarhical structure is as follows.
-Note that we drop the intercept (0) and slope (1) subscripts and instead refer to a vector of coefficients, $\boldsymbol{\beta}$.
+Formally, our hierarhical structure is as follows, where we drop the intercept (0) and slope (1) subscripts and instead refer to a vector of coefficients, $\boldsymbol{\beta}$:
 
@@
 \begin{align}
-\boldsymbol{\beta}_{j(k)} &\sim \mathcal{N} \left( \boldsymbol{\beta}_k, \Sigma(k) \right), \\
-\boldsymbol{\beta}_{k} &\sim \mathcal{N} \left( \boldsymbol{\beta}, \Sigma  \right), \\
-\boldsymbol{\beta} &\sim \mathcal{N} \left( 0, 1 \right).
+\boldsymbol{\beta}_{j(k)} &\sim \text{MVN} \left( \boldsymbol{\beta}_k, \Sigma(k) \right), \\
+\boldsymbol{\beta}_{k} &\sim \text{MVN} \left( \boldsymbol{\beta}, \Sigma  \right), \\
+\boldsymbol{\beta} &\sim \text{Normal} \left( 0, 1 \right),
 \end{align}
+
+\noindent{} where $\boldsymbol{\beta}_{j(k)}$ is the vector of regression coefficients (intercept and slope) for plot *j* associated with treatment *k*, $\boldsymbol{\beta}_{k}$ is the vector of coefficients for each treatment, and $\boldsymbol{\beta}$ is the vector of overall coefficients.
+The plot- and treatment-level coefficients are drawn from multivariate normal distributions with covariance matrix $\Sigma$.
+For the plot-level coefficients, each treatment has its own variance-covariance matrix (i.e., $\Sigma(k)$).
+The overall coefficients are drawn from a normal prior with mean 0 and standard deviation 1.
+A full description of model is in Appendix 2.
 
@@
 All R code and data necessary to reproduce our analysis has been archived on Figshare (\emph{link here after acceptance}) and released on GitHub (https://github.com/atredennick/usses_water/releases/v0.1).
-We also include annotated Stan code in our model description in Appendix 1.
+We also include annotated Stan code in our model description in Appendix 2.
 
@@
 Thus, those three years represent a lower magnitude of absolute change in precipitation experienced by the treatments.
-Averaged across treatments, ANPP varied from a minimum of xxx g m$^{-2}$ in 2014 to a maximum of xxx g m$^{-2}$ in 2016 (Fig. 1B).
-ANPP was slightly higher in irrigation plots and slightly lower in drought plots (Fig. 1B), corresponding to estimated soil volumetric water content (VWC) differences among treatments (Fig. 1C).
+Averaged across treatments, ANPP varied from a minimum of 74.5 g m$^{-2}$ in 2014 to a maximum of 237.1 g m$^{-2}$ in 2016 (Fig. 1C).
+ANPP was slightly higher in irrigation plots and slightly lower in drought plots (Fig. 1C), corresponding to estimated soil volumetric water content (VWC) differences among treatments (Fig. 1B).
 Such differences in soil VWC indicate our treatment infrastructure was successful.
 
-Growing season precipitation had a positive effect on ANPP (mean of $\beta_{1}$ = 0.67; 80\% BCI = 0.24, 1.10; 95\% BCI = -0.11, 1.34).
+Growing season precipitation had a positive effect on ANPP (mean of $\beta_{1}$ = 0.67; 80\% BCI = 0.24, 1.10; 95\% BCI = -0.11, 1.34) (Fig. 1D).
 Average ANPP was similar among treatments (similar intercepts, Fig. 2A), as was the effect of precipitation (similar slopes, Fig. 2B).
@@
 1), the probability that ANPP in a drought plot is less than ANPP in a control plot was 0.58, and the probability that ANPP in an irrigation plot is higher than in a control plot was 0.57.
-In other words, the posterior values of $\beta_0(\text{control}) - \beta_0(\text{drought})$ and $\beta_0(\text{control}) - \beta_0(\text{drought})$ broadly overlapped zero.
+In other words, the posterior distributions of $\beta_0(\text{control}) - \beta_0(\text{drought})$ and $\beta_0(\text{control}) - \beta_0(\text{drought})$ broadly overlapped zero.
 There was also no evidence that the treatment effects became more important over time because differences between average treatment and control plots were centered on zero regardless of year of the experiment (Fig.
@@
 ----------
+We manipulated growing season precipitation for five years in a sagebrush steppe to reach the extremes of the historical distribution of precipitation (Fig. 1A).
+In so doing, we aimed to understand how this ecosystem might respond to periods of relatively novel precipitation regimes.
+Our results suggest that ecosystem functioning, as measured by annual net primary productivity, and community composition are not sensitive to the precipitation manipulations we imposed.
 
@@
 
+<!---
 Tables
@@
 \newpage{}
+
+--->
 

### 2017-06-05 code/look_at_weather.R

@@ -192,3 +192,13 @@ weather_table <- data.frame(avg_ann_precip = avg_mar,
 
 write.csv(x = weather_table, file = "../results/weather_summary.csv")
 
+
+
+####
+####  GET MIN AND MAX ANPP VALUES, AVERAGED OVER TREATMENT ----
+####
+source("read_format_data.R") # load data
+anpp_summary <- anpp_data %>%
+  group_by(year) %>%
+  summarise(mean_anpp = mean(anpp))
+write.csv(anpp_summary, "../results/avg_anpp_by_year.csv")

### 2017-06-06 manuscript/usses_water_ms.Rmd
trivia

@@
 \begin{align}
-\boldsymbol{\beta}_{j(k)} &\sim \text{MVN} \left( \boldsymbol{\beta}_k, \Sigma(k) \right), \\
-\boldsymbol{\beta}_{k} &\sim \text{MVN} \left( \boldsymbol{\beta}, \Sigma  \right), \\
+\boldsymbol{\beta}_{j(k)} &\sim \text{MVN} \left( \boldsymbol{\beta}_k, \boldsymbol{\Sigma}(k) \right), \\
+\boldsymbol{\beta}_{k} &\sim \text{MVN} \left( \boldsymbol{\beta}, \boldsymbol{\Sigma}  \right), \\
 \boldsymbol{\beta} &\sim \text{Normal} \left( 0, 1 \right),
@@
 \noindent{} where $\boldsymbol{\beta}_{j(k)}$ is the vector of regression coefficients (intercept and slope) for plot *j* associated with treatment *k*, $\boldsymbol{\beta}_{k}$ is the vector of coefficients for each treatment, and $\boldsymbol{\beta}$ is the vector of overall coefficients.
-The plot- and treatment-level coefficients are drawn from multivariate normal distributions with covariance matrix $\Sigma$.
-For the plot-level coefficients, each treatment has its own variance-covariance matrix (i.e., $\Sigma(k)$).
+The plot- and treatment-level coefficients are drawn from multivariate normal distributions with covariance matrix $\boldsymbol{\Sigma}$.
+For the plot-level coefficients, each treatment has its own variance-covariance matrix (i.e., $\boldsymbol{\Sigma}(k)$).
 The overall coefficients are drawn from a normal prior with mean 0 and standard deviation 1.
@@
 
-We fit the above model for the drought and irrigation treatments independently because we are only interested in comparing each treatment to the control, not to eachother.
 We fit the model using a Bayesian approach, obtaining posterior estimates of all unkowns via the No-U-Turn Hamiltonian Monte Carlo sampler in Stan [@stan2016].

### 2017-06-06 manuscript/usses_water_ms.Rmd
bad intro writing

@@
 Precipitation is predicted to become more variable in the western U.S., meaning years of above and below average precipitation will become more common.
-Such periods of drought and deluge could become major drivers of plant community dynamics and ecosystem functioning.
+Such periods of drought and deluge could become major drivers of plant community dynamics and ecosystem functioning in water limited grasslands.
 Here we report the results of a five-year experiment where we used drought and irrigation treatments (50\% decrease/increase) to see how a sagebrush steppe plant community in Idaho will respond to future climate changes.
@@
 
-\emph{Key words: drought, aboveground net primary productivity, ecosystem resistance, climate change, species composition, sagebrush steppe}
+\emph{Key words: aboveground net primary productivity, climate change, community dynamics, drought, ecosystem resistance, sagebrush steppe}
 \end{abstract}
@@
 In the near term, ecosystem functioning such as annual net primary productivity (ANPP) will reflect the physiological responses of individual species to the manipulated resource level.
-For example, ANPP may decline under simulated drought because the initial community consisted of drought-intolerant species.
-Over longer time spans, ecosystem functioning may recover as new species, better suited to take advantage of the new resource regime, colonize local communities.
-For example, ANPP may initially decline, but eventually rise back to pre-treatment levels once drought-tolerant species colonize.
-It is also possible that ecosystem functioning shifts to a new mean state, reflective of the suite of species in the new community.
+For example, ANPP may decline under simulated drought because the initial community consisted of drought-intolerant species [@Hoover2014].
+Over longer time spans, ecosystem functioning may recover as new species colonize or initial species reorder in relative abundance.
+For example, ANPP may initially decline, but eventually rise back to pre-treatment levels once drought-tolerant species become more abundant and compensate for drought-intolerant species [@Hoover2014].
+It is also possible that ecosystem functioning shifts to a new mean state, reflective of the suite of species in the new community [@Knapp2012].
 
-Much of the research on ecosystem and community responses to global climate change has focused on grassland systems, where water is typically a limiting resource.
+Much of the research on ecosystem and community responses to global climate change has focused on grassland systems, where soil water is typically a limiting resource.
 The sensitivity of ANPP to water availability in grasslands has been characterized spatially (across sites) and temporally (within sites).
-Spatially, there is a strong positive relationship between the amount of precipitation at a given site and mean ANPP (Knapp citation).
-Temporally, however, the response of ANPP to interannual precipitation variability is much weaker (Hsu and Adler).
+Across sites, there is a strong positive relationship between the amount of precipitation at a given site and mean ANPP [@Sala1988; @Huxman2004].
+Within sites, however, the response of ANPP to annual precipitation amount is much weaker [@Huxman2004; @Hsu2012].
+Furthermore, it is becoming clear that grassland ecosystem response to altered water availability is likely to be idiosyncratic, conforming to neither cross-site nor within-site ANPP-precipitation relationships [@Wilcox2016].
+This is because within-site relationships do not account for species reordering under chronic resource alterations, while cross-site relationships may overestimate the ability of species compensation.
+Thus, discovering which ecosystems are most sensitive to chronic alterations of water availability, and the time scales at which ecosystem and community responses occur, is critical to avoid surprises in the face of climate change.
 
-In many areas of the western United States, precipitation is likely to become more variable.
-This will result in swings back-and-forth between multiyear periods of above and above average precipitation.
-An obvious question is, how will semiarid ecosystems respond to periods of drought and deluge?
-A naive expectation is that alterations in precipitation, and subsequently available soil moisture, should cause changes in plant community composition and increase or decrease ANPP dependending on the direction of precipitation change.
-But our emerging qualitative understanding suggests an alternative expectation: altering soil moisture may have little to no effect on plant community composition and ecosystem functioning.
-This expectation stems from the fact that precipitation is already a variable resource in semiarid systems, meaning the plant community is not sensitive to realistic increases/decreases in precipitation.
+Here we report the results of a five-year precipitation manipulation experiment in a sagebrush steppe grassland.
+We imposed drought and irrigation treatments (50\% decrease/increase) and measured ecosystem (ANPP) and community (species composition) responses.
+We focus on how the drought and irrigation treatments affect the linear relationship between ANPP and growing season precipitation, and if community dynamics underly the ecosystem responses.
+In particular, we sought to answer the following questions:
 
-Here we test our qualitative understanding of how altering precipitation will impact a sagebrush steppe ecosystem by imposing drought and irrigation for five years.
-In particular, we test the following competing predictions:
-
-\begin{description}
-\item [P1.] Altering precipitation will favor certain species over others, resulting in a shift in the plant community composition.
-ANPP will increase (irrigation) or decrease (drought), and the treatment effects will get stronger over time (a treatment$\times$year interaction).
-
-\item[P2.] Altering precipitation will not favor certain species and plant community composition will not change.
-ANPP will increase (irrigation) or decrease (drought), and the treatment effects will get stronger over time.
-\end{description}
-
-\noindent{}Note that both predictions assume ANPP will respond to precipitation alteration.
-Under **P1**, ANPP response stems directly from plant community shifts to species able to take advantage of the new precipitation regime.
-Under **P2**, ANPP response stems from all species responding similarly to changes in precipitation.
+1. Does drought or irrigation impact mean ANPP (i.e., different intercepts)?
+2. Does drought or irrigation impact the effect of ambient precipitation on ANPP (i.e., different slopes)?
+3. Is the resistance or sensitivity of ANPP to altered precipitation related to underlying community dynamics?
 

### 2017-06-06 manuscript/Appendix1.Rmd
beginnings of appendix 1

@@
   - \usepackage{lineno}
-  - \linenumbers
   - \usepackage{setspace}
   - \usepackage{todonotes}
-  - \onehalfspacing
+  - \linespread{1.25}
   - \usepackage{rotating}
@@
 
-\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
@@
 
+\linenumbers
+
 \section{Estimating ANPP}
+We used a radiometer to nondestructively estimate aboveground net primary productivity.
+Our approach relies on relating greenness in a plot to aboveground biomass.
+
+\section{Additional Figures}

### 2017-06-07 manuscript/Appendix1.Rmd
writing appendix 1

@@
   - \usepackage[sc]{mathpazo}
+  - \usepackage{mathtools}
 bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
@@
 Our approach relies on relating greenness in a plot to aboveground biomass.
+In each year we recorded ground reflectances at four bands, two associated with the red spectrum and two associated with the near-infrared spectrum (Table A1-1).
+We took four readings per plot that were averaged.
+Bands 1 and 3 correspond to wavelengths collected by the MODIS satellite and bands 2 and 4 correspond to wavelengths collected by the AVHRR satellite.
 
-\section{Additional Figures}
+| Band number | Spectrum name | Spectrum wavelengths | Corresponding satellite |
+| ----------- | ------------- | -------------------- | ----------------------- |
+| 1 | red | xx-xx $\mu$m | AVHRR |
+| 2 | red | xx-xx $\mu$m | MODIS |
+| 3 | near-infrared | xx-xx $\mu$m | AVHRR |
+| 4 | near-infrared | xx-xx $\mu$m | MODIS |
+
+Table: Radiometer specifications.
+
+Using the RED and NIR reflectance values, we calculate the normalized difference vegetation index (NDVI) for each plot based on both AVHRR- and MODIS-based wavelengths.
+We calculated NDVI as:
+
+\begin{align}
+\text{NDVI}_{\text{AVHRR}} &= \frac{b_3\delta_{1(\text{AVHRR})} - b_1\delta_{2(\text{AVHRR})}}{b_3\delta_{1(\text{AVHRR})} + b_1\delta_{2(\text{AVHRR})}} \\[10pt]
+\text{NDVI}_{\text{MODIS}} &= \frac{b_4\delta_{1(\text{MODIS})} - b_2\delta_{2(\text{MODIS})}}{b_4\delta_{1(\text{MODIS})} + b_2\delta_{2(\text{MODIS})}}
+\end{align}
+
+\noindent{} where $b_x$ refers to band *x* (*x* = 1,2,3,4) in Table A1-1 and $\delta$s are scaling factors unique to each band.
+The values for the scaling factors come from \textcolor{red}{(PETER?)} and are as follows: $\delta_{1(AVHRR)} = 0.77$, $\delta_{2(AVHRR)} = 1$, $\delta_{1(MODIS)} = 0.95$, $\delta_{2(MODIS)} = 0.96$.
+
+To convert plot NDVI to biomass, we regressed known biomass values from calibration plots against NDVI calculate for those plots.
+Calibration plots were located near our experiment plots, and each year we located a new set of 10 plots in which we clipped all aboveground biomass, dried it to a constant weight at 60$^\circ$ C, and the weighed.
+We used these biomass values to estimate regression parameters for both AVHRR- and MODIS-based NDVI.
+We assessed model fit using $R^2$ and, for each year, we used the regression parameters associated with the best fit model to estimate biomass in the experimental plots based on their NDVI values (Table A1-2).
+R code for this procedure is in the file "calibrate_radiometer_by_year.R" in the supplemental code set.
+
+```{r ndvi_table, include=T, echo=F, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
+library(xtable)
+library(tidyverse)
+library(dplyr)
+tab_dat <- read.csv("../data/estimated_biomass/ndvi_biomass_regression_parameters.csv", row.names = 1) %>%
+  filter(year > 2011)
+colnames(tab_dat) <- c("Year", "Intercept", "NDVI Slope", "$R^2$", "Min(biomass)", "Max(biomass)", "Algorithm")
+
+my_caption <- "Details of regression models used to estimate biomass each year."
+print(xtable(tab_dat, caption = my_caption, digits = 2),
+      caption.placement="top",
+      include.rownames = F,
+      sanitize.colnames.function = identity,
+      comment = FALSE,
+      size = "normalsize")
+```

### 2017-06-09 manuscript/Appendix2.Rmd
more results and writing
writing app2

--- 2017-06-05 manuscript/Appendix1.Rmd
@@
   - \usepackage[sc]{mathpazo}
-bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
-csl: components/ecology_letters.csl
+bibliography: ~/Dropbox/Bibliography/usses_water.bib
+csl: components/ecology.csl
 
@@
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
-\renewcommand\thefigure{A1-\arabic{figure}}  
-\renewcommand\thetable{A1-\arabic{table}}  
-\renewcommand\thesection{Section A1.\arabic{section}}
+\renewcommand\thefigure{A2-\arabic{figure}}
+\renewcommand\theequation{A2.\arabic{equation}}
+\renewcommand\thetable{A2-\arabic{table}}  
+\renewcommand\thesection{Section A2.\arabic{section}}
 
 \begin{center}
-\textbf{\Large{Appendix 1}} \\
+\textbf{\Large{Appendix 2}} \\
 A.T. Tredennick, A.R. Kleinhesselink, \& P.B. Adler \\
@@
 
-\section{Estimating ANPP}
+\section{Random Slopes, Random Intercepts Model}
+\subsection{Model Description}
+Our hierarchical Bayesian model with random intercepts and random slopes allowed us to test for differences among treatments in the relationship between ANPP and growing season precipitation, and allowed us to account for the non-independence of observations through time within a plot.
+The random effects structure mirrors the nested structure of the data: plot-level parameters are nested within treatment-level parameters, which are nested within overall parameters.
+In what follows, $\textbf{X}$ is the design matrix including a column of 1s for intercepts and a column of precipitation values for each observation, $\textbf{y}$ is a row vector of log(ANPP) values for each observation, *i* references observations, *j* references plots, and *k* references treatment.
+The notation *i(j(k))* reads as 'observation *i* associated with plot *j* associated with treatment *k*'.
+Recall we have three treatments: drought (~50\% decrease in precipitation), control (ambient precipitation), and irrigation (~50\% increase in precipitation).
+
+We assume the observations are conditionally Gaussian,
+
+\vspace{-2em}
+
+\begin{align}
+y_{i(j(k))} \sim \text{Normal}\left(\mu_{i(j(k))}, \sigma^2_k \right),
+\end{align}
+
+\noindent{}where $\mu_{i(j(k))}$ is the determinstic expectation from the regression model,
+
+\vspace{-2em}
+
+\begin{align}
+\mu_{i(j(k))} = \textbf{x}_i'\boldsymbol{\beta_{j(k)}}.
+\end{align}
+
+Regression parameters (the intercept and slope in the parameter vector $\boldsymbol{\beta}$) were modeled hierarchically,
+
+\vspace{-2em}
+
+\begin{align}
+\boldsymbol{\beta}_{j(k)} &\sim \text{MVN} \left( \boldsymbol{\beta}_k, \boldsymbol{\Sigma}(k) \right), \\
+\boldsymbol{\beta}_{k} &\sim \text{MVN} \left( \boldsymbol{\beta}, \boldsymbol{\Sigma}  \right), \\
+\boldsymbol{\beta} &\sim \text{Normal} \left( 0, 1 \right),
+\end{align}
+
+\noindent{} where $\boldsymbol{\beta}_{j(k)}$ is the vector of regression coefficients (intercept and slope) for plot *j* associated with treatment *k*, $\boldsymbol{\beta}_{k}$ is the vector of coefficients for each treatment, and $\boldsymbol{\beta}$ is the vector of overall coefficients.
+The plot- and treatment-level coefficients are drawn from multivariate normal distributions with covariance matrix $\boldsymbol{\Sigma}$.
+For the plot-level coefficients, each treatment has its own variance-covariance matrix (i.e., $\boldsymbol{\Sigma}(k)$).
+The overall coefficients are drawn from a normal prior with mean 0 and standard deviation 1.
+
+We fit the model using MCMC as implemented in the software Stan [@stan2016].
+Our Stan code is below.
+All code necessary to reproduce our results is in the supplementary code set (Appendix 3).
+
+\subsection{Stan Code}
+```{r stan1, echo=T, eval=F}
+data {
+  int<lower=0> Npreds;        # number of covariates, including intercept
+  int<lower=0> Nplots;        # number of plots
+  int<lower=0> Ntreats;       # number of treatments
+  int<lower=0> Nobs;          # number of observations
+  int<lower=0> Nppts;         # Number of precip levels to predict
+  vector[Nobs] y;             # vector of observations
+  row_vector[Npreds] x[Nobs]; # design matrix
+  matrix[Nppts,Npreds] newx;  # design matrix for predictions 
+  matrix[Npreds,Npreds] R;	  # priors for covariance matrix
+  int plot_id[Nobs];          # vector of plot ids
+  int treat_id[Nobs];         # vector of treatment ids
+}
+
+parameters {
+  vector[Npreds] beta_plot[Nplots]; # a unique vector matrix for each plot
+  vector[Npreds] beta_treat[Ntreats]; # a unique vector matrix for each treatment
+	vector[Npreds] beta_mu; # overall coefficients
+	cov_matrix[Npreds] Sigma; # covariance matrix for treatment-level coefficients
+	cov_matrix[Npreds] Sigma_plot[Ntreats]; # unique covariance matrix for plot-level coefs by treatment
+	vector<lower=0>[Ntreats] sd_y; # treatment-level observation std. dev.
+}
+
+transformed parameters {
+  vector[Nobs] yhat; # vector of expected values (predictions)
+  for (i in 1:Nobs)
+    # regression model for expected values (one for each plot-year)
+    yhat[i] = x[i]*beta_plot[plot_id[i]]; 
+}
+
+model {
+  ####  PRIORS
+  for(i in 1:Nplots)
+    # plot-level coefficients vary normally around treatment coefs
+		beta_plot[i] ~ multi_normal(beta_treat[treat_id[i]], Sigma_plot[treat_id[i]]);
+
+	for(i in 1:Ntreats){
+	  # treatment-level coefficients vary normally around overall coefficients
+	  beta_treat[i] ~ multi_normal(beta_mu, Sigma);
+
+	  # priors on covariance of effects at plot-level
+	  Sigma_plot[i] ~ inv_wishart(Npreds+1, R);
+	}
+
+	beta_mu ~ normal(0,1); # priors on overall effects
+	Sigma ~ inv_wishart(Npreds+1, R); # priors on covariance of effects at treatment-level
+	sd_y ~ weibull(2,1); # priors on observation std. dev. for each treatment
+
+	####  LIKELIHOOD
+	for(i in 1:Nobs)
+	  # observations vary normally around expected values
+	  y[i] ~ normal(yhat[i], sd_y[treat_id[i]]);
+}
+
+generated quantities {
+  vector[Nppts] ypreds[Ntreats];
+  vector[Nppts] ydiff_control_drought;
+  vector[Nppts] ydiff_control_irrigate;
+  vector[2] inter_diffs;
+  for(i in 1:Ntreats)
+    ypreds[i] = newx*beta_treat[i]; # mean predictions for each treatment
+
+  # difference between mean predictions
+  ydiff_control_drought = ypreds[1] - ypreds[2]; 
+  ydiff_control_irrigate = ypreds[1] - ypreds[3];
+
+  # difference between control and treatment in avg ppt year
+  inter_diffs[1] = beta_treat[1][1] - beta_treat[2][1]; # control - drought
+  inter_diffs[2] = beta_treat[1][1] - beta_treat[3][1]; # control - irrigation
+}
+
+```
+
+# References

### 2017-06-09 code/anpp_randyears_model.R
adds year*treatment anova-style model and results

@@ -1,0 +1,85 @@
+##  anpp_randcoefs_model.R: Script to run GLMM analysis to test for treatment 
+##  effects on the relationship between precipitation and ANPP.
+##
+##  NOTE: Stan will issue a couple warnings after running the MCMC that, as
+##  the messages state, can be safely ignored. Just rejects a couple proposals
+##  that result in ill-formed covariance matrices.
+##
+##  Author: Andrew Tredennick
+##  Date created: June 2, 2017
+
+##  Clear everything
+rm(list=ls(all.names = T))
+
+##  Set working directory to source file location
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only for RStudio
+
+
+
+####
+####  LOAD LIBRARIES ----
+####
+library(tidyverse)    # Data munging
+library(dplyr)        # Data summarizing
+library(ggthemes)     # Pleasing ggplot themes
+library(stringr)      # Working with strings
+library(rstan)        # For MCMC
+
+
+
+
+####
+####  READ IN AND EXTRACT EXPERIMENT ANPP DATA ----
+####
+source("read_format_data.R")
+
+
+
+####
+####  SET UP AND FIT MODEL IN STAN ----
+####
+fit_stan_model <- function(model_data, check_diags=FALSE, treattype){
+  lmod <- lm(log(anpp) ~ Treatment, model_data)
+  x <- model.matrix(lmod)
+  anppdat <- list(Nobs = nrow(model_data),
+                  Npreds = ncol(x),
+                  Nplots = length(unique(model_data$quadname)),
+                  Nyears = length(unique(model_data$year_id)),
+                  y = as.numeric(scale(log(model_data$anpp))),
+                  x = x,
+                  plot_id = as.numeric(as.factor(model_data$quadname)),
+                  year_id = model_data$year_id,
+                  R = diag(1,ncol(x)))
+  
+  rstan_options(auto_write = TRUE)
+  options(mc.cores = parallel::detectCores())
+  rt <- stanc("anpp_treatyr.stan")
+  sm <- stan_model(stanc_ret = rt, verbose=FALSE)
+  set.seed(123)
+  fit <- sampling(sm, data=anppdat, iter=10000, chains=4, thin=10)
+  
+  if(check_diags){
+    pnames <- c("sigmaeps", "sigmaint", "sigmaslope","beta[1]",
[…]

### 2017-06-09 code/plot_model_results.R
updates year effects figure

@@ -20,6 +20,7 @@ library(ggthemes)     # Pleasing ggplot themes
 library(stringr)      # Working with strings
 library(rstan)        # For MCMC and Stan objects
 library(gridExtra)    # For combining ggplot objects
+library(viridis)
 
 
 
@@ -28,6 +29,7 @@ library(gridExtra)    # For combining ggplot objects
 ####
 source("read_format_data.R") # load data
 all_fit <- readRDS("../results/randcoefs_alltreatments_fit.RDS")
+year_fit <- readRDS("../results/randyears_alltreatments_fit.RDS")
 
 
 
@@ -76,28 +78,49 @@ ggsave("../figures/glmm_treatment_posteriors.png",plot = g1, width = 6, height =
 ####
 ####  PLOT TREATMENT DIFFERENCES BY YEAR POSTERIORS ----
 ####
-ydiffs_drought_summary <- as.data.frame(summary(all_fit, pars = c("ydiff_control_drought"), probs = c(0.025,0.1,0.5,0.9,0.975))$summary) %>%
-  mutate(pptyear = 1:5, Treatment = "Drought")
-ydiffs_irrigate_summary <- as.data.frame(summary(all_fit, pars = c("ydiff_control_irrigate"), probs = c(0.025,0.1,0.5,0.9,0.975))$summary) %>%
-  mutate(pptyear = 1:5, Treatment = "Irrigation")
-ydiffs <- rbind(ydiffs_drought_summary, ydiffs_irrigate_summary)
+param_id_names <- data.frame(param_id = c(1:3),
+                             param_name = c("Control", "Drought", "Irrigation"))
+year_id_names <- data.frame(year_id = c(1:5),
+                            year_name = c(1:5))
+betas <- reshape2::melt(rstan::extract(year_fit, pars="beta")) %>%
+  rename(iteration = iterations, year_id = Var2, param_id = Var3, estimate = value, stan_name = L1) %>%
+  left_join(param_id_names, by="param_id") %>%
+  left_join(year_id_names, by="year_id")
 
-dodge <- position_dodge(width=0.3)
-g2 <- ggplot(ydiffs, aes(y=`50%`, x=pptyear, color=Treatment))+
-  geom_hline(aes(yintercept=0), linetype=2, size=0.2)+
-  geom_errorbar(aes(ymin = `2.5%`, ymax=`97.5%`), width=0, position = dodge)+
[…]

### 2017-06-09 manuscript/usses_water_ms.Rmd
trivia
updates year effects figure
updates model description for ANOVA thingy

@@
 ### Data Analysis
-Our goal was to test whether the relationship between ANPP and growing season precipitation (hereafter, precipiation) differed among the drought, control, and irrigation treatments.
-To achieve this goal, we fit a multi-level random intercept and random slope regression with log(ANPP) as the response variabile and precipitation as the sole predictor.
+Our main goal was to test whether the relationship between ANPP and growing season precipitation (hereafter, precipiation) differed among the drought, control, and irrigation treatments.
+To achieve this goal, we fit a multi-level random intercept and random slope regression with log(ANPP) as the response variabile and precipitation as the sole predictor (hereafter, 'ANPP-Precipitation' model).
 We fit the model under a Bayesian framework, allowing us to test for treatment differences by comparing the posterior distributions of the treatment-level coefficients [e.g., @Tredennick2013].
@@
 
-We fit the model using a Bayesian approach, obtaining posterior estimates of all unkowns via the No-U-Turn Hamiltonian Monte Carlo sampler in Stan [@stan2016].
+Unfortunately, we cannot detect whether treatments grow stronger over time in the ANPP-Precipitation model described above because year and precipitation are perfectly colinear.
+Therefore, we fit simpler model with log(ANPP) as the response and treatment effects for drought and irrigation (hereafter, 'ANPP-Treatment' model). 
+We again use a hierarchical structure, where the treatment effects were allowed to vary by year (i.e., year random effects).
+This allowed us to test whether treatment effects got stronger over time.
+We also included a plot random effect to account for non-independence of observations through time in a plot.
+This model is also fully described in Appendix 2.
+
+We fit both models using a Bayesian approach, obtaining posterior estimates of all unkowns via the No-U-Turn Hamiltonian Monte Carlo sampler in Stan [@stan2016].
 We used the R package 'rstan' [@rstan2016] to link R [@R2016] to Stan.
@@
 In other words, the posterior distributions of $\beta_0(\text{control}) - \beta_0(\text{drought})$ and $\beta_0(\text{control}) - \beta_0(\text{drought})$ broadly overlapped zero.
-There was also no evidence that the treatment effects became more important over time because differences between average treatment and control plots were centered on zero regardless of year of the experiment (Fig.
+There was also no evidence that the treatment effects became more important over time because there was no directional trend in the yearly treatment effects estimated from the ANPP-Treatment model (Fig.
 3).
@@
       \includegraphics[width=6in]{../figures/glmm_treatment_posteriors.png}
-  \caption{(A) Relationship between ANPP and growing season precipitation over the course of the experiment.
-Regression lines are indpendent linear fits for each treatment for visual clarity of the mean trends.
-(B) Posterior medians (points), 80\% BCIs (heavy lines), and 95\% BCIs (light lines) of effects from the fitted repeated measures generalized linear mixed-effects model for each treatment-control comparison.
-BCI refers to 'Bayesian Credible Interval', which is the upper and lower quantiles of the posterior distribution at the specified level.}
+  \caption{Posterior distributions of treatment-level parameters ('Intercept' and the effect of 'Precipitation') from the ANPP-Precipitation model.}
 \end{figure}
@@
   \centering
-      \includegraphics[width=4in]{../figures/glmm_yeardiffs.png}
-  \caption{Posterior medians (points), 80\% BCIs (heavy lines), and 95\% BCIs (light lines) of the difference between log(ANPP) in control and treatment plots over time.
-BCI refers to 'Bayesian Credible Interval', which are the upper and lower quantiles of the posterior distribution at the specified level.}
+      \includegraphics[width=6in]{../figures/glmm_yeardiffs.png}
+  \caption{Posterior distributions of yearly treatment effects from the ANPP-Treatment model.}
 \end{figure}

### 2017-06-09 code/plot_model_results.R

@@ -98,8 +98,8 @@ ggplot(filter(betas, param_name!="Control"), aes(x=estimate))+
   xlab("Parameter Value")+
   ylab("Probability Density")+
   theme_few()+
-  theme(legend.position = c(0.4,0.8),
-        legend.key.size = unit(6,"pt"),
+  theme(legend.position = c(0.4,0.7),
+        legend.key.size = unit(10,"pt"),
         legend.title = element_text(size=10),
         legend.text = element_text(size = 8),
         legend.key.height = unit(0.8,"line"))

### 2017-06-12 manuscript/Appendix3.Rmd
adds stuff on hrv and extreme years

---
layout: 11pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \onehalfspacing
  - \usepackage{rotating}
  - \usepackage{color, soul}
  - \usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{tikz}
  - \usepackage{bm,mathrsfs}
  - \usepackage[sc]{mathpazo}
bibliography: ~/Dropbox/Bibliography/usses_water.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
    number_sections: true
fontsize: 11pt
geometry: margin=1in
linkcolor: black
urlcolor: black
---

\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{A3-\arabic{figure}}
\renewcommand\theequation{A3.\arabic{equation}}
\renewcommand\thetable{A3-\arabic{table}}  
\renewcommand\thesection{Section A3.\arabic{section}}

\begin{center}
\textbf{\Large{Appendix 3}} \\
A.T. Tredennick, A.R. Kleinhesselink, \& P.B. Adler \\
``Ecosystem and community resistance...'' \\
\emph{PeerJ}
\end{center}

\section{Characterizing Extreme Precipitation Amounts}

Following the proposed methods of @Lemoine2016, we calculated quantiles from the empirical distribution of growing season precipitation at Dubios, ID.
We chose the 1\% quantile to be indicative of extreme dry conditions (drought) and the 99\% quantile to be indicative of extreme wet conditions (irrigation).
The data consist of 91 yearly records, which we assume are approximately normally distributed.
The R code below shows our procedure, and Fig. A3-1 shows the results.

```{r hrv, fig.cap="Density of the empirical distribution of growing season precipitation at Dubois, ID. Dashed vertical lines show the 1% and 99% quantiles, assuming a normal distribution."}
library(ggplot2)
weather    <- read.csv("../data/weather/ClimateIPM.csv")
mean_ppt   <- mean(weather$ppt1)
quants_ppt <- quantile(weather$ppt1,probs = c(0.01,0.99))
quants_ppt[1]/mean_ppt*100 # percent of mean ppt for drought
quants_ppt[2]/mean_ppt*100 # percent of mean ppt for irrigation

ggplot(weather, aes(x=ppt1))+
  geom_histogram(bins=20, color="dodgerblue", fill="dodgerblue", aes(y=..density..))+
  geom_line(stat="density", color="blue")+
  geom_vline(aes(xintercept=quants_ppt[1]), linetype=2)+
  geom_vline(aes(xintercept=quants_ppt[2]), linetype=2)+
  ylab("Density")+
  xlab("Growing Season Precipitation (mm)")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())
```

# References

### 2017-06-12 manuscript/usses_water_ms.Rmd
adds stuff on hrv and extreme years

@@
 ----------
-We manipulated growing season precipitation for five years in a sagebrush steppe to reach the extremes of the historical distribution of precipitation (Fig. 1A).
-In so doing, we aimed to understand how this ecosystem might respond to periods of relatively novel precipitation regimes.
-Our results suggest that ecosystem functioning, as measured by annual net primary productivity, and community composition are not sensitive to the precipitation manipulations we imposed.
+Ecosystem response to chronic resource alteration is expected to follow a temporal trend.
+Initially, ecosystem response will be modest and reflect the physiological responses of constituent species.
+Over longer time periods, species reordering will cause greater responses as species better able to take advantage of, or cope with, new resource levels become more abundant.
+This temporal trend is formalized by the 'Hierarchical Response Framework' [HRF, @Smith2009], and has been empirically supported [@Knapp2012; @Wilcox2016].
+A lingering question is not whether the HRF is a reasonable model of ecosystem dynamics, but rather how the time scales of ecosystem response and community change differ among ecosystems.
+Indeed, previous work has shown that community compositional shifts can be both rapid, on the order of years [@Hoover2014], and slow, on order of decades [@Knapp2012; @Wilcox2016].
+To add to this growing body of knowledge, we performed a precipitation manipulation experiment in a sagebrush steppe ecosystem.
+
+Surprisingly, ANPP was not sensitive to either drought or irrigation treatments, whether in the context of growing season precipitation (Fig.
+2) or in the context of year since the experiment began (Fig.
+3).
+This is surprising because grasslands generally, and sagebrush steppe specifically, are considered water-limited systems.
+Thus, our expectation was that ANPP would be tightly linked to soil water availability.
+ANPP was slightly lower in drought plots and slightly higher in irrigation plots, but these differences were not statistically important.
+
+In the absence of community composition data, an obvious hypothesis to explain the lack of ANPP response is that species reordered rapidly.
+In so doing, newly abundant species could compensate for the loss of ecosystem functioning of previously abundant species.
+We did not find this to be the case, as specie composition was remarkably stable over the course of the experiment (Fig. 3).
+
+In combination, the lack of ecosystem and community response to drought or irrigation shows that this sagebrush steppe ecosystem is resistant to chronic alterations in water availability.
+There are three possible explanations for the resistance we found.
+First, it could be that our experiment simply was not long enough to induce responses.
+This may be true for community responses, which can take decades [@Wilcox2016], but the response of grassland ANPP to water manipulations is typically immediate [e.g., @Hoover2014].
+
+Second, it could be that our manipulations were not large enough to induce a response.
+That is, maybe a 50\% decrease/increase in any given year is not abnormal give our sites historical range of variability [@Knapp2017].
+We cannot definitively rule out this possibility, but we have reason to believe our manipulations *should* have been large enough.
+Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of mean growing season precipitation neccesary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site.
+The 1\% quantile of precipitation at our site is 110 mm, a 47% reduction from the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean growing season precipitation (Appendix 3).
+Thus, our drought treatment represented extreme precipitation amounts, especially in years where ambient precipitation was below average (Fig. 1A).
+The irrigation treatment may have been too small, however.
+
+Third, the ecosystem and community resistance to drought and deluge we observed could be a real phenomenon in this system.
 

### 2017-06-12 code/look_at_weather.R

@@ -202,3 +202,30 @@ anpp_summary <- anpp_data %>%
   group_by(year) %>%
   summarise(mean_anpp = mean(anpp))
 write.csv(anpp_summary, "../results/avg_anpp_by_year.csv")
+
+
+
+####
+####  CHARACTERIZE HRV OF PRECIP ----
+####
+weather <- read.csv("../data/weather/ClimateIPM.csv")
+mean_ppt <- mean(weather$ppt1)
+quants_ppt <- quantile(weather$ppt1,probs = c(0.01,0.99))
+quants_ppt[1]/mean_ppt*100
+quants_ppt[2]/mean_ppt*100
+
+ggplot(weather, aes(x=ppt1))+
+  geom_histogram(bins=20, color="dodgerblue", fill="dodgerblue", aes(y=..density..))+
+  geom_line(stat="density", color="blue")+
+  geom_vline(aes(xintercept=quants_ppt[1]), linetype=2)+
+  geom_vline(aes(xintercept=quants_ppt[2]), linetype=2)+
+  ylab("Density")+
+  xlab("Growing Season Precipitation (mm)")+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank())
+
+
+
+
+
+

### 2017-06-13 code/community_dynamics.R
attempting new community dynamics analysis for issue 3

@@ -267,12 +267,48 @@ annual_community_matrix <- read.csv("../data/vital_rates/idaho_annuals.csv") %>%
   separate(species,c("genus","species1")) %>%
   unite(species,genus,species1,sep = " ") %>%
   spread(species,abundance, fill = 0)%>%
-  select(-QuadName,-Grazing,-paddock,-Group) %>%
+  select(-QuadName,-Grazing,-paddock,-Group,-`Lappula occidentalis`,-`Tragopogon dubius`) %>%
   ungroup()
 
 full_community_matrix <- cover_community_matrix %>%
   left_join(density_community_matrix) %>%
   left_join(annual_community_matrix)
+
+# set missing species abundances as 0, which == absent
+full_community_matrix[is.na(full_community_matrix)] <- 0 
+
+full_community_matrix_scaled <- cbind(full_community_matrix[,1:3,drop=F], 
+                                      scale(full_community_matrix[,4:ncol(full_community_matrix)]))
+
+spp_to_remove <- names(which(is.nan(colMeans(full_community_matrix_scaled[4:ncol(full_community_matrix_scaled)]))))
+ids_to_remove <- which(colnames(full_community_matrix_scaled) %in% spp_to_remove)
+full_community_matrix_scaled <- full_community_matrix_scaled %>%
+  select(-ids_to_remove)
+
+
+nmds_df <- {}
+pdf("../figures/rda_raw.pdf",onefile = TRUE)
+for(doyr in unique(full_community_matrix_scaled$year)){
+  tmp <- filter(full_community_matrix_scaled, year == doyr) %>% select(-year)
+  tmp_rda <- rda(tmp[3:ncol(tmp)])
+  nmds_points <- as.data.frame(summary(tmp_rda)$sites[,1:2])
+  nmds_points$Treatment <- tmp$Treatment
+  nmds_points$Year <- doyr
+  nmds_df <- rbind(nmds_df, nmds_points)
+  plot(tmp_rda, main=doyr)
+}
+dev.off()
+
+ggplot(nmds_df, aes(x=PC1, y=PC2, color=Treatment))+
+  geom_point()+
+  geom_point(color="grey35",shape=1)+
+  scale_color_brewer(palette = "Set2")+
+  ylab("PCA 2")+
+  xlab("PCA 1")+
+  facet_wrap("Year")+
+  theme_few()
+ggsave(paste0(figure_path,"sppcomp_rda.png"), width=6, height = 4, units = "in", dpi = 120)
+
 
 
 

### 2017-06-13 manuscript/Appendix1.Rmd
trivia
adds wavelength info for radiometer

@@
 In each year we recorded ground reflectances at four bands, two associated with the red spectrum and two associated with the near-infrared spectrum (Table A1-1).
-We took four readings per plot that were averaged.
+We took four readings per plot that were averaged for each band.
 Bands 1 and 3 correspond to wavelengths collected by the MODIS satellite and bands 2 and 4 correspond to wavelengths collected by the AVHRR satellite.
 
-| Band number | Spectrum name | Spectrum wavelengths | Corresponding satellite |
-| ----------- | ------------- | -------------------- | ----------------------- |
-| 1 | red | xx-xx $\mu$m | AVHRR |
-| 2 | red | xx-xx $\mu$m | MODIS |
-| 3 | near-infrared | xx-xx $\mu$m | AVHRR |
-| 4 | near-infrared | xx-xx $\mu$m | MODIS |
+| Band number | Spectrum name | Center wavelength | Corresponding satellite |
+| ----------- | ------------- | ----------------- | ----------------------- |
+| 1 | red | 626 nm | AVHRR |
+| 2 | red | 652 nm | MODIS |
+| 3 | near-infrared | 875 nm | AVHRR |
+| 4 | near-infrared | 859 nm | MODIS |
 

### 2017-06-15 code/cache/community_dynamics_old.R
cleans up comm dynamics script and uses full scaled abundance matrix for rank clocks

@@ -1,0 +1,457 @@
+##  usses_stability.R: R script to analyze the temporal stability of plots at
+##    USSES in Idaho. Plots received irrigation/drought treatments starting
+##    in 2011.
+##
+##  Author:       Andrew Tredennick
+##  Date created: April 24, 2017
+##
+##  UPDATES  ###################################################################
+##
+##
+##  ############################################################################
+
+##  Clear the workspace
+rm(list=ls(all.names = TRUE))
+
+
+
+####
+####  LOAD PACKAGES ----
+####
+library(tidyverse)
+library(dplyr)
+library(stringr)
+library(ggthemes)
+library(ggalt)
+library(broom)
+library(RColorBrewer)
+library(viridis)
+library(lme4)
+library(codyn)
+library(vegan)
+library(gridExtra)
+library(car)
+
+
+
+####
+####  SET DIRECTORIES; READ IN DATA ----
+####
+# Set working dir to source file location
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only works in RStudio
+
+# Set paths
+biomass_path <- "../data/estimated_biomass/"
+weather_path <- "../data/weather/"
+output_path  <- "../output/"
+figure_path  <- "../figures/"
+
+# Read in data
+plot_biomass <- readRDS(paste0(biomass_path,"permanent_plots_estimated_biomass.RDS"))
+weather_data <- read.csv(paste0(weather_path,"ClimateIPM.csv"))
+plot_info    <- read.csv("../data/estimated_biomass/quad_info.csv")
+
+cover_species <- read.csv("../data/vital_rates/allrecords_cover_v24.csv") %>%
+  group_by(quad, year, species) %>%
+  summarise(total_area = sum(area)) %>%
+  left_join(plot_info) %>%
+  filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
+  filter(!str_detect(QuadName, 'P1|P7')) %>%
+  group_by(year, species, Treatment) %>%
+  summarise(avg_area = mean(total_area)) %>%
+  ungroup()
+
+
+
+####
+####  RANK CLOCKS (COVER) ----
+####
+domspp <- c("Artemisia tripartita", "Poa secunda", "Pseudoroegneria spicata", "Hesperostipa comata")
+cover_ts <- cover_species %>%
[…]

### 2017-06-15 code/community_dynamics.R

@@ -4,42 +4,30 @@
 ##
 ##  Author:       Andrew Tredennick
 ##  Date created: April 24, 2017
-##
-##  UPDATES  ###################################################################
-##
-##
-##  ############################################################################
 
 ##  Clear the workspace
 rm(list=ls(all.names = TRUE))
+
+##  Set working dir to source file location
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only for RStudio
 
 
 
 ####
 ####  LOAD PACKAGES ----
 ####
-library(tidyverse)
-library(dplyr)
-library(stringr)
-library(ggthemes)
-library(ggalt)
-library(broom)
-library(RColorBrewer)
-library(viridis)
-library(lme4)
-library(codyn)
-library(vegan)
-library(gridExtra)
-library(car)
+library(tidyverse) # all the good things from Wikham
+library(dplyr)     # data munging
+library(stringr)   # working with strings
+library(ggthemes)  # pleasing ggplot2 themes
+library(viridis)   # color color scheme
+library(vegan)     # community similarity metrics
 
 
 
 ####
 ####  SET DIRECTORIES; READ IN DATA ----
 ####
-# Set working dir to source file location
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only works in RStudio
-
 # Set paths
 biomass_path <- "../data/estimated_biomass/"
 weather_path <- "../data/weather/"
@@ -51,28 +39,148 @@ plot_biomass <- readRDS(paste0(biomass_path,"permanent_plots_estimated_biomass.R
 weather_data <- read.csv(paste0(weather_path,"ClimateIPM.csv"))
 plot_info    <- read.csv("../data/estimated_biomass/quad_info.csv")
 
-cover_species <- read.csv("../data/vital_rates/allrecords_cover_v24.csv") %>%
-  group_by(quad, year, species) %>%
-  summarise(total_area = sum(area)) %>%
-  left_join(plot_info) %>%
-  filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
-  filter(!str_detect(QuadName, 'P1|P7')) %>%
-  group_by(year, species, Treatment) %>%
-  summarise(avg_area = mean(total_area)) %>%
-  ungroup()
[…]

### 2017-06-15 manuscript/Appendix1.Rmd
updates calibration info to deal with scling factors

@@
 We calculated NDVI as:
-
+\vspace{-2em}
 \begin{align}
-\text{NDVI}_{\text{AVHRR}} &= \frac{b_3\delta_{1(\text{AVHRR})} - b_1\delta_{2(\text{AVHRR})}}{b_3\delta_{1(\text{AVHRR})} + b_1\delta_{2(\text{AVHRR})}} \\[10pt]
-\text{NDVI}_{\text{MODIS}} &= \frac{b_4\delta_{1(\text{MODIS})} - b_2\delta_{2(\text{MODIS})}}{b_4\delta_{1(\text{MODIS})} + b_2\delta_{2(\text{MODIS})}}
+\text{NDVI}_{\text{AVHRR}} &= \frac{b_3 - b_1}{b_3 + b_1} \\[10pt]
+\text{NDVI}_{\text{MODIS}} &= \frac{b_4 - b_2}{b_4 + b_2}
 \end{align}
 
-\noindent{} where $b_x$ refers to band *x* (*x* = 1,2,3,4) in Table A1-1 and $\delta$s are scaling factors unique to each band.
-The values for the scaling factors come from \textcolor{red}{(PETER?)} and are as follows: $\delta_{1(AVHRR)} = 0.77$, $\delta_{2(AVHRR)} = 1$, $\delta_{1(MODIS)} = 0.95$, $\delta_{2(MODIS)} = 0.96$.
+\noindent{} where $b_x$ refers to band *x* (*x* = 1,2,3,4) in Table A1-1.
 

### 2017-06-15 manuscript/usses_water_ms.Rmd
updates calibration info to deal with scling factors

@@
   \centering
-      \includegraphics[width=5in]{../figures/nmds_comp_combined.png}
-  \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot, colored by treatment.
-(A) NMDS results using basal cover data; (B) NMDS results using individual density data.
-2011 is a pre-treatment year.}
+      \includegraphics[width=5in]{../figures/sppcomp_bray_all.png}
+  \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot, colored by treatment.}
 \end{figure}
@@
   \centering
-      \includegraphics[width=5in]{../figures/rank_clocks_cover.png}
-  \caption{Rank clocks of average species' basal cover by treatment. (A) All species. (B) Dominant species. Cover is log-transformed to improve visualization of species with low cover.}
+      \includegraphics[width=5in]{../figures/rank_clocks_scaled_abundance.png}
+  \caption{Rank clocks of average species' scaled abundance by treatment. (A) All species. (B) Dominant species.}
 \end{figure}

### 2017-06-15 code/calibrate_radiometer_by_year.R

@@ -37,6 +37,9 @@ library(stringr)   # tidy tools for strings
 ####
 ####  FUNCTIONS TO FIT AND OUTPUT CALIBRATION REGRESSSION ----
 ####
+# NDVI calculations include "sens_*" scaling factors because we did not
+# use the diffuser on the radiometer. These scaling factors are based on the
+# factory calibration tests and specifications.
 calculate_ndvi <- function(radiometer_data, col_starts = "ch", 
                            red_num = 2, nir_num = 4, 
                            sens_red = 0.96, sens_nir = 0.95, digits = 2){

### 2017-06-20 code/anpp_randcoefs_model.R
adds random year effect to anpp-vwc model

@@ -47,11 +47,13 @@ fit_stan_model <- function(model_data, check_diags=FALSE, treattype){
                   Nplots = length(unique(model_data$quadname)),
                   Ntreats = length(unique(model_data$Treatment)),
                   Nppts = nrow(newx),
+                  Nyears = length(unique(model_data$year)),
                   y = as.numeric(scale(log(model_data$anpp))),
                   x = x,
                   newx = newx,
                   plot_id = as.numeric(as.factor(model_data$quadname)),
                   treat_id = as.numeric(as.factor(as.character(model_data$Treatment))),
+                  year_id = as.numeric(as.factor(model_data$year)),
                   R = diag(1,ncol(x)))
   
   rstan_options(auto_write = TRUE)

### 2017-06-20 manuscript/usses_water_ms.Rmd
updates figures; caches some stuff

@@
       \includegraphics[width=6in]{../figures/glmm_treatment_posteriors.png}
-  \caption{Posterior distributions of treatment-level parameters ('Intercept' and the effect of 'Precipitation') from the ANPP-Precipitation model.}
+  \caption{Posterior distributions of treatment-level parameters ('Intercept' and the effect of 'Soil Moisture').
+Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.}
 \end{figure}
 
+<!---
 \newpage{}
@@
   \caption{Posterior distributions of yearly treatment effects from the ANPP-Treatment model.}
-\end{figure}
+\end{figure}\
+--->
 

### 2017-06-20 code/anpp_randcoefs_model.R

@@ -61,7 +61,8 @@ fit_stan_model <- function(model_data, check_diags=FALSE, treattype){
   rt <- stanc("anpp_randcoefs.stan")
   sm <- stan_model(stanc_ret = rt, verbose=FALSE)
   set.seed(123)
-  fit <- sampling(sm, data=anppdat, iter=2000, chains=4, thin=2,  control = list(adapt_delta = 0.99))
+  fit <- sampling(sm, data=anppdat, iter=10000, chains=4, 
+                  thin=10, control = list(adapt_delta = 0.99))
   
   if(check_diags){
     pnames <- c("sigmaeps", "sigmaint", "sigmaslope","beta[1]",

### 2017-06-20 code/look_at_weather.R

@@ -116,7 +116,9 @@ soil_vwc <- ggplot(soil_moisture, aes(x=julian_date, y=VWC, group=Treatment, col
   scale_y_continuous(breaks=seq(0,24,8))+
   facet_grid(year~.)+
   theme_bw()+
-  theme(panel.grid.minor = element_blank(),strip.background = element_blank(), strip.text = element_text(size=6))+
+  theme(panel.grid.minor = element_blank(),
+        strip.background = element_blank(), 
+        strip.text = element_text(size=6))+
   guides(color=FALSE)
 
 

### 2017-06-20 code/plot_model_results.R

@@ -29,7 +29,7 @@ library(viridis)
 ####
 source("read_format_data.R") # load data
 all_fit <- readRDS("../results/randcoefs_alltreatments_fit.RDS")
-year_fit <- readRDS("../results/randyears_alltreatments_fit.RDS")
+# year_fit <- readRDS("../results/randyears_alltreatments_fit.RDS")
 
 
 
@@ -48,7 +48,7 @@ get_one_tailed <- function(values){
 ####  PLOT TREATMENT-LEVEL POSTERIOR DISTRIBUTIONS ----
 ####
 param_id_names <- data.frame(param_id = c(1,2),
-                             param_name = c("Intercept", "Precipitation"))
+                             param_name = c("Intercept", "Soil Moisture"))
 treat_id_names <- data.frame(treat_id = c(1,2,3),
                              treatment = c("Control","Drought","Irrigation"))
 betas <- reshape2::melt(rstan::extract(all_fit, pars="beta_treat")) %>%
@@ -60,17 +60,21 @@ g1 <- ggplot(betas, aes(x=estimate))+
   geom_hline(aes(yintercept=0), color="grey45", size=0.1)+
   geom_line(stat="density",aes(color=treatment),adjust=5,size=1)+
   scale_color_brewer(palette = "Set2",name="Treatment")+
-  scale_x_continuous(breaks=seq(-1,2,0.5))+
+  scale_x_continuous(breaks=seq(-1.5,2,0.5))+
   scale_y_continuous(breaks=seq(0,2,0.25))+
   facet_wrap(~param_name)+
   xlab("Parameter Value")+
   ylab("Probability Density")+
-  theme_few()+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank())+
   theme(legend.position = c(0.4,0.8),
         legend.key.size = unit(6,"pt"),
         legend.title = element_text(size=10),
         legend.text = element_text(size = 8),
-        legend.key.height = unit(0.8,"line"))
+        legend.key.height = unit(0.8,"line"),
+        legend.box.background = element_rect(),
+        strip.background = element_blank(), 
+        strip.text = element_text(size=10))
 ggsave("../figures/glmm_treatment_posteriors.png",plot = g1, width = 6, height = 3, units = "in", dpi =120)
 
 

### 2017-06-21 manuscript/Appendix2.Rmd
updates model description with random year effects

@@
 The random effects structure mirrors the nested structure of the data: plot-level parameters are nested within treatment-level parameters, which are nested within overall parameters.
-In what follows, $\textbf{X}$ is the design matrix including a column of 1s for intercepts and a column of precipitation values for each observation, $\textbf{y}$ is a row vector of log(ANPP) values for each observation, *i* references observations, *j* references plots, and *k* references treatment.
-The notation *i(j(k))* reads as 'observation *i* associated with plot *j* associated with treatment *k*'.
+In what follows, $\textbf{X}$ is the design matrix including a column of 1s for intercepts and a column of precipitation values for each observation, $\textbf{y}$ is a row vector of log(ANPP) values for each observation, *i* references observations, *j* references plots, *k* references treatment, and *t* references year.
+The notation *i(j(k(t)))* reads as 'observation *i* associated with plot *j* associated with treatment *k* within year *t*'.
 Recall we have three treatments: drought (~50\% decrease in precipitation), control (ambient precipitation), and irrigation (~50\% increase in precipitation).
@@
 \begin{align}
-y_{i(j(k))} \sim \text{Normal}\left(\mu_{i(j(k))}, \sigma^2_k \right),
+y_{i(j(k(t)))} \sim \text{Normal}\left(\mu_{i(j(k(t)))}, \sigma^2_k \right),
 \end{align}
 
-\noindent{}where $\mu_{i(j(k))}$ is the determinstic expectation from the regression model,
+\noindent{}where $\mu_{i(j(k(t)))}$ is the determinstic expectation from the regression model,
 
@@
 \begin{align}
-\mu_{i(j(k))} = \textbf{x}_i'\boldsymbol{\beta_{j(k)}}.
+\mu_{i(j(k(t)))} = \textbf{x}_i'\boldsymbol{\beta_{j(k)} + \gamma_{t}}.
 \end{align}
 
-Regression parameters (the intercept and slope in the parameter vector $\boldsymbol{\beta}$) were modeled hierarchically,
+The intercept and slope in the parameter vector $\boldsymbol{\beta}$ were modeled hierarchically,
 
@@
 
+The random year effects ($\boldsymbol{\gamma}$) are modeled as intercept offsets centered on zero with a shared variance ($\sigma_{yr}$),
+
+\vspace{-2em}
+
+\begin{align}
+\boldsymbol{\gamma} \sim \text{Normal} \left( 0, \sigma_{yr}  \right).
+\end{align}
+
 We fit the model using MCMC as implemented in the software Stan [@stan2016].
@@
   int<lower=0> Nppts;         # Number of precip levels to predict
+  int<lower=0> Nyears;        # Number of years
   vector[Nobs] y;             # vector of observations
@@
   int treat_id[Nobs];         # vector of treatment ids
+  int year_id[Nobs];          # vector of year ids
 }
@@
 	vector[Npreds] beta_mu; # overall coefficients
+	vector[Nyears] year_off; # vector of year effects
 	cov_matrix[Npreds] Sigma; # covariance matrix for treatment-level coefficients
-	cov_matrix[Npreds] Sigma_plot[Ntreats]; # unique covariance matrix for plot-level coefs by treatment
+	cov_matrix[Npreds] Sigma_plot[Ntreats]; # unique covariance matrix for plot-level coefficients for each treatment
 	vector<lower=0>[Ntreats] sd_y; # treatment-level observation std. dev.
+	vector<lower=0>[Nyears] sigma_year; # year std. dev. hyperprior
 }
@@
     # regression model for expected values (one for each plot-year)
-    yhat[i] = x[i]*beta_plot[plot_id[i]]; 
+    yhat[i] = x[i]*beta_plot[plot_id[i]] + year_off[year_id[i]];
 }
@@
 	  beta_treat[i] ~ multi_normal(beta_mu, Sigma);
-
-	  # priors on covariance of effects at plot-level
-	  Sigma_plot[i] ~ inv_wishart(Npreds+1, R);
+	  Sigma_plot[i] ~ inv_wishart(Npreds+1, R); # priors on covariance of effects at plot-level
 	}
@@
 	beta_mu ~ normal(0,1); # priors on overall effects
+	year_off ~ normal(0,sigma_year); # priors on year effects, shared variance
 	Sigma ~ inv_wishart(Npreds+1, R); # priors on covariance of effects at treatment-level
 	sd_y ~ weibull(2,1); # priors on observation std. dev. for each treatment
+	sigma_year ~ weibull(2,1); # hyperprior on year std. dev.
 

### 2017-06-21 manuscript/usses_water_ms.Rmd
updates stats in ms

@@
 \begin{align}
-\mu_{i(j(k))} &= \beta_{0,j(k)} + \beta_{1,j(k)}x_i, \\
-y_{i(j(k))} &\sim \text{Normal} \left(\mu_{i(j(k))}, \sigma^2_{k} \right),
+\mu_{i(j(k(t)))} &= \beta_{0,j(k)} + \beta_{1,j(k)}x_i + \gamma_t, \\
+y_{i(j(k(t)))} &\sim \text{Normal} \left(\mu_{i(j(k(t)))}, \sigma^2_{k} \right),
 \end{align}
 
-\noindent{} where $\mu_{i(j(k))}$ is the deterministic prediction from the regression model for observation *i* for plot *j* associated with treatment *k*, $\beta_{0,j(k)}$ is the intercept for plot *j* associated with treatment *k*, $\beta_{1,j(k)}$ is the slope term for the effect of precipitation for plot *j* associated with treatment *k*, and $\sigma^2_k$ is the process variance for treatment *k*.
-Data include the standardized log(ANPP) observations ($y_{i(j(k))}$) and precipitation ($x_i$).
+\noindent{} where $\mu_{i(j(k))}$ is the deterministic prediction from the regression model for observation *i* for plot *j* associated with treatment *k* in year *t*, $\beta_{0,j(k)}$ is the intercept for plot *j* associated with treatment *k*, $\beta_{1,j(k)}$ is the slope term for the effect of precipitation for plot *j* associated with treatment *k*, $\gamma_t$ is the intercept offset for year *t*, and $\sigma^2_k$ is the process variance for treatment *k*.
+Data include the standardized log(ANPP) observations ($y_{i(j(k(t)))}$) and precipitation ($x_i$).
 Although we include observation subscript *i* on the *x*s, observations within a year all share the same precipitation values.
@@
 
+<!---
 Unfortunately, we cannot detect whether treatments grow stronger over time in the ANPP-Precipitation model described above because year and precipitation are perfectly colinear.
@@
 This model is also fully described in Appendix 2.
+---->
 
-We fit both models using a Bayesian approach, obtaining posterior estimates of all unkowns via the No-U-Turn Hamiltonian Monte Carlo sampler in Stan [@stan2016].
+We fit the model using a Bayesian approach, obtaining posterior estimates of all unkowns via the No-U-Turn Hamiltonian Monte Carlo sampler in Stan [@stan2016].
 We used the R package 'rstan' [@rstan2016] to link R [@R2016] to Stan.
@@
 
-Growing season precipitation had a positive effect on ANPP (mean of $\beta_{1}$ = 0.67; 80\% BCI = 0.24, 1.10; 95\% BCI = -0.11, 1.34) (Fig. 1D).
+Cumulative growing season soil moisture had a positive effect on ANPP (mean of $\beta_{1}$ = 0.43; 80\% BCI = -0.05, 0.88; 95\% BCI = -0.36, 1.15) (Fig. 1D).
 Average ANPP was similar among treatments (similar intercepts, Fig. 2A), as was the effect of precipitation (similar slopes, Fig. 2B).
 In an average precipitation year (i.e., $x$ = 0 in Eq.
-1), the probability that ANPP in a drought plot is less than ANPP in a control plot was 0.58, and the probability that ANPP in an irrigation plot is higher than in a control plot was 0.57.
+1), the probability that ANPP in a conrol plot is greater than ANPP in a drought plot was 0.71, and the probability that ANPP in a control plot is lower than in an irrigation plot was 0.41.
 In other words, the posterior distributions of $\beta_0(\text{control}) - \beta_0(\text{drought})$ and $\beta_0(\text{control}) - \beta_0(\text{drought})$ broadly overlapped zero.
@@
 
-Community composition was similar among treatments, whether looking at basal cover or density (Fig. 4).
+Community composition was similar among treatments.
+In no year did community composition among treatments not overlap, and they were equally dispersed in all years (Fig. 4).
 Likewise, community composition was remarkably stable over time, with no evidence of divergence among treatments (Fig. 4).

### 2017-06-21 code/plot_model_results.R

@@ -82,32 +82,32 @@ ggsave("../figures/glmm_treatment_posteriors.png",plot = g1, width = 6, height =
 ####
 ####  PLOT TREATMENT DIFFERENCES BY YEAR POSTERIORS ----
 ####
-param_id_names <- data.frame(param_id = c(1:3),
-                             param_name = c("Control", "Drought", "Irrigation"))
-year_id_names <- data.frame(year_id = c(1:5),
-                            year_name = c(1:5))
-betas <- reshape2::melt(rstan::extract(year_fit, pars="beta")) %>%
-  rename(iteration = iterations, year_id = Var2, param_id = Var3, estimate = value, stan_name = L1) %>%
-  left_join(param_id_names, by="param_id") %>%
-  left_join(year_id_names, by="year_id")
-
-ggplot(filter(betas, param_name!="Control"), aes(x=estimate))+
-  geom_vline(aes(xintercept=0), linetype=2, color="grey45")+
-  geom_line(stat="density",aes(color=as.character(year_id)),adjust=5,size=1)+
-  # scale_color_brewer(palette = "Set1",name="Year")+
-  scale_color_viridis(end=0.8,discrete=T,name="Year")+
-  scale_x_continuous(breaks=seq(-3,3,0.5), limits=c(-2,2))+
-  scale_y_continuous(breaks=seq(0,1.25,0.25))+
-  facet_wrap(~param_name)+
-  xlab("Parameter Value")+
-  ylab("Probability Density")+
-  theme_few()+
-  theme(legend.position = c(0.4,0.7),
-        legend.key.size = unit(10,"pt"),
-        legend.title = element_text(size=10),
-        legend.text = element_text(size = 8),
-        legend.key.height = unit(0.8,"line"))
-ggsave("../figures/glmm_yeardiffs.png", width = 6, height = 3, units = "in", dpi =120)
+# param_id_names <- data.frame(param_id = c(1:3),
+#                              param_name = c("Control", "Drought", "Irrigation"))
+# year_id_names <- data.frame(year_id = c(1:5),
+#                             year_name = c(1:5))
+# betas <- reshape2::melt(rstan::extract(year_fit, pars="beta")) %>%
+#   rename(iteration = iterations, year_id = Var2, param_id = Var3, estimate = value, stan_name = L1) %>%
+#   left_join(param_id_names, by="param_id") %>%
[…]

### 2017-06-21 code/read_format_data.R

@@ -5,12 +5,12 @@
 ##  Date created: May 25, 2017
 
 
-####
-####  LOAD LIBRARIES ----
-####
-library(tidyverse)    # Data munging
-library(dplyr)        # Data summarizing
-library(stringr)      # Working with strings
+# ####
+# ####  LOAD LIBRARIES ----
+# ####
+# library(tidyverse)    # Data munging
+# library(dplyr)        # Data summarizing
+# library(stringr)      # Working with strings
 
 
 

### 2017-06-22 manuscript/usses_water_ms.Rmd
trivia

@@
 We estimated aboveground net primary productivity (ANPP) using a radiometer to relate ground reflectance to plant biomass [see @Byrne2011 for a review].
-We recorded ground reflectance at four wavelengths, two associated with red reflectance (xxx and xxx) and two associated with near-infrared reflectance (xxx and xxx).
+We recorded ground reflectance at four wavelengths, two associated with red reflectance (626 nm and 652 nm) and two associated with near-infrared reflectance (875 nm and 859 nm).
 At each plot in each year, we took four readings of ground reflectances at the above wavelengths.
@@
 (C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.
-(D) Scatterplot of ANPP versus growing season precipitation.}
+(D) Scatterplot of ANPP versus growing season precipitation.
+Colored regression lines are independently-fit linear models for each treatment with no random effects structure; dark black regression line is a linear regression through all the point with no random effects structure.
+Our analysis seeks to find if the data supports a model with different intercepts and/or slopes for each treatment, while also accounting for nonindependence of samples within plots across years.}
 \end{figure}

### 2017-06-22 manuscript/usses_water_ms.Rmd
adds info on standardizing species abundances

@@
 We coupled this temporally intensive soil moisture sampling with spatially extensive readings taken at six points within all 16 plots and associated ambient measurement areas.
-These snapshot data were collected on 06/06/2012, 04/29/2015, 05/07/2015, 06/09/2015, and 05/10/2016 using a handheld EC-5 sensor.
-
-\textcolor{red}{[ANDY: paragraph here on SOILWAT. Is this the right thing to show/do? See Fig. 1C.]}
+These snapshot data were collected on 06/06/2012, 04/29/2015, 05/07/2015, 06/09/2015, and 05/10/2016 using a handheld EC-5 sensor.\todo{ANDY: add paragraph on VWC model}
 
@@
 
-Species composition data came from annual census maps for each plot made using a pantograph [@Hill1920].
-The maps record the spatial location and size of each individual plant.
+\todo{PETER: check this section}Species composition data came from two sources: yearly census maps for each plot made using a pantograph [@Hill1920] and yearly counts of annual species in each plot.
+The maps record the spatial location of all individuals in the plot and the basal cover of each individual with cover greater than 1 cm.
 Using those annual maps, we aggregated over individuals to calculate total basal cover for each species in each plot.
+We made a large plot-treatment-year by species matrix, where columns were filled with either basal cover or density, depending on the measurement made for the particular species.
+So we could analyze the different types of data together, we standardized the values in each column.
+This puts all abundance values on the same scale, but comes with the limitation that all species are weighted equally.
+Nonetheless, this scaling approach allows a comprehensive view of community composition dynamics through time.
 
@@
 Our main goal was to test whether the relationship between ANPP and growing season precipitation (hereafter, precipiation) differed among the drought, control, and irrigation treatments.
-To achieve this goal, we fit a multi-level random intercept and random slope regression with log(ANPP) as the response variabile and precipitation as the sole predictor (hereafter, 'ANPP-Precipitation' model).
+To achieve this goal, we fit a multi-level random intercept and random slope regression with log(ANPP) as the response variabile and soil moisture (volumetric water content) as the sole predictor.
 We fit the model under a Bayesian framework, allowing us to test for treatment differences by comparing the posterior distributions of the treatment-level coefficients [e.g., @Tredennick2013].
-Both log(ANPP) and precipitation were standardized to have mean 0 and unit variance before fitting the model [i.e., $(x_i - \bar{x})/\sigma_x$].
+Both log(ANPP) and soil moisture were standardized to have mean 0 and unit variance before fitting the model [i.e., $(x_i - \bar{x})/\sigma_x$].
 
@@
 
-\noindent{} where $\mu_{i(j(k))}$ is the deterministic prediction from the regression model for observation *i* for plot *j* associated with treatment *k* in year *t*, $\beta_{0,j(k)}$ is the intercept for plot *j* associated with treatment *k*, $\beta_{1,j(k)}$ is the slope term for the effect of precipitation for plot *j* associated with treatment *k*, $\gamma_t$ is the intercept offset for year *t*, and $\sigma^2_k$ is the process variance for treatment *k*.
-Data include the standardized log(ANPP) observations ($y_{i(j(k(t)))}$) and precipitation ($x_i$).
-Although we include observation subscript *i* on the *x*s, observations within a year all share the same precipitation values.
+\noindent{} where $\mu_{i(j(k))}$ is the deterministic prediction from the regression model for observation *i* for plot *j* associated with treatment *k* in year *t*, $\beta_{0,j(k)}$ is the intercept for plot *j* associated with treatment *k*, $\beta_{1,j(k)}$ is the slope term for the effect of soil moisture for plot *j* associated with treatment *k*, $\gamma_t$ is the intercept offset for year *t*, and $\sigma^2_k$ is the process variance for treatment *k*.
+Data include the standardized log(ANPP) observations $\left( y_{i(j(k(t)))} \right)$ and soil moisture ($x_i$).
+Although we include observation subscript *i* on the *x*s, observations within a treatment-year all share the same soil moisture values.
 
-The intercept and slope terms are modeled hierarchichally to account for the non-independence of observations across years within plots and to allow us to test the hypothesis that our treatments alter the ANPP-precipitation relationship.
-As noted above, plot-level coefficients are drawn from treatment-level coeffecients, which are drawn from overall coefficients.
+The intercept and slope terms are modeled hierarchichally to account for the non-independence of observations across years within plots and to allow us to test the hypothesis that our treatments alter the ANPP-soil moisture relationship.
+As noted above, plot-level coefficients are drawn from treatment-level coefficients, which are drawn from overall coefficients.
 We also include a covariance structure among the intercept and slope at each level.
@@
 The overall coefficients are drawn from a normal prior with mean 0 and standard deviation 1.
+The random year effects ($\boldsymbol{\gamma}$) are drawn from a normal prior with mean 0 and standard deviation $\sigma_{\text{year}}$, which was drawn from a weibull distribution.
 A full description of model is in Appendix 2.
@@
 For each year of the experiment, we first calculated Bray-Curtis distances among all plots, and then extracted those distances for use in the NMDS.
+Because we standardized species' abundances, some values were negative, which is not allowed for calculating Bray-Curtis distances.
+We simply added '2' to each abundance value to ensure all values were greater than zero.
 We plotted the first two axes of NMDS scores to see if community composition overlapped, or not, among treatments in each year.
-We used functions in the R package 'vegan' [@Oksanen2016] to calculate Bray-Curtis distances and then to run the NMDS analysis.
+We used the 'metaMDS()' function in the R package 'vegan' [@Oksanen2016] to calculate Bray-Curtis distances and then to run the NMDS analysis.
 Lastly, we examined rank clocks of species' abundances through time to assess the stability of community composition over the course of the experiment [@Collins2008].
-Rank clocks were made using 'ggplot2' [@Wickham2009] and R code from @Hallett2016.
+Rank clocks were made using 'ggplot2' [@Wickham2009] and R code from the 'codyn' package [@Hallett2016].
 

### 2017-06-22 manuscript/usses_water_ms.Rmd
updates year effects plots and caption

@@
 We thank the many summer research technicians who collected the data reported in this paper and the US Experimental Sheep Station for facilitating work on their property.
-We also thank Susan Durham for clarifying our thinking on the statistical analyses.
+We also thank Susan Durham for clarifying our thinking on the statistical analyses and Kevin Wilcox for helpful discussions on analyzing community composition data.
 
@@
       \includegraphics[height=7in]{../figures/Figure1.png}
-  \caption{(A) Histogram of historical precipitation from 1926-2016, with the years of the experiment overlaid as colored points.
-``Drought'' treatments (squares) are shown as a 50\% reduction from the observed precipitation.
-``Irrigation'' treatment precipitation levels are not shown.
-(B) Monthly average estimated soil volumetric water content (VWC) from SOILWAT model fit using soil moisture data from experimental plots.
+  \caption{(A) Probability density of historical precipitation from 1926-2016, with the years of the experiment shown with arrows on the *x*-axis.
+(B) Observed soil volumetric water content (VWC) over the course of the experiment.
 (C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.
@@
 Colored regression lines are independently-fit linear models for each treatment with no random effects structure; dark black regression line is a linear regression through all the point with no random effects structure.
-Our analysis seeks to find if the data supports a model with different intercepts and/or slopes for each treatment, while also accounting for nonindependence of samples within plots across years.}
+Our analysis seeks to find if the data supports a model with different intercepts and/or slopes for each treatment, while also accounting for nonindependence of samples within plots across years.
+Color mapping for panels B-D is shown in the legend for panel C.}
 \end{figure}
@@
 
-<!---
 \newpage{}
@@
       \includegraphics[width=6in]{../figures/glmm_yeardiffs.png}
-  \caption{Posterior distributions of yearly treatment effects from the ANPP-Treatment model.}
-\end{figure}\
---->
+  \caption{Posterior distributions of random year effects (intercept offsets).}
+\end{figure}
 

### 2017-06-22 code/plot_model_results.R

@@ -82,32 +82,33 @@ ggsave("../figures/glmm_treatment_posteriors.png",plot = g1, width = 6, height =
 ####
 ####  PLOT TREATMENT DIFFERENCES BY YEAR POSTERIORS ----
 ####
-# param_id_names <- data.frame(param_id = c(1:3),
-#                              param_name = c("Control", "Drought", "Irrigation"))
-# year_id_names <- data.frame(year_id = c(1:5),
-#                             year_name = c(1:5))
-# betas <- reshape2::melt(rstan::extract(year_fit, pars="beta")) %>%
-#   rename(iteration = iterations, year_id = Var2, param_id = Var3, estimate = value, stan_name = L1) %>%
-#   left_join(param_id_names, by="param_id") %>%
-#   left_join(year_id_names, by="year_id")
-# 
-# ggplot(filter(betas, param_name!="Control"), aes(x=estimate))+
-#   geom_vline(aes(xintercept=0), linetype=2, color="grey45")+
-#   geom_line(stat="density",aes(color=as.character(year_id)),adjust=5,size=1)+
-#   # scale_color_brewer(palette = "Set1",name="Year")+
-#   scale_color_viridis(end=0.8,discrete=T,name="Year")+
-#   scale_x_continuous(breaks=seq(-3,3,0.5), limits=c(-2,2))+
-#   scale_y_continuous(breaks=seq(0,1.25,0.25))+
-#   facet_wrap(~param_name)+
-#   xlab("Parameter Value")+
-#   ylab("Probability Density")+
-#   theme_few()+
-#   theme(legend.position = c(0.4,0.7),
-#         legend.key.size = unit(10,"pt"),
-#         legend.title = element_text(size=10),
-#         legend.text = element_text(size = 8),
-#         legend.key.height = unit(0.8,"line"))
-# ggsave("../figures/glmm_yeardiffs.png", width = 6, height = 3, units = "in", dpi =120)
+param_id_names <- data.frame(param_id = c(1:3),
+                             param_name = c("Control", "Drought", "Irrigation"))
+year_id_names <- data.frame(year_id = c(1:5),
+                            year_name = c(1:5))
+betas <- reshape2::melt(rstan::extract(all_fit, pars="year_off")) %>%
+  rename(iteration = iterations, year_id = Var2, estimate = value, stan_name = L1) %>%
[…]

### 2017-06-22 manuscript/usses_water_ms.Rmd
adds text on stats for species comp

@@
 We used the 'metaMDS()' function in the R package 'vegan' [@Oksanen2016] to calculate Bray-Curtis distances and then to run the NMDS analysis.
+We used the 'vegan::adonis()' function [@Oksanen2016] to perform permutational multivariate analysis of variance to test whether treatment plots formed distinct groupings.
+To test whether treatment plots were equally dispersed, or not, we used the 'vegan::betadisper' function [@Oksanen2016].
 Lastly, we examined rank clocks of species' abundances through time to assess the stability of community composition over the course of the experiment [@Collins2008].
@@
 Community composition was similar among treatments.
-In no year did community composition among treatments not overlap, and they were equally dispersed in all years (Fig. 4).
-Likewise, community composition was remarkably stable over time, with no evidence of divergence among treatments (Fig. 4).
+In no year did community composition among treatments not overlap, and they were equally dispersed in all years (Table 1; Fig. 4).
+Likewise, community composition was remarkably stable over time, with no evidence of divergence among treatments (Table 1; Fig. 4).
 Species' abundances and ranks showed little deviation over the five-year experiment, regardless of treatment (Fig. 5).
@@
 ---------------
-We gratefully acknowledge the support of the Utah Agricultural Experiment Station (journal paper xxxx).
+We gratefully acknowledge the support of the Utah Agricultural Experiment Station (journal paper xxxx\todo{get paper number}).
 We thank the many summer research technicians who collected the data reported in this paper and the US Experimental Sheep Station for facilitating work on their property.
@@
 
-<!---
 Tables
@@
 library(xtable)
-# Read in within-year ANOVA results table
-# inyear_anova <- readRDS("../results/within_year_anova_table.RDS")
-# 
-# inyear_caption <- "ANOVA for treatment effects on ANPP for each year of the experiment."
-# print(xtable(inyear_anova, caption = inyear_caption, digits = 2),
-#       caption.placement="top",
-#       include.rownames = F, 
-#       sanitize.colnames.function = identity,
-#       comment = FALSE,
-#       size = "normalsize")
+# Read in species composition stats
+sppcomp_stats <- readRDS("../results/sppcomp_stats.RDS")
+colnames(sppcomp_stats) <- c("Year", "Test", "n", "d.f.","Sums","Means","\\emph{F}", "\\emph{P}")
+sppcomp_stats$d.f. <- as.integer(sppcomp_stats$d.f.)
+
+# Make the table
+inyear_caption <- "Results from statistical tests for clustering and dispersion of community composition among precipitation treatments.
+`adonis` tests whether treatments form unique clusters in multidimensial space; `betadisper` tests whether treatments have similar dispersion.
+For both tests, \\emph{P} values greater than 0.05 indicate there is no support that the treatments differ."
+print(xtable(dplyr::select(sppcomp_stats, -Sums, -Means), caption = inyear_caption, digits = 2),
+       caption.placement="top",
+       include.rownames = F,
+       sanitize.colnames.function = identity,
+       comment = FALSE,
+       size = "normalsize")
 
@@
 
---->
-
 Figures
@@
       \includegraphics[height=7in]{../figures/Figure1.png}
-  \caption{(A) Probability density of historical precipitation from 1926-2016, with the years of the experiment shown with arrows on the *x*-axis.
+  \caption{(A) Probability density of historical precipitation from 1926-2016, with the years of the experiment shown with arrows on the \emph{x}-axis.
 (B) Observed soil volumetric water content (VWC) over the course of the experiment.
@@
       \includegraphics[width=6in]{../figures/glmm_yeardiffs.png}
-  \caption{Posterior distributions of random year effects (intercept offsets).}
+  \caption{Posterior distributions of random year effects (intercept offsets). Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.}
 \end{figure}
@@
   \centering
-      \includegraphics[width=5in]{../figures/sppcomp_bray_all.png}
+      \includegraphics[width=6in]{../figures/sppcomp_bray_all.png}
   \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot, colored by treatment.}
@@
   \centering
-      \includegraphics[width=5in]{../figures/rank_clocks_scaled_abundance.png}
+      \includegraphics[width=6in]{../figures/rank_clocks_scaled_abundance.png}
   \caption{Rank clocks of average species' scaled abundance by treatment. (A) All species. (B) Dominant species.}

### 2017-06-23 manuscript/usses_water_ms.Rmd
trivia and log regressions for fig 1

@@
 We imposed drought and irrigation treatments (50\% decrease/increase) and measured ecosystem (ANPP) and community (species composition) responses.
-We focus on how the drought and irrigation treatments affect the linear relationship between ANPP and growing season precipitation, and if community dynamics underly the ecosystem responses.
-In particular, we sought to answer the following questions:
-
-1. Does drought or irrigation impact mean ANPP (i.e., different intercepts)?
-2. Does drought or irrigation impact the effect of ambient precipitation on ANPP (i.e., different slopes)?
-3. Is the resistance or sensitivity of ANPP to altered precipitation related to underlying community dynamics?
+We focus on how the drought and irrigation treatments affect the linear relationship between ANPP and available soil moisture, and if community dynamics underly the ecosystem responses.
+In particular, we are interested in the consistency of the ANPP-soil moisture relationship among treatments.
+Is the relationship steeper under the drought treatment, at low soil moisture?
+Does the relationship saturate under the irrigation treatment, at high soil moisture?
+To answer these questions we fit a random intercepts, random slopes model to test whether the regressions differed among treatments.
 

### 2017-06-23 code/look_at_weather.R

@@ -129,8 +129,10 @@ soil_vwc <- ggplot(soil_moisture, aes(x=julian_date, y=VWC, group=Treatment, col
 source("read_format_data.R") # load data
 data_plot <- ggplot(anpp_data, aes(x=total_seasonal_vwc,y=anpp))+
   geom_point(shape=21,color="grey25",alpha=0.8,aes(fill=Treatment))+
-  stat_smooth(method="lm", aes(color=Treatment), se=FALSE, size=0.5)+
-  stat_smooth(method="lm", color="black", se=FALSE, size=0.7)+
+  #stat_smooth(method="lm", aes(color=Treatment), se=FALSE, size=0.5)+
+  stat_smooth(method="glm", formula=y~x, method.args=list(family=quasi(link='log')),aes(color=Treatment), se=FALSE, size=0.5)+
+  #stat_smooth(method="lm", color="black", se=FALSE, size=0.7)+
+  stat_smooth(method="glm", formula=y~x, method.args=list(family=quasi(link='log')),color="black", se=FALSE, size=0.5)+
   scale_fill_brewer(palette = "Set2")+
   scale_color_brewer(palette = "Set2")+
   #scale_x_continuous(limits=c(100,300),breaks=seq(100,300,50))+

### 2017-07-06 manuscript/usses_water_ms.Rmd
adds outcomes figure
tries new title...

@@
 
-\Large{\textbf{Ecosystem and community resistance to five years of drought and deluge in a sagebrush steppe}}
+\Large{\textbf{Consistent ecosystem functional response to soil moisture during five years of drought and deluge in a sagebrush steppe}}
 
@@
   \centering
+      \includegraphics[width=4in]{../figures/hypothesis_figtable.png}
+  \caption{Possible outcomes of chronic resource alteration.}
+\end{figure}
+
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
       \includegraphics[height=7in]{../figures/Figure1.png}
@@
 
+<!---
 \newpage{}
@@
 \end{figure}
+---->
 

### 2017-07-10 manuscript/usses_water_ms.Rmd
start of intro rewrite

@@
 ------------
+Water availability is a strong driver of aboveground net primary productivity (ANPP) in grassland ecosystems.
+Mean annual precipitation (MAP) is strongly correlated with ANPP across sites [@Huxman2004], and, in water-limited ecosystems, interannual variation in MAP is also correlated with interannual variation in ANPP [@Hsu2012].
+These relationships suggest that chronic alterations in levels of precipitation, associated with global climate change, will impact ANPP.
+
+At a given site, the functional response of ANPP to water availability (e.g., soil moisture) can be characterized by fitting a model to historical observations of ANPP and soil moisture.
+However, the fitted functional response may provide an incomplete picture because future conditions are likely to be outside the historical range of variability [@Smith2011].
+For example, historical trends may show a linear soil moisture-ANPP relationship, but that relationship could potentially saturate if soil moisture is pushed far beyond typical levels.
+Saturating relationships are actually common [@Hsu2012], perhaps because other resources, like nitrogen, become more limiting in wet years than dry years.
+Knowing the curvature of the soil moisture-ANPP relationship is critical for understanding how ecosystems will respond to chronic alterations in water availability.
+
+Manipulating potentially limiting resources offers a route to understanding how ecosystems will respond to resource levels that fall outside the historical range of variability [@Knapp2017].
+Chronic alterations to resource availability should cause community composition to shift over time, the focal ecosystem is sensitive to the manipulated resource.
+Species' relative abundances are expected to re-order and, eventually, some species will be lost from the local community and new species will be gained [@Smith2009; @Avolio2015].
+On the heels of changes in community composition, subsequent changes in the level of ecosystem functioning may occur.
+
+@Smith2009 introduced the 'Hierarchical Response Framework' for understanding the interplay of community composition and ecosystem functioning in response to resource manipulations over time.
+In the near term, ecosystem functioning such as annual net primary productivity (ANPP) will reflect the physiological responses of individual species to the manipulated resource level.
+For example, ANPP may decline under simulated drought because the initial community consisted of drought-intolerant species [@Hoover2014].
+Over longer time spans, ecosystem functioning may recover as new species colonize or initial species reorder in relative abundance.
+For example, ANPP may initially decline, but eventually rise back to pre-treatment levels once drought-tolerant species become more abundant and compensate for drought-intolerant species [@Hoover2014].
+It is also possible that ecosystem functioning shifts to a new mean state, reflective of the suite of species in the new community [@Knapp2012].
+
+Here we report the results of a five-year precipitation manipulation experiment in a sagebrush steppe grassland.
+We imposed drought and irrigation treatments (50\% decrease/increase) and measured ecosystem (ANPP) and community (species composition) responses.
+We focus on how the drought and irrigation treatments affect the relationship between ANPP and available soil moisture, and if community dynamics underly the ecosystem responses.
+In particular, we are interested in the consistency of the ANPP-soil moisture relationship among treatments.
+Is the relationship steeper under the drought treatment, at low soil moisture?
+Does the relationship saturate under the irrigation treatment, at high soil moisture?
+To answer these questions we fit a random intercepts, random slopes model to test whether the regressions differed among treatments.
+
+<!---
 As the rate of climate change accelerates, ecologists are being pressed to forecast its impacts on ecosystem functions and services.
@@
 Thus, discovering which ecosystems are most sensitive to chronic alterations of water availability, and the time scales at which ecosystem and community responses occur, is critical to avoid surprises in the face of climate change.
-
-Here we report the results of a five-year precipitation manipulation experiment in a sagebrush steppe grassland.
-We imposed drought and irrigation treatments (50\% decrease/increase) and measured ecosystem (ANPP) and community (species composition) responses.
-We focus on how the drought and irrigation treatments affect the linear relationship between ANPP and available soil moisture, and if community dynamics underly the ecosystem responses.
-In particular, we are interested in the consistency of the ANPP-soil moisture relationship among treatments.
-Is the relationship steeper under the drought treatment, at low soil moisture?
-Does the relationship saturate under the irrigation treatment, at high soil moisture?
-To answer these questions we fit a random intercepts, random slopes model to test whether the regressions differed among treatments.
+--->
 

### 2017-07-11 manuscript/usses_water_ms.Rmd
updates ms to peerJ formatting using rticles

@@
 ---
-layout: 11pt
-header-includes:
-  - \usepackage{lineno}
-  - \usepackage{setspace}
-  - \usepackage{todonotes}
-  - \linespread{1.25}
-  - \usepackage{rotating}
-  - \usepackage{color, soul}
-  - \usepackage{sectsty}
-  - \usepackage{bm,mathrsfs}
-  - \usepackage[sc]{mathpazo}
+title: Consistent ecosystem functional response during five years of drought and deluge in a sagebrush steppe
+preprint: true
+author: 
+  - name: Andrew T. Tredennick
+    affiliation: 1
+    corresponding: true
+    email: atredenn@gmail.com
+  - name: Andrew R. Kleinhesselink
+    affiliation: 2
+  - name: Peter B. Adler
+    affiliation: 1
+affiliation:
+  - code: 1
+    address: Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322
+  - code: 2
+    address: Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California xxxxx
+abstract: >
+  Precipitation is predicted to become more variable in the western U.S., meaning years of above and below average precipitation will become more common.
+Such periods of drought and deluge could become major drivers of plant community dynamics and ecosystem functioning in water limited grasslands.
+Here we report the results of a five-year experiment where we used drought and irrigation treatments (50% decrease/increase) to see how a sagebrush steppe plant community in Idaho will respond to future climate changes.
+The ecosystem was suprisingly resistant to both drought and irrigation.
+Aboveground net primary productivity (ANPP) responded positively to ambient growing season precipitation, but the response did not vary across treatments.
+There was also no evidence that treatment effects grew over time.
+The similarity of ecosystem functioning was not due to compensatory shifts at the plant community level, where species composition among treatments was similar and remarkably stable over the five years.
+At least in the short-term, ecosystem functioning and community composition in this sagebrush steppe system is resistant to increases and decreases in growing season precipitation.
+header-includes: >
+  \usepackage{lipsum}
+  \usepackage{setspace}
+  \usepackage{todonotes}
+  \usepackage{rotating}
+  \usepackage{color, soul}
+  \usepackage{sectsty}
+  \usepackage{bm,mathrsfs}
 bibliography: ~/Dropbox/Bibliography/usses_water.bib
-csl: components/ecology.csl
-
-## rmarkdown render options
 output: 
-  pdf_document:
-    fig_caption: true
-    keep_tex: false
-fontsize: 11pt
-geometry: margin=1in
-linkcolor: black
-urlcolor: black
+  rticles::peerj_article:
+    base_format: rmarkdown::pdf_document # bookdown::pdf_document2 # for using \@ref()
 ---
 
-\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
 \definecolor{blue}{rgb}{0,0,0.7} 
@@
 \newcommand{\ark}{\textcolor{red}}
-
-\begin{singlespace}
-
-\begin{centering}
-
-\Large{\textbf{Consistent ecosystem functional response to soil moisture during five years of drought and deluge in a sagebrush steppe}}
-
-\bigskip{} \bigskip{}
-
-\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
-
-\normalsize{Andrew T. Tredennick$^{1,}$\footnote{Corresponding author; e-mail: atredenn@gmail.com}, Andrew R. Kleinhesselink$^{2}$, and Peter B. Adler$^{1}$}
-
-\bigskip{}
-
-\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah}} \\
-\textit{\small{\textsuperscript{2}Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California}}
-
-\bigskip{}
-
-Last compile: \today
-
-\end{centering}
-
-\vspace{3em}
-
-\end{singlespace}
 
@@
 
-\begin{abstract}
-Precipitation is predicted to become more variable in the western U.S., meaning years of above and below average precipitation will become more common.
-Such periods of drought and deluge could become major drivers of plant community dynamics and ecosystem functioning in water limited grasslands.
-Here we report the results of a five-year experiment where we used drought and irrigation treatments (50\% decrease/increase) to see how a sagebrush steppe plant community in Idaho will respond to future climate changes.
-The ecosystem was suprisingly resistant to both drought and irrigation.
-Aboveground net primary productivity (ANPP) responded positively to ambient growing season precipitation, but the response did not vary across treatments.
-There was also no evidence that treatment effects grew over time.
-The similarity of ecosystem functioning was not due to compensatory shifts at the plant community level, where species composition among treatments was similar and remarkably stable over the five years.
-At least in the short-term, ecosystem functioning and community composition in this sagebrush steppe system is resistant to increases and decreases in growing season precipitation.
+# Introduction
 
-\vspace{2em}
-
-\emph{Key words: aboveground net primary productivity, climate change, community dynamics, drought, ecosystem resistance, sagebrush steppe}
-\end{abstract}
-
-\setlength{\parindent}{5ex}
-
-\linenumbers
-
-Introduction
-------------
 Water availability is a strong driver of aboveground net primary productivity (ANPP) in grassland ecosystems.
@@
 Thus, discovering which ecosystems are most sensitive to chronic alterations of water availability, and the time scales at which ecosystem and community responses occur, is critical to avoid surprises in the face of climate change.
---->
+---->
 
-Materials and Methods
----------------------
-### Study Area
+# Materials and Methods
+
+## Study Area
 We conducted our precipitation manipulation experiment at the United States Sheep Experimental Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
@@
 
-### Precipitation Experiment
+## Precipitation Experiment
 Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
@@
 We coupled this temporally intensive soil moisture sampling with spatially extensive readings taken at six points within all 16 plots and associated ambient measurement areas.
-These snapshot data were collected on 06/06/2012, 04/29/2015, 05/07/2015, 06/09/2015, and 05/10/2016 using a handheld EC-5 sensor.\todo{ANDY: add paragraph on VWC model}
+These snapshot data were collected on 06/06/2012, 04/29/2015, 05/07/2015, 06/09/2015, and 05/10/2016 using a handheld EC-5 sensor.
 
-### Data Collection
+## Data Collection
 We estimated aboveground net primary productivity (ANPP) using a radiometer to relate ground reflectance to plant biomass [see @Byrne2011 for a review].
@@
 
-\todo{PETER: check this section}Species composition data came from two sources: yearly census maps for each plot made using a pantograph [@Hill1920] and yearly counts of annual species in each plot.
+Species composition data came from two sources: yearly census maps for each plot made using a pantograph [@Hill1920] and yearly counts of annual species in each plot.
 The maps record the spatial location of all individuals in the plot and the basal cover of each individual with cover greater than 1 cm.
@@
 
-### Data Analysis
+## Data Analysis
 Our main goal was to test whether the relationship between ANPP and growing season precipitation (hereafter, precipiation) differed among the drought, control, and irrigation treatments.
@@
 
-\vspace{-3em}
+\vspace{-2em}
 
@@
 
-<!---
-Unfortunately, we cannot detect whether treatments grow stronger over time in the ANPP-Precipitation model described above because year and precipitation are perfectly colinear.
-Therefore, we fit simpler model with log(ANPP) as the response and treatment effects for drought and irrigation (hereafter, 'ANPP-Treatment' model). 
-We again use a hierarchical structure, where the treatment effects were allowed to vary by year (i.e., year random effects).
-This allowed us to test whether treatment effects got stronger over time.
-We also included a plot random effect to account for non-independence of observations through time in a plot.
-This model is also fully described in Appendix 2.
----->
-
 We fit the model using a Bayesian approach, obtaining posterior estimates of all unkowns via the No-U-Turn Hamiltonian Monte Carlo sampler in Stan [@stan2016].
@@
 
-Results
--------
+# Results
+
 Three of our five treatment years fell in years of below average rainfall (Fig. 1A).
@@
 
-Discussion
-----------
+# Discussion
+
 Ecosystem response to chronic resource alteration is expected to follow a temporal trend.
@@
 
-Acknowledgments
----------------
-We gratefully acknowledge the support of the Utah Agricultural Experiment Station (journal paper xxxx\todo{get paper number}).
+# Acknowledgments
+
+We gratefully acknowledge the support of the Utah Agricultural Experiment Station (journal paper xxxx).
 We thank the many summer research technicians who collected the data reported in this paper and the US Experimental Sheep Station for facilitating work on their property.
@@
 
-Funding
--------
+# Funding
+
 NSF DBI-1400370 to Andrew Tredennick.  
@@
 
-\singlespace
-\setlength{\parindent}{0ex}
-
 \newpage{}
 
-Tables
-------
+# Tables
 
@@
 
-Figures
--------
+# Figures
 
@@
   \centering
-      \includegraphics[height=7in]{../figures/Figure1.png}
+      \includegraphics[height=5in]{../figures/Figure1.png}
   \caption{(A) Probability density of historical precipitation from 1926-2016, with the years of the experiment shown with arrows on the \emph{x}-axis.
@@
   \centering
-      \includegraphics[width=6in]{../figures/glmm_treatment_posteriors.png}
+      \includegraphics[width=5in]{../figures/glmm_treatment_posteriors.png}
   \caption{Posterior distributions of treatment-level parameters ('Intercept' and the effect of 'Soil Moisture').
@@
   \centering
-      \includegraphics[width=6in]{../figures/glmm_yeardiffs.png}
+      \includegraphics[width=5in]{../figures/glmm_yeardiffs.png}
   \caption{Posterior distributions of random year effects (intercept offsets). Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.}
@@
   \centering
-      \includegraphics[width=6in]{../figures/sppcomp_bray_all.png}
+      \includegraphics[width=5in]{../figures/sppcomp_bray_all.png}
   \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot, colored by treatment.}
@@
 <!---
-\newpage{}
-
 \begin{figure}[!ht]
@@
 
-References
-----------
+# References

### 2017-07-11 manuscript/usses_water_ms.tex
updates ms to peerJ formatting using rticles

--- 2017-05-05 manuscript/usses_water_ms.tex
@@
-\documentclass[12pt,]{article}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
-\hypersetup{unicode=true,
-            colorlinks=true,
-            linkcolor=black,
-            citecolor=Blue,
-            urlcolor=black,
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
+%% Submissions for peer-review must enable line-numbering
+%% using the lineno option in the \documentclass command.
+%%
+%% Preprints and camera-ready submissions do not need
+%% line numbers, and should have this option removed.
+%%
+%% Please note that the line numbering option requires
+%% version 1.1 or newer of the wlpeerj.cls file, and
+%% the corresponding author info requires v1.2
+
+\documentclass[fleqn,10pt]{wlpeerj} % for preprint submissions
+
+% ZNK -- Adding headers for pandoc
+
+\setlength{\emergencystretch}{3em} 
+\providecommand{\tightlist}{
 \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
+\usepackage{lipsum}
+\usepackage[unicode=true]{hyperref}
+\usepackage{longtable}
 
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
+\usepackage{lipsum} \usepackage{setspace} \usepackage{todonotes}
+\usepackage{rotating} \usepackage{color, soul} \usepackage{sectsty}
+\usepackage{bm,mathrsfs}
 
-%%% Change title format to be more compact
-\usepackage{titling}
+\title{Consistent ecosystem functional response during five years of drought
+and deluge in a sagebrush steppe}
 
-% Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
+\author[1]{Andrew T. Tredennick}
 
-\setlength{\droptitle}{-2em}
-  \title{}
-  \pretitle{\vspace{\droptitle}}
-  \posttitle{}
-  \author{}
-  \preauthor{}\postauthor{}
-  \date{}
-  \predate{}\postdate{}
+\corrauthor[1]{Andrew T. Tredennick}{\href{mailto:atredenn@gmail.com}{\nolinkurl{atredenn@gmail.com}}}
+\author[2]{Andrew R. Kleinhesselink}
 
-\usepackage{lineno}
-\linenumbers
-\usepackage{setspace}
-\usepackage{todonotes}
-\doublespacing
-\usepackage{rotating}
-\usepackage{color, soul}
-\usepackage[font={normalsize},labelfont={bf},labelsep=quad]{caption}
-\usepackage{sectsty}
-\usepackage{bm,mathrsfs}
-\usepackage{mathptmx}
+\author[1]{Peter B. Adler}
+
+\affil[1]{Department of Wildland Resources and the Ecology Center, Utah State
+University, Logan, Utah 84322}
+\affil[2]{Department of Ecology and Evolutionary Biology, University of
+California, Los Angeles, Los Angeles, California xxxxx}
+
+%
+% \author[1]{First Author}
+% \author[2]{Second Author}
+% \affil[1]{Address of first author}
+% \affil[2]{Address of second author}
+% \corrauthor[1]{First Author}{f.author@email.com}
+
+% 
+\begin{abstract}
+Precipitation is predicted to become more variable in the western U.S.,
+meaning years of above and below average precipitation will become more
+common. Such periods of drought and deluge could become major drivers of
+plant community dynamics and ecosystem functioning in water limited
+grasslands. Here we report the results of a five-year experiment where
+we used drought and irrigation treatments (50\% decrease/increase) to
+see how a sagebrush steppe plant community in Idaho will respond to
+future climate changes. The ecosystem was suprisingly resistant to both
+drought and irrigation. Aboveground net primary productivity (ANPP)
+responded positively to ambient growing season precipitation, but the
+response did not vary across treatments. There was also no evidence that
+treatment effects grew over time. The similarity of ecosystem
+functioning was not due to compensatory shifts at the plant community
+level, where species composition among treatments was similar and
+remarkably stable over the five years. At least in the short-term,
+ecosystem functioning and community composition in this sagebrush steppe
+system is resistant to increases and decreases in growing season
+precipitation.
+% Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+\end{abstract}
 
@@
 
-\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
+\flushbottom
+\maketitle
+\thispagestyle{empty}
 
-\definecolor{blue}{rgb}{0,0,0.7} \newcommand{\new}{\textcolor{blue}}
+\definecolor{blue}{rgb}{0,0,0.7} \definecolor{red}{rgb}{0.7,0,0}
+\newcommand{\pba}{\textcolor{blue}} \newcommand{\ark}{\textcolor{red}}
 
-\begin{singlespace}
+\section{Introduction}\label{introduction}
 
-\begin{centering}
+Water availability is a strong driver of aboveground net primary
+productivity (ANPP) in grassland ecosystems. Mean annual precipitation
+(MAP) is strongly correlated with ANPP across sites \citep{Huxman2004},
+and, in water-limited ecosystems, interannual variation in MAP is also
+correlated with interannual variation in ANPP \citep{Hsu2012}. These
+relationships suggest that chronic alterations in levels of
+precipitation, associated with global climate change, will impact ANPP.
 
-\Large{\textbf{Ecosystem and community resistance to six years of drought and deluge in the Sagebrush Steppe}}
+At a given site, the functional response of ANPP to water availability
+(e.g., soil moisture) can be characterized by fitting a model to
+historical observations of ANPP and soil moisture. However, the fitted
+functional response may provide an incomplete picture because future
+conditions are likely to be outside the historical range of variability
+\citep{Smith2011}. For example, historical trends may show a linear soil
+moisture-ANPP relationship, but that relationship could potentially
+saturate if soil moisture is pushed far beyond typical levels.
+Saturating relationships are actually common \citep{Hsu2012}, perhaps
+because other resources, like nitrogen, become more limiting in wet
+years than dry years. Knowing the curvature of the soil moisture-ANPP
+relationship is critical for understanding how ecosystems will respond
+to chronic alterations in water availability.
 
-\bigskip{} \bigskip{}
+Manipulating potentially limiting resources offers a route to
+understanding how ecosystems will respond to resource levels that fall
+outside the historical range of variability \citep{Knapp2017}. Chronic
+alterations to resource availability should cause community composition
+to shift over time, the focal ecosystem is sensitive to the manipulated
+resource. Species' relative abundances are expected to re-order and,
+eventually, some species will be lost from the local community and new
+species will be gained \citep{Smith2009, Avolio2015}. On the heels of
+changes in community composition, subsequent changes in the level of
+ecosystem functioning may occur.
 
-\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
+\citet{Smith2009} introduced the `Hierarchical Response Framework' for
+understanding the interplay of community composition and ecosystem
+functioning in response to resource manipulations over time. In the near
+term, ecosystem functioning such as annual net primary productivity
+(ANPP) will reflect the physiological responses of individual species to
+the manipulated resource level. For example, ANPP may decline under
+simulated drought because the initial community consisted of
+drought-intolerant species \citep{Hoover2014}. Over longer time spans,
+ecosystem functioning may recover as new species colonize or initial
+species reorder in relative abundance. For example, ANPP may initially
+decline, but eventually rise back to pre-treatment levels once
+drought-tolerant species become more abundant and compensate for
+drought-intolerant species \citep{Hoover2014}. It is also possible that
+ecosystem functioning shifts to a new mean state, reflective of the
+suite of species in the new community \citep{Knapp2012}.
 
-\normalsize{Andrew T. Tredennick\textsuperscript{1}, Peter B. Adler\textsuperscript{1}, and Andrew R. Kleinhesselink\textsuperscript{2}}
+Here we report the results of a five-year precipitation manipulation
+experiment in a sagebrush steppe grassland. We imposed drought and
+irrigation treatments (50\% decrease/increase) and measured ecosystem
+(ANPP) and community (species composition) responses. We focus on how
+the drought and irrigation treatments affect the relationship between
+ANPP and available soil moisture, and if community dynamics underly the
+ecosystem responses. In particular, we are interested in the consistency
+of the ANPP-soil moisture relationship among treatments. Is the
+relationship steeper under the drought treatment, at low soil moisture?
+Does the relationship saturate under the irrigation treatment, at high
+soil moisture? To answer these questions we fit a random intercepts,
+random slopes model to test whether the regressions differed among
+treatments.
 
-\bigskip{}
+\section{Materials and Methods}\label{materials-and-methods}
 
-\textit{\small{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, Utah State University, Logan, Utah 84322}}
-\textit{\small{\textsuperscript{2}Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California xxxxx}}
-
-\end{centering}
-
-\vspace{3em}
-
-Last compile: \today
-
-\end{singlespace}
-
-\begin{abstract}
-Summarize it here
-\vspace{2em}
-\end{abstract}
-
-\setlength{\parindent}{5ex}
-
-\subsection{Introduction}\label{introduction}
-
-Start it here
-
-\subsection{Methods}\label{methods}
-
-\subsubsection{Study Area}\label{study-area}
+\subsection{Study Area}\label{study-area}
 
@@
 secunda}, and \emph{Hesperostipa comata}. During the period of our
-experiment, average mean annual precipitation was xxx mm and mean
-monthly temperature ranged from -x\(^{\circ}\)C in January to
-x\(^{\circ}\)C in July.
+experiment (2011 -- 2015), average mean annual precipitation was 265 mm
+year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
+-5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July.
 
-\subsubsection{Precipitation Experiment}\label{precipitation-experiment}
+\subsection{Precipitation Experiment}\label{precipitation-experiment}
 
@@
 this, we used a system of rain-out shelters and automatic irrigation
-(Gherardi \& Sala 2013). The rain-out shelters consisted of transparent
-acrylic shingles 1-1.5 m above the ground that covered an area of 2.5 by
-2 m. The shingles intercepted approximately 50\% of incoming rainfall,
-which was channeled into 75 liter containers. Captured rainfall was then
-pumped out of the containers and sprayed on to the adjacent irrigation
-plot via two suspended sprinklers. Pumping was triggered by float
-switches once water levels reached about 20 liters. We disconnected the
-irrigation pumps each fall and reconnected them, often with difficulty,
-each spring. The rain-out shelters remained in place throughout the
-year.
+\citep{Gherardi2013}. The rain-out shelters consisted of transparent
+acrylic shingles 1-1.5 m above the ground that covered an area of
+\(2.5\times2\) m. The shingles intercepted approximately 50\% of
+incoming rainfall, which was channeled into 75 liter containers.
+Captured rainfall was then pumped out of the containers and sprayed on
+to the adjacent irrigation plot via two suspended sprinklers. Pumping
+was triggered by float switches once water levels reached about 20
+liters. We disconnected the irrigation pumps each fall and reconnected
+them, often with difficulty, each spring. The rain-out shelters remained
+in place throughout the year.
 
@@
 
-\subsubsection{Data Collection}\label{data-collection}
-
-\paragraph{Aboveground Net Primary
-Productivity}\label{aboveground-net-primary-productivity}
+\subsection{Data Collection}\label{data-collection}
 
 We estimated aboveground net primary productivity (ANPP) using a
-radiometer to related ground reflectance to plant biomass (see Byrne et
-al. 2011 for a review).
+radiometer to relate ground reflectance to plant biomass \citep[see][
+for a review]{Byrne2011}. We recorded ground reflectance at four
+wavelengths, two associated with red reflectance (626 nm and 652 nm) and
+two associated with near-infrared reflectance (875 nm and 859 nm). At
+each plot in each year, we took four readings of ground reflectances at
+the above wavelengths. We also took readings in ten calibration plots
+adjacent to the experimental site, in which we harvested all aboveground
+biomass, dried it to a constant weight at 60\(^{\circ}\)C, and weighed
+it to estimate ANPP.
 
-\subsection{Acknowledgments}\label{acknowledgments}
+For each plot and year, we averaged the four readings for each
+wavelength and then calculated NDVI using the MODIS and AVHRR
+algorithms. To convert NDVI to ANPP we regressed NDVI against the dry
+biomass weight from the ten calibration plots. We fit regressions to
+MODIS-based NDVI and AVHRR-based NDVI for each year and retained the
+regression with the better fit. Using the best regression equation for
+each year, we predicted ANPP (Appendix 1).
 
-This work was funded by the National Science Foundation through grants
-DEB-1353078 and DEB-1054040 to PBA, a Graduate Research Fellowship to
-ARK, and a Postdoctoral Resaerch Fellowship in Biology to ATT
-(DBI-1400370). Additional support came from the Utah Agricultural
+Species composition data came from two sources: yearly census maps for
+each plot made using a pantograph \citep{Hill1920} and yearly counts of
+annual species in each plot. The maps record the spatial location of all
+individuals in the plot and the basal cover of each individual with
+cover greater than 1 cm. Using those annual maps, we aggregated over
+individuals to calculate total basal cover for each species in each
+plot. We made a large plot-treatment-year by species matrix, where
+columns were filled with either basal cover or density, depending on the
+measurement made for the particular species. So we could analyze the
+different types of data together, we standardized the values in each
+column. This puts all abundance values on the same scale, but comes with
+the limitation that all species are weighted equally. Nonetheless, this
+scaling approach allows a comprehensive view of community composition
+dynamics through time.
+
+\subsection{Data Analysis}\label{data-analysis}
+
+Our main goal was to test whether the relationship between ANPP and
+growing season precipitation (hereafter, precipiation) differed among
+the drought, control, and irrigation treatments. To achieve this goal,
+we fit a multi-level random intercept and random slope regression with
+log(ANPP) as the response variabile and soil moisture (volumetric water
+content) as the sole predictor. We fit the model under a Bayesian
+framework, allowing us to test for treatment differences by comparing
+the posterior distributions of the treatment-level coefficients
+\citep[e.g.,][]{Tredennick2013}. Both log(ANPP) and soil moisture were
+standardized to have mean 0 and unit variance before fitting the model
+{[}i.e., \((x_i - \bar{x})/\sigma_x\){]}.
+
+Our multi-level model has three grouping levels for coefficients,
+representing the nested structure of the data: (i) overall coefficients,
+(ii) treatment coefficients, and (iii) plot coefficients. Each
+subsequent level is drawn from the distribution of coefficients at the
+previous level. Formally, our model is defined as follows:
+
+\vspace{-2em}
+
+\begin{align}
+\mu_{i(j(k(t)))} &= \beta_{0,j(k)} + \beta_{1,j(k)}x_i + \gamma_t, \\
+y_{i(j(k(t)))} &\sim \text{Normal} \left(\mu_{i(j(k(t)))}, \sigma^2_{k} \right),
+\end{align}
+
+\noindent{} where \(\mu_{i(j(k))}\) is the deterministic prediction from
+the regression model for observation \emph{i} for plot \emph{j}
+associated with treatment \emph{k} in year \emph{t}, \(\beta_{0,j(k)}\)
+is the intercept for plot \emph{j} associated with treatment \emph{k},
+\(\beta_{1,j(k)}\) is the slope term for the effect of soil moisture for
+plot \emph{j} associated with treatment \emph{k}, \(\gamma_t\) is the
+intercept offset for year \emph{t}, and \(\sigma^2_k\) is the process
+variance for treatment \emph{k}. Data include the standardized log(ANPP)
+observations \(\left( y_{i(j(k(t)))} \right)\) and soil moisture
+(\(x_i\)). Although we include observation subscript \emph{i} on the
+\emph{x}s, observations within a treatment-year all share the same soil
+moisture values.
+
+The intercept and slope terms are modeled hierarchichally to account for
+the non-independence of observations across years within plots and to
+allow us to test the hypothesis that our treatments alter the ANPP-soil
+moisture relationship. As noted above, plot-level coefficients are drawn
+from treatment-level coefficients, which are drawn from overall
+coefficients. We also include a covariance structure among the intercept
+and slope at each level. Formally, our hierarhical structure is as
+follows, where we drop the intercept (0) and slope (1) subscripts and
+instead refer to a vector of coefficients, \(\boldsymbol{\beta}\):
+
+\vspace{-2em}
+
+\begin{align}
+\boldsymbol{\beta}_{j(k)} &\sim \text{MVN} \left( \boldsymbol{\beta}_k, \boldsymbol{\Sigma}(k) \right), \\
+\boldsymbol{\beta}_{k} &\sim \text{MVN} \left( \boldsymbol{\beta}, \boldsymbol{\Sigma}  \right), \\
+\boldsymbol{\beta} &\sim \text{Normal} \left( 0, 1 \right),
+\end{align}
+
+\noindent{} where \(\boldsymbol{\beta}_{j(k)}\) is the vector of
+regression coefficients (intercept and slope) for plot \emph{j}
+associated with treatment \emph{k}, \(\boldsymbol{\beta}_{k}\) is the
+vector of coefficients for each treatment, and \(\boldsymbol{\beta}\) is
+the vector of overall coefficients. The plot- and treatment-level
+coefficients are drawn from multivariate normal distributions with
+covariance matrix \(\boldsymbol{\Sigma}\). For the plot-level
+coefficients, each treatment has its own variance-covariance matrix
+(i.e., \(\boldsymbol{\Sigma}(k)\)). The overall coefficients are drawn
+from a normal prior with mean 0 and standard deviation 1. The random
+year effects (\(\boldsymbol{\gamma}\)) are drawn from a normal prior
+with mean 0 and standard deviation \(\sigma_{\text{year}}\), which was
+drawn from a weibull distribution. A full description of model is in
+Appendix 2.
+
+We fit the model using a Bayesian approach, obtaining posterior
+estimates of all unkowns via the No-U-Turn Hamiltonian Monte Carlo
+sampler in Stan \citep{stan2016}. We used the R package `rstan'
+\citep{rstan2016} to link R \citep{R2016} to Stan. We obtained samples
+from the posterior distribution for all model parameters from four
+parallel MCMC chains run for 10,000 iterations, saving every
+\(10^{\text{th}}\) sample. Traceplots of all parameters were visually
+inspected to ensure well-mixed chains and convergence. We also made sure
+all scale reduction factors (\(\hat{R}\)) were less than 1.1.
+
+To see if community composition differed among treatments through time,
+we used non-dimensional multivariate scaling (NMDS) based on Bray-Curtis
+distances. For each year of the experiment, we first calculated
+Bray-Curtis distances among all plots, and then extracted those
+distances for use in the NMDS. Because we standardized species'
+abundances, some values were negative, which is not allowed for
+calculating Bray-Curtis distances. We simply added `2' to each abundance
+value to ensure all values were greater than zero. We plotted the first
+two axes of NMDS scores to see if community composition overlapped, or
+not, among treatments in each year. We used the `metaMDS()' function in
+the R package `vegan' \citep{Oksanen2016} to calculate Bray-Curtis
+distances and then to run the NMDS analysis. We used the
+`vegan::adonis()' function \citep{Oksanen2016} to perform permutational
+multivariate analysis of variance to test whether treatment plots formed
+distinct groupings. To test whether treatment plots were equally
+dispersed, or not, we used the `vegan::betadisper' function
+\citep{Oksanen2016}. Lastly, we examined rank clocks of species'
+abundances through time to assess the stability of community composition
+over the course of the experiment \citep{Collins2008}. Rank clocks were
+made using `ggplot2' \citep{Wickham2009} and R code from the `codyn'
+package \citep{Hallett2016}.
+
+All R code and data necessary to reproduce our analysis has been
+archived on Figshare (\emph{link here after acceptance}) and released on
+GitHub (\url{https://github.com/atredennick/usses_water/releases/v0.1}).
+We also include annotated Stan code in our model description in Appendix
+2.
+
+\section{Results}\label{results}
+
+Three of our five treatment years fell in years of below average
+rainfall (Fig. 1A). Thus, those three years represent a lower magnitude
+of absolute change in precipitation experienced by the treatments.
+Averaged across treatments, ANPP varied from a minimum of 74.5 g
+m\(^{-2}\) in 2014 to a maximum of 237.1 g m\(^{-2}\) in 2016 (Fig. 1C).
+ANPP was slightly higher in irrigation plots and slightly lower in
+drought plots (Fig. 1C), corresponding to estimated soil volumetric
+water content (VWC) differences among treatments (Fig. 1B). Such
+differences in soil VWC indicate our treatment infrastructure was
+successful.
+
+Cumulative growing season soil moisture had a positive effect on ANPP
+(mean of \(\beta_{1}\) = 0.43; 80\% BCI = -0.05, 0.88; 95\% BCI = -0.36,
+1.15) (Fig. 1D). Average ANPP was similar among treatments (similar
+intercepts, Fig. 2A), as was the effect of precipitation (similar
+slopes, Fig. 2B). In an average precipitation year (i.e., \(x\) = 0 in
+Eq. 1), the probability that ANPP in a conrol plot is greater than ANPP
+in a drought plot was 0.71, and the probability that ANPP in a control
+plot is lower than in an irrigation plot was 0.41. In other words, the
+posterior distributions of
+\(\beta_0(\text{control}) - \beta_0(\text{drought})\) and
+\(\beta_0(\text{control}) - \beta_0(\text{drought})\) broadly overlapped
+zero. There was also no evidence that the treatment effects became more
+important over time because there was no directional trend in the yearly
+treatment effects estimated from the ANPP-Treatment model (Fig. 3).
+
+Community composition was similar among treatments. In no year did
+community composition among treatments not overlap, and they were
+equally dispersed in all years (Table 1; Fig. 4). Likewise, community
+composition was remarkably stable over time, with no evidence of
+divergence among treatments (Table 1; Fig. 4). Species' abundances and
+ranks showed little deviation over the five-year experiment, regardless
+of treatment (Fig. 5).
+
+\section{Discussion}\label{discussion}
+
+Ecosystem response to chronic resource alteration is expected to follow
+a temporal trend. Initially, ecosystem response will be modest and
+reflect the physiological responses of constituent species. Over longer
+time periods, species reordering will cause greater responses as species
+better able to take advantage of, or cope with, new resource levels
+become more abundant. This temporal trend is formalized by the
+`Hierarchical Response Framework' \citep[HRF,][]{Smith2009}, and has
+been empirically supported \citep{Knapp2012, Wilcox2016}. A lingering
+question is not whether the HRF is a reasonable model of ecosystem
+dynamics, but rather how the time scales of ecosystem response and
+community change differ among ecosystems. Indeed, previous work has
+shown that community compositional shifts can be both rapid, on the
+order of years \citep{Hoover2014}, and slow, on order of decades
+\citep{Knapp2012, Wilcox2016}. To add to this growing body of knowledge,
+we performed a precipitation manipulation experiment in a sagebrush
+steppe ecosystem.
+
+Surprisingly, ANPP was not sensitive to either drought or irrigation
+treatments, whether in the context of growing season precipitation (Fig.
+2) or in the context of year since the experiment began (Fig. 3). This
+is surprising because grasslands generally, and sagebrush steppe
+specifically, are considered water-limited systems. Thus, our
+expectation was that ANPP would be tightly linked to soil water
+availability. ANPP was slightly lower in drought plots and slightly
+higher in irrigation plots, but these differences were not statistically
+important.
+
+In the absence of community composition data, an obvious hypothesis to
+explain the lack of ANPP response is that species reordered rapidly. In
+so doing, newly abundant species could compensate for the loss of
+ecosystem functioning of previously abundant species. We did not find
+this to be the case, as specie composition was remarkably stable over
+the course of the experiment (Fig. 3).
+
+In combination, the lack of ecosystem and community response to drought
+or irrigation shows that this sagebrush steppe ecosystem is resistant to
+chronic alterations in water availability. There are three possible
+explanations for the resistance we found. First, it could be that our
+experiment simply was not long enough to induce responses. This may be
+true for community responses, which can take decades \citep{Wilcox2016},
+but the response of grassland ANPP to water manipulations is typically
+immediate \citep[e.g.,][]{Hoover2014}.
+
+Second, it could be that our manipulations were not large enough to
+induce a response. That is, maybe a 50\% decrease/increase in any given
+year is not abnormal give our sites historical range of variability
+\citep{Knapp2017}. We cannot definitively rule out this possibility, but
+we have reason to believe our manipulations \emph{should} have been
+large enough. Using the methods described by \citet{Lemoine2016}, we
+calculated the percent reduction and increase of mean growing season
+precipitation neccesary to reach the 1\% and 99\% extremes of the
+historical precipitation regime at our site. The 1\% quantile of
+precipitation at our site is 110 mm, a 47\% reduction from the mean, and
+the 99\% quantile is 414 mm, a 77\% increase from mean growing season
+precipitation (Appendix 3). Thus, our drought treatment represented
+extreme precipitation amounts, especially in years where ambient
+precipitation was below average (Fig. 1A). The irrigation treatment may
+have been too small, however.
+
+Third, the ecosystem and community resistance to drought and deluge we
+observed could be a real phenomenon in this system.
+
+\section{Acknowledgments}\label{acknowledgments}
+
+We gratefully acknowledge the support of the Utah Agricultural
 Experiment Station (journal paper xxxx). We thank the many summer
-research technicians who collected the data reported in this paper.
+research technicians who collected the data reported in this paper and
+the US Experimental Sheep Station for facilitating work on their
+property. We also thank Susan Durham for clarifying our thinking on the
+statistical analyses and Kevin Wilcox for helpful discussions on
+analyzing community composition data.
 
-\singlespace
-\setlength{\parindent}{0ex}
+\section{Funding}\label{funding}
 
-\subsection*{References}\label{references}
-\addcontentsline{toc}{subsection}{References}
+NSF DBI-1400370 to Andrew Tredennick.\\
+NSF Graduate Research Fellowship to Andrew Kleinhesselink.\\
+NSF DEB-1353078 and DEB-1054040 to Peter Adler.
 
-\hypertarget{refs}{}
-\hypertarget{ref-Byrne2011}{}
-Byrne, K.M., Lauenroth, W.K., Adler, P.B. \& Byrne, C.M. (2011).
-Estimating Aboveground Net Primary Production in Grasslands: A
-Comparison of Nondestructive Methods. \emph{Rangeland Ecology \&
-Management}, 64, 498--505.
+\newpage{}
 
-\hypertarget{ref-Gherardi2013}{}
-Gherardi, L.A. \& Sala, O.E. (2013). Automated rainfall manipulation
-system: a reliable and inexpensive tool for ecologists.
-\emph{Ecosphere}, 4, 1--10.
+\section{Tables}\label{tables}
+
+\begin{table}[ht]
+\centering
+\caption{Results from statistical tests for clustering and dispersion of community composition among precipitation treatments.
+`adonis` tests whether treatments form unique clusters in multidimensial space; `betadisper` tests whether treatments have similar dispersion.
+For both tests, \emph{P} values greater than 0.05 indicate there is no support that the treatments differ.}
+\begingroup\normalsize
+\begin{tabular}{rlrrrr}
+  \hline
+Year & Test & n & d.f. & \emph{F} & \emph{P} \\ 
+  \hline
+2011 & adonis &  21 &   2 & 1.02 & 0.45 \\ 
+  2011 & betadisper &  21 &   2 & 2.23 & 0.14 \\ 
+  2012 & adonis &  22 &   2 & 1.10 & 0.29 \\ 
+  2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
+  2013 & adonis &  22 &   2 & 1.23 & 0.14 \\ 
+  2013 & betadisper &  22 &   2 & 0.28 & 0.76 \\ 
+  2014 & adonis &  22 &   2 & 0.95 & 0.58 \\ 
+  2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
+  2015 & adonis &  21 &   2 & 1.05 & 0.37 \\ 
+  2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 
+  2016 & adonis &  21 &   2 & 1.07 & 0.35 \\ 
+  2016 & betadisper &  21 &   2 & 0.50 & 0.62 \\ 
+   \hline
+\end{tabular}
+\endgroup
+\end{table}
+
+\newpage{}
+
+\section{Figures}\label{figures}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=4in]{../figures/hypothesis_figtable.png}
+  \caption{Possible outcomes of chronic resource alteration.}
+\end{figure}
+
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[height=5in]{../figures/Figure1.png}
+  \caption{(A) Probability density of historical precipitation from 1926-2016, with the years of the experiment shown with arrows on the \emph{x}-axis.
+(B) Observed soil volumetric water content (VWC) over the course of the experiment.
+(C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.
+(D) Scatterplot of ANPP versus growing season precipitation.
+Colored regression lines are independently-fit linear models for each treatment with no random effects structure; dark black regression line is a linear regression through all the point with no random effects structure.
+Our analysis seeks to find if the data supports a model with different intercepts and/or slopes for each treatment, while also accounting for nonindependence of samples within plots across years.
+Color mapping for panels B-D is shown in the legend for panel C.}
+\end{figure}
+
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=5in]{../figures/glmm_treatment_posteriors.png}
+  \caption{Posterior distributions of treatment-level parameters ('Intercept' and the effect of 'Soil Moisture').
+Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.}
+\end{figure}
+
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=5in]{../figures/glmm_yeardiffs.png}
+  \caption{Posterior distributions of random year effects (intercept offsets). Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.}
+\end{figure}
+
+\newpage{}
+
+\begin{figure}[!ht]
+  \centering
+      \includegraphics[width=5in]{../figures/sppcomp_bray_all.png}
+  \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot, colored by treatment.}
+\end{figure}
+
+\newpage{}
+
+\bibliography{/Users/atredenn/Dropbox/Bibliography/usses_water.bib}
 

### 2017-07-11 manuscript/usses_water_ms.Rmd
trivia
more intro writing

@@
 title: Consistent ecosystem functional response during five years of drought and deluge in a sagebrush steppe
-preprint: true
+preprint: false
 author: 
@@
 
-# Introduction
+# INTRODUCTION
 
@@
 
-Manipulating potentially limiting resources offers a route to understanding how ecosystems will respond to resource levels that fall outside the historical range of variability [@Knapp2017].
-Chronic alterations to resource availability should cause community composition to shift over time, the focal ecosystem is sensitive to the manipulated resource.
-Species' relative abundances are expected to re-order and, eventually, some species will be lost from the local community and new species will be gained [@Smith2009; @Avolio2015].
-On the heels of changes in community composition, subsequent changes in the level of ecosystem functioning may occur.
-
-@Smith2009 introduced the 'Hierarchical Response Framework' for understanding the interplay of community composition and ecosystem functioning in response to resource manipulations over time.
+Another problem with relying on historical ecoystem functional responses is that they are not static.
+Changes in species identities and abundances can alter an ecosystem's functional response to water availability because different species have different physiological thresholds for producting biomass.
+@Smith2009 introduced the 'Hierarchical Response Framework' (HRF) for understanding the interplay of community composition and ecosystem functioning in response to resource manipulations over time.
 In the near term, ecosystem functioning such as annual net primary productivity (ANPP) will reflect the physiological responses of individual species to the manipulated resource level.
@@
 For example, ANPP may initially decline, but eventually rise back to pre-treatment levels once drought-tolerant species become more abundant and compensate for drought-intolerant species [@Hoover2014].
-It is also possible that ecosystem functioning shifts to a new mean state, reflective of the suite of species in the new community [@Knapp2012].
+It is also possible that ecosystem functioning shifts to a new mean state, reflecting the suite of species in the new community [@Knapp2012].
 
-Here we report the results of a five-year precipitation manipulation experiment in a sagebrush steppe grassland.
+Manipulating potentially limiting resources offers a route to understanding how ecosystems will respond to resource levels that fall outside the historical range of variability [@Avolio2015; @Gherardi2015; @Knapp2017].
+Altering the amount of precipitation over many years should provide insight into the time scales at which water-limited ecosystems respond to chronic resource alteration.
+Following the HRF, we propose four alternative predictions for the effect of precipitation manipulation on the ecosystem functional response to soil moisture, that is, the soil moisture-ANPP relationship (Figure 1).
+The four predictions are based on possible outcomes at the community (e.g., community composition) and ecosystem (e.g., soil moisture-ANPP regression) levels.
+
+First, altered precipitation changes neither ecosystem functional response nor community composition (Fig. 1, top left).
+In this case, changes in ANPP simply follow the soil moisture-ANPP relationship under ambient conditions.
+This corresponds to the early phases of the HRF, where ecosystem response is due to the physiological responses of individual species.
+Second, the ecosystem functional response changes but community composition remains the same (Fig. 1, top right).
+A saturating soil moisture-ANPP response fits this scenario, where individual species hit physiological thresholds or are limited by some other resource.
+Third, the ecosystem functional response is consistent but underlying community composition changes (Fig. 1, bottom left).
+In this case, changes in species' indentities or abundances occur in response to altered precipitation levels and species more suited to the new conditions compensate for reduced function of initial species.
+Fourth, and last, both ecosystem functional response and community composition change (Fig. 1, bottom right).
+New species, or newly abundant species, with different physiological responses completely reshape the ecosystem functional response.
+
+All four outcomes are possible in any given ecosystem, but the time scales at which the different scenarios play out likely differ [@Smith2009; @Wilcox2016; @Knapp2017].
+Thus, our task is not to test the validity of the HRF, but rather to amass information on how quickly ecosystem functional responses change in different ecosystems.
+Likewise, we need to understand whether changes at the ecosystem level are driven community level changes or individual level responses.
+
+To that end, here we report the results of a five-year precipitation manipulation experiment in a sagebrush steppe grassland.
 We imposed drought and irrigation treatments (50\% decrease/increase) and measured ecosystem (ANPP) and community (species composition) responses.
-We focus on how the drought and irrigation treatments affect the relationship between ANPP and available soil moisture, and if community dynamics underly the ecosystem responses.
-In particular, we are interested in the consistency of the ANPP-soil moisture relationship among treatments.
+We focus on how the drought and irrigation treatments affect the relationship between available soil moisture and ANPP, and if community dynamics underly the ecosystem responses.
+In particular, we are interested in the consistency of the soil moisture-ANPP relationship among treatments.
 Is the relationship steeper under the drought treatment, at low soil moisture?
@@
 To answer these questions we fit a random intercepts, random slopes model to test whether the regressions differed among treatments.
+We also analyzed community composition over time, allowing us to place our experimental results within the framework of the HRF and our competing predictions (Fig. 1).
 
 <!---
+Chronic alterations to resource availability should cause community composition to shift over time, if the focal ecosystem is sensitive to the manipulated resource.
+Species' relative abundances are expected to re-order and, eventually, some species will be lost from the local community and new species will be gained [@Smith2009; @Avolio2015].
+On the heels of changes in community composition, subsequent changes in the level of ecosystem functioning may occur.
+
 As the rate of climate change accelerates, ecologists are being pressed to forecast its impacts on ecosystem functions and services.
@@
 
-# Materials and Methods
+# METHODS
 
@@
 To test whether treatment plots were equally dispersed, or not, we used the 'vegan::betadisper' function [@Oksanen2016].
-Lastly, we examined rank clocks of species' abundances through time to assess the stability of community composition over the course of the experiment [@Collins2008].
-Rank clocks were made using 'ggplot2' [@Wickham2009] and R code from the 'codyn' package [@Hallett2016].
 
@@
 
-# Results
+# RESULTS
 
@@
 
-# Discussion
+# DISCUSSION
 
@@
 
-# Acknowledgments
+# ACKNOWLEDGEMENTS
 
@@
 
-# Funding
+# FUNDING
 
@@
 
-# Tables
+# TABLES
 
@@
 
-# Figures
+# FIGURES
 
@@
 
-# References
+# REFERENCES

### 2017-07-11 manuscript/usses_water_ms.tex
trivia
more intro writing

@@
 
-\documentclass[fleqn,10pt]{wlpeerj} % for preprint submissions
+\documentclass[fleqn,10pt,lineno]{wlpeerj} % for journal submissions
 
@@
 
-\section{Introduction}\label{introduction}
+\section{INTRODUCTION}\label{introduction}
 
@@
 
+Another problem with relying on historical ecoystem functional responses
+is that they are not static. Changes in species identities and
+abundances can alter an ecosystem's functional response to water
+availability because different species have different physiological
+thresholds for producting biomass. \citet{Smith2009} introduced the
+`Hierarchical Response Framework' (HRF) for understanding the interplay
+of community composition and ecosystem functioning in response to
+resource manipulations over time. In the near term, ecosystem
+functioning such as annual net primary productivity (ANPP) will reflect
+the physiological responses of individual species to the manipulated
+resource level. For example, ANPP may decline under simulated drought
+because the initial community consisted of drought-intolerant species
+\citep{Hoover2014}. Over longer time spans, ecosystem functioning may
+recover as new species colonize or initial species reorder in relative
+abundance. For example, ANPP may initially decline, but eventually rise
+back to pre-treatment levels once drought-tolerant species become more
+abundant and compensate for drought-intolerant species
+\citep{Hoover2014}. It is also possible that ecosystem functioning
+shifts to a new mean state, reflecting the suite of species in the new
+community \citep{Knapp2012}.
+
 Manipulating potentially limiting resources offers a route to
 understanding how ecosystems will respond to resource levels that fall
-outside the historical range of variability \citep{Knapp2017}. Chronic
-alterations to resource availability should cause community composition
-to shift over time, the focal ecosystem is sensitive to the manipulated
-resource. Species' relative abundances are expected to re-order and,
-eventually, some species will be lost from the local community and new
-species will be gained \citep{Smith2009, Avolio2015}. On the heels of
-changes in community composition, subsequent changes in the level of
-ecosystem functioning may occur.
+outside the historical range of variability
+\citep{Avolio2015, Gherardi2015, Knapp2017}. Altering the amount of
+precipitation over many years should provide insight into the time
+scales at which water-limited ecosystems respond to chronic resource
+alteration. Following the HRF, we propose four alternative predictions
+for the effect of precipitation manipulation on the ecosystem functional
+response to soil moisture, that is, the soil moisture-ANPP relationship
+(Figure 1). The four predictions are based on possible outcomes at the
+community (e.g., community composition) and ecosystem (e.g., soil
+moisture-ANPP regression) levels.
 
-\citet{Smith2009} introduced the `Hierarchical Response Framework' for
-understanding the interplay of community composition and ecosystem
-functioning in response to resource manipulations over time. In the near
-term, ecosystem functioning such as annual net primary productivity
-(ANPP) will reflect the physiological responses of individual species to
-the manipulated resource level. For example, ANPP may decline under
-simulated drought because the initial community consisted of
-drought-intolerant species \citep{Hoover2014}. Over longer time spans,
-ecosystem functioning may recover as new species colonize or initial
-species reorder in relative abundance. For example, ANPP may initially
-decline, but eventually rise back to pre-treatment levels once
-drought-tolerant species become more abundant and compensate for
-drought-intolerant species \citep{Hoover2014}. It is also possible that
-ecosystem functioning shifts to a new mean state, reflective of the
-suite of species in the new community \citep{Knapp2012}.
+First, altered precipitation changes neither ecosystem functional
+response nor community composition (Fig. 1, top left). In this case,
+changes in ANPP simply follow the soil moisture-ANPP relationship under
+ambient conditions. This corresponds to the early phases of the HRF,
+where ecosystem response is due to the physiological responses of
+individual species. Second, the ecosystem functional response changes
+but community composition remains the same (Fig. 1, top right). A
+saturating soil moisture-ANPP response fits this scenario, where
+individual species hit physiological thresholds or are limited by some
+other resource. Third, the ecosystem functional response is consistent
+but underlying community composition changes (Fig. 1, bottom left). In
+this case, changes in species' indentities or abundances occur in
+response to altered precipitation levels and species more suited to the
+new conditions compensate for reduced function of initial species.
+Fourth, and last, both ecosystem functional response and community
+composition change (Fig. 1, bottom right). New species, or newly
+abundant species, with different physiological responses completely
+reshape the ecosystem functional response.
 
-Here we report the results of a five-year precipitation manipulation
-experiment in a sagebrush steppe grassland. We imposed drought and
-irrigation treatments (50\% decrease/increase) and measured ecosystem
-(ANPP) and community (species composition) responses. We focus on how
-the drought and irrigation treatments affect the relationship between
-ANPP and available soil moisture, and if community dynamics underly the
-ecosystem responses. In particular, we are interested in the consistency
-of the ANPP-soil moisture relationship among treatments. Is the
-relationship steeper under the drought treatment, at low soil moisture?
-Does the relationship saturate under the irrigation treatment, at high
-soil moisture? To answer these questions we fit a random intercepts,
-random slopes model to test whether the regressions differed among
-treatments.
+All four outcomes are possible in any given ecosystem, but the time
+scales at which the different scenarios play out likely differ
+\citep{Smith2009, Wilcox2016, Knapp2017}. Thus, our task is not to test
+the validity of the HRF, but rather to amass information on how quickly
+ecosystem functional responses change in different ecosystems. Likewise,
+we need to understand whether changes at the ecosystem level are driven
+community level changes or individual level responses.
 
-\section{Materials and Methods}\label{materials-and-methods}
+To that end, here we report the results of a five-year precipitation
+manipulation experiment in a sagebrush steppe grassland. We imposed
+drought and irrigation treatments (50\% decrease/increase) and measured
+ecosystem (ANPP) and community (species composition) responses. We focus
+on how the drought and irrigation treatments affect the relationship
+between available soil moisture and ANPP, and if community dynamics
+underly the ecosystem responses. In particular, we are interested in the
+consistency of the soil moisture-ANPP relationship among treatments. Is
+the relationship steeper under the drought treatment, at low soil
+moisture? Does the relationship saturate under the irrigation treatment,
+at high soil moisture? To answer these questions we fit a random
+intercepts, random slopes model to test whether the regressions differed
+among treatments. We also analyzed community composition over time,
+allowing us to place our experimental results within the framework of
+the HRF and our competing predictions (Fig. 1).
+
+\section{METHODS}\label{methods}
 
@@
 dispersed, or not, we used the `vegan::betadisper' function
-\citep{Oksanen2016}. Lastly, we examined rank clocks of species'
-abundances through time to assess the stability of community composition
-over the course of the experiment \citep{Collins2008}. Rank clocks were
-made using `ggplot2' \citep{Wickham2009} and R code from the `codyn'
-package \citep{Hallett2016}.
+\citep{Oksanen2016}.
 
@@
 
-\section{Results}\label{results}
+\section{RESULTS}\label{results}
 
@@
 
-\section{Discussion}\label{discussion}
+\section{DISCUSSION}\label{discussion}
 
@@
 
-\section{Acknowledgments}\label{acknowledgments}
+\section{ACKNOWLEDGEMENTS}\label{acknowledgements}
 
@@
 
-\section{Funding}\label{funding}
+\section{FUNDING}\label{funding}
 
@@
 
-\section{Tables}\label{tables}
+\section{TABLES}\label{tables}
 
@@
 
-\section{Figures}\label{figures}
+\section{FIGURES}\label{figures}
 

### 2017-07-11 manuscript/usses_water_ms.Rmd
finished discussion draft; ready for comments

@@
   - code: 2
-    address: Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California xxxxx
+    address: Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California 90095
 abstract: >
@@
 Here we report the results of a five-year experiment where we used drought and irrigation treatments (50% decrease/increase) to see how a sagebrush steppe plant community in Idaho will respond to future climate changes.
-The ecosystem was suprisingly resistant to both drought and irrigation.
-Aboveground net primary productivity (ANPP) responded positively to ambient growing season precipitation, but the response did not vary across treatments.
+Ecosystem functional response, defined as the relationship between soil moisture and aboveground net primary productivity, was suprisingly consistent across drought, control, and irrigation treatments.
+Aboveground net primary productivity (ANPP) responded positively to availabile soil moisture, but the response did not vary across treatments.
 There was also no evidence that treatment effects grew over time.
-The similarity of ecosystem functioning was not due to compensatory shifts at the plant community level, where species composition among treatments was similar and remarkably stable over the five years.
-At least in the short-term, ecosystem functioning and community composition in this sagebrush steppe system is resistant to increases and decreases in growing season precipitation.
+The similarity of ecosystem functional response across treatment was not due to compensatory shifts at the plant community level, where species composition among treatments was similar and remarkably stable over the five years.
+At least in the short-term, ecosystem functional response and community composition in this sagebrush steppe system is resistant to increases and decreases in growing season precipitation.
 header-includes: >
@@
 
+\ark{ANDY: add paragraph on the statistical modeling of soil moisture to produce Figure 2B (same as your Figure 2 in the precip-demography-prediction dissertation chapter).}
+
 ## Data Collection
@@
 
-\noindent{} where $\mu_{i(j(k))}$ is the deterministic prediction from the regression model for observation *i* for plot *j* associated with treatment *k* in year *t*, $\beta_{0,j(k)}$ is the intercept for plot *j* associated with treatment *k*, $\beta_{1,j(k)}$ is the slope term for the effect of soil moisture for plot *j* associated with treatment *k*, $\gamma_t$ is the intercept offset for year *t*, and $\sigma^2_k$ is the process variance for treatment *k*.
+\noindent{}where $\mu_{i(j(k))}$ is the deterministic prediction from the regression model for observation *i* for plot *j* associated with treatment *k* in year *t*, $\beta_{0,j(k)}$ is the intercept for plot *j* associated with treatment *k*, $\beta_{1,j(k)}$ is the slope term for the effect of soil moisture for plot *j* associated with treatment *k*, $\gamma_t$ is the intercept offset for year *t*, and $\sigma^2_k$ is the process variance for treatment *k*.
 Data include the standardized log(ANPP) observations $\left( y_{i(j(k(t)))} \right)$ and soil moisture ($x_i$).
@@
 
-\vspace{-2em}
+\vspace{-1em}
 
@@
 
-\noindent{} where $\boldsymbol{\beta}_{j(k)}$ is the vector of regression coefficients (intercept and slope) for plot *j* associated with treatment *k*, $\boldsymbol{\beta}_{k}$ is the vector of coefficients for each treatment, and $\boldsymbol{\beta}$ is the vector of overall coefficients.
+\noindent{}where $\boldsymbol{\beta}_{j(k)}$ is the vector of regression coefficients (intercept and slope) for plot *j* associated with treatment *k*, $\boldsymbol{\beta}_{k}$ is the vector of coefficients for each treatment, and $\boldsymbol{\beta}$ is the vector of overall coefficients.
 The plot- and treatment-level coefficients are drawn from multivariate normal distributions with covariance matrix $\boldsymbol{\Sigma}$.
@@
 
-Three of our five treatment years fell in years of below average rainfall (Fig. 1A).
+Three of our five treatment years fell in years of below average rainfall (Fig. 2A).
 Thus, those three years represent a lower magnitude of absolute change in precipitation experienced by the treatments.
-Averaged across treatments, ANPP varied from a minimum of 74.5 g m$^{-2}$ in 2014 to a maximum of 237.1 g m$^{-2}$ in 2016 (Fig. 1C).
-ANPP was slightly higher in irrigation plots and slightly lower in drought plots (Fig. 1C), corresponding to estimated soil volumetric water content (VWC) differences among treatments (Fig. 1B).
+Averaged across treatments, ANPP varied from a minimum of 74.5 g m$^{-2}$ in 2014 to a maximum of 237.1 g m$^{-2}$ in 2016 (Fig. 2C).
+ANPP was slightly higher in irrigation plots and slightly lower in drought plots (Fig. 2C), corresponding to estimated soil volumetric water content (VWC) differences among treatments (Fig. 2B).
 Such differences in soil VWC indicate our treatment infrastructure was successful.
 
-Cumulative growing season soil moisture had a positive effect on ANPP (mean of $\beta_{1}$ = 0.43; 80\% BCI = -0.05, 0.88; 95\% BCI = -0.36, 1.15) (Fig. 1D).
-Average ANPP was similar among treatments (similar intercepts, Fig. 2A), as was the effect of precipitation (similar slopes, Fig. 2B).
-In an average precipitation year (i.e., $x$ = 0 in Eq.
-1), the probability that ANPP in a conrol plot is greater than ANPP in a drought plot was 0.71, and the probability that ANPP in a control plot is lower than in an irrigation plot was 0.41.
-In other words, the posterior distributions of $\beta_0(\text{control}) - \beta_0(\text{drought})$ and $\beta_0(\text{control}) - \beta_0(\text{drought})$ broadly overlapped zero.
-There was also no evidence that the treatment effects became more important over time because there was no directional trend in the yearly treatment effects estimated from the ANPP-Treatment model (Fig.
-3).
+Cumulative growing season soil moisture had a positive effect on ANPP (mean of $\beta_{1}$ = 0.43; 80\% BCI = -0.05, 0.88; 95\% BCI = -0.36, 1.15) (Fig. 2D).
+Ecosystem functional response was similar among treatments, with treatment level regressions between soil moisture and log(ANPP) having similar intercepts (Fig. 3A) and slopes (Fig. 3B).
+There was also no evidence that the treatment effects became more important over time because there was no directional trend in the random year effects (Fig. 4).
 
 Community composition was similar among treatments.
-In no year did community composition among treatments not overlap, and they were equally dispersed in all years (Table 1; Fig. 4).
-Likewise, community composition was remarkably stable over time, with no evidence of divergence among treatments (Table 1; Fig. 4).
-Species' abundances and ranks showed little deviation over the five-year experiment, regardless of treatment (Fig. 5).
+In no year did community composition among treatments not overlap, and they were equally dispersed in all years (Table 1; Fig. 5).
+Likewise, community composition was remarkably stable over time, with no evidence of divergence among treatments (Table 1; Fig. 5).
 
@@
 
-Surprisingly, ANPP was not sensitive to either drought or irrigation treatments, whether in the context of growing season precipitation (Fig.
-2) or in the context of year since the experiment began (Fig.
+Ecosystem functional response (i.e., the relationship between soil moisture and ANPP) and community composition in this sagebrush steppe ecosystem was consistent across drought, control, and irrigation treatments (Fig.
+2D, Fig.
 3).
+In other words, functional responses estimated from any precipitation level sampled apply to any other level of precipitation, and the consistent functional response was not due to species reordering and compensation.
+Thus, our results conform to the first of our four predictions: neither ecosystem functional response nor community composition changed with chronic alteration in water availability (Fig.
+1, top left).
+
 This is surprising because grasslands generally, and sagebrush steppe specifically, are considered water-limited systems.
-Thus, our expectation was that ANPP would be tightly linked to soil water availability.
-ANPP was slightly lower in drought plots and slightly higher in irrigation plots, but these differences were not statistically important.
+Thus, our naive expecatations were that either (i) ecosystem functional response would differ among treatments as species reach physiological thresholds and resource limitations (Fig.
+1, top right), or (ii) ecosystem functional response stays the same due to species reordering and compensation (Fig.
+1, bottom left).
 
-In the absence of community composition data, an obvious hypothesis to explain the lack of ANPP response is that species reordered rapidly.
+In the absence of community composition data, an obvious hypothesis to explain the consistent ecosystem functional response is that species reordered rapidly.
 In so doing, newly abundant species could compensate for the loss of ecosystem functioning of previously abundant species.
-We did not find this to be the case, as specie composition was remarkably stable over the course of the experiment (Fig. 3).
+We did not find this to be the case, as species composition was remarkably stable over the course of the experiment (Fig. 5).
 
@@
 First, it could be that our experiment simply was not long enough to induce responses.
-This may be true for community responses, which can take decades [@Wilcox2016], but the response of grassland ANPP to water manipulations is typically immediate [e.g., @Hoover2014].
+This may be true for community responses, which can take decades [@Wilcox2016].
 
@@
 The 1\% quantile of precipitation at our site is 110 mm, a 47% reduction from the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean growing season precipitation (Appendix 3).
-Thus, our drought treatment represented extreme precipitation amounts, especially in years where ambient precipitation was below average (Fig. 1A).
+Thus, our drought treatment represented extreme precipitation amounts, especially in years where ambient precipitation was below average (Fig. 2A).
 The irrigation treatment may have been too small, however.
 
-Third, the ecosystem and community resistance to drought and deluge we observed could be a real phenomenon in this system.
+Third, the ecosystem and community resistance to drought and deluge we observed could be a real phenomenon in this system, at least over the five year time span we observed.
+The perennial plants in this cold desert ecosystem are tolerant to drought conditions [@Bazzaz1979; @Franks2011, A.R.
+Kleinhesselink, unpublished data], meaning a consistent ecosystem functional response at low soil moisture is perfectly reasonable.
+For example, many of the perennial grasses in our focal ecosystem avoid drought stress by growing early in the growing season (A.R. Kleinhesselink, personal observation).
+Likewise, the dominant shrub in our focal ecosystem, *Artemisia tripartita*, has access to water deep in the soil profile thanks to a deep root system [@Germino2014].
+As for the irrigation treatments, the lack of a saturating response (e.g., a lower regression slope) suggests water availability remained a limiting factor at the levels of precipitation we were able to manipulate.
+
+In conclusion, our results suggest that five years of $\pm$ 50\% ambient precipitation is not enough to induce a shift in ecosystem functional response in a sagebrush steppe.
+This is likely because component species are tolerant to drought conditions and water remains co-limiting at higher precipitation levels.
+Longer time series of chronic precipitation alteration may reveal plant community shifts that we did not observe [e.g., @Wilcox2016].
+Our results suggest compositional shifts would have the largest impact at high rainfall because the current community maintained consistent ecosystem functional response at very low water availability.
 

### 2017-07-11 manuscript/usses_water_ms.tex
finished discussion draft; ready for comments

@@
 \affil[2]{Department of Ecology and Evolutionary Biology, University of
-California, Los Angeles, Los Angeles, California xxxxx}
+California, Los Angeles, Los Angeles, California 90095}
 
@@
 see how a sagebrush steppe plant community in Idaho will respond to
-future climate changes. The ecosystem was suprisingly resistant to both
-drought and irrigation. Aboveground net primary productivity (ANPP)
-responded positively to ambient growing season precipitation, but the
-response did not vary across treatments. There was also no evidence that
-treatment effects grew over time. The similarity of ecosystem
-functioning was not due to compensatory shifts at the plant community
-level, where species composition among treatments was similar and
-remarkably stable over the five years. At least in the short-term,
-ecosystem functioning and community composition in this sagebrush steppe
-system is resistant to increases and decreases in growing season
-precipitation.
+future climate changes. Ecosystem functional response, defined as the
+relationship between soil moisture and aboveground net primary
+productivity, was suprisingly consistent across drought, control, and
+irrigation treatments. Aboveground net primary productivity (ANPP)
+responded positively to availabile soil moisture, but the response did
+not vary across treatments. There was also no evidence that treatment
+effects grew over time. The similarity of ecosystem functional response
+across treatment was not due to compensatory shifts at the plant
+community level, where species composition among treatments was similar
+and remarkably stable over the five years. At least in the short-term,
+ecosystem functional response and community composition in this
+sagebrush steppe system is resistant to increases and decreases in
+growing season precipitation.
 % Dummy abstract text.
@@
 
+\ark{ANDY: add paragraph on the statistical modeling of soil moisture to produce Figure 2B (same as your Figure 2 in the precip-demography-prediction dissertation chapter).}
+
 \subsection{Data Collection}\label{data-collection}
@@
 
-\noindent{} where \(\mu_{i(j(k))}\) is the deterministic prediction from
+\noindent{}where \(\mu_{i(j(k))}\) is the deterministic prediction from
 the regression model for observation \emph{i} for plot \emph{j}
@@
 
-\noindent{} where \(\boldsymbol{\beta}_{j(k)}\) is the vector of
+\noindent{}where \(\boldsymbol{\beta}_{j(k)}\) is the vector of
 regression coefficients (intercept and slope) for plot \emph{j}
@@
 Three of our five treatment years fell in years of below average
-rainfall (Fig. 1A). Thus, those three years represent a lower magnitude
+rainfall (Fig. 2A). Thus, those three years represent a lower magnitude
 of absolute change in precipitation experienced by the treatments.
 Averaged across treatments, ANPP varied from a minimum of 74.5 g
-m\(^{-2}\) in 2014 to a maximum of 237.1 g m\(^{-2}\) in 2016 (Fig. 1C).
+m\(^{-2}\) in 2014 to a maximum of 237.1 g m\(^{-2}\) in 2016 (Fig. 2C).
 ANPP was slightly higher in irrigation plots and slightly lower in
-drought plots (Fig. 1C), corresponding to estimated soil volumetric
-water content (VWC) differences among treatments (Fig. 1B). Such
+drought plots (Fig. 2C), corresponding to estimated soil volumetric
+water content (VWC) differences among treatments (Fig. 2B). Such
 differences in soil VWC indicate our treatment infrastructure was
@@
 (mean of \(\beta_{1}\) = 0.43; 80\% BCI = -0.05, 0.88; 95\% BCI = -0.36,
-1.15) (Fig. 1D). Average ANPP was similar among treatments (similar
-intercepts, Fig. 2A), as was the effect of precipitation (similar
-slopes, Fig. 2B). In an average precipitation year (i.e., \(x\) = 0 in
-Eq. 1), the probability that ANPP in a conrol plot is greater than ANPP
-in a drought plot was 0.71, and the probability that ANPP in a control
-plot is lower than in an irrigation plot was 0.41. In other words, the
-posterior distributions of
-\(\beta_0(\text{control}) - \beta_0(\text{drought})\) and
-\(\beta_0(\text{control}) - \beta_0(\text{drought})\) broadly overlapped
-zero. There was also no evidence that the treatment effects became more
-important over time because there was no directional trend in the yearly
-treatment effects estimated from the ANPP-Treatment model (Fig. 3).
+1.15) (Fig. 2D). Ecosystem functional response was similar among
+treatments, with treatment level regressions between soil moisture and
+log(ANPP) having similar intercepts (Fig. 3A) and slopes (Fig. 3B).
+There was also no evidence that the treatment effects became more
+important over time because there was no directional trend in the random
+year effects (Fig. 4).
 
@@
 community composition among treatments not overlap, and they were
-equally dispersed in all years (Table 1; Fig. 4). Likewise, community
+equally dispersed in all years (Table 1; Fig. 5). Likewise, community
 composition was remarkably stable over time, with no evidence of
-divergence among treatments (Table 1; Fig. 4). Species' abundances and
-ranks showed little deviation over the five-year experiment, regardless
-of treatment (Fig. 5).
+divergence among treatments (Table 1; Fig. 5).
 
@@
 
-Surprisingly, ANPP was not sensitive to either drought or irrigation
-treatments, whether in the context of growing season precipitation (Fig.
-2) or in the context of year since the experiment began (Fig. 3). This
-is surprising because grasslands generally, and sagebrush steppe
-specifically, are considered water-limited systems. Thus, our
-expectation was that ANPP would be tightly linked to soil water
-availability. ANPP was slightly lower in drought plots and slightly
-higher in irrigation plots, but these differences were not statistically
-important.
+Ecosystem functional response (i.e., the relationship between soil
+moisture and ANPP) and community composition in this sagebrush steppe
+ecosystem was consistent across drought, control, and irrigation
+treatments (Fig. 2D, Fig. 3). In other words, functional responses
+estimated from any precipitation level sampled apply to any other level
+of precipitation, and the consistent functional response was not due to
+species reordering and compensation. Thus, our results conform to the
+first of our four predictions: neither ecosystem functional response nor
+community composition changed with chronic alteration in water
+availability (Fig. 1, top left).
+
+This is surprising because grasslands generally, and sagebrush steppe
+specifically, are considered water-limited systems. Thus, our naive
+expecatations were that either (i) ecosystem functional response would
+differ among treatments as species reach physiological thresholds and
+resource limitations (Fig. 1, top right), or (ii) ecosystem functional
+response stays the same due to species reordering and compensation (Fig.
+1, bottom left).
 
 In the absence of community composition data, an obvious hypothesis to
-explain the lack of ANPP response is that species reordered rapidly. In
-so doing, newly abundant species could compensate for the loss of
-ecosystem functioning of previously abundant species. We did not find
-this to be the case, as specie composition was remarkably stable over
-the course of the experiment (Fig. 3).
+explain the consistent ecosystem functional response is that species
+reordered rapidly. In so doing, newly abundant species could compensate
+for the loss of ecosystem functioning of previously abundant species. We
+did not find this to be the case, as species composition was remarkably
+stable over the course of the experiment (Fig. 5).
 
@@
 experiment simply was not long enough to induce responses. This may be
-true for community responses, which can take decades \citep{Wilcox2016},
-but the response of grassland ANPP to water manipulations is typically
-immediate \citep[e.g.,][]{Hoover2014}.
+true for community responses, which can take decades \citep{Wilcox2016}.
 
@@
 extreme precipitation amounts, especially in years where ambient
-precipitation was below average (Fig. 1A). The irrigation treatment may
+precipitation was below average (Fig. 2A). The irrigation treatment may
 have been too small, however.
@@
 Third, the ecosystem and community resistance to drought and deluge we
-observed could be a real phenomenon in this system.
+observed could be a real phenomenon in this system, at least over the
+five year time span we observed. The perennial plants in this cold
+desert ecosystem are tolerant to drought conditions \citep[A.R.
+Kleinhesselink, unpublished data]{Bazzaz1979, Franks2011}, meaning a
+consistent ecosystem functional response at low soil moisture is
+perfectly reasonable. For example, many of the perennial grasses in our
+focal ecosystem avoid drought stress by growing early in the growing
+season (A.R. Kleinhesselink, personal observation). Likewise, the
+dominant shrub in our focal ecosystem, \emph{Artemisia tripartita}, has
+access to water deep in the soil profile thanks to a deep root system
+\citep{Germino2014}. As for the irrigation treatments, the lack of a
+saturating response (e.g., a lower regression slope) suggests water
+availability remained a limiting factor at the levels of precipitation
+we were able to manipulate.
+
+In conclusion, our results suggest that five years of \(\pm\) 50\%
+ambient precipitation is not enough to induce a shift in ecosystem
+functional response in a sagebrush steppe. This is likely because
+component species are tolerant to drought conditions and water remains
+co-limiting at higher precipitation levels. Longer time series of
+chronic precipitation alteration may reveal plant community shifts that
+we did not observe \citep[e.g.,][]{Wilcox2016}. Our results suggest
+compositional shifts would have the largest impact at high rainfall
+because the current community maintained consistent ecosystem functional
+response at very low water availability.
 

### 2017-07-19 manuscript/usses_water_ms.Rmd
adds BT to author list

@@
     affiliation: 2
+  - name: Bret Taylor
+    affiliation: 3
   - name: Peter B. Adler
@@
     address: Department of Ecology and Evolutionary Biology, University of California, Los Angeles, Los Angeles, California 90095
+  - code: 3
+    address: USSES, address...
 abstract: >

### 2017-07-19 manuscript/usses_water_ms.Rmd
adds discussion points

@@
 
+<!---
+NOTES FROM MEETING WITH PETER
+1.
+Bet hedgers -- dominant species here deal well with low water and interannual variability; they live in dry and variable in environment, which has selected for tratis (not hitting physiological thresholds which only becom apparaent when species really go for it).
+Bet hedging = do the same things every year.
+2. Statistical explanation -- either (1) above explains our weak and variable relationship, or we just looked at the wrong thing; lots of other things beyond VWC that are influencing ANPP.
+3. Limitation -- snow, we don't know what happens. Its blowing around, who knows how big of an impact this has.
+4. Show Fig 2D with real model fits and move to results figure 3.
+---->
+
 Ecosystem response to chronic resource alteration is expected to follow a temporal trend.

### 2017-07-21 manuscript/usses_water_ms.Rmd
title change

@@
 ---
-title: Consistent ecosystem functional response during five years of drought and deluge in a sagebrush steppe
+title: Linear ecosystem functional response across precipitation extremes in a sagebrush steppe
 preprint: false
@@
   Precipitation is predicted to become more variable in the western U.S., meaning years of above and below average precipitation will become more common.
-Such periods of drought and deluge could become major drivers of plant community dynamics and ecosystem functioning in water limited grasslands.
+Such periods of drought and deluge are major drivers of plant community dynamics and ecosystem functioning in water limited grasslands.
 Here we report the results of a five-year experiment where we used drought and irrigation treatments (50% decrease/increase) to see how a sagebrush steppe plant community in Idaho will respond to future climate changes.

### 2017-07-21 code/look_at_weather.R
looking at new model structures

@@ -87,7 +87,7 @@ anpp_means <- ggplot(biomass_yr_trt_summ, aes(x=year, y=mean_biomass, color=Trea
   theme_bw()+
   theme(panel.grid.minor = element_blank())+
   guides(color = guide_legend(override.aes = list(size=1)))+
-  theme(legend.position = c(0.2, 0.8),legend.key.size = unit(1,"pt"),legend.title = element_text(size=10),
+  theme(legend.position = c(0.2, 0.75),legend.key.size = unit(1,"pt"),legend.title = element_text(size=10),
         legend.text = element_text(size = 8),
         legend.key.height = unit(0.8,"line"),
         legend.box.background = element_rect())
@@ -127,29 +127,38 @@ soil_vwc <- ggplot(soil_moisture, aes(x=julian_date, y=VWC, group=Treatment, col
 ####  MAKE SCATTERPLOT OF ANPP vs. PRECIP ----
 ####
 source("read_format_data.R") # load data
-data_plot <- ggplot(anpp_data, aes(x=total_seasonal_vwc,y=anpp))+
-  geom_point(shape=21,color="grey25",alpha=0.8,aes(fill=Treatment))+
-  #stat_smooth(method="lm", aes(color=Treatment), se=FALSE, size=0.5)+
-  stat_smooth(method="glm", formula=y~x, method.args=list(family=quasi(link='log')),aes(color=Treatment), se=FALSE, size=0.5)+
-  #stat_smooth(method="lm", color="black", se=FALSE, size=0.7)+
-  stat_smooth(method="glm", formula=y~x, method.args=list(family=quasi(link='log')),color="black", se=FALSE, size=0.5)+
-  scale_fill_brewer(palette = "Set2")+
-  scale_color_brewer(palette = "Set2")+
-  #scale_x_continuous(limits=c(100,300),breaks=seq(100,300,50))+
-  scale_y_continuous(breaks=seq(50,350,50))+
-  xlab("March-June Cumulative VWC")+
-  ylab(expression(paste("ANPP (g ",m^2,")")))+
-  ggtitle("D")+
-  guides(fill=FALSE,color=FALSE)+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())
+mod <- lm(log(anpp)~total_seasonal_vwc, data = dplyr::filter(anpp_data, Treatment == "Control"))
[…]

### 2017-07-25 manuscript/usses_water_ms.Rmd
updates figs and a little writing

@@
 ---
-title: Linear ecosystem functional response across precipitation extremes in a sagebrush steppe
+title: Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe
 preprint: false
@@
 
+```{r get_results, echo=FALSE}
+betas <- read.csv("../results/beta_summaries.csv")
+```
+
 # RESULTS
@@
 
-Cumulative growing season soil moisture had a positive effect on ANPP (mean of $\beta_{1}$ = 0.43; 80\% BCI = -0.05, 0.88; 95\% BCI = -0.36, 1.15) (Fig. 2D).
-Ecosystem functional response was similar among treatments, with treatment level regressions between soil moisture and log(ANPP) having similar intercepts (Fig. 3A) and slopes (Fig. 3B).
+Cumulative growing season soil moisture had a positive effect on ANPP (mean of $\beta_{1}$ = `r round(betas$mean[8],2)`; 80\% BCI = `r round(betas$X10.[8],2)`, `r round(betas$X90.[8],2)`; 95\% BCI = `r round(betas$X2.5.[8],2)`, `r round(betas$X97.5.[8],2)`) (Fig.
+3).
+Ecosystem functional response was similar among treatments, with treatment level regressions between soil moisture and log(ANPP) having similar intercepts and slopes (Fig. 3A).
 There was also no evidence that the treatment effects became more important over time because there was no directional trend in the random year effects (Fig. 4).
@@
 
+# AUTHOR CONTRIBUTIONS
+\begin{itemize}
+  \item Andrew T. Tredennick collected data, analyzed the data, wrote the paper, prepared figures and/or tables, reviewed
+drafts of the paper.
+  \item Andrew R. Kleinhesselink conceived and designed the experiments, performed the experiments, collected data, reviewed drafts of the paper.
+  \item Peter B. Adler conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed
+drafts of the paper.
+\end{itemize}
+
 \newpage{}
@@
 (B) Observed soil volumetric water content (VWC) over the course of the experiment.
-(C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.
-(D) Scatterplot of ANPP versus growing season precipitation.
-Colored regression lines are independently-fit linear models for each treatment with no random effects structure; dark black regression line is a linear regression through all the point with no random effects structure.
-Our analysis seeks to find if the data supports a model with different intercepts and/or slopes for each treatment, while also accounting for nonindependence of samples within plots across years.
-Color mapping for panels B-D is shown in the legend for panel C.}
+(C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.}
 \end{figure}
@@
   \centering
-      \includegraphics[width=5in]{../figures/glmm_treatment_posteriors.png}
-  \caption{Posterior distributions of treatment-level parameters ('Intercept' and the effect of 'Soil Moisture').
-Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.}
+      \includegraphics[width=5in]{../figures/glmm_main_results.png}
+  \caption{Results from the generalized linear mixed effects model.
+(A) Posterior distributions of treatment-level parameters ('Intercept' and the effect of 'Soil Moisture').
+Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.
+(B) Scatterplot of the data and model estimates shown a solid lines.
+Model estimates come from treatment level coefficients (colored lines) and the overall mean coefficients (heavy black line).
+Note that we show log(ANPP) on the \emph{y}-axis of panel B; this same plot can be seen on the arithmetic scale in supporting material Fig.
+S2-1.}
 \end{figure}

### 2017-07-25 manuscript/usses_water_ms.tex
updates figs and a little writing

@@
 
-\title{Consistent ecosystem functional response during five years of drought
-and deluge in a sagebrush steppe}
+\title{Consistent ecosystem functional response across precipitation extremes
+in a sagebrush steppe}
 
@@
 \author[2]{Andrew R. Kleinhesselink}
+
+\author[3]{Bret Taylor}
 
@@
 California, Los Angeles, Los Angeles, California 90095}
+\affil[3]{USSES, address\ldots{}}
 
@@
 meaning years of above and below average precipitation will become more
-common. Such periods of drought and deluge could become major drivers of
-plant community dynamics and ecosystem functioning in water limited
+common. Such periods of drought and deluge are major drivers of plant
+community dynamics and ecosystem functioning in water limited
 grasslands. Here we report the results of a five-year experiment where
@@
 Cumulative growing season soil moisture had a positive effect on ANPP
-(mean of \(\beta_{1}\) = 0.43; 80\% BCI = -0.05, 0.88; 95\% BCI = -0.36,
-1.15) (Fig. 2D). Ecosystem functional response was similar among
+(mean of \(\beta_{1}\) = 0.41; 80\% BCI = -0.08, 0.86; 95\% BCI = -0.45,
+1.13) (Fig. 3). Ecosystem functional response was similar among
 treatments, with treatment level regressions between soil moisture and
-log(ANPP) having similar intercepts (Fig. 3A) and slopes (Fig. 3B).
-There was also no evidence that the treatment effects became more
-important over time because there was no directional trend in the random
-year effects (Fig. 4).
+log(ANPP) having similar intercepts and slopes (Fig. 3A). There was also
+no evidence that the treatment effects became more important over time
+because there was no directional trend in the random year effects (Fig.
+4).
 
@@
 
+\section{AUTHOR CONTRIBUTIONS}\label{author-contributions}
+
+\begin{itemize}
+  \item Andrew T. Tredennick collected data, analyzed the data, wrote the paper, prepared figures and/or tables, reviewed
+drafts of the paper.
+  \item Andrew R. Kleinhesselink conceived and designed the experiments, performed the experiments, collected data, reviewed drafts of the paper.
+  \item Peter B. Adler conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed
+drafts of the paper.
+\end{itemize}
+
 \newpage{}

### 2017-07-25 code/anpp_randcoefs_model.R

@@ -84,8 +84,24 @@ fit_stan_model <- function(model_data, check_diags=FALSE, treattype){
 all_fit <- fit_stan_model(anpp_data,check_diags = FALSE, treattype = "all")
 saveRDS(all_fit, "../results/randcoefs_alltreatments_fit.RDS")
 
-# drought_fit <- fit_stan_model(drought_data,check_diags = T,treattype = "drought")
-# irrigate_fit <- fit_stan_model(irrigate_data,check_diags = T,treattype = "irrigate")
-# saveRDS(drought_fit, "../results/randcoefs_drought_fit.RDS")
-# saveRDS(irrigate_fit, "../results/randcoefs_irrigate_fit.RDS")
 
+
+####
+####  PLOT RESIDUALS ----
+####
+# predictions <- reshape2::melt(rstan::extract(all_fit, pars="yhat")) %>%
+#   rename(iteration = iterations, obs_id = Var2, estimate = value, stan_name = L1) %>%
+#   group_by(obs_id) %>%
+#   summarise(mean_estimate = mean(estimate))
+# 
+# resids <- reshape2::melt(rstan::extract(all_fit, pars="resid")) %>%
+#   rename(iteration = iterations, obs_id = Var2, resid = value, stan_name = L1) %>%
+#   left_join(predictions, by="obs_id") %>%
+#   group_by(obs_id) %>%
+#   summarise(mean_resid = mean(resid),
+#             mean_est_anpp = mean(mean_estimate)) %>%
+#   ungroup()
+# 
+# ggplot(resids, aes(x=mean_est_anpp, y=mean_resid))+
+#   geom_hline(aes(yintercept=0))+
+#   geom_point()

### 2017-07-25 code/plot_model_results.R

@@ -60,22 +60,94 @@ g1 <- ggplot(betas, aes(x=estimate))+
   geom_hline(aes(yintercept=0), color="grey45", size=0.1)+
   geom_line(stat="density",aes(color=treatment),adjust=5,size=1)+
   scale_color_brewer(palette = "Set2",name="Treatment")+
-  scale_x_continuous(breaks=seq(-1.5,2,0.5))+
-  scale_y_continuous(breaks=seq(0,2,0.25))+
+  #scale_x_continuous(breaks=seq(-1.5,2,0.5))+
+  #scale_y_continuous(breaks=seq(0,2,0.25))+
   facet_wrap(~param_name)+
   xlab("Parameter Value")+
   ylab("Probability Density")+
   theme_bw()+
   theme(panel.grid.minor = element_blank())+
-  theme(legend.position = c(0.4,0.8),
+  theme(legend.position = c(0.4,0.78),
         legend.key.size = unit(6,"pt"),
         legend.title = element_text(size=10),
         legend.text = element_text(size = 8),
         legend.key.height = unit(0.8,"line"),
         legend.box.background = element_rect(),
         strip.background = element_blank(), 
-        strip.text = element_text(size=10))
-ggsave("../figures/glmm_treatment_posteriors.png",plot = g1, width = 6, height = 3, units = "in", dpi =120)
+        strip.text = element_text(size=10))+
+  ggtitle("A)")
+#ggsave("../figures/glmm_treatment_posteriors.png",plot = g1, width = 6, height = 3, units = "in", dpi =120)
+
+
+
+####
+####  PLOT MODEL PREDICTIONS WITH DATA ----
+####
+lmod <- lm(log(anpp) ~ total_seasonal_vwc, anpp_data)
+x <- model.matrix(lmod)
+newx <- as.data.frame(unique(x)) %>% select(total_seasonal_vwc)
+newx$vwc_id <- c(1:nrow(newx))
+mean_log_anpp <- mean(log(anpp_data$anpp))
+sd_log_anpp <- sd(log(anpp_data$anpp))
+model_preds <- reshape2::melt(rstan::extract(all_fit, pars="ypreds_mu")) %>%
+  rename(iteration = iterations, vwc_id = Var2, estimate = value, stan_name = L1) %>%
+  left_join(newx, by="vwc_id") %>%
[…]

### 2017-07-26 manuscript/usses_water_ms.Rmd
addresses PBA minor comments

@@
 abstract: >
-  Precipitation is predicted to become more variable in the western U.S., meaning years of above and below average precipitation will become more common.
-Such periods of drought and deluge are major drivers of plant community dynamics and ecosystem functioning in water limited grasslands.
-Here we report the results of a five-year experiment where we used drought and irrigation treatments (50% decrease/increase) to see how a sagebrush steppe plant community in Idaho will respond to future climate changes.
+  Precipitation is predicted to become more variable in the western United States, meaning years of above and below average precipitation will become more common.
+Periods of drought and deluge are major drivers of ecosystem functioning in water limited grasslands, but how ecosystems respond to precipitation may change as the duration of above and below average periods increases.
+Changes in ecosystem functional response could reflect underlying changes in species composition or species reaching physiological thresholds at extreme precipitation levels.
+Here we report the results of a five-year experiment where we used drought and irrigation treatments (approximately 50% decrease/increase) to investigate whether ecosystem functional response remains consistent at extreme precipitation levels.
+We also use species-level adundance data to link ecosystem functional response to community dynamics.
 Ecosystem functional response, defined as the relationship between soil moisture and aboveground net primary productivity, was suprisingly consistent across drought, control, and irrigation treatments.
-Aboveground net primary productivity (ANPP) responded positively to availabile soil moisture, but the response did not vary across treatments.
+Aboveground net primary productivity (ANPP) responded positively to availabile soil moisture, but the response was weak and did not vary across treatments.
 There was also no evidence that treatment effects grew over time.
-The similarity of ecosystem functional response across treatment was not due to compensatory shifts at the plant community level, where species composition among treatments was similar and remarkably stable over the five years.
-At least in the short-term, ecosystem functional response and community composition in this sagebrush steppe system is resistant to increases and decreases in growing season precipitation.
+The similarity of ecosystem functional response across treatments was not due to compensatory shifts at the plant community level, where species composition among treatments was similar and remarkably stable over the five years.
+At least in the short-term, ecosystem functional response and community composition in this sagebrush steppe system is resistant to increases and decreases in growing season precipitation, perhaps due to bet-hedging strategies of constituent species.
 header-includes: >
-  \usepackage{lipsum}
   \usepackage{setspace}
@@
 
+\reversemarginpar
+
 ```{r caching, include=FALSE}
@@
 # INTRODUCTION
+
+Projected changes in precipitation regimes, associated with global climate change, will affect
 
@@
 To that end, here we report the results of a five-year precipitation manipulation experiment in a sagebrush steppe grassland.
-We imposed drought and irrigation treatments (50\% decrease/increase) and measured ecosystem (ANPP) and community (species composition) responses.
+We imposed drought and irrigation treatments (approximately $\pm50\%$) and measured ecosystem (ANPP) and community (species composition) responses.
 We focus on how the drought and irrigation treatments affect the relationship between available soil moisture and ANPP, and if community dynamics underly the ecosystem responses.
@@
 We also analyzed community composition over time, allowing us to place our experimental results within the framework of the HRF and our competing predictions (Fig. 1).
-
-<!---
-Chronic alterations to resource availability should cause community composition to shift over time, if the focal ecosystem is sensitive to the manipulated resource.
-Species' relative abundances are expected to re-order and, eventually, some species will be lost from the local community and new species will be gained [@Smith2009; @Avolio2015].
-On the heels of changes in community composition, subsequent changes in the level of ecosystem functioning may occur.
-
-As the rate of climate change accelerates, ecologists are being pressed to forecast its impacts on ecosystem functions and services.
-Making such forecasts faces two challenges.
-First, future conditions are likely to be outside the historical range of variability [@Smith2011], meaning we cannot simply look to the past to predict the future.
-Second, ecosystems will likely exhibit unique responses to climate change induced resource alterations [e.g., @Byrne2017], meaning we cannot simply look across space to predict the future.
-These two challenges motivate the use of *in situ* experimental manipulations of resource availability [@Avolio2015].
-
-Manipulating potentially limiting resources offers a route to understanding how ecosystems will respond to resource levels that fall outside the historical range of variability [@Knapp2017].
-Chronic alterations to resource availability should cause community composition to shift over time, the focal ecosystem is sensitive to the manipulated resource.
-Species' relative abundances are expected to re-order and, eventually, some species will be lost from the local community and new species will be gained [@Smith2009; @Avolio2015].
-On the heels of changes in community composition, subsequent changes in the level of ecosystem functioning may occur.
-
-@Smith2009 introduced the 'Hierarchical Response Framework' for understanding the interplay of community composition and ecosystem functioning in response to resource manipulations over time.
-In the near term, ecosystem functioning such as annual net primary productivity (ANPP) will reflect the physiological responses of individual species to the manipulated resource level.
-For example, ANPP may decline under simulated drought because the initial community consisted of drought-intolerant species [@Hoover2014].
-Over longer time spans, ecosystem functioning may recover as new species colonize or initial species reorder in relative abundance.
-For example, ANPP may initially decline, but eventually rise back to pre-treatment levels once drought-tolerant species become more abundant and compensate for drought-intolerant species [@Hoover2014].
-It is also possible that ecosystem functioning shifts to a new mean state, reflective of the suite of species in the new community [@Knapp2012].
-
-Much of the research on ecosystem and community responses to global climate change has focused on grassland systems, where soil water is typically a limiting resource.
-The sensitivity of ANPP to water availability in grasslands has been characterized spatially (across sites) and temporally (within sites).
-Across sites, there is a strong positive relationship between the amount of precipitation at a given site and mean ANPP [@Sala1988; @Huxman2004].
-Within sites, however, the response of ANPP to annual precipitation amount is much weaker [@Huxman2004; @Hsu2012].
-Furthermore, it is becoming clear that grassland ecosystem response to altered water availability is likely to be idiosyncratic, conforming to neither cross-site nor within-site ANPP-precipitation relationships [@Wilcox2016].
-This is because within-site relationships do not account for species reordering under chronic resource alterations, while cross-site relationships may overestimate the ability of species compensation.
-Thus, discovering which ecosystems are most sensitive to chronic alterations of water availability, and the time scales at which ecosystem and community responses occur, is critical to avoid surprises in the face of climate change.
----->
 
@@
 At each plot in each year, we took four readings of ground reflectances at the above wavelengths.
-We also took readings in ten calibration plots adjacent to the experimental site, in which we harvested all aboveground biomass, dried it to a constant weight at 60$^{\circ}$C, and weighed it to estimate ANPP.
+We also took readings in 12 (2015), 15 (2012, 2013, 2014), or 16 (2016) calibration plots adjacent to the experimental site, in which we harvested all aboveground biomass, dried it to a constant weight at 60$^{\circ}$C, and weighed it to estimate ANPP.
 
-For each plot and year, we averaged the four readings for each wavelength and then calculated NDVI using the MODIS and AVHRR algorithms.
-To convert NDVI to ANPP we regressed NDVI against the dry biomass weight from the ten calibration plots.
-We fit regressions to MODIS-based NDVI and AVHRR-based NDVI for each year and retained the regression with the better fit.
+For each plot and year, we averaged the four readings for each wavelength and then calculated a greenness index based on the same bands used to calculate NDVI using the MODIS and AVHRR bands for NDVI.
+To convert the greenness index to ANPP we regressed the greenness index against the dry biomass weight from the ten calibration plots.
+We fit regressions to MODIS-based index and AVHRR-based index for each year and retained the regression with the better fit.
 Using the best regression equation for each year, we predicted ANPP (Appendix 1).
@@
 Species composition data came from two sources: yearly census maps for each plot made using a pantograph [@Hill1920] and yearly counts of annual species in each plot.
-The maps record the spatial location of all individuals in the plot and the basal cover of each individual with cover greater than 1 cm.
-Using those annual maps, we aggregated over individuals to calculate total basal cover for each species in each plot.
+From those maps, we have data on the density of all annuals and perennials forbs, basal cover of perennial grasses, and canopy cover of shrubs.
 We made a large plot-treatment-year by species matrix, where columns were filled with either basal cover or density, depending on the measurement made for the particular species.
 So we could analyze the different types of data together, we standardized the values in each column.
-This puts all abundance values on the same scale, but comes with the limitation that all species are weighted equally.
-Nonetheless, this scaling approach allows a comprehensive view of community composition dynamics through time.
+This puts all abundance values on the same scale, meaning that common and rare species are weighted equally.
+Nonetheless, if we assume that rare species will respond more than common ones, then this approach is anti-conservative for detecting community change.
+This means that our approach is biased toward detecting compositional changes.\todo{Did I say this right?}
 
 ## Data Analysis
-Our main goal was to test whether the relationship between ANPP and growing season precipitation (hereafter, precipiation) differed among the drought, control, and irrigation treatments.
-To achieve this goal, we fit a multi-level random intercept and random slope regression with log(ANPP) as the response variabile and soil moisture (volumetric water content) as the sole predictor.
+Our main goal was to test whether the relationship between ANPP and soil moisture differed among the drought, control, and irrigation treatments.
+Based on our own observations and previous work at our study site [@Blaisdell1958; @Dalgleish2011; @Adler2012], we chose to use cumulative volumetric water content from March through June as our metric of soil moisture (hereafter referred to simply as `soil moisture').
+To achieve this goal, we fit a multi-level random intercept and random slope regression with log(ANPP) as the response variabile and soil moisture as the sole predictor.
+We log-transformed ANPP to account for heteroscedasticity.
 We fit the model under a Bayesian framework, allowing us to test for treatment differences by comparing the posterior distributions of the treatment-level coefficients [e.g., @Tredennick2013].
@@
 
-Our multi-level model has three grouping levels for coefficients, representing the nested structure of the data: (i) overall coefficients, (ii) treatment coefficients, and (iii) plot coefficients.
-Each subsequent level is drawn from the distribution of coefficients at the previous level.
-Formally, our model is defined as follows:
+Our model is defined as follows:
 
@@
 \begin{align}
-\mu_{i(j(k(t)))} &= \beta_{0,j(k)} + \beta_{1,j(k)}x_i + \gamma_t, \\
-y_{i(j(k(t)))} &\sim \text{Normal} \left(\mu_{i(j(k(t)))}, \sigma^2_{k} \right),
+\mu_{j(k(t))} &= \beta_{0,j(k)} + \beta_{1,j(k)}x_{k(t)} + \gamma_t, \\
+y_{j(k(t))} &\sim \text{Normal} \left(\mu_{j(k(t))}, \sigma^2_{k} \right),
 \end{align}
 
-\noindent{}where $\mu_{i(j(k))}$ is the deterministic prediction from the regression model for observation *i* for plot *j* associated with treatment *k* in year *t*, $\beta_{0,j(k)}$ is the intercept for plot *j* associated with treatment *k*, $\beta_{1,j(k)}$ is the slope term for the effect of soil moisture for plot *j* associated with treatment *k*, $\gamma_t$ is the intercept offset for year *t*, and $\sigma^2_k$ is the process variance for treatment *k*.
-Data include the standardized log(ANPP) observations $\left( y_{i(j(k(t)))} \right)$ and soil moisture ($x_i$).
-Although we include observation subscript *i* on the *x*s, observations within a treatment-year all share the same soil moisture values.
+\noindent{}where $\mu_{j(k(t))}$ is the deterministic prediction from the regression model for plot *j* associated with treatment *k* in year *t*, $\beta_{0,j(k)}$ is the intercept for plot *j* associated with treatment *k*, $\beta_{1,j(k)}$ is the slope term for the effect of soil moisture for plot *j* associated with treatment *k*, $\gamma_t$ is the intercept offset for year *t*, and $\sigma^2_k$ is the process variance for treatment *k*.
+Data include the standardized log(ANPP) observations $\left( y_{j(k(t))} \right)$ and soil moisture ($x_{k(t)}$).
 
 The intercept and slope terms are modeled hierarchichally to account for the non-independence of observations across years within plots and to allow us to test the hypothesis that our treatments alter the ANPP-soil moisture relationship.
-As noted above, plot-level coefficients are drawn from treatment-level coefficients, which are drawn from overall coefficients.
+Thus, our multi-level model has three grouping levels for coefficients, representing the nested structure of the data: (i) overall coefficients, (ii) treatment coefficients, and (iii) plot coefficients.
+Each subsequent level is drawn from the distribution of coefficients at the previous level, meaning plot-level coefficients are drawn from treatment-level coefficients, which are drawn from overall coefficients.
 We also include a covariance structure among the intercept and slope at each level.
@@
 
-Cumulative growing season soil moisture had a positive effect on ANPP (mean of $\beta_{1}$ = `r round(betas$mean[8],2)`; 80\% BCI = `r round(betas$X10.[8],2)`, `r round(betas$X90.[8],2)`; 95\% BCI = `r round(betas$X2.5.[8],2)`, `r round(betas$X97.5.[8],2)`) (Fig.
+Cumulative March-June soil moisture had a weak positive effect on ANPP (mean of $\beta_{1}$ = `r round(betas$mean[8],2)`), but the effect is associated with high uncertainty, with credible intervals that overlap zerp (80\% BCI = `r round(betas$X10.[8],2)`, `r round(betas$X90.[8],2)`; 95\% BCI = `r round(betas$X2.5.[8],2)`, `r round(betas$X97.5.[8],2)`) (Fig.
 3).
@@
 Kleinhesselink, unpublished data], meaning a consistent ecosystem functional response at low soil moisture is perfectly reasonable.
-For example, many of the perennial grasses in our focal ecosystem avoid drought stress by growing early in the growing season (A.R. Kleinhesselink, personal observation).
-Likewise, the dominant shrub in our focal ecosystem, *Artemisia tripartita*, has access to water deep in the soil profile thanks to a deep root system [@Germino2014].
+For example, many of the perennial grasses in our focal ecosystem avoid drought stress by growing early in the growing season [@Blaisdell1958, A.R. Kleinhesselink, personal observation].
+Likewise, the dominant shrub in our focal ecosystem, *Artemisia tripartita*, has access to water deep in the soil profile thanks to a deep root system (CITATION).
 As for the irrigation treatments, the lack of a saturating response (e.g., a lower regression slope) suggests water availability remained a limiting factor at the levels of precipitation we were able to manipulate.

### 2017-07-26 manuscript/usses_water_ms.tex
addresses PBA minor comments

@@
 
-\usepackage{lipsum} \usepackage{setspace} \usepackage{todonotes}
-\usepackage{rotating} \usepackage{color, soul} \usepackage{sectsty}
-\usepackage{bm,mathrsfs}
+\usepackage{setspace} \usepackage{todonotes} \usepackage{rotating}
+\usepackage{color, soul} \usepackage{sectsty} \usepackage{bm,mathrsfs}
 
@@
 \begin{abstract}
-Precipitation is predicted to become more variable in the western U.S.,
-meaning years of above and below average precipitation will become more
-common. Such periods of drought and deluge are major drivers of plant
-community dynamics and ecosystem functioning in water limited
-grasslands. Here we report the results of a five-year experiment where
-we used drought and irrigation treatments (50\% decrease/increase) to
-see how a sagebrush steppe plant community in Idaho will respond to
-future climate changes. Ecosystem functional response, defined as the
-relationship between soil moisture and aboveground net primary
-productivity, was suprisingly consistent across drought, control, and
-irrigation treatments. Aboveground net primary productivity (ANPP)
-responded positively to availabile soil moisture, but the response did
-not vary across treatments. There was also no evidence that treatment
-effects grew over time. The similarity of ecosystem functional response
-across treatment was not due to compensatory shifts at the plant
-community level, where species composition among treatments was similar
-and remarkably stable over the five years. At least in the short-term,
-ecosystem functional response and community composition in this
-sagebrush steppe system is resistant to increases and decreases in
-growing season precipitation.
+Precipitation is predicted to become more variable in the western United
+States, meaning years of above and below average precipitation will
+become more common. Periods of drought and deluge are major drivers of
+ecosystem functioning in water limited grasslands, but how ecosystems
+respond to precipitation may change as the duration of above and below
+average periods increases. Changes in ecosystem functional response
+could reflect underlying changes in species composition or species
+reaching physiological thresholds at extreme precipitation levels. Here
+we report the results of a five-year experiment where we used drought
+and irrigation treatments (approximately 50\% decrease/increase) to
+investigate whether ecosystem functional response remains consistent at
+extreme precipitation levels. We also use species-level adundance data
+to link ecosystem functional response to community dynamics. Ecosystem
+functional response, defined as the relationship between soil moisture
+and aboveground net primary productivity, was suprisingly consistent
+across drought, control, and irrigation treatments. Aboveground net
+primary productivity (ANPP) responded positively to availabile soil
+moisture, but the response was weak and did not vary across treatments.
+There was also no evidence that treatment effects grew over time. The
+similarity of ecosystem functional response across treatments was not
+due to compensatory shifts at the plant community level, where species
+composition among treatments was similar and remarkably stable over the
+five years. At least in the short-term, ecosystem functional response
+and community composition in this sagebrush steppe system is resistant
+to increases and decreases in growing season precipitation, perhaps due
+to bet-hedging strategies of constituent species.
 % Dummy abstract text.
@@
 
+\reversemarginpar
+
 \section{INTRODUCTION}\label{introduction}
+
+Projected changes in precipitation regimes, associated with global
+climate change, will affect
 
@@
 manipulation experiment in a sagebrush steppe grassland. We imposed
-drought and irrigation treatments (50\% decrease/increase) and measured
-ecosystem (ANPP) and community (species composition) responses. We focus
-on how the drought and irrigation treatments affect the relationship
-between available soil moisture and ANPP, and if community dynamics
-underly the ecosystem responses. In particular, we are interested in the
-consistency of the soil moisture-ANPP relationship among treatments. Is
-the relationship steeper under the drought treatment, at low soil
-moisture? Does the relationship saturate under the irrigation treatment,
-at high soil moisture? To answer these questions we fit a random
-intercepts, random slopes model to test whether the regressions differed
-among treatments. We also analyzed community composition over time,
-allowing us to place our experimental results within the framework of
-the HRF and our competing predictions (Fig. 1).
+drought and irrigation treatments (approximately \(\pm50\%\)) and
+measured ecosystem (ANPP) and community (species composition) responses.
+We focus on how the drought and irrigation treatments affect the
+relationship between available soil moisture and ANPP, and if community
+dynamics underly the ecosystem responses. In particular, we are
+interested in the consistency of the soil moisture-ANPP relationship
+among treatments. Is the relationship steeper under the drought
+treatment, at low soil moisture? Does the relationship saturate under
+the irrigation treatment, at high soil moisture? To answer these
+questions we fit a random intercepts, random slopes model to test
+whether the regressions differed among treatments. We also analyzed
+community composition over time, allowing us to place our experimental
+results within the framework of the HRF and our competing predictions
+(Fig. 1).
 
@@
 each plot in each year, we took four readings of ground reflectances at
-the above wavelengths. We also took readings in ten calibration plots
-adjacent to the experimental site, in which we harvested all aboveground
-biomass, dried it to a constant weight at 60\(^{\circ}\)C, and weighed
-it to estimate ANPP.
+the above wavelengths. We also took readings in 12 (2015), 15 (2012,
+2013, 2014), or 16 (2016) calibration plots adjacent to the experimental
+site, in which we harvested all aboveground biomass, dried it to a
+constant weight at 60\(^{\circ}\)C, and weighed it to estimate ANPP.
 
 For each plot and year, we averaged the four readings for each
-wavelength and then calculated NDVI using the MODIS and AVHRR
-algorithms. To convert NDVI to ANPP we regressed NDVI against the dry
-biomass weight from the ten calibration plots. We fit regressions to
-MODIS-based NDVI and AVHRR-based NDVI for each year and retained the
-regression with the better fit. Using the best regression equation for
-each year, we predicted ANPP (Appendix 1).
+wavelength and then calculated a greenness index based on the same bands
+used to calculate NDVI using the MODIS and AVHRR bands for NDVI. To
+convert the greenness index to ANPP we regressed the greenness index
+against the dry biomass weight from the ten calibration plots. We fit
+regressions to MODIS-based index and AVHRR-based index for each year and
+retained the regression with the better fit. Using the best regression
+equation for each year, we predicted ANPP (Appendix 1).
 
@@
 each plot made using a pantograph \citep{Hill1920} and yearly counts of
-annual species in each plot. The maps record the spatial location of all
-individuals in the plot and the basal cover of each individual with
-cover greater than 1 cm. Using those annual maps, we aggregated over
-individuals to calculate total basal cover for each species in each
-plot. We made a large plot-treatment-year by species matrix, where
-columns were filled with either basal cover or density, depending on the
-measurement made for the particular species. So we could analyze the
-different types of data together, we standardized the values in each
-column. This puts all abundance values on the same scale, but comes with
-the limitation that all species are weighted equally. Nonetheless, this
-scaling approach allows a comprehensive view of community composition
-dynamics through time.
+annual species in each plot. From those maps, we have data on the
+density of all annuals and perennials forbs, basal cover of perennial
+grasses, and canopy cover of shrubs. We made a large plot-treatment-year
+by species matrix, where columns were filled with either basal cover or
+density, depending on the measurement made for the particular species.
+So we could analyze the different types of data together, we
+standardized the values in each column. This puts all abundance values
+on the same scale, meaning that common and rare species are weighted
+equally. Nonetheless, if we assume that rare species will respond more
+than common ones, then this approach is anti-conservative for detecting
+community change. This means that our approach is biased toward
+detecting compositional changes.\todo{Did I say this right?}
 
@@
 
-Our main goal was to test whether the relationship between ANPP and
-growing season precipitation (hereafter, precipiation) differed among
-the drought, control, and irrigation treatments. To achieve this goal,
-we fit a multi-level random intercept and random slope regression with
-log(ANPP) as the response variabile and soil moisture (volumetric water
-content) as the sole predictor. We fit the model under a Bayesian
-framework, allowing us to test for treatment differences by comparing
-the posterior distributions of the treatment-level coefficients
+Our main goal was to test whether the relationship between ANPP and soil
+moisture differed among the drought, control, and irrigation treatments.
+Based on our own observations and previous work at our study site
+\citep{Blaisdell1958, Dalgleish2011, Adler2012}, we chose to use
+cumulative volumetric water content from March through June as our
+metric of soil moisture (hereafter referred to simply as `soil
+moisture'). To achieve this goal, we fit a multi-level random intercept
+and random slope regression with log(ANPP) as the response variabile and
+soil moisture as the sole predictor. We log-transformed ANPP to account
+for heteroscedasticity. We fit the model under a Bayesian framework,
+allowing us to test for treatment differences by comparing the posterior
+distributions of the treatment-level coefficients
 \citep[e.g.,][]{Tredennick2013}. Both log(ANPP) and soil moisture were
@@
 
-Our multi-level model has three grouping levels for coefficients,
-representing the nested structure of the data: (i) overall coefficients,
-(ii) treatment coefficients, and (iii) plot coefficients. Each
-subsequent level is drawn from the distribution of coefficients at the
-previous level. Formally, our model is defined as follows:
+Our model is defined as follows:
 
@@
 \begin{align}
-\mu_{i(j(k(t)))} &= \beta_{0,j(k)} + \beta_{1,j(k)}x_i + \gamma_t, \\
-y_{i(j(k(t)))} &\sim \text{Normal} \left(\mu_{i(j(k(t)))}, \sigma^2_{k} \right),
+\mu_{j(k(t))} &= \beta_{0,j(k)} + \beta_{1,j(k)}x_{k(t)} + \gamma_t, \\
+y_{j(k(t))} &\sim \text{Normal} \left(\mu_{j(k(t))}, \sigma^2_{k} \right),
 \end{align}
 
-\noindent{}where \(\mu_{i(j(k))}\) is the deterministic prediction from
-the regression model for observation \emph{i} for plot \emph{j}
-associated with treatment \emph{k} in year \emph{t}, \(\beta_{0,j(k)}\)
-is the intercept for plot \emph{j} associated with treatment \emph{k},
-\(\beta_{1,j(k)}\) is the slope term for the effect of soil moisture for
-plot \emph{j} associated with treatment \emph{k}, \(\gamma_t\) is the
-intercept offset for year \emph{t}, and \(\sigma^2_k\) is the process
-variance for treatment \emph{k}. Data include the standardized log(ANPP)
-observations \(\left( y_{i(j(k(t)))} \right)\) and soil moisture
-(\(x_i\)). Although we include observation subscript \emph{i} on the
-\emph{x}s, observations within a treatment-year all share the same soil
-moisture values.
+\noindent{}where \(\mu_{j(k(t))}\) is the deterministic prediction from
+the regression model for plot \emph{j} associated with treatment
+\emph{k} in year \emph{t}, \(\beta_{0,j(k)}\) is the intercept for plot
+\emph{j} associated with treatment \emph{k}, \(\beta_{1,j(k)}\) is the
+slope term for the effect of soil moisture for plot \emph{j} associated
+with treatment \emph{k}, \(\gamma_t\) is the intercept offset for year
+\emph{t}, and \(\sigma^2_k\) is the process variance for treatment
+\emph{k}. Data include the standardized log(ANPP) observations
+\(\left( y_{j(k(t))} \right)\) and soil moisture (\(x_{k(t)}\)).
 
@@
 allow us to test the hypothesis that our treatments alter the ANPP-soil
-moisture relationship. As noted above, plot-level coefficients are drawn
-from treatment-level coefficients, which are drawn from overall
+moisture relationship. Thus, our multi-level model has three grouping
+levels for coefficients, representing the nested structure of the data:
+(i) overall coefficients, (ii) treatment coefficients, and (iii) plot
+coefficients. Each subsequent level is drawn from the distribution of
+coefficients at the previous level, meaning plot-level coefficients are
+drawn from treatment-level coefficients, which are drawn from overall
 coefficients. We also include a covariance structure among the intercept
@@
 
-Cumulative growing season soil moisture had a positive effect on ANPP
-(mean of \(\beta_{1}\) = 0.41; 80\% BCI = -0.08, 0.86; 95\% BCI = -0.45,
-1.13) (Fig. 3). Ecosystem functional response was similar among
-treatments, with treatment level regressions between soil moisture and
-log(ANPP) having similar intercepts and slopes (Fig. 3A). There was also
-no evidence that the treatment effects became more important over time
-because there was no directional trend in the random year effects (Fig.
-4).
+Cumulative March-June soil moisture had a weak positive effect on ANPP
+(mean of \(\beta_{1}\) = 0.41), but the effect is associated with high
+uncertainty, with credible intervals that overlap zerp (80\% BCI =
+-0.08, 0.86; 95\% BCI = -0.45, 1.13) (Fig. 3). Ecosystem functional
+response was similar among treatments, with treatment level regressions
+between soil moisture and log(ANPP) having similar intercepts and slopes
+(Fig. 3A). There was also no evidence that the treatment effects became
+more important over time because there was no directional trend in the
+random year effects (Fig. 4).
 
@@
 focal ecosystem avoid drought stress by growing early in the growing
-season (A.R. Kleinhesselink, personal observation). Likewise, the
-dominant shrub in our focal ecosystem, \emph{Artemisia tripartita}, has
-access to water deep in the soil profile thanks to a deep root system
-\citep{Germino2014}. As for the irrigation treatments, the lack of a
-saturating response (e.g., a lower regression slope) suggests water
-availability remained a limiting factor at the levels of precipitation
-we were able to manipulate.
+season \citep[A.R. Kleinhesselink, personal observation]{Blaisdell1958}.
+Likewise, the dominant shrub in our focal ecosystem, \emph{Artemisia
+tripartita}, has access to water deep in the soil profile thanks to a
+deep root system \citep{Germino2014}. As for the irrigation treatments,
+the lack of a saturating response (e.g., a lower regression slope)
+suggests water availability remained a limiting factor at the levels of
+precipitation we were able to manipulate.
 

### 2017-07-26 code/calibrate_radiometer_by_year.R

@@ -119,7 +119,8 @@ get_regression <- function(ndvi_data, biomass_data, plots = FALSE){
                          ndvi_slope  = round(fit_summary[which(fit_summary$term=="ndvi"), "estimate"],2),
                          rsquare     = round(glance(fit)$r.squared,2),
                          min_biomass = min(combo_data$biomass_grams),
-                         max_biomass = max(combo_data$biomass_grams))
+                         max_biomass = max(combo_data$biomass_grams),
+                         num_plots   = nrow(combo_data))
   
   if(plots){
     gg <- ggplot(combo_data, aes(x=ndvi, y=biomass_grams)) +

### 2017-07-26 manuscript/usses_water_ms.Rmd
addresses ARK minor comments

@@
 Periods of drought and deluge are major drivers of ecosystem functioning in water limited grasslands, but how ecosystems respond to precipitation may change as the duration of above and below average periods increases.
-Changes in ecosystem functional response could reflect underlying changes in species composition or species reaching physiological thresholds at extreme precipitation levels.
+Changes in ecosystem functional response could reflect compensatory changes in species composition or species reaching physiological thresholds at extreme precipitation levels.
 Here we report the results of a five-year experiment where we used drought and irrigation treatments (approximately 50% decrease/increase) to investigate whether ecosystem functional response remains consistent at extreme precipitation levels.
@@
 The similarity of ecosystem functional response across treatments was not due to compensatory shifts at the plant community level, where species composition among treatments was similar and remarkably stable over the five years.
-At least in the short-term, ecosystem functional response and community composition in this sagebrush steppe system is resistant to increases and decreases in growing season precipitation, perhaps due to bet-hedging strategies of constituent species.
+At least in the short-term, ecosystem functional response and community composition in this sagebrush steppe system is consistent, even at extreme precipitation levels.
+Such consistency may be due to bet-hedging strategies of constituent species.
 header-includes: >
@@
 For example, historical trends may show a linear soil moisture-ANPP relationship, but that relationship could potentially saturate if soil moisture is pushed far beyond typical levels.
-Saturating relationships are actually common [@Hsu2012], perhaps because other resources, like nitrogen, become more limiting in wet years than dry years.
+Saturating relationships are actually common [@Hsu2012; @Gherardi2015a], perhaps because other resources, like nitrogen, become more limiting in wet years than dry years.
 Knowing the curvature of the soil moisture-ANPP relationship is critical for understanding how ecosystems will respond to chronic alterations in water availability.
@@
 
-Manipulating potentially limiting resources offers a route to understanding how ecosystems will respond to resource levels that fall outside the historical range of variability [@Avolio2015; @Gherardi2015; @Knapp2017].
+Manipulating potentially limiting resources like precipitation offers a route to understanding how ecosystems will respond to resource levels that fall outside the historical range of variability [@Avolio2015; @Gherardi2015; @Knapp2017].
 Altering the amount of precipitation over many years should provide insight into the time scales at which water-limited ecosystems respond to chronic resource alteration.
@@
 Thus, our task is not to test the validity of the HRF, but rather to amass information on how quickly ecosystem functional responses change in different ecosystems.
-Likewise, we need to understand whether changes at the ecosystem level are driven community level changes or individual level responses.
+Likewise, we need to understand whether changes at the ecosystem level are driven by community level changes or individual level responses.
 
@@
 ## Study Area
-We conducted our precipitation manipulation experiment at the United States Sheep Experimental Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
-The vegetation is typical of high elevation sagebrush steppe.
+We conducted our precipitation manipulation experiment in a sagebrush steppe community at the United States Sheep Experimental Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
 The plant community is dominated by the shrub *Artemesia tripartita* and three perennial bunchgrasses, *Pseudoroegneria spicata*, *Poa secunda*, and *Hesperostipa comata*.
@@
 We use these six plots as control plots that have recieved no treatment, just ambient precipitation.
-In spring 2011, we (well, two of us [A. Kleinhesselink and P. Adler]) established 16 new 1 m$^2$ plots.
+In spring 2011, we (well, two of us [A. Kleinhesselink and P. Adler]) established 16 new 1 m$^2$ plots located in one of the same exclosures as 6 of the original vegetation monitoring plots.
 We avoided areas on steep hill slopes, areas with greater than 20\% cover of bare rock, and areas with greater than 10\% cover of the shrubs *Purshia tridentata* and/or *Amelanchier utahensis*.
@@
 
-To make sure the treatments were having the desired effects, we monitored soil moisture in four of the drought-irrigation pairs using Decagon Devices (Pulman, Washington) 5TM and EC-5 soil moisture sensors.
-We installed four sensors in each plot, two at 5 cm soil depth and two at 25 cm soil depth.
+We monitored soil moisture in four of the drought-irrigation pairs using Decagon Devices (Pulman, Washington) 5TM and EC-5 soil moisture sensors.
+We installed four sensors around the edges of each plot, two at 5 cm soil depth and two at 25 cm soil depth.
 We also installed four sensors in areas nearby the four selected plot pairs to measure ambient soil moisture at the same depths.
@@
 
-\ark{ANDY: add paragraph on the statistical modeling of soil moisture to produce Figure 2B (same as your Figure 2 in the precip-demography-prediction dissertation chapter).}
+Analyzing the response to experimental treatments was complicated by the fact that we did not directly monitor soil moisture in each plot on each day of the experiment.  
+Only a subset of plots were equiped with soil moisture sensors, and within those plots, one or more of the sensors frequently failed to collect data. 
+To remedy these problems, and to produce average daily soil moisture values for the ambient, drought, and irrigation conditions, we used a statistical model to predict the average treatment effects on soil moisture during the course of the experiment.
+
+We first averaged the observed soil moisture for each day and within each plot. 
+Then we standardized the averages within each plot group by subtracting the average ambient soil moisture in that plot group and dividing by the standard deviation of the ambient soil moisture in that plot group.
+We then found the difference between the standardized ambient soil moisture and the standardized drought and irrigation soil moisture within each plot group. 
+These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.
+
+We then modeled the daily deviation from ambient conditions of the drought and irrigation treatments using a linear mixed effects model with independent variables for treatment, season (winter, spring, summer, fall), rainfall, and all two-way interactions.
+Rainy days were defined as any day in which precipitation was recorded and average temperature was above 3$^{\circ}$C. 
+The day immediately following rainfall was also classified as rainy. 
+We fit the model using the 'lme4::lmer()' function [@Bates2015] in R [@R2016], with random effects for plot group and date.
+We weighted observations by the number of unique sensors or spot measurements that were taken in each plot on that day. 
+We then used the model to predict the average daily soil moisture in the treated plots based on the average daily ambient soil moisture.  
+We could only predict soil moisture in the treated plots on days for which we took at least one ambient soil moisture measurement (Fig. 2B).
 
@@
   \centering
-      \includegraphics[height=5in]{../figures/Figure1.png}
+      \includegraphics[height=7in]{../figures/Figure1.png}
   \caption{(A) Probability density of historical precipitation from 1926-2016, with the years of the experiment shown with arrows on the \emph{x}-axis.

### 2017-07-26 manuscript/usses_water_ms.tex
addresses ARK minor comments

@@
 average periods increases. Changes in ecosystem functional response
-could reflect underlying changes in species composition or species
+could reflect compensatory changes in species composition or species
 reaching physiological thresholds at extreme precipitation levels. Here
@@
 five years. At least in the short-term, ecosystem functional response
-and community composition in this sagebrush steppe system is resistant
-to increases and decreases in growing season precipitation, perhaps due
-to bet-hedging strategies of constituent species.
+and community composition in this sagebrush steppe system is consistent,
+even at extreme precipitation levels. Such consistency may be due to
+bet-hedging strategies of constituent species.
 % Dummy abstract text.
@@
 saturate if soil moisture is pushed far beyond typical levels.
-Saturating relationships are actually common \citep{Hsu2012}, perhaps
-because other resources, like nitrogen, become more limiting in wet
-years than dry years. Knowing the curvature of the soil moisture-ANPP
-relationship is critical for understanding how ecosystems will respond
-to chronic alterations in water availability.
+Saturating relationships are actually common
+\citep{Hsu2012, Gherardi2015a}, perhaps because other resources, like
+nitrogen, become more limiting in wet years than dry years. Knowing the
+curvature of the soil moisture-ANPP relationship is critical for
+understanding how ecosystems will respond to chronic alterations in
+water availability.
 
@@
 
-Manipulating potentially limiting resources offers a route to
-understanding how ecosystems will respond to resource levels that fall
-outside the historical range of variability
+Manipulating potentially limiting resources like precipitation offers a
+route to understanding how ecosystems will respond to resource levels
+that fall outside the historical range of variability
 \citep{Avolio2015, Gherardi2015, Knapp2017}. Altering the amount of
@@
 we need to understand whether changes at the ecosystem level are driven
-community level changes or individual level responses.
+by community level changes or individual level responses.
 
@@
 
-We conducted our precipitation manipulation experiment at the United
-States Sheep Experimental Station (USSES) near Dubois, Idaho
-(44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m above sea level. The
-vegetation is typical of high elevation sagebrush steppe. The plant
-community is dominated by the shrub \emph{Artemesia tripartita} and
-three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa
-secunda}, and \emph{Hesperostipa comata}. During the period of our
-experiment (2011 -- 2015), average mean annual precipitation was 265 mm
+We conducted our precipitation manipulation experiment in a sagebrush
+steppe community at the United States Sheep Experimental Station (USSES)
+near Dubois, Idaho (44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m
+above sea level. The plant community is dominated by the shrub
+\emph{Artemesia tripartita} and three perennial bunchgrasses,
+\emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and
+\emph{Hesperostipa comata}. During the period of our experiment (2011 --
+2015), average mean annual precipitation was 265 mm
 year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
@@
 us {[}A. Kleinhesselink and P. Adler{]}) established 16 new 1 m\(^2\)
-plots. We avoided areas on steep hill slopes, areas with greater than
-20\% cover of bare rock, and areas with greater than 10\% cover of the
-shrubs \emph{Purshia tridentata} and/or \emph{Amelanchier utahensis}. We
-established the new plots in pairs and randomly assigned each plot in a
-pair to receive a ``drought'' or ``irrigation'' treatment.
+plots located in one of the same exclosures as 6 of the original
+vegetation monitoring plots. We avoided areas on steep hill slopes,
+areas with greater than 20\% cover of bare rock, and areas with greater
+than 10\% cover of the shrubs \emph{Purshia tridentata} and/or
+\emph{Amelanchier utahensis}. We established the new plots in pairs and
+randomly assigned each plot in a pair to receive a ``drought'' or
+``irrigation'' treatment.
 
@@
 
-To make sure the treatments were having the desired effects, we
-monitored soil moisture in four of the drought-irrigation pairs using
+We monitored soil moisture in four of the drought-irrigation pairs using
 Decagon Devices (Pulman, Washington) 5TM and EC-5 soil moisture sensors.
-We installed four sensors in each plot, two at 5 cm soil depth and two
-at 25 cm soil depth. We also installed four sensors in areas nearby the
-four selected plot pairs to measure ambient soil moisture at the same
-depths. Soil moisture measurements were automatically logged every four
-hours. We coupled this temporally intensive soil moisture sampling with
-spatially extensive readings taken at six points within all 16 plots and
-associated ambient measurement areas. These snapshot data were collected
-on 06/06/2012, 04/29/2015, 05/07/2015, 06/09/2015, and 05/10/2016 using
-a handheld EC-5 sensor.
+We installed four sensors around the edges of each plot, two at 5 cm
+soil depth and two at 25 cm soil depth. We also installed four sensors
+in areas nearby the four selected plot pairs to measure ambient soil
+moisture at the same depths. Soil moisture measurements were
+automatically logged every four hours. We coupled this temporally
+intensive soil moisture sampling with spatially extensive readings taken
+at six points within all 16 plots and associated ambient measurement
+areas. These snapshot data were collected on 06/06/2012, 04/29/2015,
+05/07/2015, 06/09/2015, and 05/10/2016 using a handheld EC-5 sensor.
 
-\ark{ANDY: add paragraph on the statistical modeling of soil moisture to produce Figure 2B (same as your Figure 2 in the precip-demography-prediction dissertation chapter).}
+Analyzing the response to experimental treatments was complicated by the
+fact that we did not directly monitor soil moisture in each plot on each
+day of the experiment.\\
+Only a subset of plots were equiped with soil moisture sensors, and
+within those plots, one or more of the sensors frequently failed to
+collect data. To remedy these problems, and to produce average daily
+soil moisture values for the ambient, drought, and irrigation
+conditions, we used a statistical model to predict the average treatment
+effects on soil moisture during the course of the experiment.
+
+We first averaged the observed soil moisture for each day and within
+each plot. Then we standardized the averages within each plot group by
+subtracting the average ambient soil moisture in that plot group and
+dividing by the standard deviation of the ambient soil moisture in that
+plot group. We then found the difference between the standardized
+ambient soil moisture and the standardized drought and irrigation soil
+moisture within each plot group. These transformations ensured that the
+treatment effects in each plot were appropriately scaled by the local
+ambient conditions within each plot group.
+
+We then modeled the daily deviation from ambient conditions of the
+drought and irrigation treatments using a linear mixed effects model
+with independent variables for treatment, season (winter, spring,
+summer, fall), rainfall, and all two-way interactions.\\
+Rainy days were defined as any day in which precipitation was recorded
+and average temperature was above 3\(^{\circ}\)C. The day immediately
+following rainfall was also classified as rainy. We fit the model using
+the `lme4::lmer()' function \citep{Bates2015} in R \citep{R2016}, with
+random effects for plot group and date. We weighted observations by the
+number of unique sensors or spot measurements that were taken in each
+plot on that day. We then used the model to predict the average daily
+soil moisture in the treated plots based on the average daily ambient
+soil moisture.\\
+We could only predict soil moisture in the treated plots on days for
+which we took at least one ambient soil moisture measurement (Fig. 2B).
 
@@
 tripartita}, has access to water deep in the soil profile thanks to a
-deep root system \citep{Germino2014}. As for the irrigation treatments,
-the lack of a saturating response (e.g., a lower regression slope)
-suggests water availability remained a limiting factor at the levels of
-precipitation we were able to manipulate.
+deep root system (CITATION). As for the irrigation treatments, the lack
+of a saturating response (e.g., a lower regression slope) suggests water
+availability remained a limiting factor at the levels of precipitation
+we were able to manipulate.
 

### 2017-07-26 manuscript/usses_water_ms.Rmd
rewriting discussion

@@
   - code: 3
-    address: USSES, address...
+    address: United States Department of Agriculture, Agriculture Research Station, U.S. Sheep Experiment Station, Dubois, Idaho 83423
 abstract: >
@@
 
-Projected changes in precipitation regimes, associated with global climate change, will affect
-
-Water availability is a strong driver of aboveground net primary productivity (ANPP) in grassland ecosystems.
-Mean annual precipitation (MAP) is strongly correlated with ANPP across sites [@Huxman2004], and, in water-limited ecosystems, interannual variation in MAP is also correlated with interannual variation in ANPP [@Hsu2012].
-These relationships suggest that chronic alterations in levels of precipitation, associated with global climate change, will impact ANPP.
+Ecosystems worldwide are projected to experience either wetter or drier conditions in the future [@Zhang2007; @Greve2014].
+This has important implications for ecosystem functioning because mean annual net primary productivity (ANPP) positively covaries with mean annual precipitation across ecosystems [@Sala1988; @Huxman2004] and interannual responses of ANPP are linked to interannual variability in precipitation within ecosystems [@Lauenroth1992; @Hsu2012; @Gherardi2015a].
+To predict how ecosystems might respond to extreme precipitation levels, we need to understand whether ecosystems tend to follow temporal trends, which assume relatively constant plant community composition, or spatial trends, which allow community composition to shift [@Wilcox2016].
 
@@
 However, the fitted functional response may provide an incomplete picture because future conditions are likely to be outside the historical range of variability [@Smith2011].
-For example, historical trends may show a linear soil moisture-ANPP relationship, but that relationship could potentially saturate if soil moisture is pushed far beyond typical levels.
+For example, historical trends may underestimate the potential for the soil moisture-ANPP relationship to saturate if soil moisture is pushed far beyond typical levels.
 Saturating relationships are actually common [@Hsu2012; @Gherardi2015a], perhaps because other resources, like nitrogen, become more limiting in wet years than dry years.
-Knowing the curvature of the soil moisture-ANPP relationship is critical for understanding how ecosystems will respond to chronic alterations in water availability.
+Knowing the curvature of the soil moisture-ANPP relationship at extreme precipitation levels is critical for understanding how ecosystems will respond to chronic alterations in water availability.
 
@@
 
-Manipulating potentially limiting resources like precipitation offers a route to understanding how ecosystems will respond to resource levels that fall outside the historical range of variability [@Avolio2015; @Gherardi2015; @Knapp2017].
+Manipulating potentially limiting resources, like precipitation, offers a route to understanding how ecosystems will respond to resource levels that fall outside the historical range of variability [@Avolio2015; @Gherardi2015; @Knapp2017].
 Altering the amount of precipitation over many years should provide insight into the time scales at which water-limited ecosystems respond to chronic resource alteration.
-Following the HRF, we propose four alternative predictions for the effect of precipitation manipulation on the ecosystem functional response to soil moisture, that is, the soil moisture-ANPP relationship (Figure 1).
+Following the HRF, we propose four alternative predictions for the effect of precipitation manipulation on the ecosystem functional response to soil moisture, that is, the soil moisture-ANPP relationship (Fig.
+1).
 The four predictions are based on possible outcomes at the community (e.g., community composition) and ecosystem (e.g., soil moisture-ANPP regression) levels.
@@
 We use these six plots as control plots that have recieved no treatment, just ambient precipitation.
-In spring 2011, we (well, two of us [A. Kleinhesselink and P. Adler]) established 16 new 1 m$^2$ plots located in one of the same exclosures as 6 of the original vegetation monitoring plots.
+In spring 2011, we (well, two of us [A. Kleinhesselink and P. Adler]) established 16 new 1 m$^2$ plots located in the same exclosure as the six control plots.
 We avoided areas on steep hill slopes, areas with greater than 20\% cover of bare rock, and areas with greater than 10\% cover of the shrubs *Purshia tridentata* and/or *Amelanchier utahensis*.
@@
 
-Cumulative March-June soil moisture had a weak positive effect on ANPP (mean of $\beta_{1}$ = `r round(betas$mean[8],2)`), but the effect is associated with high uncertainty, with credible intervals that overlap zerp (80\% BCI = `r round(betas$X10.[8],2)`, `r round(betas$X90.[8],2)`; 95\% BCI = `r round(betas$X2.5.[8],2)`, `r round(betas$X97.5.[8],2)`) (Fig.
-3).
+Cumulative March-June soil moisture had a weak positive effect on ANPP (mean of $\beta_{1}$ = `r round(betas$mean[8],2)`; Fig.
+3), but the effect is associated with high uncertainty, with credible intervals that overlap zerp (80\% BCI = `r round(betas$X10.[8],2)`, `r round(betas$X90.[8],2)`; 95\% BCI = `r round(betas$X2.5.[8],2)`, `r round(betas$X97.5.[8],2)`).
 Ecosystem functional response was similar among treatments, with treatment level regressions between soil moisture and log(ANPP) having similar intercepts and slopes (Fig. 3A).
@@
 
-Ecosystem response to chronic resource alteration is expected to follow a temporal trend.
-Initially, ecosystem response will be modest and reflect the physiological responses of constituent species.
-Over longer time periods, species reordering will cause greater responses as species better able to take advantage of, or cope with, new resource levels become more abundant.
-This temporal trend is formalized by the 'Hierarchical Response Framework' [HRF, @Smith2009], and has been empirically supported [@Knapp2012; @Wilcox2016].
-A lingering question is not whether the HRF is a reasonable model of ecosystem dynamics, but rather how the time scales of ecosystem response and community change differ among ecosystems.
-Indeed, previous work has shown that community compositional shifts can be both rapid, on the order of years [@Hoover2014], and slow, on order of decades [@Knapp2012; @Wilcox2016].
-To add to this growing body of knowledge, we performed a precipitation manipulation experiment in a sagebrush steppe ecosystem.
-
-Ecosystem functional response (i.e., the relationship between soil moisture and ANPP) and community composition in this sagebrush steppe ecosystem was consistent across drought, control, and irrigation treatments (Fig.
-2D, Fig.
-3).
-In other words, functional responses estimated from any precipitation level sampled apply to any other level of precipitation, and the consistent functional response was not due to species reordering and compensation.
-Thus, our results conform to the first of our four predictions: neither ecosystem functional response nor community composition changed with chronic alteration in water availability (Fig.
+Ecosystem response to precipitation extremes depends on the physiological responses of constituent species and the rate at which community composition shifts to favor species better able to take advantage of, or cope with, new resource levels [@Smith2009].
+Previous work has shown that community compositional shifts can be both rapid, on the order of years [@Hoover2014], and slow, on order of decades [@Knapp2012; @Wilcox2016].
+Thus, a lingering question is how the time scales of ecosystem response and community change differ among ecosystems, which can be answered by manipulating precipitation to reach extreme levels.
+The results of our five year experiment in a sagebrush steppe conform to the first of our four predictions: neither ecosystem functional response nor community composition changed with chronic alteration in water availability (Fig.
 1, top left).
 
-This is surprising because grasslands generally, and sagebrush steppe specifically, are considered water-limited systems.
-Thus, our naive expecatations were that either (i) ecosystem functional response would differ among treatments as species reach physiological thresholds and resource limitations (Fig.
-1, top right), or (ii) ecosystem functional response stays the same due to species reordering and compensation (Fig.
-1, bottom left).
+The similarity of ecosystem functional response (Fig.
+3A) and community composition (Fig.
+5) across treatments is surprising because grasslands generally, and sagebrush steppe specifically, are considered water-limited systems.
+Indeed, we expected ecosystem functional response, community composition, or both to change with precipitation treatment, landing us in any box of Fig. 1 \emph{except} the top left.
+So, why did our treatments fail to induce ecosystem or community responses?
+We can think of three reasons; two are limitations of our study, and one is the life history traits of the species in our focal communities.
+We first discuss the potential limitations of our study, and then discuss biological explanations.
 
-In the absence of community composition data, an obvious hypothesis to explain the consistent ecosystem functional response is that species reordered rapidly.
-In so doing, newly abundant species could compensate for the loss of ecosystem functioning of previously abundant species.
-We did not find this to be the case, as species composition was remarkably stable over the course of the experiment (Fig. 5).
-
-In combination, the lack of ecosystem and community response to drought or irrigation shows that this sagebrush steppe ecosystem is resistant to chronic alterations in water availability.
-There are three possible explanations for the resistance we found.
-First, it could be that our experiment simply was not long enough to induce responses.
-This may be true for community responses, which can take decades [@Wilcox2016].
-
-Second, it could be that our manipulations were not large enough to induce a response.
-That is, maybe a 50\% decrease/increase in any given year is not abnormal give our sites historical range of variability [@Knapp2017].
+First, it could be that our manipulations were not large enough to induce a response.
+That is, maybe a 50\% decrease/increase in any given year is not abnormal given our site's historical range of variability [@Knapp2017].
 We cannot definitively rule out this possibility, but we have reason to believe our manipulations *should* have been large enough.
@@
 The irrigation treatment may have been too small, however.
+
+Second,
 
@@
 In conclusion, our results suggest that five years of $\pm$ 50\% ambient precipitation is not enough to induce a shift in ecosystem functional response in a sagebrush steppe.
-This is likely because component species are tolerant to drought conditions and water remains co-limiting at higher precipitation levels.
-Longer time series of chronic precipitation alteration may reveal plant community shifts that we did not observe [e.g., @Wilcox2016].
+This is likely because the species in our focal plant community are bet hedgers, maintaining relatively consistent responses to interannual variation in precipitation to avoid booms and busts.
+Longer time series of chronic precipitation alteration may reveal plant community shifts that we did not observe [e.g., @Wilcox2016], in which case species that do not bet hedge may gain prominence and dominate the ecosystem functional response.
 Our results suggest compositional shifts would have the largest impact at high rainfall because the current community maintained consistent ecosystem functional response at very low water availability.
@@
 
-NSF DBI-1400370 to Andrew Tredennick.  
-NSF Graduate Research Fellowship to Andrew Kleinhesselink.  
-NSF DEB-1353078 and DEB-1054040 to Peter Adler.
+NSF DBI-1400370 to Andrew T. Tredennick.  
+NSF Graduate Research Fellowship to Andrew R. Kleinhesselink.  
+NSF DEB-1353078 and DEB-1054040 to Peter B. Adler.
 
@@
   \item Andrew R. Kleinhesselink conceived and designed the experiments, performed the experiments, collected data, reviewed drafts of the paper.
+  \item Bret Taylor...
   \item Peter B. Adler conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed

### 2017-07-26 manuscript/usses_water_ms.tex
rewriting discussion

@@
 California, Los Angeles, Los Angeles, California 90095}
-\affil[3]{USSES, address\ldots{}}
+\affil[3]{United States Department of Agriculture, Agriculture Research Station,
+U.S. Sheep Experiment Station, Dubois, Idaho 83423}
 
@@
 
-Projected changes in precipitation regimes, associated with global
-climate change, will affect
-
-Water availability is a strong driver of aboveground net primary
-productivity (ANPP) in grassland ecosystems. Mean annual precipitation
-(MAP) is strongly correlated with ANPP across sites \citep{Huxman2004},
-and, in water-limited ecosystems, interannual variation in MAP is also
-correlated with interannual variation in ANPP \citep{Hsu2012}. These
-relationships suggest that chronic alterations in levels of
-precipitation, associated with global climate change, will impact ANPP.
+Ecosystems worldwide are projected to experience either wetter or drier
+conditions in the future \citep{Zhang2007, Greve2014}. This has
+important implications for ecosystem functioning because mean annual net
+primary productivity (ANPP) positively covaries with mean annual
+precipitation across ecosystems \citep{Sala1988, Huxman2004} and
+interannual responses of ANPP are linked to interannual variability in
+precipitation within ecosystems
+\citep{Lauenroth1992, Hsu2012, Gherardi2015a}. To predict how ecosystems
+might respond to extreme precipitation levels, we need to understand
+whether ecosystems tend to follow temporal trends, which assume
+relatively constant plant community composition, or spatial trends,
+which allow community composition to shift \citep{Wilcox2016}.
 
@@
 conditions are likely to be outside the historical range of variability
-\citep{Smith2011}. For example, historical trends may show a linear soil
-moisture-ANPP relationship, but that relationship could potentially
-saturate if soil moisture is pushed far beyond typical levels.
-Saturating relationships are actually common
-\citep{Hsu2012, Gherardi2015a}, perhaps because other resources, like
-nitrogen, become more limiting in wet years than dry years. Knowing the
-curvature of the soil moisture-ANPP relationship is critical for
-understanding how ecosystems will respond to chronic alterations in
-water availability.
+\citep{Smith2011}. For example, historical trends may underestimate the
+potential for the soil moisture-ANPP relationship to saturate if soil
+moisture is pushed far beyond typical levels. Saturating relationships
+are actually common \citep{Hsu2012, Gherardi2015a}, perhaps because
+other resources, like nitrogen, become more limiting in wet years than
+dry years. Knowing the curvature of the soil moisture-ANPP relationship
+at extreme precipitation levels is critical for understanding how
+ecosystems will respond to chronic alterations in water availability.
 
@@
 
-Manipulating potentially limiting resources like precipitation offers a
-route to understanding how ecosystems will respond to resource levels
+Manipulating potentially limiting resources, like precipitation, offers
+a route to understanding how ecosystems will respond to resource levels
 that fall outside the historical range of variability
@@
 response to soil moisture, that is, the soil moisture-ANPP relationship
-(Figure 1). The four predictions are based on possible outcomes at the
+(Fig. 1). The four predictions are based on possible outcomes at the
 community (e.g., community composition) and ecosystem (e.g., soil
@@
 us {[}A. Kleinhesselink and P. Adler{]}) established 16 new 1 m\(^2\)
-plots located in one of the same exclosures as 6 of the original
-vegetation monitoring plots. We avoided areas on steep hill slopes,
-areas with greater than 20\% cover of bare rock, and areas with greater
-than 10\% cover of the shrubs \emph{Purshia tridentata} and/or
-\emph{Amelanchier utahensis}. We established the new plots in pairs and
-randomly assigned each plot in a pair to receive a ``drought'' or
-``irrigation'' treatment.
+plots located in the same exclosure as the six control plots. We avoided
+areas on steep hill slopes, areas with greater than 20\% cover of bare
+rock, and areas with greater than 10\% cover of the shrubs \emph{Purshia
+tridentata} and/or \emph{Amelanchier utahensis}. We established the new
+plots in pairs and randomly assigned each plot in a pair to receive a
+``drought'' or ``irrigation'' treatment.
 
@@
 fact that we did not directly monitor soil moisture in each plot on each
-day of the experiment.\\
-Only a subset of plots were equiped with soil moisture sensors, and
-within those plots, one or more of the sensors frequently failed to
-collect data. To remedy these problems, and to produce average daily
-soil moisture values for the ambient, drought, and irrigation
-conditions, we used a statistical model to predict the average treatment
-effects on soil moisture during the course of the experiment.
+day of the experiment. Only a subset of plots were equiped with soil
+moisture sensors, and within those plots, one or more of the sensors
+frequently failed to collect data. To remedy these problems, and to
+produce average daily soil moisture values for the ambient, drought, and
+irrigation conditions, we used a statistical model to predict the
+average treatment effects on soil moisture during the course of the
+experiment.
 
@@
 with independent variables for treatment, season (winter, spring,
-summer, fall), rainfall, and all two-way interactions.\\
-Rainy days were defined as any day in which precipitation was recorded
-and average temperature was above 3\(^{\circ}\)C. The day immediately
-following rainfall was also classified as rainy. We fit the model using
-the `lme4::lmer()' function \citep{Bates2015} in R \citep{R2016}, with
+summer, fall), rainfall, and all two-way interactions. Rainy days were
+defined as any day in which precipitation was recorded and average
+temperature was above 3\(^{\circ}\)C. The day immediately following
+rainfall was also classified as rainy. We fit the model using the
+`lme4::lmer()' function \citep{Bates2015} in R \citep{R2016}, with
 random effects for plot group and date. We weighted observations by the
@@
 soil moisture in the treated plots based on the average daily ambient
-soil moisture.\\
-We could only predict soil moisture in the treated plots on days for
-which we took at least one ambient soil moisture measurement (Fig. 2B).
+soil moisture. We could only predict soil moisture in the treated plots
+on days for which we took at least one ambient soil moisture measurement
+(Fig. 2B).
 
@@
 Cumulative March-June soil moisture had a weak positive effect on ANPP
-(mean of \(\beta_{1}\) = 0.41), but the effect is associated with high
-uncertainty, with credible intervals that overlap zerp (80\% BCI =
--0.08, 0.86; 95\% BCI = -0.45, 1.13) (Fig. 3). Ecosystem functional
+(mean of \(\beta_{1}\) = 0.41; Fig. 3), but the effect is associated
+with high uncertainty, with credible intervals that overlap zerp (80\%
+BCI = -0.08, 0.86; 95\% BCI = -0.45, 1.13). Ecosystem functional
 response was similar among treatments, with treatment level regressions
@@
 
-Ecosystem response to chronic resource alteration is expected to follow
-a temporal trend. Initially, ecosystem response will be modest and
-reflect the physiological responses of constituent species. Over longer
-time periods, species reordering will cause greater responses as species
-better able to take advantage of, or cope with, new resource levels
-become more abundant. This temporal trend is formalized by the
-`Hierarchical Response Framework' \citep[HRF,][]{Smith2009}, and has
-been empirically supported \citep{Knapp2012, Wilcox2016}. A lingering
-question is not whether the HRF is a reasonable model of ecosystem
-dynamics, but rather how the time scales of ecosystem response and
-community change differ among ecosystems. Indeed, previous work has
-shown that community compositional shifts can be both rapid, on the
-order of years \citep{Hoover2014}, and slow, on order of decades
-\citep{Knapp2012, Wilcox2016}. To add to this growing body of knowledge,
-we performed a precipitation manipulation experiment in a sagebrush
-steppe ecosystem.
+Ecosystem response to precipitation extremes depends on the
+physiological responses of constituent species and the rate at which
+community composition shifts to favor species better able to take
+advantage of, or cope with, new resource levels \citep{Smith2009}.
+Previous work has shown that community compositional shifts can be both
+rapid, on the order of years \citep{Hoover2014}, and slow, on order of
+decades \citep{Knapp2012, Wilcox2016}. Thus, a lingering question is how
+the time scales of ecosystem response and community change differ among
+ecosystems, which can be answered by manipulating precipitation to reach
+extreme levels. The results of our five year experiment in a sagebrush
+steppe conform to the first of our four predictions: neither ecosystem
+functional response nor community composition changed with chronic
+alteration in water availability (Fig. 1, top left).
 
-Ecosystem functional response (i.e., the relationship between soil
-moisture and ANPP) and community composition in this sagebrush steppe
-ecosystem was consistent across drought, control, and irrigation
-treatments (Fig. 2D, Fig. 3). In other words, functional responses
-estimated from any precipitation level sampled apply to any other level
-of precipitation, and the consistent functional response was not due to
-species reordering and compensation. Thus, our results conform to the
-first of our four predictions: neither ecosystem functional response nor
-community composition changed with chronic alteration in water
-availability (Fig. 1, top left).
+The similarity of ecosystem functional response (Fig. 3A) and community
+composition (Fig. 5) across treatments is surprising because grasslands
+generally, and sagebrush steppe specifically, are considered
+water-limited systems. Indeed, we expected ecosystem functional
+response, community composition, or both to change with precipitation
+treatment, landing us in any box of Fig. 1 \emph{except} the top left.
+So, why did our treatments fail to induce ecosystem or community
+responses? We can think of three reasons; two are limitations of our
+study, and one is the life history traits of the species in our focal
+communities. We first discuss the potential limitations of our study,
+and then discuss biological explanations.
 
-This is surprising because grasslands generally, and sagebrush steppe
-specifically, are considered water-limited systems. Thus, our naive
-expecatations were that either (i) ecosystem functional response would
-differ among treatments as species reach physiological thresholds and
-resource limitations (Fig. 1, top right), or (ii) ecosystem functional
-response stays the same due to species reordering and compensation (Fig.
-1, bottom left).
-
-In the absence of community composition data, an obvious hypothesis to
-explain the consistent ecosystem functional response is that species
-reordered rapidly. In so doing, newly abundant species could compensate
-for the loss of ecosystem functioning of previously abundant species. We
-did not find this to be the case, as species composition was remarkably
-stable over the course of the experiment (Fig. 5).
-
-In combination, the lack of ecosystem and community response to drought
-or irrigation shows that this sagebrush steppe ecosystem is resistant to
-chronic alterations in water availability. There are three possible
-explanations for the resistance we found. First, it could be that our
-experiment simply was not long enough to induce responses. This may be
-true for community responses, which can take decades \citep{Wilcox2016}.
-
-Second, it could be that our manipulations were not large enough to
+First, it could be that our manipulations were not large enough to
 induce a response. That is, maybe a 50\% decrease/increase in any given
-year is not abnormal give our sites historical range of variability
+year is not abnormal given our site's historical range of variability
 \citep{Knapp2017}. We cannot definitively rule out this possibility, but
@@
 have been too small, however.
+
+Second,
 
@@
 ambient precipitation is not enough to induce a shift in ecosystem
-functional response in a sagebrush steppe. This is likely because
-component species are tolerant to drought conditions and water remains
-co-limiting at higher precipitation levels. Longer time series of
-chronic precipitation alteration may reveal plant community shifts that
-we did not observe \citep[e.g.,][]{Wilcox2016}. Our results suggest
-compositional shifts would have the largest impact at high rainfall
-because the current community maintained consistent ecosystem functional
-response at very low water availability.
+functional response in a sagebrush steppe. This is likely because the
+species in our focal plant community are bet hedgers, maintaining
+relatively consistent responses to interannual variation in
+precipitation to avoid booms and busts. Longer time series of chronic
+precipitation alteration may reveal plant community shifts that we did
+not observe \citep[e.g.,][]{Wilcox2016}, in which case species that do
+not bet hedge may gain prominence and dominate the ecosystem functional
+response. Our results suggest compositional shifts would have the
+largest impact at high rainfall because the current community maintained
+consistent ecosystem functional response at very low water availability.
 
@@
 
-NSF DBI-1400370 to Andrew Tredennick.\\
-NSF Graduate Research Fellowship to Andrew Kleinhesselink.\\
-NSF DEB-1353078 and DEB-1054040 to Peter Adler.
+NSF DBI-1400370 to Andrew T. Tredennick.\\
+NSF Graduate Research Fellowship to Andrew R. Kleinhesselink.\\
+NSF DEB-1353078 and DEB-1054040 to Peter B. Adler.
 

### 2017-07-27 manuscript/usses_water_ms.Rmd
more discussion writing

@@
 
-<!---
-NOTES FROM MEETING WITH PETER
-1.
-Bet hedgers -- dominant species here deal well with low water and interannual variability; they live in dry and variable in environment, which has selected for tratis (not hitting physiological thresholds which only becom apparaent when species really go for it).
-Bet hedging = do the same things every year.
-2. Statistical explanation -- either (1) above explains our weak and variable relationship, or we just looked at the wrong thing; lots of other things beyond VWC that are influencing ANPP.
-3. Limitation -- snow, we don't know what happens. Its blowing around, who knows how big of an impact this has.
-4. Show Fig 2D with real model fits and move to results figure 3.
----->
-
 Ecosystem response to precipitation extremes depends on the physiological responses of constituent species and the rate at which community composition shifts to favor species better able to take advantage of, or cope with, new resource levels [@Smith2009].
@@
 We can think of three reasons; two are limitations of our study, and one is the life history traits of the species in our focal communities.
-We first discuss the potential limitations of our study, and then discuss biological explanations.
+We first discuss the potential limitations of our study, and then discuss a biological explanation.
 
@@
 
-Second,
+Second, ANPP at our site may be influenced by factors beyond the window of soil moisture we included in our statistical model.
+For example, temperature can also impact community composition and ANPP [@Hoover2014].
+Measurements of soil moisture likely contain a signal of temperature, through its impact on evaporation and infiltration, but the measurements will not capture the direct effect of temperature on metabolic and physiological processes.
+Likewise, we did not redistribute snow across our treatments in the winter, and snowmelt may spur early spring growth (CITATION).
 
-Third, the ecosystem and community resistance to drought and deluge we observed could be a real phenomenon in this system, at least over the five year time span we observed.
-The perennial plants in this cold desert ecosystem are tolerant to drought conditions [@Bazzaz1979; @Franks2011, A.R.
-Kleinhesselink, unpublished data], meaning a consistent ecosystem functional response at low soil moisture is perfectly reasonable.
+Third, the life history traits of the dominant species in our study ecosystem may explain the consistent and weak effect of soil moisture on ANPP (Fig. 3).
+Species that live in variable environments, such as cold deserts, must have strategies to ensure long-term success as conditions vary.
+One strategy is bet hedging, where species forego short-term gains to reduce the variance of long-term success (CITATION).
+The dry and variable environment of the sagebrush steppe has likely selected for bet hedging species that can maintain function at low water availability and have weak responses to high water availability.
+In so doing, the dominant species in our ecosystem avoid "boom and bust" cycles, which corresponds to the weak effect of soil moisture on ANPP (i.e., the credible intervals for $\beta_1$ overlapping zero), even at precipitation extremes.
+
+Another strategy to deal with variable environmental conditions is avoidance (CITATION), which would result in a consistent ecosystem functional response at low soil moisture.
+The perennial plants in this cold desert ecosystem are tolerant to drought conditions [@Bazzaz1979; @Franks2011, A.R. Kleinhesselink, unpublished data].
 For example, many of the perennial grasses in our focal ecosystem avoid drought stress by growing early in the growing season [@Blaisdell1958, A.R. Kleinhesselink, personal observation].
 Likewise, the dominant shrub in our focal ecosystem, *Artemisia tripartita*, has access to water deep in the soil profile thanks to a deep root system (CITATION).
-As for the irrigation treatments, the lack of a saturating response (e.g., a lower regression slope) suggests water availability remained a limiting factor at the levels of precipitation we were able to manipulate.
 
 In conclusion, our results suggest that five years of $\pm$ 50\% ambient precipitation is not enough to induce a shift in ecosystem functional response in a sagebrush steppe.
-This is likely because the species in our focal plant community are bet hedgers, maintaining relatively consistent responses to interannual variation in precipitation to avoid booms and busts.
+This is likely because the species in our focal plant community are tolerant of drought conditions and bet hedgers in wet conditions, maintaining relatively consistent responses to interannual variation in precipitation to avoid booms and busts.
 Longer time series of chronic precipitation alteration may reveal plant community shifts that we did not observe [e.g., @Wilcox2016], in which case species that do not bet hedge may gain prominence and dominate the ecosystem functional response.

### 2017-07-27 manuscript/usses_water_ms.tex
more discussion writing

@@
 communities. We first discuss the potential limitations of our study,
-and then discuss biological explanations.
+and then discuss a biological explanation.
 
@@
 
-Second,
+Second, ANPP at our site may be influenced by factors beyond the window
+of soil moisture we included in our statistical model. For example,
+temperature can also impact community composition and ANPP
+\citep{Hoover2014}. Measurements of soil moisture likely contain a
+signal of temperature, through its impact on evaporation and
+infiltration, but the measurements will not capture the direct effect of
+temperature on metabolic and physiological processes. Likewise, we did
+not redistribute snow across our treatments in the winter, and snowmelt
+may spur early spring growth (CITATION).
 
-Third, the ecosystem and community resistance to drought and deluge we
-observed could be a real phenomenon in this system, at least over the
-five year time span we observed. The perennial plants in this cold
-desert ecosystem are tolerant to drought conditions \citep[A.R.
-Kleinhesselink, unpublished data]{Bazzaz1979, Franks2011}, meaning a
-consistent ecosystem functional response at low soil moisture is
-perfectly reasonable. For example, many of the perennial grasses in our
-focal ecosystem avoid drought stress by growing early in the growing
-season \citep[A.R. Kleinhesselink, personal observation]{Blaisdell1958}.
-Likewise, the dominant shrub in our focal ecosystem, \emph{Artemisia
-tripartita}, has access to water deep in the soil profile thanks to a
-deep root system (CITATION). As for the irrigation treatments, the lack
-of a saturating response (e.g., a lower regression slope) suggests water
-availability remained a limiting factor at the levels of precipitation
-we were able to manipulate.
+Third, the life history traits of the dominant species in our study
+ecosystem may explain the consistent and weak effect of soil moisture on
+ANPP (Fig. 3). Species that live in variable environments, such as cold
+deserts, must have strategies to ensure long-term success as conditions
+vary. One strategy is bet hedging, where species forego short-term gains
+to reduce the variance of long-term success (CITATION). The dry and
+variable environment of the sagebrush steppe has likely selected for bet
+hedging species that can maintain function at low water availability and
+have weak responses to high water availability. In so doing, the
+dominant species in our ecosystem avoid ``boom and bust'' cycles, which
+corresponds to the weak effect of soil moisture on ANPP (i.e., the
+credible intervals for \(\beta_1\) overlapping zero), even at
+precipitation extremes.
+
+Another strategy to deal with variable environmental conditions is
+avoidance (CITATION), which would result in a consistent ecosystem
+functional response at low soil moisture. The perennial plants in this
+cold desert ecosystem are tolerant to drought conditions \citep[A.R.
+Kleinhesselink, unpublished data]{Bazzaz1979, Franks2011}. For example,
+many of the perennial grasses in our focal ecosystem avoid drought
+stress by growing early in the growing season \citep[A.R.
+Kleinhesselink, personal observation]{Blaisdell1958}. Likewise, the
+dominant shrub in our focal ecosystem, \emph{Artemisia tripartita}, has
+access to water deep in the soil profile thanks to a deep root system
+(CITATION).
 
@@
 functional response in a sagebrush steppe. This is likely because the
-species in our focal plant community are bet hedgers, maintaining
-relatively consistent responses to interannual variation in
-precipitation to avoid booms and busts. Longer time series of chronic
-precipitation alteration may reveal plant community shifts that we did
-not observe \citep[e.g.,][]{Wilcox2016}, in which case species that do
-not bet hedge may gain prominence and dominate the ecosystem functional
-response. Our results suggest compositional shifts would have the
-largest impact at high rainfall because the current community maintained
-consistent ecosystem functional response at very low water availability.
+species in our focal plant community are tolerant of drought conditions
+and bet hedgers in wet conditions, maintaining relatively consistent
+responses to interannual variation in precipitation to avoid booms and
+busts. Longer time series of chronic precipitation alteration may reveal
+plant community shifts that we did not observe
+\citep[e.g.,][]{Wilcox2016}, in which case species that do not bet hedge
+may gain prominence and dominate the ecosystem functional response. Our
+results suggest compositional shifts would have the largest impact at
+high rainfall because the current community maintained consistent
+ecosystem functional response at very low water availability.
 

### 2017-07-27 manuscript/usses_water_ms.Rmd
adds references

@@
 Second, ANPP at our site may be influenced by factors beyond the window of soil moisture we included in our statistical model.
-For example, temperature can also impact community composition and ANPP [@Hoover2014].
+For example, temperature can impact ANPP directly [@Epstein1997] and by exacerbating the effects of soil moisture [@DeBoeck2011].
 Measurements of soil moisture likely contain a signal of temperature, through its impact on evaporation and infiltration, but the measurements will not capture the direct effect of temperature on metabolic and physiological processes.
-Likewise, we did not redistribute snow across our treatments in the winter, and snowmelt may spur early spring growth (CITATION).
+Likewise, we did not redistribute snow across our treatments in the winter, and snowmelt may spur early spring growth.
+These statistical issues of missing potentially important covariates could explain the weak and uncertain relationship we observed between soil moisture and ANPP.
 
-Third, the life history traits of the dominant species in our study ecosystem may explain the consistent and weak effect of soil moisture on ANPP (Fig. 3).
+Third, the life history traits of the dominant species in our study ecosystem may explain the consistent but weak and uncertain effect of soil moisture on ANPP (Fig. 3).
 Species that live in variable environments, such as cold deserts, must have strategies to ensure long-term success as conditions vary.
-One strategy is bet hedging, where species forego short-term gains to reduce the variance of long-term success (CITATION).
+One strategy is bet hedging, where species forego short-term gains to reduce the variance of long-term success [@Seger1987].
+In other words, species do the same thing every year, with only minimal response to environmental conditions.
 The dry and variable environment of the sagebrush steppe has likely selected for bet hedging species that can maintain function at low water availability and have weak responses to high water availability.
@@
 
-Another strategy to deal with variable environmental conditions is avoidance (CITATION), which would result in a consistent ecosystem functional response at low soil moisture.
+Another strategy to deal with variable environmental conditions is avoidance, which would result in a consistent ecosystem functional response at low soil moisture.
 The perennial plants in this cold desert ecosystem are tolerant to drought conditions [@Bazzaz1979; @Franks2011, A.R. Kleinhesselink, unpublished data].
 For example, many of the perennial grasses in our focal ecosystem avoid drought stress by growing early in the growing season [@Blaisdell1958, A.R. Kleinhesselink, personal observation].
-Likewise, the dominant shrub in our focal ecosystem, *Artemisia tripartita*, has access to water deep in the soil profile thanks to a deep root system (CITATION).
+Likewise, the dominant shrub in our focal ecosystem, *Artemisia tripartita*, has access to water deep in the soil profile thanks to a deep root system [@Kulmatiski2017a].
 

### 2017-07-27 manuscript/usses_water_ms.tex
adds references

@@
 of soil moisture we included in our statistical model. For example,
-temperature can also impact community composition and ANPP
-\citep{Hoover2014}. Measurements of soil moisture likely contain a
-signal of temperature, through its impact on evaporation and
-infiltration, but the measurements will not capture the direct effect of
-temperature on metabolic and physiological processes. Likewise, we did
-not redistribute snow across our treatments in the winter, and snowmelt
-may spur early spring growth (CITATION).
+temperature can impact ANPP directly \citep{Epstein1997} and by
+exacerbating the effects of soil moisture \citep{DeBoeck2011}.
+Measurements of soil moisture likely contain a signal of temperature,
+through its impact on evaporation and infiltration, but the measurements
+will not capture the direct effect of temperature on metabolic and
+physiological processes. Likewise, we did not redistribute snow across
+our treatments in the winter, and snowmelt may spur early spring growth.
+These statistical issues of missing potentially important covariates
+could explain the weak and uncertain relationship we observed between
+soil moisture and ANPP.
 
 Third, the life history traits of the dominant species in our study
-ecosystem may explain the consistent and weak effect of soil moisture on
-ANPP (Fig. 3). Species that live in variable environments, such as cold
-deserts, must have strategies to ensure long-term success as conditions
-vary. One strategy is bet hedging, where species forego short-term gains
-to reduce the variance of long-term success (CITATION). The dry and
-variable environment of the sagebrush steppe has likely selected for bet
-hedging species that can maintain function at low water availability and
-have weak responses to high water availability. In so doing, the
-dominant species in our ecosystem avoid ``boom and bust'' cycles, which
-corresponds to the weak effect of soil moisture on ANPP (i.e., the
-credible intervals for \(\beta_1\) overlapping zero), even at
+ecosystem may explain the consistent but weak and uncertain effect of
+soil moisture on ANPP (Fig. 3). Species that live in variable
+environments, such as cold deserts, must have strategies to ensure
+long-term success as conditions vary. One strategy is bet hedging, where
+species forego short-term gains to reduce the variance of long-term
+success \citep{Seger1987}. In other words, species do the same thing
+every year, with only minimal response to environmental conditions. The
+dry and variable environment of the sagebrush steppe has likely selected
+for bet hedging species that can maintain function at low water
+availability and have weak responses to high water availability. In so
+doing, the dominant species in our ecosystem avoid ``boom and bust''
+cycles, which corresponds to the weak effect of soil moisture on ANPP
+(i.e., the credible intervals for \(\beta_1\) overlapping zero), even at
 precipitation extremes.
@@
 Another strategy to deal with variable environmental conditions is
-avoidance (CITATION), which would result in a consistent ecosystem
-functional response at low soil moisture. The perennial plants in this
-cold desert ecosystem are tolerant to drought conditions \citep[A.R.
-Kleinhesselink, unpublished data]{Bazzaz1979, Franks2011}. For example,
-many of the perennial grasses in our focal ecosystem avoid drought
-stress by growing early in the growing season \citep[A.R.
-Kleinhesselink, personal observation]{Blaisdell1958}. Likewise, the
-dominant shrub in our focal ecosystem, \emph{Artemisia tripartita}, has
-access to water deep in the soil profile thanks to a deep root system
-(CITATION).
+avoidance, which would result in a consistent ecosystem functional
+response at low soil moisture. The perennial plants in this cold desert
+ecosystem are tolerant to drought conditions \citep[A.R. Kleinhesselink,
+unpublished data]{Bazzaz1979, Franks2011}. For example, many of the
+perennial grasses in our focal ecosystem avoid drought stress by growing
+early in the growing season \citep[A.R. Kleinhesselink, personal
+observation]{Blaisdell1958}. Likewise, the dominant shrub in our focal
+ecosystem, \emph{Artemisia tripartita}, has access to water deep in the
+soil profile thanks to a deep root system \citep{Kulmatiski2017a}.
 

### 2017-07-27 manuscript/Appendix2.Rmd
code and paper updates

@@
 The random effects structure mirrors the nested structure of the data: plot-level parameters are nested within treatment-level parameters, which are nested within overall parameters.
-In what follows, $\textbf{X}$ is the design matrix including a column of 1s for intercepts and a column of precipitation values for each observation, $\textbf{y}$ is a row vector of log(ANPP) values for each observation, *i* references observations, *j* references plots, *k* references treatment, and *t* references year.
-The notation *i(j(k(t)))* reads as 'observation *i* associated with plot *j* associated with treatment *k* within year *t*'.
+In what follows, $\textbf{X}$ is the design matrix including a column of 1s for intercepts and a column of precipitation values for each observation, $\textbf{y}$ is a row vector of log(ANPP) values for each observation, *j* references plots, *k* references treatment, and *t* references year.
+The notation *j(k(t))* reads as 'observation of plot *j* associated with treatment *k* within year *t*'.
 Recall we have three treatments: drought (~50\% decrease in precipitation), control (ambient precipitation), and irrigation (~50\% increase in precipitation).
@@
 \begin{align}
-y_{i(j(k(t)))} \sim \text{Normal}\left(\mu_{i(j(k(t)))}, \sigma^2_k \right),
+y_{j(k(t))} \sim \text{Normal}\left(\mu_{j(k(t))}, \sigma^2_k \right),
 \end{align}
 
-\noindent{}where $\mu_{i(j(k(t)))}$ is the determinstic expectation from the regression model,
+\noindent{}where $\mu_{j(k(t))}$ is the determinstic expectation from the regression model,
 
@@
 \begin{align}
-\mu_{i(j(k(t)))} = \textbf{x}_i'\boldsymbol{\beta_{j(k)} + \gamma_{t}}.
+\mu_{j(k(t))} = \textbf{x}_{k(t)}'\boldsymbol{\beta_{j(k)} + \gamma_{t}}.
 \end{align}
@@
 Our Stan code is below.
-All code necessary to reproduce our results is in the supplementary code set (Appendix 3).
+All code necessary to reproduce our results has been archived on Figshare (*link here*) and released on GitHub (https://github.com/atredennick/usses_water/releases/v0.1).
+
+\newpage{}
 
@@
 
+\newpage{}
+
+# Figures
+
+```{r hrv, fig.cap="Scatterplot, on the arithmetic scale, of the data and model estimates shown a solid lines.
+Model estimates come from treatment level coefficients (colored lines) and the overall mean coefficients (heavy black line).", echo=FALSE, message=FALSE}
+library(tidyverse)
+library(dplyr)
+library(stringr)
+library(ggplot2)
+library(ggthemes)
+
+source("../code/read_format_data.R") # load data
+all_fit <- readRDS("../results/randcoefs_alltreatments_fit.RDS")
+
+lmod <- lm(log(anpp) ~ total_seasonal_vwc, anpp_data)
+x <- model.matrix(lmod)
+newx <- as.data.frame(unique(x)) %>% select(total_seasonal_vwc)
+newx$vwc_id <- c(1:nrow(newx))
+mean_log_anpp <- mean(log(anpp_data$anpp))
+sd_log_anpp <- sd(log(anpp_data$anpp))
+model_preds <- reshape2::melt(rstan::extract(all_fit, pars="ypreds_mu")) %>%
+  rename(iteration = iterations, vwc_id = Var2, estimate = value, stan_name = L1) %>%
+  left_join(newx, by="vwc_id") %>%
+  group_by(total_seasonal_vwc) %>%
+  summarise(mean_estimate = mean(estimate*sd_log_anpp+mean_log_anpp)) %>%
+  mutate(backtrans_estimate = mean_estimate) %>%
+  ungroup()
+
+trt_vwc <- data.frame(treatment = rep(c("Control","Drought","Irrigation"), each = nrow(newx)),
+                      treat_id = rep(c(1:3), each = nrow(newx)),
+                      vwc = rep(newx$total_seasonal_vwc, times = 3),
+                      vwc_id = rep(newx$vwc_id, times = 3))
+treat_preds <- reshape2::melt(rstan::extract(all_fit, pars="ypreds")) %>%
+  rename(iteration = iterations, treat_id = Var2, vwc_id = Var3, estimate = value, stan_name = L1) %>%
+  left_join(trt_vwc, by = c("vwc_id","treat_id")) %>%
+  group_by(treatment, vwc) %>%
+  summarise(mean_estimate = mean(estimate*sd_log_anpp+mean_log_anpp)) %>%
+  mutate(backtrans_estimate = mean_estimate) %>%
+  ungroup()
+
+ggplot(anpp_data, aes(x=total_seasonal_vwc,y=anpp))+
+  geom_point(shape=21,color="grey25",alpha=0.8,aes(fill=Treatment))+
+  geom_line(data=treat_preds, aes(x=vwc, y=exp(backtrans_estimate), color=treatment))+
+  geom_line(data=model_preds, aes(x=total_seasonal_vwc, y=exp(backtrans_estimate)), size=1)+
+  scale_fill_brewer(palette = "Set2")+
+  scale_color_brewer(palette = "Set2")+
+  #scale_x_continuous(limits=c(100,300),breaks=seq(100,300,50))+
+  #scale_y_continuous(breaks=seq(50,350,50))+
+  xlab("March-June Cumulative VWC")+
+  ylab(expression(paste("ANPP (g ",m^2,")")))+
+  guides(fill=FALSE,color=FALSE)+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank())
+
+```
+
 # References

### 2017-07-27 manuscript/usses_water_ms.Rmd
code and paper updates

@@
     address: United States Department of Agriculture, Agriculture Research Station, U.S. Sheep Experiment Station, Dubois, Idaho 83423
-abstract: >
-  Precipitation is predicted to become more variable in the western United States, meaning years of above and below average precipitation will become more common.
-Periods of drought and deluge are major drivers of ecosystem functioning in water limited grasslands, but how ecosystems respond to precipitation may change as the duration of above and below average periods increases.
+abstract: |
+  **Background.** Precipitation is predicted to become more variable in the western United States, meaning years of above and below average precipitation will become more common.
+Periods of extreme precipitation are major drivers of ecosystem functioning in water limited grasslands, but how ecosystems respond to precipitation may change as the duration of above and below average periods increases.
 Changes in ecosystem functional response could reflect compensatory changes in species composition or species reaching physiological thresholds at extreme precipitation levels.
-Here we report the results of a five-year experiment where we used drought and irrigation treatments (approximately 50% decrease/increase) to investigate whether ecosystem functional response remains consistent at extreme precipitation levels.
-We also use species-level adundance data to link ecosystem functional response to community dynamics.
-Ecosystem functional response, defined as the relationship between soil moisture and aboveground net primary productivity, was suprisingly consistent across drought, control, and irrigation treatments.
-Aboveground net primary productivity (ANPP) responded positively to availabile soil moisture, but the response was weak and did not vary across treatments.
+
+  **Methods.** We conducted a five year precipitation manipulation experiment in a sagebrush steppe ecosystem in Idaho, US.
+We used drought and irrigation treatments (approximately 50% decrease/increase) to investigate whether ecosystem functional response remains consistent at extreme precipitation levels.
+We recorded data on aboveground net primary productivity (ANPP), species abundance, and soil moisture.
+We fit a random intercept, random slope model to determine if the relationship between ANPP and soil moisture differed among treatments.
+We used nondimensional multivariate scaling to quantify community composition over the five years.
+
+  **Results.** Ecosystem functional response, defined as the relationship between soil moisture and ANPP, did not vary across drought, control, and irrigation treatments.
+ANPP responded positively to availabile soil moisture, but the response was weak and uncertain.
 There was also no evidence that treatment effects grew over time.
-The similarity of ecosystem functional response across treatments was not due to compensatory shifts at the plant community level, where species composition among treatments was similar and remarkably stable over the five years.
-At least in the short-term, ecosystem functional response and community composition in this sagebrush steppe system is consistent, even at extreme precipitation levels.
-Such consistency may be due to bet-hedging strategies of constituent species.
+Plant community composition was remarkably stable over the course of the experiment and did not differ among treatments.
+
+  **Discussion.** At least in the short-term, ecosystem functional response and community composition in this sagebrush steppe system is consistent, even at extreme precipitation levels.
+The similarity of ecosystem functional response across treatments was not due to compensatory shifts at the plant community level, but instead reflects the consistent responses of individual species.
+Such consistency may be due to bet-hedging strategies where species respond weakly to soil moisture to ensure low variance of long-term success.
 header-includes: >
@@
   rticles::peerj_article:
+    number_sections: false
     base_format: rmarkdown::pdf_document # bookdown::pdf_document2 # for using \@ref()
@@
 
+<!---
 Ecosystems worldwide are projected to experience either wetter or drier conditions in the future [@Zhang2007; @Greve2014].
@@
 To predict how ecosystems might respond to extreme precipitation levels, we need to understand whether ecosystems tend to follow temporal trends, which assume relatively constant plant community composition, or spatial trends, which allow community composition to shift [@Wilcox2016].
+--->
 
-At a given site, the functional response of ANPP to water availability (e.g., soil moisture) can be characterized by fitting a model to historical observations of ANPP and soil moisture.
+At a given site, the functional response of aboveground net primary productivity (ANPP) to water availability (e.g., soil moisture) can be characterized by fitting a model to historical observations of ANPP and soil moisture.
 However, the fitted functional response may provide an incomplete picture because future conditions are likely to be outside the historical range of variability [@Smith2011].
@@
 
-Another problem with relying on historical ecoystem functional responses is that they are not static.
-Changes in species identities and abundances can alter an ecosystem's functional response to water availability because different species have different physiological thresholds for producting biomass.
+Another problem with relying on historical ecosystem functional responses is that they are not static.
+Changes in species identities and abundances can alter an ecosystem's functional response to water availability because different species have different physiological thresholds for producing biomass.
 @Smith2009 introduced the 'Hierarchical Response Framework' (HRF) for understanding the interplay of community composition and ecosystem functioning in response to resource manipulations over time.
-In the near term, ecosystem functioning such as annual net primary productivity (ANPP) will reflect the physiological responses of individual species to the manipulated resource level.
+In the near term, ecosystem functioning such as ANPP will reflect the physiological responses of individual species to the manipulated resource level.
 For example, ANPP may decline under simulated drought because the initial community consisted of drought-intolerant species [@Hoover2014].
@@
 Third, the ecosystem functional response is consistent but underlying community composition changes (Fig. 1, bottom left).
-In this case, changes in species' indentities or abundances occur in response to altered precipitation levels and species more suited to the new conditions compensate for reduced function of initial species.
+In this case, changes in species' identities or abundances occur in response to altered precipitation levels and species more suited to the new conditions compensate for reduced function of initial species.
 Fourth, and last, both ecosystem functional response and community composition change (Fig. 1, bottom right).
@@
 We imposed drought and irrigation treatments (approximately $\pm50\%$) and measured ecosystem (ANPP) and community (species composition) responses.
-We focus on how the drought and irrigation treatments affect the relationship between available soil moisture and ANPP, and if community dynamics underly the ecosystem responses.
+We focus on how the drought and irrigation treatments affect the relationship between available soil moisture and ANPP, and if community dynamics underlie the ecosystem responses.
 In particular, we are interested in the consistency of the soil moisture-ANPP relationship among treatments.
@@
 Does the relationship saturate under the irrigation treatment, at high soil moisture?
-To answer these questions we fit a random intercepts, random slopes model to test whether the regressions differed among treatments.
+To answer these questions we fit a random intercept, random slope model to test whether the regressions differed among treatments.
 We also analyzed community composition over time, allowing us to place our experimental results within the framework of the HRF and our competing predictions (Fig. 1).
@@
 During the period of our experiment (2011 -- 2015), average mean annual precipitation was 265 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
+Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
+In 2007, we (well, one of us [P. Adler]) relocated 14 of the original quadrats, six of which were inside a large, permanent livestock exclosure.
+We use these six plots as control plots that have received no treatment, just ambient precipitation, in the experiment described below.
 
 ## Precipitation Experiment
-Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
-In 2007, we (well, one of us [P. Adler]) relocated 14 of the original quadrats, six of which were inside a large, permanent livestock exclosure.
-We use these six plots as control plots that have recieved no treatment, just ambient precipitation.
 In spring 2011, we (well, two of us [A. Kleinhesselink and P. Adler]) established 16 new 1 m$^2$ plots located in the same exclosure as the six control plots.
@@
 
-We monitored soil moisture in four of the drought-irrigation pairs using Decagon Devices (Pulman, Washington) 5TM and EC-5 soil moisture sensors.
+We monitored soil moisture in four of the drought-irrigation pairs using Decagon Devices (Pullman, Washington) 5TM and EC-5 soil moisture sensors.
 We installed four sensors around the edges of each plot, two at 5 cm soil depth and two at 25 cm soil depth.
@@
 Analyzing the response to experimental treatments was complicated by the fact that we did not directly monitor soil moisture in each plot on each day of the experiment.
-Only a subset of plots were equiped with soil moisture sensors, and within those plots, one or more of the sensors frequently failed to collect data.
+Only a subset of plots were equipped with soil moisture sensors, and within those plots, one or more of the sensors frequently failed to collect data.
 To remedy these problems, and to produce average daily soil moisture values for the ambient, drought, and irrigation conditions, we used a statistical model to predict the average treatment effects on soil moisture during the course of the experiment.
@@
 We then used the model to predict the average daily soil moisture in the treated plots based on the average daily ambient soil moisture.
-We could only predict soil moisture in the treated plots on days for which we took at least one ambient soil moisture measurement (Fig. 2B).
+We could only predict soil moisture in the treated plots on days for which we took at least one ambient soil moisture measurement.
 
@@
 Nonetheless, if we assume that rare species will respond more than common ones, then this approach is anti-conservative for detecting community change.
-This means that our approach is biased toward detecting compositional changes.\todo{Did I say this right?}
+This means that our approach is biased toward detecting compositional changes.
 
@@
 Based on our own observations and previous work at our study site [@Blaisdell1958; @Dalgleish2011; @Adler2012], we chose to use cumulative volumetric water content from March through June as our metric of soil moisture (hereafter referred to simply as `soil moisture').
-To achieve this goal, we fit a multi-level random intercept and random slope regression with log(ANPP) as the response variabile and soil moisture as the sole predictor.
+To achieve this goal, we fit a multi-level random intercept and random slope regression with log(ANPP) as the response variable and soil moisture as the sole predictor.
 We log-transformed ANPP to account for heteroscedasticity.
@@
 
-The intercept and slope terms are modeled hierarchichally to account for the non-independence of observations across years within plots and to allow us to test the hypothesis that our treatments alter the ANPP-soil moisture relationship.
+The intercept and slope terms are modeled hierarchically to account for the non-independence of observations across years within plots and to allow us to test the hypothesis that our treatments alter the ANPP-soil moisture relationship.
 Thus, our multi-level model has three grouping levels for coefficients, representing the nested structure of the data: (i) overall coefficients, (ii) treatment coefficients, and (iii) plot coefficients.
@@
 We also include a covariance structure among the intercept and slope at each level.
-Formally, our hierarhical structure is as follows, where we drop the intercept (0) and slope (1) subscripts and instead refer to a vector of coefficients, $\boldsymbol{\beta}$:
+Formally, our hierarchical structure is as follows, where we drop the intercept (0) and slope (1) subscripts and instead refer to a vector of coefficients, $\boldsymbol{\beta}$:
 
@@
 
-We fit the model using a Bayesian approach, obtaining posterior estimates of all unkowns via the No-U-Turn Hamiltonian Monte Carlo sampler in Stan [@stan2016].
+We fit the model using a Bayesian approach, obtaining posterior estimates of all unknowns via the No-U-Turn Hamiltonian Monte Carlo sampler in Stan [@stan2016].
 We used the R package 'rstan' [@rstan2016] to link R [@R2016] to Stan.
 We obtained samples from the posterior distribution for all model parameters from four parallel MCMC chains run for 10,000 iterations, saving every $10^{\text{th}}$ sample.
-Traceplots of all parameters were visually inspected to ensure well-mixed chains and convergence.
+Trace plots of all parameters were visually inspected to ensure well-mixed chains and convergence.
 We also made sure all scale reduction factors ($\hat{R}$) were less than 1.1.
@@
 Cumulative March-June soil moisture had a weak positive effect on ANPP (mean of $\beta_{1}$ = `r round(betas$mean[8],2)`; Fig.
-3), but the effect is associated with high uncertainty, with credible intervals that overlap zerp (80\% BCI = `r round(betas$X10.[8],2)`, `r round(betas$X90.[8],2)`; 95\% BCI = `r round(betas$X2.5.[8],2)`, `r round(betas$X97.5.[8],2)`).
+3), but the effect is associated with high uncertainty, with credible intervals that overlap zero (80\% BCI = `r round(betas$X10.[8],2)`, `r round(betas$X90.[8],2)`; 95\% BCI = `r round(betas$X2.5.[8],2)`, `r round(betas$X97.5.[8],2)`).
 Ecosystem functional response was similar among treatments, with treatment level regressions between soil moisture and log(ANPP) having similar intercepts and slopes (Fig. 3A).
@@
 The similarity of ecosystem functional response (Fig.
-3A) and community composition (Fig.
+3) and community composition (Fig.
 5) across treatments is surprising because grasslands generally, and sagebrush steppe specifically, are considered water-limited systems.
@@
 We can think of three reasons; two are limitations of our study, and one is the life history traits of the species in our focal communities.
-We first discuss the potential limitations of our study, and then discuss a biological explanation.
+We first discuss the potential limitations of our study, and then discuss the biological explanation.
 
@@
 We cannot definitively rule out this possibility, but we have reason to believe our manipulations *should* have been large enough.
-Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of mean growing season precipitation neccesary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site.
+Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of mean growing season precipitation necessary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site.
 The 1\% quantile of precipitation at our site is 110 mm, a 47% reduction from the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean growing season precipitation (Appendix 3).
@@
 Measurements of soil moisture likely contain a signal of temperature, through its impact on evaporation and infiltration, but the measurements will not capture the direct effect of temperature on metabolic and physiological processes.
-Likewise, we did not redistribute snow across our treatments in the winter, and snowmelt may spur early spring growth.
+Likewise, we did not redistribute snow across our treatments in the winter, and snow melt may spur early spring growth.
 These statistical issues of missing potentially important covariates could explain the weak and uncertain relationship we observed between soil moisture and ANPP.
@@
 
-We gratefully acknowledge the support of the Utah Agricultural Experiment Station (journal paper xxxx).
 We thank the many summer research technicians who collected the data reported in this paper and the US Experimental Sheep Station for facilitating work on their property.
@@
 
-NSF DBI-1400370 to Andrew T. Tredennick.  
-NSF Graduate Research Fellowship to Andrew R. Kleinhesselink.  
-NSF DEB-1353078 and DEB-1054040 to Peter B. Adler.
+This research was supported by the Utah Agricultural Experiment Station, Utah State University, and approved as journal paper number XXXX.
+The research was also supported by the National Science Foundation, through a Postdoctoral Research Fellowship in Biology and Mathematics to ATT (DBI-1400370), a Graduate Research Fellowship to ARK, and grants DEB-1353078 and DEB-1054040 to PBA.
 
@@
   \item Andrew R. Kleinhesselink conceived and designed the experiments, performed the experiments, collected data, reviewed drafts of the paper.
-  \item Bret Taylor...
+  \item Bret Taylor contributed reagents/materials/analysis tools, reviewed drafts of the paper.
   \item Peter B. Adler conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed
@@
 \end{itemize}
+
+# SUPPLEMENTAL INFORMATION
+
+\begin{description}
+\item [Appendix 1.] Additional methods and information on estimating aboveground net primary productivity.
+\item [Appendix 2.] Details of the hierarchical Bayesian model and Fig. A2-1.
+\item [Appendix 3.] Details on analysis of precipitation historical range of variability and Fig. A3-1.
+\end{description}
 
@@
   \centering
-      \includegraphics[height=7in]{../figures/Figure1.png}
+      \includegraphics[height=7in]{../figures/data_panels.png}
   \caption{(A) Probability density of historical precipitation from 1926-2016, with the years of the experiment shown with arrows on the \emph{x}-axis.
@@
 Note that we show log(ANPP) on the \emph{y}-axis of panel B; this same plot can be seen on the arithmetic scale in supporting material Fig.
-S2-1.}
+A2-1.}
 \end{figure}

### 2017-07-27 manuscript/usses_water_ms.tex
code and paper updates

@@
 \begin{abstract}
-Precipitation is predicted to become more variable in the western United
-States, meaning years of above and below average precipitation will
-become more common. Periods of drought and deluge are major drivers of
-ecosystem functioning in water limited grasslands, but how ecosystems
-respond to precipitation may change as the duration of above and below
-average periods increases. Changes in ecosystem functional response
-could reflect compensatory changes in species composition or species
-reaching physiological thresholds at extreme precipitation levels. Here
-we report the results of a five-year experiment where we used drought
+\textbf{Background.} Precipitation is predicted to become more variable
+in the western United States, meaning years of above and below average
+precipitation will become more common. Periods of extreme precipitation
+are major drivers of ecosystem functioning in water limited grasslands,
+but how ecosystems respond to precipitation may change as the duration
+of above and below average periods increases. Changes in ecosystem
+functional response could reflect compensatory changes in species
+composition or species reaching physiological thresholds at extreme
+precipitation levels.
+
+\textbf{Methods.} We conducted a five year precipitation manipulation
+experiment in a sagebrush steppe ecosystem in Idaho, US. We used drought
 and irrigation treatments (approximately 50\% decrease/increase) to
 investigate whether ecosystem functional response remains consistent at
-extreme precipitation levels. We also use species-level adundance data
-to link ecosystem functional response to community dynamics. Ecosystem
-functional response, defined as the relationship between soil moisture
-and aboveground net primary productivity, was suprisingly consistent
-across drought, control, and irrigation treatments. Aboveground net
-primary productivity (ANPP) responded positively to availabile soil
-moisture, but the response was weak and did not vary across treatments.
-There was also no evidence that treatment effects grew over time. The
-similarity of ecosystem functional response across treatments was not
-due to compensatory shifts at the plant community level, where species
-composition among treatments was similar and remarkably stable over the
-five years. At least in the short-term, ecosystem functional response
-and community composition in this sagebrush steppe system is consistent,
-even at extreme precipitation levels. Such consistency may be due to
-bet-hedging strategies of constituent species.
+extreme precipitation levels. We recorded data on aboveground net
+primary productivity (ANPP), species abundance, and soil moisture. We
+fit a random intercept, random slope model to determine if the
+relationship between ANPP and soil moisture differed among treatments.
+We used nondimensional multivariate scaling to quantify community
+composition over the five years.
+
+\textbf{Results.} Ecosystem functional response, defined as the
+relationship between soil moisture and ANPP, did not vary across
+drought, control, and irrigation treatments. ANPP responded positively
+to availabile soil moisture, but the response was weak and uncertain.
+There was also no evidence that treatment effects grew over time. Plant
+community composition was remarkably stable over the course of the
+experiment and did not differ among treatments.
+
+\textbf{Discussion.} At least in the short-term, ecosystem functional
+response and community composition in this sagebrush steppe system is
+consistent, even at extreme precipitation levels. The similarity of
+ecosystem functional response across treatments was not due to
+compensatory shifts at the plant community level, but instead reflects
+the consistent responses of individual species. Such consistency may be
+due to bet-hedging strategies where species respond weakly to soil
+moisture to ensure low variance of long-term success.
 % Dummy abstract text.
@@
 
-Ecosystems worldwide are projected to experience either wetter or drier
-conditions in the future \citep{Zhang2007, Greve2014}. This has
-important implications for ecosystem functioning because mean annual net
-primary productivity (ANPP) positively covaries with mean annual
-precipitation across ecosystems \citep{Sala1988, Huxman2004} and
-interannual responses of ANPP are linked to interannual variability in
-precipitation within ecosystems
-\citep{Lauenroth1992, Hsu2012, Gherardi2015a}. To predict how ecosystems
-might respond to extreme precipitation levels, we need to understand
-whether ecosystems tend to follow temporal trends, which assume
-relatively constant plant community composition, or spatial trends,
-which allow community composition to shift \citep{Wilcox2016}.
+At a given site, the functional response of aboveground net primary
+productivity (ANPP) to water availability (e.g., soil moisture) can be
+characterized by fitting a model to historical observations of ANPP and
+soil moisture. However, the fitted functional response may provide an
+incomplete picture because future conditions are likely to be outside
+the historical range of variability \citep{Smith2011}. For example,
+historical trends may underestimate the potential for the soil
+moisture-ANPP relationship to saturate if soil moisture is pushed far
+beyond typical levels. Saturating relationships are actually common
+\citep{Hsu2012, Gherardi2015a}, perhaps because other resources, like
+nitrogen, become more limiting in wet years than dry years. Knowing the
+curvature of the soil moisture-ANPP relationship at extreme
+precipitation levels is critical for understanding how ecosystems will
+respond to chronic alterations in water availability.
 
-At a given site, the functional response of ANPP to water availability
-(e.g., soil moisture) can be characterized by fitting a model to
-historical observations of ANPP and soil moisture. However, the fitted
-functional response may provide an incomplete picture because future
-conditions are likely to be outside the historical range of variability
-\citep{Smith2011}. For example, historical trends may underestimate the
-potential for the soil moisture-ANPP relationship to saturate if soil
-moisture is pushed far beyond typical levels. Saturating relationships
-are actually common \citep{Hsu2012, Gherardi2015a}, perhaps because
-other resources, like nitrogen, become more limiting in wet years than
-dry years. Knowing the curvature of the soil moisture-ANPP relationship
-at extreme precipitation levels is critical for understanding how
-ecosystems will respond to chronic alterations in water availability.
-
-Another problem with relying on historical ecoystem functional responses
-is that they are not static. Changes in species identities and
+Another problem with relying on historical ecosystem functional
+responses is that they are not static. Changes in species identities and
 abundances can alter an ecosystem's functional response to water
 availability because different species have different physiological
-thresholds for producting biomass. \citet{Smith2009} introduced the
+thresholds for producing biomass. \citet{Smith2009} introduced the
 `Hierarchical Response Framework' (HRF) for understanding the interplay
@@
 resource manipulations over time. In the near term, ecosystem
-functioning such as annual net primary productivity (ANPP) will reflect
-the physiological responses of individual species to the manipulated
-resource level. For example, ANPP may decline under simulated drought
-because the initial community consisted of drought-intolerant species
-\citep{Hoover2014}. Over longer time spans, ecosystem functioning may
-recover as new species colonize or initial species reorder in relative
-abundance. For example, ANPP may initially decline, but eventually rise
-back to pre-treatment levels once drought-tolerant species become more
-abundant and compensate for drought-intolerant species
-\citep{Hoover2014}. It is also possible that ecosystem functioning
-shifts to a new mean state, reflecting the suite of species in the new
-community \citep{Knapp2012}.
+functioning such as ANPP will reflect the physiological responses of
+individual species to the manipulated resource level. For example, ANPP
+may decline under simulated drought because the initial community
+consisted of drought-intolerant species \citep{Hoover2014}. Over longer
+time spans, ecosystem functioning may recover as new species colonize or
+initial species reorder in relative abundance. For example, ANPP may
+initially decline, but eventually rise back to pre-treatment levels once
+drought-tolerant species become more abundant and compensate for
+drought-intolerant species \citep{Hoover2014}. It is also possible that
+ecosystem functioning shifts to a new mean state, reflecting the suite
+of species in the new community \citep{Knapp2012}.
 
@@
 but underlying community composition changes (Fig. 1, bottom left). In
-this case, changes in species' indentities or abundances occur in
+this case, changes in species' identities or abundances occur in
 response to altered precipitation levels and species more suited to the
@@
 relationship between available soil moisture and ANPP, and if community
-dynamics underly the ecosystem responses. In particular, we are
+dynamics underlie the ecosystem responses. In particular, we are
 interested in the consistency of the soil moisture-ANPP relationship
@@
 the irrigation treatment, at high soil moisture? To answer these
-questions we fit a random intercepts, random slopes model to test
-whether the regressions differed among treatments. We also analyzed
-community composition over time, allowing us to place our experimental
-results within the framework of the HRF and our competing predictions
-(Fig. 1).
+questions we fit a random intercept, random slope model to test whether
+the regressions differed among treatments. We also analyzed community
+composition over time, allowing us to place our experimental results
+within the framework of the HRF and our competing predictions (Fig. 1).
 
@@
 year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
--5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July.
+-5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July. Between 1926
+and 1932, range scientists at the USSES established 26 permanent 1
+m\(^2\) quadrats to track vegetation change over time. In 2007, we
+(well, one of us {[}P. Adler{]}) relocated 14 of the original quadrats,
+six of which were inside a large, permanent livestock exclosure. We use
+these six plots as control plots that have received no treatment, just
+ambient precipitation, in the experiment described below.
 
@@
 
-Between 1926 and 1932, range scientists at the USSES established 26
-permanent 1 m\(^2\) quadrats to track vegetation change over time. In
-2007, we (well, one of us {[}P. Adler{]}) relocated 14 of the original
-quadrats, six of which were inside a large, permanent livestock
-exclosure. We use these six plots as control plots that have recieved no
-treatment, just ambient precipitation. In spring 2011, we (well, two of
-us {[}A. Kleinhesselink and P. Adler{]}) established 16 new 1 m\(^2\)
-plots located in the same exclosure as the six control plots. We avoided
-areas on steep hill slopes, areas with greater than 20\% cover of bare
-rock, and areas with greater than 10\% cover of the shrubs \emph{Purshia
-tridentata} and/or \emph{Amelanchier utahensis}. We established the new
-plots in pairs and randomly assigned each plot in a pair to receive a
-``drought'' or ``irrigation'' treatment.
+In spring 2011, we (well, two of us {[}A. Kleinhesselink and P.
+Adler{]}) established 16 new 1 m\(^2\) plots located in the same
+exclosure as the six control plots. We avoided areas on steep hill
+slopes, areas with greater than 20\% cover of bare rock, and areas with
+greater than 10\% cover of the shrubs \emph{Purshia tridentata} and/or
+\emph{Amelanchier utahensis}. We established the new plots in pairs and
+randomly assigned each plot in a pair to receive a ``drought'' or
+``irrigation'' treatment.
 
@@
 We monitored soil moisture in four of the drought-irrigation pairs using
-Decagon Devices (Pulman, Washington) 5TM and EC-5 soil moisture sensors.
-We installed four sensors around the edges of each plot, two at 5 cm
-soil depth and two at 25 cm soil depth. We also installed four sensors
-in areas nearby the four selected plot pairs to measure ambient soil
-moisture at the same depths. Soil moisture measurements were
+Decagon Devices (Pullman, Washington) 5TM and EC-5 soil moisture
+sensors. We installed four sensors around the edges of each plot, two at
+5 cm soil depth and two at 25 cm soil depth. We also installed four
+sensors in areas nearby the four selected plot pairs to measure ambient
+soil moisture at the same depths. Soil moisture measurements were
 automatically logged every four hours. We coupled this temporally
@@
 fact that we did not directly monitor soil moisture in each plot on each
-day of the experiment. Only a subset of plots were equiped with soil
+day of the experiment. Only a subset of plots were equipped with soil
 moisture sensors, and within those plots, one or more of the sensors
@@
 soil moisture. We could only predict soil moisture in the treated plots
-on days for which we took at least one ambient soil moisture measurement
-(Fig. 2B).
+on days for which we took at least one ambient soil moisture
+measurement.
 
@@
 community change. This means that our approach is biased toward
-detecting compositional changes.\todo{Did I say this right?}
+detecting compositional changes.
 
@@
 moisture'). To achieve this goal, we fit a multi-level random intercept
-and random slope regression with log(ANPP) as the response variabile and
+and random slope regression with log(ANPP) as the response variable and
 soil moisture as the sole predictor. We log-transformed ANPP to account
@@
 
-The intercept and slope terms are modeled hierarchichally to account for
+The intercept and slope terms are modeled hierarchically to account for
 the non-independence of observations across years within plots and to
@@
 coefficients. We also include a covariance structure among the intercept
-and slope at each level. Formally, our hierarhical structure is as
+and slope at each level. Formally, our hierarchical structure is as
 follows, where we drop the intercept (0) and slope (1) subscripts and
@@
 We fit the model using a Bayesian approach, obtaining posterior
-estimates of all unkowns via the No-U-Turn Hamiltonian Monte Carlo
+estimates of all unknowns via the No-U-Turn Hamiltonian Monte Carlo
 sampler in Stan \citep{stan2016}. We used the R package `rstan'
@@
 parallel MCMC chains run for 10,000 iterations, saving every
-\(10^{\text{th}}\) sample. Traceplots of all parameters were visually
+\(10^{\text{th}}\) sample. Trace plots of all parameters were visually
 inspected to ensure well-mixed chains and convergence. We also made sure
@@
 (mean of \(\beta_{1}\) = 0.41; Fig. 3), but the effect is associated
-with high uncertainty, with credible intervals that overlap zerp (80\%
+with high uncertainty, with credible intervals that overlap zero (80\%
 BCI = -0.08, 0.86; 95\% BCI = -0.45, 1.13). Ecosystem functional
@@
 
-The similarity of ecosystem functional response (Fig. 3A) and community
+The similarity of ecosystem functional response (Fig. 3) and community
 composition (Fig. 5) across treatments is surprising because grasslands
@@
 communities. We first discuss the potential limitations of our study,
-and then discuss a biological explanation.
+and then discuss the biological explanation.
 
@@
 calculated the percent reduction and increase of mean growing season
-precipitation neccesary to reach the 1\% and 99\% extremes of the
+precipitation necessary to reach the 1\% and 99\% extremes of the
 historical precipitation regime at our site. The 1\% quantile of
@@
 physiological processes. Likewise, we did not redistribute snow across
-our treatments in the winter, and snowmelt may spur early spring growth.
-These statistical issues of missing potentially important covariates
-could explain the weak and uncertain relationship we observed between
-soil moisture and ANPP.
+our treatments in the winter, and snow melt may spur early spring
+growth. These statistical issues of missing potentially important
+covariates could explain the weak and uncertain relationship we observed
+between soil moisture and ANPP.
 
@@
 
-We gratefully acknowledge the support of the Utah Agricultural
-Experiment Station (journal paper xxxx). We thank the many summer
-research technicians who collected the data reported in this paper and
-the US Experimental Sheep Station for facilitating work on their
-property. We also thank Susan Durham for clarifying our thinking on the
-statistical analyses and Kevin Wilcox for helpful discussions on
-analyzing community composition data.
+We thank the many summer research technicians who collected the data
+reported in this paper and the US Experimental Sheep Station for
+facilitating work on their property. We also thank Susan Durham for
+clarifying our thinking on the statistical analyses and Kevin Wilcox for
+helpful discussions on analyzing community composition data.
 
@@
 
-NSF DBI-1400370 to Andrew T. Tredennick.\\
-NSF Graduate Research Fellowship to Andrew R. Kleinhesselink.\\
-NSF DEB-1353078 and DEB-1054040 to Peter B. Adler.
+This research was supported by the Utah Agricultural Experiment Station,
+Utah State University, and approved as journal paper number XXXX. The
+research was also supported by the National Science Foundation, through
+a Postdoctoral Research Fellowship in Biology and Mathematics to ATT
+(DBI-1400370), a Graduate Research Fellowship to ARK, and grants
+DEB-1353078 and DEB-1054040 to PBA.
 
@@
 \end{itemize}
+
+\section{SUPPLEMENTAL INFORMATION}\label{supplemental-information}
+
+\begin{description}
+\item [Appendix 1.] Additional methods and information on estimating aboveground net primary productivity.
+\item [Appendix 2.] Details of the hierarchical Bayesian model and Fig. A2-1.
+\item [Appendix 3.] Details on analysis of precipitation historical range of variability and Fig. A3-1.
+\end{description}
 
@@
   \hline
-2011 & adonis &  21 &   2 & 1.02 & 0.45 \\ 
+2011 & adonis &  21 &   2 & 1.02 & 0.42 \\ 
   2011 & betadisper &  21 &   2 & 2.23 & 0.14 \\ 
-  2012 & adonis &  22 &   2 & 1.10 & 0.29 \\ 
+  2012 & adonis &  22 &   2 & 1.10 & 0.34 \\ 
   2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
@@
   2013 & betadisper &  22 &   2 & 0.28 & 0.76 \\ 
-  2014 & adonis &  22 &   2 & 0.95 & 0.58 \\ 
+  2014 & adonis &  22 &   2 & 0.95 & 0.54 \\ 
   2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
-  2015 & adonis &  21 &   2 & 1.05 & 0.37 \\ 
+  2015 & adonis &  21 &   2 & 1.05 & 0.40 \\ 
   2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 
-  2016 & adonis &  21 &   2 & 1.07 & 0.35 \\ 
+  2016 & adonis &  21 &   2 & 1.07 & 0.33 \\ 
   2016 & betadisper &  21 &   2 & 0.50 & 0.62 \\ 

### 2017-07-27 code/00_source_usses_water_scripts.R

@@ -34,19 +34,19 @@ for(pack in 1:length(packages)){
 }
 
 ####  2. FIT ANPP-NDVI REGRESSIONS ----
-source("calibrate_radiometer_by_year.R")
+source("01_calibrate_radiometer_by_year.R")
 
 ####  3. MAKE DATA FIGURE (Fig. 1) ----
-source("look_at_weather.R")
+source("02_plot_data.R")
 
 ####  4. FIT ANPP-PRECIP MODEL ----
-source("anpp_randcoefs_model.R")
+source("03_anpp_randcoefs_model.R")
 
 ####  5. PLOT MODEL RESULTS (Figs. 2 and 3) ----
-source("plot_model_results.R")
+source("04_plot_model_results.R")
 
 ####  6. RUN NMDS ANALYSIS AND PLOTS (Figs. 4 and 5, Table 1) ----
-source("community_dynamics.R")
+source("04_community_dynamics.R")
 
 
 

### 2017-07-27 code/02_plot_data.R

@@ -46,51 +46,7 @@ ppt_histogram <- ggplot(weather, aes(x=ppt1))+
   xlab("Growing Season Precipitation (mm)")+
   theme_bw()+
   theme(panel.grid.minor = element_blank())+
-  ggtitle("A")
-
-####
-####  READ IN ANPP DATA AND PLOT TREND ----
-####
-permanent_quad_biomass <- readRDS("../data/estimated_biomass/permanent_plots_estimated_biomass.RDS")
-
-
-
-####
-####  INITIAL PLOTS ----
-####
-permanent_quad_biomass <- permanent_quad_biomass %>% filter(Treatment %in% c("Control","Drought","Irrigation"))
-biomass_year_treatment <- permanent_quad_biomass %>%
-  group_by(Treatment,year) %>%
-  summarise(mean_biomass = mean(biomass_grams_est))
-
-
-biomass_yr_trt_summ <- permanent_quad_biomass %>%
-  filter(!str_detect(quadname, 'P1|P7')) %>%
-  group_by(Treatment,year) %>%
-  summarise(mean_biomass = mean(biomass_grams_est),
-            sd_biomass = sd(biomass_grams_est)) %>%
-  filter(year > 2011)
-
-stats_lab <- expression(paste("year*treatment  ", italic("P"), " = 0.67"))
-anpp_means <- ggplot(biomass_yr_trt_summ, aes(x=year, y=mean_biomass, color=Treatment))+
-  geom_line()+
-  geom_errorbar(aes(ymin=mean_biomass-sd_biomass, ymax=mean_biomass+sd_biomass), width=0.05)+
-  geom_point(color="white", size=3)+
-  geom_point()+
-  geom_point(color="grey35", shape=1)+
-  scale_color_brewer(palette = "Set2", name="Treatment")+
-  scale_x_continuous(breaks=c(2011:2016))+
-  scale_y_continuous(breaks=seq(50,350,50))+
-  ylab(expression(paste("ANPP (g ", m^-2,")")))+
-  xlab("Year")+
-  ggtitle("C")+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())+
-  guides(color = guide_legend(override.aes = list(size=1)))+
-  theme(legend.position = c(0.2, 0.75),legend.key.size = unit(1,"pt"),legend.title = element_text(size=10),
-        legend.text = element_text(size = 8),
[…]

### 2017-07-27 code/04_plot_model_results.R

@@ -131,23 +131,23 @@ ggsave("../figures/glmm_main_results.png",plot = gout, width = 5.5, height = 6, 
 
 
 
-####
-####  DATA AND REGRESSION PLOT ON ARITHMETIC SCALE FOR SI ----
-####
-ggplot(anpp_data, aes(x=total_seasonal_vwc,y=anpp))+
-  geom_point(shape=21,color="grey25",alpha=0.8,aes(fill=Treatment))+
-  geom_line(data=treat_preds, aes(x=vwc, y=exp(backtrans_estimate), color=treatment))+
-  geom_line(data=model_preds, aes(x=total_seasonal_vwc, y=exp(backtrans_estimate)), size=1)+
-  scale_fill_brewer(palette = "Set2")+
-  scale_color_brewer(palette = "Set2")+
-  #scale_x_continuous(limits=c(100,300),breaks=seq(100,300,50))+
-  #scale_y_continuous(breaks=seq(50,350,50))+
-  xlab("March-June Cumulative VWC")+
-  ylab(expression(paste("ANPP (g ",m^2,")")))+
-  guides(fill=FALSE,color=FALSE)+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())
-ggsave("../figures/glmm_main_results_arithmetic.png", width = 4, height = 3, units = "in", dpi =120)
+# ####
+# ####  DATA AND REGRESSION PLOT ON ARITHMETIC SCALE FOR SI ----
+# ####
+# ggplot(anpp_data, aes(x=total_seasonal_vwc,y=anpp))+
+#   geom_point(shape=21,color="grey25",alpha=0.8,aes(fill=Treatment))+
+#   geom_line(data=treat_preds, aes(x=vwc, y=exp(backtrans_estimate), color=treatment))+
+#   geom_line(data=model_preds, aes(x=total_seasonal_vwc, y=exp(backtrans_estimate)), size=1)+
+#   scale_fill_brewer(palette = "Set2")+
+#   scale_color_brewer(palette = "Set2")+
+#   #scale_x_continuous(limits=c(100,300),breaks=seq(100,300,50))+
+#   #scale_y_continuous(breaks=seq(50,350,50))+
+#   xlab("March-June Cumulative VWC")+
+#   ylab(expression(paste("ANPP (g ",m^2,")")))+
+#   guides(fill=FALSE,color=FALSE)+
+#   theme_bw()+
[…]

### 2017-07-27 code/05_community_dynamics.R

@@ -129,7 +129,7 @@ full_community_matrix_scaled <- full_community_matrix_scaled %>%
 
 nmds_df <- {}
 out_stats <- {}
-pdf("../figures/bray_raw.pdf",onefile = TRUE)
+# pdf("../figures/bray_raw.pdf",onefile = TRUE)
 for(doyr in unique(full_community_matrix_scaled$year)){
   tmp <- filter(full_community_matrix_scaled, year == doyr) %>% select(-year)
  invisible( capture.output( # already checked for convergence
@@ -139,7 +139,7 @@ for(doyr in unique(full_community_matrix_scaled$year)){
   nmds_points$Treatment <- tmp$Treatment
   nmds_points$Year <- doyr
   nmds_df <- rbind(nmds_df, nmds_points)
-  plot(tmp_bray, main=doyr)
+  # plot(tmp_bray, main=doyr)
   adonis_stats <- as.data.frame(vegan::adonis(tmp[3:ncol(tmp)]+2~tmp$Treatment)$aov.tab)[1,]
   betadisper_stats <- anova(vegan::betadisper(vegan::vegdist(tmp[3:ncol(tmp)]+2),tmp$Treatment))[1,]
   tmp_out <- rbind(adonis_stats[,-which(colnames(adonis_stats)=="R2")],
@@ -150,7 +150,7 @@ for(doyr in unique(full_community_matrix_scaled$year)){
                                            n = nrow(tmp),
                                            tmp_out))
 }
-dev.off()
+# dev.off()
 saveRDS(object = out_stats, file = "../results/sppcomp_stats.RDS")
 
 ggplot(nmds_df, aes(x=MDS1, y=MDS2, color=Treatment))+
@@ -173,64 +173,64 @@ ggsave(paste0(figure_path,"sppcomp_bray_all.png"), width=6, height = 4, units = 
 ####
 ####  RANK CLOCKS ON SCALED COMMUNITY DATA ----
 ####
-scaled_comp_ts <- full_community_matrix_scaled %>%
-  gather(key = species, value = scaled_abundance, -quad, -year, -Treatment) %>%
-  group_by(year, Treatment, species) %>%
-  summarise(mean_scaled_abundance = mean(scaled_abundance))
-
-## Create the graph
-g1 <- ggplot(scaled_comp_ts, aes(year, mean_scaled_abundance, color = species)) + 
[…]

### 2017-08-07 code/03_anpp_randcoefs_model.R
minor updates to andys model

@@ -1,7 +1,7 @@
 ##  anpp_randcoefs_model.R: Script to run GLMM analysis to test for treatment 
 ##  effects on the relationship between precipitation and ANPP.
 ##
-##  NOTE: Stan will issue a couple warnings after running the MCMC that, as
+##  NOTE: Stan may issue a couple warnings after running the MCMC that, as
 ##  the messages state, can be safely ignored. Just rejects a couple proposals
 ##  that result in ill-formed covariance matrices.
 ##
@@ -24,6 +24,7 @@ library(dplyr)        # Data summarizing
 library(ggthemes)     # Pleasing ggplot themes
 library(stringr)      # Working with strings
 library(rstan)        # For MCMC
+library(lme4)         # Mixed-effects modeling
 
 
 
@@ -32,87 +33,64 @@ library(rstan)        # For MCMC
 ####
 source("read_format_data.R")
 
-library(lme4)
-anpp_data$x <- anpp_data$vwc_scaled
-anpp_data$y <- as.numeric(scale(log(anpp_data$anpp)))
-
-m1 <- lmer(data = anpp_data, y ~ Treatment*x + (1|year) + (x|quadname) )
-model.matrix(m1)
-
-summary(m1)
 
 
 ####
 ####  SET UP AND FIT MODEL IN STAN ----
 ####
-fit_stan_model <- function(model_data, model_file, check_diags=FALSE, treattype){
-  lmod <- lm(log(anpp) ~ Treatment*vwc_scaled, model_data)
-  x <- model.matrix(lmod)
-  newx <- unique(x)
-  
-  lmod <- lm(log(anpp) ~ vwc_scaled, model_data)
-  z <- model.matrix(lmod)
-  
-  anppdat <- list(Nobs = nrow(model_data),
-                  Npreds = ncol(x),
-                  Npreds2 = ncol(z),
-                  Nplots = length(unique(model_data$quadname)),
-                  Ntreats = length(unique(model_data$Treatment)),
-                  Nppts = nrow(newx),
-                  Nyears = length(unique(model_data$year)),
-                  y = as.numeric(scale(log(model_data$anpp))),
-                  x = x,
-                  z = z,
-                  newx = newx,
[…]

### 2017-08-07 manuscript/usses_water_ms.Rmd

@@
 \begin{itemize}
-  \item Andrew T. Tredennick collected data, analyzed the data, wrote the paper, prepared figures and/or tables, reviewed
-drafts of the paper.
-  \item Andrew R. Kleinhesselink conceived and designed the experiments, performed the experiments, collected data, reviewed drafts of the paper.
+  \item Andrew T. Tredennick collected data, analyzed the data, wrote the paper, prepared figures and/or tables, reviewed drafts of the paper.
+  \item Andrew R. Kleinhesselink conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed drafts of the paper.
   \item Bret Taylor contributed reagents/materials/analysis tools, reviewed drafts of the paper.
-  \item Peter B. Adler conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed
-drafts of the paper.
+  \item Peter B. Adler conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed drafts of the paper.
 \end{itemize}
@@
 library(xtable)
+coef_probs <- readRDS("../results/betadiff_probabilities.RDS")
+coef_caption <- "Probability that the difference between treatment-level coefficients is greater or less than zero for each of the precipitation treatments.
+Probabilities denote max(Pr(difference) $>$ 0, Pr(difference) $<$ 0), where the sign of the coefficient was ignored.
+Thus, probabilities represent the probability of the compared coefficients being different from one another.
+Pr(difference) values far from 0 or 1 indicate no support for the difference being greater than zero.
+Bold coefficients denote Pr(difference) $\\geq$ 0.95 (there are none)."
+print(xtable(coef_probs, caption = coef_caption, digits = 2),
+      caption.placement="top",
+      include.rownames = F,
+      sanitize.colnames.function = identity,
+      comment = FALSE,
+      size = "normalsize")
+
 # Read in species composition stats
@@
        size = "normalsize")
-
-# coef_probs <- readRDS("../results/coef_probabilities.RDS")
-#
-# coef_caption <- "Probability that each log(ANPP) coefficient is greater or less than zero for each of the precipitation treatments.
-# Probabilities denote max(Pr(coefficient) > 0, Pr(coefficient) < 0), where the sign of the coefficient was ignored.
-Thus, probabilities represent the probability of the coefficient being important in the model.
-Bold coefficients denote Pr(coefficient) $\\geq$ 0.95."
-# print(xtable(coef_probs, caption = coef_caption, digits = 2),
-#       caption.placement="top",
-#       include.rownames = F,
-#       sanitize.colnames.function = identity,
-#       comment = FALSE,
-#       size = "normalsize")
 ```

### 2017-08-07 manuscript/usses_water_ms.tex

@@
 drafts of the paper.
-  \item Andrew R. Kleinhesselink conceived and designed the experiments, performed the experiments, collected data, reviewed drafts of the paper.
+  \item Andrew R. Kleinhesselink conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed drafts of the paper.
   \item Bret Taylor contributed reagents/materials/analysis tools, reviewed drafts of the paper.
@@
 \begin{table}[ht]
+\centering
+\caption{Probability that the difference between treatment-level coefficients is greater or less than zero for each of the precipitation treatments.
+Probabilities denote max(Pr(difference) $>$ 0, Pr(difference) $<$ 0), where the sign of the coefficient was ignored.
+Thus, probabilities represent the probability of the compared coefficients being different from one another.
+Pr(difference) values far from 0 or 1 indicate no support for the difference being greater than zero.
+Bold coefficients denote Pr(difference) $\geq$ 0.95 (there are none).}
+\begingroup\normalsize
+\begin{tabular}{lrr}
+  \hline
+Comparison & Intercept & Soil Moisture Effect \\ 
+  \hline
+Control - Drought & 0.70 & 0.72 \\ 
+  Control - Irrigation & 0.57 & 0.85 \\ 
+  Drought - Irrigation & 0.65 & 0.70 \\ 
+   \hline
+\end{tabular}
+\endgroup
+\end{table}\begin{table}[ht]
 \centering

### 2017-08-15 manuscript/Appendix2.Rmd
updates all model related stuff to ARK specification

--- 2017-06-09 manuscript/Appendix2.Rmd
@@
 \textbf{\Large{Appendix 2}} \\
-A.T. Tredennick, A.R. Kleinhesselink, \& P.B. Adler \\
-``Ecosystem and community resistance...'' \\
+A.T. Tredennick, A.R. Kleinhesselink, B. Taylor \& P.B. Adler \\
+``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe'' \\
 \emph{PeerJ}
@@
 \subsection{Model Description}
-Our hierarchical Bayesian model with random intercepts and random slopes allowed us to test for differences among treatments in the relationship between ANPP and growing season precipitation, and allowed us to account for the non-independence of observations through time within a plot.
-The random effects structure mirrors the nested structure of the data: plot-level parameters are nested within treatment-level parameters, which are nested within overall parameters.
-In what follows, $\textbf{X}$ is the design matrix including a column of 1s for intercepts and a column of precipitation values for each observation, $\textbf{y}$ is a row vector of log(ANPP) values for each observation, *i* references observations, *j* references plots, and *k* references treatment.
-The notation *i(j(k))* reads as 'observation *i* associated with plot *j* associated with treatment *k*'.
-Recall we have three treatments: drought (~50\% decrease in precipitation), control (ambient precipitation), and irrigation (~50\% increase in precipitation).
+Our hierarchical Bayesian model allows us to test for differences among treatments in the relationship between ANPP and soil moisture, and allowed us to account for the non-independence of observations through time within a plot.
+Treatment differences are modeled as fixed effects, which are modified by plot-level random effects.
+In what follows, $\textbf{X}$ is the fixed effects design matrix including:
 
+1. a column of 1s for intercepts
+2. a column of continuous values for the scaled volumetric water content for each observation
+3. a column of 0s or 1s indicating whether the observation is from a drought treatment 
+4. a column of 0s or 1s indicating whether the observation is from an irrigation treatment
+5. a column of continuous values for the interaction between scaled volumetric water content and the drought treatment indicator (column 2)
+6. a column of continuous values for the interaction between scaled volumetric water content and the irrigation treatment indicator (column 3)
+
+The first seven rows of the fixed effects design matrix (six control plots and one drought plot) is as follows:
+
+```{r fixed_design, echo=F, eval=T, message=F, warning=F}
+library(tidyverse)    # Data munging
+library(dplyr)        # Data summarizing
+library(ggthemes)     # Pleasing ggplot themes
+library(ggjoy)        # Joy plots!
+library(stringr)      # Working with strings
+library(rstan)        # For MCMC and Stan objects
+library(gridExtra)    # For combining ggplot objects
+library(viridis)      # Pleasing color palette
+library(RColorBrewer) # More color palettes
+library(cowplot)      # For combining ggplots
+source("../code/read_format_data.R") # load data
+lmod <- lm(log(anpp) ~ vwc_scaled*Treatment, anpp_data)
+x    <- model.matrix(lmod) 
+colnames(x) <- c("Int", "VWC", "Drought", "Irrig", "Drought:VWC", "Irrig:VWC")
+head(x, 7)
+```
+
+\noindent{}Note that within a year, all plots within a treatment share the same value of volumetric water content.
+This is because we could not monitor soil moisture in each plot, and instead used sparse observations to model average soil moisture in each treatment in each year (see main text).
+Using this design matrix, we can estimate six fixed effects ($\boldsymbol{\beta}$s):
+
+1. the intercept of the soil moisture-ANPP relationship for control plots
+2. the slope of the soil moisture-ANPP relationship for control plots
+3. the intercept offset for drought plots
+4. the intercept offset for irrigation plots
+5. the slope offset for drought plots
+6. the slope offset for irrigation plots
+
+\noindent{}We are particularly interested in the slope offsets, because these allow us to test whether the slopes for drought or irrigation are different from the control slope.
+If the slope offset is different from zero, this indicates the slopes are different.
+We assess whether the slope offsets are different from zero by calculating the probability that the posterior distribution is less or greater than zero (one-tailed tests).
+If the probability is greater than 0.95, then there is strong evidence that the slope offset is different from zero, and thus different from the control treatment slope.
+
+To account for the fact that observations within plots through time are not independent, we include random effects that modify the fixed effects in each plot.
+We model these random effects ($\boldsymbol{\gamma}$s) as offsets drawn from a multivariate normal distribution with mean 0 and a variance-covariance matrix ($\Sigma$) that includes a covariance between the intercept and slope offsets.
+We implement this random effects structure by including a random effects design matrix ($\textbf{Z}$) with a column of 1s for intercept offsets and a column of continuous values for the volumetric water content for each observation.
+
+Lastly, to account for unkown variation across years, we include random year effects. These year effects ($\boldsymbol{\eta}$s) act as offsets on the intercept.
+
+Putting it all together, our model is defined mathematically as follows, where *i* denotes observation, *j* denotes plot, and *t* denotes year.
 We assume the observations are conditionally Gaussian,
@@
 \begin{align}
-y_{i(j(k))} \sim \text{Normal}\left(\mu_{i(j(k))}, \sigma^2_k \right),
+\textbf{y} &\sim \text{Normal} \left(\boldsymbol{\mu}, \sigma^2 \right),
 \end{align}
 
-\noindent{}where $\mu_{i(j(k))}$ is the determinstic expectation from the regression model,
+\noindent{}where $\boldsymbol{\mu}$ is the determinstic expectation from the regression model,
 
@@
 \begin{align}
-\mu_{i(j(k))} = \textbf{x}_i'\boldsymbol{\beta_{j(k)}}.
+\mu_{i} &= \boldsymbol{\beta}\textbf{x}_i + \boldsymbol{\gamma}_{j(i)}\textbf{z}_i + \eta_t.
 \end{align}
 
-Regression parameters (the intercept and slope in the parameter vector $\boldsymbol{\beta}$) were modeled hierarchically,
+\noindent{}All fixed effect $\beta$s were drawn from normally distributed priors with mean 0 and standard deviation 1: $\boldsymbol{\beta} \sim \text{Normal}(0,1)$.
+$\gamma$ random effects were drawn from a multivariate prior centered on zero with a shared variance covariance matrix:
 
@@
 \begin{align}
-\boldsymbol{\beta}_{j(k)} &\sim \text{MVN} \left( \boldsymbol{\beta}_k, \boldsymbol{\Sigma}(k) \right), \\
-\boldsymbol{\beta}_{k} &\sim \text{MVN} \left( \boldsymbol{\beta}, \boldsymbol{\Sigma}  \right), \\
-\boldsymbol{\beta} &\sim \text{Normal} \left( 0, 1 \right),
+\boldsymbol{\gamma} \sim \text{MVN} \left( 0, \Sigma \right), \\
+\Sigma \sim \text{Cholesky...}
 \end{align}
 
-\noindent{} where $\boldsymbol{\beta}_{j(k)}$ is the vector of regression coefficients (intercept and slope) for plot *j* associated with treatment *k*, $\boldsymbol{\beta}_{k}$ is the vector of coefficients for each treatment, and $\boldsymbol{\beta}$ is the vector of overall coefficients.
-The plot- and treatment-level coefficients are drawn from multivariate normal distributions with covariance matrix $\boldsymbol{\Sigma}$.
-For the plot-level coefficients, each treatment has its own variance-covariance matrix (i.e., $\boldsymbol{\Sigma}(k)$).
-The overall coefficients are drawn from a normal prior with mean 0 and standard deviation 1.
+\noindent{The random year effects ($\boldsymbol{\eta}$) are modeled as intercept offsets centered on zero with a shared variance ($\sigma_{yr}$): $\boldsymbol{\gamma} \sim \text{Normal} \left( 0, \sigma_{yr}  \right)$.
 
@@
   int<lower=0> Npreds;         # number of covariates, including intercept
+  int<lower=0> Npreds2;        # number of random effect covariates
   int<lower=0> Nplots;         # number of plots
@@
   int<lower=0> Nobs;           # number of observations
-  int<lower=0> Nppts;         # Number of precip levels to predict
+  int<lower=0> Nppts;          # number of precip levels to predict
+  int<lower=0> Nyears;         # number of years
   vector[Nobs] y;              # vector of observations
-  row_vector[Npreds] x[Nobs]; # design matrix
+  row_vector[Npreds] x[Nobs];	 # design matrix for fixed effects
+  row_vector[Npreds2] z[Nobs]; # simple design matrix for random effects 
   matrix[Nppts,Npreds] newx;   # design matrix for predictions 
-  matrix[Npreds,Npreds] R;	  # priors for covariance matrix
+  matrix[Npreds2,Npreds2] R;	 # priors for covariance matrix
   int plot_id[Nobs];           # vector of plot ids
   int treat_id[Nobs];          # vector of treatment ids
+  int year_id[Nobs];           # vector of year ids
 }
@@
 parameters {
-  vector[Npreds] beta_plot[Nplots]; # a unique vector matrix for each plot
-  vector[Npreds] beta_treat[Ntreats]; # a unique vector matrix for each treatment
-	vector[Npreds] beta_mu; # overall coefficients
-	cov_matrix[Npreds] Sigma; # covariance matrix for treatment-level coefficients
-	cov_matrix[Npreds] Sigma_plot[Ntreats]; # unique covariance matrix for plot-level coefs by treatment
-	vector<lower=0>[Ntreats] sd_y; # treatment-level observation std. dev.
+	vector[Npreds] beta;	             # overall coefficients
+	vector[Nyears] year_off;           # vector of year effects
+	cholesky_factor_corr[Npreds2] L_u; # cholesky factor of plot random effect corr matrix
+  vector[Npreds2] beta_plot[Nplots]; # plot level random effects
+	vector<lower=0>[Npreds2] sigma_u;  # plot random effect std. dev.
+	real<lower=0> sd_y;                # treatment-level observation std. dev.
+	real<lower=0> sigma_year;          # year std. dev. hyperprior
 }
@@
 transformed parameters {
-  vector[Nobs] yhat; # vector of expected values (predictions)
+  vector[Nobs] yhat;               # vector of expected values
+  vector[Npreds2] u[Nplots];       # transformed plot random effects
+  matrix[Npreds2,Npreds2] Sigma_u; # plot ranef cov matrix
+
+  Sigma_u = diag_pre_multiply(sigma_u, L_u); # plot-level covariance matrix 
+  for(j in 1:Nplots)
+    u[j] = Sigma_u * beta_plot[j]; # plot random intercepts and slopes
+
+  # regression model for expected values (one for each plot-year)
   for (i in 1:Nobs)
-    # regression model for expected values (one for each plot-year)
-    yhat[i] = x[i]*beta_plot[plot_id[i]]; 
+    yhat[i] = x[i]*beta + z[i]*u[plot_id[i]] + year_off[year_id[i]]; 
 }
@@
   ####  PRIORS
+  sigma_year ~ cauchy(0,2)
+  year_off ~ normal(0,sigma_year); # priors on year effects, shared variance
+  beta ~ normal(0,1);	             # priors on treatment coefficients
+  L_u ~ lkj_corr_cholesky(1);      # prior on the cholesky factor which controls the 
+                                   # correlation between plot level treatment effects
+
   for(i in 1:Nplots)
-    # plot-level coefficients vary normally around treatment coefs
-		beta_plot[i] ~ multi_normal(beta_treat[treat_id[i]], Sigma_plot[treat_id[i]]);
-
-	for(i in 1:Ntreats){
-	  # treatment-level coefficients vary normally around overall coefficients
-	  beta_treat[i] ~ multi_normal(beta_mu, Sigma);
-
-	  # priors on covariance of effects at plot-level
-	  Sigma_plot[i] ~ inv_wishart(Npreds+1, R);
-	}
-
-	beta_mu ~ normal(0,1); # priors on overall effects
-	Sigma ~ inv_wishart(Npreds+1, R); # priors on covariance of effects at treatment-level
-	sd_y ~ weibull(2,1); # priors on observation std. dev. for each treatment
+		beta_plot[i] ~ normal(0,1); # plot-level coefficients for intercept and slope
 
@@
   for(i in 1:Nobs)
-	  # observations vary normally around expected values
-	  y[i] ~ normal(yhat[i], sd_y[treat_id[i]]);
+    y[i] ~ normal(yhat[i], sd_y); # observations vary normally around expected values
 }
+```
 
-generated quantities {
-  vector[Nppts] ypreds[Ntreats];
-  vector[Nppts] ydiff_control_drought;
-  vector[Nppts] ydiff_control_irrigate;
-  vector[2] inter_diffs;
-  for(i in 1:Ntreats)
-    ypreds[i] = newx*beta_treat[i]; # mean predictions for each treatment
+\newpage{}
 
-  # difference between mean predictions
-  ydiff_control_drought = ypreds[1] - ypreds[2]; 
-  ydiff_control_irrigate = ypreds[1] - ypreds[3];
+# Figures
 
-  # difference between control and treatment in avg ppt year
-  inter_diffs[1] = beta_treat[1][1] - beta_treat[2][1]; # control - drought
-  inter_diffs[2] = beta_treat[1][1] - beta_treat[3][1]; # control - irrigation
-}
+```{r scatterplot, fig.cap="Scatterplot, on the arithmetic scale, of the data and model estimates shown as solid lines.
+Model estimates come from treatment level coefficients (colored lines).", echo=FALSE, message=FALSE}
+library(tidyverse)    # Data munging
+library(dplyr)        # Data summarizing
+library(ggthemes)     # Pleasing ggplot themes
+library(ggjoy)        # Joy plots!
+library(stringr)      # Working with strings
+library(rstan)        # For MCMC and Stan objects
+library(gridExtra)    # For combining ggplot objects
+library(viridis)      # Pleasing color palette
+library(RColorBrewer) # More color palettes
+library(cowplot)      # For combining ggplots
+
+source("../code/read_format_data.R") # load data
+all_fit <- readRDS("../results/randcoefs_alltreatments_fit.RDS")
+
+param_labels <- c("Control::Intercept",
+                  "Drought::Intercept", 
+                  "Irrigation::Intercept",
+                  "Control::Slope", 
+                  "Drought::Slope", 
+                  "Irrigation::Slope")
+
+treat_effects <- data.frame(extract(all_fit, pars = 'beta')) %>% 
+  mutate(iteration = row_number()) %>% 
+  gather(param_name, estimate, starts_with('beta')) %>%
+  mutate(param_name = factor(param_name, labels = param_labels)) %>% 
+  spread(param_name, estimate) %>% 
+  mutate(`Drought::Intercept` = `Drought::Intercept` + `Control::Intercept`, 
+         `Irrigation::Intercept` = `Irrigation::Intercept` + `Control::Intercept`,  
+         `Drought::Slope` = `Drought::Slope` + `Control::Slope`, 
+         `Irrigation::Slope` = `Irrigation::Slope` + `Control::Slope`) %>%
+  gather(param_name, estimate, -iteration)
+
+mean_effects <- treat_effects %>%
+  group_by(param_name) %>%
+  summarise(mean_estimate = mean(estimate))
+
+regression <- mean_effects %>%
+  separate(param_name, into = c("Treatment","Coefficient"), sep = "::") %>%
+  spread(Coefficient, mean_estimate)
+
+lmod <- lm(log(anpp) ~ vwc_scaled, anpp_data)
+x    <- model.matrix(lmod) 
+newx <- data.frame(x = seq(min(x[,2]), max(x[,2]), length.out = 50)) %>%
+  mutate(id = paste("x",1:50,sep = "::")) %>%
+  spread(id, x)
+newx <- rbind(rbind(newx,newx),newx)
+newx$Treatment <- regression$Treatment
+
+regression <- regression %>%
+  left_join(newx, by="Treatment") %>%
+  gather(xid, newx, -Treatment, -Intercept, -Slope)
+regression$estimate <- with(regression, Intercept+Slope*newx)
+
+mean_log_anpp <- mean(log(anpp_data$anpp))
+sd_log_anpp <- sd(log(anpp_data$anpp))
+mean_vwc <- mean(anpp_data$total_seasonal_vwc)
+sd_vwc <- sd(anpp_data$total_seasonal_vwc)
+
+regression <- regression %>%
+  mutate(backtrans_estimate = estimate*sd_log_anpp+mean_log_anpp,
+         backtrans_vwc = newx*sd_vwc+mean_vwc)
+
+ggplot(anpp_data, aes(x=total_seasonal_vwc,y=anpp))+
+  geom_point(shape=21,color="grey25",alpha=0.8,aes(fill=Treatment))+
+  geom_line(data=regression, aes(x=backtrans_vwc, y=exp(backtrans_estimate), color=Treatment), size=1)+
+  scale_fill_brewer(palette = "Set2")+
+  scale_color_brewer(palette = "Set2")+
+  xlab("March-June Cumulative VWC")+
+  ylab(expression(paste("ANPP (g ",m^2,")")))+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank())
 
@@
 
+\newpage{}
+
+```{r year_effects, fig.cap="Posterior distributions of random year effects (intercept offsets).
+Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.", echo=FALSE, message=FALSE, warning=FALSE}
+library(tidyverse)    # Data munging
+library(dplyr)        # Data summarizing
+library(ggthemes)     # Pleasing ggplot themes
+library(ggjoy)        # Joy plots!
+library(stringr)      # Working with strings
+library(rstan)        # For MCMC and Stan objects
+library(gridExtra)    # For combining ggplot objects
+library(viridis)      # Pleasing color palette
+library(RColorBrewer) # More color palettes
+library(cowplot)      # For combining ggplots
+
+all_fit <- readRDS("../results/randcoefs_alltreatments_fit.RDS")
+
+year_id_names <- data.frame(year_id = c(1:5),
+                            year_name = c(1:5))
+betas <- reshape2::melt(rstan::extract(all_fit, pars="year_off")) %>%
+  rename(iteration = iterations, year_id = Var2, estimate = value, stan_name = L1) %>%
+  left_join(year_id_names, by="year_id")
+
+ggplot(betas, aes(x=estimate, y=as.factor(year_id), fill=as.factor(year_id), color=as.factor(year_id), height = ..density..))+
+  geom_vline(aes(xintercept=0), linetype=2, color="grey45")+
+  geom_joy(stat="density", adjust=3, alpha=0.9)+
+  scale_x_continuous(breaks=seq(-3,3,0.5), limits=c(-3,3))+
+  scale_y_discrete(labels = c("2012","2013","2014","2015","2016"))+
+  scale_fill_viridis(end=0.8,discrete=T)+
+  scale_color_viridis(end=0.8,discrete=T)+
+  guides(fill=FALSE, color=FALSE)+
+  xlab("Parameter Value")+
+  ylab("Year")+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank(),
+        panel.border     = element_blank(),
+        axis.line        = element_blank(),
+        axis.ticks       = element_blank())
+
+```
+
+\newpage{}
+
 # References

### 2017-08-15 manuscript/usses_water_ms.Rmd
updates all model related stuff to ARK specification

@@
 Based on our own observations and previous work at our study site [@Blaisdell1958; @Dalgleish2011; @Adler2012], we chose to use cumulative volumetric water content from March through June as our metric of soil moisture (hereafter referred to simply as `soil moisture').
-To achieve this goal, we fit a multi-level random intercept and random slope regression with log(ANPP) as the response variable and soil moisture as the sole predictor.
+To achieve this goal, we fit a hierarchical regression model with log(ANPP) as the response variable and soil moisture as the sole predictor.
 We log-transformed ANPP to account for heteroscedasticity.
-We fit the model under a Bayesian framework, allowing us to test for treatment differences by comparing the posterior distributions of the treatment-level coefficients [e.g., @Tredennick2013].
 Both log(ANPP) and soil moisture were standardized to have mean 0 and unit variance before fitting the model [i.e., $(x_i - \bar{x})/\sigma_x$].
@@
 
+<!---
 \begin{align}
@@
 \end{align}
+---->
 
-\noindent{}where $\mu_{j(k(t))}$ is the deterministic prediction from the regression model for plot *j* associated with treatment *k* in year *t*, $\beta_{0,j(k)}$ is the intercept for plot *j* associated with treatment *k*, $\beta_{1,j(k)}$ is the slope term for the effect of soil moisture for plot *j* associated with treatment *k*, $\gamma_t$ is the intercept offset for year *t*, and $\sigma^2_k$ is the process variance for treatment *k*.
+\begin{align}
+\mu_{i} &= \boldsymbol{\beta}\textbf{x}_i + \boldsymbol{\gamma}_{j(i)}\textbf{z}_i + \eta_t, \\
+\textbf{y} &\sim \text{Normal} \left(\boldsymbol{\mu}, \sigma^2 \right),
+\end{align}
+
+\noindent{}where $\mu_{i}$ is the deterministic prediction from the regression model for observation *i*, which is associated with plot *j* and treatment year *t*.
+$\boldsymbol{\beta}$ is the vector of coefficients for the fixed effects in the design matrix $\textbf{X}$.
+Each row of the design matrix represents a single observation ($\textbf{x}_i$) and is a vector with the following elements: 1 for the intercept, a binary 0 or 1 if the treatment is "drought", a binary 0 or 1 if the treatment is "irrigation", the scaled value of volumetric water content (VWC), binary "drought" value times VWC, and binary "irrigation" value times VWC.
+Thus, our model treats "control" observations as the main treatment and then estimates intercept and slope offsets for the "drought" and "irrigation" treatments.
+In reference to our model, the hypotheses we wish to test are: (1) the coefficient for drought$\times$VWC is positive and different from zero, and (2) the coefficient for irrigation$\times$VWC is negative and different from zero
+
+To account for the fact that observations within plots and years are not independent, we include three random effects.
+Specifically, we include plot-specific offsets ($\boldsymbol{\gamma}$) for the intercept and slope terms and year-specific intercept offsets ($\eta_t$).
+The covariate vector $\textbf{z}_i$ for each observation *i* has two elements: a 1 for the intercept and the scaled value of VWC for that plot and year.
+The plot-specific coefficients are modeled hierarchically, where plot level coefficients are drawn from a multivariate normal distribution with mean 0 and a variance-covariance strucutre that allows the intercept and slope terms to be correlated:
+
+\vspace{-1em}
+
+\begin{align}
+\boldsymbol{\gamma}_{j(i)} &\sim \text{MVN} \left( 0, \Sigma  \right),
+\end{align}
+
+\noindent{}where $\Sigma$ is the variance-covariance matrix and *j(i)* reads as "plot *j* associated with observation *i*".
+The random year effects ($\boldsymbol{\eta}$) are drawn from a normal prior with mean 0 and standard deviation $\sigma_{\text{year}}$, which was drawn from a half-cauchy distribution.
+A full description of our model is in Appendix 2.
+
+<!----
+is the intercept for plot *j* associated with treatment *k*, $\beta_{1,j(k)}$ is the slope term for the effect of soil moisture for plot *j* associated with treatment *k*, $\gamma_t$ is the intercept offset for year *t*, and $\sigma^2_k$ is the process variance for treatment *k*.
 Data include the standardized log(ANPP) observations $\left( y_{j(k(t))} \right)$ and soil moisture ($x_{k(t)}$).
@@
 A full description of model is in Appendix 2.
+---->
 
@@
 \item [Appendix 1.] Additional methods and information on estimating aboveground net primary productivity.
-\item [Appendix 2.] Details of the hierarchical Bayesian model and Fig. A2-1.
+\item [Appendix 2.] Details of the hierarchical Bayesian model, Fig. A2-1, and Fig. A2-2.
 \item [Appendix 3.] Details on analysis of precipitation historical range of variability and Fig. A3-1.
@@
 library(xtable)
-coef_probs <- readRDS("../results/betadiff_probabilities.RDS")
-coef_caption <- "Probability that the difference between treatment-level coefficients is greater or less than zero for each of the precipitation treatments.
-Probabilities denote max(Pr(difference) $>$ 0, Pr(difference) $<$ 0), where the sign of the coefficient was ignored.
-Thus, probabilities represent the probability of the compared coefficients being different from one another.
-Pr(difference) values far from 0 or 1 indicate no support for the difference being greater than zero.
-Bold coefficients denote Pr(difference) $\\geq$ 0.95 (there are none)."
-print(xtable(coef_probs, caption = coef_caption, digits = 2),
+effect_summary <- readRDS("../results/effect_summary.RDS")
+colnames(effect_summary) <- c("Coefficient", "Treatment", "Posterior Mean", "Posterior Median",
+                              "Lower 95\\% BCI", "Upper 95\\% BCI")
+coef_caption <- "Summary statistics from the posterior distributions of coefficients for each treatment."
+print(xtable(effect_summary, caption = coef_caption, digits = 2),
       caption.placement="top",
@@
 inyear_caption <- "Results from statistical tests for clustering and dispersion of community composition among precipitation treatments.
-`adonis` tests whether treatments form unique clusters in multidimensial space; `betadisper` tests whether treatments have similar dispersion.
+`adonis' tests whether treatments form unique clusters in multidimensial space; `betadisper' tests whether treatments have similar dispersion.
 For both tests, \\emph{P} values greater than 0.05 indicate there is no support that the treatments differ."
@@
   \caption{Results from the generalized linear mixed effects model.
-(A) Posterior distributions of treatment-level parameters ('Intercept' and the effect of 'Soil Moisture').
+(A) Posterior distributions of the intercept and slope offsets for the drought and irrigation treatments.
+Offsets indicate the amount to which the coefficients for drought or irrigation treatments differ from the control treatment estimates.
+Probabilities (``Pr $<$ 0 ='') for each distribution indicate the probability that coefficient is less than zero.
+Probabilities greater than 0.95 indicate strong support for the coefficient being less than zero.
+We only show the one-tailed probability for the value being less than zero because the median of each distribution is less than zero.
 Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.
 (B) Scatterplot of the data and model estimates shown a solid lines.
-Model estimates come from treatment level coefficients (colored lines) and the overall mean coefficients (heavy black line).
+Model estimates come from treatment level coefficients (colored lines).
 Note that we show log(ANPP) on the \emph{y}-axis of panel B; this same plot can be seen on the arithmetic scale in supporting material Fig.
@@
 \end{figure}
+--->
 
@@
 
-<!---
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=6in]{../figures/rank_clocks_scaled_abundance.png}
-  \caption{Rank clocks of average species' scaled abundance by treatment. (A) All species. (B) Dominant species.}
-\end{figure}
----->
-
 \newpage{}

### 2017-08-15 manuscript/usses_water_ms.tex
updates all model related stuff to ARK specification

--- 2017-07-27 manuscript/usses_water_ms.tex
@@
 metric of soil moisture (hereafter referred to simply as `soil
-moisture'). To achieve this goal, we fit a multi-level random intercept
-and random slope regression with log(ANPP) as the response variable and
-soil moisture as the sole predictor. We log-transformed ANPP to account
-for heteroscedasticity. We fit the model under a Bayesian framework,
-allowing us to test for treatment differences by comparing the posterior
-distributions of the treatment-level coefficients
-\citep[e.g.,][]{Tredennick2013}. Both log(ANPP) and soil moisture were
-standardized to have mean 0 and unit variance before fitting the model
-{[}i.e., \((x_i - \bar{x})/\sigma_x\){]}.
+moisture'). To achieve this goal, we fit a hierarchical regression model
+with log(ANPP) as the response variable and soil moisture as the sole
+predictor. We log-transformed ANPP to account for heteroscedasticity.
+Both log(ANPP) and soil moisture were standardized to have mean 0 and
+unit variance before fitting the model {[}i.e.,
+\((x_i - \bar{x})/\sigma_x\){]}.
 
@@
 \begin{align}
-\mu_{j(k(t))} &= \beta_{0,j(k)} + \beta_{1,j(k)}x_{k(t)} + \gamma_t, \\
-y_{j(k(t))} &\sim \text{Normal} \left(\mu_{j(k(t))}, \sigma^2_{k} \right),
+\mu_{i} &= \boldsymbol{\beta}\textbf{x}_i + \boldsymbol{\gamma}_{j(i)}\textbf{z}_i + \eta_t, \\
+\textbf{y} &\sim \text{Normal} \left(\boldsymbol{\mu}, \sigma^2 \right),
 \end{align}
 
-\noindent{}where \(\mu_{j(k(t))}\) is the deterministic prediction from
-the regression model for plot \emph{j} associated with treatment
-\emph{k} in year \emph{t}, \(\beta_{0,j(k)}\) is the intercept for plot
-\emph{j} associated with treatment \emph{k}, \(\beta_{1,j(k)}\) is the
-slope term for the effect of soil moisture for plot \emph{j} associated
-with treatment \emph{k}, \(\gamma_t\) is the intercept offset for year
-\emph{t}, and \(\sigma^2_k\) is the process variance for treatment
-\emph{k}. Data include the standardized log(ANPP) observations
-\(\left( y_{j(k(t))} \right)\) and soil moisture (\(x_{k(t)}\)).
+\noindent{}where \(\mu_{i}\) is the deterministic prediction from the
+regression model for observation \emph{i}, which is associated with plot
+\emph{j} and treatment year \emph{t}. \(\boldsymbol{\beta}\) is the
+vector of coefficients for the fixed effects in the design matrix
+\(\textbf{X}\). Each row of the design matrix represents a single
+observation (\(\textbf{x}_i\)) and is a vector with the following
+elements: 1 for the intercept, a binary 0 or 1 if the treatment is
+``drought'', a binary 0 or 1 if the treatment is ``irrigation'', the
+scaled value of volumetric water content (VWC), binary ``drought'' value
+times VWC, and binary ``irrigation'' value times VWC. Thus, our model
+treats ``control'' observations as the main treatment and then estimates
+intercept and slope offsets for the ``drought'' and ``irrigation''
+treatments. In reference to our model, the hypotheses we wish to test
+are: (1) the coefficient for drought\(\times\)VWC is positive and
+different from zero, and (2) the coefficient for irrigation\(\times\)VWC
+is negative and different from zero
 
-The intercept and slope terms are modeled hierarchically to account for
-the non-independence of observations across years within plots and to
-allow us to test the hypothesis that our treatments alter the ANPP-soil
-moisture relationship. Thus, our multi-level model has three grouping
-levels for coefficients, representing the nested structure of the data:
-(i) overall coefficients, (ii) treatment coefficients, and (iii) plot
-coefficients. Each subsequent level is drawn from the distribution of
-coefficients at the previous level, meaning plot-level coefficients are
-drawn from treatment-level coefficients, which are drawn from overall
-coefficients. We also include a covariance structure among the intercept
-and slope at each level. Formally, our hierarchical structure is as
-follows, where we drop the intercept (0) and slope (1) subscripts and
-instead refer to a vector of coefficients, \(\boldsymbol{\beta}\):
+To account for the fact that observations within plots and years are not
+independent, we include three random effects. Specifically, we include
+plot-specific offsets (\(\boldsymbol{\gamma}\)) for the intercept and
+slope terms and year-specific intercept offsets (\(\eta_t\)). The
+covariate vector \(\textbf{z}_i\) for each observation \emph{i} has two
+elements: a 1 for the intercept and the scaled value of VWC for that
+plot and year. The plot-specific coefficients are modeled
+hierarchically, where plot level coefficients are drawn from a
+multivariate normal distribution with mean 0 and a variance-covariance
+strucutre that allows the intercept and slope terms to be correlated:
 
@@
 \begin{align}
-\boldsymbol{\beta}_{j(k)} &\sim \text{MVN} \left( \boldsymbol{\beta}_k, \boldsymbol{\Sigma}(k) \right), \\
-\boldsymbol{\beta}_{k} &\sim \text{MVN} \left( \boldsymbol{\beta}, \boldsymbol{\Sigma}  \right), \\
-\boldsymbol{\beta} &\sim \text{Normal} \left( 0, 1 \right),
+\boldsymbol{\gamma}_{j(i)} &\sim \text{MVN} \left( 0, \Sigma  \right),
 \end{align}
 
-\noindent{}where \(\boldsymbol{\beta}_{j(k)}\) is the vector of
-regression coefficients (intercept and slope) for plot \emph{j}
-associated with treatment \emph{k}, \(\boldsymbol{\beta}_{k}\) is the
-vector of coefficients for each treatment, and \(\boldsymbol{\beta}\) is
-the vector of overall coefficients. The plot- and treatment-level
-coefficients are drawn from multivariate normal distributions with
-covariance matrix \(\boldsymbol{\Sigma}\). For the plot-level
-coefficients, each treatment has its own variance-covariance matrix
-(i.e., \(\boldsymbol{\Sigma}(k)\)). The overall coefficients are drawn
-from a normal prior with mean 0 and standard deviation 1. The random
-year effects (\(\boldsymbol{\gamma}\)) are drawn from a normal prior
-with mean 0 and standard deviation \(\sigma_{\text{year}}\), which was
-drawn from a weibull distribution. A full description of model is in
-Appendix 2.
+\noindent{}where \(\Sigma\) is the variance-covariance matrix and
+\emph{j(i)} reads as ``plot \emph{j} associated with observation
+\emph{i}''. The random year effects (\(\boldsymbol{\eta}\)) are drawn
+from a normal prior with mean 0 and standard deviation
+\(\sigma_{\text{year}}\), which was drawn from a half-cauchy
+distribution. A full description of our model is in Appendix 2.
 
@@
 \centering
+\caption{Summary statistics from the posterior distributions of coefficients for each treatment.} 
+\begingroup\normalsize
+\begin{tabular}{llrrrr}
+  \hline
+Coefficient & Treatment & Posterior Mean & Posterior Median & Lower 95\% BCI & Upper 95\% BCI \\ 
+  \hline
+Intercept & Control & 0.02 & 0.03 & -1.06 & 1.01 \\ 
+  Intercept & Drought & -0.04 & -0.04 & -1.39 & 1.21 \\ 
+  Intercept & Irrigation & -0.13 & -0.11 & -1.35 & 1.09 \\ 
+  Slope & Control & 0.63 & 0.64 & -0.33 & 1.61 \\ 
+  Slope & Drought & 0.46 & 0.47 & -0.59 & 1.52 \\ 
+  Slope & Irrigation & 0.26 & 0.26 & -0.60 & 1.12 \\ 
+   \hline
+\end{tabular}
+\endgroup
+\end{table}\begin{table}[ht]
+\centering
 \caption{Results from statistical tests for clustering and dispersion of community composition among precipitation treatments.
-`adonis` tests whether treatments form unique clusters in multidimensial space; `betadisper` tests whether treatments have similar dispersion.
+`adonis' tests whether treatments form unique clusters in multidimensial space; `betadisper' tests whether treatments have similar dispersion.
 For both tests, \emph{P} values greater than 0.05 indicate there is no support that the treatments differ.}
@@
   \caption{Results from the generalized linear mixed effects model.
-(A) Posterior distributions of treatment-level parameters ('Intercept' and the effect of 'Soil Moisture').
+(A) Posterior distributions of the intercept and slope offsets for the drought and irrigation treatments.
+Offsets indicate the amount to which the coefficients for drought or irrigation treatments differ from the control treatment estimates.
+Probabilities (``Pr $<$ 0 ='') for each distribution indicate the probability that coefficient is less than zero.
+Probabilities greater than 0.95 indicate strong support for the coefficient being less than zero.
+We only show the one-tailed probability for the value being less than zero because the median of each distribution is less than zero.
 Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.
 (B) Scatterplot of the data and model estimates shown a solid lines.
-Model estimates come from treatment level coefficients (colored lines) and the overall mean coefficients (heavy black line).
+Model estimates come from treatment level coefficients (colored lines).
 Note that we show log(ANPP) on the \emph{y}-axis of panel B; this same plot can be seen on the arithmetic scale in supporting material Fig.
 A2-1.}
-\end{figure}
-
-\newpage{}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=5in]{../figures/glmm_yeardiffs.png}
-  \caption{Posterior distributions of random year effects (intercept offsets). Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.}
 \end{figure}

### 2017-08-15 code/03_anpp_randcoefs_model.R

@@ -9,7 +9,7 @@
 ##  Date created: June 2, 2017
 
 ##  Clear everything
-rm(list=ls(all.names = T))
+rm(list = ls(all.names = TRUE))
 
 ##  Set working directory to source file location
 setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only for RStudio
@@ -65,13 +65,14 @@ anppdat <- list(
   R        = diag(1,ncol(z))
   ) # close list
 
-##  Fit the model in Stan
+##  Set Stan options
 rstan_options(auto_write = TRUE)
 options(mc.cores = parallel::detectCores())
 set.seed(123)
 
+##  Fit the model in Stan
 rt  <- stanc("anpp_randcoefs.stan")
-sm  <- stan_model(stanc_ret = rt, verbose=FALSE)
+sm  <- stan_model(stanc_ret = rt, verbose = FALSE)
 fit <- sampling(sm, data = anppdat, iter = 10000, chains = 4, thin = 10)
 
 ##  Save the model fit

### 2017-08-15 code/04_plot_model_results.R

@@ -22,6 +22,8 @@ library(stringr)      # Working with strings
 library(rstan)        # For MCMC and Stan objects
 library(gridExtra)    # For combining ggplot objects
 library(viridis)      # Pleasing color palette
+library(RColorBrewer) # More color palettes
+library(cowplot)      # For combining ggplots
 
 
 
@@ -30,12 +32,11 @@ library(viridis)      # Pleasing color palette
 ####
 source("read_format_data.R") # load data
 all_fit <- readRDS("../results/randcoefs_alltreatments_fit.RDS")
-# year_fit <- readRDS("../results/randyears_alltreatments_fit.RDS")
 
 
 
 ####
-####  CALCULATE MAX(Pr(effect) > 0, Pr(effect) < 0) FUNCTION ----
+####  FUNCTION TO CALCULATE MAX(Pr(effect) > 0, Pr(effect) < 0) ----
 ####
 get_one_tailed <- function(values){
   above <- 1 - ecdf(values)(0)
@@ -46,202 +47,375 @@ get_one_tailed <- function(values){
 
 
 ####
-####  CALCULATE PROBABILITIES OF TREATMENT DIFFERENCES ----
-####
-comp_id_names <- data.frame(comp_id    = c(1,2,3),
-                            Comparison = c("Control - Drought",
-                                           "Control - Irrigation",
-                                           "Drought - Irrigation"))
-
-inter_diffs <- reshape2::melt(rstan::extract(all_fit, pars="inter_diffs")) %>%
-  rename(iteration = iterations, comp_id = Var2, estimate = value, stan_name = L1) %>%
-  left_join(comp_id_names, by="comp_id")
-intercept_probs <- inter_diffs %>%
-  group_by(Comparison) %>%
-  summarise(prob = get_one_tailed(estimate)) %>%
-  mutate(beta = "Intercept")
-
-slope_diffs <- reshape2::melt(rstan::extract(all_fit, pars="vwc_diffs")) %>%
-  rename(iteration = iterations, comp_id = Var2, estimate = value, stan_name = L1) %>%
-  left_join(comp_id_names, by="comp_id")
-slope_probs <- slope_diffs %>%
-  group_by(Comparison) %>%
-  summarise(prob = get_one_tailed(estimate)) %>%
-  mutate(beta = "Soil Moisture Effect")
-
[…]

### 2017-08-15 manuscript/Appendix2.Rmd
adds bayes posterior expression

@@
 \boldsymbol{\gamma} \sim \text{MVN} \left( 0, \Sigma \right), \\
-\Sigma \sim \text{Cholesky...}
+\Sigma = \sigma_{plt}^2 \textbf{R}, \\
+\sigma_{plt} \sim \text{Uniform}(0,1), \\
+\textbf{R} \sim \text{Cholesky}(1)
 \end{align}
 
-\noindent{The random year effects ($\boldsymbol{\eta}$) are modeled as intercept offsets centered on zero with a shared variance ($\sigma_{yr}$): $\boldsymbol{\gamma} \sim \text{Normal} \left( 0, \sigma_{yr}  \right)$.
+\noindent{}The random year effects ($\boldsymbol{\eta}$) are modeled as intercept offsets centered on zero with a shared variance ($\sigma_{yr}$): $\boldsymbol{\gamma} \sim \text{Normal} \left( 0, \sigma_{yr}  \right)$.
+
+The Bayesian posterior distribution of our model can be expressed as:
+
+\vspace{-2em}
+
+\begin{align}
+\left[ \boldsymbol{\beta}, \boldsymbol{\gamma}, \boldsymbol{\eta}, \sigma_{yr}, \sigma_{plt}, \textbf{R}, \sigma \right | \textbf{y}] &\propto \left( \prod_{i=1}^{n} \left[ y_i | \boldsymbol{\beta}, \boldsymbol{\gamma}, \boldsymbol{\eta}, \sigma \right] \right) \left( \prod_{j=1}^{J} \left[ \gamma_j | \sigma_{plt}, \textbf{R} \right] \right) \\
+&\times \left( \prod_{t=1}^{T} \left[ \eta_t | \sigma_{yr} \right] \right)  [\boldsymbol{\beta}] [\sigma_{plt}] [\textbf{R}] [\sigma_{yr}] [\sigma]
+\end{align}
 

### 2017-08-15 manuscript/usses_water_ms.Rmd
updates results with new model calculations

@@
 
-Cumulative March-June soil moisture had a weak positive effect on ANPP (mean of $\beta_{1}$ = `r round(betas$mean[8],2)`; Fig.
-3), but the effect is associated with high uncertainty, with credible intervals that overlap zero (80\% BCI = `r round(betas$X10.[8],2)`, `r round(betas$X90.[8],2)`; 95\% BCI = `r round(betas$X2.5.[8],2)`, `r round(betas$X97.5.[8],2)`).
-Ecosystem functional response was similar among treatments, with treatment level regressions between soil moisture and log(ANPP) having similar intercepts and slopes (Fig. 3A).
-There was also no evidence that the treatment effects became more important over time because there was no directional trend in the random year effects (Fig. 4).
+Cumulative March-June soil moisture had a weak positive effect on ANPP (Table 1; Fig. 3), but the effect is associated with high uncertainty, with credible intervals that overlap zero (Table 1).
+Ecosystem functional response was similar among treatments (Table 1; Fig. 3), but there is evidence that the slope for the irrigation treatment is less than the slope for the control treatment.
+This evidence comes from interpreting the posterior distribution of the slope offset for the irrigation treatment, from which we calculate a 98% probability that the estimate is less than zero (Fig.
+3A, right panel).
+There was no evidence that the treatment effects became more important over time because there was no directional trend in the random year effects (Fig. A2-2).
 
 Community composition was similar among treatments.
-In no year did community composition among treatments not overlap, and they were equally dispersed in all years (Table 1; Fig. 5).
-Likewise, community composition was remarkably stable over time, with no evidence of divergence among treatments (Table 1; Fig. 5).
+In no year did community composition among treatments not overlap, and they were equally dispersed in all years (Table 2; Fig. 5).
+Likewise, community composition was remarkably stable over time, with no evidence of divergence among treatments (Table 2; Fig. 5).
 

### 2017-08-15 manuscript/usses_water_ms.tex
updates results with new model calculations

@@
 Cumulative March-June soil moisture had a weak positive effect on ANPP
-(mean of \(\beta_{1}\) = 0.41; Fig. 3), but the effect is associated
-with high uncertainty, with credible intervals that overlap zero (80\%
-BCI = -0.08, 0.86; 95\% BCI = -0.45, 1.13). Ecosystem functional
-response was similar among treatments, with treatment level regressions
-between soil moisture and log(ANPP) having similar intercepts and slopes
-(Fig. 3A). There was also no evidence that the treatment effects became
-more important over time because there was no directional trend in the
-random year effects (Fig. 4).
+(Table 1; Fig. 3), but the effect is associated with high uncertainty,
+with credible intervals that overlap zero (Table 1). Ecosystem
+functional response was similar among treatments (Table 1; Fig. 3), but
+there is evidence that the slope for the irrigation treatment is less
+than the slope for the control treatment. This evidence comes from
+interpreting the posterior distribution of the slope offset for the
+irrigation treatment, from which we calculate a 98\% probability that
+the estimate is less than zero (Fig. 3A, right panel). There was no
+evidence that the treatment effects became more important over time
+because there was no directional trend in the random year effects (Fig.
+A2-2).
 
@@
 community composition among treatments not overlap, and they were
-equally dispersed in all years (Table 1; Fig. 5). Likewise, community
+equally dispersed in all years (Table 2; Fig. 5). Likewise, community
 composition was remarkably stable over time, with no evidence of
-divergence among treatments (Table 1; Fig. 5).
+divergence among treatments (Table 2; Fig. 5).
 

### 2017-08-15 manuscript/usses_water_ms.Rmd
updates data plots with cowplot labels

@@
 Thus, a lingering question is how the time scales of ecosystem response and community change differ among ecosystems, which can be answered by manipulating precipitation to reach extreme levels.
+
+The results of our five year experiment in a sagebrush steppe conform to two of our four predictions, depending on treatment.
+Under chronic drought, neither ecosystem functional response nor community composition changed (Fig. 1, top left).
+But under irrigation, ecosystem functional response was different from the control treatment while community composition remained unchanged (Fig. 1, top right).
+Thus, we found some evidence for a saturating response at high water availability (Fig. 3), but the response of ANPP to soil moisture was weak overall (Table 1).
+<!---
 The results of our five year experiment in a sagebrush steppe conform to the first of our four predictions: neither ecosystem functional response nor community composition changed with chronic alteration in water availability (Fig.
 1, top left).
+--->
 
@@
 3) and community composition (Fig.
-5) across treatments is surprising because grasslands generally, and sagebrush steppe specifically, are considered water-limited systems.
-Indeed, we expected ecosystem functional response, community composition, or both to change with precipitation treatment, landing us in any box of Fig. 1 \emph{except} the top left.
-So, why did our treatments fail to induce ecosystem or community responses?
+5) between drought and control treatments is surprising because grasslands generally, and sagebrush steppe specifically, are considered water-limited systems.
+Indeed, we expected ecosystem functional response, community composition, or both to change under the drought treatment, landing us in any box of Fig. 1 \emph{except} the top left.
+So, why did our dourhgt treatment fail to induce ecosystem or community responses?
 We can think of three reasons; two are limitations of our study, and one is the life history traits of the species in our focal communities.
@@
 Thus, our drought treatment represented extreme precipitation amounts, especially in years where ambient precipitation was below average (Fig. 2A).
-The irrigation treatment may have been too small, however.
+The irrigation treatment may have been too small, yet that is the treatment where we observed an effect (Fig. 3).
 

### 2017-08-15 code/02_plot_data.R

@@ -21,6 +21,7 @@ library(stringr)
 library(ggthemes)
 library(ggalt)
 library(gridExtra)
+library(cowplot)
 
 
 
@@ -45,8 +46,7 @@ ppt_histogram <- ggplot(weather, aes(x=ppt1))+
   ylab("Density")+
   xlab("Growing Season Precipitation (mm)")+
   theme_bw()+
-  theme(panel.grid.minor = element_blank())+
-  ggtitle("A)")
+  theme(panel.grid.minor = element_blank())
 
 
 
@@ -68,7 +68,6 @@ soil_vwc <- ggplot(soil_moisture, aes(x=julian_date, y=VWC, group=Treatment, col
   scale_color_brewer(palette = "Set2", name="Treatment")+
   ylab(expression(paste("Mean Soil VWC (ml ", ml^-1,")")))+
   xlab("Julian Day")+
-  ggtitle("B)")+
   scale_y_continuous(breaks=seq(0,24,8))+
   facet_grid(year~.)+
   theme_bw()+
@@ -106,7 +105,6 @@ anpp_means <- ggplot(biomass_yr_trt_summ, aes(x=year, y=mean_biomass, color=Trea
   scale_y_continuous(breaks=seq(50,350,50))+
   ylab(expression(paste("ANPP (g ", m^-2,")")))+
   xlab("Year")+
-  ggtitle("C)")+
   theme_bw()+
   theme(panel.grid.minor = element_blank())+
   guides(color = guide_legend(override.aes = list(size=1)))+
@@ -120,8 +118,11 @@ anpp_means <- ggplot(biomass_yr_trt_summ, aes(x=year, y=mean_biomass, color=Trea
 ####
 ####  COMBINE PLOTS AND SAVE ----
 ####
-gout <- grid.arrange(ppt_histogram,soil_vwc,anpp_means,ncol=1,nrow=3)
-ggsave("../figures/data_panels.png", gout, width=3.2, height=8, units="in", dpi=120)
+# gout <- grid.arrange(ppt_histogram,soil_vwc,anpp_means,ncol=1,nrow=3)
+gout <- cowplot::plot_grid(ppt_histogram,soil_vwc,anpp_means,
+                           ncol=1, labels = c("A)","B)","C)"), hjust = -0.4)
+
+ggsave("../figures/data_panels.png", gout, width=3.3, height=8, units="in", dpi=120)
 
 
 

### 2017-08-15 code/00_source_usses_water_scripts.R
trivia

@@ -21,11 +21,14 @@ packages <- c(
   "dplyr",
   "ggthemes",
   "ggalt",
+  "ggjoy",
   "gridExtra",
   "stringr",
   "rstan",
   "vegan",
-  "viridis"
+  "viridis",
+  "RColorBrewer",
+  "cowplot"
 )
 
 missing_packages <- packages[ !packages %in% installed.packages() ] 
@@ -36,16 +39,16 @@ if (length(missing_packages) > 0 ){
 ####  2. FIT ANPP-NDVI REGRESSIONS ----
 source("01_calibrate_radiometer_by_year.R")
 
-####  3. MAKE DATA FIGURE (Fig. 1) ----
+####  3. MAKE DATA FIGURE (Fig. 2) ----
 source("02_plot_data.R")
 
 ####  4. FIT ANPP-PRECIP MODEL ----
 source("03_anpp_randcoefs_model.R")
 
-####  5. PLOT MODEL RESULTS (Figs. 2 and 3) ----
+####  5. PLOT MODEL RESULTS (Fig. 3, Table 1) ----
 source("04_plot_model_results.R")
 
-####  6. RUN NMDS ANALYSIS AND PLOTS (Figs. 4 and 5, Table 1) ----
+####  6. RUN NMDS ANALYSIS AND PLOTS (Fig. 4, Table 2) ----
 source("05_community_dynamics.R")
 
 

### 2017-08-21 manuscript/usses_water_ms.tex
updates regression figure with lines trimmed to data ranges

@@
 ecosystems, which can be answered by manipulating precipitation to reach
-extreme levels. The results of our five year experiment in a sagebrush
-steppe conform to the first of our four predictions: neither ecosystem
-functional response nor community composition changed with chronic
-alteration in water availability (Fig. 1, top left).
+extreme levels.
+
+The results of our five year experiment in a sagebrush steppe conform to
+two of our four predictions, depending on treatment. Under chronic
+drought, neither ecosystem functional response nor community composition
+changed (Fig. 1, top left). But under irrigation, ecosystem functional
+response was different from the control treatment while community
+composition remained unchanged (Fig. 1, top right). Thus, we found some
+evidence for a saturating response at high water availability (Fig. 3),
+but the response of ANPP to soil moisture was weak overall (Table 1).
 
 The similarity of ecosystem functional response (Fig. 3) and community
-composition (Fig. 5) across treatments is surprising because grasslands
-generally, and sagebrush steppe specifically, are considered
-water-limited systems. Indeed, we expected ecosystem functional
-response, community composition, or both to change with precipitation
-treatment, landing us in any box of Fig. 1 \emph{except} the top left.
-So, why did our treatments fail to induce ecosystem or community
-responses? We can think of three reasons; two are limitations of our
-study, and one is the life history traits of the species in our focal
-communities. We first discuss the potential limitations of our study,
-and then discuss the biological explanation.
+composition (Fig. 5) between drought and control treatments is
+surprising because grasslands generally, and sagebrush steppe
+specifically, are considered water-limited systems. Indeed, we expected
+ecosystem functional response, community composition, or both to change
+under the drought treatment, landing us in any box of Fig. 1
+\emph{except} the top left. So, why did our drought treatment fail to
+induce ecosystem or community responses? We can think of three reasons;
+two are limitations of our study, and one is the life history traits of
+the species in our focal communities. We first discuss the potential
+limitations of our study, and then discuss the biological explanation.
 
@@
 precipitation was below average (Fig. 2A). The irrigation treatment may
-have been too small, however.
+have been too small, yet that is the treatment where we observed an
+effect (Fig. 3).
 

### 2017-08-21 code/04_plot_model_results.R

@@ -4,7 +4,7 @@
 ##  Date created: May 25, 2017
 
 ##  Clear everything
-rm(list=ls(all.names = T))
+rm(list=ls(all.names = TRUE))
 
 ##  Set working directory to source file location
 setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only for RStudio
@@ -151,16 +151,26 @@ regression <- regression %>%
   mutate(backtrans_estimate = estimate*sd_log_anpp+mean_log_anpp,
          backtrans_vwc = newx*sd_vwc+mean_vwc)
 
-regress_plot <- ggplot(anpp_data, aes(x=total_seasonal_vwc,y=log(anpp)))+
-  geom_point(shape=21,color="grey25",alpha=0.8,aes(fill=Treatment))+
-  geom_line(data=regression, aes(x=backtrans_vwc, y=backtrans_estimate, color=Treatment), size=1)+
-  scale_fill_brewer(palette = "Set2")+
-  scale_color_brewer(palette = "Set2")+
-  xlab("March-June Cumulative VWC")+
-  ylab(expression(paste("log[ANPP (g ",m^2,")]")))+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())#+
-  #ggtitle("B)")
+regression_limited <- {}
+for(dotrt in unique(regression$Treatment)){
+  tmpdat <- filter(regression, Treatment == dotrt)
+  tmprange <- range(anpp_data[which(anpp_data$Treatment==dotrt),"total_seasonal_vwc"])
+  tmpdat$backtrans_vwc[tmpdat$backtrans_vwc < tmprange[1] | tmpdat$backtrans_vwc > tmprange[2]] <- NA
+  regression_limited <- rbind(regression_limited, tmpdat)
+}
+
+suppressWarnings( # ignore wanrnings about NA values
+  regress_plot <- ggplot(anpp_data, aes(x=total_seasonal_vwc,y=log(anpp)))+
+    geom_point(shape=21,color="grey25",alpha=0.8,aes(fill=Treatment))+
+    geom_line(data=regression_limited, aes(x=backtrans_vwc, y=backtrans_estimate, color=Treatment), size=1)+
+    scale_fill_brewer(palette = "Set2")+
+    scale_color_brewer(palette = "Set2")+
+    xlab("March-June Cumulative VWC")+
+    ylab(expression(paste("log[ANPP (g ",m^2,")]")))+
+    theme_bw()+
[…]

### 2017-08-21 manuscript/usses_water_ms.Rmd
trivia

@@
 
-Cumulative March-June soil moisture had a weak positive effect on ANPP (Table 1; Fig. 3), but the effect is associated with high uncertainty, with credible intervals that overlap zero (Table 1).
+Cumulative March-June soil moisture had a weak positive effect on ANPP (Table 1; Fig.
+3), but the effect is associated with high uncertainty, with credible intervals that overlap zero, regardless of treatment (Table 1).
 Ecosystem functional response was similar among treatments (Table 1; Fig. 3), but there is evidence that the slope for the irrigation treatment is less than the slope for the control treatment.
@@
 But under irrigation, ecosystem functional response was different from the control treatment while community composition remained unchanged (Fig. 1, top right).
-Thus, we found some evidence for a saturating response at high water availability (Fig. 3), but the response of ANPP to soil moisture was weak overall (Table 1).
+Thus, we found some evidence for a saturating response at high water availability (Fig. 3), but the response of ANPP to soil moisture was weak overall, regardless of treatment (Table 1).
 <!---
@@
 Indeed, we expected ecosystem functional response, community composition, or both to change under the drought treatment, landing us in any box of Fig. 1 \emph{except} the top left.
-So, why did our dourhgt treatment fail to induce ecosystem or community responses?
+So, why did our drought treatment fail to induce ecosystem or community responses?
 We can think of three reasons; two are limitations of our study, and one is the life history traits of the species in our focal communities.

### 2017-08-22 manuscript/usses_water_ms.Rmd
tiny edits to methods in ms

@@
 Our main goal was to test whether the relationship between ANPP and soil moisture differed among the drought, control, and irrigation treatments.
-Based on our own observations and previous work at our study site [@Blaisdell1958; @Dalgleish2011; @Adler2012], we chose to use cumulative volumetric water content from March through June as our metric of soil moisture (hereafter referred to simply as `soil moisture').
-To achieve this goal, we fit a hierarchical regression model with log(ANPP) as the response variable and soil moisture as the sole predictor.
+Based on our own observations and previous work at our study site [@Blaisdell1958; @Dalgleish2011; @Adler2012], we chose to use cumulative volumetric water content from March through June as our metric of soil moisture (hereafter referred to as `VWC').
+To achieve this goal, we fit a generalized linear mixed effects regression model with log(ANPP) as the response variable and VWC and treatment as fixed effects.
+Plot and year of treatment were included as random effects to account for non-independence of observations, as described below.
 We log-transformed ANPP to account for heteroscedasticity.
-Both log(ANPP) and soil moisture were standardized to have mean 0 and unit variance before fitting the model [i.e., $(x_i - \bar{x})/\sigma_x$].
+Both log(ANPP) and VWC were standardized to have mean 0 and unit variance before fitting the model [i.e., $(x_i - \bar{x})/\sigma_x$].
 
@@
 $\boldsymbol{\beta}$ is the vector of coefficients for the fixed effects in the design matrix $\textbf{X}$.
-Each row of the design matrix represents a single observation ($\textbf{x}_i$) and is a vector with the following elements: 1 for the intercept, a binary 0 or 1 if the treatment is "drought", a binary 0 or 1 if the treatment is "irrigation", the scaled value of volumetric water content (VWC), binary "drought" value times VWC, and binary "irrigation" value times VWC.
+Each row of the design matrix represents a single observation ($\textbf{x}_i$) and is a vector with the following elements: 1 for the intercept, a binary 0 or 1 if the treatment is "drought", a binary 0 or 1 if the treatment is "irrigation", the scaled value of VWC, binary "drought" value times VWC, and binary "irrigation" value times VWC.
 Thus, our model treats "control" observations as the main treatment and then estimates intercept and slope offsets for the "drought" and "irrigation" treatments.
-In reference to our model, the hypotheses we wish to test are: (1) the coefficient for drought$\times$VWC is positive and different from zero, and (2) the coefficient for irrigation$\times$VWC is negative and different from zero
+In reference to our model, the hypotheses we wish to test are:
 
-To account for the fact that observations within plots and years are not independent, we include three random effects.
+\begin{description}
+\item [H1] The coefficient for drought$\times$VWC is positive and different from zero.
+\item [H2] The coefficient for irrigation$\times$VWC is negative and different from zero.
+\end{description}
+
+To account for the fact that observations within plots and years are not independent, we include two random effects.
 Specifically, we include plot-specific offsets ($\boldsymbol{\gamma}$) for the intercept and slope terms and year-specific intercept offsets ($\eta_t$).
@@
 
+<!---
 ```{r get_results, echo=FALSE}
@@
 ```
+--->
 
@@
 3), but the effect is associated with high uncertainty, with credible intervals that overlap zero, regardless of treatment (Table 1).
-Ecosystem functional response was similar among treatments (Table 1; Fig. 3), but there is evidence that the slope for the irrigation treatment is less than the slope for the control treatment.
+Ecosystem functional response was similar among treatments (Table 1; Fig. 3B), but there is evidence that the slope for the irrigation treatment is less than the slope for the control treatment.
 This evidence comes from interpreting the posterior distribution of the slope offset for the irrigation treatment, from which we calculate a 98% probability that the estimate is less than zero (Fig.

### 2017-08-22 manuscript/usses_water_ms.tex
tiny edits to methods in ms

@@
 cumulative volumetric water content from March through June as our
-metric of soil moisture (hereafter referred to simply as `soil
-moisture'). To achieve this goal, we fit a hierarchical regression model
-with log(ANPP) as the response variable and soil moisture as the sole
-predictor. We log-transformed ANPP to account for heteroscedasticity.
-Both log(ANPP) and soil moisture were standardized to have mean 0 and
-unit variance before fitting the model {[}i.e.,
-\((x_i - \bar{x})/\sigma_x\){]}.
+metric of soil moisture (hereafter referred to as `VWC'). To achieve
+this goal, we fit a generalized linear mixed effects regression model
+with log(ANPP) as the response variable and VWC and treatment as fixed
+effects. Plot and year of treatment were included as random effects to
+account for non-independence of observations, as described below. We
+log-transformed ANPP to account for heteroscedasticity. Both log(ANPP)
+and VWC were standardized to have mean 0 and unit variance before
+fitting the model {[}i.e., \((x_i - \bar{x})/\sigma_x\){]}.
 
@@
 ``drought'', a binary 0 or 1 if the treatment is ``irrigation'', the
-scaled value of volumetric water content (VWC), binary ``drought'' value
-times VWC, and binary ``irrigation'' value times VWC. Thus, our model
-treats ``control'' observations as the main treatment and then estimates
-intercept and slope offsets for the ``drought'' and ``irrigation''
-treatments. In reference to our model, the hypotheses we wish to test
-are: (1) the coefficient for drought\(\times\)VWC is positive and
-different from zero, and (2) the coefficient for irrigation\(\times\)VWC
-is negative and different from zero
+scaled value of VWC, binary ``drought'' value times VWC, and binary
+``irrigation'' value times VWC. Thus, our model treats ``control''
+observations as the main treatment and then estimates intercept and
+slope offsets for the ``drought'' and ``irrigation'' treatments. In
+reference to our model, the hypotheses we wish to test are:
+
+\begin{description}
+\item [H1] The coefficient for drought$\times$VWC is positive and different from zero.
+\item [H2] The coefficient for irrigation$\times$VWC is negative and different from zero.
+\end{description}
 
 To account for the fact that observations within plots and years are not
-independent, we include three random effects. Specifically, we include
+independent, we include two random effects. Specifically, we include
 plot-specific offsets (\(\boldsymbol{\gamma}\)) for the intercept and
@@
 (Table 1; Fig. 3), but the effect is associated with high uncertainty,
-with credible intervals that overlap zero (Table 1). Ecosystem
-functional response was similar among treatments (Table 1; Fig. 3), but
-there is evidence that the slope for the irrigation treatment is less
-than the slope for the control treatment. This evidence comes from
-interpreting the posterior distribution of the slope offset for the
-irrigation treatment, from which we calculate a 98\% probability that
-the estimate is less than zero (Fig. 3A, right panel). There was no
-evidence that the treatment effects became more important over time
-because there was no directional trend in the random year effects (Fig.
-A2-2).
+with credible intervals that overlap zero, regardless of treatment
+(Table 1). Ecosystem functional response was similar among treatments
+(Table 1; Fig. 3B), but there is evidence that the slope for the
+irrigation treatment is less than the slope for the control treatment.
+This evidence comes from interpreting the posterior distribution of the
+slope offset for the irrigation treatment, from which we calculate a
+98\% probability that the estimate is less than zero (Fig. 3A, right
+panel). There was no evidence that the treatment effects became more
+important over time because there was no directional trend in the random
+year effects (Fig. A2-2).
 
@@
 evidence for a saturating response at high water availability (Fig. 3),
-but the response of ANPP to soil moisture was weak overall (Table 1).
+but the response of ANPP to soil moisture was weak overall, regardless
+of treatment (Table 1).
 

### 2017-08-22 manuscript/usses_water_ms.Rmd
updates abstract

@@
 We recorded data on aboveground net primary productivity (ANPP), species abundance, and soil moisture.
-We fit a random intercept, random slope model to determine if the relationship between ANPP and soil moisture differed among treatments.
+We fit a generalized linear mixed effects model to determine if the relationship between ANPP and soil moisture differed among treatments.
 We used nondimensional multivariate scaling to quantify community composition over the five years.
 
-  **Results.** Ecosystem functional response, defined as the relationship between soil moisture and ANPP, did not vary across drought, control, and irrigation treatments.
-ANPP responded positively to availabile soil moisture, but the response was weak and uncertain.
+  **Results.** Ecosystem functional response, defined as the relationship between soil moisture and ANPP was similar among drought and control treatments, but the irrigation treatment had a lower slope than the control treatment.
+However, ANPP response to availabile soil moisture was weak and uncertain regardless of treatment, with all slopes overlapping zero.
 There was also no evidence that treatment effects grew over time.
@@
 
-  **Discussion.** At least in the short-term, ecosystem functional response and community composition in this sagebrush steppe system is consistent, even at extreme precipitation levels.
-The similarity of ecosystem functional response across treatments was not due to compensatory shifts at the plant community level, but instead reflects the consistent responses of individual species.
-Such consistency may be due to bet-hedging strategies where species respond weakly to soil moisture to ensure low variance of long-term success.
+  **Discussion.** Despite some evidence that ecosystem functional response saturated under extreme wet conditions, the response of ANPP to soil moisture and community compositional changes in this sagebrush steppe system were consistently small.
+The similarity, or difference, of ecosystem functional response across treatments was not due to compensatory shifts at the plant community level, but instead reflects the weak responses of individual species.
+Such responses may be due to bet-hedging strategies where species respond weakly to even extreme levels of soil moisture to ensure low variance of long-term success.
 header-includes: >

### 2017-08-22 manuscript/usses_water_ms.tex
updates abstract

@@
 primary productivity (ANPP), species abundance, and soil moisture. We
-fit a random intercept, random slope model to determine if the
+fit a generalized linear mixed effects model to determine if the
 relationship between ANPP and soil moisture differed among treatments.
@@
 \textbf{Results.} Ecosystem functional response, defined as the
-relationship between soil moisture and ANPP, did not vary across
-drought, control, and irrigation treatments. ANPP responded positively
-to availabile soil moisture, but the response was weak and uncertain.
-There was also no evidence that treatment effects grew over time. Plant
-community composition was remarkably stable over the course of the
-experiment and did not differ among treatments.
+relationship between soil moisture and ANPP was similar among drought
+and control treatments, but the irrigation treatment had a lower slope
+than the control treatment. However, ANPP response to availabile soil
+moisture was weak and uncertain regardless of treatment, with all slopes
+overlapping zero. There was also no evidence that treatment effects grew
+over time. Plant community composition was remarkably stable over the
+course of the experiment and did not differ among treatments.
 
-\textbf{Discussion.} At least in the short-term, ecosystem functional
-response and community composition in this sagebrush steppe system is
-consistent, even at extreme precipitation levels. The similarity of
+\textbf{Discussion.} Despite some evidence that ecosystem functional
+response saturated under extreme wet conditions, the response of ANPP to
+soil moisture and community compositional changes in this sagebrush
+steppe system were consistently small. The similarity, or difference, of
 ecosystem functional response across treatments was not due to
 compensatory shifts at the plant community level, but instead reflects
-the consistent responses of individual species. Such consistency may be
-due to bet-hedging strategies where species respond weakly to soil
-moisture to ensure low variance of long-term success.
+the weak responses of individual species. Such responses may be due to
+bet-hedging strategies where species respond weakly to even extreme
+levels of soil moisture to ensure low variance of long-term success.
 % Dummy abstract text.

### 2017-08-22 manuscript/usses_water_ms.Rmd
more abstract edits

@@
 ---
-title: Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe
+title: Consistent, and weak, ecosystem functional response across precipitation extremes in a sagebrush steppe
 preprint: false
@@
 
-  **Discussion.** Despite some evidence that ecosystem functional response saturated under extreme wet conditions, the response of ANPP to soil moisture and community compositional changes in this sagebrush steppe system were consistently small.
+  **Discussion.** Despite some evidence that ecosystem functional response saturated under extreme wet conditions, the response of ANPP to soil moisture was consistintly weak and community composition was remarkably stable.
 The similarity, or difference, of ecosystem functional response across treatments was not due to compensatory shifts at the plant community level, but instead reflects the weak responses of individual species.
@@
 But under irrigation, ecosystem functional response was different from the control treatment while community composition remained unchanged (Fig. 1, top right).
-Thus, we found some evidence for a saturating response at high water availability (Fig. 3), but the response of ANPP to soil moisture was weak overall, regardless of treatment (Table 1).
+Thus, we found some evidence for a saturating response at high water availability (Fig. 3A, right panel).
+However, the response of ANPP to soil moisture was consitently weak, regardless of treatment (Table 1).
 <!---

### 2017-08-23 manuscript/usses_water_ms.Rmd
tiny ms update

--- 2017-08-22 manuscript/usses_water_ms.Rmd
@@
 We recorded data on aboveground net primary productivity (ANPP), species abundance, and soil moisture.
-We fit a random intercept, random slope model to determine if the relationship between ANPP and soil moisture differed among treatments.
+We fit a generalized linear mixed effects model to determine if the relationship between ANPP and soil moisture differed among treatments.
 We used nondimensional multivariate scaling to quantify community composition over the five years.
 
-  **Results.** Ecosystem functional response, defined as the relationship between soil moisture and ANPP, did not vary across drought, control, and irrigation treatments.
-ANPP responded positively to availabile soil moisture, but the response was weak and uncertain.
+  **Results.** Ecosystem functional response, defined as the linear relationship between soil moisture and ANPP, was similar among drought and control treatments, but the irrigation treatment had a slightly lower slope.
+ANPP responded positively to availabile soil moisture, but the response was weak and uncertain regardless of treatment.
 There was also no evidence that treatment effects grew over time.

### 2017-08-23 manuscript/usses_water_ms.tex

@@
 
-\title{Consistent ecosystem functional response across precipitation extremes
-in a sagebrush steppe}
+\title{Consistent, and weak, ecosystem functional response across precipitation
+extremes in a sagebrush steppe}
 
@@
 response saturated under extreme wet conditions, the response of ANPP to
-soil moisture and community compositional changes in this sagebrush
-steppe system were consistently small. The similarity, or difference, of
-ecosystem functional response across treatments was not due to
-compensatory shifts at the plant community level, but instead reflects
-the weak responses of individual species. Such responses may be due to
-bet-hedging strategies where species respond weakly to even extreme
-levels of soil moisture to ensure low variance of long-term success.
+soil moisture was consistintly weak and community composition was
+remarkably stable. The similarity, or difference, of ecosystem
+functional response across treatments was not due to compensatory shifts
+at the plant community level, but instead reflects the weak responses of
+individual species. Such responses may be due to bet-hedging strategies
+where species respond weakly to even extreme levels of soil moisture to
+ensure low variance of long-term success.
 % Dummy abstract text.
@@
 composition remained unchanged (Fig. 1, top right). Thus, we found some
-evidence for a saturating response at high water availability (Fig. 3),
-but the response of ANPP to soil moisture was weak overall, regardless
-of treatment (Table 1).
+evidence for a saturating response at high water availability (Fig. 3A,
+right panel). However, the response of ANPP to soil moisture was
+consitently weak, regardless of treatment (Table 1).
 

### 2017-08-23 manuscript/usses_water_ms.tex
trivia

@@
   \hline
-Intercept & Control & 0.02 & 0.03 & -1.06 & 1.01 \\ 
-  Intercept & Drought & -0.04 & -0.04 & -1.39 & 1.21 \\ 
-  Intercept & Irrigation & -0.13 & -0.11 & -1.35 & 1.09 \\ 
-  Slope & Control & 0.63 & 0.64 & -0.33 & 1.61 \\ 
-  Slope & Drought & 0.46 & 0.47 & -0.59 & 1.52 \\ 
-  Slope & Irrigation & 0.26 & 0.26 & -0.60 & 1.12 \\ 
+Intercept & Control & 0.05 & 0.06 & -0.91 & 0.97 \\ 
+  Intercept & Drought & -0.01 & -0.02 & -1.27 & 1.25 \\ 
+  Intercept & Irrigation & -0.09 & -0.08 & -1.25 & 1.04 \\ 
+  Slope & Control & 0.62 & 0.60 & -0.27 & 1.60 \\ 
+  Slope & Drought & 0.45 & 0.45 & -0.57 & 1.52 \\ 
+  Slope & Irrigation & 0.26 & 0.25 & -0.53 & 1.09 \\ 
    \hline

### 2017-08-25 manuscript/Appendix2.Rmd
figure updates; edits results section

@@
 
-\noindent{}All fixed effect $\beta$s were drawn from normally distributed priors with mean 0 and standard deviation 1: $\boldsymbol{\beta} \sim \text{Normal}(0,1)$.
+\noindent{}All fixed effect $\beta$s were drawn from normally distributed priors with mean 0 and standard deviation 5: $\boldsymbol{\beta} \sim \text{Normal}(0,5)$.
 $\gamma$ random effects were drawn from a multivariate prior centered on zero with a shared variance covariance matrix:
@@
 \Sigma = \sigma_{plt}^2 \textbf{R}, \\
-\sigma_{plt} \sim \text{Uniform}(0,1), \\
-\textbf{R} \sim \text{Cholesky}(1)
+\sigma_{plt} \sim \text{Cauchy}(0,2.5), \\
+\textbf{R} \sim \text{LKJ}(2.0)
 \end{align}
@@
   int<lower=0> Nobs;           # number of observations
-  int<lower=0> Nppts;          # number of precip levels to predict
   int<lower=0> Nyears;         # number of years
@@
   row_vector[Npreds2] z[Nobs]; # simple design matrix for random effects 
-  matrix[Nppts,Npreds] newx;   # design matrix for predictions 
-  matrix[Npreds2,Npreds2] R;	 # priors for covariance matrix
   int plot_id[Nobs];           # vector of plot ids
@@
   ####  PRIORS
-  sigma_year ~ cauchy(0,2)
+  sigma_u ~ cauchy(0,2.5)
+  sigma_year ~ cauchy(0,2.5)
   year_off ~ normal(0,sigma_year); # priors on year effects, shared variance
-  beta ~ normal(0,1);	             # priors on treatment coefficients
-  L_u ~ lkj_corr_cholesky(1);      # prior on the cholesky factor which controls the 
+  beta ~ normal(0,5);	             # priors on treatment coefficients
+  L_u ~ lkj_corr_cholesky(2.0);      # prior on the cholesky factor which controls the 
                                    # correlation between plot level treatment effects
@@
 }
+
+generated quantities{
+  corr_matrix[Npreds2] R = multiply_lower_tri_self_transpose(L_u);
+  cov_matrix[Npreds2] V = quad_form_diag(R,sigma_u);
+}
 ```
@@
 ```{r scatterplot, fig.cap="Scatterplot, on the arithmetic scale, of the data and model estimates shown as solid lines.
-Model estimates come from treatment level coefficients (colored lines).", echo=FALSE, message=FALSE}
+Model estimates come from treatment level coefficients (colored lines).", echo=FALSE, message=FALSE, warning=FALSE}
 library(tidyverse)    # Data munging
@@
 
+regression_limited <- {}
+for(dotrt in unique(regression$Treatment)){
+  tmpdat <- filter(regression, Treatment == dotrt)
+  tmprange <- range(anpp_data[which(anpp_data$Treatment==dotrt),"total_seasonal_vwc"])
+  tmpdat$backtrans_vwc[tmpdat$backtrans_vwc < tmprange[1] | tmpdat$backtrans_vwc > tmprange[2]] <- NA
+  regression_limited <- rbind(regression_limited, tmpdat)
+}
+
 ggplot(anpp_data, aes(x=total_seasonal_vwc,y=anpp))+
   geom_point(shape=21,color="grey25",alpha=0.8,aes(fill=Treatment))+
-  geom_line(data=regression, aes(x=backtrans_vwc, y=exp(backtrans_estimate), color=Treatment), size=1)+
+  geom_line(data=regression_limited, aes(x=backtrans_vwc, y=exp(backtrans_estimate), color=Treatment), size=1)+
   scale_fill_brewer(palette = "Set2")+
@@
 
+```{r posteriors, fig.cap="Prior (dashed lines) and posterior (solid line) distributions of intercepts and slopes for each treatment.
+The slope represents the main effect of soil moisture on log(ANPP).
+The red line marks 0.
+Shrinkage of the posterior distribution and/or changes in the mean indicate the data informed model parameters beyond the information contained in the prior for all coefficients.",echo=FALSE, message=FALSE, warning=FALSE}
+posteriors <- treat_effects %>%
+  separate(param_name, into = c("Treatment", "Coefficient")) %>%
+  mutate(prior = rnorm(n = nrow(treat_effects), 0, 5)) %>%
+  gather(Distribution, Value, -iteration, -Treatment, -Coefficient)
+
+ggplot(posteriors, aes(x=Value, linetype=Distribution))+
+    geom_line(stat = "density")+
+    geom_vline(aes(xintercept=0), color="red", size=0.2)+
+    scale_x_continuous(limits=c(-5,5))+
+    scale_linetype_discrete(labels = c("Posterior","Prior"), name = NULL)+
+    facet_grid(Coefficient~Treatment)+
+    ylab("Density")+
+    theme_few()+
+    theme(axis.text = element_text(size=7))
+```
+
+\newpage{}
+
 ```{r year_effects, fig.cap="Posterior distributions of random year effects (intercept offsets).
-Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.", echo=FALSE, message=FALSE, warning=FALSE}
+Kernel bandwidths of posterior densities were adjusted by a factor of 5 to smooth the density curves for visual clarity.", echo=FALSE, message=FALSE, warning=FALSE}
 library(tidyverse)    # Data munging

### 2017-08-25 manuscript/usses_water_ms.Rmd
figure updates; edits results section

--- 2017-08-22 manuscript/usses_water_ms.Rmd
@@
 Thus, those three years represent a lower magnitude of absolute change in precipitation experienced by the treatments.
-Averaged across treatments, ANPP varied from a minimum of 74.5 g m$^{-2}$ in 2014 to a maximum of 237.1 g m$^{-2}$ in 2016 (Fig. 2C).
+ANPP varied from a minimum of 74.5 g m$^{-2}$ in 2014 to a maximum of 237.1 g m$^{-2}$ in 2016 when averaged across treatments (Fig. 2C).
 ANPP was slightly higher in irrigation plots and slightly lower in drought plots (Fig. 2C), corresponding to estimated soil volumetric water content (VWC) differences among treatments (Fig. 2B).
@@
 
-Cumulative March-June soil moisture had a weak positive effect on ANPP (Table 1; Fig.
-3), but the effect is associated with high uncertainty, with credible intervals that overlap zero, regardless of treatment (Table 1).
+Cumulative March-June soil moisture had a weak positive effect on ANPP (Table 1; Fig. 3).
+The effect of soil moisture for each treatment is associated with high uncertainty, however, with credible intervals that overlap zero (Table 1).
+Although the parameter estimates for the effect of soil moisture overlap zero, the posterior distributions of the slopes all shrank and shifted to more positive values relative to the prior distributions (Fig.
+A2-2).
 Ecosystem functional response was similar among treatments (Table 1; Fig. 3B), but there is evidence that the slope for the irrigation treatment is less than the slope for the control treatment.
-This evidence comes from interpreting the posterior distribution of the slope offset for the irrigation treatment, from which we calculate a 98% probability that the estimate is less than zero (Fig.
+This evidence comes from interpreting the posterior distribution of the slope offset for the irrigation treatment, from which we calculate a 99% one-tailed probability that the estimate is less than zero (Fig.
 3A, right panel).
@@
 \item [Appendix 1.] Additional methods and information on estimating aboveground net primary productivity.
-\item [Appendix 2.] Details of the hierarchical Bayesian model, Fig. A2-1, and Fig. A2-2.
+\item [Appendix 2.] Details of the hierarchical Bayesian model, Fig. A2-1, Fig. A2-2, and Fig. A2-3.
 \item [Appendix 3.] Details on analysis of precipitation historical range of variability and Fig. A3-1.
@@
 We only show the one-tailed probability for the value being less than zero because the median of each distribution is less than zero.
-Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.
+Kernel bandwidths of posterior densities were adjusted by a factor of 5 to smooth the distribution for visual clarity.
 (B) Scatterplot of the data and model estimates shown a solid lines.

### 2017-08-25 manuscript/usses_water_ms.tex
figure updates; edits results section

@@
   \hline
-Intercept & Control & 0.05 & 0.06 & -0.91 & 0.97 \\ 
-  Intercept & Drought & -0.01 & -0.02 & -1.27 & 1.25 \\ 
-  Intercept & Irrigation & -0.09 & -0.08 & -1.25 & 1.04 \\ 
-  Slope & Control & 0.62 & 0.60 & -0.27 & 1.60 \\ 
-  Slope & Drought & 0.45 & 0.45 & -0.57 & 1.52 \\ 
-  Slope & Irrigation & 0.26 & 0.25 & -0.53 & 1.09 \\ 
+Intercept & Control & 0.22 & 0.16 & -1.25 & 2.16 \\ 
+  Intercept & Drought & 0.64 & 0.44 & -1.37 & 3.79 \\ 
+  Intercept & Irrigation & -0.51 & -0.37 & -2.99 & 1.26 \\ 
+  Slope & Control & 1.31 & 1.14 & -0.30 & 3.83 \\ 
+  Slope & Drought & 1.18 & 1.00 & -0.62 & 4.12 \\ 
+  Slope & Irrigation & 0.83 & 0.68 & -0.61 & 3.12 \\ 
    \hline

### 2017-08-25 code/04_plot_model_results.R

@@ -173,32 +173,35 @@ suppressWarnings( # ignore wanrnings about NA values
 )
 
 suppressWarnings( # ignore wanrnings about NA values
-  gridplot <- cowplot::plot_grid(treat_posteriors, regress_plot, nrow = 2, labels = c("A)","B)"))
+  gridplot <- cowplot::plot_grid(treat_posteriors, NULL, regress_plot, 
+                                 rel_heights = c(1, 0.08, 1),
+                                 nrow = 3, 
+                                 labels = c("A)","","B)"))
 )
 ggsave("../figures/glmm_main_results.png", plot = gridplot, width = 7, height = 5, units = "in", dpi =120)
 
 
 
 ####
-####  PLOT PRIORS AND POSTERIORS ----
+####  PLOT PRIORS AND POSTERIORS (now in App. 2) ----
 ####
-posteriors <- treat_effects %>%
-  separate(param_name, into = c("Treatment", "Coefficient")) %>%
-  mutate(prior = rnorm(n = nrow(treat_effects), 0, 5)) %>%
-  gather(Distribution, Value, -iteration, -Treatment, -Coefficient)
-
-suppressWarnings( # ignore wanrnings about NA values
-  ggplot(posteriors, aes(x=Value, linetype=Distribution))+
-    geom_line(stat = "density")+
-    geom_vline(aes(xintercept=0), color="red", size=0.2)+
-    scale_x_continuous(limits=c(-5,5))+
-    scale_linetype_discrete(labels = c("Posterior","Prior"), name = NULL)+
-    facet_grid(Coefficient~Treatment)+
-    ylab("Density")+
-    theme_few()+
-    theme(axis.text = element_text(size=7))
-)
-ggsave("../figures/priors_and_posteriors.png", width = 7, height = 4, units = "in", dpi =120)
+# posteriors <- treat_effects %>%
+#   separate(param_name, into = c("Treatment", "Coefficient")) %>%
+#   mutate(prior = rnorm(n = nrow(treat_effects), 0, 5)) %>%
+#   gather(Distribution, Value, -iteration, -Treatment, -Coefficient)
+# 
+# suppressWarnings( # ignore wanrnings about NA values
+#   ggplot(posteriors, aes(x=Value, linetype=Distribution))+
+#     geom_line(stat = "density")+
+#     geom_vline(aes(xintercept=0), color="red", size=0.2)+
+#     scale_x_continuous(limits=c(-5,5))+
[…]

### 2017-08-25 manuscript/usses_water_ms.Rmd
writing the discussion

@@
 
-At a given site, the functional response of aboveground net primary productivity (ANPP) to water availability (e.g., soil moisture) can be characterized by fitting a model to historical observations of ANPP and soil moisture.
+The functional response of aboveground net primary productivity (ANPP) to water availability (e.g., soil moisture) can be characterized by fitting a model to historical observations of ANPP and soil moisture at a given site.
 However, the fitted functional response may provide an incomplete picture because future conditions are likely to be outside the historical range of variability [@Smith2011].
-For example, historical trends may underestimate the potential for the soil moisture-ANPP relationship to saturate if soil moisture is pushed far beyond typical levels.
+For example, historical trends may underestimate the potential for the soil moisture-ANPP relationship to saturate or weaken if soil moisture is pushed far beyond typical levels.
 Saturating relationships are actually common [@Hsu2012; @Gherardi2015a], perhaps because other resources, like nitrogen, become more limiting in wet years than dry years.
@@
 Changes in species identities and abundances can alter an ecosystem's functional response to water availability because different species have different physiological thresholds for producing biomass.
-@Smith2009 introduced the 'Hierarchical Response Framework' (HRF) for understanding the interplay of community composition and ecosystem functioning in response to resource manipulations over time.
+@Smith2009 introduced the 'Hierarchical Response Framework' for understanding the interplay of community composition and ecosystem functioning in response to resource manipulations over time.
 In the near term, ecosystem functioning such as ANPP will reflect the physiological responses of individual species to the manipulated resource level.
 For example, ANPP may decline under simulated drought because the initial community consisted of drought-intolerant species [@Hoover2014].
-Over longer time spans, ecosystem functioning may recover as new species colonize or initial species reorder in relative abundance.
+Ecosystem functioning may recover over longer time spans as new species colonize or initial species reorder in relative abundance.
 For example, ANPP may initially decline, but eventually rise back to pre-treatment levels once drought-tolerant species become more abundant and compensate for drought-intolerant species [@Hoover2014].
@@
 Altering the amount of precipitation over many years should provide insight into the time scales at which water-limited ecosystems respond to chronic resource alteration.
-Following the HRF, we propose four alternative predictions for the effect of precipitation manipulation on the ecosystem functional response to soil moisture, that is, the soil moisture-ANPP relationship (Fig.
-1).
+We propose four alternative predictions for the effect of precipitation manipulation on the ecosystem functional response to soil moisture based on the Hierarchical Response Framework (Fig. 1).
+We define 'ecosystem functional response' as the relationship between available soil moisture and ANPP.
 The four predictions are based on possible outcomes at the community (e.g., community composition) and ecosystem (e.g., soil moisture-ANPP regression) levels.
@@
 In this case, changes in ANPP simply follow the soil moisture-ANPP relationship under ambient conditions.
-This corresponds to the early phases of the HRF, where ecosystem response is due to the physiological responses of individual species.
+This corresponds to the early phases of the Hierarchical Respone Framework, where ecosystem response follows the physiological responses of individual species.
 Second, the ecosystem functional response changes but community composition remains the same (Fig. 1, top right).
@@
 All four outcomes are possible in any given ecosystem, but the time scales at which the different scenarios play out likely differ [@Smith2009; @Wilcox2016; @Knapp2017].
-Thus, our task is not to test the validity of the HRF, but rather to amass information on how quickly ecosystem functional responses change in different ecosystems.
-Likewise, we need to understand whether changes at the ecosystem level are driven by community level changes or individual level responses.
+Thus, our task is not to test the validity of the Hierarchical Response Framework, but rather to amass information on how quickly ecosystem functional responses change in different ecosystems.
+We also need to understand whether changes at the ecosystem level are driven by community level changes or individual level responses.
 
@@
 Does the relationship saturate under the irrigation treatment, at high soil moisture?
-To answer these questions we fit a random intercept, random slope model to test whether the regressions differed among treatments.
-We also analyzed community composition over time, allowing us to place our experimental results within the framework of the HRF and our competing predictions (Fig. 1).
+To answer these questions we fit a generalized linear mixed effects model to test whether the regressions differed among treatments.
+We also analyzed community composition over time, allowing us to place our experimental results within the framework our competing predictions (Fig. 1).
 
@@
 
-Drought and irrigation treatments were designed to decrease and increase the amount of ambient precipitation by 50\%, respectively.
+Drought and irrigation treatments were designed to decrease and increase the amount of ambient precipitation by 50\%.
 To achieve this, we used a system of rain-out shelters and automatic irrigation [@Gherardi2013].
@@
 We coupled this temporally intensive soil moisture sampling with spatially extensive readings taken at six points within all 16 plots and associated ambient measurement areas.
-These snapshot data were collected on 06/06/2012, 04/29/2015, 05/07/2015, 06/09/2015, and 05/10/2016 using a handheld EC-5 sensor.
+These snapshot data were collected on 06-06-2012, 04-29-2015, 05-07-2015, 06-09-2015, and 05-10-2016\footnote{Dates formatted as: mm-dd-yyyy.} using a handheld EC-5 sensor.
 
@@
 For each plot and year, we averaged the four readings for each wavelength and then calculated a greenness index based on the same bands used to calculate NDVI using the MODIS and AVHRR bands for NDVI.
-To convert the greenness index to ANPP we regressed the greenness index against the dry biomass weight from the ten calibration plots.
-We fit regressions to MODIS-based index and AVHRR-based index for each year and retained the regression with the better fit.
-Using the best regression equation for each year, we predicted ANPP (Appendix 1).
+We regressed the greenness index against the dry biomass weight from the ten calibration plots to convert the greenness index to ANPP.
+We fit regressions to a MODIS-based index and an AVHRR-based index for each year and retained the regression with the better fit based on $R^2$ values.
+We then predicted ANPP using the best regression equation for each year (Appendix 1).
 
 Species composition data came from two sources: yearly census maps for each plot made using a pantograph [@Hill1920] and yearly counts of annual species in each plot.
-From those maps, we have data on the density of all annuals and perennials forbs, basal cover of perennial grasses, and canopy cover of shrubs.
+We extracted the mapped information to obtain data on the density of all annuals and perennials forbs, the basal cover of perennial grasses, and the canopy cover of shrubs.
 We made a large plot-treatment-year by species matrix, where columns were filled with either basal cover or density, depending on the measurement made for the particular species.
-So we could analyze the different types of data together, we standardized the values in each column.
+We standardized the values in each column so we could analyze the different types of data together.
 This puts all abundance values on the same scale, meaning that common and rare species are weighted equally.
-Nonetheless, if we assume that rare species will respond more than common ones, then this approach is anti-conservative for detecting community change.
-This means that our approach is biased toward detecting compositional changes.
+This approach is anti-conservative for detecting community change if we assume that rare species will respond more than common ones, which biases our approach toward detecting compositional changes.
 
@@
 \begin{description}
-\item [H1] The coefficient for drought$\times$VWC is positive and different from zero.
-\item [H2] The coefficient for irrigation$\times$VWC is negative and different from zero.
+\item [H1.] The coefficient for drought$\times$VWC is positive and different from zero.
+\item [H2.] The coefficient for irrigation$\times$VWC is negative and different from zero.
 \end{description}
 
-To account for the fact that observations within plots and years are not independent, we include two random effects.
+We include two random effects to account for the fact that observations within plots and years are not independent. 
 Specifically, we include plot-specific offsets ($\boldsymbol{\gamma}$) for the intercept and slope terms and year-specific intercept offsets ($\eta_t$). 
 The covariate vector $\textbf{z}_i$ for each observation *i* has two elements: a 1 for the intercept and the scaled value of VWC for that plot and year. 
-The plot-specific coefficients are modeled hierarchically, where plot level coefficients are drawn from a multivariate normal distribution with mean 0 and a variance-covariance strucutre that allows the intercept and slope terms to be correlated:
+The plot-specific coefficients are modeled hierarchically, where plot level coefficients are drawn from a multivariate normal distribution with mean 0 and a variance-covariance structure that allows the intercept and slope terms to be correlated:
 
@@
 
-We fit the model using a Bayesian approach, obtaining posterior estimates of all unknowns via the No-U-Turn Hamiltonian Monte Carlo sampler in Stan [@stan2016].
+We fit the model using a Bayesian approach and obtained posterior estimates of all unknowns via the No-U-Turn Hamiltonian Monte Carlo sampler in Stan [@stan2016].
 We used the R package 'rstan' [@rstan2016] to link R [@R2016] to Stan.
@@
 Trace plots of all parameters were visually inspected to ensure well-mixed chains and convergence.
-We also made sure all scale reduction factors ($\hat{R}$) were less than 1.1.
+We also made sure all scale reduction factors ($\hat{R}$ values) were less than 1.1 [@Gelman2009].
 
-To see if community composition differed among treatments through time, we used non-dimensional multivariate scaling (NMDS) based on Bray-Curtis distances.
-For each year of the experiment, we first calculated Bray-Curtis distances among all plots, and then extracted those distances for use in the NMDS.
-Because we standardized species' abundances, some values were negative, which is not allowed for calculating Bray-Curtis distances.
+We used non-dimensional multivariate scaling (NMDS) based on Bray-Curtis distances to see if community composition differed among treatments through time.
+We first calculated Bray-Curtis distances among all plots for each year of the experiment and then extracted those distances for use in the NMDS.
+Some values of standardized species' abundances were negative, which is not allowed for calculating Bray-Curtis distances.
 We simply added '2' to each abundance value to ensure all values were greater than zero.
@@
 Cumulative March-June soil moisture had a weak positive effect on ANPP (Table 1; Fig. 3).
-The effect of soil moisture for each treatment is associated with high uncertainty, however, with credible intervals that overlap zero (Table 1).
+The effect of soil moisture for each treatment is associated with high uncertainty, however, with 95\% Bayesian credible intervals that overlap zero (Table 1).
 Although the parameter estimates for the effect of soil moisture overlap zero, the posterior distributions of the slopes all shrank and shifted to more positive values relative to the prior distributions (Fig.
-A2-2).
+A2-2), which indicates the data did influence parameter estimates beyond the information from the uninformative priors.
 Ecosystem functional response was similar among treatments (Table 1; Fig. 3B), but there is evidence that the slope for the irrigation treatment is less than the slope for the control treatment.
@@
 3A, right panel).
-There was no evidence that the treatment effects became more important over time because there was no directional trend in the random year effects (Fig. A2-2).
+There was no evidence that the treatment effects became more important over time because there was no directional trend in the random year effects (Fig. A2-3).
 
@@
 In no year did community composition among treatments not overlap, and they were equally dispersed in all years (Table 2; Fig. 5).
-Likewise, community composition was remarkably stable over time, with no evidence of divergence among treatments (Table 2; Fig. 5).
+Community composition was also remarkably stable over time, with no evidence of divergence among treatments (Table 2; Fig. 5).
 
@@
 The results of our five year experiment in a sagebrush steppe conform to two of our four predictions, depending on treatment.
-Under chronic drought, neither ecosystem functional response nor community composition changed (Fig. 1, top left).
-But under irrigation, ecosystem functional response was different from the control treatment while community composition remained unchanged (Fig. 1, top right).
-Thus, we found some evidence for a saturating response at high water availability (Fig. 3A, right panel).
-However, the response of ANPP to soil moisture was consitently weak, regardless of treatment (Table 1).
+Neither ecosystem functional response nor community composition changed under chronic drought (Fig. 3A, Fig. 4), consistent with the top left prediction in Fig. 1.
+Ecosystem functional response under chronic irrigation was different from the control treatment, but community composition remained unchanged (Fig.
+3A, Fig.
+4), consistent with the top right prediction in Fig.
+1.
+The altered ecosystem functional response under irrigation matched our expectations becuase we found evidence for a saturating response for the irrigation treatment (Fig. 3A, right panel).
+However, the response of ANPP to soil moisture was consitently weak, regardless of treatment (Table 1, Fig. A2-2).
 <!---
@@
 
-First, it could be that our manipulations were not large enough to induce a response.
-That is, maybe a 50\% decrease/increase in any given year is not abnormal given our site's historical range of variability [@Knapp2017].
-We cannot definitively rule out this possibility, but we have reason to believe our manipulations *should* have been large enough.
+First, it could be that our drought manipulation was not large enough to induce a response.
+That is, maybe a 50\% decrease in any given year is not abnormal given our site's historical range of variability [@Knapp2017].
+We cannot definitively rule out this possibility, but we have reason to believe our drought treatment *should* have been large enough.
 Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of mean growing season precipitation necessary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site.
@@
 Thus, our drought treatment represented extreme precipitation amounts, especially in years where ambient precipitation was below average (Fig. 2A).
-The irrigation treatment may have been too small, yet that is the treatment where we observed an effect (Fig. 3).
+The irrigation treatment may have been too small, yet that is the treatment where we did observe an effect (Fig. 3).
 
-Second, ANPP at our site may be influenced by factors beyond the window of soil moisture we included in our statistical model.
+Second, ANPP at our site may be influenced by factors beyond the window of soil moisture (cumulative March-June) we included in our statistical model.
 For example, temperature can impact ANPP directly [@Epstein1997] and by exacerbating the effects of soil moisture [@DeBoeck2011].
 Measurements of soil moisture likely contain a signal of temperature, through its impact on evaporation and infiltration, but the measurements will not capture the direct effect of temperature on metabolic and physiological processes.
-Likewise, we did not redistribute snow across our treatments in the winter, and snow melt may spur early spring growth.
-These statistical issues of missing potentially important covariates could explain the weak and uncertain relationship we observed between soil moisture and ANPP.
+We also did not redistribute snow across our treatments in the winter, and snow melt may spur early spring growth.
+These statistical issues of missing potentially important covariates could explain the weak and uncertain relationship we observed between soil moisture and ANPP across all treatments (Table 1, Fig.
+A2-2).
 
-Third, the life history traits of the dominant species in our study ecosystem may explain the consistent but weak and uncertain effect of soil moisture on ANPP (Fig. 3).
+Third, the life history traits of the dominant species in our study ecosystem may explain the consistently positive, but weak and uncertain, effect of soil moisture on ANPP (Fig. 3).
 Species that live in variable environments, such as cold deserts, must have strategies to ensure long-term success as conditions vary.
@@
 Likewise, the dominant shrub in our focal ecosystem, *Artemisia tripartita*, has access to water deep in the soil profile thanks to a deep root system [@Kulmatiski2017a].
+
+The weaker soil moisture-ANPP relationship we observed for the irrigation treatment, evidenced by the less positive slope relative to controls (Fig.
+3A, right panel), has three non-exclusive explanations.
+First, species' biomass production may have hit physiological thresholds and/or become limited by other factors, such as nitrogen availability [@LeBauer2008].
+\pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES? WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY THIS IS...}
+Second, the less positive slope could be another indicator of bet-hedging in this community.
+Avoiding high biomass production in wet years could benefit species in the long run if climate conditions are variable, and reducing growth
 

### 2017-08-25 manuscript/usses_water_ms.tex
writing the discussion

@@
 
-At a given site, the functional response of aboveground net primary
-productivity (ANPP) to water availability (e.g., soil moisture) can be
-characterized by fitting a model to historical observations of ANPP and
-soil moisture. However, the fitted functional response may provide an
+The functional response of aboveground net primary productivity (ANPP)
+to water availability (e.g., soil moisture) can be characterized by
+fitting a model to historical observations of ANPP and soil moisture at
+a given site. However, the fitted functional response may provide an
 incomplete picture because future conditions are likely to be outside
@@
 historical trends may underestimate the potential for the soil
-moisture-ANPP relationship to saturate if soil moisture is pushed far
-beyond typical levels. Saturating relationships are actually common
-\citep{Hsu2012, Gherardi2015a}, perhaps because other resources, like
-nitrogen, become more limiting in wet years than dry years. Knowing the
-curvature of the soil moisture-ANPP relationship at extreme
+moisture-ANPP relationship to saturate or weaken if soil moisture is
+pushed far beyond typical levels. Saturating relationships are actually
+common \citep{Hsu2012, Gherardi2015a}, perhaps because other resources,
+like nitrogen, become more limiting in wet years than dry years. Knowing
+the curvature of the soil moisture-ANPP relationship at extreme
 precipitation levels is critical for understanding how ecosystems will
@@
 thresholds for producing biomass. \citet{Smith2009} introduced the
-`Hierarchical Response Framework' (HRF) for understanding the interplay
-of community composition and ecosystem functioning in response to
-resource manipulations over time. In the near term, ecosystem
-functioning such as ANPP will reflect the physiological responses of
-individual species to the manipulated resource level. For example, ANPP
-may decline under simulated drought because the initial community
-consisted of drought-intolerant species \citep{Hoover2014}. Over longer
-time spans, ecosystem functioning may recover as new species colonize or
-initial species reorder in relative abundance. For example, ANPP may
-initially decline, but eventually rise back to pre-treatment levels once
+`Hierarchical Response Framework' for understanding the interplay of
+community composition and ecosystem functioning in response to resource
+manipulations over time. In the near term, ecosystem functioning such as
+ANPP will reflect the physiological responses of individual species to
+the manipulated resource level. For example, ANPP may decline under
+simulated drought because the initial community consisted of
+drought-intolerant species \citep{Hoover2014}. Ecosystem functioning may
+recover over longer time spans as new species colonize or initial
+species reorder in relative abundance. For example, ANPP may initially
+decline, but eventually rise back to pre-treatment levels once
 drought-tolerant species become more abundant and compensate for
@@
 scales at which water-limited ecosystems respond to chronic resource
-alteration. Following the HRF, we propose four alternative predictions
-for the effect of precipitation manipulation on the ecosystem functional
-response to soil moisture, that is, the soil moisture-ANPP relationship
-(Fig. 1). The four predictions are based on possible outcomes at the
-community (e.g., community composition) and ecosystem (e.g., soil
-moisture-ANPP regression) levels.
+alteration. We propose four alternative predictions for the effect of
+precipitation manipulation on the ecosystem functional response to soil
+moisture based on the Hierarchical Response Framework (Fig. 1). We
+define `ecosystem functional response' as the relationship between
+available soil moisture and ANPP. The four predictions are based on
+possible outcomes at the community (e.g., community composition) and
+ecosystem (e.g., soil moisture-ANPP regression) levels.
 
@@
 changes in ANPP simply follow the soil moisture-ANPP relationship under
-ambient conditions. This corresponds to the early phases of the HRF,
-where ecosystem response is due to the physiological responses of
-individual species. Second, the ecosystem functional response changes
-but community composition remains the same (Fig. 1, top right). A
-saturating soil moisture-ANPP response fits this scenario, where
-individual species hit physiological thresholds or are limited by some
-other resource. Third, the ecosystem functional response is consistent
-but underlying community composition changes (Fig. 1, bottom left). In
-this case, changes in species' identities or abundances occur in
-response to altered precipitation levels and species more suited to the
-new conditions compensate for reduced function of initial species.
-Fourth, and last, both ecosystem functional response and community
-composition change (Fig. 1, bottom right). New species, or newly
-abundant species, with different physiological responses completely
-reshape the ecosystem functional response.
+ambient conditions. This corresponds to the early phases of the
+Hierarchical Respone Framework, where ecosystem response follows the
+physiological responses of individual species. Second, the ecosystem
+functional response changes but community composition remains the same
+(Fig. 1, top right). A saturating soil moisture-ANPP response fits this
+scenario, where individual species hit physiological thresholds or are
+limited by some other resource. Third, the ecosystem functional response
+is consistent but underlying community composition changes (Fig. 1,
+bottom left). In this case, changes in species' identities or abundances
+occur in response to altered precipitation levels and species more
+suited to the new conditions compensate for reduced function of initial
+species. Fourth, and last, both ecosystem functional response and
+community composition change (Fig. 1, bottom right). New species, or
+newly abundant species, with different physiological responses
+completely reshape the ecosystem functional response.
 
@@
 \citep{Smith2009, Wilcox2016, Knapp2017}. Thus, our task is not to test
-the validity of the HRF, but rather to amass information on how quickly
-ecosystem functional responses change in different ecosystems. Likewise,
-we need to understand whether changes at the ecosystem level are driven
-by community level changes or individual level responses.
+the validity of the Hierarchical Response Framework, but rather to amass
+information on how quickly ecosystem functional responses change in
+different ecosystems. We also need to understand whether changes at the
+ecosystem level are driven by community level changes or individual
+level responses.
 
@@
 the irrigation treatment, at high soil moisture? To answer these
-questions we fit a random intercept, random slope model to test whether
-the regressions differed among treatments. We also analyzed community
-composition over time, allowing us to place our experimental results
-within the framework of the HRF and our competing predictions (Fig. 1).
+questions we fit a generalized linear mixed effects model to test
+whether the regressions differed among treatments. We also analyzed
+community composition over time, allowing us to place our experimental
+results within the framework our competing predictions (Fig. 1).
 
@@
 Drought and irrigation treatments were designed to decrease and increase
-the amount of ambient precipitation by 50\%, respectively. To achieve
-this, we used a system of rain-out shelters and automatic irrigation
+the amount of ambient precipitation by 50\%. To achieve this, we used a
+system of rain-out shelters and automatic irrigation
 \citep{Gherardi2013}. The rain-out shelters consisted of transparent
@@
 at six points within all 16 plots and associated ambient measurement
-areas. These snapshot data were collected on 06/06/2012, 04/29/2015,
-05/07/2015, 06/09/2015, and 05/10/2016 using a handheld EC-5 sensor.
+areas. These snapshot data were collected on 06-06-2012, 04-29-2015,
+05-07-2015, 06-09-2015, and
+05-10-2016\footnote{Dates formatted as: mm-dd-yyyy.} using a handheld
+EC-5 sensor.
 
@@
 rainfall (Fig. 2A). Thus, those three years represent a lower magnitude
-of absolute change in precipitation experienced by the treatments.
-Averaged across treatments, ANPP varied from a minimum of 74.5 g
-m\(^{-2}\) in 2014 to a maximum of 237.1 g m\(^{-2}\) in 2016 (Fig. 2C).
-ANPP was slightly higher in irrigation plots and slightly lower in
-drought plots (Fig. 2C), corresponding to estimated soil volumetric
-water content (VWC) differences among treatments (Fig. 2B). Such
-differences in soil VWC indicate our treatment infrastructure was
-successful.
+of absolute change in precipitation experienced by the treatments. ANPP
+varied from a minimum of 74.5 g m\(^{-2}\) in 2014 to a maximum of 237.1
+g m\(^{-2}\) in 2016 when averaged across treatments (Fig. 2C). ANPP was
+slightly higher in irrigation plots and slightly lower in drought plots
+(Fig. 2C), corresponding to estimated soil volumetric water content
+(VWC) differences among treatments (Fig. 2B). Such differences in soil
+VWC indicate our treatment infrastructure was successful.
 
 Cumulative March-June soil moisture had a weak positive effect on ANPP
-(Table 1; Fig. 3), but the effect is associated with high uncertainty,
-with credible intervals that overlap zero, regardless of treatment
-(Table 1). Ecosystem functional response was similar among treatments
-(Table 1; Fig. 3B), but there is evidence that the slope for the
-irrigation treatment is less than the slope for the control treatment.
-This evidence comes from interpreting the posterior distribution of the
-slope offset for the irrigation treatment, from which we calculate a
-98\% probability that the estimate is less than zero (Fig. 3A, right
-panel). There was no evidence that the treatment effects became more
-important over time because there was no directional trend in the random
-year effects (Fig. A2-2).
+(Table 1; Fig. 3). The effect of soil moisture for each treatment is
+associated with high uncertainty, however, with credible intervals that
+overlap zero (Table 1). Although the parameter estimates for the effect
+of soil moisture overlap zero, the posterior distributions of the slopes
+all shrank and shifted to more positive values relative to the prior
+distributions (Fig. A2-2). Ecosystem functional response was similar
+among treatments (Table 1; Fig. 3B), but there is evidence that the
+slope for the irrigation treatment is less than the slope for the
+control treatment. This evidence comes from interpreting the posterior
+distribution of the slope offset for the irrigation treatment, from
+which we calculate a 99\% one-tailed probability that the estimate is
+less than zero (Fig. 3A, right panel). There was no evidence that the
+treatment effects became more important over time because there was no
+directional trend in the random year effects (Fig. A2-2).
 

### 2017-08-25 manuscript/usses_water_ms.Rmd
adds bib file

@@
 ---
-title: Consistent, and weak, ecosystem functional response across precipitation extremes in a sagebrush steppe
+title: Ecosystem functional response across precipitation extremes in a sagebrush steppe
 preprint: false
@@
 However, ANPP response to availabile soil moisture was weak and uncertain regardless of treatment, with all slopes overlapping zero.
-There was also no evidence that treatment effects grew over time.
+There was also large variation in ANPP within-years.
 Plant community composition was remarkably stable over the course of the experiment and did not differ among treatments.
 
-  **Discussion.** Despite some evidence that ecosystem functional response saturated under extreme wet conditions, the response of ANPP to soil moisture was consistintly weak and community composition was remarkably stable.
-The similarity, or difference, of ecosystem functional response across treatments was not due to compensatory shifts at the plant community level, but instead reflects the weak responses of individual species.
-Such responses may be due to bet-hedging strategies where species respond weakly to even extreme levels of soil moisture to ensure low variance of long-term success.
+  **Discussion.** Despite some evidence that ecosystem functional response was less positive under extreme wet conditions, the response of ANPP to soil moisture was consistently weak and community composition was remarkably stable.
+Ecosystem functional responses across treatments was not related to compensatory shifts at the plant community level, but instead reflects the weak and variable responses of individual species.
+Such responses may be due to bet hedging strategies where species respond weakly to even extreme levels of soil moisture to ensure low variance of long-term success.
 header-includes: >
@@
   \usepackage{bm,mathrsfs}
-bibliography: ~/Dropbox/Bibliography/usses_water.bib
+bibliography: ~./usses_water.bib
 output: 
@@
 The day immediately following rainfall was also classified as rainy.
-We fit the model using the 'lme4::lmer()' function [@Bates2015] in R [@R2016], with random effects for plot group and date.
+We fit the model using the ```lme4::lmer()``` function [@Bates2015] in R [@R2016], with random effects for plot group and date.
 We weighted observations by the number of unique sensors or spot measurements that were taken in each plot on that day.
@@
 We plotted the first two axes of NMDS scores to see if community composition overlapped, or not, among treatments in each year.
-We used the 'metaMDS()' function in the R package 'vegan' [@Oksanen2016] to calculate Bray-Curtis distances and then to run the NMDS analysis.
-We used the 'vegan::adonis()' function [@Oksanen2016] to perform permutational multivariate analysis of variance to test whether treatment plots formed distinct groupings.
-To test whether treatment plots were equally dispersed, or not, we used the 'vegan::betadisper' function [@Oksanen2016].
+We used the ```vegan::metaMDS()``` function [@Oksanen2016] to calculate Bray-Curtis distances and then to run the NMDS analysis.
+We used the ```vegan::adonis()``` function [@Oksanen2016] to perform permutational multivariate analysis of variance to test whether treatment plots formed distinct groupings.
+To test whether treatment plots were equally dispersed, or not, we used the ```vegan::betadisper()``` function [@Oksanen2016].
 
@@
 The altered ecosystem functional response under irrigation matched our expectations becuase we found evidence for a saturating response for the irrigation treatment (Fig. 3A, right panel).
-However, the response of ANPP to soil moisture was consitently weak, regardless of treatment (Table 1, Fig. A2-2).
+However, the response of ANPP to soil moisture in all treatments was consistently weak (Table 1, Fig. A2-2) and within-year variation of ANPP was large (Fig. 3B)
+
 <!---
@@
 Indeed, we expected ecosystem functional response, community composition, or both to change under the drought treatment, landing us in any box of Fig. 1 \emph{except} the top left.
-So, why did our drought treatment fail to induce ecosystem or community responses?
+Why did our drought treatment fail to induce ecosystem or community responses?
 We can think of three reasons; two are limitations of our study, and one is the life history traits of the species in our focal communities.
@@
 First, it could be that our drought manipulation was not large enough to induce a response.
-That is, maybe a 50\% decrease in any given year is not abnormal given our site's historical range of variability [@Knapp2017].
+That is, a 50\% decrease in any given year may not be abnormal given our site's historical range of variability [@Knapp2017].
 We cannot definitively rule out this possibility, but we have reason to believe our drought treatment *should* have been large enough.
-Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of mean growing season precipitation necessary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site.
+Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of mean growing season precipitation necessary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site (Fig.
+A3-3).
 The 1\% quantile of precipitation at our site is 110 mm, a 47% reduction from the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean growing season precipitation (Appendix 3).
@@
 We also did not redistribute snow across our treatments in the winter, and snow melt may spur early spring growth.
-These statistical issues of missing potentially important covariates could explain the weak and uncertain relationship we observed between soil moisture and ANPP across all treatments (Table 1, Fig.
+These statistical issues of missing potentially important covariates could explain the within-year spread of ANPP (Fig 2C, Fig.
+3B) and the resulting uncertain relationship we observed between soil moisture and ANPP across all treatments (Table 1, Fig.
 A2-2).
@@
 The dry and variable environment of the sagebrush steppe has likely selected for bet hedging species that can maintain function at low water availability and have weak responses to high water availability.
-In so doing, the dominant species in our ecosystem avoid "boom and bust" cycles, which corresponds to the weak effect of soil moisture on ANPP (i.e., the credible intervals for $\beta_1$ overlapping zero), even at precipitation extremes.
+In so doing, the dominant species in our ecosystem avoid "boom and bust" cycles, which corresponds to the weak effect of soil moisture on ANPP (i.e., the Bayesian credible intervals for $\beta$ in Eq.
+1 overlapping zero).
 
-Another strategy to deal with variable environmental conditions is avoidance, which would result in a consistent ecosystem functional response at low soil moisture.
+Another strategy to deal with variable environmental conditions is avoidance, which would also result in a consistent ecosystem functional response between drought and control treatments.
 The perennial plants in this cold desert ecosystem are tolerant to drought conditions [@Bazzaz1979; @Franks2011, A.R. Kleinhesselink, unpublished data].
 For example, many of the perennial grasses in our focal ecosystem avoid drought stress by growing early in the growing season [@Blaisdell1958, A.R. Kleinhesselink, personal observation].
-Likewise, the dominant shrub in our focal ecosystem, *Artemisia tripartita*, has access to water deep in the soil profile thanks to a deep root system [@Kulmatiski2017a].
+Furthermore, the dominant shrub in our focal ecosystem, *Artemisia tripartita*, has access to water deep in the soil profile thanks to a deep root system [@Kulmatiski2017a].
 
@@
 \pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES? WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY THIS IS...}
-Second, the less positive slope could be another indicator of bet-hedging in this community.
-Avoiding high biomass production in wet years could benefit species in the long run if climate conditions are variable, and reducing growth
+Second, the less positive slope could be another indicator of bet hedging in this community.
+Avoiding high biomass production in wet years could benefit species in the long run if climate conditions are variable.
+Third, the less positive slope may be a statistical artifact due the variability of ANPP within years and our low sample size.
 
-In conclusion, our results suggest that five years of $\pm$ 50\% ambient precipitation is not enough to induce a shift in ecosystem functional response in a sagebrush steppe.
-This is likely because the species in our focal plant community are tolerant of drought conditions and bet hedgers in wet conditions, maintaining relatively consistent responses to interannual variation in precipitation to avoid booms and busts.
+In conclusion, our results suggest the species in our focal plant community are tolerant of drought conditions and bet hedgers in wet conditions.
 Longer time series of chronic precipitation alteration may reveal plant community shifts that we did not observe [e.g., @Wilcox2016], in which case species that do not bet hedge may gain prominence and dominate the ecosystem functional response.
@@
       \includegraphics[width=4in]{../figures/hypothesis_figtable.png}
-  \caption{Possible outcomes of chronic resource alteration.}
+  \caption{Possible outcomes of chronic resource alteration based on the 'Hierachical Response Framework' (Smith et al. 2009).}
 \end{figure}

### 2017-08-25 manuscript/usses_water_ms.tex
adds bib file

@@
 
-\title{Consistent, and weak, ecosystem functional response across precipitation
-extremes in a sagebrush steppe}
+\title{Ecosystem functional response across precipitation extremes in a
+sagebrush steppe}
 
@@
 moisture was weak and uncertain regardless of treatment, with all slopes
-overlapping zero. There was also no evidence that treatment effects grew
-over time. Plant community composition was remarkably stable over the
-course of the experiment and did not differ among treatments.
+overlapping zero. There was also large variation in ANPP within-years.
+Plant community composition was remarkably stable over the course of the
+experiment and did not differ among treatments.
 
 \textbf{Discussion.} Despite some evidence that ecosystem functional
-response saturated under extreme wet conditions, the response of ANPP to
-soil moisture was consistintly weak and community composition was
-remarkably stable. The similarity, or difference, of ecosystem
-functional response across treatments was not due to compensatory shifts
-at the plant community level, but instead reflects the weak responses of
-individual species. Such responses may be due to bet-hedging strategies
-where species respond weakly to even extreme levels of soil moisture to
-ensure low variance of long-term success.
+response was less positive under extreme wet conditions, the response of
+ANPP to soil moisture was consistently weak and community composition
+was remarkably stable. Ecosystem functional responses across treatments
+was not related to compensatory shifts at the plant community level, but
+instead reflects the weak and variable responses of individual species.
+Such responses may be due to bet hedging strategies where species
+respond weakly to even extreme levels of soil moisture to ensure low
+variance of long-term success.
 % Dummy abstract text.
@@
 rainfall was also classified as rainy. We fit the model using the
-`lme4::lmer()' function \citep{Bates2015} in R \citep{R2016}, with
-random effects for plot group and date. We weighted observations by the
-number of unique sensors or spot measurements that were taken in each
-plot on that day. We then used the model to predict the average daily
-soil moisture in the treated plots based on the average daily ambient
-soil moisture. We could only predict soil moisture in the treated plots
-on days for which we took at least one ambient soil moisture
-measurement.
+\texttt{lme4::lmer()} function \citep{Bates2015} in R \citep{R2016},
+with random effects for plot group and date. We weighted observations by
+the number of unique sensors or spot measurements that were taken in
+each plot on that day. We then used the model to predict the average
+daily soil moisture in the treated plots based on the average daily
+ambient soil moisture. We could only predict soil moisture in the
+treated plots on days for which we took at least one ambient soil
+moisture measurement.
 
@@
 wavelength and then calculated a greenness index based on the same bands
-used to calculate NDVI using the MODIS and AVHRR bands for NDVI. To
-convert the greenness index to ANPP we regressed the greenness index
-against the dry biomass weight from the ten calibration plots. We fit
-regressions to MODIS-based index and AVHRR-based index for each year and
-retained the regression with the better fit. Using the best regression
-equation for each year, we predicted ANPP (Appendix 1).
+used to calculate NDVI using the MODIS and AVHRR bands for NDVI. We
+regressed the greenness index against the dry biomass weight from the
+ten calibration plots to convert the greenness index to ANPP. We fit
+regressions to a MODIS-based index and an AVHRR-based index for each
+year and retained the regression with the better fit based on \(R^2\)
+values. We then predicted ANPP using the best regression equation for
+each year (Appendix 1).
 
@@
 each plot made using a pantograph \citep{Hill1920} and yearly counts of
-annual species in each plot. From those maps, we have data on the
-density of all annuals and perennials forbs, basal cover of perennial
-grasses, and canopy cover of shrubs. We made a large plot-treatment-year
-by species matrix, where columns were filled with either basal cover or
-density, depending on the measurement made for the particular species.
-So we could analyze the different types of data together, we
-standardized the values in each column. This puts all abundance values
-on the same scale, meaning that common and rare species are weighted
-equally. Nonetheless, if we assume that rare species will respond more
-than common ones, then this approach is anti-conservative for detecting
-community change. This means that our approach is biased toward
+annual species in each plot. We extracted the mapped information to
+obtain data on the density of all annuals and perennials forbs, the
+basal cover of perennial grasses, and the canopy cover of shrubs. We
+made a large plot-treatment-year by species matrix, where columns were
+filled with either basal cover or density, depending on the measurement
+made for the particular species. We standardized the values in each
+column so we could analyze the different types of data together. This
+puts all abundance values on the same scale, meaning that common and
+rare species are weighted equally. This approach is anti-conservative
+for detecting community change if we assume that rare species will
+respond more than common ones, which biases our approach toward
 detecting compositional changes.
@@
 \begin{description}
-\item [H1] The coefficient for drought$\times$VWC is positive and different from zero.
-\item [H2] The coefficient for irrigation$\times$VWC is negative and different from zero.
+\item [H1.] The coefficient for drought$\times$VWC is positive and different from zero.
+\item [H2.] The coefficient for irrigation$\times$VWC is negative and different from zero.
 \end{description}
 
-To account for the fact that observations within plots and years are not
-independent, we include two random effects. Specifically, we include
+We include two random effects to account for the fact that observations
+within plots and years are not independent. Specifically, we include
 plot-specific offsets (\(\boldsymbol{\gamma}\)) for the intercept and
@@
 multivariate normal distribution with mean 0 and a variance-covariance
-strucutre that allows the intercept and slope terms to be correlated:
+structure that allows the intercept and slope terms to be correlated:
 
@@
 
-We fit the model using a Bayesian approach, obtaining posterior
+We fit the model using a Bayesian approach and obtained posterior
 estimates of all unknowns via the No-U-Turn Hamiltonian Monte Carlo
@@
 inspected to ensure well-mixed chains and convergence. We also made sure
-all scale reduction factors (\(\hat{R}\)) were less than 1.1.
+all scale reduction factors (\(\hat{R}\) values) were less than 1.1
+\citep{Gelman2009}.
 
-To see if community composition differed among treatments through time,
-we used non-dimensional multivariate scaling (NMDS) based on Bray-Curtis
-distances. For each year of the experiment, we first calculated
-Bray-Curtis distances among all plots, and then extracted those
-distances for use in the NMDS. Because we standardized species'
-abundances, some values were negative, which is not allowed for
-calculating Bray-Curtis distances. We simply added `2' to each abundance
-value to ensure all values were greater than zero. We plotted the first
-two axes of NMDS scores to see if community composition overlapped, or
-not, among treatments in each year. We used the `metaMDS()' function in
-the R package `vegan' \citep{Oksanen2016} to calculate Bray-Curtis
-distances and then to run the NMDS analysis. We used the
-`vegan::adonis()' function \citep{Oksanen2016} to perform permutational
-multivariate analysis of variance to test whether treatment plots formed
-distinct groupings. To test whether treatment plots were equally
-dispersed, or not, we used the `vegan::betadisper' function
-\citep{Oksanen2016}.
+We used non-dimensional multivariate scaling (NMDS) based on Bray-Curtis
+distances to see if community composition differed among treatments
+through time. We first calculated Bray-Curtis distances among all plots
+for each year of the experiment and then extracted those distances for
+use in the NMDS. Some values of standardized species' abundances were
+negative, which is not allowed for calculating Bray-Curtis distances. We
+simply added `2' to each abundance value to ensure all values were
+greater than zero. We plotted the first two axes of NMDS scores to see
+if community composition overlapped, or not, among treatments in each
+year. We used the \texttt{vegan::metaMDS()} function \citep{Oksanen2016}
+to calculate Bray-Curtis distances and then to run the NMDS analysis. We
+used the \texttt{vegan::adonis()} function \citep{Oksanen2016} to
+perform permutational multivariate analysis of variance to test whether
+treatment plots formed distinct groupings. To test whether treatment
+plots were equally dispersed, or not, we used the
+\texttt{vegan::betadisper()} function \citep{Oksanen2016}.
 
@@
 (Table 1; Fig. 3). The effect of soil moisture for each treatment is
-associated with high uncertainty, however, with credible intervals that
-overlap zero (Table 1). Although the parameter estimates for the effect
-of soil moisture overlap zero, the posterior distributions of the slopes
-all shrank and shifted to more positive values relative to the prior
-distributions (Fig. A2-2). Ecosystem functional response was similar
+associated with high uncertainty, however, with 95\% Bayesian credible
+intervals that overlap zero (Table 1). Although the parameter estimates
+for the effect of soil moisture overlap zero, the posterior
+distributions of the slopes all shrank and shifted to more positive
+values relative to the prior distributions (Fig. A2-2), which indicates
+the data did influence parameter estimates beyond the information from
+the uninformative priors. Ecosystem functional response was similar
 among treatments (Table 1; Fig. 3B), but there is evidence that the
@@
 treatment effects became more important over time because there was no
-directional trend in the random year effects (Fig. A2-2).
+directional trend in the random year effects (Fig. A2-3).
 
@@
 community composition among treatments not overlap, and they were
-equally dispersed in all years (Table 2; Fig. 5). Likewise, community
-composition was remarkably stable over time, with no evidence of
-divergence among treatments (Table 2; Fig. 5).
+equally dispersed in all years (Table 2; Fig. 5). Community composition
+was also remarkably stable over time, with no evidence of divergence
+among treatments (Table 2; Fig. 5).
 
@@
 The results of our five year experiment in a sagebrush steppe conform to
-two of our four predictions, depending on treatment. Under chronic
-drought, neither ecosystem functional response nor community composition
-changed (Fig. 1, top left). But under irrigation, ecosystem functional
-response was different from the control treatment while community
-composition remained unchanged (Fig. 1, top right). Thus, we found some
-evidence for a saturating response at high water availability (Fig. 3A,
-right panel). However, the response of ANPP to soil moisture was
-consitently weak, regardless of treatment (Table 1).
+two of our four predictions, depending on treatment. Neither ecosystem
+functional response nor community composition changed under chronic
+drought (Fig. 3A, Fig. 4), consistent with the top left prediction in
+Fig. 1. Ecosystem functional response under chronic irrigation was
+different from the control treatment, but community composition remained
+unchanged (Fig. 3A, Fig. 4), consistent with the top right prediction in
+Fig. 1. The altered ecosystem functional response under irrigation
+matched our expectations becuase we found evidence for a saturating
+response for the irrigation treatment (Fig. 3A, right panel). However,
+the response of ANPP to soil moisture in all treatments was consistently
+weak (Table 1, Fig. A2-2) and within-year variation of ANPP was large
+(Fig. 3B)
 
@@
 under the drought treatment, landing us in any box of Fig. 1
-\emph{except} the top left. So, why did our drought treatment fail to
-induce ecosystem or community responses? We can think of three reasons;
-two are limitations of our study, and one is the life history traits of
-the species in our focal communities. We first discuss the potential
+\emph{except} the top left. Why did our drought treatment fail to induce
+ecosystem or community responses? We can think of three reasons; two are
+limitations of our study, and one is the life history traits of the
+species in our focal communities. We first discuss the potential
 limitations of our study, and then discuss the biological explanation.
 
-First, it could be that our manipulations were not large enough to
-induce a response. That is, maybe a 50\% decrease/increase in any given
-year is not abnormal given our site's historical range of variability
+First, it could be that our drought manipulation was not large enough to
+induce a response. That is, a 50\% decrease in any given year may not be
+abnormal given our site's historical range of variability
 \citep{Knapp2017}. We cannot definitively rule out this possibility, but
-we have reason to believe our manipulations \emph{should} have been
+we have reason to believe our drought treatment \emph{should} have been
 large enough. Using the methods described by \citet{Lemoine2016}, we
@@
 precipitation necessary to reach the 1\% and 99\% extremes of the
-historical precipitation regime at our site. The 1\% quantile of
-precipitation at our site is 110 mm, a 47\% reduction from the mean, and
-the 99\% quantile is 414 mm, a 77\% increase from mean growing season
-precipitation (Appendix 3). Thus, our drought treatment represented
-extreme precipitation amounts, especially in years where ambient
-precipitation was below average (Fig. 2A). The irrigation treatment may
-have been too small, yet that is the treatment where we observed an
-effect (Fig. 3).
+historical precipitation regime at our site (Fig. A3-3). The 1\%
+quantile of precipitation at our site is 110 mm, a 47\% reduction from
+the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean
+growing season precipitation (Appendix 3). Thus, our drought treatment
+represented extreme precipitation amounts, especially in years where
+ambient precipitation was below average (Fig. 2A). The irrigation
+treatment may have been too small, yet that is the treatment where we
+did observe an effect (Fig. 3).
 
 Second, ANPP at our site may be influenced by factors beyond the window
-of soil moisture we included in our statistical model. For example,
-temperature can impact ANPP directly \citep{Epstein1997} and by
-exacerbating the effects of soil moisture \citep{DeBoeck2011}.
-Measurements of soil moisture likely contain a signal of temperature,
-through its impact on evaporation and infiltration, but the measurements
-will not capture the direct effect of temperature on metabolic and
-physiological processes. Likewise, we did not redistribute snow across
-our treatments in the winter, and snow melt may spur early spring
-growth. These statistical issues of missing potentially important
-covariates could explain the weak and uncertain relationship we observed
-between soil moisture and ANPP.
+of soil moisture (cumulative March-June) we included in our statistical
+model. For example, temperature can impact ANPP directly
+\citep{Epstein1997} and by exacerbating the effects of soil moisture
+\citep{DeBoeck2011}. Measurements of soil moisture likely contain a
+signal of temperature, through its impact on evaporation and
+infiltration, but the measurements will not capture the direct effect of
+temperature on metabolic and physiological processes. We also did not
+redistribute snow across our treatments in the winter, and snow melt may
+spur early spring growth. These statistical issues of missing
+potentially important covariates could explain the within-year spread of
+ANPP (Fig 2C, Fig. 3B) and the resulting uncertain relationship we
+observed between soil moisture and ANPP across all treatments (Table 1,
+Fig. A2-2).
 
 Third, the life history traits of the dominant species in our study
-ecosystem may explain the consistent but weak and uncertain effect of
-soil moisture on ANPP (Fig. 3). Species that live in variable
+ecosystem may explain the consistently positive, but weak and uncertain,
+effect of soil moisture on ANPP (Fig. 3). Species that live in variable
 environments, such as cold deserts, must have strategies to ensure
@@
 cycles, which corresponds to the weak effect of soil moisture on ANPP
-(i.e., the credible intervals for \(\beta_1\) overlapping zero), even at
-precipitation extremes.
+(i.e., the Bayesian credible intervals for \(\beta\) in Eq. 1
+overlapping zero).
 
 Another strategy to deal with variable environmental conditions is
-avoidance, which would result in a consistent ecosystem functional
-response at low soil moisture. The perennial plants in this cold desert
-ecosystem are tolerant to drought conditions \citep[A.R. Kleinhesselink,
-unpublished data]{Bazzaz1979, Franks2011}. For example, many of the
-perennial grasses in our focal ecosystem avoid drought stress by growing
-early in the growing season \citep[A.R. Kleinhesselink, personal
-observation]{Blaisdell1958}. Likewise, the dominant shrub in our focal
-ecosystem, \emph{Artemisia tripartita}, has access to water deep in the
-soil profile thanks to a deep root system \citep{Kulmatiski2017a}.
+avoidance, which would also result in a consistent ecosystem functional
+response between drought and control treatments. The perennial plants in
+this cold desert ecosystem are tolerant to drought conditions
+\citep[A.R. Kleinhesselink, unpublished data]{Bazzaz1979, Franks2011}.
+For example, many of the perennial grasses in our focal ecosystem avoid
+drought stress by growing early in the growing season \citep[A.R.
+Kleinhesselink, personal observation]{Blaisdell1958}. Furthermore, the
+dominant shrub in our focal ecosystem, \emph{Artemisia tripartita}, has
+access to water deep in the soil profile thanks to a deep root system
+\citep{Kulmatiski2017a}.
 
-In conclusion, our results suggest that five years of \(\pm\) 50\%
-ambient precipitation is not enough to induce a shift in ecosystem
-functional response in a sagebrush steppe. This is likely because the
-species in our focal plant community are tolerant of drought conditions
-and bet hedgers in wet conditions, maintaining relatively consistent
-responses to interannual variation in precipitation to avoid booms and
-busts. Longer time series of chronic precipitation alteration may reveal
-plant community shifts that we did not observe
+The weaker soil moisture-ANPP relationship we observed for the
+irrigation treatment, evidenced by the less positive slope relative to
+controls (Fig. 3A, right panel), has three non-exclusive explanations.
+First, species' biomass production may have hit physiological thresholds
+and/or become limited by other factors, such as nitrogen availability
+\citep{LeBauer2008}.
+\pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES? WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY THIS IS...}
+Second, the less positive slope could be another indicator of bet
+hedging in this community. Avoiding high biomass production in wet years
+could benefit species in the long run if climate conditions are
+variable. Third, the less positive slope may be a statistical artifact
+due the variability of ANPP within years and our low sample size.
+
+In conclusion, our results suggest the species in our focal plant
+community are tolerant of drought conditions and bet hedgers in wet
+conditions. Longer time series of chronic precipitation alteration may
+reveal plant community shifts that we did not observe
 \citep[e.g.,][]{Wilcox2016}, in which case species that do not bet hedge

### 2017-08-25 manuscript/usses_water_ms.Rmd
removes all old text

@@
 # INTRODUCTION
-
-<!---
-Ecosystems worldwide are projected to experience either wetter or drier conditions in the future [@Zhang2007; @Greve2014].
-This has important implications for ecosystem functioning because mean annual net primary productivity (ANPP) positively covaries with mean annual precipitation across ecosystems [@Sala1988; @Huxman2004] and interannual responses of ANPP are linked to interannual variability in precipitation within ecosystems [@Lauenroth1992; @Hsu2012; @Gherardi2015a].
-To predict how ecosystems might respond to extreme precipitation levels, we need to understand whether ecosystems tend to follow temporal trends, which assume relatively constant plant community composition, or spatial trends, which allow community composition to shift [@Wilcox2016].
---->
 
@@
 
-<!---
-\begin{align}
-\mu_{j(k(t))} &= \beta_{0,j(k)} + \beta_{1,j(k)}x_{k(t)} + \gamma_t, \\
-y_{j(k(t))} &\sim \text{Normal} \left(\mu_{j(k(t))}, \sigma^2_{k} \right),
-\end{align}
----->
-
 \begin{align}
@@
 
-<!----
-is the intercept for plot *j* associated with treatment *k*, $\beta_{1,j(k)}$ is the slope term for the effect of soil moisture for plot *j* associated with treatment *k*, $\gamma_t$ is the intercept offset for year *t*, and $\sigma^2_k$ is the process variance for treatment *k*.
-Data include the standardized log(ANPP) observations $\left( y_{j(k(t))} \right)$ and soil moisture ($x_{k(t)}$).
-
-The intercept and slope terms are modeled hierarchically to account for the non-independence of observations across years within plots and to allow us to test the hypothesis that our treatments alter the ANPP-soil moisture relationship.
-Thus, our multi-level model has three grouping levels for coefficients, representing the nested structure of the data: (i) overall coefficients, (ii) treatment coefficients, and (iii) plot coefficients.
-Each subsequent level is drawn from the distribution of coefficients at the previous level, meaning plot-level coefficients are drawn from treatment-level coefficients, which are drawn from overall coefficients.
-We also include a covariance structure among the intercept and slope at each level.
-Formally, our hierarchical structure is as follows, where we drop the intercept (0) and slope (1) subscripts and instead refer to a vector of coefficients, $\boldsymbol{\beta}$:
-
-\vspace{-1em}
-
-\begin{align}
-\boldsymbol{\beta}_{j(k)} &\sim \text{MVN} \left( \boldsymbol{\beta}_k, \boldsymbol{\Sigma}(k) \right), \\
-\boldsymbol{\beta}_{k} &\sim \text{MVN} \left( \boldsymbol{\beta}, \boldsymbol{\Sigma}  \right), \\
-\boldsymbol{\beta} &\sim \text{Normal} \left( 0, 1 \right),
-\end{align}
-
-\noindent{}where $\boldsymbol{\beta}_{j(k)}$ is the vector of regression coefficients (intercept and slope) for plot *j* associated with treatment *k*, $\boldsymbol{\beta}_{k}$ is the vector of coefficients for each treatment, and $\boldsymbol{\beta}$ is the vector of overall coefficients.
-The plot- and treatment-level coefficients are drawn from multivariate normal distributions with covariance matrix $\boldsymbol{\Sigma}$.
-For the plot-level coefficients, each treatment has its own variance-covariance matrix (i.e., $\boldsymbol{\Sigma}(k)$).
-The overall coefficients are drawn from a normal prior with mean 0 and standard deviation 1.
-The random year effects ($\boldsymbol{\gamma}$) are drawn from a normal prior with mean 0 and standard deviation $\sigma_{\text{year}}$, which was drawn from a weibull distribution.
-A full description of model is in Appendix 2.
----->
-
 We fit the model using a Bayesian approach and obtained posterior estimates of all unknowns via the No-U-Turn Hamiltonian Monte Carlo sampler in Stan [@stan2016].
@@
 We also include annotated Stan code in our model description in Appendix 2.
-
-<!---
-```{r get_results, echo=FALSE}
-betas <- read.csv("../results/beta_summaries.csv")
-```
---->
 
@@
 However, the response of ANPP to soil moisture in all treatments was consistently weak (Table 1, Fig. A2-2) and within-year variation of ANPP was large (Fig. 3B)
-
-<!---
-The results of our five year experiment in a sagebrush steppe conform to the first of our four predictions: neither ecosystem functional response nor community composition changed with chronic alteration in water availability (Fig.
-1, top left).
---->
 
@@
 First, species' biomass production may have hit physiological thresholds and/or become limited by other factors, such as nitrogen availability [@LeBauer2008].
-\pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES? WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY THIS IS...}
+\pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES? WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY OR UNLIKEL THIS IS...}
 Second, the less positive slope could be another indicator of bet hedging in this community.
@@
 
-<!---
-\newpage{}
-
-\begin{figure}[!ht]
-  \centering
-      \includegraphics[width=5in]{../figures/glmm_yeardiffs.png}
-  \caption{Posterior distributions of random year effects (intercept offsets). Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.}
-\end{figure}
---->
-
 \newpage{}

### 2017-08-25 manuscript/usses_water_ms.tex
removes all old text

@@
 \citep{LeBauer2008}.
-\pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES? WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY THIS IS...}
+\pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES? WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY OR UNLIKEL THIS IS...}
 Second, the less positive slope could be another indicator of bet
@@
 
-\bibliography{/Users/atredenn/Dropbox/Bibliography/usses_water.bib}
+\bibliography{~./usses_water.bib}
 

### 2017-08-25 manuscript/usses_water_ms.Rmd
fixes typos

@@
 In this case, changes in ANPP simply follow the soil moisture-ANPP relationship under ambient conditions.
-This corresponds to the early phases of the Hierarchical Respone Framework, where ecosystem response follows the physiological responses of individual species.
+This corresponds to the early phases of the Hierarchical Response Framework, where ecosystem response follows the physiological responses of individual species.
 Second, the ecosystem functional response changes but community composition remains the same (Fig. 1, top right).
@@
 \noindent{}where $\Sigma$ is the variance-covariance matrix and *j(i)* reads as "plot *j* associated with observation *i*". 
-The random year effects ($\boldsymbol{\eta}$) are drawn from a normal prior with mean 0 and standard deviation $\sigma_{\text{year}}$, which was drawn from a half-cauchy distribution. 
+The random year effects ($\boldsymbol{\eta}$) are drawn from a normal prior with mean 0 and standard deviation $\sigma_{\text{year}}$, which was drawn from a half-Cauchy distribution. 
 A full description of our model is in Appendix 2.
@@
 1.
-The altered ecosystem functional response under irrigation matched our expectations becuase we found evidence for a saturating response for the irrigation treatment (Fig. 3A, right panel).
+The altered ecosystem functional response under irrigation matched our expectations because we found evidence for a saturating response for the irrigation treatment (Fig. 3A, right panel).
 However, the response of ANPP to soil moisture in all treatments was consistently weak (Table 1, Fig. A2-2) and within-year variation of ANPP was large (Fig. 3B)
@@
       \includegraphics[width=4in]{../figures/hypothesis_figtable.png}
-  \caption{Possible outcomes of chronic resource alteration based on the 'Hierachical Response Framework' (Smith et al. 2009).}
+  \caption{Possible outcomes of chronic resource alteration based on the 'Hierarchical Response Framework' (Smith et al. 2009).}
 \end{figure}

### 2017-08-25 manuscript/usses_water_ms.tex
fixes typos

@@
 ambient conditions. This corresponds to the early phases of the
-Hierarchical Respone Framework, where ecosystem response follows the
+Hierarchical Response Framework, where ecosystem response follows the
 physiological responses of individual species. Second, the ecosystem
@@
 from a normal prior with mean 0 and standard deviation
-\(\sigma_{\text{year}}\), which was drawn from a half-cauchy
+\(\sigma_{\text{year}}\), which was drawn from a half-Cauchy
 distribution. A full description of our model is in Appendix 2.
@@
 Fig. 1. The altered ecosystem functional response under irrigation
-matched our expectations becuase we found evidence for a saturating
+matched our expectations because we found evidence for a saturating
 response for the irrigation treatment (Fig. 3A, right panel). However,

### 2017-08-25 manuscript/usses_water_ms.Rmd
final edits for coauthors

@@
 
-  **Methods.** We conducted a five year precipitation manipulation experiment in a sagebrush steppe ecosystem in Idaho, US.
+  **Methods.** We conducted a five year precipitation manipulation experiment in a sagebrush steppe ecosystem in Idaho, United States.
 We used drought and irrigation treatments (approximately 50% decrease/increase) to investigate whether ecosystem functional response remains consistent at extreme precipitation levels.
@@
 We fit a generalized linear mixed effects model to determine if the relationship between ANPP and soil moisture differed among treatments.
-We used nondimensional multivariate scaling to quantify community composition over the five years.
+We used nonmetric multidimensional scaling to quantify community composition over the five years.
 
@@
 
-We used non-dimensional multivariate scaling (NMDS) based on Bray-Curtis distances to see if community composition differed among treatments through time.
+We used nonmetric multidimensional scaling (NMDS) based on Bray-Curtis distances to see if community composition differed among treatments through time.
 We first calculated Bray-Curtis distances among all plots for each year of the experiment and then extracted those distances for use in the NMDS.
@@
 Such differences in soil VWC indicate our treatment infrastructure was successful.
+ANPP was highly variable within years (Fig. 2C).
 
@@
 The dry and variable environment of the sagebrush steppe has likely selected for bet hedging species that can maintain function at low water availability and have weak responses to high water availability.
-In so doing, the dominant species in our ecosystem avoid "boom and bust" cycles, which corresponds to the weak effect of soil moisture on ANPP (i.e., the Bayesian credible intervals for $\beta$ in Eq.
+In so doing, the dominant species in our ecosystem avoid "boom and bust" cycles, which corresponds to the weak effect of soil moisture on ANPP (i.e., the Bayesian credible intervals for the slopes in $\boldsymbol{\beta}$ in Eq.
 1 overlapping zero).
@@
 Second, the less positive slope could be another indicator of bet hedging in this community.
-Avoiding high biomass production in wet years could benefit species in the long run if climate conditions are variable.
+Avoiding high biomass and seed production in wet years could benefit species in the long run if climate conditions are variable.
 Third, the less positive slope may be a statistical artifact due the variability of ANPP within years and our low sample size.

### 2017-08-25 manuscript/usses_water_ms.tex
final edits for coauthors

@@
 \textbf{Methods.} We conducted a five year precipitation manipulation
-experiment in a sagebrush steppe ecosystem in Idaho, US. We used drought
-and irrigation treatments (approximately 50\% decrease/increase) to
-investigate whether ecosystem functional response remains consistent at
-extreme precipitation levels. We recorded data on aboveground net
-primary productivity (ANPP), species abundance, and soil moisture. We
-fit a generalized linear mixed effects model to determine if the
-relationship between ANPP and soil moisture differed among treatments.
-We used nondimensional multivariate scaling to quantify community
-composition over the five years.
+experiment in a sagebrush steppe ecosystem in Idaho, United States. We
+used drought and irrigation treatments (approximately 50\%
+decrease/increase) to investigate whether ecosystem functional response
+remains consistent at extreme precipitation levels. We recorded data on
+aboveground net primary productivity (ANPP), species abundance, and soil
+moisture. We fit a generalized linear mixed effects model to determine
+if the relationship between ANPP and soil moisture differed among
+treatments. We used nonmetric multidimensional scaling to quantify
+community composition over the five years.
 
@@
 
-We used non-dimensional multivariate scaling (NMDS) based on Bray-Curtis
+We used nonmetric multidimensional scaling (NMDS) based on Bray-Curtis
 distances to see if community composition differed among treatments
@@
 (VWC) differences among treatments (Fig. 2B). Such differences in soil
-VWC indicate our treatment infrastructure was successful.
+VWC indicate our treatment infrastructure was successful. ANPP was
+highly variable within years (Fig. 2C).
 
@@
 cycles, which corresponds to the weak effect of soil moisture on ANPP
-(i.e., the Bayesian credible intervals for \(\beta\) in Eq. 1
-overlapping zero).
+(i.e., the Bayesian credible intervals for the slopes in
+\(\boldsymbol{\beta}\) in Eq. 1 overlapping zero).
 
@@
 Second, the less positive slope could be another indicator of bet
-hedging in this community. Avoiding high biomass production in wet years
-could benefit species in the long run if climate conditions are
-variable. Third, the less positive slope may be a statistical artifact
-due the variability of ANPP within years and our low sample size.
+hedging in this community. Avoiding high biomass and seed production in
+wet years could benefit species in the long run if climate conditions
+are variable. Third, the less positive slope may be a statistical
+artifact due the variability of ANPP within years and our low sample
+size.
 

### 2017-08-25 manuscript/merge_pdfs.sh

@@ -1,0 +1,3 @@
+#!/bin/bash 
+
+"/System/Library/Automator/Combine PDF Pages.action/Contents/Resources/join.py" -o usses_water_ms_appendices.pdf usses_water_ms.pdf Appendix1.pdf Appendix2.pdf Appendix3.pdf

### 2017-08-26 manuscript/Appendix1.Rmd
updates all model related stuff to ARK specification
fixes table A1-2 column names

@@
 \textbf{\Large{Appendix 1}} \\
-A.T. Tredennick, A.R. Kleinhesselink, \& P.B. Adler \\
-``Ecosystem and community resistance...'' \\
+A.T. Tredennick, A.R. Kleinhesselink, B. Taylor \& P.B. Adler \\
+``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe''  \\
 \emph{PeerJ}
@@
 tab_dat <- read.csv("../data/estimated_biomass/ndvi_biomass_regression_parameters.csv", row.names = 1) %>%
-  filter(year > 2011)
-colnames(tab_dat) <- c("Year", "Intercept", "NDVI Slope", "$R^2$", "Min(biomass)", "Max(biomass)", "Algorithm")
+  filter(year > 2011 & year < 2017)
+colnames(tab_dat) <- c("Year", "Intercept", "NDVI Slope", "$R^2$", "Min(biomass)", "Max(biomass)", "\\emph{N}", "Algorithm")
 

### 2017-09-18 manuscript/usses_water_ms.Rmd
adds pba edits

@@
   - name: Andrew R. Kleinhesselink
-    affiliation: 2
+    affiliation: 1,2
   - name: Bret Taylor
@@
   **Background.** Precipitation is predicted to become more variable in the western United States, meaning years of above and below average precipitation will become more common.
-Periods of extreme precipitation are major drivers of ecosystem functioning in water limited grasslands, but how ecosystems respond to precipitation may change as the duration of above and below average periods increases.
-Changes in ecosystem functional response could reflect compensatory changes in species composition or species reaching physiological thresholds at extreme precipitation levels.
+Periods of extreme precipitation are major drivers of interannual variability in ecosystem functioning in water limited communities, but how ecosystems respond to these extremes over the long-term may shift with precipitation means and variances.
+Long-term changes in ecosystem functional response could reflect compensatory changes in species composition or species reaching physiological thresholds at extreme precipitation levels.
 
   **Methods.** We conducted a five year precipitation manipulation experiment in a sagebrush steppe ecosystem in Idaho, United States.
-We used drought and irrigation treatments (approximately 50% decrease/increase) to investigate whether ecosystem functional response remains consistent at extreme precipitation levels.
+We used drought and irrigation treatments (approximately 50% decrease/increase) to investigate whether ecosystem functional response remains consistent under sustained high or low precipitation.
 We recorded data on aboveground net primary productivity (ANPP), species abundance, and soil moisture.
@@
 However, ANPP response to availabile soil moisture was weak and uncertain regardless of treatment, with all slopes overlapping zero.
-There was also large variation in ANPP within-years.
+There was also large spatial variation in ANPP within-years.
 Plant community composition was remarkably stable over the course of the experiment and did not differ among treatments.
 
-  **Discussion.** Despite some evidence that ecosystem functional response was less positive under extreme wet conditions, the response of ANPP to soil moisture was consistently weak and community composition was remarkably stable.
-Ecosystem functional responses across treatments was not related to compensatory shifts at the plant community level, but instead reflects the weak and variable responses of individual species.
-Such responses may be due to bet hedging strategies where species respond weakly to even extreme levels of soil moisture to ensure low variance of long-term success.
+  **Discussion.** Despite some evidence that ecosystem functional response became less sensitive under sustained wet conditions, the response of ANPP to soil moisture was consistently weak and community composition was stable.
+Differences in ecosystem functional responses across treatments were not related to compensatory shifts at the plant community level, but instead may reflect the insensitivity of the dominant species to soil moisture.
+These species may be successful precisely because they have evolved life history strategies which buffer them against precipitation variability.
 header-includes: >
@@
 
-The functional response of aboveground net primary productivity (ANPP) to water availability (e.g., soil moisture) can be characterized by fitting a model to historical observations of ANPP and soil moisture at a given site.
-However, the fitted functional response may provide an incomplete picture because future conditions are likely to be outside the historical range of variability [@Smith2011].
-For example, historical trends may underestimate the potential for the soil moisture-ANPP relationship to saturate or weaken if soil moisture is pushed far beyond typical levels.
-Saturating relationships are actually common [@Hsu2012; @Gherardi2015a], perhaps because other resources, like nitrogen, become more limiting in wet years than dry years.
-Knowing the curvature of the soil moisture-ANPP relationship at extreme precipitation levels is critical for understanding how ecosystems will respond to chronic alterations in water availability.
+At any given site, the relationship between aboveground net primary productivity (ANPP) and water availability (e.g., soil moisture) can be characterized by regressing historical observations of ANPP on observations of soil moisture.
+The fitted functional response can then be used to project ANPP under future precipitation regimes (give example REF).
+A problem with this approach is that it requires extrapolation if future precipitation falls outside outside the historical range of variability [@Smith2011].
+For example, the soil moisture-ANPP relationship may be linear within the historical range of interannual variation, but could saturate at higher levels of soil moisture.
+In fact, saturating relationships are actually common [@Hsu2012; @Gherardi2015a], perhaps because other resources, like nitrogen, become more limiting in wet years than dry years.
+Failure to accurately estimate the curvature of the soil moisture-ANPP relationship will lead to over- or underprediction of ANNP under extreme precipitation.
 
-Another problem with relying on historical ecosystem functional responses is that they are not static.
-Changes in species identities and abundances can alter an ecosystem's functional response to water availability because different species have different physiological thresholds for producing biomass.
-@Smith2009 introduced the 'Hierarchical Response Framework' for understanding the interplay of community composition and ecosystem functioning in response to resource manipulations over time.
+Another problem with relying on historical ecosystem functional responses to predict impacts of altered precipitation regimes is that these relationships themselves might shift over the long-term.
+Changes in species identities and abundances can alter an ecosystem's functional response to water availability because different species have different physiological thresholds.
+@Smith2009 introduced the 'Hierarchical Response Framework' for understanding the interplay of community composition and ecosystem functioning in response to long-term shifts in resources.
 In the near term, ecosystem functioning such as ANPP will reflect the physiological responses of individual species to the manipulated resource level.
@@
 Ecosystem functioning may recover over longer time spans as new species colonize or initial species reorder in relative abundance.
-For example, ANPP may initially decline, but eventually rise back to pre-treatment levels once drought-tolerant species become more abundant and compensate for drought-intolerant species [@Hoover2014].
 It is also possible that ecosystem functioning shifts to a new mean state, reflecting the suite of species in the new community [@Knapp2012].
 
-Manipulating potentially limiting resources, like precipitation, offers a route to understanding how ecosystems will respond to resource levels that fall outside the historical range of variability [@Avolio2015; @Gherardi2015; @Knapp2017].
+Experimental manipulations of limiting resources, like precipitation, offer a route to understanding how ecosystems will respond to resource levels that fall outside the historical range of variability [@Avolio2015; @Gherardi2015; @Knapp2017].
 Altering the amount of precipitation over many years should provide insight into the time scales at which water-limited ecosystems respond to chronic resource alteration.
@@
 
-First, altered precipitation changes neither ecosystem functional response nor community composition (Fig. 1, top left).
-In this case, changes in ANPP simply follow the soil moisture-ANPP relationship under ambient conditions.
+First, altered precipitation might have no effect on either ecosystem functional response or community composition (Fig. 1, top left).
+In this case, changes in ANPP would be well predicted by the current, observed soil moisture-ANPP relationship.
 This corresponds to the early phases of the Hierarchical Response Framework, where ecosystem response follows the physiological responses of individual species.
-Second, the ecosystem functional response changes but community composition remains the same (Fig. 1, top right).
+Second, the ecosystem functional response might change while community composition remains the same (Fig. 1, top right).
 A saturating soil moisture-ANPP response fits this scenario, where individual species hit physiological thresholds or are limited by some other resource.
-Third, the ecosystem functional response is consistent but underlying community composition changes (Fig. 1, bottom left).
+Third, the ecosystem functional response might be constant but community composition changes (Fig. 1, bottom left).
 In this case, changes in species' identities or abundances occur in response to altered precipitation levels and species more suited to the new conditions compensate for reduced function of initial species.
-Fourth, and last, both ecosystem functional response and community composition change (Fig. 1, bottom right).
+Fourth, and last, both ecosystem functional response and community composition could change (Fig. 1, bottom right).
 New species, or newly abundant species, with different physiological responses completely reshape the ecosystem functional response.
@@
 All four outcomes are possible in any given ecosystem, but the time scales at which the different scenarios play out likely differ [@Smith2009; @Wilcox2016; @Knapp2017].
-Thus, our task is not to test the validity of the Hierarchical Response Framework, but rather to amass information on how quickly ecosystem functional responses change in different ecosystems.
+To determine these time scales, we need to amass information on how quickly ecosystem functional responses change in different ecosystems.
 We also need to understand whether changes at the ecosystem level are driven by community level changes or individual level responses.
@@
 We imposed drought and irrigation treatments (approximately $\pm50\%$) and measured ecosystem (ANPP) and community (species composition) responses.
-We focus on how the drought and irrigation treatments affect the relationship between available soil moisture and ANPP, and if community dynamics underlie the ecosystem responses.
+We focus on how the drought and irrigation treatments affect the relationship between interannual variation in available soil moisture and interannual variation in ANPP, and if community dynamics underlie the ecosystem responses.
 In particular, we are interested in the consistency of the soil moisture-ANPP relationship among treatments.
@@
 ## Study Area
-We conducted our precipitation manipulation experiment in a sagebrush steppe community at the United States Sheep Experimental Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
+We conducted our precipitation manipulation experiment in a sagebrush steppe community at the USDA-ARS Sheep Experimental Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
 The plant community is dominated by the shrub *Artemesia tripartita* and three perennial bunchgrasses, *Pseudoroegneria spicata*, *Poa secunda*, and *Hesperostipa comata*.
@@
 Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
-In 2007, we (well, one of us [P. Adler]) relocated 14 of the original quadrats, six of which were inside a large, permanent livestock exclosure.
-We use these six plots as control plots that have received no treatment, just ambient precipitation, in the experiment described below.
+In 2007, we relocated 14 of the original quadrats, six of which were inside a large, permanent livestock exclosure.
+We use these six plots as control plots (i.e. ambient precipitation) in the experiment described below.
 
 ## Precipitation Experiment
-In spring 2011, we (well, two of us [A. Kleinhesselink and P. Adler]) established 16 new 1 m$^2$ plots located in the same exclosure as the six control plots.
+In spring 2011, we established 16 new 1 m$^2$ plots located in the same exclosure as the six control plots.
 We avoided areas on steep hill slopes, areas with greater than 20\% cover of bare rock, and areas with greater than 10\% cover of the shrubs *Purshia tridentata* and/or *Amelanchier utahensis*.
@@
 Pumping was triggered by float switches once water levels reached about 20 liters.
-We disconnected the irrigation pumps each fall and reconnected them, often with difficulty, each spring.
+We disconnected the irrigation pumps each October and reconnected them each April.
 The rain-out shelters remained in place throughout the year.
@@
 We monitored soil moisture in four of the drought-irrigation pairs using Decagon Devices (Pullman, Washington) 5TM and EC-5 soil moisture sensors.
-We installed four sensors around the edges of each plot, two at 5 cm soil depth and two at 25 cm soil depth.
+We installed four sensors around the edges of each 1x1 m census plot, two at 5 cm soil depth and two at 25 cm soil depth.
 We also installed four sensors in areas nearby the four selected plot pairs to measure ambient soil moisture at the same depths.
@@
 Only a subset of plots were equipped with soil moisture sensors, and within those plots, one or more of the sensors frequently failed to collect data.
-To remedy these problems, and to produce average daily soil moisture values for the ambient, drought, and irrigation conditions, we used a statistical model to predict the average treatment effects on soil moisture during the course of the experiment.
+To remedy these problems, and to produce average daily soil moisture values for the ambient, drought, and irrigation treatments, we used a statistical model to predict the average treatment effects on soil moisture during the course of the experiment.
 
@@
 We could only predict soil moisture in the treated plots on days for which we took at least one ambient soil moisture measurement.
+<!--- If we aren't including the code and results for this model, you should
+include a reference, probably to Andy's dissertation--->
 
@@
 At each plot in each year, we took four readings of ground reflectances at the above wavelengths.
-We also took readings in 12 (2015), 15 (2012, 2013, 2014), or 16 (2016) calibration plots adjacent to the experimental site, in which we harvested all aboveground biomass, dried it to a constant weight at 60$^{\circ}$C, and weighed it to estimate ANPP.
+We also took readings in 12 (2015), 15 (2012, 2013, 2014), or 16 (2016) calibration plots adjacent to the experimental site, in which we harvested all aboveground biomass produced in the current year (we excluded litter and standing dead material), dried it to a constant weight at 60$^{\circ}$C, and weighed it to estimate ANPP.
+We harvested at peak biomass each year, typically in late June.
 
-For each plot and year, we averaged the four readings for each wavelength and then calculated a greenness index based on the same bands used to calculate NDVI using the MODIS and AVHRR bands for NDVI.
+For each plot and year, we averaged the four readings for each wavelength and then calculated a greenness index based on the same bands used to calculate NDVI using the MODIS and AVHRR [YOU HAVEN'T DEFINED THESE] bands for NDVI.
 We regressed the greenness index against the dry biomass weight from the ten calibration plots to convert the greenness index to ANPP.
@@
 Species composition data came from two sources: yearly census maps for each plot made using a pantograph [@Hill1920] and yearly counts of annual species in each plot.
-We extracted the mapped information to obtain data on the density of all annuals and perennials forbs, the basal cover of perennial grasses, and the canopy cover of shrubs.
+From these sources, we determined the density of all annuals and perennials forbs, the basal cover of perennial grasses, and the canopy cover of shrubs.
 We made a large plot-treatment-year by species matrix, where columns were filled with either basal cover or density, depending on the measurement made for the particular species.
-We standardized the values in each column so we could analyze the different types of data together.
+We standardized the values in each column so we could directly compare species quantified with different metrics (density, basal cover, and canopy cover).
 This puts all abundance values on the same scale, meaning that common and rare species are weighted equally.
-This approach is anti-conservative for detecting community change if we assume that rare species will respond more than common ones, which biases our approach toward detecting compositional changes.
+Assuming that rare species will respond to treatments more than common ones, our approach is biased towards detecting compositional changes.
 
@@
 Thus, our model treats "control" observations as the main treatment and then estimates intercept and slope offsets for the "drought" and "irrigation" treatments. 
-In reference to our model, the hypotheses we wish to test are:
+We use our model to test the following hypotheses:
 
@@
 \end{description}
+
+These hypotheses are based on evidence that precipitation-ANPP relationships often saturate with increasing precipitation (REFERENCES).
+<!---these should probably be stated in the Intro--->
 
@@
 A full description of our model is in Appendix 2.
+<!---call out available computer code here?--->
 
@@
 
-We used nonmetric multidimensional scaling (NMDS) based on Bray-Curtis distances to see if community composition differed among treatments through time.
+We used nonmetric multidimensional scaling (NMDS) based on Bray-Curtis distances to identify temporal changes in community composition among treatments.
 We first calculated Bray-Curtis distances among all plots for each year of the experiment and then extracted those distances for use in the NMDS.
@@
 
-Three of our five treatment years fell in years of below average rainfall (Fig. 2A).
-Thus, those three years represent a lower magnitude of absolute change in precipitation experienced by the treatments.
+Three of our five study years received below average precipitation (Fig. 2A).
+Because our treatments are proportional changes in ambient precipitation, those three years represent a lower magnitude of absolute difference in precipitation among the treatments.
+<!---why is the previous point important here? should it go in Methods?--->
 ANPP varied from a minimum of 74.5 g m$^{-2}$ in 2014 to a maximum of 237.1 g m$^{-2}$ in 2016 when averaged across treatments (Fig. 2C).
@@
 Such differences in soil VWC indicate our treatment infrastructure was successful.
-ANPP was highly variable within years (Fig. 2C).
+ANPP was highly variable across plots within years (Fig. 2C).
 
@@
 Community composition was similar among treatments.
-In no year did community composition among treatments not overlap, and they were equally dispersed in all years (Table 2; Fig. 5).
+Community composition among treatments overlapped in all years and was equally dispersed in all years (Table 2; Fig. 5).
 Community composition was also remarkably stable over time, with no evidence of divergence among treatments (Table 2; Fig. 5).
@@
 Previous work has shown that community compositional shifts can be both rapid, on the order of years [@Hoover2014], and slow, on order of decades [@Knapp2012; @Wilcox2016].
-Thus, a lingering question is how the time scales of ecosystem response and community change differ among ecosystems, which can be answered by manipulating precipitation to reach extreme levels.
+A lingering question is how the time scales of ecosystem response and community change vary among ecosystems.
+Precipitation manipulation experiments can help answer this question, especially if they push water availability outside the historical range of variability for long periods.
 
 The results of our five year experiment in a sagebrush steppe conform to two of our four predictions, depending on treatment.
-Neither ecosystem functional response nor community composition changed under chronic drought (Fig. 3A, Fig. 4), consistent with the top left prediction in Fig. 1.
+Neither ecosystem functional response nor community composition changed under chronic drought (Fig. 3A, Fig. 4), representin the top left SCENARIO in Fig. 1.
 Ecosystem functional response under chronic irrigation was different from the control treatment, but community composition remained unchanged (Fig.
 3A, Fig.
-4), consistent with the top right prediction in Fig.
+4), representing the top right SCENARIO in Fig.
 1.
-The altered ecosystem functional response under irrigation matched our expectations because we found evidence for a saturating response for the irrigation treatment (Fig. 3A, right panel).
-However, the response of ANPP to soil moisture in all treatments was consistently weak (Table 1, Fig. A2-2) and within-year variation of ANPP was large (Fig. 3B)
+The altered ecosystem functional response under irrigation matched our expectations because we found evidence for a saturating response for the irrigation treatment (Fig. 3).
+<!--- Can we really call it saturating if it is concave up? More accurate to say weaker response than control? --->
+However, the response of ANPP to soil moisture in all treatments was consistently weak (Table 1, Fig. A2-2) and within-year variation of ANPP was large (Fig. 3B).
+<!---Last sentence leaves me hanging...what's your point? Did it the response change or not? Statistically but not ecologically significant?--->
 
@@
 Thus, our drought treatment represented extreme precipitation amounts, especially in years where ambient precipitation was below average (Fig. 2A).
-The irrigation treatment may have been too small, yet that is the treatment where we did observe an effect (Fig. 3).
+The irrigation treatment may not qualify as extreme, yet that is the treatment where we did observe an effect (Fig. 3).
 
-Second, ANPP at our site may be influenced by factors beyond the window of soil moisture (cumulative March-June) we included in our statistical model.
+Second, ANPP at our site may be influenced by additional factors, not only the cumulative March-June soil moisture covariate we included in our statistical model.
 For example, temperature can impact ANPP directly [@Epstein1997] and by exacerbating the effects of soil moisture [@DeBoeck2011].
@@
 We also did not redistribute snow across our treatments in the winter, and snow melt may spur early spring growth.
-These statistical issues of missing potentially important covariates could explain the within-year spread of ANPP (Fig 2C, Fig.
+Failure to account for potentially important covariates could explain the within-year spread of ANPP (Fig 2C, Fig.
 3B) and the resulting uncertain relationship we observed between soil moisture and ANPP across all treatments (Table 1, Fig.
 A2-2).
+<!---so why didn't we do this? sample size limitations, power? --->
 
@@
 One strategy is bet hedging, where species forego short-term gains to reduce the variance of long-term success [@Seger1987].
-In other words, species do the same thing every year, with only minimal response to environmental conditions.
+In other words, species follow the same conservative strategy every year, designed to minimize response to environmental conditions.
 The dry and variable environment of the sagebrush steppe has likely selected for bet hedging species that can maintain function at low water availability and have weak responses to high water availability.
@@
 Another strategy to deal with variable environmental conditions is avoidance, which would also result in a consistent ecosystem functional response between drought and control treatments.
-The perennial plants in this cold desert ecosystem are tolerant to drought conditions [@Bazzaz1979; @Franks2011, A.R. Kleinhesselink, unpublished data].
 For example, many of the perennial grasses in our focal ecosystem avoid drought stress by growing early in the growing season [@Blaisdell1958, A.R. Kleinhesselink, personal observation].
@@
 
-The weaker soil moisture-ANPP relationship we observed for the irrigation treatment, evidenced by the less positive slope relative to controls (Fig.
-3A, right panel), has three non-exclusive explanations.
+The weaker soil moisture-ANPP relationship we observed for the irrigation treatment has three non-exclusive explanations.
 First, species' biomass production may have hit physiological thresholds and/or become limited by other factors, such as nitrogen availability [@LeBauer2008].
-\pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES? WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY OR UNLIKEL THIS IS...}
-Second, the less positive slope could be another indicator of bet hedging in this community.
+\pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES?
+WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY OR UNLIKEL THIS IS...NUTNET SHOWS POSITIVE EFFECTS OF N*P ON PRODUCTION, EFFECT OF N ALONE NOT AS CLEAR.
+YOU COULD CITE THIS AS UNPUBLISHED DATA}
+
+Second, the less positive slope could be an indicator of bet hedging in this community.
 Avoiding high biomass and seed production in wet years could benefit species in the long run if climate conditions are variable.
+\pba{but why would this have cause a change relative to controls? this should also be true for controls...}
 Third, the less positive slope may be a statistical artifact due the variability of ANPP within years and our low sample size.
 
-In conclusion, our results suggest the species in our focal plant community are tolerant of drought conditions and bet hedgers in wet conditions.
-Longer time series of chronic precipitation alteration may reveal plant community shifts that we did not observe [e.g., @Wilcox2016], in which case species that do not bet hedge may gain prominence and dominate the ecosystem functional response.
+In conclusion, our results suggest the species in our focal plant community are tolerant of drought conditions and unresponsive to wet conditions.
+\pba{no, you can't be a bet hedger in wet years but not in dry years...}
+Longer, chronic precipitation alteration might reveal plant community shifts that we did not observe [e.g., @Wilcox2016].
+For example, a long-term increase in water availability could allow species that do not bet hedge to gain prominence and dominate the ecosystem functional response.
 Our results suggest compositional shifts would have the largest impact at high rainfall because the current community maintained consistent ecosystem functional response at very low water availability.

### 2017-09-19 manuscript/usses_water_ms.Rmd

--- 2017-08-25 manuscript/usses_water_ms.Rmd
@@
   \caption{(A) Probability density of historical precipitation from 1926-2016, with the years of the experiment shown with arrows on the \emph{x}-axis.
-(B) Observed soil volumetric water content (VWC) over the course of the experiment.
-(C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.}
+(B) Mean soil volumetric water content (VWC) in each of the three treatments during the course of the experiment.
+(C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.
+**\emph{note plot b doesn't have a treatment legend, consider adding to the side of (B) and removing from (C)} }
 \end{figure}
@@
   \caption{Results from the generalized linear mixed effects model.
-(A) Posterior distributions of the intercept and slope offsets for the drought and irrigation treatments.
-Offsets indicate the amount to which the coefficients for drought or irrigation treatments differ from the control treatment estimates.
-Probabilities (``Pr $<$ 0 ='') for each distribution indicate the probability that coefficient is less than zero.
-Probabilities greater than 0.95 indicate strong support for the coefficient being less than zero.
+(A) Posterior distributions for the effects of drought and irrigation on the intercept and slope of the productivity-soil moisture relationship.
+Treatment effects show the difference between the coefficients estimated in the treated plots and the control plots.
+Probabilities (``Pr $<$ 0 ='') for each coefficient indicate the probability that the coefficient is less than zero.
 We only show the one-tailed probability for the value being less than zero because the median of each distribution is less than zero.
-Kernel bandwidths of posterior densities were adjusted by a factor of 5 to smooth the distribution for visual clarity.
-(B) Scatterplot of the data and model estimates shown a solid lines.
+The posterior densities were smoothed for visual clarity by increasing kernal bandwidth by a factor of five.
+(B) Scatterplot of the data and model estimates shown as solid lines.
 Model estimates come from treatment level coefficients (colored lines).
@@
       \includegraphics[width=5in]{../figures/sppcomp_bray_all.png}
-  \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot, colored by treatment.}
+  \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot, colored by treatment.
+**\emph{Worth adding arrow vectors showing association between axes and dominant species?}}
 \end{figure}

### 2017-09-19 manuscript/usses_water_ms.Rmd
adds most of JBT edits; working on PBA edits; updates Fig 3 probability sign

@@
     affiliation: 1,2
-  - name: Bret Taylor
+  - name: J. Bret Taylor
     affiliation: 3
@@
   - code: 3
-    address: United States Department of Agriculture, Agriculture Research Station, U.S. Sheep Experiment Station, Dubois, Idaho 83423
+    address: United States Department of Agriculture, Agricultural Research Service, U.S. Sheep Experiment Station, Dubois, Idaho 83423
 abstract: |
@@
   \usepackage{bm,mathrsfs}
-bibliography: ~./usses_water.bib
+bibliography: /Users/atredenn/Dropbox/Bibliography/usses_water.bib
 output: 
@@
 At any given site, the relationship between aboveground net primary productivity (ANPP) and water availability (e.g., soil moisture) can be characterized by regressing historical observations of ANPP on observations of soil moisture.
-The fitted functional response can then be used to project ANPP under future precipitation regimes (give example REF).
-A problem with this approach is that it requires extrapolation if future precipitation falls outside outside the historical range of variability [@Smith2011].
+The fitted functional response can then be used to project ANPP under future precipitation regimes (REFERENCE). 
+A problem with this approach is that it requires extrapolation if future precipitation falls outside outside the historical range of variability [@Smith2011; @Peters2012].
 For example, the soil moisture-ANPP relationship may be linear within the historical range of interannual variation, but could saturate at higher levels of soil moisture. 
 In fact, saturating relationships are actually common [@Hsu2012; @Gherardi2015a], perhaps because other resources, like nitrogen, become more limiting in wet years than dry years.
-Failure to accurately estimate the curvature of the soil moisture-ANPP relationship will lead to over- or underprediction of ANNP under extreme precipitation.
+Failure to accurately estimate the curvature of the soil moisture-ANPP relationship will lead to over- or underprediction of ANNP under extreme precipitation [@Peters2012].
 
 Another problem with relying on historical ecosystem functional responses to predict impacts of altered precipitation regimes is that these relationships themselves might shift over the long-term.
-Changes in species identities and abundances can alter an ecosystem's functional response to water availability because different species have different physiological thresholds.
+Shifts in species identities and/or abundances can alter an ecosystem's functional response to water availability because different species have different physiological thresholds.
 @Smith2009 introduced the 'Hierarchical Response Framework' for understanding the interplay of community composition and ecosystem functioning in response to long-term shifts in resources.
@@
 Altering the amount of precipitation over many years should provide insight into the time scales at which water-limited ecosystems respond to chronic resource alteration.
-We propose four alternative predictions for the effect of precipitation manipulation on the ecosystem functional response to soil moisture based on the Hierarchical Response Framework (Fig. 1).
+We propose four alternative scenarios for the effect of precipitation manipulation on the ecosystem functional response to soil moisture based on the Hierarchical Response Framework (Fig. 1).
 We define 'ecosystem functional response' as the relationship between available soil moisture and ANPP.
-The four predictions are based on possible outcomes at the community (e.g., community composition) and ecosystem (e.g., soil moisture-ANPP regression) levels.
+The four scenarios are based on possible outcomes at the community (e.g., community composition) and ecosystem (e.g., soil moisture-ANPP regression) levels.
 
@@
 Third, the ecosystem functional response might be constant but community composition changes (Fig. 1, bottom left).
-In this case, changes in species' identities or abundances occur in response to altered precipitation levels and species more suited to the new conditions compensate for reduced function of initial species.
+In this case, changes in species' identities and/or abundances occur in response to altered precipitation levels and species more suited to the new conditions compensate for reduced function of initial species.
 Fourth, and last, both ecosystem functional response and community composition could change (Fig. 1, bottom right).
@@
 In particular, we are interested in the consistency of the soil moisture-ANPP relationship among treatments.
-Is the relationship steeper under the drought treatment, at low soil moisture?
-Does the relationship saturate under the irrigation treatment, at high soil moisture?
+Is the relationship steeper under the drought treatment at low soil moisture?
+Does the relationship saturate under the irrigation treatment at high soil moisture?
 To answer these questions we fit a generalized linear mixed effects model to test whether the regressions differed among treatments.
-We also analyzed community composition over time, allowing us to place our experimental results within the framework our competing predictions (Fig. 1).
+We also analyzed community composition over time, allowing us to place our experimental results within the framework our scenarios (Fig. 1).
 
@@
 ## Study Area
-We conducted our precipitation manipulation experiment in a sagebrush steppe community at the USDA-ARS Sheep Experimental Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
+We conducted our precipitation manipulation experiment in a sagebrush steppe community at the USDA, U.S.
+Sheep Experiment Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
 The plant community is dominated by the shrub *Artemesia tripartita* and three perennial bunchgrasses, *Pseudoroegneria spicata*, *Poa secunda*, and *Hesperostipa comata*.
@@
 
-## Precipitation Experiment
 In spring 2011, we established 16 new 1 m$^2$ plots located in the same exclosure as the six control plots.
@@
 We established the new plots in pairs and randomly assigned each plot in a pair to receive a "drought" or "irrigation" treatment.
+
+## Precipitation Experiment
 
@@
 We could only predict soil moisture in the treated plots on days for which we took at least one ambient soil moisture measurement.
-<!--- If we aren't including the code and results for this model, you should
-include a reference, probably to Andy's dissertation--->
+See @Kleinhesselink2017b for complete details on our approach to estimating soil moisture.
 
@@
 
-We thank the many summer research technicians who collected the data reported in this paper and the US Experimental Sheep Station for facilitating work on their property.
+We thank the many summer research technicians who collected the data reported in this paper and the USDA, U.S. Sheep Experiment Station (Dubois, ID) for facilitating work on their property.
 We also thank Susan Durham for clarifying our thinking on the statistical analyses and Kevin Wilcox for helpful discussions on analyzing community composition data.
@@
   \item Andrew R. Kleinhesselink conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed drafts of the paper.
-  \item Bret Taylor contributed reagents/materials/analysis tools, reviewed drafts of the paper.
+  \item J. Bret Taylor contributed reagents/materials/analysis tools, reviewed drafts of the paper.
   \item Peter B. Adler conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed drafts of the paper.
@@
 Treatment effects show the difference between the coefficients estimated in the treated plots and the control plots.
-Probabilities (``Pr $<$ 0 ='') for each coefficient indicate the probability that the coefficient is less than zero.
-We only show the one-tailed probability for the value being less than zero because the median of each distribution is less than zero.
+Probabilities (``Pr $\lessgtr$ 0 ='') for each coefficient indicate the one-tailed probability that the coefficient is less than or greater than zero, depending on whether the median of the distribution is less than or greater than zero.
 The posterior densities were smoothed for visual clarity by increasing kernal bandwidth by a factor of five.

### 2017-09-19 manuscript/usses_water_ms.tex
adds most of JBT edits; working on PBA edits; updates Fig 3 probability sign

@@
 \corrauthor[1]{Andrew T. Tredennick}{\href{mailto:atredenn@gmail.com}{\nolinkurl{atredenn@gmail.com}}}
-\author[2]{Andrew R. Kleinhesselink}
+\author[1,2]{Andrew R. Kleinhesselink}
 
-\author[3]{Bret Taylor}
+\author[3]{J. Bret Taylor}
 
@@
 California, Los Angeles, Los Angeles, California 90095}
-\affil[3]{United States Department of Agriculture, Agriculture Research Station,
+\affil[3]{United States Department of Agriculture, Agricultural Research Service,
 U.S. Sheep Experiment Station, Dubois, Idaho 83423}
@@
 precipitation will become more common. Periods of extreme precipitation
-are major drivers of ecosystem functioning in water limited grasslands,
-but how ecosystems respond to precipitation may change as the duration
-of above and below average periods increases. Changes in ecosystem
-functional response could reflect compensatory changes in species
-composition or species reaching physiological thresholds at extreme
-precipitation levels.
+are major drivers of interannual variability in ecosystem functioning in
+water limited communities, but how ecosystems respond to these extremes
+over the long-term may shift with precipitation means and variances.
+Long-term changes in ecosystem functional response could reflect
+compensatory changes in species composition or species reaching
+physiological thresholds at extreme precipitation levels.
 
@@
 decrease/increase) to investigate whether ecosystem functional response
-remains consistent at extreme precipitation levels. We recorded data on
-aboveground net primary productivity (ANPP), species abundance, and soil
-moisture. We fit a generalized linear mixed effects model to determine
-if the relationship between ANPP and soil moisture differed among
-treatments. We used nonmetric multidimensional scaling to quantify
-community composition over the five years.
+remains consistent under sustained high or low precipitation. We
+recorded data on aboveground net primary productivity (ANPP), species
+abundance, and soil moisture. We fit a generalized linear mixed effects
+model to determine if the relationship between ANPP and soil moisture
+differed among treatments. We used nonmetric multidimensional scaling to
+quantify community composition over the five years.
 
@@
 moisture was weak and uncertain regardless of treatment, with all slopes
-overlapping zero. There was also large variation in ANPP within-years.
-Plant community composition was remarkably stable over the course of the
-experiment and did not differ among treatments.
+overlapping zero. There was also large spatial variation in ANPP
+within-years. Plant community composition was remarkably stable over the
+course of the experiment and did not differ among treatments.
 
 \textbf{Discussion.} Despite some evidence that ecosystem functional
-response was less positive under extreme wet conditions, the response of
-ANPP to soil moisture was consistently weak and community composition
-was remarkably stable. Ecosystem functional responses across treatments
-was not related to compensatory shifts at the plant community level, but
-instead reflects the weak and variable responses of individual species.
-Such responses may be due to bet hedging strategies where species
-respond weakly to even extreme levels of soil moisture to ensure low
-variance of long-term success.
+response became less sensitive under sustained wet conditions, the
+response of ANPP to soil moisture was consistently weak and community
+composition was stable. Differences in ecosystem functional responses
+across treatments were not related to compensatory shifts at the plant
+community level, but instead may reflect the insensitivity of the
+dominant species to soil moisture. These species may be successful
+precisely because they have evolved life history strategies which buffer
+them against precipitation variability.
 % Dummy abstract text.
@@
 
-The functional response of aboveground net primary productivity (ANPP)
-to water availability (e.g., soil moisture) can be characterized by
-fitting a model to historical observations of ANPP and soil moisture at
-a given site. However, the fitted functional response may provide an
-incomplete picture because future conditions are likely to be outside
-the historical range of variability \citep{Smith2011}. For example,
-historical trends may underestimate the potential for the soil
-moisture-ANPP relationship to saturate or weaken if soil moisture is
-pushed far beyond typical levels. Saturating relationships are actually
-common \citep{Hsu2012, Gherardi2015a}, perhaps because other resources,
-like nitrogen, become more limiting in wet years than dry years. Knowing
-the curvature of the soil moisture-ANPP relationship at extreme
-precipitation levels is critical for understanding how ecosystems will
-respond to chronic alterations in water availability.
+At any given site, the relationship between aboveground net primary
+productivity (ANPP) and water availability (e.g., soil moisture) can be
+characterized by regressing historical observations of ANPP on
+observations of soil moisture. The fitted functional response can then
+be used to project ANPP under future precipitation regimes (REFERENCE).
+A problem with this approach is that it requires extrapolation if future
+precipitation falls outside outside the historical range of variability
+\citep{Smith2011, Peters2012}. For example, the soil moisture-ANPP
+relationship may be linear within the historical range of interannual
+variation, but could saturate at higher levels of soil moisture. In
+fact, saturating relationships are actually common
+\citep{Hsu2012, Gherardi2015a}, perhaps because other resources, like
+nitrogen, become more limiting in wet years than dry years. Failure to
+accurately estimate the curvature of the soil moisture-ANPP relationship
+will lead to over- or underprediction of ANNP under extreme
+precipitation \citep{Peters2012}.
 
 Another problem with relying on historical ecosystem functional
-responses is that they are not static. Changes in species identities and
-abundances can alter an ecosystem's functional response to water
-availability because different species have different physiological
-thresholds for producing biomass. \citet{Smith2009} introduced the
-`Hierarchical Response Framework' for understanding the interplay of
-community composition and ecosystem functioning in response to resource
-manipulations over time. In the near term, ecosystem functioning such as
-ANPP will reflect the physiological responses of individual species to
-the manipulated resource level. For example, ANPP may decline under
+responses to predict impacts of altered precipitation regimes is that
+these relationships themselves might shift over the long-term. Shifts in
+species identities and/or abundances can alter an ecosystem's functional
+response to water availability because different species have different
+physiological thresholds. \citet{Smith2009} introduced the `Hierarchical
+Response Framework' for understanding the interplay of community
+composition and ecosystem functioning in response to long-term shifts in
+resources. In the near term, ecosystem functioning such as ANPP will
+reflect the physiological responses of individual species to the
+manipulated resource level. For example, ANPP may decline under
 simulated drought because the initial community consisted of
@@
 recover over longer time spans as new species colonize or initial
-species reorder in relative abundance. For example, ANPP may initially
-decline, but eventually rise back to pre-treatment levels once
-drought-tolerant species become more abundant and compensate for
-drought-intolerant species \citep{Hoover2014}. It is also possible that
+species reorder in relative abundance. It is also possible that
 ecosystem functioning shifts to a new mean state, reflecting the suite
@@
 
-Manipulating potentially limiting resources, like precipitation, offers
-a route to understanding how ecosystems will respond to resource levels
-that fall outside the historical range of variability
+Experimental manipulations of limiting resources, like precipitation,
+offer a route to understanding how ecosystems will respond to resource
+levels that fall outside the historical range of variability
 \citep{Avolio2015, Gherardi2015, Knapp2017}. Altering the amount of
@@
 scales at which water-limited ecosystems respond to chronic resource
-alteration. We propose four alternative predictions for the effect of
+alteration. We propose four alternative scenarios for the effect of
 precipitation manipulation on the ecosystem functional response to soil
@@
 define `ecosystem functional response' as the relationship between
-available soil moisture and ANPP. The four predictions are based on
+available soil moisture and ANPP. The four scenarios are based on
 possible outcomes at the community (e.g., community composition) and
@@
 
-First, altered precipitation changes neither ecosystem functional
-response nor community composition (Fig. 1, top left). In this case,
-changes in ANPP simply follow the soil moisture-ANPP relationship under
-ambient conditions. This corresponds to the early phases of the
-Hierarchical Response Framework, where ecosystem response follows the
-physiological responses of individual species. Second, the ecosystem
-functional response changes but community composition remains the same
-(Fig. 1, top right). A saturating soil moisture-ANPP response fits this
-scenario, where individual species hit physiological thresholds or are
-limited by some other resource. Third, the ecosystem functional response
-is consistent but underlying community composition changes (Fig. 1,
-bottom left). In this case, changes in species' identities or abundances
-occur in response to altered precipitation levels and species more
-suited to the new conditions compensate for reduced function of initial
-species. Fourth, and last, both ecosystem functional response and
-community composition change (Fig. 1, bottom right). New species, or
-newly abundant species, with different physiological responses
-completely reshape the ecosystem functional response.
+First, altered precipitation might have no effect on either ecosystem
+functional response or community composition (Fig. 1, top left). In this
+case, changes in ANPP would be well predicted by the current, observed
+soil moisture-ANPP relationship. This corresponds to the early phases of
+the Hierarchical Response Framework, where ecosystem response follows
+the physiological responses of individual species. Second, the ecosystem
+functional response might change while community composition remains the
+same (Fig. 1, top right). A saturating soil moisture-ANPP response fits
+this scenario, where individual species hit physiological thresholds or
+are limited by some other resource. Third, the ecosystem functional
+response might be constant but community composition changes (Fig. 1,
+bottom left). In this case, changes in species' identities and/or
+abundances occur in response to altered precipitation levels and species
+more suited to the new conditions compensate for reduced function of
+initial species. Fourth, and last, both ecosystem functional response
+and community composition could change (Fig. 1, bottom right). New
+species, or newly abundant species, with different physiological
+responses completely reshape the ecosystem functional response.
 
@@
 scales at which the different scenarios play out likely differ
-\citep{Smith2009, Wilcox2016, Knapp2017}. Thus, our task is not to test
-the validity of the Hierarchical Response Framework, but rather to amass
-information on how quickly ecosystem functional responses change in
-different ecosystems. We also need to understand whether changes at the
-ecosystem level are driven by community level changes or individual
-level responses.
+\citep{Smith2009, Wilcox2016, Knapp2017}. To determine these time
+scales, we need to amass information on how quickly ecosystem functional
+responses change in different ecosystems. We also need to understand
+whether changes at the ecosystem level are driven by community level
+changes or individual level responses.
 
@@
 We focus on how the drought and irrigation treatments affect the
-relationship between available soil moisture and ANPP, and if community
-dynamics underlie the ecosystem responses. In particular, we are
-interested in the consistency of the soil moisture-ANPP relationship
-among treatments. Is the relationship steeper under the drought
-treatment, at low soil moisture? Does the relationship saturate under
-the irrigation treatment, at high soil moisture? To answer these
-questions we fit a generalized linear mixed effects model to test
-whether the regressions differed among treatments. We also analyzed
-community composition over time, allowing us to place our experimental
-results within the framework our competing predictions (Fig. 1).
+relationship between interannual variation in available soil moisture
+and interannual variation in ANPP, and if community dynamics underlie
+the ecosystem responses. In particular, we are interested in the
+consistency of the soil moisture-ANPP relationship among treatments. Is
+the relationship steeper under the drought treatment at low soil
+moisture? Does the relationship saturate under the irrigation treatment
+at high soil moisture? To answer these questions we fit a generalized
+linear mixed effects model to test whether the regressions differed
+among treatments. We also analyzed community composition over time,
+allowing us to place our experimental results within the framework our
+scenarios (Fig. 1).
 
@@
 We conducted our precipitation manipulation experiment in a sagebrush
-steppe community at the United States Sheep Experimental Station (USSES)
-near Dubois, Idaho (44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m
-above sea level. The plant community is dominated by the shrub
-\emph{Artemesia tripartita} and three perennial bunchgrasses,
-\emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and
-\emph{Hesperostipa comata}. During the period of our experiment (2011 --
-2015), average mean annual precipitation was 265 mm
-year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
--5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July. Between 1926
-and 1932, range scientists at the USSES established 26 permanent 1
-m\(^2\) quadrats to track vegetation change over time. In 2007, we
-(well, one of us {[}P. Adler{]}) relocated 14 of the original quadrats,
-six of which were inside a large, permanent livestock exclosure. We use
-these six plots as control plots that have received no treatment, just
-ambient precipitation, in the experiment described below.
+steppe community at the USDA, U.S. Sheep Experiment Station (USSES) near
+Dubois, Idaho (44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m above
+sea level. The plant community is dominated by the shrub \emph{Artemesia
+tripartita} and three perennial bunchgrasses, \emph{Pseudoroegneria
+spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata}. During the
+period of our experiment (2011 -- 2015), average mean annual
+precipitation was 265 mm year\(\phantom{}^{-1}\) and mean monthly
+temperature ranged from -5.2\(^{\circ}\)C in January to
+21.8\(^{\circ}\)C in July. Between 1926 and 1932, range scientists at
+the USSES established 26 permanent 1 m\(^2\) quadrats to track
+vegetation change over time. In 2007, we relocated 14 of the original
+quadrats, six of which were inside a large, permanent livestock
+exclosure. We use these six plots as control plots (i.e.~ambient
+precipitation) in the experiment described below.
 
-\subsection{Precipitation Experiment}\label{precipitation-experiment}
-
-In spring 2011, we (well, two of us {[}A. Kleinhesselink and P.
-Adler{]}) established 16 new 1 m\(^2\) plots located in the same
-exclosure as the six control plots. We avoided areas on steep hill
+In spring 2011, we established 16 new 1 m\(^2\) plots located in the
+same exclosure as the six control plots. We avoided areas on steep hill
 slopes, areas with greater than 20\% cover of bare rock, and areas with
@@
 ``irrigation'' treatment.
+
+\subsection{Precipitation Experiment}\label{precipitation-experiment}
 
@@
 was triggered by float switches once water levels reached about 20
-liters. We disconnected the irrigation pumps each fall and reconnected
-them, often with difficulty, each spring. The rain-out shelters remained
-in place throughout the year.
+liters. We disconnected the irrigation pumps each October and
+reconnected them each April. The rain-out shelters remained in place
+throughout the year.
 
@@
 Decagon Devices (Pullman, Washington) 5TM and EC-5 soil moisture
-sensors. We installed four sensors around the edges of each plot, two at
-5 cm soil depth and two at 25 cm soil depth. We also installed four
-sensors in areas nearby the four selected plot pairs to measure ambient
-soil moisture at the same depths. Soil moisture measurements were
-automatically logged every four hours. We coupled this temporally
-intensive soil moisture sampling with spatially extensive readings taken
-at six points within all 16 plots and associated ambient measurement
-areas. These snapshot data were collected on 06-06-2012, 04-29-2015,
-05-07-2015, 06-09-2015, and
+sensors. We installed four sensors around the edges of each 1x1 m census
+plot, two at 5 cm soil depth and two at 25 cm soil depth. We also
+installed four sensors in areas nearby the four selected plot pairs to
+measure ambient soil moisture at the same depths. Soil moisture
+measurements were automatically logged every four hours. We coupled this
+temporally intensive soil moisture sampling with spatially extensive
+readings taken at six points within all 16 plots and associated ambient
+measurement areas. These snapshot data were collected on 06-06-2012,
+04-29-2015, 05-07-2015, 06-09-2015, and
 05-10-2016\footnote{Dates formatted as: mm-dd-yyyy.} using a handheld
@@
 produce average daily soil moisture values for the ambient, drought, and
-irrigation conditions, we used a statistical model to predict the
+irrigation treatments, we used a statistical model to predict the
 average treatment effects on soil moisture during the course of the
@@
 treated plots on days for which we took at least one ambient soil
-moisture measurement.
+moisture measurement. See \citet{Kleinhesselink2017b} for complete
+details on our approach to estimating soil moisture.
 
@@
 2013, 2014), or 16 (2016) calibration plots adjacent to the experimental
-site, in which we harvested all aboveground biomass, dried it to a
-constant weight at 60\(^{\circ}\)C, and weighed it to estimate ANPP.
+site, in which we harvested all aboveground biomass produced in the
+current year (we excluded litter and standing dead material), dried it
+to a constant weight at 60\(^{\circ}\)C, and weighed it to estimate
+ANPP. We harvested at peak biomass each year, typically in late June.
 
@@
 wavelength and then calculated a greenness index based on the same bands
-used to calculate NDVI using the MODIS and AVHRR bands for NDVI. We
-regressed the greenness index against the dry biomass weight from the
-ten calibration plots to convert the greenness index to ANPP. We fit
-regressions to a MODIS-based index and an AVHRR-based index for each
-year and retained the regression with the better fit based on \(R^2\)
-values. We then predicted ANPP using the best regression equation for
-each year (Appendix 1).
+used to calculate NDVI using the MODIS and AVHRR {[}YOU HAVEN'T DEFINED
+THESE{]} bands for NDVI. We regressed the greenness index against the
+dry biomass weight from the ten calibration plots to convert the
+greenness index to ANPP. We fit regressions to a MODIS-based index and
+an AVHRR-based index for each year and retained the regression with the
+better fit based on \(R^2\) values. We then predicted ANPP using the
+best regression equation for each year (Appendix 1).
 
@@
 each plot made using a pantograph \citep{Hill1920} and yearly counts of
-annual species in each plot. We extracted the mapped information to
-obtain data on the density of all annuals and perennials forbs, the
-basal cover of perennial grasses, and the canopy cover of shrubs. We
-made a large plot-treatment-year by species matrix, where columns were
-filled with either basal cover or density, depending on the measurement
-made for the particular species. We standardized the values in each
-column so we could analyze the different types of data together. This
-puts all abundance values on the same scale, meaning that common and
-rare species are weighted equally. This approach is anti-conservative
-for detecting community change if we assume that rare species will
-respond more than common ones, which biases our approach toward
-detecting compositional changes.
+annual species in each plot. From these sources, we determined the
+density of all annuals and perennials forbs, the basal cover of
+perennial grasses, and the canopy cover of shrubs. We made a large
+plot-treatment-year by species matrix, where columns were filled with
+either basal cover or density, depending on the measurement made for the
+particular species. We standardized the values in each column so we
+could directly compare species quantified with different metrics
+(density, basal cover, and canopy cover). This puts all abundance values
+on the same scale, meaning that common and rare species are weighted
+equally. Assuming that rare species will respond to treatments more than
+common ones, our approach is biased towards detecting compositional
+changes.
 
@@
 observations as the main treatment and then estimates intercept and
-slope offsets for the ``drought'' and ``irrigation'' treatments. In
-reference to our model, the hypotheses we wish to test are:
+slope offsets for the ``drought'' and ``irrigation'' treatments. We use
+our model to test the following hypotheses:
 
@@
 \end{description}
+
+These hypotheses are based on evidence that precipitation-ANPP
+relationships often saturate with increasing precipitation (REFERENCES).
 
@@
 We used nonmetric multidimensional scaling (NMDS) based on Bray-Curtis
-distances to see if community composition differed among treatments
-through time. We first calculated Bray-Curtis distances among all plots
+distances to identify temporal changes in community composition among
+treatments. We first calculated Bray-Curtis distances among all plots
 for each year of the experiment and then extracted those distances for
@@
 
-Three of our five treatment years fell in years of below average
-rainfall (Fig. 2A). Thus, those three years represent a lower magnitude
-of absolute change in precipitation experienced by the treatments. ANPP
-varied from a minimum of 74.5 g m\(^{-2}\) in 2014 to a maximum of 237.1
-g m\(^{-2}\) in 2016 when averaged across treatments (Fig. 2C). ANPP was
-slightly higher in irrigation plots and slightly lower in drought plots
-(Fig. 2C), corresponding to estimated soil volumetric water content
-(VWC) differences among treatments (Fig. 2B). Such differences in soil
-VWC indicate our treatment infrastructure was successful. ANPP was
-highly variable within years (Fig. 2C).
+Three of our five study years received below average precipitation (Fig.
+2A). Because our treatments are proportional changes in ambient
+precipitation, those three years represent a lower magnitude of absolute
+difference in precipitation among the treatments. ANPP varied from a
+minimum of 74.5 g m\(^{-2}\) in 2014 to a maximum of 237.1 g m\(^{-2}\)
+in 2016 when averaged across treatments (Fig. 2C). ANPP was slightly
+higher in irrigation plots and slightly lower in drought plots (Fig.
+2C), corresponding to estimated soil volumetric water content (VWC)
+differences among treatments (Fig. 2B). Such differences in soil VWC
+indicate our treatment infrastructure was successful. ANPP was highly
+variable across plots within years (Fig. 2C).
 
@@
 
-Community composition was similar among treatments. In no year did
-community composition among treatments not overlap, and they were
-equally dispersed in all years (Table 2; Fig. 5). Community composition
-was also remarkably stable over time, with no evidence of divergence
-among treatments (Table 2; Fig. 5).
+Community composition was similar among treatments. Community
+composition among treatments overlapped in all years and was equally
+dispersed in all years (Table 2; Fig. 5). Community composition was also
+remarkably stable over time, with no evidence of divergence among
+treatments (Table 2; Fig. 5).
 
@@
 rapid, on the order of years \citep{Hoover2014}, and slow, on order of
-decades \citep{Knapp2012, Wilcox2016}. Thus, a lingering question is how
-the time scales of ecosystem response and community change differ among
-ecosystems, which can be answered by manipulating precipitation to reach
-extreme levels.
+decades \citep{Knapp2012, Wilcox2016}. A lingering question is how the
+time scales of ecosystem response and community change vary among
+ecosystems. Precipitation manipulation experiments can help answer this
+question, especially if they push water availability outside the
+historical range of variability for long periods.
 
@@
 functional response nor community composition changed under chronic
-drought (Fig. 3A, Fig. 4), consistent with the top left prediction in
-Fig. 1. Ecosystem functional response under chronic irrigation was
-different from the control treatment, but community composition remained
-unchanged (Fig. 3A, Fig. 4), consistent with the top right prediction in
-Fig. 1. The altered ecosystem functional response under irrigation
-matched our expectations because we found evidence for a saturating
-response for the irrigation treatment (Fig. 3A, right panel). However,
-the response of ANPP to soil moisture in all treatments was consistently
-weak (Table 1, Fig. A2-2) and within-year variation of ANPP was large
-(Fig. 3B)
+drought (Fig. 3A, Fig. 4), representin the top left SCENARIO in Fig. 1.
+Ecosystem functional response under chronic irrigation was different
+from the control treatment, but community composition remained unchanged
+(Fig. 3A, Fig. 4), representing the top right SCENARIO in Fig. 1. The
+altered ecosystem functional response under irrigation matched our
+expectations because we found evidence for a saturating response for the
+irrigation treatment (Fig. 3). However, the response of ANPP to soil
+moisture in all treatments was consistently weak (Table 1, Fig. A2-2)
+and within-year variation of ANPP was large (Fig. 3B).
 
@@
 ambient precipitation was below average (Fig. 2A). The irrigation
-treatment may have been too small, yet that is the treatment where we
+treatment may not qualify as extreme, yet that is the treatment where we
 did observe an effect (Fig. 3).
 
-Second, ANPP at our site may be influenced by factors beyond the window
-of soil moisture (cumulative March-June) we included in our statistical
-model. For example, temperature can impact ANPP directly
+Second, ANPP at our site may be influenced by additional factors, not
+only the cumulative March-June soil moisture covariate we included in
+our statistical model. For example, temperature can impact ANPP directly
 \citep{Epstein1997} and by exacerbating the effects of soil moisture
@@
 redistribute snow across our treatments in the winter, and snow melt may
-spur early spring growth. These statistical issues of missing
-potentially important covariates could explain the within-year spread of
-ANPP (Fig 2C, Fig. 3B) and the resulting uncertain relationship we
-observed between soil moisture and ANPP across all treatments (Table 1,
-Fig. A2-2).
+spur early spring growth. Failure to account for potentially important
+covariates could explain the within-year spread of ANPP (Fig 2C, Fig.
+3B) and the resulting uncertain relationship we observed between soil
+moisture and ANPP across all treatments (Table 1, Fig. A2-2).
 
@@
 species forego short-term gains to reduce the variance of long-term
-success \citep{Seger1987}. In other words, species do the same thing
-every year, with only minimal response to environmental conditions. The
-dry and variable environment of the sagebrush steppe has likely selected
-for bet hedging species that can maintain function at low water
-availability and have weak responses to high water availability. In so
-doing, the dominant species in our ecosystem avoid ``boom and bust''
-cycles, which corresponds to the weak effect of soil moisture on ANPP
-(i.e., the Bayesian credible intervals for the slopes in
-\(\boldsymbol{\beta}\) in Eq. 1 overlapping zero).
+success \citep{Seger1987}. In other words, species follow the same
+conservative strategy every year, designed to minimize response to
+environmental conditions. The dry and variable environment of the
+sagebrush steppe has likely selected for bet hedging species that can
+maintain function at low water availability and have weak responses to
+high water availability. In so doing, the dominant species in our
+ecosystem avoid ``boom and bust'' cycles, which corresponds to the weak
+effect of soil moisture on ANPP (i.e., the Bayesian credible intervals
+for the slopes in \(\boldsymbol{\beta}\) in Eq. 1 overlapping zero).
 
@@
 avoidance, which would also result in a consistent ecosystem functional
-response between drought and control treatments. The perennial plants in
-this cold desert ecosystem are tolerant to drought conditions
-\citep[A.R. Kleinhesselink, unpublished data]{Bazzaz1979, Franks2011}.
-For example, many of the perennial grasses in our focal ecosystem avoid
-drought stress by growing early in the growing season \citep[A.R.
-Kleinhesselink, personal observation]{Blaisdell1958}. Furthermore, the
-dominant shrub in our focal ecosystem, \emph{Artemisia tripartita}, has
-access to water deep in the soil profile thanks to a deep root system
+response between drought and control treatments. For example, many of
+the perennial grasses in our focal ecosystem avoid drought stress by
+growing early in the growing season \citep[A.R. Kleinhesselink, personal
+observation]{Blaisdell1958}. Furthermore, the dominant shrub in our
+focal ecosystem, \emph{Artemisia tripartita}, has access to water deep
+in the soil profile thanks to a deep root system
 \citep{Kulmatiski2017a}.
@@
 The weaker soil moisture-ANPP relationship we observed for the
-irrigation treatment, evidenced by the less positive slope relative to
-controls (Fig. 3A, right panel), has three non-exclusive explanations.
-First, species' biomass production may have hit physiological thresholds
-and/or become limited by other factors, such as nitrogen availability
+irrigation treatment has three non-exclusive explanations. First,
+species' biomass production may have hit physiological thresholds and/or
+become limited by other factors, such as nitrogen availability
 \citep{LeBauer2008}.
-\pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES? WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY OR UNLIKEL THIS IS...}
-Second, the less positive slope could be another indicator of bet
-hedging in this community. Avoiding high biomass and seed production in
-wet years could benefit species in the long run if climate conditions
-are variable. Third, the less positive slope may be a statistical
-artifact due the variability of ANPP within years and our low sample
-size.
+\pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES?
+WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY OR UNLIKEL THIS IS...NUTNET SHOWS POSITIVE EFFECTS OF N*P ON PRODUCTION, EFFECT OF N ALONE NOT AS CLEAR.
+YOU COULD CITE THIS AS UNPUBLISHED DATA}
+
+Second, the less positive slope could be an indicator of bet hedging in
+this community. Avoiding high biomass and seed production in wet years
+could benefit species in the long run if climate conditions are
+variable.
+\pba{but why would this have cause a change relative to controls? this should also be true for controls...}
+Third, the less positive slope may be a statistical artifact due the
+variability of ANPP within years and our low sample size.
 
 In conclusion, our results suggest the species in our focal plant
-community are tolerant of drought conditions and bet hedgers in wet
-conditions. Longer time series of chronic precipitation alteration may
-reveal plant community shifts that we did not observe
-\citep[e.g.,][]{Wilcox2016}, in which case species that do not bet hedge
-may gain prominence and dominate the ecosystem functional response. Our
-results suggest compositional shifts would have the largest impact at
-high rainfall because the current community maintained consistent
-ecosystem functional response at very low water availability.
+community are tolerant of drought conditions and unresponsive to wet
+conditions.
+\pba{no, you can't be a bet hedger in wet years but not in dry years...}
+Longer, chronic precipitation alteration might reveal plant community
+shifts that we did not observe \citep[e.g.,][]{Wilcox2016}. For example,
+a long-term increase in water availability could allow species that do
+not bet hedge to gain prominence and dominate the ecosystem functional
+response. Our results suggest compositional shifts would have the
+largest impact at high rainfall because the current community maintained
+consistent ecosystem functional response at very low water availability.
 
@@
 We thank the many summer research technicians who collected the data
-reported in this paper and the US Experimental Sheep Station for
-facilitating work on their property. We also thank Susan Durham for
-clarifying our thinking on the statistical analyses and Kevin Wilcox for
-helpful discussions on analyzing community composition data.
+reported in this paper and the USDA, U.S. Sheep Experiment Station
+(Dubois, ID) for facilitating work on their property. We also thank
+Susan Durham for clarifying our thinking on the statistical analyses and
+Kevin Wilcox for helpful discussions on analyzing community composition
+data.
 
@@
   \caption{Results from the generalized linear mixed effects model.
-(A) Posterior distributions of the intercept and slope offsets for the drought and irrigation treatments.
-Offsets indicate the amount to which the coefficients for drought or irrigation treatments differ from the control treatment estimates.
-Probabilities (``Pr $<$ 0 ='') for each distribution indicate the probability that coefficient is less than zero.
-Probabilities greater than 0.95 indicate strong support for the coefficient being less than zero.
-We only show the one-tailed probability for the value being less than zero because the median of each distribution is less than zero.
-Kernel bandwidths of posterior densities were adjusted by a factor of 5 to smooth the distribution for visual clarity.
-(B) Scatterplot of the data and model estimates shown a solid lines.
+(A) Posterior distributions for the effects of drought and irrigation on the intercept and slope of the productivity-soil moisture relationship.
+Treatment effects show the difference between the coefficients estimated in the treated plots and the control plots.
+Probabilities (``Pr $\lessgtr$ 0 ='') for each coefficient indicate the one-tailed probability that the coefficient is less than or greater than zero, depending on whether the median of the distribution is less than or greater than zero.
+The posterior densities were smoothed for visual clarity by increasing kernal bandwidth by a factor of five.
+(B) Scatterplot of the data and model estimates shown as solid lines.
 Model estimates come from treatment level coefficients (colored lines).
@@
 
-\bibliography{~./usses_water.bib}
+\bibliography{/Users/atredenn/Dropbox/Bibliography/usses_water.bib}
 

### 2017-09-19 code/04_plot_model_results.R

@@ -70,6 +70,7 @@ slope_probs <- treat_slopes %>%
   group_by(Treatment, Type) %>%
   summarise(probs = round(get_one_tailed(estimate),2)) %>%
   mutate(prob_text = paste("Pr < 0 =", probs))
+slope_probs$prob_text[1] <- paste("Pr > 0 =", slope_probs$probs[1]) # switch direction of drought intercpet since greater than 0
 slope_probs$xpos <- c(-2.8, -1, -3.6, -1.15)
 slope_probs$ypos <- c(1.3,1.3,2.55,2.7)
 

### 2017-09-22 manuscript/usses_water_ms.Rmd
rewrite discussion after PBA comments

@@
   **Discussion.** Despite some evidence that ecosystem functional response became less sensitive under sustained wet conditions, the response of ANPP to soil moisture was consistently weak and community composition was stable.
-Differences in ecosystem functional responses across treatments were not related to compensatory shifts at the plant community level, but instead may reflect the insensitivity of the dominant species to soil moisture.
+The similarity of ecosystem functional responses across treatments was not related to compensatory shifts at the plant community level, but instead may reflect the insensitivity of the dominant species to soil moisture.
 These species may be successful precisely because they have evolved life history strategies which buffer them against precipitation variability.
@@
 Thus, our model treats "control" observations as the main treatment and then estimates intercept and slope offsets for the "drought" and "irrigation" treatments. 
-We use our model to test the following hypotheses:
+We use our model to test two statistical hypotheses based on the questions in our Introduction:
 
@@
 
-Ecosystem response to precipitation extremes depends on the physiological responses of constituent species and the rate at which community composition shifts to favor species better able to take advantage of, or cope with, new resource levels [@Smith2009].
+Ecosystem response to a new precipitation regime depends on the physiological responses of constituent species and the rate at which community composition shifts to favor species better able to take advantage of, or cope with, new resource levels [@Smith2009].
 Previous work has shown that community compositional shifts can be both rapid, on the order of years [@Hoover2014], and slow, on order of decades [@Knapp2012; @Wilcox2016].
@@
 The results of our five year experiment in a sagebrush steppe conform to two of our four predictions, depending on treatment.
-Neither ecosystem functional response nor community composition changed under chronic drought (Fig. 3A, Fig. 4), representin the top left SCENARIO in Fig. 1.
+Ecosystem functional response and community composition did not change under chronic drought (Fig. 3A, Fig. 4), representing the top left scenario in Fig. 1.
 Ecosystem functional response under chronic irrigation was different from the control treatment, but community composition remained unchanged (Fig.
 3A, Fig.
-4), representing the top right SCENARIO in Fig.
+4), representing the top right scenario in Fig.
 1.
-The altered ecosystem functional response under irrigation matched our expectations because we found evidence for a saturating response for the irrigation treatment (Fig. 3).
-<!--- Can we really call it saturating if it is concave up? More accurate to say weaker response than control? --->
-However, the response of ANPP to soil moisture in all treatments was consistently weak (Table 1, Fig. A2-2) and within-year variation of ANPP was large (Fig. 3B).
-<!---Last sentence leaves me hanging...what's your point? Did it the response change or not? Statistically but not ecologically significant?--->
+The altered ecosystem functional response under irrigation matched our expectations because we found evidence for a weaker response for the irrigation treatment (Fig. 3).
+Although the difference between the control and irrigation slopes is likely different from zero (Fig.
+3A), the posterior distribution of the slopes (not their difference) are similar and overlap broadly (Table1, Fig.
+A2-2).
+Thus, the statistical difference between the slopes is not ecologically meaningful.
+We therefore conclude that ecosystem functional response is consistent and weak across all precipitation treatments.
 
@@
 3) and community composition (Fig.
-5) between drought and control treatments is surprising because grasslands generally, and sagebrush steppe specifically, are considered water-limited systems.
-Indeed, we expected ecosystem functional response, community composition, or both to change under the drought treatment, landing us in any box of Fig. 1 \emph{except} the top left.
-Why did our drought treatment fail to induce ecosystem or community responses?
+5) among treatments is surprising because grasslands generally, and sagebrush steppe specifically, are considered water-limited systems.
+Indeed, we expected ecosystem functional response, community composition, or both to change under our treatments.
+Why did our treatments fail to induce ecosystem or community responses?
 We can think of three reasons; two are limitations of our study, and one is the life history traits of the species in our focal communities.
@@
 
-First, it could be that our drought manipulation was not large enough to induce a response.
+First, it could be that our precipitation manipulations were not large enough to induce a response.
 That is, a 50\% decrease in any given year may not be abnormal given our site's historical range of variability [@Knapp2017].
-We cannot definitively rule out this possibility, but we have reason to believe our drought treatment *should* have been large enough.
+We cannot definitively rule out this possibility, but we have reason to believe our treatments *should* have been large enough.
 Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of mean growing season precipitation necessary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site (Fig.
@@
 Thus, our drought treatment represented extreme precipitation amounts, especially in years where ambient precipitation was below average (Fig. 2A).
-The irrigation treatment may not qualify as extreme, yet that is the treatment where we did observe an effect (Fig. 3).
+The irrigation treatment may not qualify as extreme, yet that is the treatment where we did observe an effect (Fig. 3A).
 
@@
 A2-2).
-<!---so why didn't we do this? sample size limitations, power? --->
+Unfortunately, we did not collect data on additional factors because our focus was on precipitation.
 
-Third, the life history traits of the dominant species in our study ecosystem may explain the consistently positive, but weak and uncertain, effect of soil moisture on ANPP (Fig. 3).
+Third, the life history traits of the dominant species in our study ecosystem may explain the consistently positive, but weak and uncertain, effect of soil moisture on ANPP (Table 1, Fig. 3).
 Species that live in variable environments, such as cold deserts, must have strategies to ensure long-term success as conditions vary.
@@
 
+<!---
 The weaker soil moisture-ANPP relationship we observed for the irrigation treatment has three non-exclusive explanations.
@@
 Third, the less positive slope may be a statistical artifact due the variability of ANPP within years and our low sample size.
+--->
 
-In conclusion, our results suggest the species in our focal plant community are tolerant of drought conditions and unresponsive to wet conditions.
-\pba{no, you can't be a bet hedger in wet years but not in dry years...}
+In conclusion, our results suggest the species in our focal plant community are insensitive to to changes in precipitation regime. 
+Such insensitivity can buffer species against precipitation variability in this semi-arid ecosystem, making them successful in the long run.
 Longer, chronic precipitation alteration might reveal plant community shifts that we did not observe [e.g., @Wilcox2016]. 

### 2017-09-22 manuscript/usses_water_ms.tex
rewrite discussion after PBA comments

@@
 response of ANPP to soil moisture was consistently weak and community
-composition was stable. Differences in ecosystem functional responses
-across treatments were not related to compensatory shifts at the plant
+composition was stable. The similarity of ecosystem functional responses
+across treatments was not related to compensatory shifts at the plant
 community level, but instead may reflect the insensitivity of the
@@
 slope offsets for the ``drought'' and ``irrigation'' treatments. We use
-our model to test the following hypotheses:
+our model to test two statistical hypotheses based on the questions in
+our Introduction:
 
@@
 
-Ecosystem response to precipitation extremes depends on the
+Ecosystem response to a new precipitation regime depends on the
 physiological responses of constituent species and the rate at which
@@
 The results of our five year experiment in a sagebrush steppe conform to
-two of our four predictions, depending on treatment. Neither ecosystem
-functional response nor community composition changed under chronic
-drought (Fig. 3A, Fig. 4), representin the top left SCENARIO in Fig. 1.
-Ecosystem functional response under chronic irrigation was different
-from the control treatment, but community composition remained unchanged
-(Fig. 3A, Fig. 4), representing the top right SCENARIO in Fig. 1. The
-altered ecosystem functional response under irrigation matched our
-expectations because we found evidence for a saturating response for the
-irrigation treatment (Fig. 3). However, the response of ANPP to soil
-moisture in all treatments was consistently weak (Table 1, Fig. A2-2)
-and within-year variation of ANPP was large (Fig. 3B).
+two of our four predictions, depending on treatment. Ecosystem
+functional response and community composition did not change under
+chronic drought (Fig. 3A, Fig. 4), representing the top left scenario in
+Fig. 1. Ecosystem functional response under chronic irrigation was
+different from the control treatment, but community composition remained
+unchanged (Fig. 3A, Fig. 4), representing the top right scenario in Fig.
+1. The altered ecosystem functional response under irrigation matched
+our expectations because we found evidence for a weaker response for the
+irrigation treatment (Fig. 3). Although the difference between the
+control and irrigation slopes is likely different from zero (Fig. 3A),
+the posterior distribution of the slopes (not their difference) are
+similar and overlap broadly (Table1, Fig. A2-2). Thus, the statistical
+difference between the slopes is not ecologically meaningful. We
+therefore conclude that ecosystem functional response is consistent and
+weak across all precipitation treatments.
 
 The similarity of ecosystem functional response (Fig. 3) and community
-composition (Fig. 5) between drought and control treatments is
-surprising because grasslands generally, and sagebrush steppe
-specifically, are considered water-limited systems. Indeed, we expected
-ecosystem functional response, community composition, or both to change
-under the drought treatment, landing us in any box of Fig. 1
-\emph{except} the top left. Why did our drought treatment fail to induce
-ecosystem or community responses? We can think of three reasons; two are
-limitations of our study, and one is the life history traits of the
-species in our focal communities. We first discuss the potential
-limitations of our study, and then discuss the biological explanation.
+composition (Fig. 5) among treatments is surprising because grasslands
+generally, and sagebrush steppe specifically, are considered
+water-limited systems. Indeed, we expected ecosystem functional
+response, community composition, or both to change under our treatments.
+Why did our treatments fail to induce ecosystem or community responses?
+We can think of three reasons; two are limitations of our study, and one
+is the life history traits of the species in our focal communities. We
+first discuss the potential limitations of our study, and then discuss
+the biological explanation.
 
-First, it could be that our drought manipulation was not large enough to
-induce a response. That is, a 50\% decrease in any given year may not be
-abnormal given our site's historical range of variability
+First, it could be that our precipitation manipulations were not large
+enough to induce a response. That is, a 50\% decrease in any given year
+may not be abnormal given our site's historical range of variability
 \citep{Knapp2017}. We cannot definitively rule out this possibility, but
-we have reason to believe our drought treatment \emph{should} have been
-large enough. Using the methods described by \citet{Lemoine2016}, we
+we have reason to believe our treatments \emph{should} have been large
+enough. Using the methods described by \citet{Lemoine2016}, we
 calculated the percent reduction and increase of mean growing season
@@
 treatment may not qualify as extreme, yet that is the treatment where we
-did observe an effect (Fig. 3).
+did observe an effect (Fig. 3A).
 
@@
 moisture and ANPP across all treatments (Table 1, Fig. A2-2).
+Unfortunately, we did not collect data on additional factors because our
+focus was on precipitation.
 
@@
 ecosystem may explain the consistently positive, but weak and uncertain,
-effect of soil moisture on ANPP (Fig. 3). Species that live in variable
-environments, such as cold deserts, must have strategies to ensure
-long-term success as conditions vary. One strategy is bet hedging, where
-species forego short-term gains to reduce the variance of long-term
-success \citep{Seger1987}. In other words, species follow the same
-conservative strategy every year, designed to minimize response to
+effect of soil moisture on ANPP (Table 1, Fig. 3). Species that live in
+variable environments, such as cold deserts, must have strategies to
+ensure long-term success as conditions vary. One strategy is bet
+hedging, where species forego short-term gains to reduce the variance of
+long-term success \citep{Seger1987}. In other words, species follow the
+same conservative strategy every year, designed to minimize response to
 environmental conditions. The dry and variable environment of the
@@
 
-The weaker soil moisture-ANPP relationship we observed for the
-irrigation treatment has three non-exclusive explanations. First,
-species' biomass production may have hit physiological thresholds and/or
-become limited by other factors, such as nitrogen availability
-\citep{LeBauer2008}.
-\pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES?
-WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY OR UNLIKEL THIS IS...NUTNET SHOWS POSITIVE EFFECTS OF N*P ON PRODUCTION, EFFECT OF N ALONE NOT AS CLEAR.
-YOU COULD CITE THIS AS UNPUBLISHED DATA}
-
-Second, the less positive slope could be an indicator of bet hedging in
-this community. Avoiding high biomass and seed production in wet years
-could benefit species in the long run if climate conditions are
-variable.
-\pba{but why would this have cause a change relative to controls? this should also be true for controls...}
-Third, the less positive slope may be a statistical artifact due the
-variability of ANPP within years and our low sample size.
-
 In conclusion, our results suggest the species in our focal plant
-community are tolerant of drought conditions and unresponsive to wet
-conditions.
-\pba{no, you can't be a bet hedger in wet years but not in dry years...}
+community are insensitive to to changes in precipitation regime. Such
+insensitivity can buffer species against precipitation variability in
+this semi-arid ecosystem, making them successful in the long run.
 Longer, chronic precipitation alteration might reveal plant community

### 2017-09-22 manuscript/usses_water_ms.tex
updates ord plot to have coord_fixed()

@@
   \hline
-2011 & adonis &  21 &   2 & 1.02 & 0.42 \\ 
+2011 & adonis &  21 &   2 & 1.02 & 0.44 \\ 
   2011 & betadisper &  21 &   2 & 2.23 & 0.14 \\ 
-  2012 & adonis &  22 &   2 & 1.10 & 0.34 \\ 
+  2012 & adonis &  22 &   2 & 1.10 & 0.32 \\ 
   2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
@@
   2013 & betadisper &  22 &   2 & 0.28 & 0.76 \\ 
-  2014 & adonis &  22 &   2 & 0.95 & 0.54 \\ 
+  2014 & adonis &  22 &   2 & 0.95 & 0.58 \\ 
   2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
@@
   2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 
-  2016 & adonis &  21 &   2 & 1.07 & 0.33 \\ 
+  2016 & adonis &  21 &   2 & 1.07 & 0.36 \\ 
   2016 & betadisper &  21 &   2 & 0.50 & 0.62 \\ 

### 2017-09-22 code/05_community_dynamics.R

@@ -162,75 +162,8 @@ ggplot(nmds_df, aes(x=MDS1, y=MDS2, color=Treatment))+
   facet_wrap("Year")+
   scale_y_continuous(limits=c(-0.15,0.15))+
   scale_x_continuous(limits=c(-0.18,0.18))+
+  coord_fixed()+
   theme_few()+
   theme(panel.grid.major = element_line(color="grey90"))
 ggsave(paste0(figure_path,"sppcomp_bray_all.png"), width=6, height = 4, units = "in", dpi = 120)
 
-
-
-
-
-####
-####  RANK CLOCKS ON SCALED COMMUNITY DATA ----
-####
-# scaled_comp_ts <- full_community_matrix_scaled %>%
-#   gather(key = species, value = scaled_abundance, -quad, -year, -Treatment) %>%
-#   group_by(year, Treatment, species) %>%
-#   summarise(mean_scaled_abundance = mean(scaled_abundance))
-# 
-# ## Create the graph
-# g1 <- ggplot(scaled_comp_ts, aes(year, mean_scaled_abundance, color = species)) + 
-#   geom_line(size = 1, alpha = 0.8) + 
-#   facet_wrap(~Treatment) +
-#   coord_polar()+
-#   labs(x=NULL, y=NULL) + 
-#   guides(color=FALSE)+
-#   theme_dark()+
-#   theme(text = element_text(size = 14), 
-#         strip.text.x = element_text(size = 14, color="black"), 
-#         strip.background = element_blank(),
-#         panel.grid.major = element_line(size = 1, color="grey55"),
-#         panel.grid.minor = element_line(colour="grey55")) + 
-#   theme(legend.position="bottom", 
-#         legend.text=element_text(face = "italic")) +
-#   geom_segment(aes(x = 2011, y = 0, xend = 2011, yend = 0.4), color = "grey70")+
-#   scale_color_viridis(discrete=T)+
-#   scale_x_continuous(breaks=c(2011,2012,2013,2014,2015))+
-#   ggtitle("A. All species")+
-#   theme(axis.title.y=element_blank(),
-#         axis.text.y=element_blank(),
-#         axis.text.x=element_text(color="grey90"),
-#         axis.ticks.y=element_blank())
-# 
-# 
-# domspp <- c("Artemisia tripartita", "Poa secunda", "Pseudoroegneria spicata", "Hesperostipa comata")
-# scaled_dom_ts <- scaled_comp_ts %>%
-#   filter(species %in% domspp)
[…]

### 2017-09-25 manuscript/usses_water_ms.Rmd
trivia
final edits before PeerJ submission

@@
   **Results.** Ecosystem functional response, defined as the relationship between soil moisture and ANPP was similar among drought and control treatments, but the irrigation treatment had a lower slope than the control treatment.
-However, ANPP response to availabile soil moisture was weak and uncertain regardless of treatment, with all slopes overlapping zero.
+However, ANPP response to available soil moisture was weak and uncertain regardless of treatment, with all slopes overlapping zero.
 There was also large spatial variation in ANPP within-years.
@@
 At any given site, the relationship between aboveground net primary productivity (ANPP) and water availability (e.g., soil moisture) can be characterized by regressing historical observations of ANPP on observations of soil moisture.
-The fitted functional response can then be used to project ANPP under future precipitation regimes (REFERENCE). 
+Fitted functional responses can then be used to infer how ANPP may change under future precipitation regimes [e.g., @Hsu2012]. 
 A problem with this approach is that it requires extrapolation if future precipitation falls outside outside the historical range of variability [@Smith2011; @Peters2012].
@@
 ## Study Area
+
 We conducted our precipitation manipulation experiment in a sagebrush steppe community at the USDA, U.S.
@@
 
-For each plot and year, we averaged the four readings for each wavelength and then calculated a greenness index based on the same bands used to calculate NDVI using the MODIS and AVHRR [YOU HAVEN'T DEFINED THESE] bands for NDVI.
+For each plot and year, we averaged the four readings for each wavelength and then calculated a greenness index based on the same bands used to calculate NDVI using the MODIS (Moderate Resolution Imaging Spectroradiometer) and AVHRR (Advanced Very High Resolution Radiometer) bands for NDVI.
 We regressed the greenness index against the dry biomass weight from the ten calibration plots to convert the greenness index to ANPP.
@@
 
-These hypotheses are based on evidence that precipitation-ANPP relationships often saturate with increasing precipitation (REFERENCES).
-<!---these should probably be stated in the Intro--->
+\noindent{}These hypotheses are based on evidence that precipitation-ANPP relationships often saturate with increasing precipitation [@Hsu2012; @Gherardi2015a].
 
@@
 The random year effects ($\boldsymbol{\eta}$) are drawn from a normal prior with mean 0 and standard deviation $\sigma_{\text{year}}$, which was drawn from a half-Cauchy distribution. 
-A full description of our model is in Appendix 2.
-<!---call out available computer code here?--->
+A full description of our model and associated R [@R2016] code is in Appendix 2.
 
@@
 
-Three of our five study years received below average precipitation (Fig. 2A).
-Because our treatments are proportional changes in ambient precipitation, those three years represent a lower magnitude of absolute difference in precipitation among the treatments.
-<!---why is the previous point important here? should it go in Methods?--->
+Ambient precipitation was variable over the five years of the study (Fig. 2A), meaning our experiment captured a range of environmental conditions.
 ANPP varied from a minimum of 74.5 g m$^{-2}$ in 2014 to a maximum of 237.1 g m$^{-2}$ in 2016 when averaged across treatments (Fig. 2C).
@@
 Community composition was similar among treatments.
-Community composition among treatments overlapped in all years and was equally dispersed in all years (Table 2; Fig. 5).
+The multidimensional space of community composition overlapped among treatments in all years and was equally dispersed in all years (Table 2; Fig. 5).
 Community composition was also remarkably stable over time, with no evidence of divergence among treatments (Table 2; Fig. 5).
@@
 1.
-The altered ecosystem functional response under irrigation matched our expectations because we found evidence for a weaker response for the irrigation treatment (Fig. 3).
-Although the difference between the control and irrigation slopes is likely different from zero (Fig.
-3A), the posterior distribution of the slopes (not their difference) are similar and overlap broadly (Table1, Fig.
-A2-2).
-Thus, the statistical difference between the slopes is not ecologically meaningful.
+The decrease in the slope of the VWC $\times$ productivity relationship in the irrigated plots is consistent with a weaker response of productivity to additional water in this ecosystem.
+This suggests that if average soil moisture were pushed consistently higher than currently observed ambient conditions, there would be a weaker relationship between precipitation and productivity in this system.
+Nonetheless, the slopes of the VWC $\times$ productivity relationship were similar among treatments (Table 1), at least at the level of irrigation we imposed.
 We therefore conclude that ecosystem functional response is consistent and weak across all precipitation treatments.
@@
 The dry and variable environment of the sagebrush steppe has likely selected for bet hedging species that can maintain function at low water availability and have weak responses to high water availability.
-In so doing, the dominant species in our ecosystem avoid "boom and bust" cycles, which corresponds to the weak effect of soil moisture on ANPP (i.e., the Bayesian credible intervals for the slopes in $\boldsymbol{\beta}$ in Eq.
-1 overlapping zero).
+In so doing, the dominant species in our ecosystem avoid "boom and bust" cycles, which corresponds to the weak effect of soil moisture on ANPP (i.e., the Bayesian credible intervals for the slopes overlapping zero; Table 1).
 
@@
 (C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.
-**\emph{note plot b doesn't have a treatment legend, consider adding to the side of (B) and removing from (C)} }
+Colors in panels B and C identify the treatment, as specified in the legend of panel C.}
 \end{figure}
@@
       \includegraphics[width=5in]{../figures/sppcomp_bray_all.png}
-  \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot, colored by treatment.
-**\emph{Worth adding arrow vectors showing association between axes and dominant species?}}
+  \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot, colored by treatment.}
 \end{figure}

### 2017-09-25 manuscript/usses_water_ms.tex
final edits before PeerJ submission

@@
 and control treatments, but the irrigation treatment had a lower slope
-than the control treatment. However, ANPP response to availabile soil
+than the control treatment. However, ANPP response to available soil
 moisture was weak and uncertain regardless of treatment, with all slopes
@@
 characterized by regressing historical observations of ANPP on
-observations of soil moisture. The fitted functional response can then
-be used to project ANPP under future precipitation regimes (REFERENCE).
-A problem with this approach is that it requires extrapolation if future
-precipitation falls outside outside the historical range of variability
-\citep{Smith2011, Peters2012}. For example, the soil moisture-ANPP
-relationship may be linear within the historical range of interannual
-variation, but could saturate at higher levels of soil moisture. In
-fact, saturating relationships are actually common
-\citep{Hsu2012, Gherardi2015a}, perhaps because other resources, like
-nitrogen, become more limiting in wet years than dry years. Failure to
-accurately estimate the curvature of the soil moisture-ANPP relationship
-will lead to over- or underprediction of ANNP under extreme
+observations of soil moisture. Fitted functional responses can then be
+used to infer how ANPP may change under future precipitation regimes
+\citep[e.g.,][]{Hsu2012}. A problem with this approach is that it
+requires extrapolation if future precipitation falls outside outside the
+historical range of variability \citep{Smith2011, Peters2012}. For
+example, the soil moisture-ANPP relationship may be linear within the
+historical range of interannual variation, but could saturate at higher
+levels of soil moisture. In fact, saturating relationships are actually
+common \citep{Hsu2012, Gherardi2015a}, perhaps because other resources,
+like nitrogen, become more limiting in wet years than dry years. Failure
+to accurately estimate the curvature of the soil moisture-ANPP
+relationship will lead to over- or underprediction of ANNP under extreme
 precipitation \citep{Peters2012}.
@@
 wavelength and then calculated a greenness index based on the same bands
-used to calculate NDVI using the MODIS and AVHRR {[}YOU HAVEN'T DEFINED
-THESE{]} bands for NDVI. We regressed the greenness index against the
-dry biomass weight from the ten calibration plots to convert the
-greenness index to ANPP. We fit regressions to a MODIS-based index and
-an AVHRR-based index for each year and retained the regression with the
-better fit based on \(R^2\) values. We then predicted ANPP using the
-best regression equation for each year (Appendix 1).
+used to calculate NDVI using the MODIS (Moderate Resolution Imaging
+Spectroradiometer) and AVHRR (Advanced Very High Resolution Radiometer)
+bands for NDVI. We regressed the greenness index against the dry biomass
+weight from the ten calibration plots to convert the greenness index to
+ANPP. We fit regressions to a MODIS-based index and an AVHRR-based index
+for each year and retained the regression with the better fit based on
+\(R^2\) values. We then predicted ANPP using the best regression
+equation for each year (Appendix 1).
 
@@
 
-These hypotheses are based on evidence that precipitation-ANPP
-relationships often saturate with increasing precipitation (REFERENCES).
+\noindent{}These hypotheses are based on evidence that
+precipitation-ANPP relationships often saturate with increasing
+precipitation \citep{Hsu2012, Gherardi2015a}.
 
@@
 \(\sigma_{\text{year}}\), which was drawn from a half-Cauchy
-distribution. A full description of our model is in Appendix 2.
+distribution. A full description of our model and associated R
+\citep{R2016} code is in Appendix 2.
 
@@
 
-Three of our five study years received below average precipitation (Fig.
-2A). Because our treatments are proportional changes in ambient
-precipitation, those three years represent a lower magnitude of absolute
-difference in precipitation among the treatments. ANPP varied from a
-minimum of 74.5 g m\(^{-2}\) in 2014 to a maximum of 237.1 g m\(^{-2}\)
-in 2016 when averaged across treatments (Fig. 2C). ANPP was slightly
-higher in irrigation plots and slightly lower in drought plots (Fig.
-2C), corresponding to estimated soil volumetric water content (VWC)
-differences among treatments (Fig. 2B). Such differences in soil VWC
-indicate our treatment infrastructure was successful. ANPP was highly
-variable across plots within years (Fig. 2C).
+Ambient precipitation was variable over the five years of the study
+(Fig. 2A), meaning our experiment captured a range of environmental
+conditions. ANPP varied from a minimum of 74.5 g m\(^{-2}\) in 2014 to a
+maximum of 237.1 g m\(^{-2}\) in 2016 when averaged across treatments
+(Fig. 2C). ANPP was slightly higher in irrigation plots and slightly
+lower in drought plots (Fig. 2C), corresponding to estimated soil
+volumetric water content (VWC) differences among treatments (Fig. 2B).
+Such differences in soil VWC indicate our treatment infrastructure was
+successful. ANPP was highly variable across plots within years (Fig.
+2C).
 
@@
 
-Community composition was similar among treatments. Community
-composition among treatments overlapped in all years and was equally
-dispersed in all years (Table 2; Fig. 5). Community composition was also
-remarkably stable over time, with no evidence of divergence among
-treatments (Table 2; Fig. 5).
+Community composition was similar among treatments. The multidimensional
+space of community composition overlapped among treatments in all years
+and was equally dispersed in all years (Table 2; Fig. 5). Community
+composition was also remarkably stable over time, with no evidence of
+divergence among treatments (Table 2; Fig. 5).
 
@@
 unchanged (Fig. 3A, Fig. 4), representing the top right scenario in Fig.
-1. The altered ecosystem functional response under irrigation matched
-our expectations because we found evidence for a weaker response for the
-irrigation treatment (Fig. 3). Although the difference between the
-control and irrigation slopes is likely different from zero (Fig. 3A),
-the posterior distribution of the slopes (not their difference) are
-similar and overlap broadly (Table1, Fig. A2-2). Thus, the statistical
-difference between the slopes is not ecologically meaningful. We
-therefore conclude that ecosystem functional response is consistent and
-weak across all precipitation treatments.
+1. The decrease in the slope of the VWC \(\times\) productivity
+relationship in the irrigated plots is consistent with a weaker response
+of productivity to additional water in this ecosystem. This suggests
+that if average soil moisture were pushed consistently higher than
+currently observed ambient conditions, there would be a weaker
+relationship between precipitation and productivity in this system.
+Nonetheless, the slopes of the VWC \(\times\) productivity relationship
+were similar among treatments (Table 1), at least at the level of
+irrigation we imposed. We therefore conclude that ecosystem functional
+response is consistent and weak across all precipitation treatments.
 
@@
 effect of soil moisture on ANPP (i.e., the Bayesian credible intervals
-for the slopes in \(\boldsymbol{\beta}\) in Eq. 1 overlapping zero).
+for the slopes overlapping zero; Table 1).
 

### 2017-09-25 manuscript/usses_water_ms.Rmd
trivia

@@
     base_format: rmarkdown::pdf_document # bookdown::pdf_document2 # for using \@ref()
+fontsize: 12pt
 ---

### 2017-09-28 manuscript/usses_water_ms.Rmd
adds journal paper number
fixes small things for PeerJ submission

@@
 Community composition was similar among treatments.
-The multidimensional space of community composition overlapped among treatments in all years and was equally dispersed in all years (Table 2; Fig. 5).
-Community composition was also remarkably stable over time, with no evidence of divergence among treatments (Table 2; Fig. 5).
+The multidimensional space of community composition overlapped among treatments in all years and was equally dispersed in all years (Table 2; Fig. 4).
+Community composition was also remarkably stable over time, with no evidence of divergence among treatments (Table 2; Fig. 4).
 
@@
 3) and community composition (Fig.
-5) among treatments is surprising because grasslands generally, and sagebrush steppe specifically, are considered water-limited systems.
+4) among treatments is surprising because grasslands generally, and sagebrush steppe specifically, are considered water-limited systems.
 Indeed, we expected ecosystem functional response, community composition, or both to change under our treatments.
@@
 
-In conclusion, our results suggest the species in our focal plant community are insensitive to to changes in precipitation regime. 
+# CONCLUSIONS
+
+Our results suggest the species in our focal plant community are insensitive to to changes in precipitation regime. 
 Such insensitivity can buffer species against precipitation variability in this semi-arid ecosystem, making them successful in the long run.
@@
 
-This research was supported by the Utah Agricultural Experiment Station, Utah State University, and approved as journal paper number XXXX.
+This research was supported by the Utah Agricultural Experiment Station, Utah State University, and approved as journal paper number 9035.
 The research was also supported by the National Science Foundation, through a Postdoctoral Research Fellowship in Biology and Mathematics to ATT (DBI-1400370), a Graduate Research Fellowship to ARK, and grants DEB-1353078 and DEB-1054040 to PBA.

### 2017-09-28 manuscript/usses_water_ms.tex
adds journal paper number
fixes small things for PeerJ submission

@@
 space of community composition overlapped among treatments in all years
-and was equally dispersed in all years (Table 2; Fig. 5). Community
+and was equally dispersed in all years (Table 2; Fig. 4). Community
 composition was also remarkably stable over time, with no evidence of
-divergence among treatments (Table 2; Fig. 5).
+divergence among treatments (Table 2; Fig. 4).
 
@@
 The similarity of ecosystem functional response (Fig. 3) and community
-composition (Fig. 5) among treatments is surprising because grasslands
+composition (Fig. 4) among treatments is surprising because grasslands
 generally, and sagebrush steppe specifically, are considered
@@
 
-In conclusion, our results suggest the species in our focal plant
-community are insensitive to to changes in precipitation regime. Such
-insensitivity can buffer species against precipitation variability in
-this semi-arid ecosystem, making them successful in the long run.
-Longer, chronic precipitation alteration might reveal plant community
-shifts that we did not observe \citep[e.g.,][]{Wilcox2016}. For example,
-a long-term increase in water availability could allow species that do
-not bet hedge to gain prominence and dominate the ecosystem functional
-response. Our results suggest compositional shifts would have the
-largest impact at high rainfall because the current community maintained
-consistent ecosystem functional response at very low water availability.
+\section{CONCLUSIONS}\label{conclusions}
+
+Our results suggest the species in our focal plant community are
+insensitive to to changes in precipitation regime. Such insensitivity
+can buffer species against precipitation variability in this semi-arid
+ecosystem, making them successful in the long run. Longer, chronic
+precipitation alteration might reveal plant community shifts that we did
+not observe \citep[e.g.,][]{Wilcox2016}. For example, a long-term
+increase in water availability could allow species that do not bet hedge
+to gain prominence and dominate the ecosystem functional response. Our
+results suggest compositional shifts would have the largest impact at
+high rainfall because the current community maintained consistent
+ecosystem functional response at very low water availability.
 
@@
 This research was supported by the Utah Agricultural Experiment Station,
-Utah State University, and approved as journal paper number XXXX. The
+Utah State University, and approved as journal paper number 9035. The
 research was also supported by the National Science Foundation, through

### 2017-10-31 manuscript/review_response_R1.Rmd
starts response letter

---
layout: 11pt
header-includes:
  - \usepackage{lineno}
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \linespread{1}
  - \usepackage{rotating}
  - \usepackage[usenames]{color}
  - \usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{tikz}
  - \usepackage{bm,mathrsfs}
  - \usepackage[sc]{mathpazo}
  - \usepackage{mathtools}
bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
csl: components/ecology_letters.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
    number_sections: true
fontsize: 11pt
geometry: margin=1in
linkcolor: black
urlcolor: black
---

\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{R1-\arabic{figure}}  
\renewcommand\thetable{R1-\arabic{table}}

\definecolor{ForestGreen}{rgb}{0.1,0.44,0.1} 
\definecolor{blue}{rgb}{0,0,0.7} 
\definecolor{red}{rgb}{0.8,0,0}

\newcommand{\new}{\textcolor{red}} 
\newcommand{\comment}{\textcolor{ForestGreen}}
\newcommand{\response}{\textcolor{blue}}

\begin{center}
\textbf{\Large{Response to Reviews}}
\end{center}

\begin{flushright}
Date: \today
\end{flushright}

Dr. Buckley,

Thank you for the opportunity to revise our manuscript.
All three reviewers agree our paper is interesting, but point out several areas where we could improve the analysis.
We were impressed with the very constructive reviews, and thank the reviewers for taking such care when reading our paper.
We hope our paper is now suitable for publication in \emph{PeerJ}.

In our point-by-point response below, the original reviews are in black and \response{our responses are in blue.} 
Likewise, changes made in the manuscript text are \response{in blue} for clarity.
In our responses, we refer to line numbers of the revised manuscript unless otherwise stated.

\bigskip{}

\normalsize 
\hrulefill

Manuscript number: 20739  
Title: Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe  
Author(s): Tredennick, A.T.; Kleinhesselink, A.R.; Taylor, J.B.;  Adler, P.B.

\hrulefill
\bigskip{} \bigskip{}

\textbf{Editors Comments for the Author(s)}

Please pay close attention to the comments from Reviewer 1 regarding the validity of the findings. Once you have responded to these comments in a point-by-point cover letter and using tracked changes in the manuscript document, I will likely send your manuscript out for re-review.

\response{We provide point-by-point responses to reviewer comments below. We have taken care to address Reviewer 1's concerns (also echoed by Reviewer 2) by }

\bigskip{}

\textbf{Reviewer: 1}

### 2017-10-31 code/01_calibrate_radiometer_by_year.R
adds Andy's fix for messed up factor labels in 2015

@@ -15,6 +15,8 @@ setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only for RStudio
 # recent years (2014 and beyond), biomass in 0.5m^2 plots (est biomass * 2)
 # earlier years (2013 and earlier), biomass in 0.25m^2 (est biomass * 4)
 
+
+
 ####
 ####  FILE PATHS, OUTPUT PATHS, ETC. ----
 ####
@@ -141,7 +143,7 @@ check_reps <- function(x){
   ids <- which(is_letter(x$rep))
   if(length(ids) > 0) {
     x     <- x[-ids, ]
-    x$rep <- as.integer(x$rep)
+    x$rep <- as.integer(levels(x$rep)[factor(x$rep)]) # recode factor labels
   }
   return(x)
 }
@@ -277,6 +279,8 @@ permanent_quad_biomass[which(permanent_quad_biomass$year > 2013), "biomass_grams
 ####
 ####  SAVE OUTPUT ----
 ####
+all_year_params
+
 write.csv(all_year_params, paste0(out_path,"ndvi_biomass_regression_parameters.csv"))
 saveRDS(permanent_quad_biomass, paste0(out_path,"permanent_plots_estimated_biomass.RDS"))
 #write.csv(permanent_quad_biomass, paste0(out_path,"permanent_plots_estimated_biomass.csv"))

### 2017-10-31 manuscript/review_response_R1.Rmd
adds all comments for response

@@
 
-\textbf{Reviewer: 1}
+**Reviewer: 1**
+
+*Basic reporting*
+
+The basic reporting is quite good.
+The manuscript is well-written and easy to follow.
+The statistics and pre-treatment analyses were, despite being complicated, easy to understand.
+No comments or concerns here.
+
+*Experimental design*
+
+The study was well executed.
+I particularly like the use of the Sala experimental design of automated drought+irrigation treatments.
+I think this is a powerful design that enables the authors to speak more thoroughly about the role of water limitation in this system (as opposed to a drought-only experiment).
+The author also did an extraordinary amount of work on the statistal analyses and to resolve issues (VWC measurements/ANPP calculations) in a quantitatively rigorous manner.
+
+*Validity of the findings*
+
+Their results indicate that the sagebrush steppe is actually quite robust to interannual variability in precipitation.
+This comes as a surprise, as the literature suggests that arid and semi-arid ecosystems (into which the sagebrush steppe easily fits) are quite sensitive to changes in precipitation.
+Huxman et al.
+(2004) and Knapp et al.
+(2015) both showed that drought impacts semi-arid sites more intensely than mesic sites, and Wilcox et al.
+(2017, in GCB) showed that most systems are more sensitive to irrigation than drought, and this is especially true for semi-arid sites.
+So far as I can tell, this study doesn't fit neatly into either line of thought (since it was relativel insensitive to drought, and irrigation effects were actually weaker than drought - see Wilcox et al.).
+It would be beneficial for the authors to put their study into this larger context and perhaps discuss why they think their site doesn't conform to these broader patterns (i.e.
+shrub-dominated, etc.).
+
+I also have a couple of concerns that potentially influence the results that the authors need to address prior to publication.
+
+1.
+I appreciate the authors effort in validating and calibrating their NDVI measurements to ANPP, but I am quite concerned about the accuracy of these calibrations.
+The $R^2$ is quite low for those calibration regressions (0.2 - 0.7), so I wonder how accurately biomass in each plot was represented by these ANPP measures.
+This is actually quite a large source of error, and I wonder if it isn't responsible for the lack of significance and weak responses observed here.
+I have two thoughts: would it be better to use raw NDVI measures rather than ANPP conversions?
+could the authors imbed the ANPP-NDVI calibration into their regressions (see Kiona Ogle's papers on multiple datasets in Bayesian analyses)?
+This would allow full uncertainty around the estimated ANPP based on NDVI to propogate into the VWC analyses.
+
+2.
+I didn't see mention of the priors on $\beta$, but I might've missed it.
+See Lemoine et al.
+(2016) - Ecology for a discussion of this issue.
+(very minor issue, I doubt the priors make too much difference here).
+
+3.
+My biggest concern is that the statistical tests don't seem to test the HRF because they treat year/time as a "nuisance" variable that adds noise to the observations.
+But the HRF is explicitly about time and how the response of ecosystem function changes over time.
+So incoporating time as an explicit part of the analysis would be necessary.
+The easiest way would be to regress ANPP against VWC and allow the slope to vary by year to see if the slopes follow the pattern dictated by HRF (or could use a hierarchical gaussian process/LOWESS model or other non-parametric model to allow for non-lienarity along the VWC gradient - drought - control - irrigation).
+Perhaps another approach would be to calculate a 'sensitivity' for each replicate (or treatment) within a Bayesian model, then look at how sensitivity to drought/irrigation changes through time (in the same model, i.e.
+plot-level sensitivity nested within year, with fixed effects to contrast sensitivity to drought/irrigation, a la Wilcox et al.
+2017).
+That is an explicit test of HRF, where the sensitivites could show any number of temporl patterns outlined by Smith.
+
+\bigskip{}
+
+**Reviewer: 2**
+
+*Basic reporting*
+
++ Well-presented manuscript. I appreciated the additional details provided in the appendices, particularly appendix 2.
++ A more general ecology audience may benefit if the introduction would start with an additional first paragraph to provide a framing of the research in a larger picture and to explain why the ANPP--soil moisture relationship is so important.
+
+*Experimental design*
+
++ I don’t understand the rationale for why the analysis is presented in terms of ANPP responses whereas NDVI was measured.
+The ANPP values represent a linear transformation of the NDVI measurements, but with a mediocre to low $R^2$ (A1-2), e.g., $R^2$ for year 2015 was 0.21 -- an information that is somewhat buried in the supplementary files.
+This less than ideal relationship is not discussed, e.g., as potential fourth source for not finding responses (lines 285-288).
+Why not analyzing NDVI--soil moisture relationships instead/in addition to ANPP--soil moisture relationships?
+
+*Validity of the findings*
+
++ The interpretation/discussion is coherent with the results. Even though the study found no responses, I agree that this is a reasonable outcome for a shrub-dominated, semi-arid ecosystem.
++ The introduction presents the idea that there are four basic types of how ANPP responds to increases/decreases in soil moisture and that the time it takes to express a response will differ among different communities (lines 87-90).
+Even though the discussion section suggests that species life history traits (‘bet hedging’, lines 310-319) are important to understand the results of no treatment responses, I feel that it would be as important to discuss how the lifespan of species in the community in relation to the experimental duration of the treatments promotes/hinders the observation of responses.
++ Several of the dominant species (e.g., threetip sagebrush and bluebunch wheatgrass) have a lifespan that is much longer than the 5-year experiment; while other species in the community have shorter reproductive cycles.
+It seems that even with a 5-year experiment the longer-lived species of the experimental community may not have had sufficient time to (fully) respond and make it possible to observe measurable changes in the community composition/response.
++ I would find it interesting if the plant community analysis is additionally carried out separately for short-lived vs long-lived species (Fig.
+4) -- based on the expectations that short-lived species have more ‘opportunities’ to respond/turnover than long-lived species.
+
+*Comments for the Author*
+
++ Fig. 3: Misspelling in ‘increasing kernal bandwidth’ -- this should be ‘kernel’
++ Fig 4: Expand caption: explain that dots represents plots and dimensions represent species.
+
+\bigskip{}
+
+**Reviewer: 3**
+
+*Basic reporting*
+
+English is high quality with very few typos, though a few locations in text could use rephrasing to increase clarity.
+Some restructuring of the introduction is suggested to provide a more logical and engaging presentation of the topic.
+Whenever I had a question about a topic the authors were referring to, or a method they were using, their citations were sufficient for me to conduct further investigation.
+More discussion of prior work into the role of dominant species in determining ecosystem function may be warranted.
+
+There are no significant departures from standard structure.
+
+Overall figures are well designed and easy to read, but treatment colors don’t translate well into greyscale.
+More contrasting colors (in terms of light and dark) are suggested.
+Minor things about a couple of table and figure headings could be changed to improve clarity.
+See notes to authors.
+
+Data is easy to find and download.
+
+Product is a self-contained research unit that doesn’t leave the reader wondering where the rest of the results are.
+
+*Experimental design*
+
+Primary research in Environmental science.
+
+The research goal is to investigate how the sagebrush steppe ecosystem may respond to long-term climate extremes outside of historic norms.
+Since most predictive work is currently done using long-term correlations between ecosystem function and climate, but the climate is changing this is a pressing concern.
+The inherent variability of ecological responses across ecosystems means that each ecosystem needs assessment.
+
+Design of the experiment looks very much like it was completely specified ahead of time and largely adhered to throughout.
+Where adjustments to the initial design were made, (probably in the soil moisture modeling) the procedure by which this was done was completely laid out.
+Though this is one of the areas I thought could use some rephrasing to increase clarity.
+The modeling work is excellent and looks at many different aspects of the question.
+
+Generally, the methods are quite detailed, including the description of the statistical analysis so reproduction should be possible.
+There were a few sections I thought could use additional clarification and these have been addressed in the comments to the authors.
+
+*Validity of the findings*
+
+Many of the findings of this paper were negative, i.e.
+no change in ecosystem function or composition was observed.
+The authors draw the appropriate conservative conclusions of no effect.
+They do however discuss the possible implications of the positive and negative lean in the data, which might make for interesting grounds for future research.
+
+Sample size is not large, though neither is it tiny.
+It was analyzed in such a way that information could still be learned despite the smaller sample size.
+Appropriate controls were included, though I would like a bit more information on why they used the controls they did, see comments to authors.
+
+Conclusions are clear and do not reach beyond the scope of the data.
+
+Authors speculated at length regarding why negative results were found.
+Concluding that it was likely a property of the ecosystem and probably not an artifact of the way the experiment was carried out, though they were careful to state that this possibility could not be ruled out.
+
+*Comments for the Author*
+
+This is a good article that yields insight into how experimental work can allow us to test whether historic precipitation/biomass relationships will told up to potential future changes in the precipitation regime.
+Overall, this is a very strong study and paper; it will make an excellent contribution to the scientific literature.
+However, there are some small issues that should be resolved first.
+There are some organizational issues in the Introduction; particularly, it seems the motivation for the study is buried deep within the text instead of in the first paragraph.
+Revising this section would make the paper a much more compelling read.
+Additionally, there were a number of places in the methods where I felt that additional clarity was needed so that others could replicate their study.
+Below I give line-by-line suggestions on these points and other minor editing suggestions.
+This reviewer feels that this paper should be accepted with minor revisions.
+
+General comments to authors
+
+Lines 31-32: sentence beginning with “However, ANPP response…” seems fairly vague and weakly worded, consider revising to increase clarity
+
+Introduction:
+
+While the abstract opens with a strong justification tying the paper into shifting precipitation patterns as a result of climate change and the need to develop a way to predict how communities might change with long-term exposure, the introduction opens only with a statement saying that there is a relationship between soil moisture and ANPP.
+This seems incredibly weak and uncompelling.
+The authors should consider restructuring their introduction so that the justification for their study is at the beginning.
+They have several other paragraphs in the paper that could do nicely as an opening paragraph.
+The opening paragraph of the discussion Lines 262-269 would make for an excellent opening.
+Or, with a little retooling the second paragraph of the introduction (lines 53-64) would make for a better statement of research motivation than the current opening.
+
+Line 46 “outside” is repeated twice
+
+Lines 71-72.
+While this is an excellent and clear definition of ‘ecosystem functional response’, I think it might be prudent at some point to explain why the authors are using soil moisture as an explanatory variable for ANPP instead of [precipitation].
+I agree that soil moisture is generally a better predictor of ANPP than direct precipitation measures, but most of the articles the authors cite refer to
+ANPP-precipitation relationships, so taking a moment to explain why they are using soil moisture would be warranted.
+
+Lines 88-90: another good motivating factor for why this research was done that could be used earlier in the introduction.
+Final paragraph on introduction. I really like the way the authors are testing multiple hypothesis based on theory.
+
+Methods:
+
+110-113: Could the authors please explain why these long-term monitoring plots make better controls than just say a generic 1 m2 plot within the exclosures next to the experimental plots.
+Certainly the long-term record on community shifts in these plots are likely interesting in their own right, but it doesn’t seem like they add much to the current study.
+
+Lines 126-127: It does concern me that the authors didn’t account for snow/winter precipitation on their irrigation plots.
+It appears as though a large percentage of the yearly precipitation in Dubois, ID falls over winter.
+Additionally, since their drought shelters were left up over winter they may actually have attracted drifting and captured more snow on the plots.
+To their credit, the authors did address this issue to some extent in the Discussion, lines 305-306.
+
+Lines 132-135: Wasn’t initially clear that these were manually collected data instead of another set of sensors.
+Maybe mention that they were collected by handheld probe at the beginning of the description instead of at the very end.
+
+Lines 142-147: Description of standardization procedure is confusing. I had to read it three times before the procedure was clear to me. Consider rewording for clarity.
+
+Lines 148-158: I’m a bit confused on which plots this modeling was conducted for.
+Was it done only on those plots that had permanent sensors in place?
+Or when possible was it also extrapolated to include those plots that only had point in time measurements.
+Please clarify.
+
+Lines 168-174: I like the idea of using the two separate estimates of greenness and then using the one that predicts best every year. Very thorough.
+
+Line 188: later, in figure 2, the authors refer to mean volumetric water content as “VWC” also.
+It would be best to use a different abbreviation for the two separate uses of volumetric water content: mean and cumulative.
+
+Results:
+
+Lines 241-244: Even though figures are referred to here, it might be nice to have some measure of the effect size in text, perhaps percentages, just to give a numerical estimation of differences, or lack there of, between plots in ANPP and soil moisture.
+
+Lines 257-260: the stability of the community over the study was impressive particularly given a method sensitive to rare species was used.
+It would be nice to see a bit more description of what the community looked like.
+How much of the community was the dominants, vs.
+rarer species?
+
+Discussion:
+
+Lines 279-281: I like that the authors were very conservative in their eventual conclusions despite their discussion of the potential meanings of the parameter estimates when these simply leaned positively or negatively and all overlapped zero.
+
+Lines 289-299: strong discussion and support for the sizes of the treatment applied. Good work.
+
+Lines 320-325: seems like authors are indicating that ecosystem response was driven by the response of the dominant species.
+Might be good to bring in some other work that has found similar results and discuss it a bit, for example: “Drivers of Variation in Aboveground Net Primary Productivity and Plant Community Composition Differ Across a Broad Precipitation Gradient” (La Pierre et al., 2016).
+
+Figures and tables:
+
+Table 1: These are the beta coefficents from the model correct? Might be good to elaborate a bit more in the table description.
+
+Figure 2: Here is the use of VWC as “mean soil volumetric water content” I referred to earlier.
+Like I said it would probably be best to have different abbreviations for different metrics of soil moisture.
+Also interesting, though not the point of this paper, after three consecutive dry years with progressively decreasing ANPP it looks like the ecosystem was primed for recovery as soon as an average precipitation year hit in 2015.
+No evidence of drought legacy.
+
+Figures general: differences between colors used to indicate treatments are not at all distinguishable if viewed in greyscale. Authors should consider using a color scheme with more contrast.

### 2017-11-02 manuscript/review_response_R1.Rmd
more on response letter

@@
 
+\response{Thanks!}
+
 *Experimental design*
@@
 The author also did an extraordinary amount of work on the statistal analyses and to resolve issues (VWC measurements/ANPP calculations) in a quantitatively rigorous manner.
+
+\response{Thanks again!}
 
@@
 (2017, in GCB) showed that most systems are more sensitive to irrigation than drought, and this is especially true for semi-arid sites.
-So far as I can tell, this study doesn't fit neatly into either line of thought (since it was relativel insensitive to drought, and irrigation effects were actually weaker than drought - see Wilcox et al.).
+So far as I can tell, this study doesn't fit neatly into either line of thought (since it was relatively insensitive to drought, and irrigation effects were actually weaker than drought - see Wilcox et al.).
 It would be beneficial for the authors to put their study into this larger context and perhaps discuss why they think their site doesn't conform to these broader patterns (i.e.
 shrub-dominated, etc.).
+
+\response{This is an excellent suggestion, and we have broadened the scope of our Discussion to place our findings in the context suggested by the reviewer.
+In particular, we discuss why our site might be different from the large scale patterns seen in other areas (lines xxx-xxx).}
 
@@
 
+\response{
+This is a thoughtful critique, and we thank the reviewer for pushing us to think deeper about our approach.
+We did initially consider embedding ANPP-NDVI calibration into the overall model, as suggested by the reviewer.
+But, in the end, we decided this would muddle the analysis and confuse the reader, which is why we kept it separate.
+In our revision, we have included a new appendix where we conduct our regression with NDVI rather than ANPP.
+We achieved the same results as when analyzing ANPP, and keep ANPP in the main text since that variable connects more easily to previous research (and what readers might be expecting).}
+
+\response{A beneficial consequence of this comment from the reviewer is that we thoroughly checked our code for converting NDVI to ANPP and found an error for one year.
+Now our lowest $R^2$ is 0.39, rather than 0.21 (see new Table A1-2).
+This is still low, which is why we conducted the comparison analysis described above.}
+
 2.
@@
 (very minor issue, I doubt the priors make too much difference here).
+
+\response{Sorry this was not clear. We included the prior specification [$\beta \sim \text{Normal}(0,5)$] in Appendix 2 on line 59. We have now included this in the main text on line xx.}
 
@@
 + Well-presented manuscript. I appreciated the additional details provided in the appendices, particularly appendix 2.
+
+\response{Thanks. We're glad you found our model description helpful.}
+
 + A more general ecology audience may benefit if the introduction would start with an additional first paragraph to provide a framing of the research in a larger picture and to explain why the ANPP--soil moisture relationship is so important.
+
+\response{Thanks for the suggestion. We have added a new introductory paragraph (lines xx-xx).}
 
@@
 
+\response{This comment echoes one of reviewer 1's concerns.
+As we mentioned above, we have conducted a model comparison where we show that results are consistent whether we use ANPP or NDVI (see new Appendix X).
+Thus, in the main text we retain the analysis focused on ANPP because we think it connects better to previous research.
+Also, as mentioned above, there was an error in our regression for the year 2015, which has been fixed (see new Table A1-2).}
+
 *Validity of the findings*
@@
 Even though the discussion section suggests that species life history traits (‘bet hedging’, lines 310-319) are important to understand the results of no treatment responses, I feel that it would be as important to discuss how the lifespan of species in the community in relation to the experimental duration of the treatments promotes/hinders the observation of responses.
+
+\response{This is an interesting point, and one we failed to discuss. We have added some text to our discussion on this topic (lines xxx-xxx).}
+
 + Several of the dominant species (e.g., threetip sagebrush and bluebunch wheatgrass) have a lifespan that is much longer than the 5-year experiment; while other species in the community have shorter reproductive cycles.
 It seems that even with a 5-year experiment the longer-lived species of the experimental community may not have had sufficient time to (fully) respond and make it possible to observe measurable changes in the community composition/response.
+
+\response{Yes, we agree that the rate of community turnover is likely related to the lifespans of constituent species.
+The physiological responses, however, should not depend on lifespan.
+Indeed, this it at the core of the Hierarchical Response Framework.
+We have added text discussing the impact of lifespan on community turnover (lines xxx-xxx).}
+
 + I would find it interesting if the plant community analysis is additionally carried out separately for short-lived vs long-lived species (Fig.
 4) -- based on the expectations that short-lived species have more ‘opportunities’ to respond/turnover than long-lived species.
+
+\response{Separate out annuals?}
 
@@
 + Fig. 3: Misspelling in ‘increasing kernal bandwidth’ -- this should be ‘kernel’
+
+\response{Good catch; fixed.}
+
 + Fig 4: Expand caption: explain that dots represents plots and dimensions represent species.
+
+\response{Good idea; done.}
 
@@
 
+\response{These are all good suggestions.
+Reviwer 2 agreed that the paper would benefit from a broader introduction, which we have added (lines xx-xx).
+We have also added a brief introduction to the idea that domoninant species play a large role in determining ecosystem functioning (lines xx-xx).}
+
 There are no significant departures from standard structure.
@@
 
+\response{We have increased the contrast between the colors so that it should be easier to distinguish colors if printed in black and white.
+The colors aren't perfect, but we anticipate most people will read this paper online.}
+
 Data is easy to find and download.
+
+\response{We're glad this was easy to access.}
 
@@
 
+\response{We have revised our description of modeling/extrapolating the discrete soil moisture measurements (lines xxx-xxx). Hopefully it is now easier to understand.}
+
 Generally, the methods are quite detailed, including the description of the statistical analysis so reproduction should be possible.
 There were a few sections I thought could use additional clarification and these have been addressed in the comments to the authors.
+
+\response{Thanks for reading our methods carefully and providing good suggestions for clarity.}
 
@@
 Appropriate controls were included, though I would like a bit more information on why they used the controls they did, see comments to authors.
+
+\response{Good question!
+We used the control plots we did because demographic data had been collected in these plots historically (e.g., the 1930's) and our other studies rely on comparing new demographic data to historic demographic data.
+We also use the demographic data to determine species composition.
+Given that this demographic data is time-consuming to collect, we decided to use the historical plots as controls rather than adding even more plots.}
 
@@
 
+\response{Thanks for these thoughtful comments. We address your specific points below.}
+
 General comments to authors
 
-Lines 31-32: sentence beginning with “However, ANPP response…” seems fairly vague and weakly worded, consider revising to increase clarity
+Lines 31-32: sentence beginning with “However, ANPP response…” seems fairly vague and weakly worded, consider revising to increase clarity.
+
+\response{We have revised this sentence to read, "xxx" (lines xx-xx).}
 
@@
 
-Line 46 “outside” is repeated twice
+\response{Fair point, and one that echoes Reviewer 2's critique of our introduction. We have added a new opening paragraph that we hope better sets the stage.}
+
+Line 46 "outside" is repeated twice
+
+\response{Good catch; extra "outside" is removed.}
 
@@
 While this is an excellent and clear definition of ‘ecosystem functional response’, I think it might be prudent at some point to explain why the authors are using soil moisture as an explanatory variable for ANPP instead of [precipitation].
-I agree that soil moisture is generally a better predictor of ANPP than direct precipitation measures, but most of the articles the authors cite refer to
-ANPP-precipitation relationships, so taking a moment to explain why they are using soil moisture would be warranted.
+I agree that soil moisture is generally a better predictor of ANPP than direct precipitation measures, but most of the articles the authors cite refer to ANPP-precipitation relationships, so taking a moment to explain why they are using soil moisture would be warranted.
+
+\response{Good point. We have added a sentence describing why we use soil moisture on lines xx-xx.}
 
 Lines 88-90: another good motivating factor for why this research was done that could be used earlier in the introduction.
+
+\response{Thanks for the suggestion. We have alluded to this in our new opening paragraph (lines xx-xx)}
+
 Final paragraph on introduction. I really like the way the authors are testing multiple hypothesis based on theory.
+
+\response{Thanks!}
 
@@
 
+\response{From above, "We used the control plots we did because demographic data had been collected in these plots historically (e.g., the 1930's) and our other studies rely on comparing new demographic data to historic demographic data.
+We also use the demographic data to determine species composition.
+Given that this demographic data is time-consuming to collect, we decided to use the historical plots as controls rather than adding even more plots."}
+
 Lines 126-127: It does concern me that the authors didn’t account for snow/winter precipitation on their irrigation plots.
@@
 
+\response{It concerns us, too.
+But, unfortunately, it was not feasible for us to control snow/winter precipitation.
+As the reviewer notes, we do consider this a limitation of our study (lines xxx-xxx).}
+
 Lines 132-135: Wasn’t initially clear that these were manually collected data instead of another set of sensors.
@@
 
+\response{Done.}
+
 Lines 142-147: Description of standardization procedure is confusing. I had to read it three times before the procedure was clear to me. Consider rewording for clarity.
+
+\response{Upon re-reading, we completely agree! We have substantially rewritten this section for clarity (lines xxx-xxx). Hopefully it is less confusing now.}
 
@@
 
+\response{We have revised this section to make this more clear (lines xxx-xxx).}
+
 Lines 168-174: I like the idea of using the two separate estimates of greenness and then using the one that predicts best every year. Very thorough.
+
+\response{Thank you.}
 
@@
 
+\response{This is a good catch. We actually only use \emph{cumulative} volumetric water content, and we have made this consistent throughout the manuscript.}
+
 Results:
@@
 Lines 241-244: Even though figures are referred to here, it might be nice to have some measure of the effect size in text, perhaps percentages, just to give a numerical estimation of differences, or lack there of, between plots in ANPP and soil moisture.
+
+\response{Good idea. We have added the effect sizes (lines xxx-xxx).}
 
@@
 
+\response{Good idea.
+We reran the community analysis with different subsets of species grouped by functional types/rarity.
+Our results are the same: community composition is really, really stable.
+We have added a mention of this in the methods and results sections, but keep our original figure in the main text.
+If the reviewer thinks it is necessary, we are happy to make a new appendix showing this analysis.}
+
 Discussion:
@@
 
+\response{Thank you.}
+
 Lines 289-299: strong discussion and support for the sizes of the treatment applied. Good work.
+
+\response{Thanks for the kind words.}
 
@@
 
+\response{Excellent suggestion. We have added this to the discussion on lines xxx-xxx, and also allude to dominant species' effects in the introduction on lines xx-xx.}
+
 Figures and tables:
@@
 Table 1: These are the beta coefficents from the model correct? Might be good to elaborate a bit more in the table description.
+
+\response{Yes, these are the beta coefficients. We have added detail to the table description.}
 
@@
 
+\response{Sorry about this mistake. We actually only use \emph{cumulative} volumetric water content and have corrected this in the manuscript.}
+
 Figures general: differences between colors used to indicate treatments are not at all distinguishable if viewed in greyscale. Authors should consider using a color scheme with more contrast.
+
+\response{We have increased the contrast between colors so that they are easier to distinguish if printed in black and white.}

### 2017-11-02 manuscript/usses_water_ms.Rmd
more on response letter

@@
 Fitted functional responses can then be used to infer how ANPP may change under future precipitation regimes [e.g., @Hsu2012]. 
-A problem with this approach is that it requires extrapolation if future precipitation falls outside outside the historical range of variability [@Smith2011; @Peters2012].
+A problem with this approach is that it requires extrapolation if future precipitation falls outside the historical range of variability [@Smith2011; @Peters2012].
 For example, the soil moisture-ANPP relationship may be linear within the historical range of interannual variation, but could saturate at higher levels of soil moisture. 
@@
 Probabilities (``Pr $\lessgtr$ 0 ='') for each coefficient indicate the one-tailed probability that the coefficient is less than or greater than zero, depending on whether the median of the distribution is less than or greater than zero.
-The posterior densities were smoothed for visual clarity by increasing kernal bandwidth by a factor of five.
+The posterior densities were smoothed for visual clarity by increasing kernel bandwidth by a factor of five.
 (B) Scatterplot of the data and model estimates shown as solid lines.

### 2017-11-02 manuscript/usses_water_ms.tex
more on response letter

@@
 \citep[e.g.,][]{Hsu2012}. A problem with this approach is that it
-requires extrapolation if future precipitation falls outside outside the
+requires extrapolation if future precipitation falls outside the
 historical range of variability \citep{Smith2011, Peters2012}. For

### 2017-11-02 manuscript/review_response_R1.Rmd
starts new sensitivity analysis; also ids a problem with ANPP value overlap between controls and treatments

@@
 
-\response{
-This is a thoughtful critique, and we thank the reviewer for pushing us to think deeper about our approach.
+\response{This is a thoughtful critique, and we thank the reviewer for pushing us to think deeper about our approach.
 We did initially consider embedding ANPP-NDVI calibration into the overall model, as suggested by the reviewer.
@@
 That is an explicit test of HRF, where the sensitivites could show any number of temporl patterns outlined by Smith.
+
+\response{The reviewer is correct that our analysis is not an explicit test of the Hierarchical Response Framework because we did not explicitly model the effect of time on ecosystem functional response.
+Unfortunately, we are unable to do so given our data -- in each year every plot shares the same average soil moisture for its particular treatment.
+Therefore, we cannot fit year-specific slopes.
+The sensitivity analysis would also be interesting, but moves us farther away from our interest in the ecosystem functional response.
+To remedy the situation, we have revised our motivation in the introduction to focus on the mechanisms that could alter ecosystem functional response, physiological responses and changes in community composition, without attributing specific time scales to them.}
 

### 2017-11-02 code/notebooks/sensitivity_analysis.R

@@ -1,0 +1,82 @@
+##  anpp_randcoefs_model.R: Script to run GLMM analysis to test for treatment 
+##  effects on the relationship between precipitation and ANPP.
+##
+##  NOTE: Stan may issue a couple warnings after running the MCMC that, as
+##  the messages state, can be safely ignored. Just rejects a couple proposals
+##  that result in ill-formed covariance matrices.
+##
+##  Author: Andrew Tredennick
+##  Date created: November
+
+##  Clear everything
+rm(list = ls(all.names = TRUE))
+
+##  Set working directory to source file location
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only for RStudio
+setwd("../")
+
+
+####
+####  LOAD LIBRARIES ----
+####
+library(tidyverse)    # Data munging
+library(dplyr)        # Data summarizing
+library(ggthemes)     # Pleasing ggplot themes
+library(stringr)      # Working with strings
+library(rstan)        # For MCMC
+library(lme4)         # Mixed-effects modeling
+
+
+
+####
+####  READ IN AND EXTRACT EXPERIMENT ANPP DATA ----
+####
+source("read_format_data.R")
+
+##  By treatment
+sens_anpp <- anpp_data %>%
+  group_by(Treatment, year) %>%
+  summarise(avg_anpp = mean(anpp)) %>%
+  spread(Treatment, avg_anpp) %>%
+  mutate(control_drought = Control - Drought,
+         control_irrigation = Control - Irrigation) %>%
+  select(-Control,-Drought,-Irrigation) %>%
+  gather(comparison, anpp_diff, -year)
+
+sens_vwc <- anpp_data %>%
+  group_by(Treatment, year) %>%
+  summarise(vwc = mean(total_seasonal_vwc)) %>%
+  spread(Treatment, vwc) %>%
+  mutate(control_drought = Control - Drought,
+         control_irrigation = Control - Irrigation) %>%
+  select(-Control,-Drought,-Irrigation) %>%
+  gather(comparison, vwc_diff, -year)
+
+sensitivity <- left_join(sens_anpp, sens_vwc, by = c("year", "comparison")) %>%
+  mutate(sens = anpp_diff / vwc_diff)
+
+ggplot(sensitivity, aes(x=year, y=sens, color=comparison))+
+  geom_line()
+
+
+##  By plot-treatment -- compare each treatment plot to each control in each year
[…]

### 2017-11-07 manuscript/Appendix3.Rmd
updates all model related stuff to ARK specification
starts revisions; adds new appendix 4; fixes figures after bug in NDVI calcs

@@
 \textbf{\Large{Appendix 3}} \\
-A.T. Tredennick, A.R. Kleinhesselink, \& P.B. Adler \\
-``Ecosystem and community resistance...'' \\
+A.T. Tredennick, A.R. Kleinhesselink, B. Taylor \& P.B. Adler \\
+``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe'' \\
 \emph{PeerJ}
@@
 
-\section{Characterizing Extreme Precipitation Amounts}
+\section{Conducting our analysis with NDVI}
 
-Following the proposed methods of @Lemoine2016, we calculated quantiles from the empirical distribution of growing season precipitation at Dubios, ID.
-We chose the 1\% quantile to be indicative of extreme dry conditions (drought) and the 99\% quantile to be indicative of extreme wet conditions (irrigation).
-The data consist of 91 yearly records, which we assume are approximately normally distributed.
-The R code below shows our procedure, and Fig. A3-1 shows the results.
+The first step of our analysis is to convert NDVI from field-based radiometric measurements to aboveground net primary productivity.
+However, this conversion is not perfect and the associated uncertainty is sometimes high (Table A1-2).
+Therefore, we also assessed ecosystem functional response across treatments using NDVI as the response variable rather than ANPP.
+The model essentially the same as the ANPP model, except we use a Beta likelihood for NDVI since its values range from 0 to 1, not including true 0s or 1s (see Stan code, below).
 
-```{r hrv, fig.cap="Density of the empirical distribution of growing season precipitation at Dubois, ID. Dashed vertical lines show the 1% and 99% quantiles, assuming a normal distribution."}
-library(ggplot2)
-weather    <- read.csv("../data/weather/ClimateIPM.csv")
-mean_ppt   <- mean(weather$ppt1)
-quants_ppt <- quantile(weather$ppt1,probs = c(0.01,0.99))
-quants_ppt[1]/mean_ppt*100 # percent of mean ppt for drought
-quants_ppt[2]/mean_ppt*100 # percent of mean ppt for irrigation
+Our NDVI-model results are very similar to the ANPP-model results (Figure A3-1).
+The only difference is there is a higher probability that the drought slope offset is less than 1 (90\% probability).
+However, these negative offsets do not result in significantly different slopes once applied to the control slope, just as in the ANPP-model.
 
-ggplot(weather, aes(x=ppt1))+
-  geom_histogram(bins=20, color="dodgerblue", fill="dodgerblue", aes(y=..density..))+
-  geom_line(stat="density", color="blue")+
-  geom_vline(aes(xintercept=quants_ppt[1]), linetype=2)+
-  geom_vline(aes(xintercept=quants_ppt[2]), linetype=2)+
-  ylab("Density")+
-  xlab("Growing Season Precipitation (mm)")+
+```{r plot_betas, fig.cap="Posterior distributions for the effects of drought and irrigation on the intercept and slope of the NDVI-soil moisture relationship.
+Treatment effects show the difference between the coefficients estimated in the treated plots and the control plots.
+Probabilities (Pr </> 0 =) for each coefficient indicate the one-tailed probability that the coefficient is less than or greater than zero, depending on whether the median of the distribution is less than or greater than zero.
+The posterior densities were smoothed for visual clarity by increasing kernel bandwidth by a factor of five.", include=T, echo=F, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
+
+library(tidyverse)    # Data munging
+library(dplyr)        # Data summarizing
+library(ggthemes)     # Pleasing ggplot themes
+library(ggjoy)        # Joy plots!
+library(stringr)      # Working with strings
+library(rstan)        # For MCMC and Stan objects
+library(gridExtra)    # For combining ggplot objects
+library(viridis)      # Pleasing color palette
+library(RColorBrewer) # More color palettes
+library(cowplot)      # For combining ggplots
+
+all_fit <- readRDS("../results/randcoefs_alltreatments_NDVIfit.RDS")
+
+####
+####  FUNCTION TO CALCULATE MAX(Pr(effect) > 0, Pr(effect) < 0) ----
+####
+get_one_tailed <- function(values){
+  above <- 1 - ecdf(values)(0)
+  below <- ecdf(values)(0)
+  max(above,below)
+}
+
+####
+####  PLOT TREATMENT-LEVEL POSTERIOR DISTRIBUTIONS ----
+####
+mycols <- c("#009E73", "#D55E00", "#0072B2")
+param_labels <- c("Control::Intercept",
+                  "Drought::Intercept", 
+                  "Irrigation::Intercept",
+                  "Control::Slope", 
+                  "Drought::Slope", 
+                  "Irrigation::Slope")
+
+betas <- data.frame(extract(all_fit, pars = 'beta')) %>% 
+  mutate(iteration = row_number()) %>% 
+  gather(param_name, estimate, starts_with('beta')) %>%
+  mutate(param_name = factor(param_name, labels = param_labels)) %>%
+  separate(param_name, c("Treatment", "Type"), "::")
+
+treat_slopes <- betas %>%
+  filter(Treatment != "Control") %>%
+  mutate(Type = paste(Type,"Offset"))
+
+slope_probs <- treat_slopes %>%
+  group_by(Treatment, Type) %>%
+  summarise(probs = round(get_one_tailed(estimate),2)) %>%
+  mutate(prob_text = paste("Pr < 0 =", probs))
+slope_probs$prob_text[1] <- paste("Pr > 0 =", slope_probs$probs[1]) # switch direction of drought intercpet since greater than 0
+slope_probs$xpos <- c(-2.8, -0.6, -2.6, -0.6)
+slope_probs$ypos <- c(1.3,1.3,2.55,2.7)
+
+treat_cols <- mycols[2:3]
+
+ggplot(treat_slopes)+
+  geom_vline(aes(xintercept=0), linetype=2, color="grey45")+
+  geom_joy(stat="density", adjust=3, alpha=0.8, aes(x=estimate, y=Treatment, fill=Treatment, color=Treatment, height = ..density..))+
+  geom_text(data = slope_probs, aes(x=xpos, y=ypos, label = prob_text), size = 3)+
+  facet_wrap(~Type, scales = "free")+
+  scale_y_discrete(labels = c("Drought","Irrigation"))+
+  scale_fill_manual(values = treat_cols)+
+  scale_color_manual(values = treat_cols)+
+  guides(fill=FALSE, color=FALSE)+
+  xlab("Coefficient Value")+
+  ylab(NULL)+
   theme_bw()+
-  theme(panel.grid.minor = element_blank())
+  theme(panel.grid.minor = element_blank(),
+        panel.border     = element_blank(), 
+        axis.line        = element_blank(),
+        axis.ticks       = element_blank(),
+        strip.background = element_blank(),
+        strip.text       = element_text(face="bold"),
+        panel.spacing    = unit(1, units = "cm"))
 ```
 
-# References
+\newpage{}
+
+\subsection{Stan Code}
+```{r stan1, echo=T, eval=F}
+data {
+  int<lower=0> Npreds;         # number of covariates, including intercept
+  int<lower=0> Npreds2;        # number of random effect covariates
+  int<lower=0> Nplots;         # number of plots
+  int<lower=0> Ntreats;        # number of treatments
+  int<lower=0> Nobs;           # number of observations
+  int<lower=0> Nyears;         # number of years
+  vector[Nobs] y;              # vector of observations
+  row_vector[Npreds] x[Nobs];	 # design matrix for fixed effects
+  row_vector[Npreds2] z[Nobs]; # simple design matrix for random effects 
+  int plot_id[Nobs];           # vector of plot ids
+  int treat_id[Nobs];          # vector of treatment ids
+  int year_id[Nobs];           # vector of year ids
+}
+
+parameters {
+	vector[Npreds] beta;	             # overall coefficients
+	vector[Nyears] year_off;           # vector of year effects
+	cholesky_factor_corr[Npreds2] L_u; # cholesky factor of plot random effect corr matrix
+  vector[Npreds2] beta_plot[Nplots]; # plot level random effects
+	vector<lower=0>[Npreds2] sigma_u;  # plot random effect std. dev.
+	real<lower=0> sd_y;                # treatment-level observation std. dev.
+	real<lower=0> sigma_year;          # year std. dev. hyperprior
+	real<lower=0> phi;                 # dispersion parameter
+}
+
+transformed parameters {
+  vector[Nobs] A;             # parameter for beta distn
+  vector[Nobs] B;             # parameter for beta distn
+  vector[Nobs] yhat;               # vector of expected values
+  vector[Npreds2] u[Nplots];       # transformed plot random effects
+  matrix[Npreds2,Npreds2] Sigma_u; # plot ranef cov matrix
+
+  Sigma_u = diag_pre_multiply(sigma_u, L_u); # cholesky factor for plot-level covariance matrix 
+  for(j in 1:Nplots)
+    u[j] = Sigma_u * beta_plot[j]; # plot random intercepts and slopes
+
+  # regression model for expected values (one for each plot-year)
+  for (i in 1:Nobs)
+    yhat[i] = inv_logit(x[i]*beta + z[i]*u[plot_id[i]] + year_off[year_id[i]]);
+
+  A = yhat * phi;
+  B = (1.0 - yhat) * phi;
+}
+
+model {
+  ####  PRIORS
+  phi ~ cauchy(0, 5); 
+  sigma_year ~ cauchy(0,2.5);
+  sd_y ~ cauchy(0,2.5);
+  year_off ~ normal(0,sigma_year); # priors on year effects, shared variance
+  beta ~ normal(0,5);	             # priors on treatment coefficients
+  L_u ~ lkj_corr_cholesky(2.0);      # prior on the cholesky factor which controls the 
+                                    # correlation between plot level treatment effects
+  sigma_u ~ cauchy(0,2.5);
+
+  for(i in 1:Nplots)
+		beta_plot[i] ~ normal(0,1); # plot-level coefficients for intercept and slope
+
+	####  LIKELIHOOD
+  y ~ beta(A, B); # observations vary according to beta distribution
+}
+
+generated quantities{
+  corr_matrix[Npreds2] R = multiply_lower_tri_self_transpose(L_u);
+  cov_matrix[Npreds2] V = quad_form_diag(R,sigma_u);
+}
+
+```

### 2017-11-07 manuscript/Appendix4.Rmd
starts revisions; adds new appendix 4; fixes figures after bug in NDVI calcs

---
layout: 11pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \onehalfspacing
  - \usepackage{rotating}
  - \usepackage{color, soul}
  - \usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{tikz}
  - \usepackage{bm,mathrsfs}
  - \usepackage[sc]{mathpazo}
bibliography: ~/Dropbox/Bibliography/usses_water.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
    number_sections: true
fontsize: 11pt
geometry: margin=1in
linkcolor: black
urlcolor: black
---

\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{A4-\arabic{figure}}
\renewcommand\theequation{A4.\arabic{equation}}
\renewcommand\thetable{A4-\arabic{table}}  
\renewcommand\thesection{Section A4.\arabic{section}}

\begin{center}
\textbf{\Large{Appendix 4}} \\
A.T. Tredennick, A.R. Kleinhesselink, B. Taylor \& P.B. Adler \\
``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe'' \\
\emph{PeerJ}
\end{center}

\section{Characterizing Extreme Precipitation Amounts}

Following the proposed methods of @Lemoine2016, we calculated quantiles from the empirical distribution of growing season precipitation at Dubios, ID.
We chose the 1\% quantile to be indicative of extreme dry conditions (drought) and the 99\% quantile to be indicative of extreme wet conditions (irrigation).
The data consist of 91 yearly records, which we assume are approximately normally distributed.
The R code below shows our procedure, and Fig. A3-1 shows the results.

```{r hrv, fig.cap="Density of the empirical distribution of growing season precipitation at Dubois, ID. Dashed vertical lines show the 1% and 99% quantiles, assuming a normal distribution."}
library(ggplot2)
weather    <- read.csv("../data/weather/ClimateIPM.csv")
mean_ppt   <- mean(weather$ppt1)
quants_ppt <- quantile(weather$ppt1,probs = c(0.01,0.99))
quants_ppt[1]/mean_ppt*100 # percent of mean ppt for drought
quants_ppt[2]/mean_ppt*100 # percent of mean ppt for irrigation

ggplot(weather, aes(x=ppt1))+
  geom_histogram(bins=20, color="dodgerblue", fill="dodgerblue", aes(y=..density..))+
  geom_line(stat="density", color="blue")+
  geom_vline(aes(xintercept=quants_ppt[1]), linetype=2)+
  geom_vline(aes(xintercept=quants_ppt[2]), linetype=2)+
  ylab("Density")+
  xlab("Growing Season Precipitation (mm)")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())
```

# References

### 2017-11-07 manuscript/usses_water_ms.Rmd
starts revisions; adds new appendix 4; fixes figures after bug in NDVI calcs

@@
 \definecolor{blue}{rgb}{0,0,0.7} 
-\definecolor{red}{rgb}{0.7,0,0}
-\newcommand{\pba}{\textcolor{blue}}
-\newcommand{\ark}{\textcolor{red}}
+\newcommand{\new}{\textcolor{blue}}
 
@@
 library("knitr")
-opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
+opts_chunk$set(tidy = FALSE, warning = FALSE, message = FALSE, 
                comment = NA, verbose = TRUE, echo=FALSE)
@@
 
-At any given site, the relationship between aboveground net primary productivity (ANPP) and water availability (e.g., soil moisture) can be characterized by regressing historical observations of ANPP on observations of soil moisture.
+\new{Water availability is a major driver of annual net primary productivity (ANPP) in grassland ecosystems} [@Huxman2004; @Hsu2012].
+\new{}
+
+At any given site, the relationship between ANPP and water availability (e.g., soil moisture) can be characterized by regressing historical observations of ANPP on observations of soil moisture. 
 Fitted functional responses can then be used to infer how ANPP may change under future precipitation regimes [e.g., @Hsu2012]. 
@@
 Sheep Experiment Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
-The plant community is dominated by the shrub *Artemesia tripartita* and three perennial bunchgrasses, *Pseudoroegneria spicata*, *Poa secunda*, and *Hesperostipa comata*.
+The plant community is dominated by the shrub \emph{Artemesia tripartita} and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata}.
 During the period of our experiment (2011 -- 2015), average mean annual precipitation was 265 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
@@
 In spring 2011, we established 16 new 1 m$^2$ plots located in the same exclosure as the six control plots.
-We avoided areas on steep hill slopes, areas with greater than 20\% cover of bare rock, and areas with greater than 10\% cover of the shrubs *Purshia tridentata* and/or *Amelanchier utahensis*.
+We avoided areas on steep hill slopes, areas with greater than 20\% cover of bare rock, and areas with greater than 10\% cover of the shrubs \emph{Purshia tridentata} and/or \emph{Amelanchier utahensis}.
 We established the new plots in pairs and randomly assigned each plot in a pair to receive a "drought" or "irrigation" treatment.
@@
 
-\noindent{}where $\mu_{i}$ is the deterministic prediction from the regression model for observation *i*, which is associated with plot *j* and treatment year *t*. 
+\noindent{}where $\mu_{i}$ is the deterministic prediction from the regression model for observation $i$, which is associated with plot $j$ and treatment year $t$. 
 $\boldsymbol{\beta}$ is the vector of coefficients for the fixed effects in the design matrix $\textbf{X}$. 
@@
 Specifically, we include plot-specific offsets ($\boldsymbol{\gamma}$) for the intercept and slope terms and year-specific intercept offsets ($\eta_t$). 
-The covariate vector $\textbf{z}_i$ for each observation *i* has two elements: a 1 for the intercept and the scaled value of VWC for that plot and year. 
+The covariate vector $\textbf{z}_i$ for each observation $i$ has two elements: a 1 for the intercept and the scaled value of VWC for that plot and year. 
 The plot-specific coefficients are modeled hierarchically, where plot level coefficients are drawn from a multivariate normal distribution with mean 0 and a variance-covariance structure that allows the intercept and slope terms to be correlated:
@@
 
-\noindent{}where $\Sigma$ is the variance-covariance matrix and *j(i)* reads as "plot *j* associated with observation *i*". 
+\noindent{}where $\Sigma$ is the variance-covariance matrix and $j(i)$ reads as "plot $j$ associated with observation $i$". 
 The random year effects ($\boldsymbol{\eta}$) are drawn from a normal prior with mean 0 and standard deviation $\sigma_{\text{year}}$, which was drawn from a half-Cauchy distribution. 
@@
 
-Cumulative March-June soil moisture had a weak positive effect on ANPP (Table 1; Fig. 3).
+Cumulative March-June soil moisture had a weak positive effect on ANPP (Table 1; Fig. 3B).
 The effect of soil moisture for each treatment is associated with high uncertainty, however, with 95\% Bayesian credible intervals that overlap zero (Table 1).
@@
 Ecosystem functional response was similar among treatments (Table 1; Fig. 3B), but there is evidence that the slope for the irrigation treatment is less than the slope for the control treatment.
-This evidence comes from interpreting the posterior distribution of the slope offset for the irrigation treatment, from which we calculate a 99% one-tailed probability that the estimate is less than zero (Fig.
+This evidence comes from interpreting the posterior distribution of the slope offset for the irrigation treatment, from which we calculate a \new{96}% one-tailed probability that the estimate is less than zero (Fig.
 3A, right panel).
-There was no evidence that the treatment effects became more important over time because there was no directional trend in the random year effects (Fig. A2-3).
 
@@
 That is, a 50\% decrease in any given year may not be abnormal given our site's historical range of variability [@Knapp2017].
-We cannot definitively rule out this possibility, but we have reason to believe our treatments *should* have been large enough.
+We cannot definitively rule out this possibility, but we have reason to believe our treatments \emph{should} have been large enough.
 Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of mean growing season precipitation necessary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site (Fig.
-A3-3).
-The 1\% quantile of precipitation at our site is 110 mm, a 47% reduction from the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean growing season precipitation (Appendix 3).
+A4-1).
+The 1\% quantile of precipitation at our site is 110 mm, a 47% reduction from the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean growing season precipitation (Appendix 4).
 Thus, our drought treatment represented extreme precipitation amounts, especially in years where ambient precipitation was below average (Fig. 2A).
@@
 For example, many of the perennial grasses in our focal ecosystem avoid drought stress by growing early in the growing season [@Blaisdell1958, A.R. Kleinhesselink, personal observation].
-Furthermore, the dominant shrub in our focal ecosystem, *Artemisia tripartita*, has access to water deep in the soil profile thanks to a deep root system [@Kulmatiski2017a].
-
-<!---
-The weaker soil moisture-ANPP relationship we observed for the irrigation treatment has three non-exclusive explanations.
-First, species' biomass production may have hit physiological thresholds and/or become limited by other factors, such as nitrogen availability [@LeBauer2008].
-\pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES?
-WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY OR UNLIKEL THIS IS...NUTNET SHOWS POSITIVE EFFECTS OF N*P ON PRODUCTION, EFFECT OF N ALONE NOT AS CLEAR.
-YOU COULD CITE THIS AS UNPUBLISHED DATA}
-
-Second, the less positive slope could be an indicator of bet hedging in this community.
-Avoiding high biomass and seed production in wet years could benefit species in the long run if climate conditions are variable.
-\pba{but why would this have cause a change relative to controls? this should also be true for controls...}
-Third, the less positive slope may be a statistical artifact due the variability of ANPP within years and our low sample size.
---->
+Furthermore, the dominant shrub in our focal ecosystem, \emph{Artemisia tripartita}, has access to water deep in the soil profile thanks to a deep root system [@Kulmatiski2017a].
 
@@
 \item [Appendix 2.] Details of the hierarchical Bayesian model, Fig. A2-1, Fig. A2-2, and Fig. A2-3.
-\item [Appendix 3.] Details on analysis of precipitation historical range of variability and Fig. A3-1.
+\item [Appendix 3.] \new{Results from analysis of NDVI without conversion to ANPP, Fig. A3-1.}
+\item [Appendix 4.] Details on analysis of precipitation historical range of variability and Fig. A4-1.
 \end{description}

### 2017-11-07 manuscript/usses_water_ms.tex
starts revisions; adds new appendix 4; fixes figures after bug in NDVI calcs

@@
 
-\definecolor{blue}{rgb}{0,0,0.7} \definecolor{red}{rgb}{0.7,0,0}
-\newcommand{\pba}{\textcolor{blue}} \newcommand{\ark}{\textcolor{red}}
+\definecolor{blue}{rgb}{0,0,0.7} \newcommand{\new}{\textcolor{blue}}
 
@@
 Dubois, Idaho (44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m above
-sea level. The plant community is dominated by the shrub \emph{Artemesia
-tripartita} and three perennial bunchgrasses, \emph{Pseudoroegneria
-spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata}. During the
-period of our experiment (2011 -- 2015), average mean annual
-precipitation was 265 mm year\(\phantom{}^{-1}\) and mean monthly
-temperature ranged from -5.2\(^{\circ}\)C in January to
-21.8\(^{\circ}\)C in July. Between 1926 and 1932, range scientists at
-the USSES established 26 permanent 1 m\(^2\) quadrats to track
-vegetation change over time. In 2007, we relocated 14 of the original
-quadrats, six of which were inside a large, permanent livestock
-exclosure. We use these six plots as control plots (i.e.~ambient
-precipitation) in the experiment described below.
+sea level. The plant community is dominated by the shrub
+\emph{Artemesia tripartita} and three perennial bunchgrasses,
+\emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and
+\emph{Hesperostipa comata}. During the period of our experiment (2011 --
+2015), average mean annual precipitation was 265 mm
+year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
+-5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July. Between 1926
+and 1932, range scientists at the USSES established 26 permanent 1
+m\(^2\) quadrats to track vegetation change over time. In 2007, we
+relocated 14 of the original quadrats, six of which were inside a large,
+permanent livestock exclosure. We use these six plots as control plots
+(i.e.~ambient precipitation) in the experiment described below.
 
@@
 \noindent{}where \(\mu_{i}\) is the deterministic prediction from the
-regression model for observation \emph{i}, which is associated with plot
-\emph{j} and treatment year \emph{t}. \(\boldsymbol{\beta}\) is the
-vector of coefficients for the fixed effects in the design matrix
-\(\textbf{X}\). Each row of the design matrix represents a single
-observation (\(\textbf{x}_i\)) and is a vector with the following
-elements: 1 for the intercept, a binary 0 or 1 if the treatment is
-``drought'', a binary 0 or 1 if the treatment is ``irrigation'', the
-scaled value of VWC, binary ``drought'' value times VWC, and binary
-``irrigation'' value times VWC. Thus, our model treats ``control''
-observations as the main treatment and then estimates intercept and
-slope offsets for the ``drought'' and ``irrigation'' treatments. We use
-our model to test two statistical hypotheses based on the questions in
-our Introduction:
+regression model for observation \(i\), which is associated with plot
+\(j\) and treatment year \(t\). \(\boldsymbol{\beta}\) is the vector of
+coefficients for the fixed effects in the design matrix \(\textbf{X}\).
+Each row of the design matrix represents a single observation
+(\(\textbf{x}_i\)) and is a vector with the following elements: 1 for
+the intercept, a binary 0 or 1 if the treatment is ``drought'', a binary
+0 or 1 if the treatment is ``irrigation'', the scaled value of VWC,
+binary ``drought'' value times VWC, and binary ``irrigation'' value
+times VWC. Thus, our model treats ``control'' observations as the main
+treatment and then estimates intercept and slope offsets for the
+``drought'' and ``irrigation'' treatments. We use our model to test two
+statistical hypotheses based on the questions in our Introduction:
 
@@
 slope terms and year-specific intercept offsets (\(\eta_t\)). The
-covariate vector \(\textbf{z}_i\) for each observation \emph{i} has two
+covariate vector \(\textbf{z}_i\) for each observation \(i\) has two
 elements: a 1 for the intercept and the scaled value of VWC for that
@@
 \noindent{}where \(\Sigma\) is the variance-covariance matrix and
-\emph{j(i)} reads as ``plot \emph{j} associated with observation
-\emph{i}''. The random year effects (\(\boldsymbol{\eta}\)) are drawn
-from a normal prior with mean 0 and standard deviation
-\(\sigma_{\text{year}}\), which was drawn from a half-Cauchy
-distribution. A full description of our model and associated R
-\citep{R2016} code is in Appendix 2.
+\(j(i)\) reads as ``plot \(j\) associated with observation \(i\)''. The
+random year effects (\(\boldsymbol{\eta}\)) are drawn from a normal
+prior with mean 0 and standard deviation \(\sigma_{\text{year}}\), which
+was drawn from a half-Cauchy distribution. A full description of our
+model and associated R \citep{R2016} code is in Appendix 2.
 
@@
 Cumulative March-June soil moisture had a weak positive effect on ANPP
-(Table 1; Fig. 3). The effect of soil moisture for each treatment is
+(Table 1; Fig. 3B). The effect of soil moisture for each treatment is
 associated with high uncertainty, however, with 95\% Bayesian credible
@@
 distribution of the slope offset for the irrigation treatment, from
-which we calculate a 99\% one-tailed probability that the estimate is
-less than zero (Fig. 3A, right panel). There was no evidence that the
-treatment effects became more important over time because there was no
-directional trend in the random year effects (Fig. A2-3).
+which we calculate a \new{96}\% one-tailed probability that the estimate
+is less than zero (Fig. 3A, right panel).
 
@@
 precipitation necessary to reach the 1\% and 99\% extremes of the
-historical precipitation regime at our site (Fig. A3-3). The 1\%
+historical precipitation regime at our site (Fig. A4-1). The 1\%
 quantile of precipitation at our site is 110 mm, a 47\% reduction from
 the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean
-growing season precipitation (Appendix 3). Thus, our drought treatment
+growing season precipitation (Appendix 4). Thus, our drought treatment
 represented extreme precipitation amounts, especially in years where
@@
   \hline
-Intercept & Control & 0.22 & 0.16 & -1.25 & 2.16 \\ 
-  Intercept & Drought & 0.64 & 0.44 & -1.37 & 3.79 \\ 
-  Intercept & Irrigation & -0.51 & -0.37 & -2.99 & 1.26 \\ 
-  Slope & Control & 1.31 & 1.14 & -0.30 & 3.83 \\ 
-  Slope & Drought & 1.18 & 1.00 & -0.62 & 4.12 \\ 
-  Slope & Irrigation & 0.83 & 0.68 & -0.61 & 3.12 \\ 
+Intercept & Control & 0.16 & 0.14 & -1.22 & 1.69 \\ 
+  Intercept & Drought & 0.13 & 0.05 & -1.69 & 2.39 \\ 
+  Intercept & Irrigation & -0.24 & -0.18 & -2.11 & 1.53 \\ 
+  Slope & Control & 0.94 & 0.86 & -0.51 & 2.69 \\ 
+  Slope & Drought & 0.65 & 0.57 & -0.99 & 2.62 \\ 
+  Slope & Irrigation & 0.72 & 0.65 & -0.54 & 2.26 \\ 
    \hline
@@
   2011 & betadisper &  21 &   2 & 2.23 & 0.14 \\ 
-  2012 & adonis &  22 &   2 & 1.10 & 0.32 \\ 
+  2012 & adonis &  22 &   2 & 1.10 & 0.34 \\ 
   2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
-  2013 & adonis &  22 &   2 & 1.23 & 0.14 \\ 
+  2013 & adonis &  22 &   2 & 1.23 & 0.16 \\ 
   2013 & betadisper &  22 &   2 & 0.28 & 0.76 \\ 
-  2014 & adonis &  22 &   2 & 0.95 & 0.58 \\ 
+  2014 & adonis &  22 &   2 & 0.95 & 0.56 \\ 
   2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
-  2015 & adonis &  21 &   2 & 1.05 & 0.40 \\ 
+  2015 & adonis &  21 &   2 & 1.05 & 0.37 \\ 
   2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 

### 2017-11-07 code/00_source_usses_water_scripts.R

@@ -51,6 +51,9 @@ source("04_plot_model_results.R")
 ####  6. RUN NMDS ANALYSIS AND PLOTS (Fig. 4, Table 2) ----
 source("05_community_dynamics.R")
 
+####  7. FIT NDVI-PRECIP MODEL (Appendix 3) ----
+source("06_ndvi_randcoefs_model.R")
+
 
 
 

### 2017-11-07 code/04_plot_model_results.R

@@ -72,7 +72,7 @@ slope_probs <- treat_slopes %>%
   summarise(probs = round(get_one_tailed(estimate),2)) %>%
   mutate(prob_text = paste("Pr < 0 =", probs))
 slope_probs$prob_text[1] <- paste("Pr > 0 =", slope_probs$probs[1]) # switch direction of drought intercpet since greater than 0
-slope_probs$xpos <- c(-2.8, -1, -3.6, -1.15)
+slope_probs$xpos <- c(-2.8, -0.7, -2.8, -0.7)
 slope_probs$ypos <- c(1.3,1.3,2.55,2.7)
 
 treat_cols <- mycols[2:3]

### 2017-11-07 code/06_ndvi_randcoefs_model.R

@@ -61,7 +61,7 @@ anppdat <- list(
   Nplots   = length(unique(anpp_data$quadname)),
   Ntreats  = length(unique(anpp_data$Treatment)),
   Nyears   = length(unique(anpp_data$year)),
-  y        = as.numeric(scale(log(anpp_data$ndvi))),
+  y        = as.numeric(anpp_data$ndvi),
   x        = x,
   z        = z,
   plot_id  = as.numeric(as.factor(anpp_data$quadname)),
@@ -75,7 +75,7 @@ options(mc.cores = parallel::detectCores())
 set.seed(123)
 
 ##  Fit the model in Stan
-rt  <- stanc("anpp_randcoefs.stan")
+rt  <- stanc("ndvi_randcoefs.stan")
 sm  <- stan_model(stanc_ret = rt, verbose = FALSE)
 fit <- sampling(sm, data = anppdat, iter = 10000, chains = 4, thin = 10)
 

### 2017-11-13 manuscript/usses_water_ms.Rmd
adds new opening paragraph

@@
 \new{Water availability is a major driver of annual net primary productivity (ANPP) in grassland ecosystems} [@Huxman2004; @Hsu2012].
-\new{}
+\new{Therefore, projected changes in precipitation regimes, associated with global climate change, will impact grassland ecosystem functioning.
+Increases in precipitation variability will result in prolonged periods of extremely wet or dry years, and predicting the consequences of such chronic resource alterations requires understanding how ecosystems respond to above and below average precipitation levels.}
 

### 2017-11-13 manuscript/usses_water_ms.tex
adds new opening paragraph

@@
 
-At any given site, the relationship between aboveground net primary
-productivity (ANPP) and water availability (e.g., soil moisture) can be
-characterized by regressing historical observations of ANPP on
-observations of soil moisture. Fitted functional responses can then be
-used to infer how ANPP may change under future precipitation regimes
-\citep[e.g.,][]{Hsu2012}. A problem with this approach is that it
-requires extrapolation if future precipitation falls outside the
-historical range of variability \citep{Smith2011, Peters2012}. For
-example, the soil moisture-ANPP relationship may be linear within the
-historical range of interannual variation, but could saturate at higher
-levels of soil moisture. In fact, saturating relationships are actually
-common \citep{Hsu2012, Gherardi2015a}, perhaps because other resources,
-like nitrogen, become more limiting in wet years than dry years. Failure
-to accurately estimate the curvature of the soil moisture-ANPP
-relationship will lead to over- or underprediction of ANNP under extreme
+\new{Water availability is a major driver of annual net primary productivity (ANPP) in grassland ecosystems}
+\citep{Huxman2004, Hsu2012}.
+\new{Therefore, projected changes in precipitation regimes, associated with global climate change, will impact grassland ecosystem functioning.
+Increases in precipitation variability will result in prolonged periods of extremely wet or dry years, and predicting the consequences of such chronic resource alterations requires understanding how ecosystems respond to above and below average precipitation levels.}
+
+At any given site, the relationship between ANPP and water availability
+(e.g., soil moisture) can be characterized by regressing historical
+observations of ANPP on observations of soil moisture. Fitted functional
+responses can then be used to infer how ANPP may change under future
+precipitation regimes \citep[e.g.,][]{Hsu2012}. A problem with this
+approach is that it requires extrapolation if future precipitation falls
+outside the historical range of variability
+\citep{Smith2011, Peters2012}. For example, the soil moisture-ANPP
+relationship may be linear within the historical range of interannual
+variation, but could saturate at higher levels of soil moisture. In
+fact, saturating relationships are actually common
+\citep{Hsu2012, Gherardi2015a}, perhaps because other resources, like
+nitrogen, become more limiting in wet years than dry years. Failure to
+accurately estimate the curvature of the soil moisture-ANPP relationship
+will lead to over- or underprediction of ANNP under extreme
 precipitation \citep{Peters2012}.

### 2017-11-13 manuscript/review_response_R1.Rmd
starts perennial and annual comp analysis

@@
 \response{This comment echoes one of reviewer 1's concerns.
-As we mentioned above, we have conducted a model comparison where we show that results are consistent whether we use ANPP or NDVI (see new Appendix X).
+As we mention above, we have conducted a model comparison where we show that results are consistent whether we use ANPP or NDVI (see new Appendix X).
 Thus, in the main text we retain the analysis focused on ANPP because we think it connects better to previous research.
@@
 
-\response{This is an interesting point, and one we failed to discuss. We have added some text to our discussion on this topic (lines xxx-xxx).}
+\response{This is an interesting point, and one we failed to discuss. We have added a sentence about this in our Conclusion section (lines xxx-xxx).}
 
@@
 Indeed, this it at the core of the Hierarchical Response Framework.
-We have added text discussing the impact of lifespan on community turnover (lines xxx-xxx).}
+We have added text discussing the impact of lifespan on community turnover in our Conclusion (lines xxx-xxx).}
 
 + I would find it interesting if the plant community analysis is additionally carried out separately for short-lived vs long-lived species (Fig.
-4) -- based on the expectations that short-lived species have more ‘opportunities’ to respond/turnover than long-lived species.
+4) -- based on the expectations that short-lived species have more `opportunities' to respond/turnover than long-lived species.
 
-\response{Separate out annuals?}
+\response{This is a great suggestion, and one we had not thought of.
+We conducted the analysis suggested by the reviewer and found no difference among perennials and annuals (Fig.
+R1).
+We have added this result in our Results section on lines xxx-xxx, but do not inlcude the figure.}
 
@@
 \response{These are all good suggestions.
-Reviwer 2 agreed that the paper would benefit from a broader introduction, which we have added (lines xx-xx).
+Reviewer 2 agreed that the paper would benefit from a broader introduction, which we have added (lines xx-xx).
 We have also added a brief introduction to the idea that domoninant species play a large role in determining ecosystem functioning (lines xx-xx).}

### 2017-11-13 manuscript/usses_water_ms.Rmd
starts perennial and annual comp analysis

@@
 For example, a long-term increase in water availability could allow species that do not bet hedge to gain prominence and dominate the ecosystem functional response.
+\new{Time \emph{per se} may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}
 Our results suggest compositional shifts would have the largest impact at high rainfall because the current community maintained consistent ecosystem functional response at very low water availability.

### 2017-11-13 code/05_community_dynamics.R

@@ -168,3 +168,98 @@ ggplot(nmds_df, aes(x=MDS1, y=MDS2, color=Treatment))+
   theme(panel.grid.major = element_line(color="grey90"))
 ggsave(paste0(figure_path,"sppcomp_bray_all.png"), width=6, height = 4, units = "in", dpi = 120)
 
+
+
+####
+####  LOOK AT PERENNIALS AND ANNUALS SEPARATELY ----
+####
+##  Perennials
+# Set missing species abundances as 0, which == absent
+density_community_matrix[is.na(density_community_matrix)] <- 0 
+
+density_community_matrix_scaled <- cbind(density_community_matrix[,1:3,drop=F], 
+                                      scale(density_community_matrix[,4:ncol(density_community_matrix)]))
+
+spp_to_remove <- names(which(is.nan(colMeans(density_community_matrix_scaled[4:ncol(density_community_matrix_scaled)]))))
+ids_to_remove <- which(colnames(density_community_matrix_scaled) %in% spp_to_remove)
+density_community_matrix_scaled <- density_community_matrix_scaled %>%
+  select(-ids_to_remove)
+
+
+nmds_df_perennials <- {}
+out_stats <- {}
+for(doyr in unique(density_community_matrix_scaled$year)){
+  tmp <- filter(density_community_matrix_scaled, year == doyr) %>% select(-year)
+  invisible( capture.output( # already checked for convergence
+    tmp_bray <- metaMDS(tmp[3:ncol(tmp)]+2, "bray", trymax = 999)
+  ))
+  nmds_points <- as.data.frame(tmp_bray$points)
+  nmds_points$Treatment <- tmp$Treatment
+  nmds_points$Year <- doyr
+  nmds_df_perennials <- rbind(nmds_df_perennials, nmds_points)
+  adonis_stats <- as.data.frame(vegan::adonis(tmp[3:ncol(tmp)]+2~tmp$Treatment)$aov.tab)[1,]
+  betadisper_stats <- anova(vegan::betadisper(vegan::vegdist(tmp[3:ncol(tmp)]+2),tmp$Treatment))[1,]
+  tmp_out <- rbind(adonis_stats[,-which(colnames(adonis_stats)=="R2")],
+                   as.numeric(betadisper_stats))
+  row.names(tmp_out) <- c()
+  out_stats <- rbind(out_stats, data.frame(year = doyr,
+                                           test = c("adonis","betadisper"),
[…]

### 2017-11-13 manuscript/review_response_R1.Rmd
more minor revisions

@@
 Reviewer 2 agreed that the paper would benefit from a broader introduction, which we have added (lines xx-xx).
-We have also added a brief introduction to the idea that domoninant species play a large role in determining ecosystem functioning (lines xx-xx).}
+We have also added a brief reference to the idea that domoninant species play a large role in determining ecosystem functioning (lines xx-xx).}
 
@@
 
-Lines 31-32: sentence beginning with “However, ANPP response…” seems fairly vague and weakly worded, consider revising to increase clarity.
+Lines 31-32: sentence beginning with "However, ANPP response…" seems fairly vague and weakly worded, consider revising to increase clarity.
 
-\response{We have revised this sentence to read, "xxx" (lines xx-xx).}
+\response{We have revised this sentence to read, "However, all estimates for the effect of soil moisture on ANPP overlapped zero, indicating the relationship is weak and uncertain regardless of treatment." (lines xx-xx).}
 
@@
 
-\response{Fair point, and one that echoes Reviewer 2's critique of our introduction. We have added a new opening paragraph that we hope better sets the stage.}
+\response{Fair point, and one that echoes Reviewer 2's critique of our introduction. We have added a new opening paragraph that we hope better sets the stage (lines xx-xx).}
 

### 2017-11-13 manuscript/usses_water_ms.Rmd
more minor revisions

@@
   **Results.** Ecosystem functional response, defined as the relationship between soil moisture and ANPP was similar among drought and control treatments, but the irrigation treatment had a lower slope than the control treatment.
-However, ANPP response to available soil moisture was weak and uncertain regardless of treatment, with all slopes overlapping zero.
-There was also large spatial variation in ANPP within-years.
+However, all estimates for the effect of soil moisture on ANPP overlapped zero, indicating the relationship is weak and uncertain regardless of treatment.
+There was also large spatial variation in ANPP within-years, which contributes to the uncertainty of the soil moisture effect.
 Plant community composition was remarkably stable over the course of the experiment and did not differ among treatments.
@@
 Soil moisture measurements were automatically logged every four hours.
-We coupled this temporally intensive soil moisture sampling with spatially extensive readings taken at six points within all 16 plots and associated ambient measurement areas.
-These snapshot data were collected on 06-06-2012, 04-29-2015, 05-07-2015, 06-09-2015, and 05-10-2016\footnote{Dates formatted as: mm-dd-yyyy.} using a handheld EC-5 sensor.
+We coupled this temporally intensive soil moisture sampling with spatially extensive readings \new{taken with a handheld EC-5 sensor} at six points within all 16 plots and associated ambient measurement areas.
+These snapshot data were collected on 06-06-2012, 04-29-2015, 05-07-2015, 06-09-2015, and 05-10-2016\footnote{Dates formatted as: mm-dd-yyyy.}.
 
@@
 
-We first averaged the observed soil moisture for each day and within each plot.
-Then we standardized the averages within each plot group by subtracting the average ambient soil moisture in that plot group and dividing by the standard deviation of the ambient soil moisture in that plot group.
+We first averaged the observed soil moisture for each day (*d*) and plot (*i*), $x_{i,d}$.
+\new{Then we standardized the average of each plot within a plot group ($g$) by subtracting the mean ambient soil moisture of the plot group for the day ($\overline{x}_{g,d} = \sum^{i=1}_G x_{i,d} / n$) from each plot's daily average and dividing by the standard deviation of the ambient soil moisture in that plot group ($\sigma_{g,d}$): $x^*_{i,d} = \frac{x_{i,d} - \overline{x}_{g,d}}{\sigma_{g,d}}$.}
 We then found the difference between the standardized ambient soil moisture and the standardized drought and irrigation soil moisture within each plot group.
@@
 We then predicted ANPP using the best regression equation for each year (Appendix 1).
+\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 2).}
 

### 2017-11-13 manuscript/usses_water_ms.tex
more minor revisions

@@
 and control treatments, but the irrigation treatment had a lower slope
-than the control treatment. However, ANPP response to available soil
-moisture was weak and uncertain regardless of treatment, with all slopes
-overlapping zero. There was also large spatial variation in ANPP
-within-years. Plant community composition was remarkably stable over the
-course of the experiment and did not differ among treatments.
+than the control treatment. However, all estimates for the effect of
+soil moisture on ANPP overlapped zero, indicating the relationship is
+weak and uncertain regardless of treatment. There was also large spatial
+variation in ANPP within-years, which contributes to the uncertainty of
+the soil moisture effect. Plant community composition was remarkably
+stable over the course of the experiment and did not differ among
+treatments.
 
@@
 temporally intensive soil moisture sampling with spatially extensive
-readings taken at six points within all 16 plots and associated ambient
-measurement areas. These snapshot data were collected on 06-06-2012,
-04-29-2015, 05-07-2015, 06-09-2015, and
-05-10-2016\footnote{Dates formatted as: mm-dd-yyyy.} using a handheld
-EC-5 sensor.
+readings \new{taken with a handheld EC-5 sensor} at six points within
+all 16 plots and associated ambient measurement areas. These snapshot
+data were collected on 06-06-2012, 04-29-2015, 05-07-2015, 06-09-2015,
+and 05-10-2016\footnote{Dates formatted as: mm-dd-yyyy.}.
 
@@
 
-We first averaged the observed soil moisture for each day and within
-each plot. Then we standardized the averages within each plot group by
-subtracting the average ambient soil moisture in that plot group and
-dividing by the standard deviation of the ambient soil moisture in that
-plot group. We then found the difference between the standardized
-ambient soil moisture and the standardized drought and irrigation soil
-moisture within each plot group. These transformations ensured that the
-treatment effects in each plot were appropriately scaled by the local
-ambient conditions within each plot group.
+We first averaged the observed soil moisture for each day (\emph{d}) and
+plot (\emph{i}), \(x_{i,d}\).
+\new{Then we standardized the average of each plot within a plot group ($g$) by subtracting the mean ambient soil moisture of the plot group for the day ($\overline{x}_{g,d} = \sum^{i=1}_G x_{i,d} / n$) from each plot's daily average and dividing by the standard deviation of the ambient soil moisture in that plot group ($\sigma_{g,d}$): $x^*_{i,d} = \frac{x_{i,d} - \overline{x}_{g,d}}{\sigma_{g,d}}$.}
+We then found the difference between the standardized ambient soil
+moisture and the standardized drought and irrigation soil moisture
+within each plot group. These transformations ensured that the treatment
+effects in each plot were appropriately scaled by the local ambient
+conditions within each plot group.
 
@@
 equation for each year (Appendix 1).
+\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 2).}
 
@@
 increase in water availability could allow species that do not bet hedge
-to gain prominence and dominate the ecosystem functional response. Our
-results suggest compositional shifts would have the largest impact at
-high rainfall because the current community maintained consistent
+to gain prominence and dominate the ecosystem functional response.
+\new{Time \emph{per se} may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}
+Our results suggest compositional shifts would have the largest impact
+at high rainfall because the current community maintained consistent
 ecosystem functional response at very low water availability.
@@
   \hline
-2011 & adonis &  21 &   2 & 1.02 & 0.44 \\ 
-  2011 & betadisper &  21 &   2 & 2.23 & 0.14 \\ 
-  2012 & adonis &  22 &   2 & 1.10 & 0.34 \\ 
+2011 & adonis &  21 &   2 & 1.02 & 0.43 \\ 
+  2011 & betadisper &  21 &   2 & 2.22 & 0.14 \\ 
+  2012 & adonis &  22 &   2 & 1.09 & 0.28 \\ 
   2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
-  2013 & adonis &  22 &   2 & 1.23 & 0.16 \\ 
-  2013 & betadisper &  22 &   2 & 0.28 & 0.76 \\ 
-  2014 & adonis &  22 &   2 & 0.95 & 0.56 \\ 
+  2013 & adonis &  22 &   2 & 1.25 & 0.16 \\ 
+  2013 & betadisper &  22 &   2 & 0.44 & 0.65 \\ 
+  2014 & adonis &  22 &   2 & 0.95 & 0.55 \\ 
   2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
-  2015 & adonis &  21 &   2 & 1.05 & 0.37 \\ 
+  2015 & adonis &  21 &   2 & 1.05 & 0.39 \\ 
   2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 
-  2016 & adonis &  21 &   2 & 1.07 & 0.36 \\ 
-  2016 & betadisper &  21 &   2 & 0.50 & 0.62 \\ 
+  2016 & adonis &  21 &   2 & 1.07 & 0.33 \\ 
+  2016 & betadisper &  21 &   2 & 0.49 & 0.62 \\ 
    \hline

### 2017-11-13 manuscript/review_response_R1.Rmd
more revisions and updates fig 2 VWC label

@@
 
-\response{Upon re-reading, we completely agree! We have substantially rewritten this section for clarity (lines xxx-xxx). Hopefully it is less confusing now.}
+\response{Upon re-reading, we completely agree!
+We have substantially rewritten this section for clarity and have added some mathematical notation to make our description precise (lines xxx-xxx).
+Hopefully it is less confusing now.}
 
@@
 
-Line 188: later, in figure 2, the authors refer to mean volumetric water content as “VWC” also.
+Line 188: later, in figure 2, the authors refer to mean volumetric water content as ``VWC'' also.
 It would be best to use a different abbreviation for the two separate uses of volumetric water content: mean and cumulative.
@@
 Lines 320-325: seems like authors are indicating that ecosystem response was driven by the response of the dominant species.
-Might be good to bring in some other work that has found similar results and discuss it a bit, for example: “Drivers of Variation in Aboveground Net Primary Productivity and Plant Community Composition Differ Across a Broad Precipitation Gradient” (La Pierre et al., 2016).
+Might be good to bring in some other work that has found similar results and discuss it a bit, for example: ``Drivers of Variation in Aboveground Net Primary Productivity and Plant Community Composition Differ Across a Broad Precipitation Gradient'' (La Pierre et al., 2016).
 
@@
 
-\response{Sorry about this mistake. We actually only use \emph{cumulative} volumetric water content and have corrected this in the manuscript.}
+\response{Sorry about this mistake.
+We use \emph{cumulative} March-June volumetric water content for the statistical model, but in this figure we plot the estimated daily values as a time series.
+We have updated the figure axis label and the caption to clarify.}
 
@@
 
-\response{We have increased the contrast between colors so that they are easier to distinguish if printed in black and white.}
+\response{We have increased the contrast between colors so that they are easier to distinguish if printed in black and white.
+They are not perfect, but they are from a color-blind friendly palette and we anticipate most readers of \emph{PeerJ} will read the paper online.}

### 2017-11-13 manuscript/usses_water_ms.Rmd
more revisions and updates fig 2 VWC label

@@
 
-We first averaged the observed soil moisture for each day (*d*) and plot (*i*), $x_{i,d}$.
-\new{Then we standardized the average of each plot within a plot group ($g$) by subtracting the mean ambient soil moisture of the plot group for the day ($\overline{x}_{g,d} = \sum^{i=1}_G x_{i,d} / n$) from each plot's daily average and dividing by the standard deviation of the ambient soil moisture in that plot group ($\sigma_{g,d}$): $x^*_{i,d} = \frac{x_{i,d} - \overline{x}_{g,d}}{\sigma_{g,d}}$.}
-We then found the difference between the standardized ambient soil moisture and the standardized drought and irrigation soil moisture within each plot group.
+\new{We first averaged the observed soil moisture for each day ($d$) and plot ($i$), $x_{i,d}$.
+Plots were assigned to plot groups consisting of a control, a drought, and an irrigation plot.
+We standardized the average of each plot within a plot group ($g$) by subtracting the mean ambient soil moisture of the plot group for the day ($\overline{x}_{g,d} = \sum^{i=1}_n x_{i,d} / n$) from each plot's daily average and dividing by the standard deviation of the ambient soil moisture in that plot group ($\sigma_{g,d}$): $x^*_{i,d} = \frac{x_{i,d} - \overline{x}_{g,d}}{\sigma_{g,d}}$.
 These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.
+We then found the difference between the standardized ambient soil moisture and the standardized drought and irrigation soil moisture within each plot group: $\Delta x^*_{\text{trt.}} = x^*_{\text{control}} - x^*_{\text{trt.}}$.}
 
-We then modeled the daily deviation from ambient conditions of the drought and irrigation treatments using a linear mixed effects model with independent variables for treatment, season (winter, spring, summer, fall), rainfall, and all two-way interactions.
+We then modeled the daily deviation \new{($\Delta x^*_\text{trt.}$)} from ambient conditions of the drought and irrigation treatments using a linear mixed effects model with independent variables for treatment, season (winter, spring, summer, fall), rainfall, and all two-way interactions.
 Rainy days were defined as any day in which precipitation was recorded and average temperature was above 3$^{\circ}$C.
@@
 We weighted observations by the number of unique sensors or spot measurements that were taken in each plot on that day.
-We then used the model to predict the average daily soil moisture in the treated plots based on the average daily ambient soil moisture.
+We then estimated average daily soil moisture in the treated plots based on the average daily ambient soil moisture \new{and the model predicted daily deviation for each treatment ($\Delta x^*_\text{trt.}$)}.
 We could only predict soil moisture in the treated plots on days for which we took at least one ambient soil moisture measurement.
@@
 \item [Appendix 1.] Additional methods and information on estimating aboveground net primary productivity.
-\item [Appendix 2.] Details of the hierarchical Bayesian model, Fig. A2-1, Fig. A2-2, and Fig. A2-3.
-\item [Appendix 3.] \new{Results from analysis of NDVI without conversion to ANPP, Fig. A3-1.}
+\item [Appendix 2.] \new{Results from analysis of NDVI without conversion to ANPP, Fig. A3-1.}
+\item [Appendix 3.] Details of the hierarchical Bayesian model, Fig. A2-1, Fig. A2-2, and Fig. A2-3.
 \item [Appendix 4.] Details on analysis of precipitation historical range of variability and Fig. A4-1.
@@
   \caption{(A) Probability density of historical precipitation from 1926-2016, with the years of the experiment shown with arrows on the \emph{x}-axis.
-(B) Mean soil volumetric water content (VWC) in each of the three treatments during the course of the experiment.
+(B) \new{Estimated daily} soil volumetric water content (VWC) in each of the three treatments during the course of the experiment.
 (C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.

### 2017-11-13 manuscript/usses_water_ms.tex
more revisions and updates fig 2 VWC label

@@
 
-We first averaged the observed soil moisture for each day (\emph{d}) and
-plot (\emph{i}), \(x_{i,d}\).
-\new{Then we standardized the average of each plot within a plot group ($g$) by subtracting the mean ambient soil moisture of the plot group for the day ($\overline{x}_{g,d} = \sum^{i=1}_G x_{i,d} / n$) from each plot's daily average and dividing by the standard deviation of the ambient soil moisture in that plot group ($\sigma_{g,d}$): $x^*_{i,d} = \frac{x_{i,d} - \overline{x}_{g,d}}{\sigma_{g,d}}$.}
-We then found the difference between the standardized ambient soil
-moisture and the standardized drought and irrigation soil moisture
-within each plot group. These transformations ensured that the treatment
-effects in each plot were appropriately scaled by the local ambient
-conditions within each plot group.
+\new{We first averaged the observed soil moisture for each day ($d$) and plot ($i$), $x_{i,d}$.
+Plots were assigned to plot groups consisting of a control, a drought, and an irrigation plot.
+We standardized the average of each plot within a plot group ($g$) by subtracting the mean ambient soil moisture of the plot group for the day ($\overline{x}_{g,d} = \sum^{i=1}_n x_{i,d} / n$) from each plot's daily average and dividing by the standard deviation of the ambient soil moisture in that plot group ($\sigma_{g,d}$): $x^*_{i,d} = \frac{x_{i,d} - \overline{x}_{g,d}}{\sigma_{g,d}}$.
+These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.
+We then found the difference between the standardized ambient soil moisture and the standardized drought and irrigation soil moisture within each plot group: $\Delta x^*_{\text{trt.}} = x^*_{\text{control}} - x^*_{\text{trt.}}$.}
 
-We then modeled the daily deviation from ambient conditions of the
-drought and irrigation treatments using a linear mixed effects model
-with independent variables for treatment, season (winter, spring,
-summer, fall), rainfall, and all two-way interactions. Rainy days were
-defined as any day in which precipitation was recorded and average
-temperature was above 3\(^{\circ}\)C. The day immediately following
-rainfall was also classified as rainy. We fit the model using the
-\texttt{lme4::lmer()} function \citep{Bates2015} in R \citep{R2016},
-with random effects for plot group and date. We weighted observations by
-the number of unique sensors or spot measurements that were taken in
-each plot on that day. We then used the model to predict the average
-daily soil moisture in the treated plots based on the average daily
-ambient soil moisture. We could only predict soil moisture in the
-treated plots on days for which we took at least one ambient soil
-moisture measurement. See \citet{Kleinhesselink2017b} for complete
-details on our approach to estimating soil moisture.
+We then modeled the daily deviation \new{($\Delta x^*_\text{trt.}$)}
+from ambient conditions of the drought and irrigation treatments using a
+linear mixed effects model with independent variables for treatment,
+season (winter, spring, summer, fall), rainfall, and all two-way
+interactions. Rainy days were defined as any day in which precipitation
+was recorded and average temperature was above 3\(^{\circ}\)C. The day
+immediately following rainfall was also classified as rainy. We fit the
+model using the \texttt{lme4::lmer()} function \citep{Bates2015} in R
+\citep{R2016}, with random effects for plot group and date. We weighted
+observations by the number of unique sensors or spot measurements that
+were taken in each plot on that day. We then estimated average daily
+soil moisture in the treated plots based on the average daily ambient
+soil moisture
+\new{and the model predicted daily deviation for each treatment ($\Delta x^*_\text{trt.}$)}.
+We could only predict soil moisture in the treated plots on days for
+which we took at least one ambient soil moisture measurement. See
+\citet{Kleinhesselink2017b} for complete details on our approach to
+estimating soil moisture.
 

### 2017-11-13 code/02_plot_data.R

@@ -57,17 +57,14 @@ mycols <- c("#009E73", "#D55E00", "#0072B2")
 soil_moisture <- read.csv("../data/soil_moisture_data/average_seasonal_soil_moisture.csv") %>%
   select(-year) %>%
   separate(simple_date, c("year","month","day")) %>%
-  # group_by(year,month,Treatment,type,year) %>%
-  # summarise(avg_vwc = mean(VWC,na.rm=TRUE)) %>%
   filter(type=="observed") %>%
-  #mutate(month_year = as.factor(paste0(year,"-",month))) %>%
   ungroup()
 
 suppressWarnings(# ignore warnings about missing values, we know they are empty
   soil_vwc <- ggplot(soil_moisture, aes(x=julian_date, y=VWC, group=Treatment, color=Treatment))+
     geom_line(size=0.5)+
     scale_color_manual(values = mycols, name="Treatment")+
-    ylab(expression(paste("Mean Soil VWC (ml ", ml^-1,")")))+
+    ylab(expression(paste("Daily Soil VWC (ml ", ml^-1,")")))+
     xlab("Julian Day")+
     scale_y_continuous(breaks=seq(0,24,8))+
     facet_grid(year~.)+

### 2017-11-26 manuscript/usses_water_ms.Rmd

@@
 
-\new{We first averaged the observed soil moisture for each day ($d$) and plot ($i$), $x_{i,d}$.
-Plots were assigned to plot groups consisting of a control, a drought, and an irrigation plot.
-We standardized the average of each plot within a plot group ($g$) by subtracting the mean ambient soil moisture of the plot group for the day ($\overline{x}_{g,d} = \sum^{i=1}_n x_{i,d} / n$) from each plot's daily average and dividing by the standard deviation of the ambient soil moisture in that plot group ($\sigma_{g,d}$): $x^*_{i,d} = \frac{x_{i,d} - \overline{x}_{g,d}}{\sigma_{g,d}}$.
+\new{We first averaged the observed soil moisture readings for each day ($d$) and plot ($i$), $x_{i,d}$.
+Plots were organized in groups of three, with each group ($g$) containing consisting of a control, a drought, and an irrigation plot.
+Within each group of plots we standardized the irrigation and drought effects on soil moisture relative to the ambient soil moisture conditions in the control plot.
+Specifically we subtracted the ambient daily soil moisture measured in the control plot from the soil moisture measured within the drought and irrigation plots within each group and then divided by the standard deviation of daily soil moisture values measured in the control plot ($\Delta x_{g,d,\text{trt.}} = (x_{g,d,\text{trt.}} - x_{g,d,\text{control}})/sd(x_{g,d,\text{control}})$) where $\Delta x$ is the standardized treatment effect and $x_{g,d,\text{trt.}}$ is the raw soil moisture measure for plot group $g$ on day $d$ and treatment $\text{trt.}$.
 These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.
-We then found the difference between the standardized ambient soil moisture and the standardized drought and irrigation soil moisture within each plot group: $\Delta x^*_{\text{trt.}} = x^*_{\text{control}} - x^*_{\text{trt.}}$.}
 
-We then modeled the daily deviation \new{($\Delta x^*_\text{trt.}$)} from ambient conditions of the drought and irrigation treatments using a linear mixed effects model with independent variables for treatment, season (winter, spring, summer, fall), rainfall, and all two-way interactions.
+We then modeled these daily deviations \new{($\Delta x\text{trt.}$)} from ambient conditions using a linear mixed effects model with independent variables for treatment (irrigation or drought), season (winter, spring, summer, fall), rainfall, and all two-way interactions.
 Rainy days were defined as any day in which precipitation was recorded and average temperature was above 3$^{\circ}$C.
@@
 We weighted observations by the number of unique sensors or spot measurements that were taken in each plot on that day.
-We then estimated average daily soil moisture in the treated plots based on the average daily ambient soil moisture \new{and the model predicted daily deviation for each treatment ($\Delta x^*_\text{trt.}$)}.
+We then used the model to predict the deviations from ambient conditions produced in the treated plots on each day of the experiment.  
+We added these predicted deviations to the average daily ambient soil moisture to generate predictions for daily soil moisture in all of the treated plots: $\bar{x}_{d, \text{trt.}} = \Delta \bar{x}_{d,\text{trt.}} + \bar{x}_{d,\text{control}}$, where $\bar{x}_{d,\text{trt.}}$ is the average predicted daily soil moisture in the treated plots and $\bar{x}_{d,\text{control}}$ is the daily ambient soil moisture averaged across all control plots.
 We could only predict soil moisture in the treated plots on days for which we took at least one ambient soil moisture measurement.
+This procedure allowed us to predict daily soil moisture conditions for all plots, even on those days when some of our direct treatment measurements were missing due to malfunction of the sensors or dataloggers.
 See @Kleinhesselink2017b for complete details on our approach to estimating soil moisture.

### 2017-11-26 manuscript/usses_water_ms.Rmd
cleaning up merge for soil moisture modeling

@@
 Specifically we subtracted the ambient daily soil moisture measured in the control plot from the soil moisture measured within the drought and irrigation plots within each group and then divided by the standard deviation of daily soil moisture values measured in the control plot ($\Delta x_{g,d,\text{trt.}} = (x_{g,d,\text{trt.}} - x_{g,d,\text{control}})/sd(x_{g,d,\text{control}})$) where $\Delta x$ is the standardized treatment effect and $x_{g,d,\text{trt.}}$ is the raw soil moisture measure for plot group $g$ on day $d$ and treatment $\text{trt.}$.
-These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.
+These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.}
 
-We then modeled these daily deviations \new{($\Delta x\text{trt.}$)} from ambient conditions using a linear mixed effects model with independent variables for treatment (irrigation or drought), season (winter, spring, summer, fall), rainfall, and all two-way interactions.
+\new{We then modeled these daily deviations ($\Delta x_\text{trt.}$)} from ambient conditions using a linear mixed effects model with independent variables for treatment (irrigation or drought), season (winter, spring, summer, fall), rainfall, and all two-way interactions.
 Rainy days were defined as any day in which precipitation was recorded and average temperature was above 3$^{\circ}$C.
@@
 We weighted observations by the number of unique sensors or spot measurements that were taken in each plot on that day.
-We then used the model to predict the deviations from ambient conditions produced in the treated plots on each day of the experiment.  
-We added these predicted deviations to the average daily ambient soil moisture to generate predictions for daily soil moisture in all of the treated plots: $\bar{x}_{d, \text{trt.}} = \Delta \bar{x}_{d,\text{trt.}} + \bar{x}_{d,\text{control}}$, where $\bar{x}_{d,\text{trt.}}$ is the average predicted daily soil moisture in the treated plots and $\bar{x}_{d,\text{control}}$ is the daily ambient soil moisture averaged across all control plots.
+\new{We then used the model to predict the deviations from ambient conditions produced in the treated plots on each day of the experiment.  
+We added these predicted deviations to the average daily ambient soil moisture to generate predictions for daily soil moisture in all of the treated plots: $\bar{x}_{d, \text{trt.}} = \Delta \bar{x}_{d,\text{trt.}} + \bar{x}_{d,\text{control}}$, where $\bar{x}_{d,\text{trt.}}$ is the average predicted daily soil moisture in the treated plots and $\bar{x}_{d,\text{control}}$ is the daily ambient soil moisture averaged across all control plots.}
 We could only predict soil moisture in the treated plots on days for which we took at least one ambient soil moisture measurement.

### 2017-11-26 manuscript/usses_water_ms.tex
cleaning up merge for soil moisture modeling

@@
 
-\new{We first averaged the observed soil moisture for each day ($d$) and plot ($i$), $x_{i,d}$.
-Plots were assigned to plot groups consisting of a control, a drought, and an irrigation plot.
-We standardized the average of each plot within a plot group ($g$) by subtracting the mean ambient soil moisture of the plot group for the day ($\overline{x}_{g,d} = \sum^{i=1}_n x_{i,d} / n$) from each plot's daily average and dividing by the standard deviation of the ambient soil moisture in that plot group ($\sigma_{g,d}$): $x^*_{i,d} = \frac{x_{i,d} - \overline{x}_{g,d}}{\sigma_{g,d}}$.
-These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.
-We then found the difference between the standardized ambient soil moisture and the standardized drought and irrigation soil moisture within each plot group: $\Delta x^*_{\text{trt.}} = x^*_{\text{control}} - x^*_{\text{trt.}}$.}
+\new{We first averaged the observed soil moisture readings for each day ($d$) and plot ($i$), $x_{i,d}$.
+Plots were organized in groups of three, with each group ($g$) containing consisting of a control, a drought, and an irrigation plot.
+Within each group of plots we standardized the irrigation and drought effects on soil moisture relative to the ambient soil moisture conditions in the control plot.
+Specifically we subtracted the ambient daily soil moisture measured in the control plot from the soil moisture measured within the drought and irrigation plots within each group and then divided by the standard deviation of daily soil moisture values measured in the control plot ($\Delta x_{g,d,\text{trt.}} = (x_{g,d,\text{trt.}} - x_{g,d,\text{control}})/sd(x_{g,d,\text{control}})$) where $\Delta x$ is the standardized treatment effect and $x_{g,d,\text{trt.}}$ is the raw soil moisture measure for plot group $g$ on day $d$ and treatment $\text{trt.}$.
+These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.}
 
-We then modeled the daily deviation \new{($\Delta x^*_\text{trt.}$)}
-from ambient conditions of the drought and irrigation treatments using a
-linear mixed effects model with independent variables for treatment,
-season (winter, spring, summer, fall), rainfall, and all two-way
-interactions. Rainy days were defined as any day in which precipitation
-was recorded and average temperature was above 3\(^{\circ}\)C. The day
-immediately following rainfall was also classified as rainy. We fit the
-model using the \texttt{lme4::lmer()} function \citep{Bates2015} in R
-\citep{R2016}, with random effects for plot group and date. We weighted
-observations by the number of unique sensors or spot measurements that
-were taken in each plot on that day. We then estimated average daily
-soil moisture in the treated plots based on the average daily ambient
-soil moisture
-\new{and the model predicted daily deviation for each treatment ($\Delta x^*_\text{trt.}$)}.
+\new{We then modeled these daily deviations ($\Delta x_\text{trt.}$)}
+from ambient conditions using a linear mixed effects model with
+independent variables for treatment (irrigation or drought), season
+(winter, spring, summer, fall), rainfall, and all two-way interactions.
+Rainy days were defined as any day in which precipitation was recorded
+and average temperature was above 3\(^{\circ}\)C. The day immediately
+following rainfall was also classified as rainy. We fit the model using
+the \texttt{lme4::lmer()} function \citep{Bates2015} in R \citep{R2016},
+with random effects for plot group and date. We weighted observations by
+the number of unique sensors or spot measurements that were taken in
+each plot on that day.
+\new{We then used the model to predict the deviations from ambient conditions produced in the treated plots on each day of the experiment.  
+We added these predicted deviations to the average daily ambient soil moisture to generate predictions for daily soil moisture in all of the treated plots: $\bar{x}_{d, \text{trt.}} = \Delta \bar{x}_{d,\text{trt.}} + \bar{x}_{d,\text{control}}$, where $\bar{x}_{d,\text{trt.}}$ is the average predicted daily soil moisture in the treated plots and $\bar{x}_{d,\text{control}}$ is the daily ambient soil moisture averaged across all control plots.}
 We could only predict soil moisture in the treated plots on days for
-which we took at least one ambient soil moisture measurement. See
+which we took at least one ambient soil moisture measurement. This
+procedure allowed us to predict daily soil moisture conditions for all
+plots, even on those days when some of our direct treatment measurements
+were missing due to malfunction of the sensors or dataloggers. See
 \citet{Kleinhesselink2017b} for complete details on our approach to
@@
   \hline
-2011 & adonis &  21 &   2 & 1.02 & 0.43 \\ 
+2011 & adonis &  21 &   2 & 1.02 & 0.45 \\ 
   2011 & betadisper &  21 &   2 & 2.22 & 0.14 \\ 
-  2012 & adonis &  22 &   2 & 1.09 & 0.28 \\ 
+  2012 & adonis &  22 &   2 & 1.09 & 0.31 \\ 
   2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
-  2013 & adonis &  22 &   2 & 1.25 & 0.16 \\ 
+  2013 & adonis &  22 &   2 & 1.25 & 0.11 \\ 
   2013 & betadisper &  22 &   2 & 0.44 & 0.65 \\ 
@@
   2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
-  2015 & adonis &  21 &   2 & 1.05 & 0.39 \\ 
+  2015 & adonis &  21 &   2 & 1.05 & 0.38 \\ 
   2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 

### 2017-12-06 manuscript/Appendix5.Rmd
adds new appendix on plant community details and dynamics, mainly RACs

---
layout: 11pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \onehalfspacing
  - \usepackage{rotating}
  - \usepackage{color, soul}
  - \usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{tikz}
  - \usepackage{bm,mathrsfs}
  - \usepackage[sc]{mathpazo}
bibliography: ~/Dropbox/Bibliography/usses_water.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
    number_sections: true
fontsize: 11pt
geometry: margin=1in
linkcolor: black
urlcolor: black
---

\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{Ax-\arabic{figure}}
\renewcommand\theequation{Ax.\arabic{equation}}
\renewcommand\thetable{Ax-\arabic{table}}  
\renewcommand\thesection{Section Ax\arabic{section}}

\begin{center}
\textbf{\Large{Appendix X}} \\
A.T. Tredennick, A.R. Kleinhesselink, B. Taylor \& P.B. Adler \\
``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe'' \\
\emph{PeerJ}
\end{center}

# Details on the plant community and dynamics

Here we provide more details on the plant community in terms of dominance and rarity.
Averaging across time, *Artemisia tripartita* and *Balsamorhiza sagittata* are the two most dominant species in each treatment.
Combined, these two species represent 28\% of total cover in control plots, 25\% of total cover in drought plots, and 25\% of total cover in irrigation plots.
Four to five species dominate the community in general (Figure Ax-1), indicating a high level of dominance in this plant community.

```{r get_data, echo=FALSE, message=FALSE, warning=FALSE}
####
####  LOAD PACKAGES ----
####
library(tidyverse) # all the good things from Wikham
library(dplyr)     # data munging
library(stringr)   # working with strings
library(ggthemes)  # pleasing ggplot2 themes

####
####  SET DIRECTORIES; READ IN DATA ----
####
# Set paths
biomass_path <- "../data/estimated_biomass/"
weather_path <- "../data/weather/"
output_path  <- "../output/"
figure_path  <- "../figures/"

# Read in data
plot_biomass <- readRDS(paste0(biomass_path,"permanent_plots_estimated_biomass.RDS"))
weather_data <- read.csv(paste0(weather_path,"ClimateIPM.csv"))
plot_info    <- read.csv("../data/estimated_biomass/quad_info.csv")

suppressWarnings( # ignore factor to character warning. It's OK.
  cover_species <- read.csv("../data/vital_rates/allrecords_cover_v24.csv") %>%
    group_by(quad, year, species) %>%
    summarise(total_area = sum(area)) %>%
    left_join(plot_info, by = "quad") %>%
    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
    filter(!str_detect(QuadName, 'P1|P7')) %>%
    group_by(year, species, Treatment) %>%
    summarise(avg_area = mean(total_area)) %>%
    ungroup()
)

####
####  COMPARE COMMUNITY COMPOSITION AMONG TREATMENTS ----
####  USING STANDARDIZED ABUNDANCES FROM COVER AND DENSITY
####
suppressWarnings( # ignore factor to character warning. It's OK.
  cover_community_matrix <- read.csv("../data/vital_rates/allrecords_cover_v24.csv") %>%
    group_by(quad, year, species) %>%
    summarise(total_area = sum(area)) %>%
    left_join(plot_info, by = "quad") %>%
    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
    filter(!str_detect(QuadName, 'P1|P7')) %>%
    filter(year > 2010 & is.na(species)==F) %>%
    spread(species,total_area, fill = 0) %>%
    select(-QuadName,-Grazing,-paddock,-Group) %>%
    ungroup()
)

suppressWarnings( # ignore factor to character warning. It's OK.
  density_community_matrix <- read.csv("../data/vital_rates/allrecords_density_v24.csv") %>%
    mutate(ind_num = 1) %>%
    group_by(year, quad, species) %>%
    summarise(total_inds = sum(ind_num)) %>%
    left_join(plot_info, by = "quad") %>%
    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
    filter(!str_detect(QuadName, 'P1|P7')) %>%
    filter(year > 2010 & is.na(species)==F) %>%
    spread(species,total_inds, fill = 0) %>%
    select(-QuadName,-Grazing,-paddock,-Group) %>%
    ungroup()
)

density_spp <- colnames(density_community_matrix[4:ncol(density_community_matrix)])
cover_spp   <- colnames(cover_community_matrix[4:ncol(cover_community_matrix)])
overlap_spp <- density_spp[which(density_spp %in% cover_spp)]
rmids       <- which(colnames(density_community_matrix) %in% overlap_spp)
density_community_matrix <- density_community_matrix[,-rmids]

suppressWarnings( # ignore factor to character warning. It's OK.
  annual_community_matrix <- read.csv("../data/vital_rates/idaho_annuals.csv") %>%
    select(-Notes) %>%
    gather(species,abundance,-c(1:3)) %>%
    left_join(plot_info, by="quad") %>%
    select(-QuadName.y) %>%
    rename(QuadName = QuadName.x) %>%
    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
    filter(!str_detect(QuadName, 'P1|P7')) %>%
    filter(year > 2010 & is.na(species)==F) %>%
    separate(species,c("genus","species1")) %>%
    unite(species,genus,species1,sep = " ") %>%
    spread(species,abundance, fill = 0) %>%
    select(-QuadName,-Grazing,-paddock,-Group,-`Lappula occidentalis`,-`Tragopogon dubius`) %>%
    ungroup()
)

annual_data <- {}
for(dotrt in unique(annual_community_matrix$Treatment)){
  for(doyr in unique(annual_community_matrix$year)){
    tmp <- annual_community_matrix %>%
      gather(species, density, -quad, -year, -Treatment) %>%
      filter(Treatment == dotrt, year == doyr) %>%
      group_by(Treatment, year, species) %>%
      summarise(mean_density = mean(density)) %>%
      arrange(-mean_density) %>%
      mutate(mean_rank = seq_along(mean_density))

    annual_data <- rbind(annual_data, tmp)
  }
}

dom_annuals <- annual_data %>%
  filter(species %in% c("Allyssum desertorum","Bromus tectorum")) %>%
  select(-mean_rank) %>%
  left_join(weather_data, by = "year")
```

\newpage{}
# Table Ax-1: Statistical results from linear model relating density of Allysum desertorum to precipitation.
```{r report_models, echo = FALSE}
summary(lm(mean_density~ppt1*Treatment, data = filter(dom_annuals, species == "Allyssum desertorum")))
```

\newpage{}
# Table Ax-2: Statistical results from linear model relating density of Bromus tectorum to precipitation.
```{r report_models2, echo = FALSE}
summary(lm(mean_density~ppt1*Treatment, data = filter(dom_annuals, species == "Bromus tectorum")))
```

\newpage{}

```{r rac_plots, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Rank abundance curves for perennial species. Basal area of individuals was summed withnin years and plots, and then the total area values were averaged across years and plots for each treatment."}
####
####  EXTRACT COVER OF MOST DOMINANT SPECIES ----
####
cover_data <- {}
for(dotrt in unique(cover_community_matrix$Treatment)){
  tmp <- cover_community_matrix %>%
    gather(species, area, -quad, -year, -Treatment) %>%
    filter(Treatment == dotrt) %>%
    group_by(Treatment, species) %>%
    summarise(mean_area = mean(area)) %>%
    arrange(-mean_area) %>%
    mutate(mean_rank = seq_along(mean_area))

  cover_data <- rbind(cover_data, tmp)
}

ggplot(cover_data, aes(x = mean_rank, y = mean_area))+
  geom_line()+
  geom_point(size = 2)+
  geom_text(data=filter(cover_data, mean_rank < 11), aes(x = mean_rank, y = mean_area, label = species), 
            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
  scale_y_continuous(limits = c(0,0.25))+
  xlab("Species rank")+
  ylab(expression(paste("Mean basal area (",m^2,")")))+
  facet_wrap(~Treatment, ncol = 3)

cover_doms <- cover_data %>%
  filter(mean_rank < 3) %>%
  group_by(Treatment) %>%
  summarise(total_area = sum(mean_area))

```

```{r annual_racs, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Rank abundance curves for annual species. Density of individuals is averaged across years and plots for each treatment."}
annual_data <- {}
for(dotrt in unique(annual_community_matrix$Treatment)){
  tmp <- annual_community_matrix %>%
    gather(species, density, -quad, -year, -Treatment) %>%
    filter(Treatment == dotrt) %>%
    group_by(Treatment, species) %>%
    summarise(mean_density = mean(density)) %>%
    arrange(-mean_density) %>%
    mutate(mean_rank = seq_along(mean_density))

  annual_data <- rbind(annual_data, tmp)
}

ggplot(annual_data, aes(x = mean_rank, y = mean_density))+
  geom_line()+
  geom_point(size = 2)+
  geom_text(data=filter(annual_data, mean_rank < 3), aes(x = mean_rank, y = mean_density, label = species), 
            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
  scale_y_continuous(limits = c(0,200))+
  xlab("Species rank")+
  ylab("Mean density")+
  facet_wrap(~Treatment, ncol = 3)
```

```{r time_perr_racs, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=6, fig.cap="Time series of rank abundance curves for perennial species in each treatment. Values of mean area were averaged over plots. The four most dominant species are labelled in each panel."}
cover_data <- {}
for(dotrt in unique(cover_community_matrix$Treatment)){
  for(doyr in unique(cover_community_matrix$year)){
    tmp <- cover_community_matrix %>%
      gather(species, area, -quad, -year, -Treatment) %>%
      filter(Treatment == dotrt, year == doyr) %>%
      group_by(Treatment, year, species) %>%
      summarise(mean_area = mean(area)) %>%
      arrange(-mean_area) %>%
      mutate(mean_rank = seq_along(mean_area))

  cover_data <- rbind(cover_data, tmp)
  }
}
ggplot(cover_data, aes(x = mean_rank, y = mean_area))+
  geom_line()+
  geom_point(size = 2)+
  geom_text(data=filter(cover_data, mean_rank < 5), aes(x = mean_rank, y = mean_area, label = species), 
            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
  scale_y_continuous(limits = c(0,0.25))+
  xlab("Species rank")+
  ylab(expression(paste("Mean basal area (",m^2,")")))+
  facet_grid(Treatment~year)
```

```{r time_annual_racs, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=6, fig.cap="Time series of rank abundance curves for annual species in each treatment. Values of density were averaged over plots. The two most dominant species are labelled in each panel."}
annual_data <- {}
for(dotrt in unique(annual_community_matrix$Treatment)){
  for(doyr in unique(annual_community_matrix$year)){
    tmp <- annual_community_matrix %>%
      gather(species, density, -quad, -year, -Treatment) %>%
      filter(Treatment == dotrt, year == doyr) %>%
      group_by(Treatment, year, species) %>%
      summarise(mean_density = mean(density)) %>%
      arrange(-mean_density) %>%
      mutate(mean_rank = seq_along(mean_density))

    annual_data <- rbind(annual_data, tmp)
  }
}

ggplot(annual_data, aes(x = mean_rank, y = mean_density))+
  geom_line()+
  geom_point(size = 2)+
  geom_text(data=filter(annual_data, mean_rank < 3), aes(x = mean_rank, y = mean_density, label = species), 
            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
  scale_y_continuous(limits = c(0,400))+
  scale_x_continuous(breaks = c(1,6,11))+
  xlab("Species rank")+
  ylab("Mean density")+
  facet_grid(Treatment~year)
```

\newpage{}

```{r annual_weather, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Relationship between precipitation and abundance of dominant annuals."}
dom_annuals <- annual_data %>%
  filter(species %in% c("Allyssum desertorum","Bromus tectorum")) %>%
  select(-mean_rank) %>%
  rename(Species = species) %>%
  left_join(weather_data, by = "year")

ggplot(dom_annuals, aes(x = ppt1, y = mean_density, color = Species))+
  geom_point()+
  xlab("Precipitation (mm)")+
  ylab("Abundance")
```

### 2017-12-06 code/05_community_dynamics.R

@@ -167,98 +167,3 @@ ggplot(nmds_df, aes(x=MDS1, y=MDS2, fill=Treatment))+
   theme(panel.grid.major = element_line(color="grey90"))
 ggsave(paste0(figure_path,"sppcomp_bray_all.png"), width=6, height = 4, units = "in", dpi = 120)
 
-
-
-####
-####  LOOK AT PERENNIALS AND ANNUALS SEPARATELY ----
-####
-##  Perennials
-# Set missing species abundances as 0, which == absent
-density_community_matrix[is.na(density_community_matrix)] <- 0 
-
-density_community_matrix_scaled <- cbind(density_community_matrix[,1:3,drop=F], 
-                                      scale(density_community_matrix[,4:ncol(density_community_matrix)]))
-
-spp_to_remove <- names(which(is.nan(colMeans(density_community_matrix_scaled[4:ncol(density_community_matrix_scaled)]))))
-ids_to_remove <- which(colnames(density_community_matrix_scaled) %in% spp_to_remove)
-density_community_matrix_scaled <- density_community_matrix_scaled %>%
-  select(-ids_to_remove)
-
-
-nmds_df_perennials <- {}
-out_stats <- {}
-for(doyr in unique(density_community_matrix_scaled$year)){
-  tmp <- filter(density_community_matrix_scaled, year == doyr) %>% select(-year)
-  invisible( capture.output( # already checked for convergence
-    tmp_bray <- metaMDS(tmp[3:ncol(tmp)]+2, "bray", trymax = 999)
-  ))
-  nmds_points <- as.data.frame(tmp_bray$points)
-  nmds_points$Treatment <- tmp$Treatment
-  nmds_points$Year <- doyr
-  nmds_df_perennials <- rbind(nmds_df_perennials, nmds_points)
-  adonis_stats <- as.data.frame(vegan::adonis(tmp[3:ncol(tmp)]+2~tmp$Treatment)$aov.tab)[1,]
-  betadisper_stats <- anova(vegan::betadisper(vegan::vegdist(tmp[3:ncol(tmp)]+2),tmp$Treatment))[1,]
-  tmp_out <- rbind(adonis_stats[,-which(colnames(adonis_stats)=="R2")],
-                   as.numeric(betadisper_stats))
-  row.names(tmp_out) <- c()
-  out_stats <- rbind(out_stats, data.frame(year = doyr,
-                                           test = c("adonis","betadisper"),
[…]

### 2017-12-06 code/cache/community_dynamics_extra.R

@@ -1,0 +1,121 @@
+####
+####  EXTRACT COVER OF MOST DOMINANT SPECIES ----
+####
+cover_data <- {}
+for(dotrt in unique(cover_community_matrix$Treatment)){
+  tmp <- cover_community_matrix %>%
+    gather(species, area, -quad, -year, -Treatment) %>%
+    filter(Treatment == dotrt) %>%
+    group_by(Treatment, species) %>%
+    summarise(mean_area = mean(area)) %>%
+    arrange(-mean_area) %>%
+    mutate(mean_rank = seq_along(mean_area))
+  
+  cover_data <- rbind(cover_data, tmp)
+}
+
+
+ggplot(cover_data, aes(x = mean_rank, y = mean_area))+
+  geom_line()+
+  geom_point(size = 2)+
+  geom_text(data=filter(cover_data, mean_rank < 11), aes(x = mean_rank, y = mean_area, label = species), 
+            hjust = -0.05, nudge_x = 0.2, angle = 45)+
+  scale_y_continuous(limits = c(0,0.25))+
+  xlab("Mean rank")+
+  ylab("Mean area (cm*cm)")+
+  facet_wrap(~Treatment, ncol = 3)
+
+
+####
+####  LOOK AT PERENNIALS AND ANNUALS SEPARATELY ----
+####
+##  Perennials
+# Set missing species abundances as 0, which == absent
+# density_community_matrix[is.na(density_community_matrix)] <- 0 
+# 
+# density_community_matrix_scaled <- cbind(density_community_matrix[,1:3,drop=F], 
+#                                       scale(density_community_matrix[,4:ncol(density_community_matrix)]))
+# 
+# spp_to_remove <- names(which(is.nan(colMeans(density_community_matrix_scaled[4:ncol(density_community_matrix_scaled)]))))
+# ids_to_remove <- which(colnames(density_community_matrix_scaled) %in% spp_to_remove)
+# density_community_matrix_scaled <- density_community_matrix_scaled %>%
+#   select(-ids_to_remove)
+# 
+# 
+# nmds_df_perennials <- {}
+# out_stats <- {}
+# for(doyr in unique(density_community_matrix_scaled$year)){
+#   tmp <- filter(density_community_matrix_scaled, year == doyr) %>% select(-year)
+#   invisible( capture.output( # already checked for convergence
[…]

### 2017-12-06 manuscript/review_response_R1.Rmd
minor work on the response and revision

@@
 But, in the end, we decided this would muddle the analysis and confuse the reader, which is why we kept it separate.
-In our revision, we have included a new appendix where we conduct our regression with NDVI rather than ANPP.
+In our revision, we have included a new appendix (Appendix X) where we conduct our regression with NDVI rather than ANPP.
 We achieved the same results as when analyzing ANPP, and keep ANPP in the main text since that variable connects more easily to previous research (and what readers might be expecting).}
@@
 2017).
-That is an explicit test of HRF, where the sensitivites could show any number of temporl patterns outlined by Smith.
+That is an explicit test of HRF, where the sensitivites could show any number of temporal patterns outlined by Smith.
 
 \response{The reviewer is correct that our analysis is not an explicit test of the Hierarchical Response Framework because we did not explicitly model the effect of time on ecosystem functional response.
-Unfortunately, we are unable to do so given our data -- in each year every plot shares the same average soil moisture for its particular treatment.
-Therefore, we cannot fit year-specific slopes.
+Unfortunately, we are unable to do so given our data -- in each year every plot shares the same average soil moisture for its particular treatment, making it difficult to fit year-specific slopes.
 The sensitivity analysis would also be interesting, but moves us farther away from our interest in the ecosystem functional response.
-To remedy the situation, we have revised our motivation in the introduction to focus on the mechanisms that could alter ecosystem functional response, physiological responses and changes in community composition, without attributing specific time scales to them.}
+To remedy the situation, we have revised our motivation in the introduction to focus on the mechanisms that could alter ecosystem functional response, physiological responses and changes in community composition, without attributing specific time scales to them.
+Thus, we no longer pitch our paper as a test of the Hierarchical Response Framework, but we do rely on it to motivate our study.
+We also note that a community change may have to happen relatively early in the time series to impact ecosystem functional response.
+Fortunately, our community level analysis allows to investigate this (and we found that the communities do not diverge over time, or change at all).}
 
@@
 
-\response{Good idea.
-We reran the community analysis with different subsets of species grouped by functional types/rarity.
-Our results are the same: community composition is really, really stable.
-We have added a mention of this in the methods and results sections, but keep our original figure in the main text.
-If the reviewer thinks it is necessary, we are happy to make a new appendix showing this analysis.}
+\response{We have added a new appendix (Appendix X) that provides more details on the dominance and rarity of the communities across and through time.
+In brief, rank abundance curves are very similar for all treatments and through time, with the same two perennial species (\emph{Artemisia tripartita} and \emph{Balsamorhiza sagittata}) dominant in all treatments, acounting for about 25\% of total basal cover.
+Similarly, all treatments had the same dominant annual species, whose dynamics were more associated with interannual variation in precipitation than to variation among treatments.
+These ancillary results add weight to our conclusion that the ecosystem responses is driven by dominant species, which are similar among treatments.}
 

### 2017-12-06 manuscript/usses_water_ms.tex
minor work on the response and revision

@@
 data.
+\new{Two anonymous reviewers and Elsie Denton provided thoughtful reviews that improved our paper.}
 
@@
   \hline
-2011 & adonis &  21 &   2 & 1.02 & 0.45 \\ 
+2011 & adonis &  21 &   2 & 1.02 & 0.43 \\ 
   2011 & betadisper &  21 &   2 & 2.22 & 0.14 \\ 
-  2012 & adonis &  22 &   2 & 1.09 & 0.31 \\ 
+  2012 & adonis &  22 &   2 & 1.09 & 0.33 \\ 
   2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
-  2013 & adonis &  22 &   2 & 1.25 & 0.11 \\ 
+  2013 & adonis &  22 &   2 & 1.25 & 0.13 \\ 
   2013 & betadisper &  22 &   2 & 0.44 & 0.65 \\ 
-  2014 & adonis &  22 &   2 & 0.95 & 0.55 \\ 
+  2014 & adonis &  22 &   2 & 0.95 & 0.56 \\ 
   2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
-  2015 & adonis &  21 &   2 & 1.05 & 0.38 \\ 
+  2015 & adonis &  21 &   2 & 1.05 & 0.41 \\ 
   2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 
-  2016 & adonis &  21 &   2 & 1.07 & 0.33 \\ 
+  2016 & adonis &  21 &   2 & 1.07 & 0.36 \\ 
   2016 & betadisper &  21 &   2 & 0.49 & 0.62 \\ 

### 2017-12-06 code/03_anpp_randcoefs_model.R
trivia

@@ -87,6 +87,7 @@ saveRDS(fit, "../results/randcoefs_alltreatments_fit.RDS")
 ####
 ####  OPTIONAL DIAGNOSTICS ----
 ####
+# fit <- readRDS("../results/randcoefs_alltreatments_fit.RDS")
 # betas <- as.numeric( summary(fit, 'beta')$summary[,1] )
 # 
 # df <- data.frame( x = anppdat$x, predicted = anppdat$x %*% betas)

### 2017-12-07 manuscript/review_response_R1.Rmd
work on new appendix and some small revisions

@@
 \response{This is a great suggestion, and one we had not thought of.
-We conducted the analysis suggested by the reviewer and found no difference among perennials and annuals (Fig.
-R1).
-We have added this result in our Results section on lines xxx-xxx, but do not inlcude the figure.}
+We conducted the analysis suggested by the reviewer and found that community composition was stable for both annuals (short-lived) and perennials (long-lived).
+In a few years we detected signficant treatment effects on dispersion or mean of community composition.
+But upon further investigation, we found that these statistical results were detecting changes in rare species because the dominant annuals showed similar dynamics through time, regardless of treatment (Fig.
+Ax-x).
+Likewise, the stable perennial community contributes most to ANPP is this systems.
+We briefly mention this analysis in our Results section on lines xxx-xxx, and show the full analysis in Appendix X.}
 

### 2017-12-07 manuscript/Appendix5.Rmd
work on new appendix and some small revisions

@@
   - \usepackage[sc]{mathpazo}
+  - \usepackage{float}
 bibliography: ~/Dropbox/Bibliography/usses_water.bib
@@
 
+We also conducted our community composition analysis with only annual species.
+Annual species are shorter-lived than the perennial species in our community, so they may respond more quickly to alterations of precipitation.
+In general, our results for annual species conform to the results from the full community analysis in the main text.
+Annual plant community composition is relatively stable through time (Fig. Ax-5) and in most years there is no evidence that treatment differentiates community composition (Table Ax-).
+
+\newpage{}
+# Tables
+
 ```{r get_data, echo=FALSE, message=FALSE, warning=FALSE}
@@
 library(ggthemes)  # pleasing ggplot2 themes
+library(vegan)     # NMDS and community analysis
+library(xtable)    # Generating LaTeX tables in RMarkdown
 
@@
 
-\newpage{}
-# Table Ax-1: Statistical results from linear model relating density of Allysum desertorum to precipitation.
+Table Ax-1: Statistical results from linear model relating density of Allysum desertorum to precipitation.
 ```{r report_models, echo = FALSE}
@@
 \newpage{}
-# Table Ax-2: Statistical results from linear model relating density of Bromus tectorum to precipitation.
+Table Ax-2: Statistical results from linear model relating density of Bromus tectorum to precipitation.
 ```{r report_models2, echo = FALSE}
@@
 
+```{r annual_nmds, echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}
+####
+####  ANALYZE JUST THE ANNUALS ----
+####
+# Set missing species abundances as 0, which == absent
+annual_community_matrix[is.na(annual_community_matrix)] <- 0
+
+annual_community_matrix_scaled <- cbind(annual_community_matrix[,1:3,drop=F],
+                                        scale(annual_community_matrix[,4:ncol(annual_community_matrix)]))
+
+spp_to_remove <- names(which(is.nan(colMeans(annual_community_matrix_scaled[4:ncol(annual_community_matrix_scaled)]))))
+ids_to_remove <- which(colnames(annual_community_matrix_scaled) %in% spp_to_remove)
+annual_community_matrix_scaled <- annual_community_matrix_scaled %>%
+  select(-ids_to_remove)
+
+nmds_df_annuals <- {}
+out_stats <- {}
+for(doyr in unique(annual_community_matrix_scaled$year)){
+  tmp <- filter(annual_community_matrix_scaled, year == doyr) %>% select(-year)
+   invisible( capture.output( # already checked for convergence
+    tmp_bray <- metaMDS(tmp[3:ncol(tmp)]+2, "bray", trymax = 999)
+  ))
+  nmds_points <- as.data.frame(tmp_bray$points)
+  nmds_points$Treatment <- tmp$Treatment
+  nmds_points$Year <- doyr
+  nmds_df_annuals <- rbind(nmds_df_annuals, nmds_points)
+  adonis_stats <- as.data.frame(vegan::adonis(tmp[3:ncol(tmp)]+2~tmp$Treatment)$aov.tab)[1,]
+  betadisper_stats <- anova(vegan::betadisper(vegan::vegdist(tmp[3:ncol(tmp)]+2),tmp$Treatment))[1,]
+  tmp_out <- rbind(adonis_stats[,-which(colnames(adonis_stats)=="R2")],
+                   as.numeric(betadisper_stats))
+  row.names(tmp_out) <- c()
+  out_stats <- rbind(out_stats, data.frame(year = doyr,
+                                           test = c("adonis","betadisper"),
+                                           n = nrow(tmp),
+                                           tmp_out))
+}
+```
+
+\setcounter{table}{2}
+
+```{r annual_nmds_table, echo = FALSE, include=TRUE, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
+colnames(out_stats) <- c("Year", "Test", "n", "d.f.","Sums","Means","\\emph{F}", "\\emph{P}")
+out_stats$d.f. <- as.integer(out_stats$d.f.)
+out_stats <- arrange(out_stats, Year)
+
+# Make the table
+inyear_caption <- "Results from statistical tests for clustering and dispersion of community composition among precipitation treatments for annual species only.
+`adonis' tests whether treatments form unique clusters in multidimensial space; `betadisper' tests whether treatments have similar dispersion.
+For both tests, \\emph{P} values greater than 0.05 indicate there is no support that the treatments differ."
+print(xtable(dplyr::select(out_stats, -Sums, -Means), caption = inyear_caption, digits = 2),
+       caption.placement="top",
+       include.rownames = F,
+       sanitize.colnames.function = identity,
+       comment = FALSE,
+       size = "normalsize")
+```
+
+\newpage{}
+
+# Figures
+
 ```{r rac_plots, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Rank abundance curves for perennial species.
-Basal area of individuals was summed withnin years and plots, and then the total area values were averaged across years and plots for each treatment."}
+Basal area of individuals was summed withnin years and plots, and then the total area values were averaged across years and plots for each treatment.", fig.pos="h",fig.align="center"}
 ####
@@
 ```
+
+```{r annual_nmds2, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Nonmetric multidimensional scaling scores representing annual plant communities in each plot, colored by treatment."}
+####
+####  ANALYZE JUST THE ANNUALS ----
+####
+# Set missing species abundances as 0, which == absent
+annual_community_matrix[is.na(annual_community_matrix)] <- 0
+
+annual_community_matrix_scaled <- cbind(annual_community_matrix[,1:3,drop=F],
+                                        scale(annual_community_matrix[,4:ncol(annual_community_matrix)]))
+
+spp_to_remove <- names(which(is.nan(colMeans(annual_community_matrix_scaled[4:ncol(annual_community_matrix_scaled)]))))
+ids_to_remove <- which(colnames(annual_community_matrix_scaled) %in% spp_to_remove)
+annual_community_matrix_scaled <- annual_community_matrix_scaled %>%
+  select(-ids_to_remove)
+
+nmds_df_annuals <- {}
+out_stats <- {}
+for(doyr in unique(annual_community_matrix_scaled$year)){
+  tmp <- filter(annual_community_matrix_scaled, year == doyr) %>% select(-year)
+   invisible( capture.output( # already checked for convergence
+    tmp_bray <- metaMDS(tmp[3:ncol(tmp)]+2, "bray", trymax = 999)
+  ))
+  nmds_points <- as.data.frame(tmp_bray$points)
+  nmds_points$Treatment <- tmp$Treatment
+  nmds_points$Year <- doyr
+  nmds_df_annuals <- rbind(nmds_df_annuals, nmds_points)
+  adonis_stats <- as.data.frame(vegan::adonis(tmp[3:ncol(tmp)]+2~tmp$Treatment)$aov.tab)[1,]
+  betadisper_stats <- anova(vegan::betadisper(vegan::vegdist(tmp[3:ncol(tmp)]+2),tmp$Treatment))[1,]
+  tmp_out <- rbind(adonis_stats[,-which(colnames(adonis_stats)=="R2")],
+                   as.numeric(betadisper_stats))
+  row.names(tmp_out) <- c()
+  out_stats <- rbind(out_stats, data.frame(year = doyr,
+                                           test = c("adonis","betadisper"),
+                                           n = nrow(tmp),
+                                           tmp_out))
+}
+
+mycols <- c("#009E73", "#0072B2", "#D55E00")
+ggplot(nmds_df_annuals, aes(x=MDS1, y=MDS2, color=Treatment))+
+  geom_point()+
+  geom_point(color="grey35",shape=1)+
+  scale_color_manual(values=mycols)+
+  ylab("NMDS 2")+
+  xlab("NMDS 1")+
+  facet_wrap(~Year)+
+  #scale_y_continuous(limits=c(-0.15,0.15))+
+  #scale_x_continuous(limits=c(-0.18,0.18))+
+  coord_fixed()
+```

### 2017-12-07 manuscript/usses_water_ms.Rmd
minor work on the response and revision
work on new appendix and some small revisions

@@
 The random year effects ($\boldsymbol{\eta}$) are drawn from a normal prior with mean 0 and standard deviation $\sigma_{\text{year}}$, which was drawn from a half-Cauchy distribution. 
+\new{We used a vague prior distribution for each $\beta$: $\boldsymbol{\beta} \sim \text{Normal}(0,5)$.}
 A full description of our model and associated R [@R2016] code is in Appendix 2.
@@
 4) among treatments is surprising because grasslands generally, and sagebrush steppe specifically, are considered water-limited systems.
-Indeed, we expected ecosystem functional response, community composition, or both to change under our treatments.
+\new{For example,} @Huxman2004 and @Knapp2015 \new{showed that semi-arid sites are more sensitive to drought than mesic sites, and} @Wilcox2017 \new{found that semi-arid sites are particularly sensitive to irrigation treatments.}
+\new{Based on these findings}, we expected ecosystem functional response, community composition, or both to change under our treatments.
 Why did our treatments fail to induce ecosystem or community responses?
@@
 Furthermore, the dominant shrub in our focal ecosystem, \emph{Artemisia tripartita}, has access to water deep in the soil profile thanks to a deep root system [@Kulmatiski2017a].
+\new{The dominance of our site by the shrub \emph{A.
+tripartita} may explain why our results do not conform to broader patterns of grassland sensitivity to precipitation manipulations} [e.g., @Huxman2004; @Knapp2015; @Wilcox2017].
 
@@
 We also thank Susan Durham for clarifying our thinking on the statistical analyses and Kevin Wilcox for helpful discussions on analyzing community composition data.
+\new{Two anonymous reviewers and Elsie Denton provided thoughtful reviews that improved our paper.}
 

### 2017-12-07 manuscript/usses_water_ms.tex
work on new appendix and some small revisions

@@
 prior with mean 0 and standard deviation \(\sigma_{\text{year}}\), which
-was drawn from a half-Cauchy distribution. A full description of our
-model and associated R \citep{R2016} code is in Appendix 2.
+was drawn from a half-Cauchy distribution.
+\new{We used a vague prior distribution for each $\beta$: $\boldsymbol{\beta} \sim \text{Normal}(0,5)$.}
+A full description of our model and associated R \citep{R2016} code is
+in Appendix 2.
 
@@
 generally, and sagebrush steppe specifically, are considered
-water-limited systems. Indeed, we expected ecosystem functional
+water-limited systems. \new{For example,} \citet{Huxman2004} and
+\citet{Knapp2015}
+\new{showed that semi-arid sites are more sensitive to drought than mesic sites, and}
+\citet{Wilcox2017}
+\new{found that semi-arid sites are particularly sensitive to irrigation treatments.}
+\new{Based on these findings}, we expected ecosystem functional
 response, community composition, or both to change under our treatments.
@@
 \citep{Kulmatiski2017a}.
+\new{The dominance of our site by the shrub \emph{A. tripartita} may explain why our results do not conform to broader patterns of grassland sensitivity to precipitation manipulations}
+\citep[e.g.,][]{Huxman2004, Knapp2015, Wilcox2017}.
 

### 2017-12-07 code/05_community_dynamics.R

@@ -167,3 +167,4 @@ ggplot(nmds_df, aes(x=MDS1, y=MDS2, fill=Treatment))+
   theme(panel.grid.major = element_line(color="grey90"))
 ggsave(paste0(figure_path,"sppcomp_bray_all.png"), width=6, height = 4, units = "in", dpi = 120)
 
+

### 2017-12-09 manuscript/review_response_R1.Rmd
a few more revisions

@@
 
-\response{We provide point-by-point responses to reviewer comments below. We have taken care to address Reviewer 1's concerns (also echoed by Reviewer 2) by }
+\response{We provide point-by-point responses to reviewer comments below.
+We have taken care to address Reviewer 1's concerns (also echoed by Reviewer 2) by substantially revising our Introduction section.
+In particular, we have shifted our focus away from the Hierarchical Response Framework, and instead focus on how species' responses and community compositional shifts can impact ecosystem functional response.
+We do this, rather than implementing the suggestions of Reviewer 1, because our data limits the complexity of our statistical model.
+Within a year and treatment, each plot has the same value of volumetric water content.
+Therefore, we simply don't have the variation needed to fit a VWC$\times$Treatment$\times$Year interaction, which would be necessary to incorporate the effect of time on the ecosystem functional response.}
+
+\response{We therefore use the Hierarchical Response Framework to support our hypotheses about potential mechanisms driving change in ecosystem functional response (species' responses and community change), but refrain from suggesting we are explicitly testing the Hierarchical Response Framework.
+We also circle back to this point in our Discussion (lines xxx-xxx).
+Fortunately, our community-level analysis does have an explicit temporal component, which does link back to the Hierarchical Response Framework.}
 
@@
 
-\response{Sorry this was not clear. We included the prior specification [$\beta \sim \text{Normal}(0,5)$] in Appendix 2 on line 59. We have now included this in the main text on line xx.}
+\response{Sorry this was not clear. We included the prior specification [$\beta \sim \text{Normal}(0,5)$] in Appendix X on line 59. We have now included this in the main text on line xx.}
 
@@
 \response{The reviewer is correct that our analysis is not an explicit test of the Hierarchical Response Framework because we did not explicitly model the effect of time on ecosystem functional response.
-Unfortunately, we are unable to do so given our data -- in each year every plot shares the same average soil moisture for its particular treatment, making it difficult to fit year-specific slopes.
+Unfortunately, we are unable to do so given our data -- in each year every plot shares the same average soil moisture for its particular treatment, making it difficult to fit year-specific slopes for each treatment level.
 The sensitivity analysis would also be interesting, but moves us farther away from our interest in the ecosystem functional response.
-To remedy the situation, we have revised our motivation in the introduction to focus on the mechanisms that could alter ecosystem functional response, physiological responses and changes in community composition, without attributing specific time scales to them.
+To remedy the situation, we have revised our motivation in the introduction to focus on the mechanisms that could alter ecosystem functional response, physiological responses and changes in community composition, without attributing specific time scales to them (lines xxx-xxx).
 Thus, we no longer pitch our paper as a test of the Hierarchical Response Framework, but we do rely on it to motivate our study.
-We also note that a community change may have to happen relatively early in the time series to impact ecosystem functional response.
+We also note that a community change may have to happen relatively early in the time series to impact ecosystem functional response (line xxx).
 Fortunately, our community level analysis allows to investigate this (and we found that the communities do not diverge over time, or change at all).}
@@
 As we mention above, we have conducted a model comparison where we show that results are consistent whether we use ANPP or NDVI (see new Appendix X).
-Thus, in the main text we retain the analysis focused on ANPP because we think it connects better to previous research.
+In the main text we retain the analysis focused on ANPP because we think it connects better to previous research.
 Also, as mentioned above, there was an error in our regression for the year 2015, which has been fixed (see new Table A1-2).}
@@
 \response{Yes, we agree that the rate of community turnover is likely related to the lifespans of constituent species.
-The physiological responses, however, should not depend on lifespan.
-Indeed, this it at the core of the Hierarchical Response Framework.
 We have added text discussing the impact of lifespan on community turnover in our Conclusion (lines xxx-xxx).}
@@
 
-\response{Good catch; fixed.}
+\response{Good catch. Fixed.}
 
@@
 
-\response{Good idea; done.}
+\response{Good idea. Done.}
 
@@
 Reviewer 2 agreed that the paper would benefit from a broader introduction, which we have added (lines xx-xx).
-We have also added a brief reference to the idea that domoninant species play a large role in determining ecosystem functioning (lines xx-xx).}
+We have also added a reference to the idea that domoninant species play a large role in determining ecosystem functioning (lines xx-xx).}
 
@@
 
-\response{We have revised this sentence to read, "However, all estimates for the effect of soil moisture on ANPP overlapped zero, indicating the relationship is weak and uncertain regardless of treatment." (lines xx-xx).}
+\response{We have revised this sentence to read, "However, all estimates for the effect of soil moisture on ANPP broadly overlapped zero, indicating the relationship is weak and uncertain regardless of treatment." (lines xx-xx).}
 
@@
 
-\response{Good catch; extra "outside" is removed.}
+\response{Good catch. Extra "outside" has been removed.}
 
@@
 We also use the demographic data to determine species composition.
-Given that this demographic data is time-consuming to collect, we decided to use the historical plots as controls rather than adding even more plots."}
+Given that this demographic data is time-consuming to collect, we decided to use the historical plots as controls rather than adding even more plots." We have added this justification in our Methods section on lines xxx-xxx.}
 
@@
 
-\response{We have revised this section to make this more clear (lines xxx-xxx).}
+\response{We have revised this section to make this clearer (lines xxx-xxx). To answer the reviewer's question, }
 

### 2017-12-09 manuscript/usses_water_ms.Rmd
a few more revisions

@@
   - name: Andrew T. Tredennick
-    affiliation: 1
+    affiliation: 1,*
     corresponding: true
@@
   - name: Andrew R. Kleinhesselink
-    affiliation: 1,2
+    affiliation: 1,2,*
   - name: J. Bret Taylor
@@
   \usepackage{bm,mathrsfs}
+  \usepackage{lipsum}
 bibliography: /Users/atredenn/Dropbox/Bibliography/usses_water.bib
@@
 \newcommand{\new}{\textcolor{blue}}
+
+\newcommand\blfootnote[1]{%
+  \begingroup
+  \renewcommand\thefootnote{}\footnote{#1}%
+  \addtocounter{footnote}{-1}%
+  \endgroup
+}
+\blfootnote{*Authors contributed equally to this work.}
 
@@
 ANPP varied from a minimum of 74.5 g m$^{-2}$ in 2014 to a maximum of 237.1 g m$^{-2}$ in 2016 when averaged across treatments (Fig. 2C).
-ANPP was slightly higher in irrigation plots and slightly lower in drought plots (Fig. 2C), corresponding to estimated soil volumetric water content (VWC) differences among treatments (Fig. 2B).
+ANPP was slightly higher in irrigation plots \new{(on average 17\% higher)} and slightly lower in drought plots \new{(on average 23\% lower) relative to control plots} (Fig.
+2C), corresponding to estimated soil volumetric water content (VWC) differences among treatments (Fig.
+2B).
+\new{VWC in drought plots was 12\% less than in control plots on average, and VWC in irrigation plots was 19\% higher than in control plots on average.}
 Such differences in soil VWC indicate our treatment infrastructure was successful.
-ANPP was highly variable across plots within years (Fig. 2C).
+ANPP was highly variable across plots within years, \new{as indicated by the large and overlapping standard deviations} (Fig. 2C).
 

### 2017-12-09 manuscript/usses_water_ms.tex
a few more revisions

@@
 \usepackage{color, soul} \usepackage{sectsty} \usepackage{bm,mathrsfs}
+\usepackage{lipsum}
 
@@
 
-\author[1]{Andrew T. Tredennick}
+\author[1,*]{Andrew T. Tredennick}
 
-\corrauthor[1]{Andrew T. Tredennick}{\href{mailto:atredenn@gmail.com}{\nolinkurl{atredenn@gmail.com}}}
-\author[1,2]{Andrew R. Kleinhesselink}
+\corrauthor[1,*]{Andrew T. Tredennick}{\href{mailto:atredenn@gmail.com}{\nolinkurl{atredenn@gmail.com}}}
+\author[1,2,*]{Andrew R. Kleinhesselink}
 
@@
 \definecolor{blue}{rgb}{0,0,0.7} \newcommand{\new}{\textcolor{blue}}
+
+\newcommand\blfootnote[1]{%
+  \begingroup
+  \renewcommand\thefootnote{}\footnote{#1}%
+  \addtocounter{footnote}{-1}%
+  \endgroup
+}
+
+\blfootnote{*Authors contributed equally to this work.}
 
@@
 maximum of 237.1 g m\(^{-2}\) in 2016 when averaged across treatments
-(Fig. 2C). ANPP was slightly higher in irrigation plots and slightly
-lower in drought plots (Fig. 2C), corresponding to estimated soil
-volumetric water content (VWC) differences among treatments (Fig. 2B).
+(Fig. 2C). ANPP was slightly higher in irrigation plots
+\new{(on average 17\% higher)} and slightly lower in drought plots
+\new{(on average 23\% lower) relative to control plots} (Fig. 2C),
+corresponding to estimated soil volumetric water content (VWC)
+differences among treatments (Fig. 2B).
+\new{VWC in drought plots was 12\% less than in control plots on average, and VWC in irrigation plots was 19\% higher than in control plots on average.}
 Such differences in soil VWC indicate our treatment infrastructure was
-successful. ANPP was highly variable across plots within years (Fig.
-2C).
+successful. ANPP was highly variable across plots within years,
+\new{as indicated by the large and overlapping standard deviations}
+(Fig. 2C).
 
@@
   \hline
-2011 & adonis &  21 &   2 & 1.02 & 0.43 \\ 
+2011 & adonis &  21 &   2 & 1.02 & 0.45 \\ 
   2011 & betadisper &  21 &   2 & 2.22 & 0.14 \\ 
-  2012 & adonis &  22 &   2 & 1.09 & 0.33 \\ 
+  2012 & adonis &  22 &   2 & 1.09 & 0.34 \\ 
   2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
@@
   2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
-  2015 & adonis &  21 &   2 & 1.05 & 0.41 \\ 
+  2015 & adonis &  21 &   2 & 1.05 & 0.39 \\ 
   2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 

### 2017-12-09 code/02_plot_data.R

@@ -1,9 +1,13 @@
-##  look_at_weather.R: Script to look at the historical distribution of
-##  precipitation at Dubois, Idaho, and to compare our treatment years to
-##  historical events.
+################################################################################
+##  02_plot_data.R: Script to plot the historical distribution of precipitation 
+##  at Dubois, Idaho, and to compare our treatment years to historical events. 
+##  The script also plots modeled VWC for each treatment through time, and the
+##  temporal trend of ANPP for each treatment.
 ##
+##  ----------------------------------------------------------------------------
 ##  Author: Andrew Tredennick (atredenn@gmail.com)
 ##  Date created: May 10, 2017
+################################################################################
 
 ##  Clear the workspace
 rm(list = ls(all.names = T))
@@ -12,27 +16,28 @@ rm(list = ls(all.names = T))
 setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
 
 
+
 ####
-####  LIBRARIES ----
+####  LOAD LIBRARIES -----------------------------------------------------------
 ####
-library(tidyverse)
-library(dplyr)
-library(stringr)
-library(ggthemes)
-library(ggalt)
-library(gridExtra)
-library(cowplot)
+library(tidyverse) # Data science packages
+library(dplyr)     # Data summarizing and manipulating
+library(stringr)   # Working with strings
+library(ggthemes)  # Pleasing themes for ggplot
+library(cowplot)   # For combining ggplot objects
 
 
 
 ####
-####  READ IN WEATHER DATA AND PLOT ----
+####  READ IN WEATHER DATA AND PLOT --------------------------------------------
 ####
 weather <- read.csv("../data/weather/ClimateIPM.csv")
 trt_data <- weather %>%
   filter(year>2011) %>%
-  select(year, ppt1)
-trt_data$Treatment <- "Control"
+  select(year, ppt1) %>%
+  mutate(Treatment = "Control")
+
+##  Adjust 2012 and 2015 values so they are distinguishable on the plot
[…]

### 2017-12-09 code/cache/alt_anpp_randcoefs_model.R

@@ -1,0 +1,117 @@
+##  anpp_randcoefs_model.R: Script to run GLMM analysis to test for treatment 
+##  effects on the relationship between precipitation and ANPP.
+##
+##  NOTE: Stan may issue a couple warnings after running the MCMC that, as
+##  the messages state, can be safely ignored. Just rejects a couple proposals
+##  that result in ill-formed covariance matrices.
+##
+##  Author: Andrew Tredennick
+##  Date created: June 2, 2017
+
+##  Clear everything
+rm(list = ls(all.names = TRUE))
+
+##  Set working directory to source file location
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only for RStudio
+
+
+
+####
+####  LOAD LIBRARIES ----
+####
+library(tidyverse)    # Data munging
+library(dplyr)        # Data summarizing
+library(ggthemes)     # Pleasing ggplot themes
+library(stringr)      # Working with strings
+library(rstan)        # For MCMC
+library(lme4)         # Mixed-effects modeling
+
+
+
+####
+####  READ IN AND EXTRACT EXPERIMENT ANPP DATA ----
+####
+source("read_format_data.R")
+
+my_df <- anpp_data
+my_df$y <- as.numeric(scale( log( anpp_data$anpp ) ))
+m1 <- lmer( y ~ Treatment*vwc_scaled*year_id + (vwc_scaled|quadname), my_df)
+
+# attr(VarCorr(m1)$quadname, 'correlation')
+
+
+
+####
+####  SET UP AND FIT MODEL IN STAN ----
+####
+##  Set up fixed and random design matrices
+lmod <- lm(log(anpp) ~ Treatment*vwc_scaled*year_id, anpp_data)
+x    <- model.matrix(lmod)
+newx <- unique(x)
+lmod <- lm(log(anpp) ~ vwc_scaled, anpp_data)
+z    <- model.matrix(lmod)
+
+##  Make data list for Stan
+anppdat <- list(
+  Nobs     = nrow(anpp_data),
+  Npreds   = ncol(x),
+  Npreds2  = ncol(z),
+  Nplots   = length(unique(anpp_data$quadname)),
+  Ntreats  = length(unique(anpp_data$Treatment)),
+  Nyears   = length(unique(anpp_data$year)),
[…]

### 2017-12-09 code/notebooks/sensitivity_analysis.R

@@ -75,7 +75,7 @@ for(do_year in unique(droughts$year)){
   year_droughts <- filter(droughts, year==do_year)
   for(do_plot in unique(year_droughts$quadname)){
     plot_droughts <- filter(year_droughts, quadname == do_plot)
-    diffs <- year_controls$anpp - plot_droughts$anpp
+    diffs <- mean(year_controls$anpp) - plot_droughts$anpp
     tmp_out <- data.frame(treatment = "drought",
                           year = do_year,
                           quadname = do_plot,
@@ -90,7 +90,7 @@ for(do_year in unique(droughts$year)){
   year_irrigates <- filter(irrigates, year==do_year)
   for(do_plot in unique(year_irrigates$quadname)){
     plot_irrigates <- filter(year_irrigates, quadname == do_plot)
-    diffs <- year_controls$anpp - plot_irrigates$anpp
+    diffs <- mean(year_controls$anpp) - plot_irrigates$anpp
     tmp_out <- data.frame(treatment = "irrigation",
                           year = do_year,
                           quadname = do_plot,
@@ -123,8 +123,9 @@ ggplot(all_diffs, aes(x = year, y = sensitivity, color = treatment, fill = treat
         legend.background = element_rect(colour = NA, fill = NA))
 
 
-drought_fit <- lm(formula = sensitivity ~ year + quadname, data = filter(all_diffs, treatment == "drought"))
+drought_fit <- lm(formula = sensitivity ~ year, data = filter(all_diffs, treatment == "drought"))
 summary(drought_fit)
-
-irrigate_fit <- lm(formula = sensitivity ~ year + quadname, data = filter(all_diffs, treatment == "irrigation"))
+mod1 <- gls(sensitivity ~ year, data = filter(all_diffs, treatment == "drought"), correlation=corARMA(p=1,q=1))
+summary(mod1)
+irrigate_fit <- lm(formula = sensitivity ~ year, data = filter(all_diffs, treatment == "irrigation"))
 summary(irrigate_fit)

### 2017-12-09 manuscript/usses_water_ms.Rmd
more revisions; adds sensitivity analysis

@@
 To answer these questions we fit a generalized linear mixed effects model to test whether the regressions differed among treatments.
-We also analyzed community composition over time, allowing us to place our experimental results within the framework our scenarios (Fig. 1).
+We also analyzed community composition and \new{the sensitivity of ANPP to drought and irrigation treatments} over time, allowing us to place our experimental results within the framework our scenarios (Fig.
+1).
 
@@
 ## Data Analysis
+
+### Ecosystem functional response
+
 Our main goal was to test whether the relationship between ANPP and soil moisture differed among the drought, control, and irrigation treatments.
@@
 
+### Sensitivity of ANPP over time
+\new{Unfortunately, our data does not allow us to include the effect of time on ecosystem functional response in the model described above. 
+This is because we lack sufficient within-year and within-treatment variation of soil moisture (i.e., within a year and treatment each plot shares the same value of VWC). 
+Therefore, we conducted an independent analysis to determine if the sensitivity of ANPP to treatment changed over time.}
+
+\new{We define `sensitivity' as $\frac{ \text{ANPP}_{\text{control}} - \text{ANPP}_{\text{treatment}} } { \text{VWC}_{\text{control}} - \text{VWC}_{\text{treatment}} }$, following} @Wilcox2017.
+\new{This metric of sensitivity measures the response of ANPP standardized to the magnitude of the treatment.
+Because our treatment and control plots are not paired, we conducted this analysis by comparing each treatment plot to the mean of the control plots in each year.
+Then, independently for each treatment, we regressed sensitivity against year of treatment using the \texttt{lm()} function in R} [@R2016].
+\new{This analysis also allows us to link particularly sensitive treatment years to changes in community composition.}
+
+### Community composition over time
+
 We used nonmetric multidimensional scaling (NMDS) based on Bray-Curtis distances to identify temporal changes in community composition among treatments.
@@
 To test whether treatment plots were equally dispersed, or not, we used the ```vegan::betadisper()``` function [@Oksanen2016].
+
+\new{We conducted the above analysis for all species, and then conducted separate analyses for perennial species only and annual species only.
+Annual species have shorter life spans, so conducting separate analyses allows us to discover if community responses are restricted to short-lived species that might respond more quickly to altered precipitation regimes.
+Given the dominance of perennial species in our system, shifts in the annual plant community could be masked in the analysis of the full community.}
+
+### Reproducibility
 
@@
 Cumulative March-June soil moisture had a weak positive effect on ANPP (Table 1; Fig. 3B).
-The effect of soil moisture for each treatment is associated with high uncertainty, however, with 95\% Bayesian credible intervals that overlap zero (Table 1).
+The effect of soil moisture for each treatment is associated with high uncertainty, however, with 95\% Bayesian credible intervals that \new{broadly} overlap zero (Table 1).
 Although the parameter estimates for the effect of soil moisture overlap zero, the posterior distributions of the slopes all shrank and shifted to more positive values relative to the prior distributions (Fig.
 A2-2), which indicates the data did influence parameter estimates beyond the information from the uninformative priors.
-Ecosystem functional response was similar among treatments (Table 1; Fig. 3B), but there is evidence that the slope for the irrigation treatment is less than the slope for the control treatment.
-This evidence comes from interpreting the posterior distribution of the slope offset for the irrigation treatment, from which we calculate a \new{96}% one-tailed probability that the estimate is less than zero (Fig.
+Ecosystem functional response was similar among treatments (Table 1; Fig.
+3B), but there is evidence that the \new{slopes for the drought and} irrigation treatments are less than the slope for the control treatment.
+This evidence comes from interpreting the posterior distributions of the slope offsets for the treatments.
+\new{From these distributions,} we calculate a \new{91}% one-tailed probability that the estimate is less than zero for the irrigation treatment \new{and a 96\% one-tailed probability that the estimate is less than zero for the drought treatment} (Fig.
 3A, right panel).
+
+\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. x). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0002; Fig. x)}.
 
@@
 Community composition was also remarkably stable over time, with no evidence of divergence among treatments (Table 2; Fig. 4).
+\new{Splitting the analysis apart for perennials and annuals led to similar results.
+The perennial community was stable over time and consistently similar among treatments (Fig. Ax-x).
+There is some evidence that the annual community changed in response to treatment in two years (Fig.
+Ax-x), but these responses appear to come from very rare species since our analysis weights them equally (Appendix X).}
 
@@
 
-The results of our five year experiment in a sagebrush steppe conform to two of our four predictions, depending on treatment.
-Ecosystem functional response and community composition did not change under chronic drought (Fig. 3A, Fig. 4), representing the top left scenario in Fig. 1.
-Ecosystem functional response under chronic irrigation was different from the control treatment, but community composition remained unchanged (Fig.
+The results of our five year experiment in a sagebrush steppe \new{conforms to one} of our four predictions.
+Ecosystem functional response under chronic \new{drought and irrigation} was different from the control treatment, but community composition remained unchanged (Fig.
 3A, Fig.
@@
 This suggests that if average soil moisture were pushed consistently higher than currently observed ambient conditions, there would be a weaker relationship between precipitation and productivity in this system.
-Nonetheless, the slopes of the VWC $\times$ productivity relationship were similar among treatments (Table 1), at least at the level of irrigation we imposed.
-We therefore conclude that ecosystem functional response is consistent and weak across all precipitation treatments.
+
+\new{The decrease in the slope of the VWC $\times$ productivity relationship in the droughted plots is puzzling because it is counter to our expectation that the slope would be higher than the control plots.
+However, we believe the significance of the slope offsets has more to do with noise than signal.
+This is because the full distributions of} the slopes of the VWC $\times$ productivity relationship, \new{not just the offsets}, were similar among treatments (Table 1).
+We therefore conclude that ecosystem functional response is consistent (\new{similar values}) and weak (\new{all broadly overlapping zero}) across all precipitation treatments.
 
@@
 Why did our treatments fail to induce ecosystem or community responses?
-We can think of three reasons; two are limitations of our study, and one is the life history traits of the species in our focal communities.
+We can think of three reasons. 
+Two are limitations of our study, and one is the life history traits of the species in our focal communities.
 We first discuss the potential limitations of our study, and then discuss the biological explanation.
@@
 
+\new{We did find that ANPP became more sensitive to drought over time (Fig. x).
+Thus, the difference between ANPP, standardize by the difference in soil moisture, in drought and control plots grew over time.
+The latter years of our experiment were relatively wet, meaning that the drought treatment could have the largest absolute reduction of soil moisture in those years.
+But our metric of sensitivity accounts for differences among years because ANPP differences is standardized by the difference in soil moisture within the year.
+Therefore, it is likely that the increased sensitivity to drought over time is due to the cimulative impact of drought on the dominant species.
+Ecosystem functional response may soon shift if sensitivity continues to increase.}
+
 # CONCLUSIONS
 
-Our results suggest the species in our focal plant community are insensitive to to changes in precipitation regime. 
+Our results suggest the species in our focal plant community are insensitive to to changes in precipitation regime, \new{at least over the five years of our experiment}. 
 Such insensitivity can buffer species against precipitation variability in this semi-arid ecosystem, making them successful in the long run.
@@
 Our results suggest compositional shifts would have the largest impact at high rainfall because the current community maintained consistent ecosystem functional response at very low water availability.
+\new{However, the increased sensitivity of ANPP to drought conditions suggests a shift in ecosystem functional response may be near, and simply take more time than we observed here.}
 

### 2017-12-09 manuscript/usses_water_ms.tex
more revisions; adds sensitivity analysis

@@
 linear mixed effects model to test whether the regressions differed
-among treatments. We also analyzed community composition over time,
-allowing us to place our experimental results within the framework our
-scenarios (Fig. 1).
+among treatments. We also analyzed community composition and
+\new{the sensitivity of ANPP to drought and irrigation treatments} over
+time, allowing us to place our experimental results within the framework
+our scenarios (Fig. 1).
 
@@
 
+\subsubsection{Ecosystem functional
+response}\label{ecosystem-functional-response}
+
 Our main goal was to test whether the relationship between ANPP and soil
@@
 
+\subsubsection{Sensitivity of ANPP over
+time}\label{sensitivity-of-anpp-over-time}
+
+\new{Unfortunately, our data does not allow us to include the effect of time on ecosystem functional response in the model described above. 
+This is because we lack sufficient within-year and within-treatment variation of soil moisture (i.e., within a year and treatment each plot shares the same value of VWC). 
+Therefore, we conducted an independent analysis to determine if the sensitivity of ANPP to treatment changed over time.}
+
+\new{We define `sensitivity' as $\frac{ \text{ANPP}_{\text{control}} - \text{ANPP}_{\text{treatment}} } { \text{VWC}_{\text{control}} - \text{VWC}_{\text{treatment}} }$, following}
+\citet{Wilcox2017}.
+\new{This metric of sensitivity measures the response of ANPP standardized to the magnitude of the treatment.
+Because our treatment and control plots are not paired, we conducted this analysis by comparing each treatment plot to the mean of the control plots in each year.
+Then, independently for each treatment, we regressed sensitivity against year of treatment using the \texttt{lm()} function in R}
+\citep{R2016}.
+\new{This analysis also allows us to link particularly sensitive treatment years to changes in community composition.}
+
+\subsubsection{Community composition over
+time}\label{community-composition-over-time}
+
 We used nonmetric multidimensional scaling (NMDS) based on Bray-Curtis
@@
 \texttt{vegan::betadisper()} function \citep{Oksanen2016}.
+
+\new{We conducted the above analysis for all species, and then conducted separate analyses for perennial species only and annual species only.
+Annual species have shorter life spans, so conducting separate analyses allows us to discover if community responses are restricted to short-lived species that might respond more quickly to altered precipitation regimes.
+Given the dominance of perennial species in our system, shifts in the annual plant community could be masked in the analysis of the full community.}
+
+\subsubsection{Reproducibility}\label{reproducibility}
 
@@
 associated with high uncertainty, however, with 95\% Bayesian credible
-intervals that overlap zero (Table 1). Although the parameter estimates
-for the effect of soil moisture overlap zero, the posterior
-distributions of the slopes all shrank and shifted to more positive
-values relative to the prior distributions (Fig. A2-2), which indicates
-the data did influence parameter estimates beyond the information from
-the uninformative priors. Ecosystem functional response was similar
-among treatments (Table 1; Fig. 3B), but there is evidence that the
-slope for the irrigation treatment is less than the slope for the
-control treatment. This evidence comes from interpreting the posterior
-distribution of the slope offset for the irrigation treatment, from
-which we calculate a \new{96}\% one-tailed probability that the estimate
-is less than zero (Fig. 3A, right panel).
+intervals that \new{broadly} overlap zero (Table 1). Although the
+parameter estimates for the effect of soil moisture overlap zero, the
+posterior distributions of the slopes all shrank and shifted to more
+positive values relative to the prior distributions (Fig. A2-2), which
+indicates the data did influence parameter estimates beyond the
+information from the uninformative priors. Ecosystem functional response
+was similar among treatments (Table 1; Fig. 3B), but there is evidence
+that the \new{slopes for the drought and} irrigation treatments are less
+than the slope for the control treatment. This evidence comes from
+interpreting the posterior distributions of the slope offsets for the
+treatments. \new{From these distributions,} we calculate a \new{91}\%
+one-tailed probability that the estimate is less than zero for the
+irrigation treatment
+\new{and a 96\% one-tailed probability that the estimate is less than zero for the drought treatment}
+(Fig. 3A, right panel).
+
+\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. x). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0002; Fig. x)}.
 
@@
 divergence among treatments (Table 2; Fig. 4).
+\new{Splitting the analysis apart for perennials and annuals led to similar results.
+The perennial community was stable over time and consistently similar among treatments (Fig. Ax-x).
+There is some evidence that the annual community changed in response to treatment in two years (Fig.
+Ax-x), but these responses appear to come from very rare species since our analysis weights them equally (Appendix X).}
 
@@
 
-The results of our five year experiment in a sagebrush steppe conform to
-two of our four predictions, depending on treatment. Ecosystem
-functional response and community composition did not change under
-chronic drought (Fig. 3A, Fig. 4), representing the top left scenario in
-Fig. 1. Ecosystem functional response under chronic irrigation was
-different from the control treatment, but community composition remained
-unchanged (Fig. 3A, Fig. 4), representing the top right scenario in Fig.
-1. The decrease in the slope of the VWC \(\times\) productivity
-relationship in the irrigated plots is consistent with a weaker response
-of productivity to additional water in this ecosystem. This suggests
-that if average soil moisture were pushed consistently higher than
-currently observed ambient conditions, there would be a weaker
-relationship between precipitation and productivity in this system.
-Nonetheless, the slopes of the VWC \(\times\) productivity relationship
-were similar among treatments (Table 1), at least at the level of
-irrigation we imposed. We therefore conclude that ecosystem functional
-response is consistent and weak across all precipitation treatments.
+The results of our five year experiment in a sagebrush steppe
+\new{conforms to one} of our four predictions. Ecosystem functional
+response under chronic \new{drought and irrigation} was different from
+the control treatment, but community composition remained unchanged
+(Fig. 3A, Fig. 4), representing the top right scenario in Fig. 1. The
+decrease in the slope of the VWC \(\times\) productivity relationship in
+the irrigated plots is consistent with a weaker response of productivity
+to additional water in this ecosystem. This suggests that if average
+soil moisture were pushed consistently higher than currently observed
+ambient conditions, there would be a weaker relationship between
+precipitation and productivity in this system.
+
+\new{The decrease in the slope of the VWC $\times$ productivity relationship in the droughted plots is puzzling because it is counter to our expectation that the slope would be higher than the control plots.
+However, we believe the significance of the slope offsets has more to do with noise than signal.
+This is because the full distributions of} the slopes of the VWC
+\(\times\) productivity relationship, \new{not just the offsets}, were
+similar among treatments (Table 1). We therefore conclude that ecosystem
+functional response is consistent (\new{similar values}) and weak
+(\new{all broadly overlapping zero}) across all precipitation
+treatments.
 
@@
 Why did our treatments fail to induce ecosystem or community responses?
-We can think of three reasons; two are limitations of our study, and one
+We can think of three reasons. Two are limitations of our study, and one
 is the life history traits of the species in our focal communities. We
@@
 
+\new{We did find that ANPP became more sensitive to drought over time (Fig. x).
+Thus, the difference between ANPP, standardize by the difference in soil moisture, in drought and control plots grew over time.
+The latter years of our experiment were relatively wet, meaning that the drought treatment could have the largest absolute reduction of soil moisture in those years.
+But our metric of sensitivity accounts for differences among years because ANPP differences is standardized by the difference in soil moisture within the year.
+Therefore, it is likely that the increased sensitivity to drought over time is due to the cimulative impact of drought on the dominant species.
+Ecosystem functional response may soon shift if sensitivity continues to increase.}
+
 \section{CONCLUSIONS}\label{conclusions}
@@
 Our results suggest the species in our focal plant community are
-insensitive to to changes in precipitation regime. Such insensitivity
+insensitive to to changes in precipitation regime,
+\new{at least over the five years of our experiment}. Such insensitivity
 can buffer species against precipitation variability in this semi-arid
@@
 ecosystem functional response at very low water availability.
+\new{However, the increased sensitivity of ANPP to drought conditions suggests a shift in ecosystem functional response may be near, and simply take more time than we observed here.}
 

### 2017-12-09 code/04_sensitivity_analysis.R

@@ -125,7 +125,5 @@ ggplot(all_diffs, aes(x = year, y = sensitivity, color = treatment, fill = treat
 
 drought_fit <- lm(formula = sensitivity ~ year, data = filter(all_diffs, treatment == "drought"))
 summary(drought_fit)
-mod1 <- gls(sensitivity ~ year, data = filter(all_diffs, treatment == "drought"), correlation=corARMA(p=1,q=1))
-summary(mod1)
 irrigate_fit <- lm(formula = sensitivity ~ year, data = filter(all_diffs, treatment == "irrigation"))
 summary(irrigate_fit)

### 2017-12-09 code/04_sensitivity_analysis.R
adds sensitivity output to results folder

@@ -13,7 +13,6 @@ rm(list = ls(all.names = TRUE))
 
 ##  Set working directory to source file location
 setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only for RStudio
-setwd("../")
 
 
 ####
@@ -55,8 +54,8 @@ sens_vwc <- anpp_data %>%
 sensitivity <- left_join(sens_anpp, sens_vwc, by = c("year", "comparison")) %>%
   mutate(sens = anpp_diff / vwc_diff)
 
-ggplot(sensitivity, aes(x=year, y=sens, color=comparison))+
-  geom_line()
+# ggplot(sensitivity, aes(x=year, y=sens, color=comparison))+
+#   geom_line()
 
 
 ##  By plot-treatment -- compare each treatment plot to each control in each year
@@ -108,22 +107,38 @@ all_diffs <- rbind(drought_diffs, irrigate_diffs) %>%
   left_join(vwc_diffs, by = c("year","treatment")) %>%
   mutate(sensitivity = cntrl_minus_treat / vwc_diff)
 
-mycols <- c("#009E73", "#D55E00", "#0072B2")
-treat_cols <- mycols[2:3]
-ggplot(all_diffs, aes(x = year, y = sensitivity, color = treatment, fill = treatment))+
-  geom_jitter(shape = 21, width = 0.1, color = "grey35", alpha = 0.6)+
-  stat_smooth(se=F, method="lm")+
-  scale_fill_manual(values = treat_cols, name = NULL, labels = c("Drought","Irrigation"))+
-  scale_color_manual(values = treat_cols, name = NULL, labels = c("Drought","Irrigation"))+
-  ylab("Sensitivity")+
-  xlab("Year")+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank(),
-        legend.position = c(0.12, 0.9),
-        legend.background = element_rect(colour = NA, fill = NA))
+saveRDS(object = all_diffs, file = "../results/sensitvities.RDS")
 
 
-drought_fit <- lm(formula = sensitivity ~ year, data = filter(all_diffs, treatment == "drought"))
-summary(drought_fit)
-irrigate_fit <- lm(formula = sensitivity ~ year, data = filter(all_diffs, treatment == "irrigation"))
-summary(irrigate_fit)
+
+####
+####  FIT LINEAR MODELS FOR EFFECT OF TIME -------------------------------------
+####
+drought_fit <- lm(formula = sensitivity ~ year, 
[…]

### 2017-12-09 manuscript/usses_water_ms.Rmd
adds sensitivity analysis to figure 3

@@
       \includegraphics[width=5in]{../figures/glmm_main_results.png}
-  \caption{Results from the generalized linear mixed effects model.
+  \caption{Results from the generalized linear mixed effects model \new{(A-B) and the sensitivity analysis (C)}.
 (A) Posterior distributions for the effects of drought and irrigation on the intercept and slope of the productivity-soil moisture relationship.
@@
 Note that we show log(ANPP) on the \emph{y}-axis of panel B; this same plot can be seen on the arithmetic scale in supporting material Fig.
-A2-1.}
+A2-1.
+\new{(C) Regression of sensitivity against time for each treatment.
+Each point represents the sensitivity of ANPP in a plot relative to the mean of the control plots in that year.
+Only the significant regression for the drought treatment is plotted (\emph{P} = 0.0002).}}
 \end{figure}

### 2017-12-09 code/04_sensitivity_analysis.R

@@ -126,19 +126,4 @@ data.frame(treatment = c("drought", "irrigation"),
            pvalues = pvals) %>%
   write_csv("../results/sensitivity_year_pvalues.csv")
 
-# mycols <- c("#009E73", "#D55E00", "#0072B2")
-# treat_cols <- mycols[2:3]
-# ggplot(all_diffs, aes(x = year, y = sensitivity, color = treatment, fill = treatment))+
-#   geom_jitter(shape = 21, width = 0.1, color = "grey35", alpha = 0.6)+
-#   stat_smooth(se=F, method="lm")+
-#   scale_fill_manual(values = treat_cols, name = NULL, labels = c("Drought","Irrigation"))+
-#   scale_color_manual(values = treat_cols, name = NULL, labels = c("Drought","Irrigation"))+
-#   ylab("Sensitivity")+
-#   xlab("Year")+
-#   theme_bw()+
-#   theme(panel.grid.minor = element_blank(),
-#         legend.position = c(0.12, 0.9),
-#         legend.background = element_rect(colour = NA, fill = NA))
-
-
 

### 2017-12-09 code/05_plot_model_results.R

@@ -166,19 +166,63 @@ suppressWarnings( # ignore wanrnings about NA values
   regress_plot <- ggplot(anpp_data, aes(x=total_seasonal_vwc,y=log(anpp)))+
     geom_point(shape=21,color="grey25",alpha=0.8,aes(fill=Treatment))+
     geom_line(data=regression_limited, aes(x=backtrans_vwc, y=backtrans_estimate, color=Treatment), size=1)+
-    scale_fill_manual(values = mycols)+
-    scale_color_manual(values = mycols)+
+    scale_fill_manual(values = mycols, name = NULL)+
+    scale_color_manual(values = mycols, name = NULL)+
     xlab("March-June Cumulative VWC")+
     ylab(expression(paste("log[ANPP (g ",m^2,")]")))+
     theme_bw()+
-    theme(panel.grid.minor = element_blank())
+    theme(panel.grid.minor = element_blank(),
+          legend.position = c(0.12, 0.85),
+          legend.background = element_rect(colour = NA, fill = NA),
+          legend.text=element_text(size=7),
+          legend.key.size = unit(0.4, "cm"))
 )
 
-suppressWarnings( # ignore wanrnings about NA values
-  gridplot <- cowplot::plot_grid(treat_posteriors, NULL, regress_plot, 
-                                 rel_heights = c(1, 0.08, 1),
-                                 nrow = 3, 
-                                 labels = c("A)","","B)"))
+
+####
+####  SENSITIVITIES ------------------------------------------------------------
+####
+all_diffs <- readRDS("../results/sensitvities.RDS")
+sens_plot <- ggplot(all_diffs, aes(x = year, y = sensitivity, fill = treatment))+
+    geom_jitter(shape = 21, width = 0.1, color = "grey25", alpha=0.8)+
+    stat_smooth(data = filter(all_diffs, treatment == "drought"), color = treat_cols[1], se=F, method="lm")+
+    scale_fill_manual(values = treat_cols, name = NULL, labels = c("Drought","Irrigation"))+
+    scale_color_manual(values = treat_cols, name = NULL, labels = c("Drought","Irrigation"))+
+    ylab("Sensitivity")+
+    xlab("Year")+
+    guides(color = FALSE, fill = FALSE)+
+    theme_bw()+
+    theme(panel.grid.minor = element_blank(),
+          legend.position = c(0.12, 0.9),
+          legend.background = element_rect(colour = NA, fill = NA))
+
+
+
+####
[…]

### 2017-12-09 code/05_plot_model_results.R
tidys up code a bit

@@ -201,6 +201,7 @@ sens_plot <- ggplot(all_diffs, aes(x = year, y = sensitivity, fill = treatment))
 ####
 ####  COMBINE PLOTS AND SAVE ---------------------------------------------------
 ####
+##  Version without sensitivity analysis
 # suppressWarnings( # ignore warnings about NA values
 #   gridplot <- cowplot::plot_grid(treat_posteriors, NULL, regress_plot, sens_plot, 
 #                                  rel_heights = c(1, 0.08, 1),
@@ -209,6 +210,7 @@ sens_plot <- ggplot(all_diffs, aes(x = year, y = sensitivity, fill = treatment))
 #                                  labels = c("A)","","B)","C)"))
 # )
 
+##  Version with sensitivity analysis
 suppressWarnings( # ignore warnings about NA values
   bottom_row <- plot_grid(regress_plot, NULL, sens_plot, 
                           labels = c('B','', 'C'),

### 2017-12-09 manuscript/review_response_R1.Rmd
more revision letter writing

@@
 \response{We provide point-by-point responses to reviewer comments below.
-We have taken care to address Reviewer 1's concerns (also echoed by Reviewer 2) by substantially revising our Introduction section.
-In particular, we have shifted our focus away from the Hierarchical Response Framework, and instead focus on how species' responses and community compositional shifts can impact ecosystem functional response.
-We do this, rather than implementing the suggestions of Reviewer 1, because our data limits the complexity of our statistical model.
+We have taken care to address Reviewer 1's concerns (also echoed by Reviewer 2) by including the suggested sensitivity analysis (see new section in Methods, lines xxx-xxx).
+Unfortunately, we could not implement the suggestion of Reviewer 1 to include a VWC$\times$Treatment$\times$Year interaction in our model because our data limits the complexity of our statistical model.
 Within a year and treatment, each plot has the same value of volumetric water content.
@@
 
-\response{We therefore use the Hierarchical Response Framework to support our hypotheses about potential mechanisms driving change in ecosystem functional response (species' responses and community change), but refrain from suggesting we are explicitly testing the Hierarchical Response Framework.
-We also circle back to this point in our Discussion (lines xxx-xxx).
-Fortunately, our community-level analysis does have an explicit temporal component, which does link back to the Hierarchical Response Framework.}
+\response{The sensitivity analysis allows us to examine how the magnitude of the treatment effect changes over time.
+We found that the response of ANPP to drought grew over time, while the response of ANPP to irrigation is static (new panel D in Figure 3).
+We think this analysis improves the connection of our results to the Hierarchical Response Framework.}
 
@@
 \response{The reviewer is correct that our analysis is not an explicit test of the Hierarchical Response Framework because we did not explicitly model the effect of time on ecosystem functional response.
-Unfortunately, we are unable to do so given our data -- in each year every plot shares the same average soil moisture for its particular treatment, making it difficult to fit year-specific slopes for each treatment level.
-The sensitivity analysis would also be interesting, but moves us farther away from our interest in the ecosystem functional response.
-To remedy the situation, we have revised our motivation in the introduction to focus on the mechanisms that could alter ecosystem functional response, physiological responses and changes in community composition, without attributing specific time scales to them (lines xxx-xxx).
-Thus, we no longer pitch our paper as a test of the Hierarchical Response Framework, but we do rely on it to motivate our study.
-We also note that a community change may have to happen relatively early in the time series to impact ecosystem functional response (line xxx).
-Fortunately, our community level analysis allows to investigate this (and we found that the communities do not diverge over time, or change at all).}
+Unfortunately, we are unable to do so given our data -- in each year every plot shares the same average soil moisture for its particular treatment, making it difficult to fit year-specific slopes for each treatment level.}
+
+\response{That said, the reviewer made an excellent suggestion of conducting a sensitivity analysis \emph{a la} Wilcox et al.
+2017, which we have included (see lines xxx - xxx in the Methods section; new results on lines xxx - xxx; and new discussion items on lines xxx-xxx).
+The sensitivity analysis allows us to examine how the effect of treatment changes over time, adding another dimension to our analysis that better connects our findings to the Hierachical Response Framework.
+We found some evidence that the ecosystem's sensitivity to drought increased over time.
+This increase was not large enough to significantly impact ecosystem functional response, and our community composition analysis does not indicate changes in the final years of our experiment.
+However, as we note in the final line of our paper (lines xxx-xxx), the increased sensitivity in the latter years of the experiment may portend a coming shift in ecosystem functional response.}
 

### 2017-12-10 manuscript/review_response_R1.Rmd
more tiny revisions and matching up with response letter

@@
 \response{This is a great suggestion, and one we had not thought of.
-We conducted the analysis suggested by the reviewer and found that community composition was stable for both annuals (short-lived) and perennials (long-lived).
+We conducted the analysis suggested by the reviewer and found that community composition was also stable when we restricted the analysis to annual plants (short-lived).
 In a few years we detected signficant treatment effects on dispersion or mean of community composition.
@@
 
-\response{Excellent suggestion. We have added this to the discussion on lines xxx-xxx, and also allude to dominant species' effects in the introduction on lines xx-xx.}
+\response{Thanks for bringing this paper to our attention. We have included a new sentence that references this citation on line xxx in the Discussion.}
 

### 2017-12-10 manuscript/Appendix1.Rmd
more tiny revisions and matching up with response letter

@@
   - \usepackage{lineno}
+  - \linenumbers
   - \usepackage{setspace}
   - \usepackage{todonotes}
-  - \linespread{1.25}
+  - \onehalfspacing
   - \usepackage{rotating}
@@
   - \usepackage[sc]{mathpazo}
-  - \usepackage{mathtools}
-bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
-csl: components/ecology_letters.csl
+  - \usepackage{float}
+bibliography: ~/Dropbox/Bibliography/usses_water.bib
+csl: components/ecology.csl
 
@@
 
+\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
 \renewcommand\thefigure{A1-\arabic{figure}}
+\renewcommand\theequation{A1.\arabic{equation}}
 \renewcommand\thetable{A1-\arabic{table}}  
@@
 \textbf{\Large{Appendix 1}} \\
-A.T. Tredennick, A.R. Kleinhesselink, B. Taylor \& P.B. Adler \\
+A.T. Tredennick, A.R. Kleinhesselink, J.B. Taylor \& P.B. Adler \\
 ``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe'' \\
@@
 
-\linenumbers
+# Details on the plant community and dynamics
 
-\section{Estimating ANPP}
-We used a radiometer to nondestructively estimate aboveground net primary productivity.
-Our approach relies on relating greenness in a plot to aboveground biomass.
-In each year we recorded ground reflectances at four bands, two associated with the red spectrum and two associated with the near-infrared spectrum (Table A1-1).
-We took four readings per plot that were averaged for each band.
-Bands 1 and 3 correspond to wavelengths collected by the MODIS satellite and bands 2 and 4 correspond to wavelengths collected by the AVHRR satellite.
+Here we provide more details on the plant community in terms of dominance and rarity.
+Averaging across time, *Artemisia tripartita* and *Balsamorhiza sagittata* are the two most dominant species in each treatment.
+Combined, these two species represent 28\% of total cover in control plots, 25\% of total cover in drought plots, and 25\% of total cover in irrigation plots.
+Four to five species dominate the community in general (Figure A1-1), indicating a high level of dominance in this plant community.
 
-| Band number | Spectrum name | Center wavelength | Corresponding satellite |
-| ----------- | ------------- | ----------------- | ----------------------- |
-| 1 | red | 626 nm | AVHRR |
-| 2 | red | 652 nm | MODIS |
-| 3 | near-infrared | 875 nm | AVHRR |
-| 4 | near-infrared | 859 nm | MODIS |
+We also conducted our community composition analysis with only annual species.
+Annual species are shorter-lived than the perennial species in our community, so they may respond more quickly to alterations of precipitation.
+In general, our results for annual species conform to the results from the full community analysis in the main text.
+Annual plant community composition is relatively stable through time (Fig. A1-6) and in most years there is no evidence that treatment differentiates community composition (Table A1-3).
+Note that in some years the ```vegan::metaMDS()``` returned unreliable estimates of Bray-Curtis distances for the annual community because of lack of sufficient data (i.e., many annual species with 0 abundance).
 
-Table: Radiometer specifications.
+\newpage{}
+# Tables
 
-Using the RED and NIR reflectance values, we calculate the normalized difference vegetation index (NDVI) for each plot based on both AVHRR- and MODIS-based wavelengths.
-We calculated NDVI as:
-\vspace{-2em}
-\begin{align}
-\text{NDVI}_{\text{AVHRR}} &= \frac{b_3 - b_1}{b_3 + b_1} \\[10pt]
-\text{NDVI}_{\text{MODIS}} &= \frac{b_4 - b_2}{b_4 + b_2}
-\end{align}
+```{r get_data, echo=FALSE, message=FALSE, warning=FALSE}
+####
+####  LOAD PACKAGES ----
+####
+library(tidyverse) # all the good things from Wikham
+library(dplyr)     # data munging
+library(stringr)   # working with strings
+library(ggthemes)  # pleasing ggplot2 themes
+library(vegan)     # NMDS and community analysis
+library(xtable)    # Generating LaTeX tables in RMarkdown
 
-\noindent{} where $b_x$ refers to band *x* (*x* = 1,2,3,4) in Table A1-1.
+####
+####  SET DIRECTORIES; READ IN DATA ----
+####
+# Set paths
+biomass_path <- "../data/estimated_biomass/"
+weather_path <- "../data/weather/"
+output_path  <- "../output/"
+figure_path  <- "../figures/"
 
-To convert plot NDVI to biomass, we regressed known biomass values from calibration plots against NDVI calculate for those plots.
-Calibration plots were located near our experiment plots, and each year we located a new set of 10 plots in which we clipped all aboveground biomass, dried it to a constant weight at 60$^\circ$ C, and the weighed.
-We used these biomass values to estimate regression parameters for both AVHRR- and MODIS-based NDVI.
-We assessed model fit using $R^2$ and, for each year, we used the regression parameters associated with the best fit model to estimate biomass in the experimental plots based on their NDVI values (Table A1-2).
-R code for this procedure is in the file "calibrate_radiometer_by_year.R" in the supplemental code set.
+# Read in data
+plot_biomass <- readRDS(paste0(biomass_path,"permanent_plots_estimated_biomass.RDS"))
+weather_data <- read.csv(paste0(weather_path,"ClimateIPM.csv"))
+plot_info    <- read.csv("../data/estimated_biomass/quad_info.csv")
 
-```{r ndvi_table, include=T, echo=F, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
-library(xtable)
-library(tidyverse)
-library(dplyr)
-tab_dat <- read.csv("../data/estimated_biomass/ndvi_biomass_regression_parameters.csv", row.names = 1) %>%
-  filter(year > 2011 & year < 2017)
-colnames(tab_dat) <- c("Year", "Intercept", "NDVI Slope", "$R^2$", "Min(biomass)", "Max(biomass)", "\\emph{N}", "Algorithm")
+suppressWarnings( # ignore factor to character warning. It's OK.
+  cover_species <- read.csv("../data/vital_rates/allrecords_cover_v24.csv") %>%
+    group_by(quad, year, species) %>%
+    summarise(total_area = sum(area)) %>%
+    left_join(plot_info, by = "quad") %>%
+    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
+    filter(!str_detect(QuadName, 'P1|P7')) %>%
+    group_by(year, species, Treatment) %>%
+    summarise(avg_area = mean(total_area)) %>%
+    ungroup()
+)
 
-my_caption <- "Details of regression models used to estimate biomass each year."
-print(xtable(tab_dat, caption = my_caption, digits = 2),
+####
+####  COMPARE COMMUNITY COMPOSITION AMONG TREATMENTS ----
+####  USING STANDARDIZED ABUNDANCES FROM COVER AND DENSITY
+####
+suppressWarnings( # ignore factor to character warning. It's OK.
+  cover_community_matrix <- read.csv("../data/vital_rates/allrecords_cover_v24.csv") %>%
+    group_by(quad, year, species) %>%
+    summarise(total_area = sum(area)) %>%
+    left_join(plot_info, by = "quad") %>%
+    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
+    filter(!str_detect(QuadName, 'P1|P7')) %>%
+    filter(year > 2010 & is.na(species)==F) %>%
+    spread(species,total_area, fill = 0) %>%
+    select(-QuadName,-Grazing,-paddock,-Group) %>%
+    ungroup()
+)
+
+suppressWarnings( # ignore factor to character warning. It's OK.
+  density_community_matrix <- read.csv("../data/vital_rates/allrecords_density_v24.csv") %>%
+    mutate(ind_num = 1) %>%
+    group_by(year, quad, species) %>%
+    summarise(total_inds = sum(ind_num)) %>%
+    left_join(plot_info, by = "quad") %>%
+    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
+    filter(!str_detect(QuadName, 'P1|P7')) %>%
+    filter(year > 2010 & is.na(species)==F) %>%
+    spread(species,total_inds, fill = 0) %>%
+    select(-QuadName,-Grazing,-paddock,-Group) %>%
+    ungroup()
+)
+
+density_spp <- colnames(density_community_matrix[4:ncol(density_community_matrix)])
+cover_spp   <- colnames(cover_community_matrix[4:ncol(cover_community_matrix)])
+overlap_spp <- density_spp[which(density_spp %in% cover_spp)]
+rmids       <- which(colnames(density_community_matrix) %in% overlap_spp)
+density_community_matrix <- density_community_matrix[,-rmids]
+
+suppressWarnings( # ignore factor to character warning. It's OK.
+  annual_community_matrix <- read.csv("../data/vital_rates/idaho_annuals.csv") %>%
+    select(-Notes) %>%
+    gather(species,abundance,-c(1:3)) %>%
+    left_join(plot_info, by="quad") %>%
+    select(-QuadName.y) %>%
+    rename(QuadName = QuadName.x) %>%
+    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
+    filter(!str_detect(QuadName, 'P1|P7')) %>%
+    filter(year > 2010 & is.na(species)==F) %>%
+    separate(species,c("genus","species1")) %>%
+    unite(species,genus,species1,sep = " ") %>%
+    spread(species,abundance, fill = 0) %>%
+    select(-QuadName,-Grazing,-paddock,-Group,-`Lappula occidentalis`,-`Tragopogon dubius`) %>%
+    ungroup()
+)
+
+annual_data <- {}
+for(dotrt in unique(annual_community_matrix$Treatment)){
+  for(doyr in unique(annual_community_matrix$year)){
+    tmp <- annual_community_matrix %>%
+      gather(species, density, -quad, -year, -Treatment) %>%
+      filter(Treatment == dotrt, year == doyr) %>%
+      group_by(Treatment, year, species) %>%
+      summarise(mean_density = mean(density)) %>%
+      arrange(-mean_density) %>%
+      mutate(mean_rank = seq_along(mean_density))
+
+    annual_data <- rbind(annual_data, tmp)
+  }
+}
+
+dom_annuals <- annual_data %>%
+  filter(species %in% c("Allyssum desertorum","Bromus tectorum")) %>%
+  select(-mean_rank) %>%
+  left_join(weather_data, by = "year")
+```
+
+Table A1-1: Statistical results from linear model relating density of Allysum desertorum to precipitation.
+```{r report_models, echo = FALSE}
+summary(lm(mean_density~ppt1*Treatment, data = filter(dom_annuals, species == "Allyssum desertorum")))
+```
+
+\newpage{}
+Table A1-2: Statistical results from linear model relating density of Bromus tectorum to precipitation.
+```{r report_models2, echo = FALSE}
+summary(lm(mean_density~ppt1*Treatment, data = filter(dom_annuals, species == "Bromus tectorum")))
+```
+
+\newpage{}
+
+```{r annual_nmds, echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}
+####
+####  ANALYZE JUST THE ANNUALS ----
+####
+# Set missing species abundances as 0, which == absent
+annual_community_matrix[is.na(annual_community_matrix)] <- 0
+
+annual_community_matrix_scaled <- cbind(annual_community_matrix[,1:3,drop=F],
+                                        scale(annual_community_matrix[,4:ncol(annual_community_matrix)]))
+
+spp_to_remove <- names(which(is.nan(colMeans(annual_community_matrix_scaled[4:ncol(annual_community_matrix_scaled)]))))
+ids_to_remove <- which(colnames(annual_community_matrix_scaled) %in% spp_to_remove)
+annual_community_matrix_scaled <- annual_community_matrix_scaled %>%
+  select(-ids_to_remove)
+
+nmds_df_annuals <- {}
+out_stats <- {}
+for(doyr in unique(annual_community_matrix_scaled$year)){
+  tmp <- filter(annual_community_matrix_scaled, year == doyr) %>% select(-year)
+   invisible( capture.output( # already checked for convergence
+    tmp_bray <- metaMDS(tmp[3:ncol(tmp)]+2, "bray", trymax = 999)
+  ))
+  nmds_points <- as.data.frame(tmp_bray$points)
+  nmds_points$Treatment <- tmp$Treatment
+  nmds_points$Year <- doyr
+  nmds_df_annuals <- rbind(nmds_df_annuals, nmds_points)
+  adonis_stats <- as.data.frame(vegan::adonis(tmp[3:ncol(tmp)]+2~tmp$Treatment)$aov.tab)[1,]
+  betadisper_stats <- anova(vegan::betadisper(vegan::vegdist(tmp[3:ncol(tmp)]+2),tmp$Treatment))[1,]
+  tmp_out <- rbind(adonis_stats[,-which(colnames(adonis_stats)=="R2")],
+                   as.numeric(betadisper_stats))
+  row.names(tmp_out) <- c()
+  out_stats <- rbind(out_stats, data.frame(year = doyr,
+                                           test = c("adonis","betadisper"),
+                                           n = nrow(tmp),
+                                           tmp_out))
+}
+```
+
+\setcounter{table}{2}
+
+```{r annual_nmds_table, echo = FALSE, include=TRUE, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
+colnames(out_stats) <- c("Year", "Test", "n", "d.f.","Sums","Means","\\emph{F}", "\\emph{P}")
+out_stats$d.f. <- as.integer(out_stats$d.f.)
+out_stats <- arrange(out_stats, Year)
+
+# Make the table
+inyear_caption <- "Results from statistical tests for clustering and dispersion of community composition among precipitation treatments for annual species only.
+`adonis' tests whether treatments form unique clusters in multidimensial space; `betadisper' tests whether treatments have similar dispersion.
+For both tests, \\emph{P} values greater than 0.05 indicate there is no support that the treatments differ."
+print(xtable(dplyr::select(out_stats, -Sums, -Means), caption = inyear_caption, digits = 2),
        caption.placement="top",
@@
 ```
+
+\newpage{}
+
+# Figures
+
+```{r rac_plots, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Rank abundance curves for perennial species.
+Basal area of individuals was summed withnin years and plots, and then the total area values were averaged across years and plots for each treatment.", fig.pos="h",fig.align="center"}
+####
+####  EXTRACT COVER OF MOST DOMINANT SPECIES ----
+####
+cover_data <- {}
+for(dotrt in unique(cover_community_matrix$Treatment)){
+  tmp <- cover_community_matrix %>%
+    gather(species, area, -quad, -year, -Treatment) %>%
+    filter(Treatment == dotrt) %>%
+    group_by(Treatment, species) %>%
+    summarise(mean_area = mean(area)) %>%
+    arrange(-mean_area) %>%
+    mutate(mean_rank = seq_along(mean_area))
+
+  cover_data <- rbind(cover_data, tmp)
+}
+
+ggplot(cover_data, aes(x = mean_rank, y = mean_area))+
+  geom_line()+
+  geom_point(size = 2)+
+  geom_text(data=filter(cover_data, mean_rank < 11), aes(x = mean_rank, y = mean_area, label = species), 
+            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
+  scale_y_continuous(limits = c(0,0.25))+
+  xlab("Species rank")+
+  ylab(expression(paste("Mean basal area (",m^2,")")))+
+  facet_wrap(~Treatment, ncol = 3)
+
+cover_doms <- cover_data %>%
+  filter(mean_rank < 3) %>%
+  group_by(Treatment) %>%
+  summarise(total_area = sum(mean_area))
+
+```
+
+```{r annual_racs, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Rank abundance curves for annual species.
+Density of individuals is averaged across years and plots for each treatment."}
+annual_data <- {}
+for(dotrt in unique(annual_community_matrix$Treatment)){
+  tmp <- annual_community_matrix %>%
+    gather(species, density, -quad, -year, -Treatment) %>%
+    filter(Treatment == dotrt) %>%
+    group_by(Treatment, species) %>%
+    summarise(mean_density = mean(density)) %>%
+    arrange(-mean_density) %>%
+    mutate(mean_rank = seq_along(mean_density))
+
+  annual_data <- rbind(annual_data, tmp)
+}
+
+ggplot(annual_data, aes(x = mean_rank, y = mean_density))+
+  geom_line()+
+  geom_point(size = 2)+
+  geom_text(data=filter(annual_data, mean_rank < 3), aes(x = mean_rank, y = mean_density, label = species), 
+            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
+  scale_y_continuous(limits = c(0,200))+
+  xlab("Species rank")+
+  ylab("Mean density")+
+  facet_wrap(~Treatment, ncol = 3)
+```
+
+```{r time_perr_racs, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=6, fig.cap="Time series of rank abundance curves for perennial species in each treatment.
+Values of mean area were averaged over plots.
+The four most dominant species are labelled in each panel."}
+cover_data <- {}
+for(dotrt in unique(cover_community_matrix$Treatment)){
+  for(doyr in unique(cover_community_matrix$year)){
+    tmp <- cover_community_matrix %>%
+      gather(species, area, -quad, -year, -Treatment) %>%
+      filter(Treatment == dotrt, year == doyr) %>%
+      group_by(Treatment, year, species) %>%
+      summarise(mean_area = mean(area)) %>%
+      arrange(-mean_area) %>%
+      mutate(mean_rank = seq_along(mean_area))
+
+  cover_data <- rbind(cover_data, tmp)
+  }
+}
+ggplot(cover_data, aes(x = mean_rank, y = mean_area))+
+  geom_line()+
+  geom_point(size = 2)+
+  geom_text(data=filter(cover_data, mean_rank < 5), aes(x = mean_rank, y = mean_area, label = species), 
+            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
+  scale_y_continuous(limits = c(0,0.25))+
+  xlab("Species rank")+
+  ylab(expression(paste("Mean basal area (",m^2,")")))+
+  facet_grid(Treatment~year)
+```
+
+```{r time_annual_racs, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=6, fig.cap="Time series of rank abundance curves for annual species in each treatment.
+Values of density were averaged over plots.
+The two most dominant species are labelled in each panel."}
+annual_data <- {}
+for(dotrt in unique(annual_community_matrix$Treatment)){
+  for(doyr in unique(annual_community_matrix$year)){
+    tmp <- annual_community_matrix %>%
+      gather(species, density, -quad, -year, -Treatment) %>%
+      filter(Treatment == dotrt, year == doyr) %>%
+      group_by(Treatment, year, species) %>%
+      summarise(mean_density = mean(density)) %>%
+      arrange(-mean_density) %>%
+      mutate(mean_rank = seq_along(mean_density))
+
+    annual_data <- rbind(annual_data, tmp)
+  }
+}
+
+ggplot(annual_data, aes(x = mean_rank, y = mean_density))+
+  geom_line()+
+  geom_point(size = 2)+
+  geom_text(data=filter(annual_data, mean_rank < 3), aes(x = mean_rank, y = mean_density, label = species), 
+            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
+  scale_y_continuous(limits = c(0,400))+
+  scale_x_continuous(breaks = c(1,6,11))+
+  xlab("Species rank")+
+  ylab("Mean density")+
+  facet_grid(Treatment~year)
+```
+
+\newpage{}
+
+```{r annual_weather, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Relationship between precipitation and abundance of dominant annuals."}
+dom_annuals <- annual_data %>%
+  filter(species %in% c("Allyssum desertorum","Bromus tectorum")) %>%
+  select(-mean_rank) %>%
+  rename(Species = species) %>%
+  left_join(weather_data, by = "year")
+
+ggplot(dom_annuals, aes(x = ppt1, y = mean_density, color = Species))+
+  geom_point()+
+  xlab("Precipitation (mm)")+
+  ylab("Abundance")
+```
+
+```{r annual_nmds2, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Nonmetric multidimensional scaling scores representing annual plant communities in each plot, colored by treatment."}
+####
+####  ANALYZE JUST THE ANNUALS ----
+####
+# Set missing species abundances as 0, which == absent
+annual_community_matrix[is.na(annual_community_matrix)] <- 0
+
+annual_community_matrix_scaled <- cbind(annual_community_matrix[,1:3,drop=F],
+                                        scale(annual_community_matrix[,4:ncol(annual_community_matrix)]))
+
+spp_to_remove <- names(which(is.nan(colMeans(annual_community_matrix_scaled[4:ncol(annual_community_matrix_scaled)]))))
+ids_to_remove <- which(colnames(annual_community_matrix_scaled) %in% spp_to_remove)
+annual_community_matrix_scaled <- annual_community_matrix_scaled %>%
+  select(-ids_to_remove)
+
+nmds_df_annuals <- {}
+out_stats <- {}
+for(doyr in unique(annual_community_matrix_scaled$year)){
+  tmp <- filter(annual_community_matrix_scaled, year == doyr) %>% select(-year)
+   invisible( capture.output( # already checked for convergence
+    tmp_bray <- metaMDS(tmp[3:ncol(tmp)]+2, "bray", trymax = 999)
+  ))
+  nmds_points <- as.data.frame(tmp_bray$points)
+  nmds_points$Treatment <- tmp$Treatment
+  nmds_points$Year <- doyr
+  nmds_df_annuals <- rbind(nmds_df_annuals, nmds_points)
+  adonis_stats <- as.data.frame(vegan::adonis(tmp[3:ncol(tmp)]+2~tmp$Treatment)$aov.tab)[1,]
+  betadisper_stats <- anova(vegan::betadisper(vegan::vegdist(tmp[3:ncol(tmp)]+2),tmp$Treatment))[1,]
+  tmp_out <- rbind(adonis_stats[,-which(colnames(adonis_stats)=="R2")],
+                   as.numeric(betadisper_stats))
+  row.names(tmp_out) <- c()
+  out_stats <- rbind(out_stats, data.frame(year = doyr,
+                                           test = c("adonis","betadisper"),
+                                           n = nrow(tmp),
+                                           tmp_out))
+}
+
+mycols <- c("#009E73", "#0072B2", "#D55E00")
+ggplot(nmds_df_annuals, aes(x=MDS1, y=MDS2, color=Treatment))+
+  geom_point()+
+  geom_point(color="grey35",shape=1)+
+  scale_color_manual(values=mycols)+
+  ylab("NMDS 2")+
+  xlab("NMDS 1")+
+  facet_wrap(~Year)+
+  #scale_y_continuous(limits=c(-0.15,0.15))+
+  #scale_x_continuous(limits=c(-0.18,0.18))+
+  coord_fixed()
+```

### 2017-12-10 manuscript/Appendix2.Rmd
more tiny revisions and matching up with response letter

--- 2017-08-15 manuscript/Appendix2.Rmd
@@
   - \usepackage[sc]{mathpazo}
-bibliography: ~/Dropbox/Bibliography/usses_water.bib
-csl: components/ecology.csl
+  - \usepackage{mathtools}
+bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
+csl: components/ecology_letters.csl
 
@@
 
-\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
 \renewcommand\thefigure{A2-\arabic{figure}}  
-\renewcommand\theequation{A2.\arabic{equation}}
 \renewcommand\thetable{A2-\arabic{table}}  
@@
 \textbf{\Large{Appendix 2}} \\
-A.T. Tredennick, A.R. Kleinhesselink, B. Taylor \& P.B. Adler \\
+A.T. Tredennick, A.R. Kleinhesselink, J.B. Taylor \& P.B. Adler \\
 ``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe''  \\
@@
 
-\section{Random Slopes, Random Intercepts Model}
-\subsection{Model Description}
-Our hierarchical Bayesian model allows us to test for differences among treatments in the relationship between ANPP and soil moisture, and allowed us to account for the non-independence of observations through time within a plot.
-Treatment differences are modeled as fixed effects, which are modified by plot-level random effects.
-In what follows, $\textbf{X}$ is the fixed effects design matrix including:
+\linenumbers
 
-1. a column of 1s for intercepts
-2. a column of continuous values for the scaled volumetric water content for each observation
-3. a column of 0s or 1s indicating whether the observation is from a drought treatment 
-4. a column of 0s or 1s indicating whether the observation is from an irrigation treatment
-5. a column of continuous values for the interaction between scaled volumetric water content and the drought treatment indicator (column 2)
-6. a column of continuous values for the interaction between scaled volumetric water content and the irrigation treatment indicator (column 3)
+\section{Estimating ANPP}
+We used a radiometer to nondestructively estimate aboveground net primary productivity.
+Our approach relies on relating greenness in a plot to aboveground biomass.
+In each year we recorded ground reflectances at four bands, two associated with the red spectrum and two associated with the near-infrared spectrum (Table A2-1).
+We took four readings per plot that were averaged for each band.
+Bands 1 and 3 correspond to wavelengths collected by the MODIS satellite and bands 2 and 4 correspond to wavelengths collected by the AVHRR satellite.
 
-The first seven rows of the fixed effects design matrix (six control plots and one drought plot) is as follows:
+| Band number | Spectrum name | Center wavelength | Corresponding satellite |
+| ----------- | ------------- | ----------------- | ----------------------- |
+| 1 | red | 626 nm | AVHRR |
+| 2 | red | 652 nm | MODIS |
+| 3 | near-infrared | 875 nm | AVHRR |
+| 4 | near-infrared | 859 nm | MODIS |
 
-```{r fixed_design, echo=F, eval=T, message=F, warning=F}
-library(tidyverse)    # Data munging
-library(dplyr)        # Data summarizing
-library(ggthemes)     # Pleasing ggplot themes
-library(ggjoy)        # Joy plots!
-library(stringr)      # Working with strings
-library(rstan)        # For MCMC and Stan objects
-library(gridExtra)    # For combining ggplot objects
-library(viridis)      # Pleasing color palette
-library(RColorBrewer) # More color palettes
-library(cowplot)      # For combining ggplots
-source("../code/read_format_data.R") # load data
-lmod <- lm(log(anpp) ~ vwc_scaled*Treatment, anpp_data)
-x    <- model.matrix(lmod) 
-colnames(x) <- c("Int", "VWC", "Drought", "Irrig", "Drought:VWC", "Irrig:VWC")
-head(x, 7)
-```
+Table: Radiometer specifications.
 
-\noindent{}Note that within a year, all plots within a treatment share the same value of volumetric water content.
-This is because we could not monitor soil moisture in each plot, and instead used sparse observations to model average soil moisture in each treatment in each year (see main text).
-Using this design matrix, we can estimate six fixed effects ($\boldsymbol{\beta}$s):
-
-1. the intercept of the soil moisture-ANPP relationship for control plots
-2. the slope of the soil moisture-ANPP relationship for control plots
-3. the intercept offset for drought plots
-4. the intercept offset for irrigation plots
-5. the slope offset for drought plots
-6. the slope offset for irrigation plots
-
-\noindent{}We are particularly interested in the slope offsets, because these allow us to test whether the slopes for drought or irrigation are different from the control slope.
-If the slope offset is different from zero, this indicates the slopes are different.
-We assess whether the slope offsets are different from zero by calculating the probability that the posterior distribution is less or greater than zero (one-tailed tests).
-If the probability is greater than 0.95, then there is strong evidence that the slope offset is different from zero, and thus different from the control treatment slope.
-
-To account for the fact that observations within plots through time are not independent, we include random effects that modify the fixed effects in each plot.
-We model these random effects ($\boldsymbol{\gamma}$s) as offsets drawn from a multivariate normal distribution with mean 0 and a variance-covariance matrix ($\Sigma$) that includes a covariance between the intercept and slope offsets.
-We implement this random effects structure by including a random effects design matrix ($\textbf{Z}$) with a column of 1s for intercept offsets and a column of continuous values for the volumetric water content for each observation.
-
-Lastly, to account for unkown variation across years, we include random year effects. These year effects ($\boldsymbol{\eta}$s) act as offsets on the intercept.
-
-Putting it all together, our model is defined mathematically as follows, where *i* denotes observation, *j* denotes plot, and *t* denotes year.
-We assume the observations are conditionally Gaussian,
-
+Using the RED and NIR reflectance values, we calculate the normalized difference vegetation index (NDVI) for each plot based on both AVHRR- and MODIS-based wavelengths.
+We calculated NDVI as:
 \vspace{-2em}
-
 \begin{align}
-\textbf{y} &\sim \text{Normal} \left(\boldsymbol{\mu}, \sigma^2 \right),
+\text{NDVI}_{\text{AVHRR}} &= \frac{b_3 - b_1}{b_3 + b_1} \\[10pt]
+\text{NDVI}_{\text{MODIS}} &= \frac{b_4 - b_2}{b_4 + b_2}
 \end{align}
 
-\noindent{}where $\boldsymbol{\mu}$ is the determinstic expectation from the regression model,
+\noindent{} where $b_x$ refers to band *x* (*x* = 1,2,3,4) in Table A1-1.
 
-\vspace{-2em}
+To convert plot NDVI to biomass, we regressed known biomass values from calibration plots against NDVI calculate for those plots.
+Calibration plots were located near our experiment plots, and each year we located a new set of 10 plots in which we clipped all aboveground biomass, dried it to a constant weight at 60$^\circ$ C, and the weighed.
+We used these biomass values to estimate regression parameters for both AVHRR- and MODIS-based NDVI.
+We assessed model fit using $R^2$ and, for each year, we used the regression parameters associated with the best fit model to estimate biomass in the experimental plots based on their NDVI values (Table A2-2).
+R code for this procedure is in the file "01_calibrate_radiometer_by_year.R" in the code set.
 
-\begin{align}
-\mu_{i} &= \boldsymbol{\beta}\textbf{x}_i + \boldsymbol{\gamma}_{j(i)}\textbf{z}_i + \eta_t.
-\end{align}
+```{r ndvi_table, include=T, echo=F, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
+library(xtable)
+library(tidyverse)
+library(dplyr)
+tab_dat <- read.csv("../data/estimated_biomass/ndvi_biomass_regression_parameters.csv", row.names = 1) %>%
+  filter(year > 2011 & year < 2017)
+colnames(tab_dat) <- c("Year", "Intercept", "NDVI Slope", "$R^2$", "Min(biomass)", "Max(biomass)", "\\emph{N}", "Algorithm")
 
-\noindent{}All fixed effect $\beta$s were drawn from normally distributed priors with mean 0 and standard deviation 1: $\boldsymbol{\beta} \sim \text{Normal}(0,1)$.
-$\gamma$ random effects were drawn from a multivariate prior centered on zero with a shared variance covariance matrix:
-
-\vspace{-2em}
-
-\begin{align}
-\boldsymbol{\gamma} \sim \text{MVN} \left( 0, \Sigma \right), \\
-\Sigma \sim \text{Cholesky...}
-\end{align}
-
-\noindent{The random year effects ($\boldsymbol{\eta}$) are modeled as intercept offsets centered on zero with a shared variance ($\sigma_{yr}$): $\boldsymbol{\gamma} \sim \text{Normal} \left( 0, \sigma_{yr}  \right)$.
-
-We fit the model using MCMC as implemented in the software Stan [@stan2016].
-Our Stan code is below.
-All code necessary to reproduce our results has been archived on Figshare (*link here*) and released on GitHub (https://github.com/atredennick/usses_water/releases/v0.1).
-
-\newpage{}
-
-\subsection{Stan Code}
-```{r stan1, echo=T, eval=F}
-data {
-  int<lower=0> Npreds;         # number of covariates, including intercept
-  int<lower=0> Npreds2;        # number of random effect covariates
-  int<lower=0> Nplots;         # number of plots
-  int<lower=0> Ntreats;        # number of treatments
-  int<lower=0> Nobs;           # number of observations
-  int<lower=0> Nppts;          # number of precip levels to predict
-  int<lower=0> Nyears;         # number of years
-  vector[Nobs] y;              # vector of observations
-  row_vector[Npreds] x[Nobs];	 # design matrix for fixed effects
-  row_vector[Npreds2] z[Nobs]; # simple design matrix for random effects 
-  matrix[Nppts,Npreds] newx;   # design matrix for predictions 
-  matrix[Npreds2,Npreds2] R;	 # priors for covariance matrix
-  int plot_id[Nobs];           # vector of plot ids
-  int treat_id[Nobs];          # vector of treatment ids
-  int year_id[Nobs];           # vector of year ids
-}
-
-parameters {
-	vector[Npreds] beta;	             # overall coefficients
-	vector[Nyears] year_off;           # vector of year effects
-	cholesky_factor_corr[Npreds2] L_u; # cholesky factor of plot random effect corr matrix
-  vector[Npreds2] beta_plot[Nplots]; # plot level random effects
-	vector<lower=0>[Npreds2] sigma_u;  # plot random effect std. dev.
-	real<lower=0> sd_y;                # treatment-level observation std. dev.
-	real<lower=0> sigma_year;          # year std. dev. hyperprior
-}
-
-transformed parameters {
-  vector[Nobs] yhat;               # vector of expected values
-  vector[Npreds2] u[Nplots];       # transformed plot random effects
-  matrix[Npreds2,Npreds2] Sigma_u; # plot ranef cov matrix
-
-  Sigma_u = diag_pre_multiply(sigma_u, L_u); # plot-level covariance matrix 
-  for(j in 1:Nplots)
-    u[j] = Sigma_u * beta_plot[j]; # plot random intercepts and slopes
-
-  # regression model for expected values (one for each plot-year)
-  for (i in 1:Nobs)
-    yhat[i] = x[i]*beta + z[i]*u[plot_id[i]] + year_off[year_id[i]]; 
-}
-
-model {
-  ####  PRIORS
-  sigma_year ~ cauchy(0,2)
-  year_off ~ normal(0,sigma_year); # priors on year effects, shared variance
-  beta ~ normal(0,1);	             # priors on treatment coefficients
-  L_u ~ lkj_corr_cholesky(1);      # prior on the cholesky factor which controls the 
-                                   # correlation between plot level treatment effects
-
-  for(i in 1:Nplots)
-		beta_plot[i] ~ normal(0,1); # plot-level coefficients for intercept and slope
-
-	####  LIKELIHOOD
-  for(i in 1:Nobs)
-    y[i] ~ normal(yhat[i], sd_y); # observations vary normally around expected values
-}
+my_caption <- "Details of regression models used to estimate biomass each year."
+print(xtable(tab_dat, caption = my_caption, digits = 2),
+      caption.placement="top",
+      include.rownames = F,
+      sanitize.colnames.function = identity,
+      comment = FALSE,
+      size = "normalsize")
 ```
-
-\newpage{}
-
-# Figures
-
-```{r scatterplot, fig.cap="Scatterplot, on the arithmetic scale, of the data and model estimates shown as solid lines.
-Model estimates come from treatment level coefficients (colored lines).", echo=FALSE, message=FALSE}
-library(tidyverse)    # Data munging
-library(dplyr)        # Data summarizing
-library(ggthemes)     # Pleasing ggplot themes
-library(ggjoy)        # Joy plots!
-library(stringr)      # Working with strings
-library(rstan)        # For MCMC and Stan objects
-library(gridExtra)    # For combining ggplot objects
-library(viridis)      # Pleasing color palette
-library(RColorBrewer) # More color palettes
-library(cowplot)      # For combining ggplots
-
-source("../code/read_format_data.R") # load data
-all_fit <- readRDS("../results/randcoefs_alltreatments_fit.RDS")
-
-param_labels <- c("Control::Intercept",
-                  "Drought::Intercept", 
-                  "Irrigation::Intercept",
-                  "Control::Slope", 
-                  "Drought::Slope", 
-                  "Irrigation::Slope")
-
-treat_effects <- data.frame(extract(all_fit, pars = 'beta')) %>% 
-  mutate(iteration = row_number()) %>% 
-  gather(param_name, estimate, starts_with('beta')) %>%
-  mutate(param_name = factor(param_name, labels = param_labels)) %>% 
-  spread(param_name, estimate) %>% 
-  mutate(`Drought::Intercept` = `Drought::Intercept` + `Control::Intercept`, 
-         `Irrigation::Intercept` = `Irrigation::Intercept` + `Control::Intercept`,  
-         `Drought::Slope` = `Drought::Slope` + `Control::Slope`, 
-         `Irrigation::Slope` = `Irrigation::Slope` + `Control::Slope`) %>%
-  gather(param_name, estimate, -iteration)
-
-mean_effects <- treat_effects %>%
-  group_by(param_name) %>%
-  summarise(mean_estimate = mean(estimate))
-
-regression <- mean_effects %>%
-  separate(param_name, into = c("Treatment","Coefficient"), sep = "::") %>%
-  spread(Coefficient, mean_estimate)
-
-lmod <- lm(log(anpp) ~ vwc_scaled, anpp_data)
-x    <- model.matrix(lmod) 
-newx <- data.frame(x = seq(min(x[,2]), max(x[,2]), length.out = 50)) %>%
-  mutate(id = paste("x",1:50,sep = "::")) %>%
-  spread(id, x)
-newx <- rbind(rbind(newx,newx),newx)
-newx$Treatment <- regression$Treatment
-
-regression <- regression %>%
-  left_join(newx, by="Treatment") %>%
-  gather(xid, newx, -Treatment, -Intercept, -Slope)
-regression$estimate <- with(regression, Intercept+Slope*newx)
-
-mean_log_anpp <- mean(log(anpp_data$anpp))
-sd_log_anpp <- sd(log(anpp_data$anpp))
-mean_vwc <- mean(anpp_data$total_seasonal_vwc)
-sd_vwc <- sd(anpp_data$total_seasonal_vwc)
-
-regression <- regression %>%
-  mutate(backtrans_estimate = estimate*sd_log_anpp+mean_log_anpp,
-         backtrans_vwc = newx*sd_vwc+mean_vwc)
-
-ggplot(anpp_data, aes(x=total_seasonal_vwc,y=anpp))+
-  geom_point(shape=21,color="grey25",alpha=0.8,aes(fill=Treatment))+
-  geom_line(data=regression, aes(x=backtrans_vwc, y=exp(backtrans_estimate), color=Treatment), size=1)+
-  scale_fill_brewer(palette = "Set2")+
-  scale_color_brewer(palette = "Set2")+
-  xlab("March-June Cumulative VWC")+
-  ylab(expression(paste("ANPP (g ",m^2,")")))+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())
-
-```
-
-\newpage{}
-
-```{r year_effects, fig.cap="Posterior distributions of random year effects (intercept offsets).
-Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.", echo=FALSE, message=FALSE, warning=FALSE}
-library(tidyverse)    # Data munging
-library(dplyr)        # Data summarizing
-library(ggthemes)     # Pleasing ggplot themes
-library(ggjoy)        # Joy plots!
-library(stringr)      # Working with strings
-library(rstan)        # For MCMC and Stan objects
-library(gridExtra)    # For combining ggplot objects
-library(viridis)      # Pleasing color palette
-library(RColorBrewer) # More color palettes
-library(cowplot)      # For combining ggplots
-
-all_fit <- readRDS("../results/randcoefs_alltreatments_fit.RDS")
-
-year_id_names <- data.frame(year_id = c(1:5),
-                            year_name = c(1:5))
-betas <- reshape2::melt(rstan::extract(all_fit, pars="year_off")) %>%
-  rename(iteration = iterations, year_id = Var2, estimate = value, stan_name = L1) %>%
-  left_join(year_id_names, by="year_id")
-
-ggplot(betas, aes(x=estimate, y=as.factor(year_id), fill=as.factor(year_id), color=as.factor(year_id), height = ..density..))+
-  geom_vline(aes(xintercept=0), linetype=2, color="grey45")+
-  geom_joy(stat="density", adjust=3, alpha=0.9)+
-  scale_x_continuous(breaks=seq(-3,3,0.5), limits=c(-3,3))+
-  scale_y_discrete(labels = c("2012","2013","2014","2015","2016"))+
-  scale_fill_viridis(end=0.8,discrete=T)+
-  scale_color_viridis(end=0.8,discrete=T)+
-  guides(fill=FALSE, color=FALSE)+
-  xlab("Parameter Value")+
-  ylab("Year")+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank(),
-        panel.border     = element_blank(),
-        axis.line        = element_blank(),
-        axis.ticks       = element_blank())
-
-```
-
-\newpage{}
-
-# References

### 2017-12-10 manuscript/Appendix3.Rmd
more tiny revisions and matching up with response letter

--- 2017-06-12 manuscript/Appendix3.Rmd
@@
 \textbf{\Large{Appendix 3}} \\
-A.T. Tredennick, A.R. Kleinhesselink, \& P.B. Adler \\
-``Ecosystem and community resistance...'' \\
+A.T. Tredennick, A.R. Kleinhesselink, J.B. Taylor \& P.B. Adler \\
+``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe'' \\
 \emph{PeerJ}
@@
 
-\section{Characterizing Extreme Precipitation Amounts}
+\section{Random Slopes, Random Intercepts Model}
+\subsection{Model Description}
+Our hierarchical Bayesian model allows us to test for differences among treatments in the relationship between ANPP and soil moisture, and allowed us to account for the non-independence of observations through time within a plot.
+Treatment differences are modeled as fixed effects, which are modified by plot-level random effects.
+In what follows, $\textbf{X}$ is the fixed effects design matrix including:
 
-Following the proposed methods of @Lemoine2016, we calculated quantiles from the empirical distribution of growing season precipitation at Dubios, ID.
-We chose the 1\% quantile to be indicative of extreme dry conditions (drought) and the 99\% quantile to be indicative of extreme wet conditions (irrigation).
-The data consist of 91 yearly records, which we assume are approximately normally distributed.
-The R code below shows our procedure, and Fig. A3-1 shows the results.
+1. a column of 1s for intercepts
+2. a column of continuous values for the scaled volumetric water content for each observation
+3. a column of 0s or 1s indicating whether the observation is from a drought treatment 
+4. a column of 0s or 1s indicating whether the observation is from an irrigation treatment
+5. a column of continuous values for the interaction between scaled volumetric water content and the drought treatment indicator (column 2)
+6. a column of continuous values for the interaction between scaled volumetric water content and the irrigation treatment indicator (column 3)
 
-```{r hrv, fig.cap="Density of the empirical distribution of growing season precipitation at Dubois, ID. Dashed vertical lines show the 1% and 99% quantiles, assuming a normal distribution."}
-library(ggplot2)
-weather    <- read.csv("../data/weather/ClimateIPM.csv")
-mean_ppt   <- mean(weather$ppt1)
-quants_ppt <- quantile(weather$ppt1,probs = c(0.01,0.99))
-quants_ppt[1]/mean_ppt*100 # percent of mean ppt for drought
-quants_ppt[2]/mean_ppt*100 # percent of mean ppt for irrigation
+The first seven rows of the fixed effects design matrix (six control plots and one drought plot) is as follows:
 
-ggplot(weather, aes(x=ppt1))+
-  geom_histogram(bins=20, color="dodgerblue", fill="dodgerblue", aes(y=..density..))+
-  geom_line(stat="density", color="blue")+
-  geom_vline(aes(xintercept=quants_ppt[1]), linetype=2)+
-  geom_vline(aes(xintercept=quants_ppt[2]), linetype=2)+
-  ylab("Density")+
-  xlab("Growing Season Precipitation (mm)")+
+```{r fixed_design, echo=F, eval=T, message=F, warning=F}
+library(tidyverse)    # Data munging
+library(dplyr)        # Data summarizing
+library(ggthemes)     # Pleasing ggplot themes
+library(ggjoy)        # Joy plots!
+library(stringr)      # Working with strings
+library(rstan)        # For MCMC and Stan objects
+library(gridExtra)    # For combining ggplot objects
+library(viridis)      # Pleasing color palette
+library(RColorBrewer) # More color palettes
+library(cowplot)      # For combining ggplots
+source("../code/read_format_data.R") # load data
+lmod <- lm(log(anpp) ~ vwc_scaled*Treatment, anpp_data)
+x    <- model.matrix(lmod) 
+colnames(x) <- c("Int", "VWC", "Drought", "Irrig", "Drought:VWC", "Irrig:VWC")
+head(x, 7)
+```
+
+\noindent{}Note that within a year, all plots within a treatment share the same value of volumetric water content.
+This is because we could not monitor soil moisture in each plot, and instead used sparse observations to model average soil moisture in each treatment in each year (see main text).
+Using this design matrix, we can estimate six fixed effects ($\boldsymbol{\beta}$s):
+
+1. the intercept of the soil moisture-ANPP relationship for control plots
+2. the slope of the soil moisture-ANPP relationship for control plots
+3. the intercept offset for drought plots
+4. the intercept offset for irrigation plots
+5. the slope offset for drought plots
+6. the slope offset for irrigation plots
+
+\noindent{}We are particularly interested in the slope offsets, because these allow us to test whether the slopes for drought or irrigation are different from the control slope.
+If the slope offset is different from zero, this indicates the slopes are different.
+We assess whether the slope offsets are different from zero by calculating the probability that the posterior distribution is less or greater than zero (one-tailed tests).
+If the probability is greater than 0.95, then there is strong evidence that the slope offset is different from zero, and thus different from the control treatment slope.
+
+To account for the fact that observations within plots through time are not independent, we include random effects that modify the fixed effects in each plot.
+We model these random effects ($\boldsymbol{\gamma}$s) as offsets drawn from a multivariate normal distribution with mean 0 and a variance-covariance matrix ($\Sigma$) that includes a covariance between the intercept and slope offsets.
+We implement this random effects structure by including a random effects design matrix ($\textbf{Z}$) with a column of 1s for intercept offsets and a column of continuous values for the volumetric water content for each observation.
+
+Lastly, to account for unkown variation across years, we include random year effects. These year effects ($\boldsymbol{\eta}$s) act as offsets on the intercept.
+
+Putting it all together, our model is defined mathematically as follows, where *i* denotes observation, *j* denotes plot, and *t* denotes year.
+We assume the observations are conditionally Gaussian,
+
+\vspace{-2em}
+
+\begin{align}
+\textbf{y} &\sim \text{Normal} \left(\boldsymbol{\mu}, \sigma^2 \right),
+\end{align}
+
+\noindent{}where $\boldsymbol{\mu}$ is the determinstic expectation from the regression model,
+
+\vspace{-2em}
+
+\begin{align}
+\mu_{i} &= \boldsymbol{\beta}\textbf{x}_i + \boldsymbol{\gamma}_{j(i)}\textbf{z}_i + \eta_t.
+\end{align}
+
+\noindent{}All fixed effect $\beta$s were drawn from normally distributed priors with mean 0 and standard deviation 5: $\boldsymbol{\beta} \sim \text{Normal}(0,5)$.
+$\gamma$ random effects were drawn from a multivariate prior centered on zero with a shared variance covariance matrix:
+
+\vspace{-2em}
+
+\begin{align}
+\boldsymbol{\gamma} \sim \text{MVN} \left( 0, \Sigma \right), \\
+\Sigma = \sigma_{plt}^2 \textbf{R}, \\
+\sigma_{plt} \sim \text{Cauchy}(0,2.5), \\
+\textbf{R} \sim \text{LKJ}(2.0)
+\end{align}
+
+\noindent{}The random year effects ($\boldsymbol{\eta}$) are modeled as intercept offsets centered on zero with a shared variance ($\sigma_{yr}$): $\boldsymbol{\gamma} \sim \text{Normal} \left( 0, \sigma_{yr}  \right)$.
+
+The Bayesian posterior distribution of our model can be expressed as:
+
+\vspace{-2em}
+
+\begin{align}
+\left[ \boldsymbol{\beta}, \boldsymbol{\gamma}, \boldsymbol{\eta}, \sigma_{yr}, \sigma_{plt}, \textbf{R}, \sigma \right | \textbf{y}] &\propto \left( \prod_{i=1}^{n} \left[ y_i | \boldsymbol{\beta}, \boldsymbol{\gamma}, \boldsymbol{\eta}, \sigma \right] \right) \left( \prod_{j=1}^{J} \left[ \gamma_j | \sigma_{plt}, \textbf{R} \right] \right) \\
+&\times \left( \prod_{t=1}^{T} \left[ \eta_t | \sigma_{yr} \right] \right)  [\boldsymbol{\beta}] [\sigma_{plt}] [\textbf{R}] [\sigma_{yr}] [\sigma]
+\end{align}
+
+We fit the model using MCMC as implemented in the software Stan [@stan2016].
+Our Stan code is below.
+All code necessary to reproduce our results has been archived on Figshare (*link here*) and released on GitHub (https://github.com/atredennick/usses_water/releases/v0.1).
+
+\newpage{}
+
+\subsection{Stan Code}
+```{r stan1, echo=T, eval=F}
+data {
+  int<lower=0> Npreds;         # number of covariates, including intercept
+  int<lower=0> Npreds2;        # number of random effect covariates
+  int<lower=0> Nplots;         # number of plots
+  int<lower=0> Ntreats;        # number of treatments
+  int<lower=0> Nobs;           # number of observations
+  int<lower=0> Nyears;         # number of years
+  vector[Nobs] y;              # vector of observations
+  row_vector[Npreds] x[Nobs];	 # design matrix for fixed effects
+  row_vector[Npreds2] z[Nobs]; # simple design matrix for random effects 
+  int plot_id[Nobs];           # vector of plot ids
+  int treat_id[Nobs];          # vector of treatment ids
+  int year_id[Nobs];           # vector of year ids
+}
+
+parameters {
+	vector[Npreds] beta;	             # overall coefficients
+	vector[Nyears] year_off;           # vector of year effects
+	cholesky_factor_corr[Npreds2] L_u; # cholesky factor of plot random effect corr matrix
+  vector[Npreds2] beta_plot[Nplots]; # plot level random effects
+	vector<lower=0>[Npreds2] sigma_u;  # plot random effect std. dev.
+	real<lower=0> sd_y;                # treatment-level observation std. dev.
+	real<lower=0> sigma_year;          # year std. dev. hyperprior
+}
+
+transformed parameters {
+  vector[Nobs] yhat;               # vector of expected values
+  vector[Npreds2] u[Nplots];       # transformed plot random effects
+  matrix[Npreds2,Npreds2] Sigma_u; # plot ranef cov matrix
+
+  Sigma_u = diag_pre_multiply(sigma_u, L_u); # plot-level covariance matrix 
+  for(j in 1:Nplots)
+    u[j] = Sigma_u * beta_plot[j]; # plot random intercepts and slopes
+
+  # regression model for expected values (one for each plot-year)
+  for (i in 1:Nobs)
+    yhat[i] = x[i]*beta + z[i]*u[plot_id[i]] + year_off[year_id[i]]; 
+}
+
+model {
+  ####  PRIORS
+  sigma_u ~ cauchy(0,2.5)
+  sigma_year ~ cauchy(0,2.5)
+  year_off ~ normal(0,sigma_year); # priors on year effects, shared variance
+  beta ~ normal(0,5);	             # priors on treatment coefficients
+  L_u ~ lkj_corr_cholesky(2.0);      # prior on the cholesky factor which controls the 
+                                   # correlation between plot level treatment effects
+
+  for(i in 1:Nplots)
+		beta_plot[i] ~ normal(0,1); # plot-level coefficients for intercept and slope
+
+	####  LIKELIHOOD
+  for(i in 1:Nobs)
+    y[i] ~ normal(yhat[i], sd_y); # observations vary normally around expected values
+}
+
+generated quantities{
+  corr_matrix[Npreds2] R = multiply_lower_tri_self_transpose(L_u);
+  cov_matrix[Npreds2] V = quad_form_diag(R,sigma_u);
+}
+```
+
+\newpage{}
+
+# Figures
+
+```{r scatterplot, fig.cap="Scatterplot, on the arithmetic scale, of the data and model estimates shown as solid lines.
+Model estimates come from treatment level coefficients (colored lines).", echo=FALSE, message=FALSE, warning=FALSE}
+library(tidyverse)    # Data munging
+library(dplyr)        # Data summarizing
+library(ggthemes)     # Pleasing ggplot themes
+library(ggjoy)        # Joy plots!
+library(stringr)      # Working with strings
+library(rstan)        # For MCMC and Stan objects
+library(gridExtra)    # For combining ggplot objects
+library(viridis)      # Pleasing color palette
+library(RColorBrewer) # More color palettes
+library(cowplot)      # For combining ggplots
+
+source("../code/read_format_data.R") # load data
+all_fit <- readRDS("../results/randcoefs_alltreatments_fit.RDS")
+
+param_labels <- c("Control::Intercept",
+                  "Drought::Intercept", 
+                  "Irrigation::Intercept",
+                  "Control::Slope", 
+                  "Drought::Slope", 
+                  "Irrigation::Slope")
+
+treat_effects <- data.frame(extract(all_fit, pars = 'beta')) %>% 
+  mutate(iteration = row_number()) %>% 
+  gather(param_name, estimate, starts_with('beta')) %>%
+  mutate(param_name = factor(param_name, labels = param_labels)) %>% 
+  spread(param_name, estimate) %>% 
+  mutate(`Drought::Intercept` = `Drought::Intercept` + `Control::Intercept`, 
+         `Irrigation::Intercept` = `Irrigation::Intercept` + `Control::Intercept`,  
+         `Drought::Slope` = `Drought::Slope` + `Control::Slope`, 
+         `Irrigation::Slope` = `Irrigation::Slope` + `Control::Slope`) %>%
+  gather(param_name, estimate, -iteration)
+
+mean_effects <- treat_effects %>%
+  group_by(param_name) %>%
+  summarise(mean_estimate = mean(estimate))
+
+regression <- mean_effects %>%
+  separate(param_name, into = c("Treatment","Coefficient"), sep = "::") %>%
+  spread(Coefficient, mean_estimate)
+
+lmod <- lm(log(anpp) ~ vwc_scaled, anpp_data)
+x    <- model.matrix(lmod) 
+newx <- data.frame(x = seq(min(x[,2]), max(x[,2]), length.out = 50)) %>%
+  mutate(id = paste("x",1:50,sep = "::")) %>%
+  spread(id, x)
+newx <- rbind(rbind(newx,newx),newx)
+newx$Treatment <- regression$Treatment
+
+regression <- regression %>%
+  left_join(newx, by="Treatment") %>%
+  gather(xid, newx, -Treatment, -Intercept, -Slope)
+regression$estimate <- with(regression, Intercept+Slope*newx)
+
+mean_log_anpp <- mean(log(anpp_data$anpp))
+sd_log_anpp <- sd(log(anpp_data$anpp))
+mean_vwc <- mean(anpp_data$total_seasonal_vwc)
+sd_vwc <- sd(anpp_data$total_seasonal_vwc)
+
+regression <- regression %>%
+  mutate(backtrans_estimate = estimate*sd_log_anpp+mean_log_anpp,
+         backtrans_vwc = newx*sd_vwc+mean_vwc)
+
+regression_limited <- {}
+for(dotrt in unique(regression$Treatment)){
+  tmpdat <- filter(regression, Treatment == dotrt)
+  tmprange <- range(anpp_data[which(anpp_data$Treatment==dotrt),"total_seasonal_vwc"])
+  tmpdat$backtrans_vwc[tmpdat$backtrans_vwc < tmprange[1] | tmpdat$backtrans_vwc > tmprange[2]] <- NA
+  regression_limited <- rbind(regression_limited, tmpdat)
+}
+
+ggplot(anpp_data, aes(x=total_seasonal_vwc,y=anpp))+
+  geom_point(shape=21,color="grey25",alpha=0.8,aes(fill=Treatment))+
+  geom_line(data=regression_limited, aes(x=backtrans_vwc, y=exp(backtrans_estimate), color=Treatment), size=1)+
+  scale_fill_brewer(palette = "Set2")+
+  scale_color_brewer(palette = "Set2")+
+  xlab("March-June Cumulative VWC")+
+  ylab(expression(paste("ANPP (g ",m^2,")")))+
   theme_bw()+
   theme(panel.grid.minor = element_blank())
+
 ```
 
+\newpage{}
+
+```{r posteriors, fig.cap="Prior (dashed lines) and posterior (solid line) distributions of intercepts and slopes for each treatment.
+The slope represents the main effect of soil moisture on log(ANPP).
+The red line marks 0.
+Shrinkage of the posterior distribution and/or changes in the mean indicate the data informed model parameters beyond the information contained in the prior for all coefficients.",echo=FALSE, message=FALSE, warning=FALSE}
+posteriors <- treat_effects %>%
+  separate(param_name, into = c("Treatment", "Coefficient")) %>%
+  mutate(prior = rnorm(n = nrow(treat_effects), 0, 5)) %>%
+  gather(Distribution, Value, -iteration, -Treatment, -Coefficient)
+
+ggplot(posteriors, aes(x=Value, linetype=Distribution))+
+    geom_line(stat = "density")+
+    geom_vline(aes(xintercept=0), color="red", size=0.2)+
+    scale_x_continuous(limits=c(-5,5))+
+    scale_linetype_discrete(labels = c("Posterior","Prior"), name = NULL)+
+    facet_grid(Coefficient~Treatment)+
+    ylab("Density")+
+    theme_few()+
+    theme(axis.text = element_text(size=7))
+```
+
+\newpage{}
+
+```{r year_effects, fig.cap="Posterior distributions of random year effects (intercept offsets).
+Kernel bandwidths of posterior densities were adjusted by a factor of 5 to smooth the density curves for visual clarity.", echo=FALSE, message=FALSE, warning=FALSE}
+library(tidyverse)    # Data munging
+library(dplyr)        # Data summarizing
+library(ggthemes)     # Pleasing ggplot themes
+library(ggjoy)        # Joy plots!
+library(stringr)      # Working with strings
+library(rstan)        # For MCMC and Stan objects
+library(gridExtra)    # For combining ggplot objects
+library(viridis)      # Pleasing color palette
+library(RColorBrewer) # More color palettes
+library(cowplot)      # For combining ggplots
+
+all_fit <- readRDS("../results/randcoefs_alltreatments_fit.RDS")
+
+year_id_names <- data.frame(year_id = c(1:5),
+                            year_name = c(1:5))
+betas <- reshape2::melt(rstan::extract(all_fit, pars="year_off")) %>%
+  rename(iteration = iterations, year_id = Var2, estimate = value, stan_name = L1) %>%
+  left_join(year_id_names, by="year_id")
+
+ggplot(betas, aes(x=estimate, y=as.factor(year_id), fill=as.factor(year_id), color=as.factor(year_id), height = ..density..))+
+  geom_vline(aes(xintercept=0), linetype=2, color="grey45")+
+  geom_joy(stat="density", adjust=3, alpha=0.9)+
+  scale_x_continuous(breaks=seq(-3,3,0.5), limits=c(-3,3))+
+  scale_y_discrete(labels = c("2012","2013","2014","2015","2016"))+
+  scale_fill_viridis(end=0.8,discrete=T)+
+  scale_color_viridis(end=0.8,discrete=T)+
+  guides(fill=FALSE, color=FALSE)+
+  xlab("Parameter Value")+
+  ylab("Year")+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank(),
+        panel.border     = element_blank(),
+        axis.line        = element_blank(),
+        axis.ticks       = element_blank())
+
+```
+
+\newpage{}
+
 # References

### 2017-12-10 manuscript/Appendix4.Rmd
more tiny revisions and matching up with response letter

@@
 
-\section{Characterizing Extreme Precipitation Amounts}
+\section{Conducting our analysis with NDVI}
 
-Following the proposed methods of @Lemoine2016, we calculated quantiles from the empirical distribution of growing season precipitation at Dubios, ID.
-We chose the 1\% quantile to be indicative of extreme dry conditions (drought) and the 99\% quantile to be indicative of extreme wet conditions (irrigation).
-The data consist of 91 yearly records, which we assume are approximately normally distributed.
-The R code below shows our procedure, and Fig. A3-1 shows the results.
+The first step of our analysis is to convert NDVI from field-based radiometric measurements to aboveground net primary productivity.
+However, this conversion is not perfect and the associated uncertainty is sometimes high (Table A1-2).
+Therefore, we also assessed ecosystem functional response across treatments using NDVI as the response variable rather than ANPP.
+The model essentially the same as the ANPP model, except we use a Beta likelihood for NDVI since its values range from 0 to 1, not including true 0s or 1s (see Stan code, below).
 
-```{r hrv, fig.cap="Density of the empirical distribution of growing season precipitation at Dubois, ID. Dashed vertical lines show the 1% and 99% quantiles, assuming a normal distribution."}
-library(ggplot2)
-weather    <- read.csv("../data/weather/ClimateIPM.csv")
-mean_ppt   <- mean(weather$ppt1)
-quants_ppt <- quantile(weather$ppt1,probs = c(0.01,0.99))
-quants_ppt[1]/mean_ppt*100 # percent of mean ppt for drought
-quants_ppt[2]/mean_ppt*100 # percent of mean ppt for irrigation
+Our NDVI-model results are very similar to the ANPP-model results (Figure A4-1).
+The only difference is there is a higher probability that the drought slope offset is less than 1 (90\% probability).
+However, these negative offsets do not result in significantly different slopes once applied to the control slope, just as in the ANPP-model.
 
-ggplot(weather, aes(x=ppt1))+
-  geom_histogram(bins=20, color="dodgerblue", fill="dodgerblue", aes(y=..density..))+
-  geom_line(stat="density", color="blue")+
-  geom_vline(aes(xintercept=quants_ppt[1]), linetype=2)+
-  geom_vline(aes(xintercept=quants_ppt[2]), linetype=2)+
-  ylab("Density")+
-  xlab("Growing Season Precipitation (mm)")+
+```{r plot_betas, fig.cap="Posterior distributions for the effects of drought and irrigation on the intercept and slope of the NDVI-soil moisture relationship.
+Treatment effects show the difference between the coefficients estimated in the treated plots and the control plots.
+Probabilities (Pr </> 0 =) for each coefficient indicate the one-tailed probability that the coefficient is less than or greater than zero, depending on whether the median of the distribution is less than or greater than zero.
+The posterior densities were smoothed for visual clarity by increasing kernel bandwidth by a factor of five.", include=T, echo=F, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
+
+library(tidyverse)    # Data munging
+library(dplyr)        # Data summarizing
+library(ggthemes)     # Pleasing ggplot themes
+library(ggjoy)        # Joy plots!
+library(stringr)      # Working with strings
+library(rstan)        # For MCMC and Stan objects
+library(gridExtra)    # For combining ggplot objects
+library(viridis)      # Pleasing color palette
+library(RColorBrewer) # More color palettes
+library(cowplot)      # For combining ggplots
+
+all_fit <- readRDS("../results/randcoefs_alltreatments_NDVIfit.RDS")
+
+####
+####  FUNCTION TO CALCULATE MAX(Pr(effect) > 0, Pr(effect) < 0) ----
+####
+get_one_tailed <- function(values){
+  above <- 1 - ecdf(values)(0)
+  below <- ecdf(values)(0)
+  max(above,below)
+}
+
+####
+####  PLOT TREATMENT-LEVEL POSTERIOR DISTRIBUTIONS ----
+####
+mycols <- c("#009E73", "#D55E00", "#0072B2")
+param_labels <- c("Control::Intercept",
+                  "Drought::Intercept", 
+                  "Irrigation::Intercept",
+                  "Control::Slope", 
+                  "Drought::Slope", 
+                  "Irrigation::Slope")
+
+betas <- data.frame(extract(all_fit, pars = 'beta')) %>% 
+  mutate(iteration = row_number()) %>% 
+  gather(param_name, estimate, starts_with('beta')) %>%
+  mutate(param_name = factor(param_name, labels = param_labels)) %>%
+  separate(param_name, c("Treatment", "Type"), "::")
+
+treat_slopes <- betas %>%
+  filter(Treatment != "Control") %>%
+  mutate(Type = paste(Type,"Offset"))
+
+slope_probs <- treat_slopes %>%
+  group_by(Treatment, Type) %>%
+  summarise(probs = round(get_one_tailed(estimate),2)) %>%
+  mutate(prob_text = paste("Pr < 0 =", probs))
+slope_probs$prob_text[1] <- paste("Pr > 0 =", slope_probs$probs[1]) # switch direction of drought intercpet since greater than 0
+slope_probs$xpos <- c(-2.8, -0.6, -2.6, -0.6)
+slope_probs$ypos <- c(1.3,1.3,2.55,2.7)
+
+treat_cols <- mycols[2:3]
+
+ggplot(treat_slopes)+
+  geom_vline(aes(xintercept=0), linetype=2, color="grey45")+
+  geom_joy(stat="density", adjust=3, alpha=0.8, aes(x=estimate, y=Treatment, fill=Treatment, color=Treatment, height = ..density..))+
+  geom_text(data = slope_probs, aes(x=xpos, y=ypos, label = prob_text), size = 3)+
+  facet_wrap(~Type, scales = "free")+
+  scale_y_discrete(labels = c("Drought","Irrigation"))+
+  scale_fill_manual(values = treat_cols)+
+  scale_color_manual(values = treat_cols)+
+  guides(fill=FALSE, color=FALSE)+
+  xlab("Coefficient Value")+
+  ylab(NULL)+
   theme_bw()+
-  theme(panel.grid.minor = element_blank())
+  theme(panel.grid.minor = element_blank(),
+        panel.border     = element_blank(), 
+        axis.line        = element_blank(),
+        axis.ticks       = element_blank(),
+        strip.background = element_blank(),
+        strip.text       = element_text(face="bold"),
+        panel.spacing    = unit(1, units = "cm"))
 ```
 
-# References
+\newpage{}
+
+\subsection{Stan Code}
+```{r stan1, echo=T, eval=F}
+data {
+  int<lower=0> Npreds;         # number of covariates, including intercept
+  int<lower=0> Npreds2;        # number of random effect covariates
+  int<lower=0> Nplots;         # number of plots
+  int<lower=0> Ntreats;        # number of treatments
+  int<lower=0> Nobs;           # number of observations
+  int<lower=0> Nyears;         # number of years
+  vector[Nobs] y;              # vector of observations
+  row_vector[Npreds] x[Nobs];	 # design matrix for fixed effects
+  row_vector[Npreds2] z[Nobs]; # simple design matrix for random effects 
+  int plot_id[Nobs];           # vector of plot ids
+  int treat_id[Nobs];          # vector of treatment ids
+  int year_id[Nobs];           # vector of year ids
+}
+
+parameters {
+	vector[Npreds] beta;	             # overall coefficients
+	vector[Nyears] year_off;           # vector of year effects
+	cholesky_factor_corr[Npreds2] L_u; # cholesky factor of plot random effect corr matrix
+  vector[Npreds2] beta_plot[Nplots]; # plot level random effects
+	vector<lower=0>[Npreds2] sigma_u;  # plot random effect std. dev.
+	real<lower=0> sd_y;                # treatment-level observation std. dev.
+	real<lower=0> sigma_year;          # year std. dev. hyperprior
+	real<lower=0> phi;                 # dispersion parameter
+}
+
+transformed parameters {
+  vector[Nobs] A;             # parameter for beta distn
+  vector[Nobs] B;             # parameter for beta distn
+  vector[Nobs] yhat;               # vector of expected values
+  vector[Npreds2] u[Nplots];       # transformed plot random effects
+  matrix[Npreds2,Npreds2] Sigma_u; # plot ranef cov matrix
+
+  Sigma_u = diag_pre_multiply(sigma_u, L_u); # cholesky factor for plot-level covariance matrix 
+  for(j in 1:Nplots)
+    u[j] = Sigma_u * beta_plot[j]; # plot random intercepts and slopes
+
+  # regression model for expected values (one for each plot-year)
+  for (i in 1:Nobs)
+    yhat[i] = inv_logit(x[i]*beta + z[i]*u[plot_id[i]] + year_off[year_id[i]]);
+
+  A = yhat * phi;
+  B = (1.0 - yhat) * phi;
+}
+
+model {
+  ####  PRIORS
+  phi ~ cauchy(0, 5); 
+  sigma_year ~ cauchy(0,2.5);
+  sd_y ~ cauchy(0,2.5);
+  year_off ~ normal(0,sigma_year); # priors on year effects, shared variance
+  beta ~ normal(0,5);	             # priors on treatment coefficients
+  L_u ~ lkj_corr_cholesky(2.0);      # prior on the cholesky factor which controls the 
+                                    # correlation between plot level treatment effects
+  sigma_u ~ cauchy(0,2.5);
+
+  for(i in 1:Nplots)
+		beta_plot[i] ~ normal(0,1); # plot-level coefficients for intercept and slope
+
+	####  LIKELIHOOD
+  y ~ beta(A, B); # observations vary according to beta distribution
+}
+
+generated quantities{
+  corr_matrix[Npreds2] R = multiply_lower_tri_self_transpose(L_u);
+  cov_matrix[Npreds2] V = quad_form_diag(R,sigma_u);
+}
+
+```

### 2017-12-10 manuscript/Appendix5.Rmd
more tiny revisions and matching up with response letter

--- 2017-12-06 manuscript/Appendix5.Rmd
@@
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
-\renewcommand\thefigure{Ax-\arabic{figure}}
-\renewcommand\theequation{Ax.\arabic{equation}}
-\renewcommand\thetable{Ax-\arabic{table}}  
-\renewcommand\thesection{Section Ax\arabic{section}}
+\renewcommand\thefigure{A5-\arabic{figure}}
+\renewcommand\theequation{A5.\arabic{equation}}
+\renewcommand\thetable{A5-\arabic{table}}  
+\renewcommand\thesection{Section A5.\arabic{section}}
 
 \begin{center}
-\textbf{\Large{Appendix X}} \\
-A.T. Tredennick, A.R. Kleinhesselink, B. Taylor \& P.B. Adler \\
+\textbf{\Large{Appendix 5}} \\
+A.T. Tredennick, A.R. Kleinhesselink, J.B. Taylor \& P.B. Adler \\
 ``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe'' \\
@@
 
-# Details on the plant community and dynamics
+\section{Characterizing Extreme Precipitation Amounts}
 
-Here we provide more details on the plant community in terms of dominance and rarity.
-Averaging across time, *Artemisia tripartita* and *Balsamorhiza sagittata* are the two most dominant species in each treatment.
-Combined, these two species represent 28\% of total cover in control plots, 25\% of total cover in drought plots, and 25\% of total cover in irrigation plots.
-Four to five species dominate the community in general (Figure Ax-1), indicating a high level of dominance in this plant community.
+Following the proposed methods of @Lemoine2016, we calculated quantiles from the empirical distribution of growing season precipitation at Dubios, ID.
+We chose the 1\% quantile to be indicative of extreme dry conditions (drought) and the 99\% quantile to be indicative of extreme wet conditions (irrigation).
+The data consist of 91 yearly records, which we assume are approximately normally distributed.
+The R code below shows our procedure, and Fig. A5-1 shows the results.
 
-```{r get_data, echo=FALSE, message=FALSE, warning=FALSE}
-####
-####  LOAD PACKAGES ----
-####
-library(tidyverse) # all the good things from Wikham
-library(dplyr)     # data munging
-library(stringr)   # working with strings
-library(ggthemes)  # pleasing ggplot2 themes
+```{r hrv, fig.cap="Density of the empirical distribution of growing season precipitation at Dubois, ID. Dashed vertical lines show the 1% and 99% quantiles, assuming a normal distribution."}
+library(ggplot2)
+weather    <- read.csv("../data/weather/ClimateIPM.csv")
+mean_ppt   <- mean(weather$ppt1)
+quants_ppt <- quantile(weather$ppt1,probs = c(0.01,0.99))
+quants_ppt[1]/mean_ppt*100 # percent of mean ppt for drought
+quants_ppt[2]/mean_ppt*100 # percent of mean ppt for irrigation
 
-####
-####  SET DIRECTORIES; READ IN DATA ----
-####
-# Set paths
-biomass_path <- "../data/estimated_biomass/"
-weather_path <- "../data/weather/"
-output_path  <- "../output/"
-figure_path  <- "../figures/"
-
-# Read in data
-plot_biomass <- readRDS(paste0(biomass_path,"permanent_plots_estimated_biomass.RDS"))
-weather_data <- read.csv(paste0(weather_path,"ClimateIPM.csv"))
-plot_info    <- read.csv("../data/estimated_biomass/quad_info.csv")
-
-suppressWarnings( # ignore factor to character warning. It's OK.
-  cover_species <- read.csv("../data/vital_rates/allrecords_cover_v24.csv") %>%
-    group_by(quad, year, species) %>%
-    summarise(total_area = sum(area)) %>%
-    left_join(plot_info, by = "quad") %>%
-    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
-    filter(!str_detect(QuadName, 'P1|P7')) %>%
-    group_by(year, species, Treatment) %>%
-    summarise(avg_area = mean(total_area)) %>%
-    ungroup()
-)
-
-####
-####  COMPARE COMMUNITY COMPOSITION AMONG TREATMENTS ----
-####  USING STANDARDIZED ABUNDANCES FROM COVER AND DENSITY
-####
-suppressWarnings( # ignore factor to character warning. It's OK.
-  cover_community_matrix <- read.csv("../data/vital_rates/allrecords_cover_v24.csv") %>%
-    group_by(quad, year, species) %>%
-    summarise(total_area = sum(area)) %>%
-    left_join(plot_info, by = "quad") %>%
-    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
-    filter(!str_detect(QuadName, 'P1|P7')) %>%
-    filter(year > 2010 & is.na(species)==F) %>%
-    spread(species,total_area, fill = 0) %>%
-    select(-QuadName,-Grazing,-paddock,-Group) %>%
-    ungroup()
-)
-
-suppressWarnings( # ignore factor to character warning. It's OK.
-  density_community_matrix <- read.csv("../data/vital_rates/allrecords_density_v24.csv") %>%
-    mutate(ind_num = 1) %>%
-    group_by(year, quad, species) %>%
-    summarise(total_inds = sum(ind_num)) %>%
-    left_join(plot_info, by = "quad") %>%
-    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
-    filter(!str_detect(QuadName, 'P1|P7')) %>%
-    filter(year > 2010 & is.na(species)==F) %>%
-    spread(species,total_inds, fill = 0) %>%
-    select(-QuadName,-Grazing,-paddock,-Group) %>%
-    ungroup()
-)
-
-density_spp <- colnames(density_community_matrix[4:ncol(density_community_matrix)])
-cover_spp   <- colnames(cover_community_matrix[4:ncol(cover_community_matrix)])
-overlap_spp <- density_spp[which(density_spp %in% cover_spp)]
-rmids       <- which(colnames(density_community_matrix) %in% overlap_spp)
-density_community_matrix <- density_community_matrix[,-rmids]
-
-suppressWarnings( # ignore factor to character warning. It's OK.
-  annual_community_matrix <- read.csv("../data/vital_rates/idaho_annuals.csv") %>%
-    select(-Notes) %>%
-    gather(species,abundance,-c(1:3)) %>%
-    left_join(plot_info, by="quad") %>%
-    select(-QuadName.y) %>%
-    rename(QuadName = QuadName.x) %>%
-    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
-    filter(!str_detect(QuadName, 'P1|P7')) %>%
-    filter(year > 2010 & is.na(species)==F) %>%
-    separate(species,c("genus","species1")) %>%
-    unite(species,genus,species1,sep = " ") %>%
-    spread(species,abundance, fill = 0) %>%
-    select(-QuadName,-Grazing,-paddock,-Group,-`Lappula occidentalis`,-`Tragopogon dubius`) %>%
-    ungroup()
-)
-
-annual_data <- {}
-for(dotrt in unique(annual_community_matrix$Treatment)){
-  for(doyr in unique(annual_community_matrix$year)){
-    tmp <- annual_community_matrix %>%
-      gather(species, density, -quad, -year, -Treatment) %>%
-      filter(Treatment == dotrt, year == doyr) %>%
-      group_by(Treatment, year, species) %>%
-      summarise(mean_density = mean(density)) %>%
-      arrange(-mean_density) %>%
-      mutate(mean_rank = seq_along(mean_density))
-
-    annual_data <- rbind(annual_data, tmp)
-  }
-}
-
-dom_annuals <- annual_data %>%
-  filter(species %in% c("Allyssum desertorum","Bromus tectorum")) %>%
-  select(-mean_rank) %>%
-  left_join(weather_data, by = "year")
+ggplot(weather, aes(x=ppt1))+
+  geom_histogram(bins=20, color="dodgerblue", fill="dodgerblue", aes(y=..density..))+
+  geom_line(stat="density", color="blue")+
+  geom_vline(aes(xintercept=quants_ppt[1]), linetype=2)+
+  geom_vline(aes(xintercept=quants_ppt[2]), linetype=2)+
+  ylab("Density")+
+  xlab("Growing Season Precipitation (mm)")+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank())
 ```
 
-\newpage{}
-# Table Ax-1: Statistical results from linear model relating density of Allysum desertorum to precipitation.
-```{r report_models, echo = FALSE}
-summary(lm(mean_density~ppt1*Treatment, data = filter(dom_annuals, species == "Allyssum desertorum")))
-```
-
-\newpage{}
-# Table Ax-2: Statistical results from linear model relating density of Bromus tectorum to precipitation.
-```{r report_models2, echo = FALSE}
-summary(lm(mean_density~ppt1*Treatment, data = filter(dom_annuals, species == "Bromus tectorum")))
-```
-
-\newpage{}
-
-```{r rac_plots, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Rank abundance curves for perennial species.
-Basal area of individuals was summed withnin years and plots, and then the total area values were averaged across years and plots for each treatment."}
-####
-####  EXTRACT COVER OF MOST DOMINANT SPECIES ----
-####
-cover_data <- {}
-for(dotrt in unique(cover_community_matrix$Treatment)){
-  tmp <- cover_community_matrix %>%
-    gather(species, area, -quad, -year, -Treatment) %>%
-    filter(Treatment == dotrt) %>%
-    group_by(Treatment, species) %>%
-    summarise(mean_area = mean(area)) %>%
-    arrange(-mean_area) %>%
-    mutate(mean_rank = seq_along(mean_area))
-
-  cover_data <- rbind(cover_data, tmp)
-}
-
-ggplot(cover_data, aes(x = mean_rank, y = mean_area))+
-  geom_line()+
-  geom_point(size = 2)+
-  geom_text(data=filter(cover_data, mean_rank < 11), aes(x = mean_rank, y = mean_area, label = species), 
-            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
-  scale_y_continuous(limits = c(0,0.25))+
-  xlab("Species rank")+
-  ylab(expression(paste("Mean basal area (",m^2,")")))+
-  facet_wrap(~Treatment, ncol = 3)
-
-cover_doms <- cover_data %>%
-  filter(mean_rank < 3) %>%
-  group_by(Treatment) %>%
-  summarise(total_area = sum(mean_area))
-
-```
-
-```{r annual_racs, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Rank abundance curves for annual species.
-Density of individuals is averaged across years and plots for each treatment."}
-annual_data <- {}
-for(dotrt in unique(annual_community_matrix$Treatment)){
-  tmp <- annual_community_matrix %>%
-    gather(species, density, -quad, -year, -Treatment) %>%
-    filter(Treatment == dotrt) %>%
-    group_by(Treatment, species) %>%
-    summarise(mean_density = mean(density)) %>%
-    arrange(-mean_density) %>%
-    mutate(mean_rank = seq_along(mean_density))
-
-  annual_data <- rbind(annual_data, tmp)
-}
-
-ggplot(annual_data, aes(x = mean_rank, y = mean_density))+
-  geom_line()+
-  geom_point(size = 2)+
-  geom_text(data=filter(annual_data, mean_rank < 3), aes(x = mean_rank, y = mean_density, label = species), 
-            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
-  scale_y_continuous(limits = c(0,200))+
-  xlab("Species rank")+
-  ylab("Mean density")+
-  facet_wrap(~Treatment, ncol = 3)
-```
-
-```{r time_perr_racs, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=6, fig.cap="Time series of rank abundance curves for perennial species in each treatment.
-Values of mean area were averaged over plots.
-The four most dominant species are labelled in each panel."}
-cover_data <- {}
-for(dotrt in unique(cover_community_matrix$Treatment)){
-  for(doyr in unique(cover_community_matrix$year)){
-    tmp <- cover_community_matrix %>%
-      gather(species, area, -quad, -year, -Treatment) %>%
-      filter(Treatment == dotrt, year == doyr) %>%
-      group_by(Treatment, year, species) %>%
-      summarise(mean_area = mean(area)) %>%
-      arrange(-mean_area) %>%
-      mutate(mean_rank = seq_along(mean_area))
-
-  cover_data <- rbind(cover_data, tmp)
-  }
-}
-ggplot(cover_data, aes(x = mean_rank, y = mean_area))+
-  geom_line()+
-  geom_point(size = 2)+
-  geom_text(data=filter(cover_data, mean_rank < 5), aes(x = mean_rank, y = mean_area, label = species), 
-            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
-  scale_y_continuous(limits = c(0,0.25))+
-  xlab("Species rank")+
-  ylab(expression(paste("Mean basal area (",m^2,")")))+
-  facet_grid(Treatment~year)
-```
-
-```{r time_annual_racs, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=6, fig.cap="Time series of rank abundance curves for annual species in each treatment.
-Values of density were averaged over plots.
-The two most dominant species are labelled in each panel."}
-annual_data <- {}
-for(dotrt in unique(annual_community_matrix$Treatment)){
-  for(doyr in unique(annual_community_matrix$year)){
-    tmp <- annual_community_matrix %>%
-      gather(species, density, -quad, -year, -Treatment) %>%
-      filter(Treatment == dotrt, year == doyr) %>%
-      group_by(Treatment, year, species) %>%
-      summarise(mean_density = mean(density)) %>%
-      arrange(-mean_density) %>%
-      mutate(mean_rank = seq_along(mean_density))
-
-    annual_data <- rbind(annual_data, tmp)
-  }
-}
-
-ggplot(annual_data, aes(x = mean_rank, y = mean_density))+
-  geom_line()+
-  geom_point(size = 2)+
-  geom_text(data=filter(annual_data, mean_rank < 3), aes(x = mean_rank, y = mean_density, label = species), 
-            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
-  scale_y_continuous(limits = c(0,400))+
-  scale_x_continuous(breaks = c(1,6,11))+
-  xlab("Species rank")+
-  ylab("Mean density")+
-  facet_grid(Treatment~year)
-```
-
-\newpage{}
-
-```{r annual_weather, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Relationship between precipitation and abundance of dominant annuals."}
-dom_annuals <- annual_data %>%
-  filter(species %in% c("Allyssum desertorum","Bromus tectorum")) %>%
-  select(-mean_rank) %>%
-  rename(Species = species) %>%
-  left_join(weather_data, by = "year")
-
-ggplot(dom_annuals, aes(x = ppt1, y = mean_density, color = Species))+
-  geom_point()+
-  xlab("Precipitation (mm)")+
-  ylab("Abundance")
-```
+# References

### 2017-12-10 manuscript/usses_water_ms.Rmd
more tiny revisions and matching up with response letter

@@
   - name: Andrew T. Tredennick
-    affiliation: 1,*
+    affiliation: 1
     corresponding: true
@@
   - name: Andrew R. Kleinhesselink
-    affiliation: 1,2,*
+    affiliation: 1,2
   - name: J. Bret Taylor
@@
 
+<!---
 \newcommand\blfootnote[1]{%
@@
 \blfootnote{*Authors contributed equally to this work.}
+--->
 
@@
 Sheep Experiment Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
-The plant community is dominated by the shrub \emph{Artemesia tripartita} and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata}.
+The plant community is dominated by the shrub \emph{Artemesia tripartita} and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata} \new{(see Appendix 1 for rank abundance curves)}.
 During the period of our experiment (2011 -- 2015), average mean annual precipitation was 265 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
@@
 We use these six plots as control plots (i.e. ambient precipitation) in the experiment described below.
+\new{We used the relocated plots as our controls because demographic data is difficult to collect and is already being collected in these plots for other ongoing studies.}
 
@@
 
-\new{We then modeled these daily deviations ($\Delta x_\text{trt.}$)} from ambient conditions using a linear mixed effects model with independent variables for treatment (irrigation or drought), season (winter, spring, summer, fall), rainfall, and all two-way interactions.
+\new{We then modeled these daily deviations ($\Delta x_{g,d,\text{trt.}}$)} from ambient conditions using a linear mixed effects model with independent variables for treatment (irrigation or drought), season (winter, spring, summer, fall), rainfall, and all two-way interactions.
 Rainy days were defined as any day in which precipitation was recorded and average temperature was above 3$^{\circ}$C.
@@
 We fit regressions to a MODIS-based index and an AVHRR-based index for each year and retained the regression with the better fit based on $R^2$ values.
-We then predicted ANPP using the best regression equation for each year (Appendix 1).
-\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 2).}
+We then predicted ANPP using the best regression equation for each year (Appendix 2).
+\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 4).}
 
@@
 \new{We used a vague prior distribution for each $\beta$: $\boldsymbol{\beta} \sim \text{Normal}(0,5)$.}
-A full description of our model and associated R [@R2016] code is in Appendix 2.
+A full description of our model and associated R [@R2016] code is in Appendix 3.
 
@@
 
-\new{We conducted the above analysis for all species, and then conducted separate analyses for perennial species only and annual species only.
-Annual species have shorter life spans, so conducting separate analyses allows us to discover if community responses are restricted to short-lived species that might respond more quickly to altered precipitation regimes.
-Given the dominance of perennial species in our system, shifts in the annual plant community could be masked in the analysis of the full community.}
+\new{We conducted the above analysis for all species, and then conducted a separate analyses for annual species only (Appendix 1). 
+Annual species have shorter life spans, so conducting a separate analysis allows us to discover if community responses are restricted to short-lived species that might respond more quickly to altered precipitation regimes.
+Given the dominance of perennial species in our system (Appendix 1), shifts in the annual plant community could be masked in the analysis of the full community.}
 
@@
 All R code and data necessary to reproduce our analysis has been archived on Figshare (\emph{link here after acceptance}) and released on GitHub (https://github.com/atredennick/usses_water/releases/v0.1).
-We also include annotated Stan code in our model description in Appendix 2.
+We also include annotated Stan code in our model description in Appendix 3.
 
@@
 Although the parameter estimates for the effect of soil moisture overlap zero, the posterior distributions of the slopes all shrank and shifted to more positive values relative to the prior distributions (Fig.
-A2-2), which indicates the data did influence parameter estimates beyond the information from the uninformative priors.
+A3-2), which indicates the data did influence parameter estimates beyond the information from the uninformative priors.
 Ecosystem functional response was similar among treatments (Table 1; Fig.
@@
 
-\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. x). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0002; Fig. x)}.
+\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3D). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0002; Fig. 3D)}.
 
@@
 Community composition was also remarkably stable over time, with no evidence of divergence among treatments (Table 2; Fig. 4).
-\new{Splitting the analysis apart for perennials and annuals led to similar results.
-The perennial community was stable over time and consistently similar among treatments (Fig. Ax-x).
+\new{There were also no changes in the dominant species over time in any treatment (Appendix 1).
+Splitting the analysis apart for annuals led to similar results.
 There is some evidence that the annual community changed in response to treatment in two years (Fig.
-Ax-x), but these responses appear to come from very rare species since our analysis weights them equally (Appendix X).}
+A1-6), but these responses appear to come from very rare species since our analysis weights them equally (Fig.
+A1-4).}
 
@@
 Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of mean growing season precipitation necessary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site (Fig.
-A4-1).
-The 1\% quantile of precipitation at our site is 110 mm, a 47% reduction from the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean growing season precipitation (Appendix 4).
+A5-1).
+The 1\% quantile of precipitation at our site is 110 mm, a 47% reduction from the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean growing season precipitation (Appendix 5).
 Thus, our drought treatment represented extreme precipitation amounts, especially in years where ambient precipitation was below average (Fig. 2A).
@@
 Second, ANPP at our site may be influenced by additional factors, not only the cumulative March-June soil moisture covariate we included in our statistical model.
-For example, temperature can impact ANPP directly [@Epstein1997] and by exacerbating the effects of soil moisture [@DeBoeck2011].
+\new{For example,} @LaPierre2016 \new{found that site-scale ANPP is better predicted by abiotic factors such as nutrient availability than precipitation.}
+Moreover, temperature can impact ANPP directly [@Epstein1997] and by exacerbating the effects of soil moisture [@DeBoeck2011].
 Measurements of soil moisture likely contain a signal of temperature, through its impact on evaporation and infiltration, but the measurements will not capture the direct effect of temperature on metabolic and physiological processes.
@@
 3B) and the resulting uncertain relationship we observed between soil moisture and ANPP across all treatments (Table 1, Fig.
-A2-2).
+A3-2).
 Unfortunately, we did not collect data on additional factors because our focus was on precipitation.
@@
 \new{The dominance of our site by the shrub \emph{A.
-tripartita} may explain why our results do not conform to broader patterns of grassland sensitivity to precipitation manipulations} [e.g., @Huxman2004; @Knapp2015; @Wilcox2017].
+tripartita} (Appendix 1) may explain why our results do not conform to broader patterns of grassland sensitivity to precipitation manipulations} [e.g., @Huxman2004; @Knapp2015; @Wilcox2017].
 
-\new{We did find that ANPP became more sensitive to drought over time (Fig. x).
-Thus, the difference between ANPP, standardize by the difference in soil moisture, in drought and control plots grew over time.
+\new{We did find that ANPP became more sensitive to drought over time (Fig. 3D).
+Thus, the difference between ANPP, standardized by the difference in soil moisture, in drought and control plots grew over time.
 The latter years of our experiment were relatively wet, meaning that the drought treatment could have the largest absolute reduction of soil moisture in those years.
@@
 Such insensitivity can buffer species against precipitation variability in this semi-arid ecosystem, making them successful in the long run.
-Longer, chronic precipitation alteration might reveal plant community shifts that we did not observe [e.g., @Wilcox2016]. 
-For example, a long-term increase in water availability could allow species that do not bet hedge to gain prominence and dominate the ecosystem functional response.
+Longer, chronic precipitation alteration might reveal plant community shifts that we did not observe [e.g., @Wilcox2016], \new{and the increased sensitivity of ANPP to drought in the final years of our experiment may portend such a shift.}
+Likewise, a long-term increase in water availability could allow species that do not bet hedge to gain prominence and dominate the ecosystem functional response.
 \new{Time \emph{per se} may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}
 Our results suggest compositional shifts would have the largest impact at high rainfall because the current community maintained consistent ecosystem functional response at very low water availability.
-\new{However, the increased sensitivity of ANPP to drought conditions suggests a shift in ecosystem functional response may be near, and simply take more time than we observed here.}
+\new{However, the increased sensitivity of ANPP to drought conditions suggests a shift in ecosystem functional response may be near, and simply takes more time than we observed here.}
 
@@
 \begin{description}
-\item [Appendix 1.] Additional methods and information on estimating aboveground net primary productivity.
-\item [Appendix 2.] \new{Results from analysis of NDVI without conversion to ANPP, Fig. A3-1.}
+\item [Appendix 1.] \new{Additional information on plant community composition and dynamics, Table A1-1, Table A1-2, Table A1-3, Fig. A1-1, Fig. A1-2, Fig. A1-3, Fig. A1-4, Fig. A1-5, and Fig. A1-6.}
+\item [Appendix 2.] Additional methods and information on estimating aboveground net primary productivity, Table A2-1, and Table A2-2.
 \item [Appendix 3.] Details of the hierarchical Bayesian model, Fig. A2-1, Fig. A2-2, and Fig. A2-3.
-\item [Appendix 4.] Details on analysis of precipitation historical range of variability and Fig. A4-1.
+\item [Appendix 4.] \new{Results from analysis of NDVI without conversion to ANPP, Fig. A3-1.}
+\item [Appendix 5.] Details on analysis of precipitation historical range of variability and Fig. A4-1.
 \end{description}
@@
                               "Lower 95\\% BCI", "Upper 95\\% BCI")
-coef_caption <- "Summary statistics from the posterior distributions of coefficients for each treatment."
+coef_caption <- "Summary statistics from the posterior distributions of coefficients for each treatment ($\\beta$ coefficients in equation 1).
+The `Intercept' and `Slope' summaries reported here for drought and irrigation are from the posterior distributions of the intercept and slope for the control treatment plus the offsets for each treatment.
+Posterior distributions of the offsets are in Figure 3A."
 print(xtable(effect_summary, caption = coef_caption, digits = 2),
@@
       \includegraphics[width=5in]{../figures/sppcomp_bray_all.png}
-  \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot, colored by treatment.}
+  \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot, colored by treatment. \new{Each point represents a plot and the axes represent species composition.}}
 \end{figure}

### 2017-12-10 manuscript/usses_water_ms.tex
more tiny revisions and matching up with response letter

@@
 
-\author[1,*]{Andrew T. Tredennick}
+\author[1]{Andrew T. Tredennick}
 
-\corrauthor[1,*]{Andrew T. Tredennick}{\href{mailto:atredenn@gmail.com}{\nolinkurl{atredenn@gmail.com}}}
-\author[1,2,*]{Andrew R. Kleinhesselink}
+\corrauthor[1]{Andrew T. Tredennick}{\href{mailto:atredenn@gmail.com}{\nolinkurl{atredenn@gmail.com}}}
+\author[1,2]{Andrew R. Kleinhesselink}
 
@@
 \definecolor{blue}{rgb}{0,0,0.7} \newcommand{\new}{\textcolor{blue}}
-
-\newcommand\blfootnote[1]{%
-  \begingroup
-  \renewcommand\thefootnote{}\footnote{#1}%
-  \addtocounter{footnote}{-1}%
-  \endgroup
-}
-
-\blfootnote{*Authors contributed equally to this work.}
 
@@
 \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and
-\emph{Hesperostipa comata}. During the period of our experiment (2011 --
-2015), average mean annual precipitation was 265 mm
-year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
+\emph{Hesperostipa comata}
+\new{(see Appendix 1 for rank abundance curves)}. During the period of
+our experiment (2011 -- 2015), average mean annual precipitation was 265
+mm year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
 -5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July. Between 1926
@@
 (i.e.~ambient precipitation) in the experiment described below.
+\new{We used the relocated plots as our controls because demographic data is difficult to collect and is already being collected in these plots for other ongoing studies.}
 
@@
 
-\new{We then modeled these daily deviations ($\Delta x_\text{trt.}$)}
+\new{We then modeled these daily deviations ($\Delta x_{g,d,\text{trt.}}$)}
 from ambient conditions using a linear mixed effects model with
@@
 \(R^2\) values. We then predicted ANPP using the best regression
-equation for each year (Appendix 1).
-\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 2).}
+equation for each year (Appendix 2).
+\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 4).}
 
@@
 A full description of our model and associated R \citep{R2016} code is
-in Appendix 2.
+in Appendix 3.
 
@@
 We also include annotated Stan code in our model description in Appendix
-2.
+3.
 
@@
 posterior distributions of the slopes all shrank and shifted to more
-positive values relative to the prior distributions (Fig. A2-2), which
+positive values relative to the prior distributions (Fig. A3-2), which
 indicates the data did influence parameter estimates beyond the
@@
 divergence among treatments (Table 2; Fig. 4).
-\new{Splitting the analysis apart for perennials and annuals led to similar results.
-The perennial community was stable over time and consistently similar among treatments (Fig. Ax-x).
+\new{There were also no changes in the dominant species over time in any treatment (Appendix 1).
+Splitting the analysis apart for annuals led to similar results.
 There is some evidence that the annual community changed in response to treatment in two years (Fig.
-Ax-x), but these responses appear to come from very rare species since our analysis weights them equally (Appendix X).}
+A1-6), but these responses appear to come from very rare species since our analysis weights them equally (Fig.
+A1-4).}
 
@@
 precipitation necessary to reach the 1\% and 99\% extremes of the
-historical precipitation regime at our site (Fig. A4-1). The 1\%
+historical precipitation regime at our site (Fig. A5-1). The 1\%
 quantile of precipitation at our site is 110 mm, a 47\% reduction from
 the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean
-growing season precipitation (Appendix 4). Thus, our drought treatment
+growing season precipitation (Appendix 5). Thus, our drought treatment
 represented extreme precipitation amounts, especially in years where
@@
 only the cumulative March-June soil moisture covariate we included in
-our statistical model. For example, temperature can impact ANPP directly
-\citep{Epstein1997} and by exacerbating the effects of soil moisture
-\citep{DeBoeck2011}. Measurements of soil moisture likely contain a
-signal of temperature, through its impact on evaporation and
-infiltration, but the measurements will not capture the direct effect of
-temperature on metabolic and physiological processes. We also did not
-redistribute snow across our treatments in the winter, and snow melt may
-spur early spring growth. Failure to account for potentially important
-covariates could explain the within-year spread of ANPP (Fig 2C, Fig.
-3B) and the resulting uncertain relationship we observed between soil
-moisture and ANPP across all treatments (Table 1, Fig. A2-2).
-Unfortunately, we did not collect data on additional factors because our
-focus was on precipitation.
+our statistical model. \new{For example,} \citet{LaPierre2016}
+\new{found that site-scale ANPP is better predicted by abiotic factors such as nutrient availability than precipitation.}
+Moreover, temperature can impact ANPP directly \citep{Epstein1997} and
+by exacerbating the effects of soil moisture \citep{DeBoeck2011}.
+Measurements of soil moisture likely contain a signal of temperature,
+through its impact on evaporation and infiltration, but the measurements
+will not capture the direct effect of temperature on metabolic and
+physiological processes. We also did not redistribute snow across our
+treatments in the winter, and snow melt may spur early spring growth.
+Failure to account for potentially important covariates could explain
+the within-year spread of ANPP (Fig 2C, Fig. 3B) and the resulting
+uncertain relationship we observed between soil moisture and ANPP across
+all treatments (Table 1, Fig. A3-2). Unfortunately, we did not collect
+data on additional factors because our focus was on precipitation.
 
@@
 \citep{Kulmatiski2017a}.
-\new{The dominance of our site by the shrub \emph{A. tripartita} may explain why our results do not conform to broader patterns of grassland sensitivity to precipitation manipulations}
+\new{The dominance of our site by the shrub \emph{A. tripartita} (Appendix 1) may explain why our results do not conform to broader patterns of grassland sensitivity to precipitation manipulations}
 \citep[e.g.,][]{Huxman2004, Knapp2015, Wilcox2017}.
 
-\new{We did find that ANPP became more sensitive to drought over time (Fig. x).
-Thus, the difference between ANPP, standardize by the difference in soil moisture, in drought and control plots grew over time.
+\new{We did find that ANPP became more sensitive to drought over time (Fig. 3D).
+Thus, the difference between ANPP, standardized by the difference in soil moisture, in drought and control plots grew over time.
 The latter years of our experiment were relatively wet, meaning that the drought treatment could have the largest absolute reduction of soil moisture in those years.
@@
 precipitation alteration might reveal plant community shifts that we did
-not observe \citep[e.g.,][]{Wilcox2016}. For example, a long-term
-increase in water availability could allow species that do not bet hedge
-to gain prominence and dominate the ecosystem functional response.
+not observe \citep[e.g.,][]{Wilcox2016},
+\new{and the increased sensitivity of ANPP to drought in the final years of our experiment may portend such a shift.}
+Likewise, a long-term increase in water availability could allow species
+that do not bet hedge to gain prominence and dominate the ecosystem
+functional response.
 \new{Time \emph{per se} may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}
@@
 ecosystem functional response at very low water availability.
-\new{However, the increased sensitivity of ANPP to drought conditions suggests a shift in ecosystem functional response may be near, and simply take more time than we observed here.}
+\new{However, the increased sensitivity of ANPP to drought conditions suggests a shift in ecosystem functional response may be near, and simply takes more time than we observed here.}
 
@@
 \centering
-\caption{Summary statistics from the posterior distributions of coefficients for each treatment.} 
+\caption{Summary statistics from the posterior distributions of coefficients for each treatment ($\beta$ coefficients in equation 1).
+The `Intercept' and `Slope' summaries reported here for drought and irrigation are from the posterior distributions of the intercept and slope for the control treatment plus the offsets for each treatment.
+Posterior distributions of the offsets are in Figure 3A.}
 \begingroup\normalsize

### 2017-12-10 manuscript/usses_water_ms.Rmd
more revisions before sending to coauthors

@@
 
-<!---
 \newcommand\blfootnote[1]{%
@@
 \blfootnote{*Authors contributed equally to this work.}
---->
 
@@
 For example, the soil moisture-ANPP relationship may be linear within the historical range of interannual variation, but could saturate at higher levels of soil moisture. 
-In fact, saturating relationships are actually common [@Hsu2012; @Gherardi2015a], perhaps because other resources, like nitrogen, become more limiting in wet years than dry years.
+Saturating relationships are actually common [@Hsu2012; @Gherardi2015a], perhaps because other resources, like nitrogen, become more limiting in wet years than dry years.
 Failure to accurately estimate the curvature of the soil moisture-ANPP relationship will lead to over- or underprediction of ANNP under extreme precipitation [@Peters2012].
@@
 Sheep Experiment Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
-The plant community is dominated by the shrub \emph{Artemesia tripartita} and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata} \new{(see Appendix 1 for rank abundance curves)}.
+The plant community is dominated by the shrub \emph{Artemesia tripartita}, \new{the perennial forb \emph{Balsamorhiza sagittata}}, and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata} \new{(see Appendix 1 for rank abundance curves)}.
 During the period of our experiment (2011 -- 2015), average mean annual precipitation was 265 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
@@
 We use these six plots as control plots (i.e. ambient precipitation) in the experiment described below.
-\new{We used the relocated plots as our controls because demographic data is difficult to collect and is already being collected in these plots for other ongoing studies.}
+\new{We used the relocated plots as our controls because collecting demographic data is difficult and it is already being collected in these plots for other ongoing studies.}
 
@@
 Within each group of plots we standardized the irrigation and drought effects on soil moisture relative to the ambient soil moisture conditions in the control plot.
-Specifically we subtracted the ambient daily soil moisture measured in the control plot from the soil moisture measured within the drought and irrigation plots within each group and then divided by the standard deviation of daily soil moisture values measured in the control plot ($\Delta x_{g,d,\text{trt.}} = (x_{g,d,\text{trt.}} - x_{g,d,\text{control}})/sd(x_{g,d,\text{control}})$) where $\Delta x$ is the standardized treatment effect and $x_{g,d,\text{trt.}}$ is the raw soil moisture measure for plot group $g$ on day $d$ and treatment $\text{trt.}$.
+Specifically we subtracted the ambient daily soil moisture measured in the control plot from the soil moisture measured within the drought and irrigation plots within each group and then divided by the standard deviation of daily soil moisture values measured in the control plot $\left( \Delta x_{g,d,\text{trt.}} = (x_{g,d,\text{trt.}} - x_{g,d,\text{control}})/\text{s.d.}(x_{g,d,\text{control}}) \right)$ where $\Delta x$ is the standardized treatment effect and $x_{g,d,\text{trt.}}$ is the raw soil moisture measure for plot group $g$ on day $d$ and treatment $\text{trt.}$.
 These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.}
@@
 This procedure allowed us to predict daily soil moisture conditions for all plots, even on those days when some of our direct treatment measurements were missing due to malfunction of the sensors or dataloggers.
-See @Kleinhesselink2017b for complete details on our approach to estimating soil moisture.
+See @Kleinhesselink2017b for more details on our approach to estimating soil moisture.
 
@@
 \new{We define `sensitivity' as $\frac{ \text{ANPP}_{\text{control}} - \text{ANPP}_{\text{treatment}} } { \text{VWC}_{\text{control}} - \text{VWC}_{\text{treatment}} }$, following} @Wilcox2017.
-\new{This metric of sensitivity measures the response of ANPP standardized to the magnitude of the treatment. 
+\new{This metric of sensitivity measures the response of ANPP standardized by the magnitude of the treatment. 
 Because our treatment and control plots are not paired, we conducted this analysis by comparing each treatment plot to the mean of the control plots in each year. 
@@
 
-\new{We conducted the above analysis for all species, and then conducted a separate analyses for annual species only (Appendix 1). 
-Annual species have shorter life spans, so conducting a separate analysis allows us to discover if community responses are restricted to short-lived species that might respond more quickly to altered precipitation regimes.
-Given the dominance of perennial species in our system (Appendix 1), shifts in the annual plant community could be masked in the analysis of the full community.}
+\new{We conducted the above analysis for all species, and then conducted a separate analysis for annual species only (Appendix 1). 
+Annual species have shorter life spans, so conducting a separate analysis allowed us to discover if community responses are restricted to short-lived species that might respond more quickly to altered precipitation regimes.
+Given the dominance of perennial species in our system (Appendix 1), shifts in the annual plant community could be difficult to detect in the analysis of the full community.}
 
@@
 
-\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3D). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0002; Fig. 3D)}.
+\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3C). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0002; Fig. 3C)}.
 
@@
 A1-6), but these responses appear to come from very rare species since our analysis weights them equally (Fig.
-A1-4).}
+A1-4).
+By definition, thsee rare species contribute little to ANPP and thus any compositional changes have little influence on ecosystem functional response.}
 
@@
 
-\new{We did find that ANPP became more sensitive to drought over time (Fig. 3D).
-Thus, the difference between ANPP, standardized by the difference in soil moisture, in drought and control plots grew over time.
+\new{We did find that ANPP became more sensitive to drought over time (Fig. 3C).
 The latter years of our experiment were relatively wet, meaning that the drought treatment could have the largest absolute reduction of soil moisture in those years.
 But our metric of sensitivity accounts for differences among years because ANPP differences is standardized by the difference in soil moisture within the year.
-Therefore, it is likely that the increased sensitivity to drought over time is due to the cimulative impact of drought on the dominant species.
+Therefore, it is likely that the increased sensitivity to drought over time is due to the cumulative impact of drought on the dominant species.
 Ecosystem functional response may soon shift if sensitivity continues to increase.}

### 2017-12-10 manuscript/usses_water_ms.tex
more revisions before sending to coauthors

@@
 
+\newcommand\blfootnote[1]{%
+  \begingroup
+  \renewcommand\thefootnote{}\footnote{#1}%
+  \addtocounter{footnote}{-1}%
+  \endgroup
+}
+
+\blfootnote{*Authors contributed equally to this work.}
+
 \reversemarginpar
@@
 relationship may be linear within the historical range of interannual
-variation, but could saturate at higher levels of soil moisture. In
-fact, saturating relationships are actually common
+variation, but could saturate at higher levels of soil moisture.
+Saturating relationships are actually common
 \citep{Hsu2012, Gherardi2015a}, perhaps because other resources, like
@@
 sea level. The plant community is dominated by the shrub
-\emph{Artemesia tripartita} and three perennial bunchgrasses,
-\emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and
-\emph{Hesperostipa comata}
+\emph{Artemesia tripartita},
+\new{the perennial forb \emph{Balsamorhiza sagittata}}, and three
+perennial bunchgrasses, \emph{Pseudoroegneria spicata},
+\emph{Poa secunda}, and \emph{Hesperostipa comata}
 \new{(see Appendix 1 for rank abundance curves)}. During the period of
@@
 (i.e.~ambient precipitation) in the experiment described below.
-\new{We used the relocated plots as our controls because demographic data is difficult to collect and is already being collected in these plots for other ongoing studies.}
+\new{We used the relocated plots as our controls because collecting demographic data is difficult and it is already being collected in these plots for other ongoing studies.}
 
@@
 were missing due to malfunction of the sensors or dataloggers. See
-\citet{Kleinhesselink2017b} for complete details on our approach to
+\citet{Kleinhesselink2017b} for more details on our approach to
 estimating soil moisture.
@@
 \citet{Wilcox2017}.
-\new{This metric of sensitivity measures the response of ANPP standardized to the magnitude of the treatment. 
+\new{This metric of sensitivity measures the response of ANPP standardized by the magnitude of the treatment. 
 Because our treatment and control plots are not paired, we conducted this analysis by comparing each treatment plot to the mean of the control plots in each year. 

### 2017-12-22 manuscript/usses_water_ms.Rmd
adding some details on quads and sample sizes

@@
 Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
-In 2007, we relocated 14 of the original quadrats, six of which were inside a large, permanent livestock exclosure.
-We use these six plots as control plots (i.e. ambient precipitation) in the experiment described below.
+In 2007, we relocated 14 of the original quadrats in permenanent livestock exclosures, \new{which we used} as control plots (i.e. ambient precipitation) in the experiment described below.
 \new{We used the relocated plots as our controls because collecting demographic data is difficult and it is already being collected in these plots for other ongoing studies.}
 
-In spring 2011, we established 16 new 1 m$^2$ plots located in the same exclosure as the six control plots.
+In spring 2011, we established 16 new 1 m$^2$ plots located in the \new{largest exclosure at USSES, which also contained} six of our control plots.
 We avoided areas on steep hill slopes, areas with greater than 20\% cover of bare rock, and areas with greater than 10\% cover of the shrubs \emph{Purshia tridentata} and/or \emph{Amelanchier utahensis}.
 We established the new plots in pairs and randomly assigned each plot in a pair to receive a "drought" or "irrigation" treatment.
+\new{Thus, our experiment consisted of $n=14$ control plots, $n=8$ irrigation plots, and $n=8$ irrigation plots, for a total of 30 plots.}
 

### 2017-12-22 manuscript/usses_water_ms.tex
adding some details on quads and sample sizes

@@
 
-\definecolor{blue}{rgb}{0,0,0.7} \newcommand{\new}{\textcolor{blue}}
+\definecolor{blue}{rgb}{0,0,0.7}\newcommand{\new}{\textcolor{blue}}
 
@@
   \endgroup
-}
-
-\blfootnote{*Authors contributed equally to this work.}
+}\blfootnote{*Authors contributed equally to this work.}
 
@@
 
-\section{INTRODUCTION}\label{introduction}
+\hypertarget{introduction}{%
+\section{INTRODUCTION}\label{introduction}}
 
@@
 
-\section{METHODS}\label{methods}
+\hypertarget{methods}{%
+\section{METHODS}\label{methods}}
 
-\subsection{Study Area}\label{study-area}
+\hypertarget{study-area}{%
+\subsection{Study Area}\label{study-area}}
 
@@
 m\(^2\) quadrats to track vegetation change over time. In 2007, we
-relocated 14 of the original quadrats, six of which were inside a large,
-permanent livestock exclosure. We use these six plots as control plots
-(i.e.~ambient precipitation) in the experiment described below.
+relocated 14 of the original quadrats in permenanent livestock
+exclosures, \new{which we used} as control plots (i.e.~ambient
+precipitation) in the experiment described below.
 \new{We used the relocated plots as our controls because collecting demographic data is difficult and it is already being collected in these plots for other ongoing studies.}
@@
 In spring 2011, we established 16 new 1 m\(^2\) plots located in the
-same exclosure as the six control plots. We avoided areas on steep hill
-slopes, areas with greater than 20\% cover of bare rock, and areas with
-greater than 10\% cover of the shrubs \emph{Purshia tridentata} and/or
+\new{largest exclosure at USSES, which also contained} six of our
+control plots. We avoided areas on steep hill slopes, areas with greater
+than 20\% cover of bare rock, and areas with greater than 10\% cover of
+the shrubs \emph{Purshia tridentata} and/or
 \emph{Amelanchier utahensis}. We established the new plots in pairs and
@@
 ``irrigation'' treatment.
+\new{Thus, our experiment consisted of $n=14$ control plots, $n=8$ irrigation plots, and $n=8$ irrigation plots, for a total of 30 plots.}
 
-\subsection{Precipitation Experiment}\label{precipitation-experiment}
+\hypertarget{precipitation-experiment}{%
+\subsection{Precipitation Experiment}\label{precipitation-experiment}}
 
@@
 
-\subsection{Data Collection}\label{data-collection}
+\hypertarget{data-collection}{%
+\subsection{Data Collection}\label{data-collection}}
 
@@
 
-\subsection{Data Analysis}\label{data-analysis}
+\hypertarget{data-analysis}{%
+\subsection{Data Analysis}\label{data-analysis}}
 
+\hypertarget{ecosystem-functional-response}{%
 \subsubsection{Ecosystem functional
-response}\label{ecosystem-functional-response}
+response}\label{ecosystem-functional-response}}
 
@@
 
+\hypertarget{sensitivity-of-anpp-over-time}{%
 \subsubsection{Sensitivity of ANPP over
-time}\label{sensitivity-of-anpp-over-time}
+time}\label{sensitivity-of-anpp-over-time}}
 
@@
 
+\hypertarget{community-composition-over-time}{%
 \subsubsection{Community composition over
-time}\label{community-composition-over-time}
+time}\label{community-composition-over-time}}
 
@@
 
-\subsubsection{Reproducibility}\label{reproducibility}
+\hypertarget{reproducibility}{%
+\subsubsection{Reproducibility}\label{reproducibility}}
 
@@
 
-\section{RESULTS}\label{results}
+\hypertarget{results}{%
+\section{RESULTS}\label{results}}
 
@@
 
-\section{DISCUSSION}\label{discussion}
+\hypertarget{discussion}{%
+\section{DISCUSSION}\label{discussion}}
 
@@
 
-\section{CONCLUSIONS}\label{conclusions}
+\hypertarget{conclusions}{%
+\section{CONCLUSIONS}\label{conclusions}}
 
@@
 
-\section{ACKNOWLEDGEMENTS}\label{acknowledgements}
+\hypertarget{acknowledgements}{%
+\section{ACKNOWLEDGEMENTS}\label{acknowledgements}}
 
@@
 
-\section{FUNDING}\label{funding}
+\hypertarget{funding}{%
+\section{FUNDING}\label{funding}}
 
@@
 
-\section{AUTHOR CONTRIBUTIONS}\label{author-contributions}
+\hypertarget{author-contributions}{%
+\section{AUTHOR CONTRIBUTIONS}\label{author-contributions}}
 
@@
 
-\section{SUPPLEMENTAL INFORMATION}\label{supplemental-information}
+\hypertarget{supplemental-information}{%
+\section{SUPPLEMENTAL INFORMATION}\label{supplemental-information}}
 
@@
 
-\section{TABLES}\label{tables}
+\hypertarget{tables}{%
+\section{TABLES}\label{tables}}
 
@@
 
-\section{FIGURES}\label{figures}
+\hypertarget{figures}{%
+\section{FIGURES}\label{figures}}
 
@@
 
-\bibliography{/Users/atredenn/Dropbox/Bibliography/usses_water.bib}
+\bibliography{/Users/atredenn/Dropbox/Bibliography/usses\_water.bib}
 

### 2017-12-22 code/06_community_dynamics.R

@@ -127,6 +127,18 @@ full_community_matrix_scaled <- full_community_matrix_scaled %>%
          -`Linanthus pungens`)
 
 
+##  Look at quadrats in each year
+# checker <- full_community_matrix_scaled %>%
+#   select(quad,year,Treatment) %>%
+#   arrange(-year)
+# unique(filter(checker, year == 2011)$quad)
+# unique(filter(checker, year == 2012)$quad)
+# unique(filter(checker, year == 2013)$quad)
+# unique(filter(checker, year == 2014)$quad)
+# unique(filter(checker, year == 2015)$quad)
+# unique(filter(checker, year == 2016)$quad)
+
+
 nmds_df <- {}
 out_stats <- {}
 # pdf("../figures/bray_raw.pdf",onefile = TRUE)

### 2017-12-30 code/06_community_dynamics.R
updates mapping data and community analysis

@@ -40,7 +40,7 @@ weather_data <- read.csv(paste0(weather_path,"ClimateIPM.csv"))
 plot_info    <- read.csv("../data/estimated_biomass/quad_info.csv")
 
 suppressWarnings( # ignore factor to character warning. It's OK.
-  cover_species <- read.csv("../data/vital_rates/allrecords_cover_v24.csv") %>%
+  cover_species <- read.csv("../data/vital_rates/allrecords_cover.csv") %>%
     group_by(quad, year, species) %>%
     summarise(total_area = sum(area)) %>%
     left_join(plot_info, by = "quad") %>%
@@ -59,7 +59,7 @@ suppressWarnings( # ignore factor to character warning. It's OK.
 ####  USING STANDARDIZED ABUNDANCES FROM COVER AND DENSITY
 ####
 suppressWarnings( # ignore factor to character warning. It's OK.
-  cover_community_matrix <- read.csv("../data/vital_rates/allrecords_cover_v24.csv") %>%
+  cover_community_matrix <- read.csv("../data/vital_rates/allrecords_cover.csv") %>%
     group_by(quad, year, species) %>%
     summarise(total_area = sum(area)) %>%
     left_join(plot_info, by = "quad") %>%
@@ -72,7 +72,7 @@ suppressWarnings( # ignore factor to character warning. It's OK.
 )
 
 suppressWarnings( # ignore factor to character warning. It's OK.
-  density_community_matrix <- read.csv("../data/vital_rates/allrecords_density_v24.csv") %>%
+  density_community_matrix <- read.csv("../data/vital_rates/allrecords_density.csv") %>%
     mutate(ind_num = 1) %>%
     group_by(year, quad, species) %>%
     summarise(total_inds = sum(ind_num)) %>%

### 2017-12-30 manuscript/usses_water_ms.Rmd
updates ms with details of sample size for demographic data

@@
 Species composition data came from two sources: yearly census maps for each plot made using a pantograph [@Hill1920] and yearly counts of annual species in each plot.
+\new{Mapping was done for plots within one of the permanent exclosures, which included six of the control plots and all of the treatment plots, for a total of 22 plots.}
 From these sources, we determined the density of all annuals and perennials forbs, the basal cover of perennial grasses, and the canopy cover of shrubs.

### 2017-12-30 manuscript/usses_water_ms.tex
updates ms with details of sample size for demographic data

@@
 each plot made using a pantograph \citep{Hill1920} and yearly counts of
-annual species in each plot. From these sources, we determined the
-density of all annuals and perennials forbs, the basal cover of
-perennial grasses, and the canopy cover of shrubs. We made a large
-plot-treatment-year by species matrix, where columns were filled with
-either basal cover or density, depending on the measurement made for the
-particular species. We standardized the values in each column so we
-could directly compare species quantified with different metrics
-(density, basal cover, and canopy cover). This puts all abundance values
-on the same scale, meaning that common and rare species are weighted
-equally. Assuming that rare species will respond to treatments more than
-common ones, our approach is biased towards detecting compositional
-changes.
+annual species in each plot.
+\new{Mapping was done for plots within one of the permanent exclosures, which included six of the control plots and all of the treatment plots, for a total of 22 plots.}
+From these sources, we determined the density of all annuals and
+perennials forbs, the basal cover of perennial grasses, and the canopy
+cover of shrubs. We made a large plot-treatment-year by species matrix,
+where columns were filled with either basal cover or density, depending
+on the measurement made for the particular species. We standardized the
+values in each column so we could directly compare species quantified
+with different metrics (density, basal cover, and canopy cover). This
+puts all abundance values on the same scale, meaning that common and
+rare species are weighted equally. Assuming that rare species will
+respond to treatments more than common ones, our approach is biased
+towards detecting compositional changes.
 
@@
   \hline
-2011 & adonis &  21 &   2 & 1.02 & 0.45 \\ 
-  2011 & betadisper &  21 &   2 & 2.22 & 0.14 \\ 
-  2012 & adonis &  22 &   2 & 1.09 & 0.34 \\ 
+2011 & adonis &  22 &   2 & 1.04 & 0.42 \\ 
+  2011 & betadisper &  22 &   2 & 2.22 & 0.14 \\ 
+  2012 & adonis &  22 &   2 & 1.09 & 0.33 \\ 
   2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
-  2013 & adonis &  22 &   2 & 1.25 & 0.13 \\ 
-  2013 & betadisper &  22 &   2 & 0.44 & 0.65 \\ 
-  2014 & adonis &  22 &   2 & 0.95 & 0.56 \\ 
-  2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
-  2015 & adonis &  21 &   2 & 1.05 & 0.39 \\ 
-  2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 
-  2016 & adonis &  21 &   2 & 1.07 & 0.36 \\ 
-  2016 & betadisper &  21 &   2 & 0.49 & 0.62 \\ 
+  2013 & adonis &  22 &   2 & 1.26 & 0.10 \\ 
+  2013 & betadisper &  22 &   2 & 0.43 & 0.66 \\ 
+  2014 & adonis &  22 &   2 & 0.96 & 0.53 \\ 
+  2014 & betadisper &  22 &   2 & 0.36 & 0.70 \\ 
+  2015 & adonis &  22 &   2 & 1.08 & 0.31 \\ 
+  2015 & betadisper &  22 &   2 & 1.96 & 0.17 \\ 
+  2016 & adonis &  22 &   2 & 1.12 & 0.24 \\ 
+  2016 & betadisper &  22 &   2 & 0.27 & 0.76 \\ 
    \hline

### 2018-01-03 manuscript/usses_water_ms.Rmd
Edit manuscript soil moisture and sensitivity
Rewording the soil moisture description and the
description of the sensitivity analysis.

@@
 Only a subset of plots were equipped with soil moisture sensors, and within those plots, one or more of the sensors frequently failed to collect data.
-To remedy these problems, and to produce average daily soil moisture values for the ambient, drought, and irrigation treatments, we used a statistical model to predict the average treatment effects on soil moisture during the course of the experiment.
+To remedy these problems, we used a statistical model to produce average daily soil moisture values for the ambient, drought, and irrigation treatments during the course of the experiment.
 
 \new{We first averaged the observed soil moisture readings for each day ($d$) and plot ($i$), $x_{i,d}$.
-Plots were organized in groups of three, with each group ($g$) containing consisting of a control, a drought, and an irrigation plot.
-Within each group of plots we standardized the irrigation and drought effects on soil moisture relative to the ambient soil moisture conditions in the control plot.
-Specifically we subtracted the ambient daily soil moisture measured in the control plot from the soil moisture measured within the drought and irrigation plots within each group and then divided by the standard deviation of daily soil moisture values measured in the control plot $\left( \Delta x_{g,d,\text{trt.}} = (x_{g,d,\text{trt.}} - x_{g,d,\text{control}})/\text{s.d.}(x_{g,d,\text{control}}) \right)$ where $\Delta x$ is the standardized treatment effect and $x_{g,d,\text{trt.}}$ is the raw soil moisture measure for plot group $g$ on day $d$ and treatment $\text{trt.}$.
+Experimental plots were organized in groups of two, with each group ($g$) containing a drought and irrigated plot.
+In addition a nearby area outside the drought or irrigated plots was monitored for local ambient soil moisture conditions.
+Within each group of plots we standardized the irrigation and drought effects on soil moisture relative to the ambient soil moisture conditions.
+Specifically we subtracted the ambient daily soil moisture from the soil moisture measured within the drought and irrigation plots within each group and then divided by the standard deviation of daily soil moisture values measured in the ambient conditions $\left( \Delta x_{g,d,\text{trt.}} = (x_{g,d,\text{trt.}} - x_{g,d,\text{ambient}})/\text{s.d.}(x_{g,d,\text{ambient}}) \right)$ where $\Delta x$ is the standardized treatment effect and $x_{g,d,\text{trt.}}$ is the raw soil moisture measure for plot group $g$ on day $d$ and treatment $\text{trt.}$.
 These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.}
@@
 \new{We then modeled these daily deviations ($\Delta x_{g,d,\text{trt.}}$)} from ambient conditions using a linear mixed effects model with independent variables for treatment (irrigation or drought), season (winter, spring, summer, fall), rainfall, and all two-way interactions.
-Rainy days were defined as any day in which precipitation was recorded and average temperature was above 3$^{\circ}$C.
-The day immediately following rainfall was also classified as rainy.
+Using the local daily weather station data, we recorded rainy days as any day with measureable precipitation or the day after such a day and with average temperatures above 3$^{\circ}$C (to exclude days with snowfall).
 We fit the model using the ```lme4::lmer()``` function [@Bates2015] in R [@R2016], with random effects for plot group and date.
@@
 \new{We then used the model to predict the deviations from ambient conditions produced in the treated plots on each day of the experiment.  
-We added these predicted deviations to the average daily ambient soil moisture to generate predictions for daily soil moisture in all of the treated plots: $\bar{x}_{d, \text{trt.}} = \Delta \bar{x}_{d,\text{trt.}} + \bar{x}_{d,\text{control}}$, where $\bar{x}_{d,\text{trt.}}$ is the average predicted daily soil moisture in the treated plots and $\bar{x}_{d,\text{control}}$ is the daily ambient soil moisture averaged across all control plots.}
+We added these predicted deviations to the average daily ambient soil moisture to generate predictions for daily soil moisture in all of the treated plots: $\bar{x}_{d, \text{trt.}} = \Delta \bar{x}_{d,\text{trt.}} + \bar{x}_{d,\text{ambient}}$, where $\bar{x}_{d,\text{trt.}}$ is the average predicted daily soil moisture in the treated plots and $\bar{x}_{d,\text{ambient}}$ is the daily ambient soil moisture averaged across all control plots.}
 We could only predict soil moisture in the treated plots on days for which we took at least one ambient soil moisture measurement.
@@
 ### Sensitivity of ANPP over time
-\new{Unfortunately, our data does not allow us to include the effect of time on ecosystem functional response in the model described above. 
+\new{Our data does not allow us to directly test whether the ecosystem functional response to precipitation may have changed over time in each of the treatments. 
 This is because we lack sufficient within-year and within-treatment variation of soil moisture (i.e., within a year and treatment each plot shares the same value of VWC). 
-Therefore, we conducted an independent analysis to determine if the sensitivity of ANPP to treatment changed over time.}
+However, we did use a separate analysis to test whether the the sensitivity of ANPP to treatment changed over time.}
 
-\new{We define `sensitivity' as $\frac{ \text{ANPP}_{\text{control}} - \text{ANPP}_{\text{treatment}} } { \text{VWC}_{\text{control}} - \text{VWC}_{\text{treatment}} }$, following} @Wilcox2017.
-\new{This metric of sensitivity measures the response of ANPP standardized by the magnitude of the treatment. 
-Because our treatment and control plots are not paired, we conducted this analysis by comparing each treatment plot to the mean of the control plots in each year. 
-Then, independently for each treatment, we regressed sensitivity against year of treatment using the \texttt{lm()} function in R} [@R2016].
+\new{We define `sensitivity' as $\frac{ \text{ANPP}_{\text{control}} - \text{ANPP}_{\text{treatment}} } { \text{VWC}_{\text{ambient}} - \text{VWC}_{\text{treatment}} }$, following} @Wilcox2017.
+\new{This metric of sensitivity scales the difference between ANPP in the treated and control plots by the change in soil moisture in the treated plots. 
+Because our treated plots are not directly paired with individual control plots, we compare ANPP in each treated plot to the average ANPP of the control plots in each year.
+After calculating sensitivity for each drought and irrigation plot in each year, we regressed sensitivity against year of treatment using the \texttt{lm()} function in R} [@R2016].
 \new{This analysis also allows us to link particularly sensitive treatment years to changes in community composition.}
@@
 
-\new{We did find that ANPP became more sensitive to drought over time (Fig. 3C).
-The latter years of our experiment were relatively wet, meaning that the drought treatment could have the largest absolute reduction of soil moisture in those years.
-But our metric of sensitivity accounts for differences among years because ANPP differences is standardized by the difference in soil moisture within the year.
-Therefore, it is likely that the increased sensitivity to drought over time is due to the cumulative impact of drought on the dominant species.
-Ecosystem functional response may soon shift if sensitivity continues to increase.}
+\new{We found that ANPP became more sensitive to the drought over time (Fig.
+3C).
+In other words, relative to the effect of the drought shelters on soil moisture, ANPP was more reduced by the drought conditions later in the experiment than early on in the experiment.
+We interepret this increase in sensitivity as the effect of cumulative impacts of the drought on dominant species.
+Plants may not have shown a large response in terms of ANPP in the first years of the experiment if they could grow from stored carbohydrate reserves or if they abstained from flowering and reproduction.
+As the drought progressed, these same plants may have have started to reduce in size or die.
+Given the long lived perennial species at this site, this increase in sensitivity may indicate a larger future change in community composition and ecosystem functional response.
+}
 

### 2018-01-05 manuscript/usses_water_ms.tex
updates data plots with soilwat sections; adds vwc combo file for statistical modeling

--- 2017-12-10 manuscript/usses_water_ms.tex
@@
 ``irrigation'' treatment.
+\new{Thus, our experiment consisted of $n=14$ control plots, $n=8$ irrigation plots, and $n=8$ irrigation plots, for a total of 30 plots.}
 
@@
 moisture sensors, and within those plots, one or more of the sensors
-frequently failed to collect data. To remedy these problems, and to
-produce average daily soil moisture values for the ambient, drought, and
-irrigation treatments, we used a statistical model to predict the
-average treatment effects on soil moisture during the course of the
+frequently failed to collect data. To remedy these problems, we used a
+statistical model to produce average daily soil moisture values for the
+ambient, drought, and irrigation treatments during the course of the
 experiment.
@@
 \new{We first averaged the observed soil moisture readings for each day ($d$) and plot ($i$), $x_{i,d}$.
-Plots were organized in groups of three, with each group ($g$) containing consisting of a control, a drought, and an irrigation plot.
-Within each group of plots we standardized the irrigation and drought effects on soil moisture relative to the ambient soil moisture conditions in the control plot.
-Specifically we subtracted the ambient daily soil moisture measured in the control plot from the soil moisture measured within the drought and irrigation plots within each group and then divided by the standard deviation of daily soil moisture values measured in the control plot $\left( \Delta x_{g,d,\text{trt.}} = (x_{g,d,\text{trt.}} - x_{g,d,\text{control}})/\text{s.d.}(x_{g,d,\text{control}}) \right)$ where $\Delta x$ is the standardized treatment effect and $x_{g,d,\text{trt.}}$ is the raw soil moisture measure for plot group $g$ on day $d$ and treatment $\text{trt.}$.
+Experimental plots were organized in groups of two, with each group ($g$) containing a drought and irrigated plot.
+In addition a nearby area outside the drought or irrigated plots was monitored for local ambient soil moisture conditions.
+Within each group of plots we standardized the irrigation and drought effects on soil moisture relative to the ambient soil moisture conditions.
+Specifically we subtracted the ambient daily soil moisture from the soil moisture measured within the drought and irrigation plots within each group and then divided by the standard deviation of daily soil moisture values measured in the ambient conditions $\left( \Delta x_{g,d,\text{trt.}} = (x_{g,d,\text{trt.}} - x_{g,d,\text{ambient}})/\text{s.d.}(x_{g,d,\text{ambient}}) \right)$ where $\Delta x$ is the standardized treatment effect and $x_{g,d,\text{trt.}}$ is the raw soil moisture measure for plot group $g$ on day $d$ and treatment $\text{trt.}$.
 These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.}
@@
 (winter, spring, summer, fall), rainfall, and all two-way interactions.
-Rainy days were defined as any day in which precipitation was recorded
-and average temperature was above 3\(^{\circ}\)C. The day immediately
-following rainfall was also classified as rainy. We fit the model using
-the \texttt{lme4::lmer()} function \citep{Bates2015} in R \citep{R2016},
-with random effects for plot group and date. We weighted observations by
-the number of unique sensors or spot measurements that were taken in
-each plot on that day.
+Using the local daily weather station data, we recorded rainy days as
+any day with measureable precipitation or the day after such a day and
+with average temperatures above 3\(^{\circ}\)C (to exclude days with
+snowfall). We fit the model using the \texttt{lme4::lmer()} function
+\citep{Bates2015} in R \citep{R2016}, with random effects for plot group
+and date. We weighted observations by the number of unique sensors or
+spot measurements that were taken in each plot on that day.
 \new{We then used the model to predict the deviations from ambient conditions produced in the treated plots on each day of the experiment.  
-We added these predicted deviations to the average daily ambient soil moisture to generate predictions for daily soil moisture in all of the treated plots: $\bar{x}_{d, \text{trt.}} = \Delta \bar{x}_{d,\text{trt.}} + \bar{x}_{d,\text{control}}$, where $\bar{x}_{d,\text{trt.}}$ is the average predicted daily soil moisture in the treated plots and $\bar{x}_{d,\text{control}}$ is the daily ambient soil moisture averaged across all control plots.}
+We added these predicted deviations to the average daily ambient soil moisture to generate predictions for daily soil moisture in all of the treated plots: $\bar{x}_{d, \text{trt.}} = \Delta \bar{x}_{d,\text{trt.}} + \bar{x}_{d,\text{ambient}}$, where $\bar{x}_{d,\text{trt.}}$ is the average predicted daily soil moisture in the treated plots and $\bar{x}_{d,\text{ambient}}$ is the daily ambient soil moisture averaged across all control plots.}
 We could only predict soil moisture in the treated plots on days for
@@
 
-\new{Unfortunately, our data does not allow us to include the effect of time on ecosystem functional response in the model described above. 
+\new{Our data does not allow us to directly test whether the ecosystem functional response to precipitation may have changed over time in each of the treatments. 
 This is because we lack sufficient within-year and within-treatment variation of soil moisture (i.e., within a year and treatment each plot shares the same value of VWC). 
-Therefore, we conducted an independent analysis to determine if the sensitivity of ANPP to treatment changed over time.}
+However, we did use a separate analysis to test whether the the sensitivity of ANPP to treatment changed over time.}
 
-\new{We define `sensitivity' as $\frac{ \text{ANPP}_{\text{control}} - \text{ANPP}_{\text{treatment}} } { \text{VWC}_{\text{control}} - \text{VWC}_{\text{treatment}} }$, following}
+\new{We define `sensitivity' as $\frac{ \text{ANPP}_{\text{control}} - \text{ANPP}_{\text{treatment}} } { \text{VWC}_{\text{ambient}} - \text{VWC}_{\text{treatment}} }$, following}
 \citet{Wilcox2017}.
-\new{This metric of sensitivity measures the response of ANPP standardized by the magnitude of the treatment. 
-Because our treatment and control plots are not paired, we conducted this analysis by comparing each treatment plot to the mean of the control plots in each year. 
-Then, independently for each treatment, we regressed sensitivity against year of treatment using the \texttt{lm()} function in R}
+\new{This metric of sensitivity scales the difference between ANPP in the treated and control plots by the change in soil moisture in the treated plots. 
+Because our treated plots are not directly paired with individual control plots, we compare ANPP in each treated plot to the average ANPP of the control plots in each year.
+After calculating sensitivity for each drought and irrigation plot in each year, we regressed sensitivity against year of treatment using the \texttt{lm()} function in R}
 \citep{R2016}.
@@
   \hline
-2011 & adonis &  21 &   2 & 1.02 & 0.45 \\ 
-  2011 & betadisper &  21 &   2 & 2.22 & 0.14 \\ 
-  2012 & adonis &  22 &   2 & 1.09 & 0.34 \\ 
-  2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
-  2013 & adonis &  22 &   2 & 1.25 & 0.13 \\ 
-  2013 & betadisper &  22 &   2 & 0.44 & 0.65 \\ 
-  2014 & adonis &  22 &   2 & 0.95 & 0.56 \\ 
-  2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
-  2015 & adonis &  21 &   2 & 1.05 & 0.39 \\ 
-  2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 
-  2016 & adonis &  21 &   2 & 1.07 & 0.36 \\ 
-  2016 & betadisper &  21 &   2 & 0.49 & 0.62 \\ 
+2011 & adonis &  30 &   2 & 1.92 & 0.01 \\ 
+  2011 & betadisper &  30 &   2 & 0.64 & 0.53 \\ 
+  2012 & adonis &  30 &   2 & 1.40 & 0.06 \\ 
+  2012 & betadisper &  30 &   2 & 0.30 & 0.75 \\ 
+  2013 & adonis &  30 &   2 & 1.70 & 0.01 \\ 
+  2013 & betadisper &  30 &   2 & 0.59 & 0.56 \\ 
+  2014 & adonis &  30 &   2 & 1.47 & 0.06 \\ 
+  2014 & betadisper &  30 &   2 & 2.23 & 0.13 \\ 
+  2015 & adonis &  30 &   2 & 1.22 & 0.14 \\ 
+  2015 & betadisper &  30 &   2 & 1.78 & 0.19 \\ 
+  2016 & adonis &  30 &   2 & 1.83 & 0.00 \\ 
+  2016 & betadisper &  30 &   2 & 0.24 & 0.79 \\ 
    \hline

### 2018-01-05 code/02_plot_data.R

@@ -82,15 +82,17 @@ soil_moisture <- read.csv("../data/soil_moisture_data/average_seasonal_soil_mois
   ungroup() %>%
   left_join(soilwat, by = c("year","Treatment","julian_date")) %>%
   mutate(VWC_combo = ifelse(is.na(VWC)==TRUE, VWC_raw, VWC),
-         VWC_source = ifelse(is.na(VWC)==TRUE, "soilwat", "observed"))
+         VWC_source = ifelse(is.na(VWC)==TRUE, "soilwat", "observed")) %>%
+  filter(month %in% c("03","04","05","06"))
 
 
-
+soilwat_inserts <- soil_moisture %>%
+  filter(is.na(VWC) == TRUE)
 
 suppressWarnings(# ignore warnings about missing values, we know they are empty
   soil_vwc <- ggplot(soil_moisture, aes(x=julian_date, y=VWC, group=Treatment, color=Treatment))+
-    geom_line(size=0.5, linetype=2, aes(x=julian_date, y=VWC_raw, group=Treatment, color=Treatment))+
-    geom_line(size=0.5)+
+    geom_line(data = soilwat_inserts, size=0.3, linetype=2, aes(x=julian_date, y=VWC_raw, group=Treatment, color=Treatment))+
+    geom_line(size=0.3)+
     scale_color_manual(values = mycols, name="Treatment")+
     ylab(expression(paste("Daily Soil VWC (ml ", ml^-1,")")))+
     xlab("Julian Day")+
@@ -103,6 +105,11 @@ suppressWarnings(# ignore warnings about missing values, we know they are empty
     guides(color=FALSE, linetype = FALSE)
 )
 
+##  Save VWC for statistical modeling that matches the above figure
+vwc_for_stats <- soil_moisture %>%
+  select(year, month, day, Treatment, VWC_combo, VWC_source) %>%
+  rename(VWC = VWC_combo)
+saveRDS(vwc_for_stats, "../data/soil_moisture_data/vwc_for_stats.RDS")
 
 
 ####

### 2018-01-05 manuscript/usses_water_ms.Rmd
big update with plug in VWC values from SOILWAT

@@
 The plant community is dominated by the shrub \emph{Artemesia tripartita}, \new{the perennial forb \emph{Balsamorhiza sagittata}}, and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata} \new{(see Appendix 1 for rank abundance curves)}.
-During the period of our experiment (2011 -- 2015), average mean annual precipitation was 265 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
+During the period of our experiment (2011 -- 2016), mean annual precipitation was 250 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
 Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
@@
 Ambient precipitation was variable over the five years of the study (Fig. 2A), meaning our experiment captured a range of environmental conditions.
-ANPP varied from a minimum of 74.5 g m$^{-2}$ in 2014 to a maximum of 237.1 g m$^{-2}$ in 2016 when averaged across treatments (Fig. 2C).
-ANPP was slightly higher in irrigation plots \new{(on average 17\% higher)} and slightly lower in drought plots \new{(on average 23\% lower) relative to control plots} (Fig.
-2C), corresponding to estimated soil volumetric water content (VWC) differences among treatments (Fig.
+ANPP varied from a minimum of 77.4 g m$^{-2}$ in 2014 to a maximum of 239.3 g m$^{-2}$ in 2016 when averaged across treatments (Fig. 2C).
+ANPP was slightly higher in irrigation plots \new{(on average 15\% higher)} and slightly lower in drought plots \new{(on average 25\% lower) relative to control plots} (Fig.
+2C), corresponding to observed and estimated soil volumetric water content (VWC) differences among treatments (Fig.
 2B).
@@
 
-\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3C). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0002; Fig. 3C)}.
+\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3C). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0007; Fig. 3C)}.
 

### 2018-01-05 manuscript/usses_water_ms.tex
big update with plug in VWC values from SOILWAT

@@
 \new{(see Appendix 1 for rank abundance curves)}. During the period of
-our experiment (2011 -- 2015), average mean annual precipitation was 265
-mm year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
+our experiment (2011 -- 2016), mean annual precipitation was 250 mm
+year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
 -5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July. Between 1926
@@
 (Fig. 2A), meaning our experiment captured a range of environmental
-conditions. ANPP varied from a minimum of 74.5 g m\(^{-2}\) in 2014 to a
-maximum of 237.1 g m\(^{-2}\) in 2016 when averaged across treatments
+conditions. ANPP varied from a minimum of 77.4 g m\(^{-2}\) in 2014 to a
+maximum of 239.3 g m\(^{-2}\) in 2016 when averaged across treatments
 (Fig. 2C). ANPP was slightly higher in irrigation plots
-\new{(on average 17\% higher)} and slightly lower in drought plots
-\new{(on average 23\% lower) relative to control plots} (Fig. 2C),
-corresponding to estimated soil volumetric water content (VWC)
-differences among treatments (Fig. 2B).
+\new{(on average 15\% higher)} and slightly lower in drought plots
+\new{(on average 25\% lower) relative to control plots} (Fig. 2C),
+corresponding to observed and estimated soil volumetric water content
+(VWC) differences among treatments (Fig. 2B).
 \new{VWC in drought plots was 12\% less than in control plots on average, and VWC in irrigation plots was 19\% higher than in control plots on average.}
@@
   \hline
-Intercept & Control & 0.16 & 0.14 & -1.22 & 1.69 \\ 
-  Intercept & Drought & 0.13 & 0.05 & -1.69 & 2.39 \\ 
-  Intercept & Irrigation & -0.24 & -0.18 & -2.11 & 1.53 \\ 
-  Slope & Control & 0.94 & 0.86 & -0.51 & 2.69 \\ 
-  Slope & Drought & 0.65 & 0.57 & -0.99 & 2.62 \\ 
-  Slope & Irrigation & 0.72 & 0.65 & -0.54 & 2.26 \\ 
+Intercept & Control & 0.14 & 0.14 & -1.24 & 1.50 \\ 
+  Intercept & Drought & 0.67 & 0.65 & -2.01 & 3.36 \\ 
+  Intercept & Irrigation & -0.39 & -0.34 & -3.51 & 2.48 \\ 
+  Slope & Control & 0.53 & 0.52 & -1.44 & 2.66 \\ 
+  Slope & Drought & 1.02 & 1.00 & -1.04 & 3.25 \\ 
+  Slope & Irrigation & 0.56 & 0.55 & -1.43 & 2.71 \\ 
    \hline
@@
   \hline
-2011 & adonis &  30 &   2 & 1.92 & 0.01 \\ 
-  2011 & betadisper &  30 &   2 & 0.64 & 0.53 \\ 
-  2012 & adonis &  30 &   2 & 1.40 & 0.06 \\ 
-  2012 & betadisper &  30 &   2 & 0.30 & 0.75 \\ 
-  2013 & adonis &  30 &   2 & 1.70 & 0.01 \\ 
-  2013 & betadisper &  30 &   2 & 0.59 & 0.56 \\ 
-  2014 & adonis &  30 &   2 & 1.47 & 0.06 \\ 
-  2014 & betadisper &  30 &   2 & 2.23 & 0.13 \\ 
-  2015 & adonis &  30 &   2 & 1.22 & 0.14 \\ 
-  2015 & betadisper &  30 &   2 & 1.78 & 0.19 \\ 
-  2016 & adonis &  30 &   2 & 1.83 & 0.00 \\ 
-  2016 & betadisper &  30 &   2 & 0.24 & 0.79 \\ 
+2011 & adonis &  22 &   2 & 1.04 & 0.41 \\ 
+  2011 & betadisper &  22 &   2 & 2.22 & 0.14 \\ 
+  2012 & adonis &  22 &   2 & 1.09 & 0.32 \\ 
+  2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
+  2013 & adonis &  22 &   2 & 1.26 & 0.14 \\ 
+  2013 & betadisper &  22 &   2 & 0.43 & 0.66 \\ 
+  2014 & adonis &  22 &   2 & 0.96 & 0.52 \\ 
+  2014 & betadisper &  22 &   2 & 0.36 & 0.70 \\ 
+  2015 & adonis &  22 &   2 & 1.08 & 0.31 \\ 
+  2015 & betadisper &  22 &   2 & 1.96 & 0.17 \\ 
+  2016 & adonis &  22 &   2 & 1.12 & 0.27 \\ 
+  2016 & betadisper &  22 &   2 & 0.27 & 0.76 \\ 
    \hline

### 2018-01-05 code/01_calibrate_radiometer_by_year.R

@@ -1,9 +1,12 @@
+################################################################################
 ##  calibrate_radiometer_by_year.R: script to fit regressions between biomass
 ##  and NDVI for Idaho plots. Regressions are fit by year, and are then used
 ##  to estimate biomass from NDVI for the permanent plots.
 ##
+##  ----------------------------------------------------------------------------
 ##  Author: Andrew Tredennick
 ##  Email:  atredenn@gmail.com
+################################################################################
 
 ##  Clear everything
 rm(list=ls(all.names = TRUE))
@@ -11,14 +14,14 @@ rm(list=ls(all.names = TRUE))
 ##  Set path to source file location
 setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only for RStudio
 
-#### SCALING FACTORS ###########################################################
-# recent years (2014 and beyond), biomass in 0.5m^2 plots (est biomass * 2)
-# earlier years (2013 and earlier), biomass in 0.25m^2 (est biomass * 4)
+### SCALING FACTORS
+# Recent years (2014 and beyond), biomass in 0.5m^2 plots (est biomass * 2)
+# Earlier years (2013 and earlier), biomass in 0.25m^2 (est biomass * 4)
 
 
 
 ####
-####  FILE PATHS, OUTPUT PATHS, ETC. ----
+####  FILE PATHS, OUTPUT PATHS, ETC. -------------------------------------------
 ####
 data_cal_path  <- "../data/radiometer/calibration/"
 data_quad_path <- "../data/radiometer/permanent_plots/"
@@ -27,7 +30,7 @@ out_path       <- "../data/estimated_biomass/"
 
 
 ####
-####  PULL SOME PACKAGES OFF THE SHELF ----
+####  PULL SOME PACKAGES OFF THE SHELF -----------------------------------------
 ####
 library(tidyverse) # all sorts of data wrangling
 library(dplyr)     # data summarizing/wrangling tools
@@ -37,7 +40,7 @@ library(stringr)   # tidy tools for strings
 
 
 ####
-####  FUNCTIONS TO FIT AND OUTPUT CALIBRATION REGRESSSION ----
+####  FUNCTIONS TO FIT AND OUTPUT CALIBRATION REGRESSSION ----------------------
 ####
 # NDVI calculations include "sens_*" scaling factors because we did not
[…]

### 2018-01-05 code/02_plot_data.R

@@ -78,7 +78,7 @@ soil_moisture <- read.csv("../data/soil_moisture_data/average_seasonal_soil_mois
   separate(simple_date, c("year","month","day")) %>%
   mutate(year = as.integer(year),
          Treatment = as.character(Treatment)) %>%
-  filter(type=="observed") %>%
+  # filter(type=="predicted") %>%
   ungroup() %>%
   left_join(soilwat, by = c("year","Treatment","julian_date")) %>%
   mutate(VWC_combo = ifelse(is.na(VWC)==TRUE, VWC_raw, VWC),
@@ -86,12 +86,12 @@ soil_moisture <- read.csv("../data/soil_moisture_data/average_seasonal_soil_mois
   filter(month %in% c("03","04","05","06"))
 
 
-soilwat_inserts <- soil_moisture %>%
-  filter(is.na(VWC) == TRUE)
+# soilwat_inserts <- soil_moisture %>%
+#   filter(is.na(VWC) == TRUE)
 
 suppressWarnings(# ignore warnings about missing values, we know they are empty
-  soil_vwc <- ggplot(soil_moisture, aes(x=julian_date, y=VWC, group=Treatment, color=Treatment))+
-    geom_line(data = soilwat_inserts, size=0.3, linetype=2, aes(x=julian_date, y=VWC_raw, group=Treatment, color=Treatment))+
+  soil_vwc <- ggplot(filter(soil_moisture, type == "observed"), aes(x=julian_date, y=VWC, group=Treatment, color=Treatment))+
+    geom_line(data = filter(soil_moisture, type == "predicted"), size=0.3, linetype=2, aes(x=julian_date, y=VWC, group=Treatment, color=Treatment))+
     geom_line(size=0.3)+
     scale_color_manual(values = mycols, name="Treatment")+
     ylab(expression(paste("Daily Soil VWC (ml ", ml^-1,")")))+
@@ -106,10 +106,10 @@ suppressWarnings(# ignore warnings about missing values, we know they are empty
 )
 
 ##  Save VWC for statistical modeling that matches the above figure
-vwc_for_stats <- soil_moisture %>%
-  select(year, month, day, Treatment, VWC_combo, VWC_source) %>%
-  rename(VWC = VWC_combo)
-saveRDS(vwc_for_stats, "../data/soil_moisture_data/vwc_for_stats.RDS")
[…]

### 2018-01-05 code/03_anpp_randcoefs_model.R

@@ -1,12 +1,15 @@
+################################################################################
 ##  anpp_randcoefs_model.R: Script to run GLMM analysis to test for treatment 
-##  effects on the relationship between precipitation and ANPP.
+##  effects on the relationship between soil moisture and ANPP.
 ##
 ##  NOTE: Stan may issue a couple warnings after running the MCMC that, as
 ##  the messages state, can be safely ignored. Just rejects a couple proposals
 ##  that result in ill-formed covariance matrices.
 ##
-##  Author: Andrew Tredennick
+##  ----------------------------------------------------------------------------
+##  Author: Andrew Tredennick and Andrew Kleinhesselink
 ##  Date created: June 2, 2017
+################################################################################
 
 ##  Clear everything
 rm(list = ls(all.names = TRUE))
@@ -17,7 +20,7 @@ setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # only for RStudio
 
 
 ####
-####  LOAD LIBRARIES ----
+####  LOAD LIBRARIES -----------------------------------------------------------
 ####
 library(tidyverse)    # Data munging
 library(dplyr)        # Data summarizing
@@ -29,21 +32,15 @@ library(lme4)         # Mixed-effects modeling
 
 
 ####
-####  READ IN AND EXTRACT EXPERIMENT ANPP DATA ----
+####  READ IN AND EXTRACT EXPERIMENT ANPP DATA ---------------------------------
 ####
 source("read_format_data.R")
 
-my_df <- anpp_data
-my_df$y <- as.numeric(scale( log( anpp_data$anpp ) ))
-m1 <- lmer( y ~ Treatment*vwc_scaled + (1|year_id) + (vwc_scaled|quadname), my_df)
-summary(m1)
-
-attr(VarCorr(m1)$quadname, 'correlation')
 
 
 
 ####
-####  SET UP AND FIT MODEL IN STAN ----
+####  SET UP AND FIT MODEL IN STAN ---------------------------------------------
 ####
 ##  Set up fixed and random design matrices
 lmod <- lm(log(anpp) ~ Treatment*vwc_scaled, anpp_data)
@@ -83,9 +80,8 @@ fit <- sampling(sm, data = anppdat, iter = 10000, chains = 4, thin = 10)
[…]

### 2018-01-05 code/04_sensitivity_analysis.R

@@ -1,12 +1,11 @@
-##  anpp_randcoefs_model.R: Script to run GLMM analysis to test for treatment 
-##  effects on the relationship between precipitation and ANPP.
+################################################################################
+##  sensitivity_analysis.R: R script to conduct a simple sensitivity analysis
+##  a la Wilcox et al. 2017, Global Change Biology.
 ##
-##  NOTE: Stan may issue a couple warnings after running the MCMC that, as
-##  the messages state, can be safely ignored. Just rejects a couple proposals
-##  that result in ill-formed covariance matrices.
-##
+##  ----------------------------------------------------------------------------
 ##  Author: Andrew Tredennick
-##  Date created: November
+##  Date created: November 2, 2017
+################################################################################
 
 ##  Clear everything
 rm(list = ls(all.names = TRUE))
@@ -28,37 +27,16 @@ library(lme4)         # Mixed-effects modeling
 
 
 ####
-####  READ IN AND EXTRACT EXPERIMENT ANPP DATA ----
+####  READ IN AND EXTRACT EXPERIMENT ANPP DATA ---------------------------------
 ####
 source("read_format_data.R")
 
-##  By treatment
-sens_anpp <- anpp_data %>%
-  group_by(Treatment, year) %>%
-  summarise(avg_anpp = mean(anpp)) %>%
-  spread(Treatment, avg_anpp) %>%
-  mutate(control_drought = Control - Drought,
-         control_irrigation = Control - Irrigation) %>%
-  dplyr::select(-Control,-Drought,-Irrigation) %>%
-  gather(comparison, anpp_diff, -year)
-
-sens_vwc <- anpp_data %>%
-  group_by(Treatment, year) %>%
-  summarise(vwc = mean(total_seasonal_vwc)) %>%
-  spread(Treatment, vwc) %>%
-  mutate(control_drought = Control - Drought,
-         control_irrigation = Control - Irrigation) %>%
-  dplyr::select(-Control,-Drought,-Irrigation) %>%
-  gather(comparison, vwc_diff, -year)
-
-sensitivity <- left_join(sens_anpp, sens_vwc, by = c("year", "comparison")) %>%
-  mutate(sens = anpp_diff / vwc_diff)
-
-# ggplot(sensitivity, aes(x=year, y=sens, color=comparison))+
-#   geom_line()
 
 
[…]

### 2018-01-05 code/05_plot_model_results.R

@@ -70,9 +70,9 @@ treat_slopes <- betas %>%
 slope_probs <- treat_slopes %>%
   group_by(Treatment, Type) %>%
   summarise(probs = round(get_one_tailed(estimate),2)) %>%
-  mutate(prob_text = paste("Pr < 0 =", probs))
-#slope_probs$prob_text[1] <- paste("Pr > 0 =", slope_probs$probs[1]) # switch direction of drought intercpet since greater than 0
-slope_probs$xpos <- c(-2.8, -0.7, -2.8, -0.7)
+  mutate(prob_text = paste("Pr > 0 =", probs))
+slope_probs$prob_text[3] <- paste("Pr < 0 =", slope_probs$probs[3]) # switch direction of drought intercpet since less than 0
+slope_probs$xpos <- c(-4.5, -0.32, -4.5, -0.27)
 slope_probs$ypos <- c(1.3,1.3,2.55,2.7)
 
 treat_cols <- mycols[2:3]

### 2018-01-05 code/06_community_dynamics.R

@@ -45,7 +45,7 @@ suppressWarnings( # ignore factor to character warning. It's OK.
     summarise(total_area = sum(area)) %>%
     left_join(plot_info, by = "quad") %>%
     filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
-    # filter(!str_detect(QuadName, 'P1|P7')) %>%
+    filter(!str_detect(QuadName, 'P1|P7')) %>% # exclude control plots in pastures different from treatment plots
     group_by(year, species, Treatment) %>%
     summarise(avg_area = mean(total_area)) %>%
     ungroup()
@@ -64,7 +64,7 @@ suppressWarnings( # ignore factor to character warning. It's OK.
     summarise(total_area = sum(area)) %>%
     left_join(plot_info, by = "quad") %>%
     filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
-    # filter(!str_detect(QuadName, 'P1|P7')) %>%
+    filter(!str_detect(QuadName, 'P1|P7')) %>% # exclude control plots in pastures different from treatment plots
     filter(year > 2010 & is.na(species)==F) %>%
     spread(species,total_area, fill = 0) %>%
     select(-QuadName,-Grazing,-paddock,-Group) %>%
@@ -78,7 +78,7 @@ suppressWarnings( # ignore factor to character warning. It's OK.
     summarise(total_inds = sum(ind_num)) %>%
     left_join(plot_info, by = "quad") %>%
     filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
-    # filter(!str_detect(QuadName, 'P1|P7')) %>%
+    filter(!str_detect(QuadName, 'P1|P7')) %>% # exclude control plots in pastures different from treatment plots
     filter(year > 2010 & is.na(species)==F) %>%
     spread(species,total_inds, fill = 0) %>%
     select(-QuadName,-Grazing,-paddock,-Group) %>%
@@ -99,7 +99,7 @@ suppressWarnings( # ignore factor to character warning. It's OK.
     select(-QuadName.y) %>%
     rename(QuadName = QuadName.x) %>%
     filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
-    # filter(!str_detect(QuadName, 'P1|P7')) %>%
[…]

### 2018-01-05 code/07_ndvi_randcoefs_model.R

@@ -33,12 +33,12 @@ library(lme4)         # Mixed-effects modeling
 ####
 source("read_format_data.R")
 
-my_df <- anpp_data
-my_df$y <- as.numeric(scale( log( anpp_data$ndvi ) ))
-m1 <- lmer( y ~ Treatment*vwc_scaled + (1|year_id) + (vwc_scaled|quadname), my_df)
-summary(m1)
-
-attr(VarCorr(m1)$quadname, 'correlation')
+# my_df <- anpp_data
+# my_df$y <- as.numeric(scale( log( anpp_data$ndvi ) ))
+# m1 <- lmer( y ~ Treatment*vwc_scaled + (1|year_id) + (vwc_scaled|quadname), my_df)
+# summary(m1)
+# 
+# attr(VarCorr(m1)$quadname, 'correlation')
 
 
 

### 2018-01-05 code/read_format_data.R

@@ -1,66 +1,102 @@
-##  Script to read in and format data for use in all analysis scripts. Subsets
-##  out some plots and split data into irrigation and drought sets.
+################################################################################
+##  Script to read in and format data for use in all analysis scripts. The raw
+##  data are:
+##    1. permanent_plots_estimated_biomass.RDS -- these estimates are produced by
+##       the R script 01_calibrate_radiometer_by_year.R, and represent estimated
+##       biomass (annual net primary productivity, ANPP) in each plot in each
+##       year.
+##    2. ClimateIPM.csv -- this is aggregated daily climate data from the MET
+##       station in Dubois, ID, near the USSES. We only use the variable 'ppt1',
+##       which is Fall-Spring precipitation for year t. E.g., if the year is 2012,
+##       then ppt1 is Fall 2011 precip through Spring 2012 precip, cumulative.
+##    3. SOILWAT_treatment_years.RDS -- these are model output from the SoilWat
+##       model using parameters and precipitation data from the USSES. The estimates
+##       are on daily time steps for the years of our treatment. We use these
+##       estimates to fill in gaps where we lack observed soil moisture data.
+##    4. average_seasonal_soil_moisture.csv -- This file contains observed and
+##       statistically modeled daily soil moisture. See the manuscript file for
+##       a description of our soil moisture data, the statistical model, and
+##       our limited use of SoilWat estimates.
 ##
+##  The main steps in this script are to summarize the soil moisture data into
+##  yearly values and merge together the soil moisture and ANPP data for
+##  statistical modeling.
+##
+## -----------------------------------------------------------------------------
 ##  Author: Andrew Tredennick
 ##  Date created: May 25, 2017
+################################################################################
 
-
-# ####
-# ####  LOAD LIBRARIES ----
-# ####
-# library(tidyverse)    # Data munging
-# library(dplyr)        # Data summarizing
[…]

### 2018-01-08 manuscript/usses_water_ms.tex
updates sensitivity p value
making the table

@@
 
-\subsection{Precipitation Experiment}\label{precipitation-experiment}
+\hypertarget{precipitation-experiment}{%
+\subsection{Precipitation Experiment}\label{precipitation-experiment}}
 
@@
   \hline
-Intercept & Control & 0.14 & 0.14 & -1.24 & 1.50 \\ 
-  Intercept & Drought & 0.67 & 0.65 & -2.01 & 3.36 \\ 
-  Intercept & Irrigation & -0.39 & -0.34 & -3.51 & 2.48 \\ 
-  Slope & Control & 0.53 & 0.52 & -1.44 & 2.66 \\ 
-  Slope & Drought & 1.02 & 1.00 & -1.04 & 3.25 \\ 
-  Slope & Irrigation & 0.56 & 0.55 & -1.43 & 2.71 \\ 
+Intercept & Control & 0.11 & 0.10 & -1.00 & 1.13 \\ 
+  Intercept & Drought & 0.06 & 0.08 & -1.42 & 1.39 \\ 
+  Intercept & Irrigation & -0.22 & -0.25 & -1.70 & 1.40 \\ 
+  Slope & Control & 0.71 & 0.75 & -0.65 & 1.87 \\ 
+  Slope & Drought & 0.72 & 0.77 & -0.64 & 1.93 \\ 
+  Slope & Irrigation & 0.67 & 0.72 & -0.68 & 1.84 \\ 
    \hline
@@
 Each point represents the sensitivity of ANPP in a plot relative to the mean of the control plots in that year.
-Only the significant regression for the drought treatment is plotted (\emph{P} = 0.0002).}}
+Only the significant regression for the drought treatment is plotted (\emph{P} = 0.0007).}}
 \end{figure}

### 2018-01-08 manuscript/usses_water_ms.tex
making table

@@
   \hline
-Intercept & Control & 0.11 & 0.10 & -1.00 & 1.13 \\ 
-  Intercept & Drought & 0.06 & 0.08 & -1.42 & 1.39 \\ 
-  Intercept & Irrigation & -0.22 & -0.25 & -1.70 & 1.40 \\ 
-  Slope & Control & 0.71 & 0.75 & -0.65 & 1.87 \\ 
-  Slope & Drought & 0.72 & 0.77 & -0.64 & 1.93 \\ 
-  Slope & Irrigation & 0.67 & 0.72 & -0.68 & 1.84 \\ 
+Intercept & Control & 0.14 & 0.14 & -1.24 & 1.50 \\ 
+  Intercept & Drought & 0.67 & 0.65 & -2.01 & 3.36 \\ 
+  Intercept & Irrigation & -0.39 & -0.34 & -3.51 & 2.48 \\ 
+  Slope & Control & 0.53 & 0.52 & -1.44 & 2.66 \\ 
+  Slope & Drought & 1.02 & 1.00 & -1.04 & 3.25 \\ 
+  Slope & Irrigation & 0.56 & 0.55 & -1.43 & 2.71 \\ 
    \hline

### 2018-01-09 manuscript/usses_water_ms.Rmd
updates sensitivity p value
starts section on soilwat; breaks bib somehow

@@
   \usepackage{lipsum}
-bibliography: /Users/atredenn/Dropbox/Bibliography/usses_water.bib
+bibliography: usses_water.bib
 output: 
@@
 See @Kleinhesselink2017b for more details on our approach to estimating soil moisture.
+
+\new{Following the above procedure, we still lacked soil moisture data for March 2012 (observations did not start unitl April) and for a string of days in 2013 during which the soil moisture probes failed take readings.
+Therefore, we used a version of the SOILWAT soil moisture model} [@Sala1992a] \new{that has been specifically designed for semiarid shrublands and grasslands} [@Bradford2014].
+\new{The model was parameterized sing generic sagebrush stepped parameters and local soil texture, soil bulk density, and precipitation data} [@Kleinhesselink2017b].
+\new{We used the locally parameterized SOILWAT model generate daily soil moisture predictions for the duration of the experiment, but only used SOILWAT predictions where there were gaps in our data (Fig.
+Sx).}
 
@@
 We also thank Susan Durham for clarifying our thinking on the statistical analyses and Kevin Wilcox for helpful discussions on analyzing community composition data.
+John Bradford and Caitlin Andrews generously provided SOILWAT model output,
 \new{Two anonymous reviewers and Elsie Denton provided thoughtful reviews that improved our paper.}
@@
 
-# REFERENCES
+# References

### 2018-01-09 manuscript/usses_water_ms.tex
starts section on soilwat; breaks bib somehow

@@
 estimating soil moisture.
+
+\new{Following the above procedure, we still lacked soil moisture data for March 2012 (observations did not start unitl April) and for a string of days in 2013 during which the soil moisture probes failed take readings.
+Therefore, we used a version of the SOILWAT soil moisture model}
+\citep{Sala1992a}
+\new{that has been specifically designed for semiarid shrublands and grasslands}
+\citep{Bradford2014}.
+\new{The model was parameterized sing generic sagebrush stepped parameters and local soil texture, soil bulk density, and precipitation data}
+\citep{Kleinhesselink2017b}.
+\new{We used the locally parameterized SOILWAT model generate daily soil moisture predictions for the duration of the experiment, but only used SOILWAT predictions where there were gaps in our data (Fig.
+Sx).}
 
@@
 Kevin Wilcox for helpful discussions on analyzing community composition
-data.
+data. John Bradford and Caitlin Andrews generously provided SOILWAT
+model output,
 \new{Two anonymous reviewers and Elsie Denton provided thoughtful reviews that improved our paper.}
@@
 
-\bibliography{/Users/atredenn/Dropbox/Bibliography/usses\_water.bib}
+\bibliography{usses\_water.bib}
 

### 2018-01-09 manuscript/usses_water_ms.Rmd
fixing some typos

@@
 Therefore, we used a version of the SOILWAT soil moisture model} [@Sala1992a] \new{that has been specifically designed for semiarid shrublands and grasslands} [@Bradford2014].
-\new{The model was parameterized sing generic sagebrush stepped parameters and local soil texture, soil bulk density, and precipitation data} [@Kleinhesselink2017b].
-\new{We used the locally parameterized SOILWAT model generate daily soil moisture predictions for the duration of the experiment, but only used SOILWAT predictions where there were gaps in our data (Fig.
-Sx).}
+\new{The model was parameterized using generic sagebrush steppe parameters and local soil texture, soil bulk density, and precipitation data} [@Kleinhesselink2017b].
+\new{We used the locally parameterized SOILWAT model to generate daily soil moisture predictions for the duration of the experiment, but only used SOILWAT predictions where there were gaps in our data (Fig.
+2B).}
 

### 2018-01-09 manuscript/usses_water_ms.tex
fixing some typos

@@
 \citep{Bradford2014}.
-\new{The model was parameterized sing generic sagebrush stepped parameters and local soil texture, soil bulk density, and precipitation data}
+\new{The model was parameterized using generic sagebrush steppe parameters and local soil texture, soil bulk density, and precipitation data}
 \citep{Kleinhesselink2017b}.
-\new{We used the locally parameterized SOILWAT model generate daily soil moisture predictions for the duration of the experiment, but only used SOILWAT predictions where there were gaps in our data (Fig.
-Sx).}
+\new{We used the locally parameterized SOILWAT model to generate daily soil moisture predictions for the duration of the experiment, but only used SOILWAT predictions where there were gaps in our data (Fig.
+2B).}
 

### 2018-01-09 manuscript/usses_water_ms.Rmd
updates data panel fig with SOILWAT estimates

@@
   \caption{(A) Probability density of historical precipitation from 1926-2016, with the years of the experiment shown with arrows on the \emph{x}-axis.
-(B) \new{Estimated daily} soil volumetric water content (VWC) in each of the three treatments during the course of the experiment.
-(C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.
+(B) \new{Observed (solid line), estimated (dashed lines), and SOILWAT-generated (dotted lines) daily} soil volumetric water content (VWC) in each of the three treatments during the course of the experiment.
+\new{We used the estimates from our statistical model, with gaps filled in by SOILWAT predictions, to calculate cumulative VWC as used in our analysis of treatment effects on ecosystem functional response.} (C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.
 Colors in panels B and C identify the treatment, as specified in the legend of panel C.}

### 2018-01-09 code/02_plot_data.R

@@ -92,6 +92,7 @@ soil_moisture <- read.csv("../data/soil_moisture_data/average_seasonal_soil_mois
 suppressWarnings(# ignore warnings about missing values, we know they are empty
   soil_vwc <- ggplot(filter(soil_moisture, type == "observed"), aes(x=julian_date, y=VWC, group=Treatment, color=Treatment))+
     geom_line(data = filter(soil_moisture, type == "predicted"), size=0.3, linetype=2, aes(x=julian_date, y=VWC, group=Treatment, color=Treatment))+
+    geom_line(data = filter(soil_moisture, VWC_source == "soilwat"), size=0.3, linetype=3, aes(x=julian_date, y=VWC_combo, group=Treatment, color=Treatment))+
     geom_line(size=0.3)+
     scale_color_manual(values = mycols, name="Treatment")+
     ylab(expression(paste("Daily Soil VWC (ml ", ml^-1,")")))+

### 2018-01-09 manuscript/usses_water_ms.Rmd
citations working again

@@
   \usepackage{lipsum}
-bibliography: usses_water.bib
+bibliography: /Users/atredenn/Dropbox/Bibliography/usses_water.bib
 output: 

### 2018-01-09 manuscript/usses_water_ms.tex
citations working again

@@
 %
+
 \begin{abstract}
@@
 \new{Water availability is a major driver of annual net primary productivity (ANPP) in grassland ecosystems}
-\citep{Huxman2004, Hsu2012}.
+(Huxman et al. 2004; Hsu, Powell, and Adler 2012).
 \new{Therefore, projected changes in precipitation regimes, associated with global climate change, will impact grassland ecosystem functioning.
@@
 responses can then be used to infer how ANPP may change under future
-precipitation regimes \citep[e.g.,][]{Hsu2012}. A problem with this
-approach is that it requires extrapolation if future precipitation falls
-outside the historical range of variability
-\citep{Smith2011, Peters2012}. For example, the soil moisture-ANPP
+precipitation regimes (e.g., Hsu, Powell, and Adler 2012). A problem
+with this approach is that it requires extrapolation if future
+precipitation falls outside the historical range of variability (Smith
+2011; Peters et al. 2012). For example, the soil moisture-ANPP
 relationship may be linear within the historical range of interannual
 variation, but could saturate at higher levels of soil moisture.
-Saturating relationships are actually common
-\citep{Hsu2012, Gherardi2015a}, perhaps because other resources, like
-nitrogen, become more limiting in wet years than dry years. Failure to
-accurately estimate the curvature of the soil moisture-ANPP relationship
-will lead to over- or underprediction of ANNP under extreme
-precipitation \citep{Peters2012}.
+Saturating relationships are actually common (Hsu, Powell, and Adler
+2012; Laureano A Gherardi and Sala 2015), perhaps because other
+resources, like nitrogen, become more limiting in wet years than dry
+years. Failure to accurately estimate the curvature of the soil
+moisture-ANPP relationship will lead to over- or underprediction of ANNP
+under extreme precipitation (Peters et al. 2012).
 
@@
 response to water availability because different species have different
-physiological thresholds. \citet{Smith2009} introduced the `Hierarchical
-Response Framework' for understanding the interplay of community
-composition and ecosystem functioning in response to long-term shifts in
-resources. In the near term, ecosystem functioning such as ANPP will
-reflect the physiological responses of individual species to the
-manipulated resource level. For example, ANPP may decline under
+physiological thresholds. Smith, Knapp, and Collins (2009) introduced
+the `Hierarchical Response Framework' for understanding the interplay of
+community composition and ecosystem functioning in response to long-term
+shifts in resources. In the near term, ecosystem functioning such as
+ANPP will reflect the physiological responses of individual species to
+the manipulated resource level. For example, ANPP may decline under
 simulated drought because the initial community consisted of
-drought-intolerant species \citep{Hoover2014}. Ecosystem functioning may
-recover over longer time spans as new species colonize or initial
-species reorder in relative abundance. It is also possible that
-ecosystem functioning shifts to a new mean state, reflecting the suite
-of species in the new community \citep{Knapp2012}.
+drought-intolerant species (Hoover, Knapp, and Smith 2014). Ecosystem
+functioning may recover over longer time spans as new species colonize
+or initial species reorder in relative abundance. It is also possible
+that ecosystem functioning shifts to a new mean state, reflecting the
+suite of species in the new community (Knapp, Briggs, and Smith 2012).
 
@@
 offer a route to understanding how ecosystems will respond to resource
-levels that fall outside the historical range of variability
-\citep{Avolio2015, Gherardi2015, Knapp2017}. Altering the amount of
-precipitation over many years should provide insight into the time
-scales at which water-limited ecosystems respond to chronic resource
-alteration. We propose four alternative scenarios for the effect of
-precipitation manipulation on the ecosystem functional response to soil
-moisture based on the Hierarchical Response Framework (Fig. 1). We
-define `ecosystem functional response' as the relationship between
-available soil moisture and ANPP. The four scenarios are based on
-possible outcomes at the community (e.g., community composition) and
+levels that fall outside the historical range of variability (Avolio et
+al. 2015; Laureano A. Gherardi and Sala 2015; Knapp et al. 2017).
+Altering the amount of precipitation over many years should provide
+insight into the time scales at which water-limited ecosystems respond
+to chronic resource alteration. We propose four alternative scenarios
+for the effect of precipitation manipulation on the ecosystem functional
+response to soil moisture based on the Hierarchical Response Framework
+(Fig. 1). We define `ecosystem functional response' as the relationship
+between available soil moisture and ANPP. The four scenarios are based
+on possible outcomes at the community (e.g., community composition) and
 ecosystem (e.g., soil moisture-ANPP regression) levels.
@@
 All four outcomes are possible in any given ecosystem, but the time
-scales at which the different scenarios play out likely differ
-\citep{Smith2009, Wilcox2016, Knapp2017}. To determine these time
-scales, we need to amass information on how quickly ecosystem functional
-responses change in different ecosystems. We also need to understand
-whether changes at the ecosystem level are driven by community level
-changes or individual level responses.
+scales at which the different scenarios play out likely differ (Smith,
+Knapp, and Collins 2009; Wilcox et al. 2016; Knapp et al. 2017). To
+determine these time scales, we need to amass information on how quickly
+ecosystem functional responses change in different ecosystems. We also
+need to understand whether changes at the ecosystem level are driven by
+community level changes or individual level responses.
 
@@
 the amount of ambient precipitation by 50\%. To achieve this, we used a
-system of rain-out shelters and automatic irrigation
-\citep{Gherardi2013}. The rain-out shelters consisted of transparent
-acrylic shingles 1-1.5 m above the ground that covered an area of
-\(2.5\times2\) m. The shingles intercepted approximately 50\% of
-incoming rainfall, which was channeled into 75 liter containers.
-Captured rainfall was then pumped out of the containers and sprayed on
-to the adjacent irrigation plot via two suspended sprinklers. Pumping
-was triggered by float switches once water levels reached about 20
-liters. We disconnected the irrigation pumps each October and
-reconnected them each April. The rain-out shelters remained in place
-throughout the year.
+system of rain-out shelters and automatic irrigation (Gherardi and Sala
+2013). The rain-out shelters consisted of transparent acrylic shingles
+1-1.5 m above the ground that covered an area of \(2.5\times2\) m. The
+shingles intercepted approximately 50\% of incoming rainfall, which was
+channeled into 75 liter containers. Captured rainfall was then pumped
+out of the containers and sprayed on to the adjacent irrigation plot via
+two suspended sprinklers. Pumping was triggered by float switches once
+water levels reached about 20 liters. We disconnected the irrigation
+pumps each October and reconnected them each April. The rain-out
+shelters remained in place throughout the year.
 
@@
 snowfall). We fit the model using the \texttt{lme4::lmer()} function
-\citep{Bates2015} in R \citep{R2016}, with random effects for plot group
-and date. We weighted observations by the number of unique sensors or
-spot measurements that were taken in each plot on that day.
+(Bates et al. 2015) in R (R Core Team 2016), with random effects for
+plot group and date. We weighted observations by the number of unique
+sensors or spot measurements that were taken in each plot on that day.
 \new{We then used the model to predict the deviations from ambient conditions produced in the treated plots on each day of the experiment.  
@@
 were missing due to malfunction of the sensors or dataloggers. See
-\citet{Kleinhesselink2017b} for more details on our approach to
-estimating soil moisture.
+Kleinhesselink (2017) for more details on our approach to estimating
+soil moisture.
 
 \new{Following the above procedure, we still lacked soil moisture data for March 2012 (observations did not start unitl April) and for a string of days in 2013 during which the soil moisture probes failed take readings.
-Therefore, we used a version of the SOILWAT soil moisture model}
-\citep{Sala1992a}
+Therefore, we used a version of the SOILWAT soil moisture model} (Sala,
+Lauenroth, and Parton 1992)
 \new{that has been specifically designed for semiarid shrublands and grasslands}
-\citep{Bradford2014}.
+(Bradford, Schlaepfer, and Lauenroth 2014).
 \new{The model was parameterized using generic sagebrush steppe parameters and local soil texture, soil bulk density, and precipitation data}
-\citep{Kleinhesselink2017b}.
+(Kleinhesselink 2017).
 \new{We used the locally parameterized SOILWAT model to generate daily soil moisture predictions for the duration of the experiment, but only used SOILWAT predictions where there were gaps in our data (Fig.
@@
 We estimated aboveground net primary productivity (ANPP) using a
-radiometer to relate ground reflectance to plant biomass \citep[see][
-for a review]{Byrne2011}. We recorded ground reflectance at four
+radiometer to relate ground reflectance to plant biomass (see Byrne et
+al. 2011 for a review). We recorded ground reflectance at four
 wavelengths, two associated with red reflectance (626 nm and 652 nm) and
@@
 Species composition data came from two sources: yearly census maps for
-each plot made using a pantograph \citep{Hill1920} and yearly counts of
+each plot made using a pantograph (Hill 1920) and yearly counts of
 annual species in each plot.
@@
 Based on our own observations and previous work at our study site
-\citep{Blaisdell1958, Dalgleish2011, Adler2012}, we chose to use
-cumulative volumetric water content from March through June as our
-metric of soil moisture (hereafter referred to as `VWC'). To achieve
-this goal, we fit a generalized linear mixed effects regression model
-with log(ANPP) as the response variable and VWC and treatment as fixed
-effects. Plot and year of treatment were included as random effects to
-account for non-independence of observations, as described below. We
-log-transformed ANPP to account for heteroscedasticity. Both log(ANPP)
-and VWC were standardized to have mean 0 and unit variance before
-fitting the model {[}i.e., \((x_i - \bar{x})/\sigma_x\){]}.
+(Blaisdell 1958; Dalgleish et al. 2011; Adler, Dalgleish, and Ellner
+2012), we chose to use cumulative volumetric water content from March
+through June as our metric of soil moisture (hereafter referred to as
+`VWC'). To achieve this goal, we fit a generalized linear mixed effects
+regression model with log(ANPP) as the response variable and VWC and
+treatment as fixed effects. Plot and year of treatment were included as
+random effects to account for non-independence of observations, as
+described below. We log-transformed ANPP to account for
+heteroscedasticity. Both log(ANPP) and VWC were standardized to have
+mean 0 and unit variance before fitting the model {[}i.e.,
+\((x_i - \bar{x})/\sigma_x\){]}.
 
@@
 precipitation-ANPP relationships often saturate with increasing
-precipitation \citep{Hsu2012, Gherardi2015a}.
+precipitation (Hsu, Powell, and Adler 2012; Laureano A Gherardi and Sala
+2015).
 
@@
 \new{We used a vague prior distribution for each $\beta$: $\boldsymbol{\beta} \sim \text{Normal}(0,5)$.}
-A full description of our model and associated R \citep{R2016} code is
-in Appendix 3.
+A full description of our model and associated R (R Core Team 2016) code
+is in Appendix 3.
 
@@
 estimates of all unknowns via the No-U-Turn Hamiltonian Monte Carlo
-sampler in Stan \citep{stan2016}. We used the R package `rstan'
-\citep{rstan2016} to link R \citep{R2016} to Stan. We obtained samples
-from the posterior distribution for all model parameters from four
-parallel MCMC chains run for 10,000 iterations, saving every
-\(10^{\text{th}}\) sample. Trace plots of all parameters were visually
-inspected to ensure well-mixed chains and convergence. We also made sure
-all scale reduction factors (\(\hat{R}\) values) were less than 1.1
-\citep{Gelman2009}.
+sampler in Stan (Stan Development Team 2016b). We used the R package
+`rstan' (Stan Development Team 2016a) to link R (R Core Team 2016) to
+Stan. We obtained samples from the posterior distribution for all model
+parameters from four parallel MCMC chains run for 10,000 iterations,
+saving every \(10^{\text{th}}\) sample. Trace plots of all parameters
+were visually inspected to ensure well-mixed chains and convergence. We
+also made sure all scale reduction factors (\(\hat{R}\) values) were
+less than 1.1 (Gelman and Hill 2009).
 
@@
 \new{We define `sensitivity' as $\frac{ \text{ANPP}_{\text{control}} - \text{ANPP}_{\text{treatment}} } { \text{VWC}_{\text{ambient}} - \text{VWC}_{\text{treatment}} }$, following}
-\citet{Wilcox2017}.
+Wilcox et al. (2017).
 \new{This metric of sensitivity scales the difference between ANPP in the treated and control plots by the change in soil moisture in the treated plots. 
@@
 After calculating sensitivity for each drought and irrigation plot in each year, we regressed sensitivity against year of treatment using the \texttt{lm()} function in R}
-\citep{R2016}.
+(R Core Team 2016).
 \new{This analysis also allows us to link particularly sensitive treatment years to changes in community composition.}
@@
 if community composition overlapped, or not, among treatments in each
-year. We used the \texttt{vegan::metaMDS()} function \citep{Oksanen2016}
-to calculate Bray-Curtis distances and then to run the NMDS analysis. We
-used the \texttt{vegan::adonis()} function \citep{Oksanen2016} to
-perform permutational multivariate analysis of variance to test whether
+year. We used the \texttt{vegan::metaMDS()} function (Oksanen 2016) to
+calculate Bray-Curtis distances and then to run the NMDS analysis. We
+used the \texttt{vegan::adonis()} function (Oksanen 2016) to perform
+permutational multivariate analysis of variance to test whether
 treatment plots formed distinct groupings. To test whether treatment
 plots were equally dispersed, or not, we used the
-\texttt{vegan::betadisper()} function \citep{Oksanen2016}.
+\texttt{vegan::betadisper()} function (Oksanen 2016).
 
@@
 community composition shifts to favor species better able to take
-advantage of, or cope with, new resource levels \citep{Smith2009}.
-Previous work has shown that community compositional shifts can be both
-rapid, on the order of years \citep{Hoover2014}, and slow, on order of
-decades \citep{Knapp2012, Wilcox2016}. A lingering question is how the
-time scales of ecosystem response and community change vary among
-ecosystems. Precipitation manipulation experiments can help answer this
-question, especially if they push water availability outside the
-historical range of variability for long periods.
+advantage of, or cope with, new resource levels (Smith, Knapp, and
+Collins 2009). Previous work has shown that community compositional
+shifts can be both rapid, on the order of years (Hoover, Knapp, and
+Smith 2014), and slow, on order of decades (Knapp, Briggs, and Smith
+2012; Wilcox et al. 2016). A lingering question is how the time scales
+of ecosystem response and community change vary among ecosystems.
+Precipitation manipulation experiments can help answer this question,
+especially if they push water availability outside the historical range
+of variability for long periods.
 
@@
 generally, and sagebrush steppe specifically, are considered
-water-limited systems. \new{For example,} \citet{Huxman2004} and
-\citet{Knapp2015}
+water-limited systems. \new{For example,} Huxman et al. (2004) and Knapp
+et al. (2015)
 \new{showed that semi-arid sites are more sensitive to drought than mesic sites, and}
-\citet{Wilcox2017}
+Wilcox et al. (2017)
 \new{found that semi-arid sites are particularly sensitive to irrigation treatments.}
@@
 may not be abnormal given our site's historical range of variability
-\citep{Knapp2017}. We cannot definitively rule out this possibility, but
-we have reason to believe our treatments \emph{should} have been large
-enough. Using the methods described by \citet{Lemoine2016}, we
+(Knapp et al. 2017). We cannot definitively rule out this possibility,
+but we have reason to believe our treatments \emph{should} have been
+large enough. Using the methods described by Lemoine et al. (2016), we
 calculated the percent reduction and increase of mean growing season
@@
 only the cumulative March-June soil moisture covariate we included in
-our statistical model. \new{For example,} \citet{LaPierre2016}
+our statistical model. \new{For example,} La Pierre et al. (2016)
 \new{found that site-scale ANPP is better predicted by abiotic factors such as nutrient availability than precipitation.}
-Moreover, temperature can impact ANPP directly \citep{Epstein1997} and
-by exacerbating the effects of soil moisture \citep{DeBoeck2011}.
-Measurements of soil moisture likely contain a signal of temperature,
-through its impact on evaporation and infiltration, but the measurements
-will not capture the direct effect of temperature on metabolic and
-physiological processes. We also did not redistribute snow across our
-treatments in the winter, and snow melt may spur early spring growth.
-Failure to account for potentially important covariates could explain
-the within-year spread of ANPP (Fig 2C, Fig. 3B) and the resulting
-uncertain relationship we observed between soil moisture and ANPP across
-all treatments (Table 1, Fig. A3-2). Unfortunately, we did not collect
-data on additional factors because our focus was on precipitation.
+Moreover, temperature can impact ANPP directly (Epstein, Lauenroth, and
+Burke 1997) and by exacerbating the effects of soil moisture (De Boeck
+et al. 2011). Measurements of soil moisture likely contain a signal of
+temperature, through its impact on evaporation and infiltration, but the
+measurements will not capture the direct effect of temperature on
+metabolic and physiological processes. We also did not redistribute snow
+across our treatments in the winter, and snow melt may spur early spring
+growth. Failure to account for potentially important covariates could
+explain the within-year spread of ANPP (Fig 2C, Fig. 3B) and the
+resulting uncertain relationship we observed between soil moisture and
+ANPP across all treatments (Table 1, Fig. A3-2). Unfortunately, we did
+not collect data on additional factors because our focus was on
+precipitation.
 
@@
 hedging, where species forego short-term gains to reduce the variance of
-long-term success \citep{Seger1987}. In other words, species follow the
-same conservative strategy every year, designed to minimize response to
+long-term success (Seger 1987). In other words, species follow the same
+conservative strategy every year, designed to minimize response to
 environmental conditions. The dry and variable environment of the
@@
 the perennial grasses in our focal ecosystem avoid drought stress by
-growing early in the growing season \citep[A.R. Kleinhesselink, personal
-observation]{Blaisdell1958}. Furthermore, the dominant shrub in our
-focal ecosystem, \emph{Artemisia tripartita}, has access to water deep
-in the soil profile thanks to a deep root system
-\citep{Kulmatiski2017a}.
+growing early in the growing season (Blaisdell 1958, A.R.
+Kleinhesselink, personal observation). Furthermore, the dominant shrub
+in our focal ecosystem, \emph{Artemisia tripartita}, has access to water
+deep in the soil profile thanks to a deep root system (Kulmatiski et al.
+2017).
 \new{The dominance of our site by the shrub \emph{A. tripartita} (Appendix 1) may explain why our results do not conform to broader patterns of grassland sensitivity to precipitation manipulations}
-\citep[e.g.,][]{Huxman2004, Knapp2015, Wilcox2017}.
+(e.g., Huxman et al. 2004; Knapp et al. 2015; Wilcox et al. 2017).
 
@@
 precipitation alteration might reveal plant community shifts that we did
-not observe \citep[e.g.,][]{Wilcox2016},
+not observe (e.g., Wilcox et al. 2016),
 \new{and the increased sensitivity of ANPP to drought in the final years of our experiment may portend such a shift.}
@@
 
-\bibliography{usses\_water.bib}
+\hypertarget{references}{%
+\section*{REFERENCES}\label{references}}
+\addcontentsline{toc}{section}{REFERENCES}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Adler2012}{}%
+Adler, P B, H J Dalgleish, and S P Ellner. 2012. ``Forecasting plant
+community impacts of climate variability and change: when do competitive
+interactions matter?'' \emph{Journal of Ecology} 100:478--87.
+\url{https://doi.org/10.1111/j.1365-2745.2011.01930.x}.
+
+\leavevmode\hypertarget{ref-Avolio2015}{}%
+Avolio, Meghan L, Kimberly J La Pierre, Gregory R Houseman, Sally E
+Koerner, Emily Grman, Forest Isbell, David Samuel Johnson, and Kevin R
+Wilcox. 2015. ``A framework for quantifying the magnitude and
+variability of community responses to global change drivers.''
+\emph{Ecosphere} 6 (12):1--14.
+\url{https://doi.org/10.1890/ES15-00317.1}.
+
+\leavevmode\hypertarget{ref-Bates2015}{}%
+Bates, D., M. Maechler, B. Bolker, and S. Walker. 2015. ``Fitting linear
+mixed-effects models using lme4.''
+\url{https://doi.org/10.18637/jss.v067.i01}.
+
+\leavevmode\hypertarget{ref-Blaisdell1958}{}%
+Blaisdell, James P. 1958. ``Seasonal development and yield of native
+plants on the upper snake river plains and their relation to certain
+climate factors.'' \emph{United States Department of Agriculture
+Technical Bulletin No. 1190}.
+
+\leavevmode\hypertarget{ref-Bradford2014}{}%
+Bradford, John B., Daniel R. Schlaepfer, and William K. Lauenroth. 2014.
+``Ecohydrology of Adjacent Sagebrush and Lodgepole Pine Ecosystems: The
+Consequences of Climate Change and Disturbance.'' \emph{Ecosystems} 17
+(4):590--605. \url{https://doi.org/10.1007/s10021-013-9745-1}.
+
+\leavevmode\hypertarget{ref-Byrne2011}{}%
+Byrne, Kerry M., William K. Lauenroth, Peter B. Adler, and Christine M.
+Byrne. 2011. ``Estimating Aboveground Net Primary Production in
+Grasslands: A Comparison of Nondestructive Methods.'' \emph{Rangeland
+Ecology and Management} 64 (5):498--505.
+\url{https://doi.org/10.2111/REM-D-10-00145.1}.
+
+\leavevmode\hypertarget{ref-Dalgleish2011}{}%
+Dalgleish, Harmony J., David N. Koons, Mevin B. Hooten, Corey A. Moffet,
+and Peter B. Adler. 2011. ``Climate influences the demography of three
+dominant sagebrush steppe plants.'' \emph{Ecology} 92 (1):75--85.
+\url{https://doi.org/10.1890/10-0780.1}.
+
+\leavevmode\hypertarget{ref-DeBoeck2011}{}%
+De Boeck, Hans J., Freja E. Dreesen, Ivan A. Janssens, and Ivan Nijs.
+2011. ``Whole-system responses of experimental plant communities to
+climate extremes imposed in different seasons.'' \emph{New Phytologist}
+189 (3):806--17. \url{https://doi.org/10.1111/j.1469-8137.2010.03515.x}.
+
+\leavevmode\hypertarget{ref-Epstein1997}{}%
+Epstein, H. E., W. K. Lauenroth, and I. C. Burke. 1997. ``Effects of
+temperature and soil texture on ANPP in the U.S. Great plains.''
+\emph{Ecology} 78 (8):2628--31.
+\href{https://doi.org/10.1890/0012-9658(1997)078\%5B2628:EOTAST\%5D2.0.CO;2}{https://doi.org/10.1890/0012-9658(1997)078{[}2628:EOTAST{]}2.0.CO;2}.
+
+\leavevmode\hypertarget{ref-Gelman2009}{}%
+Gelman, Andrew, and Jennifer Hill. 2009. \emph{Data analysis using
+regression and multilevel/hierarchical models}. Cambridge: Cambridge
+University Press.
+
+\leavevmode\hypertarget{ref-Gherardi2015a}{}%
+Gherardi, Laureano A, and Osvaldo E Sala. 2015. ``Enhanced precipitation
+variability decreases grass- and increases shrub-productivity.''
+\emph{Proceedings of the National Academy of Sciences} 112
+(41):12735--40. \url{https://doi.org/10.1073/pnas.1506433112}.
+
+\leavevmode\hypertarget{ref-Gherardi2013}{}%
+Gherardi, Laureano A., and Osvaldo E Sala. 2013. ``Automated rainfall
+manipulation system: a reliable and inexpensive tool for ecologists.''
+\emph{Ecosphere} 4 (2):1--10.
+\url{https://doi.org/10.1890/ES12-00371.1}.
+
+\leavevmode\hypertarget{ref-Gherardi2015}{}%
+Gherardi, Laureano A., and Osvaldo E. Sala. 2015. ``Enhanced interannual
+precipitation variability increases plant functional diversity that in
+turn ameliorates negative impact on productivity.'' \emph{Ecology
+Letters} 18 (12):1293--1300. \url{https://doi.org/10.1111/ele.12523}.
+
+\leavevmode\hypertarget{ref-Hill1920}{}%
+Hill, Robert R. 1920. ``Charting Quadrats with a Pantograph.''
+\emph{Ecology} 1 (4). Ecological Society of America:270--73.
+\url{https://doi.org/10.2307/1929561}.
+
+\leavevmode\hypertarget{ref-Hoover2014}{}%
+Hoover, David L., Alan K. Knapp, and Melinda D. Smith. 2014.
+``Resistance and resilience of a grassland ecosystem to climate
+extremes.'' \emph{Ecology} 95 (9):2646--56.
+\url{https://doi.org/10.1890/13-2186.1}.
+
+\leavevmode\hypertarget{ref-Hsu2012}{}%
+Hsu, Joanna S., James Powell, and Peter B. Adler. 2012. ``Sensitivity of
+mean annual primary production to precipitation.'' \emph{Global Change
+Biology} 18 (7):2246--55.
+\url{https://doi.org/10.1111/j.1365-2486.2012.02687.x}.
+
+\leavevmode\hypertarget{ref-Huxman2004}{}%
+Huxman, Travis E., Melinda D. Smith, Philip A. Fay, Alan K. Knapp, M.
+Rebecca Shaw, Michael E. Loik, Stanley D. Smith, et al. 2004.
+``Convergence across biomes to a common rain-use efficiency.''
+\emph{Nature} 429 (6992):651--54.
+\url{https://doi.org/10.1038/nature02561}.
+
+\leavevmode\hypertarget{ref-Kleinhesselink2017b}{}%
+Kleinhesselink, Andrew R. 2017. ``Direct and indirect effects of climate
+change on plant populations and communities in sagebrush steppe.''
+Dissertation, Utah State University.
+
+\leavevmode\hypertarget{ref-Knapp2017}{}%
+Knapp, Alan K, Meghan L Avolio, Claus Beier, Charles J W Carroll, Scott
+L Collins, Jeffrey S Dukes, Lauchlan H Fraser, et al. 2017. ``Pushing
+precipitation to the extremes in distributed experiments:
+recommendations for simulating wet and dry years.'' \emph{Global Change
+Biology} 23 (5):1774--82. \url{https://doi.org/10.1111/gcb.13504}.
+
+\leavevmode\hypertarget{ref-Knapp2012}{}%
+Knapp, Alan K, John M Briggs, and Melinda D Smith. 2012. ``Community
+stability does not preclude ecosystem sensitivity to chronic resource
+alteration.'' \emph{Functional Ecology} 26 (6):1231--3.
+\url{https://doi.org/10.1111/j.1365-2435.2012.02053.x}.
+
+\leavevmode\hypertarget{ref-Knapp2015}{}%
+Knapp, Alan K., Charles J.W. Carroll, Elsie M. Denton, Kimberly J. La
+Pierre, Scott L. Collins, and Melinda D. Smith. 2015. ``Differential
+sensitivity to regional-scale drought in six central US grasslands.''
+\emph{Oecologia} 177 (4):949--57.
+\url{https://doi.org/10.1007/s00442-015-3233-6}.
+
+\leavevmode\hypertarget{ref-Kulmatiski2017a}{}%
+Kulmatiski, Andrew, Peter B. Adler, John M. Stark, and Andrew T.
+Tredennick. 2017. ``Water and nitrogen uptake are better associated with
+resource availability than root biomass.'' \emph{Ecosphere} 8 (3).
+\url{https://doi.org/10.1002/ecs2.1738}.
+
+\leavevmode\hypertarget{ref-LaPierre2016}{}%
+La Pierre, Kimberly J., Dana M. Blumenthal, Cynthia S. Brown, Julia A.
+Klein, and Melinda D. Smith. 2016. ``Drivers of Variation in Aboveground
+Net Primary Productivity and Plant Community Composition Differ Across a
+Broad Precipitation Gradient.'' \emph{Ecosystems} 19 (3):521--33.
+\url{https://doi.org/10.1007/s10021-015-9949-7}.
+
+\leavevmode\hypertarget{ref-Lemoine2016}{}%
+Lemoine, Nathan P, Justin Sheffield, Jeffrey S Dukes, Alan K Knapp, and
+Melinda D Smith. 2016. ``Terrestrial Precipitation Analysis (TPA): A
+resource for characterizing long-term precipitation regimes and
+extremes.'' \emph{Methods in Ecology and Evolution} 7 (11):1396--1401.
+\url{https://doi.org/10.1111/2041-210X.12582}.
+
+\leavevmode\hypertarget{ref-Oksanen2016}{}%
+Oksanen, Jari. 2016. ``Vegan: ecological diversity.''
+\url{https://doi.org/10.1029/2006JF000545}.
+
+\leavevmode\hypertarget{ref-Peters2012}{}%
+Peters, Debra P C, Jin Yao, Osvaldo E. Sala, and John P. Anderson. 2012.
+``Directional climate change and potential reversal of desertification
+in arid and semiarid ecosystems.'' \emph{Global Change Biology} 18
+(1):151--63. \url{https://doi.org/10.1111/j.1365-2486.2011.02498.x}.
+
+\leavevmode\hypertarget{ref-R2016}{}%
+R Core Team. 2016. \emph{R: A Language and Environment for Statistical
+Computing}. Vienna, Austria.
+
+\leavevmode\hypertarget{ref-Sala1992a}{}%
+Sala, O. E., W. K. Lauenroth, and W. J. Parton. 1992. ``Long-term soil
+water dynamics in the shortgrass steppe.'' \emph{Ecology} 73
+(4):1175--81. \url{https://doi.org/10.2307/1940667}.
+
+\leavevmode\hypertarget{ref-Seger1987}{}%
+Seger, J. 1987. ``What is bet-hedging?'' In \emph{Oxford Surveys in
+Evolutionary Biology}, edited by P.H. Harvey and L. Partridge, 182--211.
+Oxford: Oxford University Press.
+\url{http://ci.nii.ac.jp/naid/10004816581/\%5Cnpapers2://publication/uuid/BE92F8E9-1B87-4E50-85C5-AA134A0EBAF0}.
+
+\leavevmode\hypertarget{ref-Smith2011}{}%
+Smith, M.D. 2011. ``An ecological perspective on extreme climatic
+events: A synthetic definition and framework to guide future research.''
+\emph{Journal of Ecology} 99 (3):656--63.
+\url{https://doi.org/10.1111/j.1365-2745.2011.01798.x}.
+
+\leavevmode\hypertarget{ref-Smith2009}{}%
+Smith, Melinda D., Alan K. Knapp, and Scott L. Collins. 2009. ``A
+framework for assessing ecosystem dynamics in response to chronic
+resource alterations induced by global change.'' \emph{Ecology} 90
+(12):3279--89. \url{https://doi.org/10.1890/08-1815.1}.
+
+\leavevmode\hypertarget{ref-rstan2016}{}%
+Stan Development Team. 2016a. ``Rstan: the R interface to Stan, Version
+2.14.1.'' \url{http://mc-stan.org/rstan.html}.
+
+\leavevmode\hypertarget{ref-stan2016}{}%
+---------. 2016b. ``Stan: A C++ Library for Probability and Sampling,
+Version 2.14.1.''
+
+\leavevmode\hypertarget{ref-Wilcox2016}{}%
+Wilcox, Kevin R, John M Blair, Melinda D Smith, and Alan K Knapp. 2016.
+``Does ecosystem sensitivity to precipitation at the site-level conform
+to regional-scale predictions?'' \emph{Ecology} 97 (3):561--68.
+\url{https://doi.org/10.1890/15-1437}.
+
+\leavevmode\hypertarget{ref-Wilcox2017}{}%
+Wilcox, Kevin R., Zheng Shi, Laureano A. Gherardi, Nathan P. Lemoine,
+Sally E. Koerner, David L. Hoover, Edward Bork, et al. 2017.
+``Asymmetric responses of primary productivity to precipitation
+extremes: A synthesis of grassland precipitation manipulation
+experiments.'' \emph{Global Change Biology} 23 (10):4376--85.
+\url{https://doi.org/10.1111/gcb.13706}.
 

### 2018-01-09 code/02_plot_data.R
adds NCDC weather data; updates data figure

@@ -32,32 +32,34 @@ library(cowplot)   # For combining ggplot objects
 ####  READ IN WEATHER DATA AND PLOT --------------------------------------------
 ####
 weather <- read.csv("../data/weather/ClimateIPM.csv")
-# weather <- read.csv("../data/weather/dubois_station_weather_01032018.csv") %>%
-#   dplyr::select(DATE, PRCP) %>%
-#   dplyr::rename("date" = DATE, "precip" = PRCP) %>%
-#   separate(date, into = c("year", "month", "day"), sep = "-") %>%
-#   mutate(precip = ifelse(is.na(precip), 0, precip)) %>% # set missing station data to 0
-#   group_by(year) %>%
-#   summarise(annual_precip = sum(precip)) 
+weather <- read.csv("../data/weather/dubois_station_weather_01092018.csv") %>%
+  dplyr::select(DATE, PRCP) %>%
+  dplyr::rename("date" = DATE, "precip" = PRCP) %>%
+  separate(date, into = c("year", "month", "day"), sep = "-") %>%
+  mutate(precip = ifelse(is.na(precip), 0, precip)) %>% # set missing station data to 0
+  mutate(year = as.numeric(year)) %>%
+  group_by(year) %>%
+  summarise(annual_precip = sum(precip))
 
 trt_data <- weather %>%
-  filter(year>2011) %>%
-  select(year, ppt1) %>%
+  filter(year > 2011) %>%
+  filter(year < 2017) %>%
+  select(year, annual_precip) %>%
   mutate(Treatment = "Control")
 
 ##  Adjust 2012 and 2015 values so they are distinguishable on the plot
-trt_data[which(trt_data$year==2015),"ppt1"] <- trt_data[which(trt_data$year==2015),"ppt1"]+10
-trt_data[which(trt_data$year==2012),"ppt1"] <- trt_data[which(trt_data$year==2012),"ppt1"]-5
+trt_data[which(trt_data$year==2015),"annual_precip"] <- trt_data[which(trt_data$year==2015),"annual_precip"]+8
+trt_data[which(trt_data$year==2014),"annual_precip"] <- trt_data[which(trt_data$year==2014),"annual_precip"]-8
 
-ppt_histogram <- ggplot(weather, aes(x=ppt1))+
+ppt_histogram <- ggplot(weather, aes(x=annual_precip))+
[…]

### 2018-01-09 manuscript/Appendix5.Rmd
updating ms text with new precip characterization

@@
 We chose the 1\% quantile to be indicative of extreme dry conditions (drought) and the 99\% quantile to be indicative of extreme wet conditions (irrigation).
-The data consist of 91 yearly records, which we assume are approximately normally distributed.
+The data consist of 91 yearly records, which we assume are approximately normally distributed for these purposes.
 The R code below shows our procedure, and Fig. A5-1 shows the results.
 
-```{r hrv, fig.cap="Density of the empirical distribution of growing season precipitation at Dubois, ID. Dashed vertical lines show the 1% and 99% quantiles, assuming a normal distribution."}
-library(ggplot2)
-weather    <- read.csv("../data/weather/ClimateIPM.csv")
-mean_ppt   <- mean(weather$ppt1)
-quants_ppt <- quantile(weather$ppt1,probs = c(0.01,0.99))
+```{r hrv, fig.cap="Density of the empirical distribution of growing season precipitation at Dubois, ID.
+Dashed vertical lines show the 1% and 99% quantiles, assuming a normal distribution.", message=FALSE}
+library(tidyverse)
+library(dplyr)
+
+weather <- read.csv("../data/weather/dubois_station_weather_01092018.csv") %>%
+  dplyr::select(DATE, PRCP) %>%
+  dplyr::rename("date" = DATE, "precip" = PRCP) %>%
+  separate(date, into = c("year", "month", "day"), sep = "-") %>%
+  mutate(precip = ifelse(is.na(precip), 0, precip)) %>% # set missing station data to 0
+  mutate(year = as.numeric(year)) %>%
+  group_by(year) %>%
+  summarise(annual_precip = sum(precip))
+
+mean_ppt   <- mean(weather$annual_precip)
+quants_ppt <- quantile(weather$annual_precip,probs = c(0.01,0.99))
 quants_ppt[1]/mean_ppt*100 # percent of mean ppt for drought
@@
 
-ggplot(weather, aes(x=ppt1))+
+ggplot(weather, aes(x=annual_precip))+
   geom_histogram(bins=20, color="dodgerblue", fill="dodgerblue", aes(y=..density..))+

### 2018-01-09 manuscript/usses_water_ms.Rmd
updating ms text with new precip characterization

@@
 We cannot definitively rule out this possibility, but we have reason to believe our treatments \emph{should} have been large enough.
-Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of mean growing season precipitation necessary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site (Fig.
+Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of annual precipitation necessary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site (Fig.
 A5-1).
-The 1\% quantile of precipitation at our site is 110 mm, a 47% reduction from the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean growing season precipitation (Appendix 5).
+The 1\% quantile of precipitation at our site is 151 mm, a 50% reduction from the mean, and the 99\% quantile is 532 mm, a 77\% increase from mean growing season precipitation (Appendix 5).
 Thus, our drought treatment represented extreme precipitation amounts, especially in years where ambient precipitation was below average (Fig. 2A).

### 2018-01-09 manuscript/usses_water_ms.tex
updating ms text with new precip characterization

--- 2018-01-09 manuscript/usses_water_ms.tex
@@
 
-\definecolor{blue}{rgb}{0,0,0.7}\newcommand{\new}{\textcolor{blue}}
+\definecolor{blue}{rgb}{0,0,0.7} \newcommand{\new}{\textcolor{blue}}
 
@@
   \endgroup
-}\blfootnote{*Authors contributed equally to this work.}
+}
+
+\blfootnote{*Authors contributed equally to this work.}
 
@@
 
-\hypertarget{introduction}{%
-\section{INTRODUCTION}\label{introduction}}
+\section{INTRODUCTION}\label{introduction}
 
@@
 
-\hypertarget{methods}{%
-\section{METHODS}\label{methods}}
+\section{METHODS}\label{methods}
 
-\hypertarget{study-area}{%
-\subsection{Study Area}\label{study-area}}
+\subsection{Study Area}\label{study-area}
 
@@
 
-\hypertarget{precipitation-experiment}{%
-\subsection{Precipitation Experiment}\label{precipitation-experiment}}
+\subsection{Precipitation Experiment}\label{precipitation-experiment}
 
@@
 
-\hypertarget{data-collection}{%
-\subsection{Data Collection}\label{data-collection}}
+\subsection{Data Collection}\label{data-collection}
 
@@
 
-\hypertarget{data-analysis}{%
-\subsection{Data Analysis}\label{data-analysis}}
+\subsection{Data Analysis}\label{data-analysis}
 
-\hypertarget{ecosystem-functional-response}{%
 \subsubsection{Ecosystem functional
-response}\label{ecosystem-functional-response}}
+response}\label{ecosystem-functional-response}
 
@@
 
-\hypertarget{sensitivity-of-anpp-over-time}{%
 \subsubsection{Sensitivity of ANPP over
-time}\label{sensitivity-of-anpp-over-time}}
+time}\label{sensitivity-of-anpp-over-time}
 
@@
 
-\hypertarget{community-composition-over-time}{%
 \subsubsection{Community composition over
-time}\label{community-composition-over-time}}
+time}\label{community-composition-over-time}
 
@@
 
-\hypertarget{reproducibility}{%
-\subsubsection{Reproducibility}\label{reproducibility}}
+\subsubsection{Reproducibility}\label{reproducibility}
 
@@
 
-\hypertarget{results}{%
-\section{RESULTS}\label{results}}
+\section{RESULTS}\label{results}
 
@@
 
-\hypertarget{discussion}{%
-\section{DISCUSSION}\label{discussion}}
+\section{DISCUSSION}\label{discussion}
 
@@
 enough. Using the methods described by \citet{Lemoine2016}, we
-calculated the percent reduction and increase of mean growing season
-precipitation necessary to reach the 1\% and 99\% extremes of the
-historical precipitation regime at our site (Fig. A5-1). The 1\%
-quantile of precipitation at our site is 110 mm, a 47\% reduction from
-the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean
-growing season precipitation (Appendix 5). Thus, our drought treatment
-represented extreme precipitation amounts, especially in years where
-ambient precipitation was below average (Fig. 2A). The irrigation
-treatment may not qualify as extreme, yet that is the treatment where we
-did observe an effect (Fig. 3A).
+calculated the percent reduction and increase of annual precipitation
+necessary to reach the 1\% and 99\% extremes of the historical
+precipitation regime at our site (Fig. A5-1). The 1\% quantile of
+precipitation at our site is 151 mm, a 50\% reduction from the mean, and
+the 99\% quantile is 532 mm, a 77\% increase from mean growing season
+precipitation (Appendix 5). Thus, our drought treatment represented
+extreme precipitation amounts, especially in years where ambient
+precipitation was below average (Fig. 2A). The irrigation treatment may
+not qualify as extreme, yet that is the treatment where we did observe
+an effect (Fig. 3A).
 
@@
 
-\hypertarget{conclusions}{%
-\section{CONCLUSIONS}\label{conclusions}}
+\section{CONCLUSIONS}\label{conclusions}
 
@@
 
-\hypertarget{acknowledgements}{%
-\section{ACKNOWLEDGEMENTS}\label{acknowledgements}}
+\section{ACKNOWLEDGEMENTS}\label{acknowledgements}
 
@@
 
-\hypertarget{funding}{%
-\section{FUNDING}\label{funding}}
+\section{FUNDING}\label{funding}
 
@@
 
-\hypertarget{author-contributions}{%
-\section{AUTHOR CONTRIBUTIONS}\label{author-contributions}}
+\section{AUTHOR CONTRIBUTIONS}\label{author-contributions}
 
@@
 
-\hypertarget{supplemental-information}{%
-\section{SUPPLEMENTAL INFORMATION}\label{supplemental-information}}
+\section{SUPPLEMENTAL INFORMATION}\label{supplemental-information}
 
@@
 
-\hypertarget{tables}{%
-\section{TABLES}\label{tables}}
+\section{TABLES}\label{tables}
 
@@
 
-\hypertarget{figures}{%
-\section{FIGURES}\label{figures}}
+\section{FIGURES}\label{figures}
 
@@
 
-\bibliography{usses\_water.bib}
+\bibliography{/Users/atredenn/Dropbox/Bibliography/usses_water.bib}
 

### 2018-01-09 code/02_plot_data.R

@@ -87,10 +87,6 @@ soil_moisture <- read.csv("../data/soil_moisture_data/average_seasonal_soil_mois
          VWC_source = ifelse(is.na(VWC)==TRUE, "soilwat", "observed")) %>%
   filter(month %in% c("03","04","05","06"))
 
-
-# soilwat_inserts <- soil_moisture %>%
-#   filter(is.na(VWC) == TRUE)
-
 suppressWarnings(# ignore warnings about missing values, we know they are empty
   soil_vwc <- ggplot(filter(soil_moisture, type == "observed"), aes(x=julian_date, y=VWC, group=Treatment, color=Treatment))+
     geom_line(data = filter(soil_moisture, type == "predicted"), size=0.3, linetype=2, aes(x=julian_date, y=VWC, group=Treatment, color=Treatment))+
@@ -109,11 +105,6 @@ suppressWarnings(# ignore warnings about missing values, we know they are empty
     guides(color=FALSE, linetype = FALSE)
 )
 
-##  Save VWC for statistical modeling that matches the above figure
-# vwc_for_stats <- soil_moisture %>%
-#   select(year, month, day, Treatment, VWC_combo, VWC_source) %>%
-#   rename(VWC = VWC_combo)
-# saveRDS(vwc_for_stats, "../data/soil_moisture_data/vwc_for_stats.RDS")
 
 
 ####

### 2018-01-17 manuscript/Appendix5.Rmd
changes precip to water-years for historical figure, fixes #12

@@
 
+##  Water year defined as precip in Oct-Dec in year t and Jan-Sept in year t+1
+##  following USGS.
+first_water_months <- c("10","11","12") # first months in water year, to be promoted a year
 weather <- read.csv("../data/weather/dubois_station_weather_01092018.csv") %>%
@@
   mutate(year = as.numeric(year)) %>%
-  group_by(year) %>%
-  summarise(annual_precip = sum(precip))
+  mutate(water_year = ifelse(month %in% first_water_months, year+1, year)) %>% # create water years, based on USGS defintion
+  filter(year != 1925) %>% # remove first year because don't have first water-year months
+  group_by(water_year) %>%
+  summarise(annual_precip = sum(precip)) %>%
+  rename(year = water_year)
 

### 2018-01-17 manuscript/usses_water_ms.Rmd
changes precip to water-years for historical figure, fixes #12

@@
 \new{We used the locally parameterized SOILWAT model to generate daily soil moisture predictions for the duration of the experiment, but only used SOILWAT predictions where there were gaps in our data (Fig.
-2B).}
+2B and Appendix X).}
 
@@
 We cannot definitively rule out this possibility, but we have reason to believe our treatments \emph{should} have been large enough.
-Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of annual precipitation necessary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site (Fig.
+Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of \new{water-year} precipitation necessary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site (Fig.
 A5-1).
-The 1\% quantile of precipitation at our site is 151 mm, a 50% reduction from the mean, and the 99\% quantile is 532 mm, a 77\% increase from mean growing season precipitation (Appendix 5).
-Thus, our drought treatment represented extreme precipitation amounts, especially in years where ambient precipitation was below average (Fig. 2A).
-The irrigation treatment may not qualify as extreme, yet that is the treatment where we did observe an effect (Fig. 3A).
+\new{The 1\% quantile of water-year precipitation at our site is 78 mm, a 26% reduction from the mean, and the 99\% quantile is 545 mm, a 84\% increase from mean growing season precipitation (Appendix 5).
+Thus, our drought treatment represented extreme precipitation amounts, and that is the treatment where we observed a small effect on the slope between soil moisture and ANPP (Fig. 3A).
+The irrigation treatment may not have been extreme enough to induce a response.}
 
@@
       \includegraphics[height=7in]{../figures/data_panels.png}
-  \caption{(A) Probability density of historical precipitation from 1926-2016, with the years of the experiment shown with arrows on the \emph{x}-axis.
-(B) \new{Observed (solid line), estimated (dashed lines), and SOILWAT-generated (dotted lines) daily} soil volumetric water content (VWC) in each of the three treatments during the course of the experiment.
-\new{We used the estimates from our statistical model, with gaps filled in by SOILWAT predictions, to calculate cumulative VWC as used in our analysis of treatment effects on ecosystem functional response.} (C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.
+  \caption{(A) Probability density of historical water-year (Oct.--Sept.) precipitation from 1927-2016, with the years of the experiment shown with arrows on the \emph{x}-axis.
+(B) \new{Satistically estimated (solid lines) and SOILWAT-generated (dashed lines) daily} soil volumetric water content (VWC) in each of the three treatments during the course of the experiment.
+\new{We used the estimates from our statistical model, with gaps filled in by SOILWAT predictions, to calculate cumulative March--June VWC as used in our analysis of treatment effects on ecosystem functional response.} (C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.
 Colors in panels B and C identify the treatment, as specified in the legend of panel C.}

### 2018-01-17 manuscript/usses_water_ms.tex
changes precip to water-years for historical figure, fixes #12

@@
 enough. Using the methods described by \citet{Lemoine2016}, we
-calculated the percent reduction and increase of annual precipitation
-necessary to reach the 1\% and 99\% extremes of the historical
-precipitation regime at our site (Fig. A5-1). The 1\% quantile of
-precipitation at our site is 151 mm, a 50\% reduction from the mean, and
-the 99\% quantile is 532 mm, a 77\% increase from mean growing season
-precipitation (Appendix 5). Thus, our drought treatment represented
-extreme precipitation amounts, especially in years where ambient
-precipitation was below average (Fig. 2A). The irrigation treatment may
-not qualify as extreme, yet that is the treatment where we did observe
-an effect (Fig. 3A).
+calculated the percent reduction and increase of \new{water-year}
+precipitation necessary to reach the 1\% and 99\% extremes of the
+historical precipitation regime at our site (Fig. A5-1).
+\new{The 1\% quantile of water-year precipitation at our site is 78 mm, a 26% reduction from the mean, and the 99\% quantile is 545 mm, a 84\% increase from mean growing season precipitation (Appendix 5).
+Thus, our drought treatment represented extreme precipitation amounts, and that is the treatment where we observed a small effect on the slope between soil moisture and ANPP (Fig. 3A).
+The irrigation treatment may not have been extreme enough to induce a response.}
 

### 2018-01-17 code/02_plot_data.R

@@ -31,15 +31,21 @@ library(cowplot)   # For combining ggplot objects
 ####
 ####  READ IN WEATHER DATA AND PLOT --------------------------------------------
 ####
-weather <- read.csv("../data/weather/ClimateIPM.csv")
+##  Water year defined as precip in Oct-Dec in year t and Jan-Sept in year t+1
+##  following USGS.
+first_water_months <- c("10","11","12") # first months in water year, to be promoted a year
+# weather <- read.csv("../data/weather/ClimateIPM.csv")
 weather <- read.csv("../data/weather/dubois_station_weather_01092018.csv") %>%
   dplyr::select(DATE, PRCP) %>%
   dplyr::rename("date" = DATE, "precip" = PRCP) %>%
   separate(date, into = c("year", "month", "day"), sep = "-") %>%
   mutate(precip = ifelse(is.na(precip), 0, precip)) %>% # set missing station data to 0
   mutate(year = as.numeric(year)) %>%
-  group_by(year) %>%
-  summarise(annual_precip = sum(precip))
+  mutate(water_year = ifelse(month %in% first_water_months, year+1, year)) %>% # create water years, based on USGS defintion
+  filter(year != 1925) %>% # remove first year because don't have first water-year months
+  group_by(water_year) %>%
+  summarise(annual_precip = sum(precip)) %>%
+  rename(year = water_year)
 
 trt_data <- weather %>%
   filter(year > 2011) %>%
@@ -48,8 +54,8 @@ trt_data <- weather %>%
   mutate(Treatment = "Control")
 
 ##  Adjust 2012 and 2015 values so they are distinguishable on the plot
-trt_data[which(trt_data$year==2015),"annual_precip"] <- trt_data[which(trt_data$year==2015),"annual_precip"]+8
-trt_data[which(trt_data$year==2014),"annual_precip"] <- trt_data[which(trt_data$year==2014),"annual_precip"]-8
+trt_data[which(trt_data$year==2015),"annual_precip"] <- trt_data[which(trt_data$year==2015),"annual_precip"]-1
+trt_data[which(trt_data$year==2014),"annual_precip"] <- trt_data[which(trt_data$year==2014),"annual_precip"]+1
 
 ppt_histogram <- ggplot(weather, aes(x=annual_precip))+
[…]

### 2018-01-18 manuscript/usses_water_ms.Rmd
adds justification for dropping some control plots from community analysis; closes #14

@@
 Species composition data came from two sources: yearly census maps for each plot made using a pantograph [@Hill1920] and yearly counts of annual species in each plot.
-\new{Mapping was done for plots within one of the permanent exclosures, which included six of the control plots and all of the treatment plots, for a total of 22 plots.}
 From these sources, we determined the density of all annuals and perennials forbs, the basal cover of perennial grasses, and the canopy cover of shrubs.
@@
 Assuming that rare species will respond to treatments more than common ones, our approach is biased towards detecting compositional changes.
+\new{We limited our community level data to observations from the permanent exclosure with our drought and irrigation treatments for the community composition analyses described below (\textbf{Community composition over time}).
+This exclosure included six of the control plots and all of the treatment plots, for a total of 22 plots.
+The other eight control plots are in other pastures, which would add a spatial bias to our community analysis focused on temporal dynamics.}
 

### 2018-01-18 manuscript/usses_water_ms.tex
adds justification for dropping some control plots from community analysis; closes #14

@@
 each plot made using a pantograph \citep{Hill1920} and yearly counts of
-annual species in each plot.
-\new{Mapping was done for plots within one of the permanent exclosures, which included six of the control plots and all of the treatment plots, for a total of 22 plots.}
-From these sources, we determined the density of all annuals and
-perennials forbs, the basal cover of perennial grasses, and the canopy
-cover of shrubs. We made a large plot-treatment-year by species matrix,
-where columns were filled with either basal cover or density, depending
-on the measurement made for the particular species. We standardized the
-values in each column so we could directly compare species quantified
-with different metrics (density, basal cover, and canopy cover). This
-puts all abundance values on the same scale, meaning that common and
-rare species are weighted equally. Assuming that rare species will
-respond to treatments more than common ones, our approach is biased
-towards detecting compositional changes.
+annual species in each plot. From these sources, we determined the
+density of all annuals and perennials forbs, the basal cover of
+perennial grasses, and the canopy cover of shrubs. We made a large
+plot-treatment-year by species matrix, where columns were filled with
+either basal cover or density, depending on the measurement made for the
+particular species. We standardized the values in each column so we
+could directly compare species quantified with different metrics
+(density, basal cover, and canopy cover). This puts all abundance values
+on the same scale, meaning that common and rare species are weighted
+equally. Assuming that rare species will respond to treatments more than
+common ones, our approach is biased towards detecting compositional
+changes.
+\new{We limited our community level data to observations from the permanent exclosure with our drought and irrigation treatments for the community composition analyses described below (\textbf{Community composition over time}).
+This exclosure included six of the control plots and all of the treatment plots, for a total of 22 plots.
+The other eight control plots are in other pastures, which would add a spatial bias to our community analysis focused on temporal dynamics.}
 

### 2018-01-18 manuscript/usses_water_ms.Rmd
trivia

@@
 A3-2), which indicates the data did influence parameter estimates beyond the information from the uninformative priors.
-Ecosystem functional response was similar among treatments (Table 1; Fig.
-3B), but there is evidence that the \new{slopes for the drought and} irrigation treatments are less than the slope for the control treatment.
+Ecosystem functional response was similar among treatments (Table 1; Fig. 3B), but there is evidence that the \new{slope for the drought treatment is greater than the slope for the control treatment.}
 This evidence comes from interpreting the posterior distributions of the slope offsets for the treatments.
-\new{From these distributions,} we calculate a \new{91}% one-tailed probability that the estimate is less than zero for the irrigation treatment \new{and a 96\% one-tailed probability that the estimate is less than zero for the drought treatment} (Fig.
+\new{From these distributions,} we calculate a \new{58}% one-tailed probability that the estimate is greater than zero for the irrigation treatment \new{and a 100\% one-tailed probability that the estimate is greater than zero for the drought treatment} (Fig.
 3A, right panel).
 
-\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3C). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0007; Fig. 3C)}.
+\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3C). 
+Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0007; Fig. 3C)}.
 
@@
 The results of our five year experiment in a sagebrush steppe \new{conforms to one} of our four predictions.
-Ecosystem functional response under chronic \new{drought and irrigation} was different from the control treatment, but community composition remained unchanged (Fig.
+Ecosystem functional response under chronic \new{drought} was different from the control treatment, but community composition remained unchanged (Fig.
 3A, Fig.
@@
 1.
-The decrease in the slope of the VWC $\times$ productivity relationship in the irrigated plots is consistent with a weaker response of productivity to additional water in this ecosystem.
-This suggests that if average soil moisture were pushed consistently higher than currently observed ambient conditions, there would be a weaker relationship between precipitation and productivity in this system.
+\new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots is consistent with a stronger response of productivity to additional water in this ecosystem.
+This suggests that if average soil moisture were pushed consistently lower than currently observed ambient conditions, there would be a stronger relationship between precipitation and productivity in this system.}
 
-\new{The decrease in the slope of the VWC $\times$ productivity relationship in the droughted plots is puzzling because it is counter to our expectation that the slope would be higher than the control plots.
-However, we believe the significance of the slope offsets has more to do with noise than signal.
+\new{However, we believe the significance of the drought slope offset has more to do with noise than signal.
 This is because the full distributions of} the slopes of the VWC $\times$ productivity relationship, \new{not just the offsets}, were similar among treatments (Table 1).

### 2018-01-18 manuscript/usses_water_ms.Rmd
adds PBA edits

@@
 In 2007, we relocated 14 of the original quadrats in permenanent livestock exclosures, \new{which we used} as control plots (i.e. ambient precipitation) in the experiment described below.
-\new{We used the relocated plots as our controls because collecting demographic data is difficult and it is already being collected in these plots for other ongoing studies.}
+\new{We used the relocated plots as our controls because collecting demographic data is time consuming and it was already being collected in these plots for other studies.}
 
@@
 Only a subset of plots were equipped with soil moisture sensors, and within those plots, one or more of the sensors frequently failed to collect data.
-To remedy these problems, we used a statistical model to produce average daily soil moisture values for the ambient, drought, and irrigation treatments during the course of the experiment.
+Therefore, we used a statistical model to estimate average daily soil moisture values for the ambient, drought, and irrigation treatments during the course of the experiment.
 
@@
 ### Sensitivity of ANPP over time
-\new{Our data does not allow us to directly test whether the ecosystem functional response to precipitation may have changed over time in each of the treatments. 
+\new{Our data do not allow us to directly test whether the ecosystem functional response to precipitation may have changed over time since the start of the experiment in each of the treatments. 
 This is because we lack sufficient within-year and within-treatment variation of soil moisture (i.e., within a year and treatment each plot shares the same value of VWC). 
@@
 2B).
-\new{VWC in drought plots was 12\% less than in control plots on average, and VWC in irrigation plots was 19\% higher than in control plots on average.}
+\new{March-June VWC in drought plots was 12\% less than in control plots on average, and March-June VWC in irrigation plots was 19\% higher than in control plots on average across the years of the experiment.}
 Such differences in soil VWC indicate our treatment infrastructure was successful.
@@
 \new{There were also no changes in the dominant species over time in any treatment (Appendix 1).
-Splitting the analysis apart for annuals led to similar results.
+Analyzing annual species on their own produced similar results.
 There is some evidence that the annual community changed in response to treatment in two years (Fig.
-A1-6), but these responses appear to come from very rare species since our analysis weights them equally (Fig.
+A1-6), but these responses appear to come from very rare species since our analysis weights common and rare species equally (Fig.
 A1-4).
@@
 
-\new{However, we believe the significance of the drought slope offset has more to do with noise than signal.
-This is because the full distributions of} the slopes of the VWC $\times$ productivity relationship, \new{not just the offsets}, were similar among treatments (Table 1).
+\new{However, we do not want to over-interpret the significance of the slope offsets given that the full distributions of} the slopes of the VWC $\times$ productivity relationship, \new{not just the offsets}, were similar among treatments (Table 1).
 We therefore conclude that ecosystem functional response is consistent (\new{similar values}) and weak (\new{all broadly overlapping zero}) across all precipitation treatments.
@@
 As the drought progressed, these same plants may have have started to reduce in size or die. 
-Given the long lived perennial species at this site, this increase in sensitivity may indicate a larger future change in community composition and ecosystem functional response.
-}
+Given the long lived perennial species at this site, this increase in sensitivity may indicate a larger future change in community composition and ecosystem functional response. }
 
@@
 Likewise, a long-term increase in water availability could allow species that do not bet hedge to gain prominence and dominate the ecosystem functional response.
-\new{Time \emph{per se} may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}
+\new{The length of the perturbation may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}
 Our results suggest compositional shifts would have the largest impact at high rainfall because the current community maintained consistent ecosystem functional response at very low water availability.

### 2018-01-18 manuscript/usses_water_ms.tex
adds PBA edits

@@
 precipitation) in the experiment described below.
-\new{We used the relocated plots as our controls because collecting demographic data is difficult and it is already being collected in these plots for other ongoing studies.}
+\new{We used the relocated plots as our controls because collecting demographic data is time consuming and it was already being collected in these plots for other studies.}
 
@@
 moisture sensors, and within those plots, one or more of the sensors
-frequently failed to collect data. To remedy these problems, we used a
-statistical model to produce average daily soil moisture values for the
-ambient, drought, and irrigation treatments during the course of the
-experiment.
+frequently failed to collect data. Therefore, we used a statistical
+model to estimate average daily soil moisture values for the ambient,
+drought, and irrigation treatments during the course of the experiment.
 
@@
 
-\new{Our data does not allow us to directly test whether the ecosystem functional response to precipitation may have changed over time in each of the treatments. 
+\new{Our data do not allow us to directly test whether the ecosystem functional response to precipitation may have changed over time since the start of the experiment in each of the treatments. 
 This is because we lack sufficient within-year and within-treatment variation of soil moisture (i.e., within a year and treatment each plot shares the same value of VWC). 
@@
 (VWC) differences among treatments (Fig. 2B).
-\new{VWC in drought plots was 12\% less than in control plots on average, and VWC in irrigation plots was 19\% higher than in control plots on average.}
+\new{March-June VWC in drought plots was 12\% less than in control plots on average, and March-June VWC in irrigation plots was 19\% higher than in control plots on average across the years of the experiment.}
 Such differences in soil VWC indicate our treatment infrastructure was
@@
 was similar among treatments (Table 1; Fig. 3B), but there is evidence
-that the \new{slopes for the drought and} irrigation treatments are less
-than the slope for the control treatment. This evidence comes from
-interpreting the posterior distributions of the slope offsets for the
-treatments. \new{From these distributions,} we calculate a \new{91}\%
-one-tailed probability that the estimate is less than zero for the
-irrigation treatment
-\new{and a 96\% one-tailed probability that the estimate is less than zero for the drought treatment}
+that the
+\new{slope for the drought treatment is greater than the slope for the control treatment.}
+This evidence comes from interpreting the posterior distributions of the
+slope offsets for the treatments. \new{From these distributions,} we
+calculate a \new{58}\% one-tailed probability that the estimate is
+greater than zero for the irrigation treatment
+\new{and a 100\% one-tailed probability that the estimate is greater than zero for the drought treatment}
 (Fig. 3A, right panel).
 
-\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3C). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0007; Fig. 3C)}.
+\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3C). 
+Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0007; Fig. 3C)}.
 
@@
 \new{conforms to one} of our four predictions. Ecosystem functional
-response under chronic \new{drought and irrigation} was different from
-the control treatment, but community composition remained unchanged
-(Fig. 3A, Fig. 4), representing the top right scenario in Fig. 1. The
-decrease in the slope of the VWC \(\times\) productivity relationship in
-the irrigated plots is consistent with a weaker response of productivity
-to additional water in this ecosystem. This suggests that if average
-soil moisture were pushed consistently higher than currently observed
-ambient conditions, there would be a weaker relationship between
-precipitation and productivity in this system.
+response under chronic \new{drought} was different from the control
+treatment, but community composition remained unchanged (Fig. 3A, Fig.
+4), representing the top right scenario in Fig. 1.
+\new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots is consistent with a stronger response of productivity to additional water in this ecosystem.
+This suggests that if average soil moisture were pushed consistently lower than currently observed ambient conditions, there would be a stronger relationship between precipitation and productivity in this system.}
 
-\new{The decrease in the slope of the VWC $\times$ productivity relationship in the droughted plots is puzzling because it is counter to our expectation that the slope would be higher than the control plots.
-However, we believe the significance of the slope offsets has more to do with noise than signal.
-This is because the full distributions of} the slopes of the VWC
-\(\times\) productivity relationship, \new{not just the offsets}, were
-similar among treatments (Table 1). We therefore conclude that ecosystem
-functional response is consistent (\new{similar values}) and weak
-(\new{all broadly overlapping zero}) across all precipitation
-treatments.
+\new{However, we do not want to over-interpret the significance of the slope offsets given that the full distributions of}
+the slopes of the VWC \(\times\) productivity relationship,
+\new{not just the offsets}, were similar among treatments (Table 1). We
+therefore conclude that ecosystem functional response is consistent
+(\new{similar values}) and weak (\new{all broadly overlapping zero})
+across all precipitation treatments.
 
@@
 functional response.
-\new{Time \emph{per se} may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}
+\new{The length of the perturbation may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}
 Our results suggest compositional shifts would have the largest impact

### 2018-01-18 manuscript/Appendix1.Rmd
adds SOILWAT appendix and updates ordering in main text. Closes #13

@@
 suppressWarnings( # ignore factor to character warning. It's OK.
-  cover_species <- read.csv("../data/vital_rates/allrecords_cover_v24.csv") %>%
+  cover_species <- read.csv("../data/vital_rates/allrecords_cover.csv") %>%
     group_by(quad, year, species) %>%
@@
 suppressWarnings( # ignore factor to character warning. It's OK.
-  cover_community_matrix <- read.csv("../data/vital_rates/allrecords_cover_v24.csv") %>%
+  cover_community_matrix <- read.csv("../data/vital_rates/allrecords_cover.csv") %>%
     group_by(quad, year, species) %>%
@@
 suppressWarnings( # ignore factor to character warning. It's OK.
-  density_community_matrix <- read.csv("../data/vital_rates/allrecords_density_v24.csv") %>%
+  density_community_matrix <- read.csv("../data/vital_rates/allrecords_density.csv") %>%
     mutate(ind_num = 1) %>%

### 2018-01-18 manuscript/Appendix2.Rmd
adds SOILWAT appendix and updates ordering in main text. Closes #13

@@
   - \usepackage[sc]{mathpazo}
-  - \usepackage{mathtools}
-bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
-csl: components/ecology_letters.csl
+bibliography: /Users/atredenn/Dropbox/Bibliography/usses_water.bib
+csl: components/ecology.csl
 
@@
 
+\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
 \renewcommand\thefigure{A2-\arabic{figure}}
+\renewcommand\theequation{A2.\arabic{equation}}
 \renewcommand\thetable{A2-\arabic{table}}  
@@
 
-\linenumbers
+\section{Details on SOILWAT predictions}
 
-\section{Estimating ANPP}
-We used a radiometer to nondestructively estimate aboveground net primary productivity.
-Our approach relies on relating greenness in a plot to aboveground biomass.
-In each year we recorded ground reflectances at four bands, two associated with the red spectrum and two associated with the near-infrared spectrum (Table A2-1).
-We took four readings per plot that were averaged for each band.
-Bands 1 and 3 correspond to wavelengths collected by the MODIS satellite and bands 2 and 4 correspond to wavelengths collected by the AVHRR satellite.
+We used a version of the SOILWAT soil moisture model [@Sala1992a] that has been developed specifically for use in semi-arid shrubland ecosystems [@Bradford2014].
+SOILWAT uses daily weather data, ecosystem specific vegetation data, and site specific soil properties to estimate water balance processes.
+Specifically, SOILWAT uses daily rainfall data to estimate rainfall interception by plants, evaporation of intercepted water, snow melt and redistribution, infiltration into the soil, percolation through the soil, evaporation from bare soil, transpiration from each soil layer, and drainage.
+We parameterized SOILWAT using the generic sagebrush steppe parameters and local soil data [@Kleinhesselink2017b].
+SOILWAT was forced by weather data collected at the US Sheep Experimental Station over the course of our experiment.
 
-| Band number | Spectrum name | Center wavelength | Corresponding satellite |
-| ----------- | ------------- | ----------------- | ----------------------- |
-| 1 | red | 626 nm | AVHRR |
-| 2 | red | 652 nm | MODIS |
-| 3 | near-infrared | 875 nm | AVHRR |
-| 4 | near-infrared | 859 nm | MODIS |
+SOILWAT generates soil moisture predictions at several soil depths.
+We averaged the daily predictions from the upper 40 cm of soil.
+These predictions represent ambient conditions, similar to our control plots.
+To generate soil moisture data for our treatment plots, we applied the statistical model described in the main text, which was also used to estimate treatment conditions from control conditions.
+The time series of those predictions, along with our observations and statistical estimates, is shown in Figure A2-1.
 
-Table: Radiometer specifications.
+```{r vwc, fig.cap="Time series of volumetric water content from March to June in each year from the observed measurements (solid lines), statistical estimates (dashed line), and SOILWAT (dotted line).", message=FALSE, echo=FALSE, comment=FALSE, warning=FALSE}
 
-Using the RED and NIR reflectance values, we calculate the normalized difference vegetation index (NDVI) for each plot based on both AVHRR- and MODIS-based wavelengths.
-We calculated NDVI as:
-\vspace{-2em}
-\begin{align}
-\text{NDVI}_{\text{AVHRR}} &= \frac{b_3 - b_1}{b_3 + b_1} \\[10pt]
-\text{NDVI}_{\text{MODIS}} &= \frac{b_4 - b_2}{b_4 + b_2}
-\end{align}
+####
+####  LOAD LIBRARIES -----------------------------------------------------------
+####
+library(tidyverse) # Data science packages
+library(dplyr)     # Data summarizing and manipulating
+library(stringr)   # Working with strings
+library(ggthemes)  # Pleasing themes for ggplot
 
-\noindent{} where $b_x$ refers to band *x* (*x* = 1,2,3,4) in Table A1-1.
+####
+####  PLOT SOIL WATER BY TREATMENT ---------------------------------------------
+####
+mycols <- c("#009E73", "#D55E00", "#0072B2")
 
-To convert plot NDVI to biomass, we regressed known biomass values from calibration plots against NDVI calculate for those plots.
-Calibration plots were located near our experiment plots, and each year we located a new set of 10 plots in which we clipped all aboveground biomass, dried it to a constant weight at 60$^\circ$ C, and the weighed.
-We used these biomass values to estimate regression parameters for both AVHRR- and MODIS-based NDVI.
-We assessed model fit using $R^2$ and, for each year, we used the regression parameters associated with the best fit model to estimate biomass in the experimental plots based on their NDVI values (Table A2-2).
-R code for this procedure is in the file "01_calibrate_radiometer_by_year.R" in the code set.
+soilwat <- readRDS("../data/soil_moisture_data/SOILWAT_treatment_years.RDS") %>%
+  select(date, year, Treatment, VWC_raw, doy) %>%
+  rename(julian_date = doy) %>%
+  mutate(julian_date = as.integer(julian_date))
 
-```{r ndvi_table, include=T, echo=F, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
-library(xtable)
-library(tidyverse)
-library(dplyr)
-tab_dat <- read.csv("../data/estimated_biomass/ndvi_biomass_regression_parameters.csv", row.names = 1) %>%
-  filter(year > 2011 & year < 2017)
-colnames(tab_dat) <- c("Year", "Intercept", "NDVI Slope", "$R^2$", "Min(biomass)", "Max(biomass)", "\\emph{N}", "Algorithm")
+soil_moisture <- read.csv("../data/soil_moisture_data/average_seasonal_soil_moisture.csv") %>%
+  select(-year) %>%
+  separate(simple_date, c("year","month","day")) %>%
+  mutate(year = as.integer(year),
+         Treatment = as.character(Treatment)) %>%
+  ungroup() %>%
+  left_join(soilwat, by = c("year","Treatment","julian_date")) %>%
+  mutate(VWC_combo = ifelse(is.na(VWC)==TRUE, VWC_raw, VWC),
+         VWC_source = ifelse(is.na(VWC)==TRUE, "soilwat", "observed")) %>%
+  filter(month %in% c("03","04","05","06"))
 
-my_caption <- "Details of regression models used to estimate biomass each year."
-print(xtable(tab_dat, caption = my_caption, digits = 2),
-      caption.placement="top",
-      include.rownames = F,
-      sanitize.colnames.function = identity,
-      comment = FALSE,
-      size = "normalsize")
+ggplot(filter(soil_moisture, type == "observed"), aes(julian_date, VWC, group=Treatment, color=Treatment))+
+  geom_line(data = soil_moisture, 
+            aes(x=julian_date, y=VWC_raw, group=Treatment, color=Treatment),
+            size=0.3, linetype=3)+
+   geom_line(data = filter(soil_moisture, type == "predicted"), 
+            aes(x=julian_date, y=VWC, group=Treatment, color=Treatment),
+            size=0.3, linetype=2)+
+  geom_line(size=0.3)+
+  scale_color_manual(values = mycols, name="Treatment")+
+  ylab(expression(paste("Daily Soil VWC (ml ", ml^-1,")")))+
+  xlab("Julian Day")+
+  scale_y_continuous(breaks=seq(0,24,8))+
+  scale_x_continuous(breaks=seq(0,250,20))+
+  facet_grid(year~.)+
+  theme_bw()+
+  theme(panel.grid.minor = element_blank(),
+        strip.background = element_blank(), 
+        strip.text = element_text(size=6))
+
 ```
+
+# References

### 2018-01-18 manuscript/Appendix3.Rmd
adds SOILWAT appendix and updates ordering in main text. Closes #13

--- 2017-06-12 manuscript/Appendix3.Rmd
@@
 
-\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
 \renewcommand\thefigure{A3-\arabic{figure}}  
-\renewcommand\theequation{A3.\arabic{equation}}
 \renewcommand\thetable{A3-\arabic{table}}  
@@
 \textbf{\Large{Appendix 3}} \\
-A.T. Tredennick, A.R. Kleinhesselink, \& P.B. Adler \\
-``Ecosystem and community resistance...'' \\
+A.T. Tredennick, A.R. Kleinhesselink, J.B. Taylor \& P.B. Adler \\
+``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe''  \\
 \emph{PeerJ}
@@
 
-\section{Characterizing Extreme Precipitation Amounts}
+\linenumbers
 
-Following the proposed methods of @Lemoine2016, we calculated quantiles from the empirical distribution of growing season precipitation at Dubios, ID.
-We chose the 1\% quantile to be indicative of extreme dry conditions (drought) and the 99\% quantile to be indicative of extreme wet conditions (irrigation).
-The data consist of 91 yearly records, which we assume are approximately normally distributed.
-The R code below shows our procedure, and Fig. A3-1 shows the results.
+\section{Estimating ANPP}
+We used a radiometer to nondestructively estimate aboveground net primary productivity.
+Our approach relies on relating greenness in a plot to aboveground biomass.
+In each year we recorded ground reflectances at four bands, two associated with the red spectrum and two associated with the near-infrared spectrum (Table A3-1).
+We took four readings per plot that were averaged for each band.
+Bands 1 and 3 correspond to wavelengths collected by the MODIS satellite and bands 2 and 4 correspond to wavelengths collected by the AVHRR satellite.
 
-```{r hrv, fig.cap="Density of the empirical distribution of growing season precipitation at Dubois, ID. Dashed vertical lines show the 1% and 99% quantiles, assuming a normal distribution."}
-library(ggplot2)
-weather    <- read.csv("../data/weather/ClimateIPM.csv")
-mean_ppt   <- mean(weather$ppt1)
-quants_ppt <- quantile(weather$ppt1,probs = c(0.01,0.99))
-quants_ppt[1]/mean_ppt*100 # percent of mean ppt for drought
-quants_ppt[2]/mean_ppt*100 # percent of mean ppt for irrigation
+| Band number | Spectrum name | Center wavelength | Corresponding satellite |
+| ----------- | ------------- | ----------------- | ----------------------- |
+| 1 | red | 626 nm | AVHRR |
+| 2 | red | 652 nm | MODIS |
+| 3 | near-infrared | 875 nm | AVHRR |
+| 4 | near-infrared | 859 nm | MODIS |
 
-ggplot(weather, aes(x=ppt1))+
-  geom_histogram(bins=20, color="dodgerblue", fill="dodgerblue", aes(y=..density..))+
-  geom_line(stat="density", color="blue")+
-  geom_vline(aes(xintercept=quants_ppt[1]), linetype=2)+
-  geom_vline(aes(xintercept=quants_ppt[2]), linetype=2)+
-  ylab("Density")+
-  xlab("Growing Season Precipitation (mm)")+
-  theme_bw()+
-  theme(panel.grid.minor = element_blank())
+Table: Radiometer specifications.
+
+Using the RED and NIR reflectance values, we calculate the normalized difference vegetation index (NDVI) for each plot based on both AVHRR- and MODIS-based wavelengths.
+We calculated NDVI as:
+\vspace{-2em}
+\begin{align}
+\text{NDVI}_{\text{AVHRR}} &= \frac{b_3 - b_1}{b_3 + b_1} \\[10pt]
+\text{NDVI}_{\text{MODIS}} &= \frac{b_4 - b_2}{b_4 + b_2}
+\end{align}
+
+\noindent{} where $b_x$ refers to band *x* (*x* = 1,2,3,4) in Table A1-1.
+
+To convert plot NDVI to biomass, we regressed known biomass values from calibration plots against NDVI calculate for those plots.
+Calibration plots were located near our experiment plots, and each year we located a new set of 10 plots in which we clipped all aboveground biomass, dried it to a constant weight at 60$^\circ$ C, and the weighed.
+We used these biomass values to estimate regression parameters for both AVHRR- and MODIS-based NDVI.
+We assessed model fit using $R^2$ and, for each year, we used the regression parameters associated with the best fit model to estimate biomass in the experimental plots based on their NDVI values (Table A3-2).
+R code for this procedure is in the file "01_calibrate_radiometer_by_year.R" in the code set.
+
+```{r ndvi_table, include=T, echo=F, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
+library(xtable)
+library(tidyverse)
+library(dplyr)
+tab_dat <- read.csv("../data/estimated_biomass/ndvi_biomass_regression_parameters.csv", row.names = 1) %>%
+  filter(year > 2011 & year < 2017)
+colnames(tab_dat) <- c("Year", "Intercept", "NDVI Slope", "$R^2$", "Min(biomass)", "Max(biomass)", "\\emph{N}", "Algorithm")
+
+my_caption <- "Details of regression models used to estimate biomass each year."
+print(xtable(tab_dat, caption = my_caption, digits = 2),
+      caption.placement="top",
+      include.rownames = F,
+      sanitize.colnames.function = identity,
+      comment = FALSE,
+      size = "normalsize")
 ```
-
-# References

### 2018-01-18 manuscript/Appendix4.Rmd
adds SOILWAT appendix and updates ordering in main text. Closes #13

--- 2017-11-07 manuscript/Appendix4.Rmd
@@
 \textbf{\Large{Appendix 4}} \\
-A.T. Tredennick, A.R. Kleinhesselink, B. Taylor \& P.B. Adler \\
+A.T. Tredennick, A.R. Kleinhesselink, J.B. Taylor \& P.B. Adler \\
 ``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe'' \\

### 2018-01-18 manuscript/Appendix5.Rmd
adds SOILWAT appendix and updates ordering in main text. Closes #13

--- 2017-12-10 manuscript/Appendix5.Rmd
@@
 \textbf{\Large{Appendix 5}} \\
-A.T. Tredennick, A.R. Kleinhesselink, J.B. Taylor \& P.B. Adler \\
+A.T. Tredennick, A.R. Kleinhesselink, B. Taylor \& P.B. Adler \\
 ``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe'' \\
@@
 
-\section{Characterizing Extreme Precipitation Amounts}
+\section{Conducting our analysis with NDVI}
 
-Following the proposed methods of @Lemoine2016, we calculated quantiles from the empirical distribution of growing season precipitation at Dubios, ID.
-We chose the 1\% quantile to be indicative of extreme dry conditions (drought) and the 99\% quantile to be indicative of extreme wet conditions (irrigation).
-The data consist of 91 yearly records, which we assume are approximately normally distributed.
-The R code below shows our procedure, and Fig. A5-1 shows the results.
+The first step of our analysis is to convert NDVI from field-based radiometric measurements to aboveground net primary productivity.
+However, this conversion is not perfect and the associated uncertainty is sometimes high (Table A1-2).
+Therefore, we also assessed ecosystem functional response across treatments using NDVI as the response variable rather than ANPP.
+The model essentially the same as the ANPP model, except we use a Beta likelihood for NDVI since its values range from 0 to 1, not including true 0s or 1s (see Stan code, below).
 
-```{r hrv, fig.cap="Density of the empirical distribution of growing season precipitation at Dubois, ID. Dashed vertical lines show the 1% and 99% quantiles, assuming a normal distribution."}
-library(ggplot2)
-weather    <- read.csv("../data/weather/ClimateIPM.csv")
-mean_ppt   <- mean(weather$ppt1)
-quants_ppt <- quantile(weather$ppt1,probs = c(0.01,0.99))
-quants_ppt[1]/mean_ppt*100 # percent of mean ppt for drought
-quants_ppt[2]/mean_ppt*100 # percent of mean ppt for irrigation
+Our NDVI-model results are very similar to the ANPP-model results (Figure A5-1).
+As with our ANPP-model results, the positive offset for the drought treatment does not result in a significantly different slope once applied to the control slope.
 
-ggplot(weather, aes(x=ppt1))+
-  geom_histogram(bins=20, color="dodgerblue", fill="dodgerblue", aes(y=..density..))+
-  geom_line(stat="density", color="blue")+
-  geom_vline(aes(xintercept=quants_ppt[1]), linetype=2)+
-  geom_vline(aes(xintercept=quants_ppt[2]), linetype=2)+
-  ylab("Density")+
-  xlab("Growing Season Precipitation (mm)")+
+```{r plot_betas, fig.cap="Posterior distributions for the effects of drought and irrigation on the intercept and slope of the NDVI-soil moisture relationship.
+Treatment effects show the difference between the coefficients estimated in the treated plots and the control plots.
+Probabilities (Pr </> 0 =) for each coefficient indicate the one-tailed probability that the coefficient is less than or greater than zero, depending on whether the median of the distribution is less than or greater than zero.
+The posterior densities were smoothed for visual clarity by increasing kernel bandwidth by a factor of five.", include=T, echo=F, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
+
+library(tidyverse)    # Data munging
+library(dplyr)        # Data summarizing
+library(ggthemes)     # Pleasing ggplot themes
+library(ggjoy)        # Joy plots!
+library(stringr)      # Working with strings
+library(rstan)        # For MCMC and Stan objects
+library(gridExtra)    # For combining ggplot objects
+library(viridis)      # Pleasing color palette
+library(RColorBrewer) # More color palettes
+library(cowplot)      # For combining ggplots
+
+all_fit <- readRDS("../results/randcoefs_alltreatments_NDVIfit.RDS")
+
+####
+####  FUNCTION TO CALCULATE MAX(Pr(effect) > 0, Pr(effect) < 0) ----
+####
+get_one_tailed <- function(values){
+  above <- 1 - ecdf(values)(0)
+  below <- ecdf(values)(0)
+  max(above,below)
+}
+
+####
+####  PLOT TREATMENT-LEVEL POSTERIOR DISTRIBUTIONS ----
+####
+mycols <- c("#009E73", "#D55E00", "#0072B2")
+param_labels <- c("Control::Intercept",
+                  "Drought::Intercept", 
+                  "Irrigation::Intercept",
+                  "Control::Slope", 
+                  "Drought::Slope", 
+                  "Irrigation::Slope")
+
+betas <- data.frame(extract(all_fit, pars = 'beta')) %>% 
+  mutate(iteration = row_number()) %>% 
+  gather(param_name, estimate, starts_with('beta')) %>%
+  mutate(param_name = factor(param_name, labels = param_labels)) %>%
+  separate(param_name, c("Treatment", "Type"), "::")
+
+treat_slopes <- betas %>%
+  filter(Treatment != "Control") %>%
+  mutate(Type = paste(Type,"Offset"))
+
+slope_probs <- treat_slopes %>%
+  group_by(Treatment, Type) %>%
+  summarise(probs = round(get_one_tailed(estimate),2)) %>%
+  mutate(prob_text = paste("Pr < 0 =", probs))
+slope_probs$prob_text[1] <- paste("Pr > 0 =", slope_probs$probs[1]) # switch direction of drought intercpet since greater than 0
+slope_probs$xpos <- c(-2.8, -0.6, -2.6, -0.6)
+slope_probs$ypos <- c(1.3,1.3,2.55,2.7)
+
+treat_cols <- mycols[2:3]
+
+ggplot(treat_slopes)+
+  geom_vline(aes(xintercept=0), linetype=2, color="grey45")+
+  geom_joy(stat="density", adjust=3, alpha=0.8, aes(x=estimate, y=Treatment, fill=Treatment, color=Treatment, height = ..density..))+
+  geom_text(data = slope_probs, aes(x=xpos, y=ypos, label = prob_text), size = 3)+
+  facet_wrap(~Type, scales = "free")+
+  scale_y_discrete(labels = c("Drought","Irrigation"))+
+  scale_fill_manual(values = treat_cols)+
+  scale_color_manual(values = treat_cols)+
+  guides(fill=FALSE, color=FALSE)+
+  xlab("Coefficient Value")+
+  ylab(NULL)+
   theme_bw()+
-  theme(panel.grid.minor = element_blank())
+  theme(panel.grid.minor = element_blank(),
+        panel.border     = element_blank(), 
+        axis.line        = element_blank(),
+        axis.ticks       = element_blank(),
+        strip.background = element_blank(),
+        strip.text       = element_text(face="bold"),
+        panel.spacing    = unit(1, units = "cm"))
 ```
 
-# References
+\newpage{}
+
+\subsection{Stan Code}
+```{r stan1, echo=T, eval=F}
+data {
+  int<lower=0> Npreds;         # number of covariates, including intercept
+  int<lower=0> Npreds2;        # number of random effect covariates
+  int<lower=0> Nplots;         # number of plots
+  int<lower=0> Ntreats;        # number of treatments
+  int<lower=0> Nobs;           # number of observations
+  int<lower=0> Nyears;         # number of years
+  vector[Nobs] y;              # vector of observations
+  row_vector[Npreds] x[Nobs];	 # design matrix for fixed effects
+  row_vector[Npreds2] z[Nobs]; # simple design matrix for random effects 
+  int plot_id[Nobs];           # vector of plot ids
+  int treat_id[Nobs];          # vector of treatment ids
+  int year_id[Nobs];           # vector of year ids
+}
+
+parameters {
+	vector[Npreds] beta;	             # overall coefficients
+	vector[Nyears] year_off;           # vector of year effects
+	cholesky_factor_corr[Npreds2] L_u; # cholesky factor of plot random effect corr matrix
+  vector[Npreds2] beta_plot[Nplots]; # plot level random effects
+	vector<lower=0>[Npreds2] sigma_u;  # plot random effect std. dev.
+	real<lower=0> sd_y;                # treatment-level observation std. dev.
+	real<lower=0> sigma_year;          # year std. dev. hyperprior
+	real<lower=0> phi;                 # dispersion parameter
+}
+
+transformed parameters {
+  vector[Nobs] A;             # parameter for beta distn
+  vector[Nobs] B;             # parameter for beta distn
+  vector[Nobs] yhat;               # vector of expected values
+  vector[Npreds2] u[Nplots];       # transformed plot random effects
+  matrix[Npreds2,Npreds2] Sigma_u; # plot ranef cov matrix
+
+  Sigma_u = diag_pre_multiply(sigma_u, L_u); # cholesky factor for plot-level covariance matrix 
+  for(j in 1:Nplots)
+    u[j] = Sigma_u * beta_plot[j]; # plot random intercepts and slopes
+
+  # regression model for expected values (one for each plot-year)
+  for (i in 1:Nobs)
+    yhat[i] = inv_logit(x[i]*beta + z[i]*u[plot_id[i]] + year_off[year_id[i]]);
+
+  A = yhat * phi;
+  B = (1.0 - yhat) * phi;
+}
+
+model {
+  ####  PRIORS
+  phi ~ cauchy(0, 5); 
+  sigma_year ~ cauchy(0,2.5);
+  sd_y ~ cauchy(0,2.5);
+  year_off ~ normal(0,sigma_year); # priors on year effects, shared variance
+  beta ~ normal(0,5);	             # priors on treatment coefficients
+  L_u ~ lkj_corr_cholesky(2.0);      # prior on the cholesky factor which controls the 
+                                    # correlation between plot level treatment effects
+  sigma_u ~ cauchy(0,2.5);
+
+  for(i in 1:Nplots)
+		beta_plot[i] ~ normal(0,1); # plot-level coefficients for intercept and slope
+
+	####  LIKELIHOOD
+  y ~ beta(A, B); # observations vary according to beta distribution
+}
+
+generated quantities{
+  corr_matrix[Npreds2] R = multiply_lower_tri_self_transpose(L_u);
+  cov_matrix[Npreds2] V = quad_form_diag(R,sigma_u);
+}
+
+```

### 2018-01-18 manuscript/Appendix6.Rmd
adds SOILWAT appendix and updates ordering in main text. Closes #13

---
layout: 11pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \onehalfspacing
  - \usepackage{rotating}
  - \usepackage{color, soul}
  - \usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{tikz}
  - \usepackage{bm,mathrsfs}
  - \usepackage[sc]{mathpazo}
bibliography: ~/Dropbox/Bibliography/usses_water.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
    number_sections: true
fontsize: 11pt
geometry: margin=1in
linkcolor: black
urlcolor: black
---

\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{A6-\arabic{figure}}
\renewcommand\theequation{A6.\arabic{equation}}
\renewcommand\thetable{A6-\arabic{table}}  
\renewcommand\thesection{Section A6.\arabic{section}}

\begin{center}
\textbf{\Large{Appendix 6}} \\
A.T. Tredennick, A.R. Kleinhesselink, J.B. Taylor \& P.B. Adler \\
``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe'' \\
\emph{PeerJ}
\end{center}

\section{Characterizing Extreme Precipitation Amounts}

Following the proposed methods of @Lemoine2016, we calculated quantiles from the empirical distribution of growing season precipitation at Dubios, ID.
We chose the 1\% quantile to be indicative of extreme dry conditions (drought) and the 99\% quantile to be indicative of extreme wet conditions (irrigation).
The data consist of 91 yearly records, which we assume are approximately normally distributed for these purposes.
The R code below shows our procedure, and Fig. A6-1 shows the results.

```{r hrv, fig.cap="Density of the empirical distribution of growing season precipitation at Dubois, ID. Dashed vertical lines show the 1% and 99% quantiles, assuming a normal distribution.", message=FALSE}
library(tidyverse)
library(dplyr)

##  Water year defined as precip in Oct-Dec in year t and Jan-Sept in year t+1
##  following USGS.
first_water_months <- c("10","11","12") # first months in water year, to be promoted a year
weather <- read.csv("../data/weather/dubois_station_weather_01092018.csv") %>%
  dplyr::select(DATE, PRCP) %>%
  dplyr::rename("date" = DATE, "precip" = PRCP) %>%
  separate(date, into = c("year", "month", "day"), sep = "-") %>%
  mutate(precip = ifelse(is.na(precip), 0, precip)) %>% # set missing station data to 0
  mutate(year = as.numeric(year)) %>%
  mutate(water_year = ifelse(month %in% first_water_months, year+1, year)) %>% # create water years, based on USGS defintion
  filter(year != 1925) %>% # remove first year because don't have first water-year months
  group_by(water_year) %>%
  summarise(annual_precip = sum(precip)) %>%
  rename(year = water_year)

mean_ppt   <- mean(weather$annual_precip)
quants_ppt <- quantile(weather$annual_precip,probs = c(0.01,0.99))
quants_ppt[1]/mean_ppt*100 # percent of mean ppt for drought
quants_ppt[2]/mean_ppt*100 # percent of mean ppt for irrigation

ggplot(weather, aes(x=annual_precip))+
  geom_histogram(bins=20, color="dodgerblue", fill="dodgerblue", aes(y=..density..))+
  geom_line(stat="density", color="blue")+
  geom_vline(aes(xintercept=quants_ppt[1]), linetype=2)+
  geom_vline(aes(xintercept=quants_ppt[2]), linetype=2)+
  ylab("Density")+
  xlab("Growing Season Precipitation (mm)")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())
```

# References

### 2018-01-18 manuscript/usses_water_ms.Rmd
adds SOILWAT appendix and updates ordering in main text. Closes #13

@@
 \new{Following the above procedure, we still lacked soil moisture data for March 2012 (observations did not start unitl April) and for a string of days in 2013 during which the soil moisture probes failed take readings.
-Therefore, we used a version of the SOILWAT soil moisture model} [@Sala1992a] \new{that has been specifically designed for semiarid shrublands and grasslands} [@Bradford2014].
+To fill in these gaps, we used a version of the SOILWAT soil moisture model} [@Sala1992a] \new{that has been specifically designed for semiarid shrublands and grasslands} [@Bradford2014].
 \new{The model was parameterized using generic sagebrush steppe parameters and local soil texture, soil bulk density, and precipitation data} [@Kleinhesselink2017b].
 \new{We used the locally parameterized SOILWAT model to generate daily soil moisture predictions for the duration of the experiment, but only used SOILWAT predictions where there were gaps in our data (Fig.
-2B and Appendix X).}
+2B and Appendix 2).
+SOILWAT predicted daily soil moisture under ambient conditions similar to our control plots.
+We applied the same statistical model and procedure described above to estimate soil moisture in drought and irrigation plots based on control plot conditions.}
 
@@
 We fit regressions to a MODIS-based index and an AVHRR-based index for each year and retained the regression with the better fit based on $R^2$ values.
-We then predicted ANPP using the best regression equation for each year (Appendix 2).
-\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 4).}
+We then predicted ANPP using the best regression equation for each year (Appendix 3).
+\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 5).}
 
@@
 \new{We used a vague prior distribution for each $\beta$: $\boldsymbol{\beta} \sim \text{Normal}(0,5)$.}
-A full description of our model and associated R [@R2016] code is in Appendix 3.
+A full description of our model and associated R [@R2016] code is in Appendix 4.
 
@@
 All R code and data necessary to reproduce our analysis has been archived on Figshare (\emph{link here after acceptance}) and released on GitHub (https://github.com/atredennick/usses_water/releases/v0.1).
-We also include annotated Stan code in our model description in Appendix 3.
+We also include annotated Stan code in our model description in Appendix 4.
 
@@
 A5-1).
-\new{The 1\% quantile of water-year precipitation at our site is 78 mm, a 26% reduction from the mean, and the 99\% quantile is 545 mm, a 84\% increase from mean growing season precipitation (Appendix 5).
+\new{The 1\% quantile of water-year precipitation at our site is 78 mm, a 26% reduction from the mean, and the 99\% quantile is 545 mm, a 84\% increase from mean growing season precipitation (Appendix 6).
 Thus, our drought treatment represented extreme precipitation amounts, and that is the treatment where we observed a small effect on the slope between soil moisture and ANPP (Fig. 3A).
@@
 \item [Appendix 1.] \new{Additional information on plant community composition and dynamics, Table A1-1, Table A1-2, Table A1-3, Fig. A1-1, Fig. A1-2, Fig. A1-3, Fig. A1-4, Fig. A1-5, and Fig. A1-6.}
-\item [Appendix 2.] Additional methods and information on estimating aboveground net primary productivity, Table A2-1, and Table A2-2.
-\item [Appendix 3.] Details of the hierarchical Bayesian model, Fig. A2-1, Fig. A2-2, and Fig. A2-3.
-\item [Appendix 4.] \new{Results from analysis of NDVI without conversion to ANPP, Fig. A3-1.}
-\item [Appendix 5.] Details on analysis of precipitation historical range of variability and Fig. A4-1.
+\item [Appendix 2.] \new{Additional details on soil moisture modeling with SOILWAT and Fig. A2-1.}
+\item [Appendix 3.] Additional methods and information on estimating aboveground net primary productivity, Table A3-1, and Table A3-2.
+\item [Appendix 4.] Details of the hierarchical Bayesian model, Fig. A4-1, Fig. A4-2, and Fig. A4-3.
+\item [Appendix 5.] \new{Results from analysis of NDVI without conversion to ANPP, Fig. A5-1.}
+\item [Appendix 6.] Details on analysis of precipitation historical range of variability and Fig. A6-1.
 \end{description}

### 2018-01-18 manuscript/usses_water_ms.tex
adds SOILWAT appendix and updates ordering in main text. Closes #13

@@
 \new{Following the above procedure, we still lacked soil moisture data for March 2012 (observations did not start unitl April) and for a string of days in 2013 during which the soil moisture probes failed take readings.
-Therefore, we used a version of the SOILWAT soil moisture model}
+To fill in these gaps, we used a version of the SOILWAT soil moisture model}
 \citep{Sala1992a}
@@
 \new{We used the locally parameterized SOILWAT model to generate daily soil moisture predictions for the duration of the experiment, but only used SOILWAT predictions where there were gaps in our data (Fig.
-2B and Appendix X).}
+2B and Appendix X).
+SOILWAT predicted daily soil moisture under ambient conditions similar to our control plots.
+We applied the same statistical model and procedure described above to estimate soil moisture in drought and irrigation plots based on control plot conditions.}
 

### 2018-01-18 manuscript/Appendix5.Rmd
updates NDVI based analysis with table of posteriors

@@
 Our NDVI-model results are very similar to the ANPP-model results (Figure A5-1).
-As with our ANPP-model results, the positive offset for the drought treatment does not result in a significantly different slope once applied to the control slope.
+As with our ANPP-model results, the positive offset for the drought treatment does not result in a significantly different slope once applied to the control slope (Table A5-1).
 
@@
 
+```{r effect_table, include=TRUE, comment=FALSE,warning=FALSE,message=FALSE,results='asis', echo=FALSE}
+library(xtable)
+treat_effects <- data.frame(extract(all_fit, pars = 'beta')) %>% 
+  mutate(iteration = row_number()) %>% 
+  gather(param_name, estimate, starts_with('beta')) %>%
+  mutate(param_name = factor(param_name, labels = param_labels)) %>% 
+  spread(param_name, estimate) %>% 
+  mutate(`Drought::Intercept` = `Drought::Intercept` + `Control::Intercept`, 
+         `Irrigation::Intercept` = `Irrigation::Intercept` + `Control::Intercept`,  
+         `Drought::Slope` = `Drought::Slope` + `Control::Slope`, 
+         `Irrigation::Slope` = `Irrigation::Slope` + `Control::Slope`) %>%
+  gather(param_name, estimate, -iteration)
+
+effect_summary <- treat_effects %>%
+  group_by(param_name) %>%
+  summarise(mean_estimate = round(mean(estimate), 2),
+            median_estimate = round(median(estimate), 2),
+            lower_95_ci = round(quantile(estimate, probs = c(0.025)), 2),
+            upper_95_ci = round(quantile(estimate, probs = c(0.975)), 2)) %>%
+  separate(param_name, into = c("Treatment", "Coefficient"), sep = "::") %>%
+  arrange(Coefficient, Treatment) %>%
+  select(Coefficient, everything())
+
+colnames(effect_summary) <- c("Coefficient", "Treatment", "Posterior Mean", "Posterior Median",
+                              "Lower 95\\% BCI", "Upper 95\\% BCI")
+coef_caption <- "Summary statistics from the posterior distributions of coefficients for each treatment ($\\beta$ coefficients in equation 1).
+The `Intercept' and `Slope' summaries reported here for drought and irrigation are from the posterior distributions of the intercept and slope for the control treatment plus the offsets for each treatment.
+Posterior distributions of the offsets are in Figure A5-1."
+print(xtable(effect_summary, caption = coef_caption, digits = 2),
+      caption.placement="top",
+      include.rownames = F,
+      sanitize.colnames.function = identity,
+      comment = FALSE,
+      size = "normalsize")
+```
+
 \newpage{}
@@
 }
-
 ```

### 2018-01-18 manuscript/review_response_R1.Rmd
more revisions before sending to coauthors
trivia

@@
 \response{The sensitivity analysis allows us to examine how the magnitude of the treatment effect changes over time.
-We found that the response of ANPP to drought grew over time, while the response of ANPP to irrigation is static (new panel D in Figure 3).
+We found that the response of ANPP to drought grew over time, while the response of ANPP to irrigation is static (new panel C in Figure 3).
 We think this analysis improves the connection of our results to the Hierarchical Response Framework.}
@@
 \response{We have added a new appendix (Appendix X) that provides more details on the dominance and rarity of the communities across and through time.
-In brief, rank abundance curves are very similar for all treatments and through time, with the same two perennial species (\emph{Artemisia tripartita} and \emph{Balsamorhiza sagittata}) dominant in all treatments, acounting for about 25\% of total basal cover.
-Similarly, all treatments had the same dominant annual species, whose dynamics were more associated with interannual variation in precipitation than to variation among treatments.
+In brief, rank abundance curves are very similar for all treatments and through time, with the same two perennial species (\emph{Artemisia tripartita} and \emph{Balsamorhiza sagittata}) dominant in all treatments, acounting for about 25\% of total cover.
+Similarly, all treatments had the same dominant annual species, whose dynamics did not vary by treatment.
 These ancillary results add weight to our conclusion that the ecosystem responses is driven by dominant species, which are similar among treatments.}

### 2018-01-18 manuscript/Appendix1.Rmd
trivia

@@
 plot_biomass <- readRDS(paste0(biomass_path,"permanent_plots_estimated_biomass.RDS"))
-weather_data <- read.csv(paste0(weather_path,"ClimateIPM.csv"))
+first_water_months <- c("10","11","12") # first months in water year, to be promoted a year
+weather_data <- read.csv("../data/weather/dubois_station_weather_01092018.csv") %>%
+  dplyr::select(DATE, PRCP) %>%
+  dplyr::rename("date" = DATE, "precip" = PRCP) %>%
+  separate(date, into = c("year", "month", "day"), sep = "-") %>%
+  mutate(precip = ifelse(is.na(precip), 0, precip)) %>% # set missing station data to 0
+  mutate(year = as.numeric(year)) %>%
+  mutate(water_year = ifelse(month %in% first_water_months, year+1, year)) %>% # create water years, based on USGS defintion
+  filter(year != 1925) %>% # remove first year because don't have first water-year months
+  group_by(water_year) %>%
+  summarise(annual_precip = sum(precip)) %>%
+  rename(year = water_year)
 plot_info    <- read.csv("../data/estimated_biomass/quad_info.csv")
@@
 
-Table A1-1: Statistical results from linear model relating density of Allysum desertorum to precipitation.
-```{r report_models, echo = FALSE}
-summary(lm(mean_density~ppt1*Treatment, data = filter(dom_annuals, species == "Allyssum desertorum")))
-```
+<!-- Table A1-1: Statistical results from linear model relating density of Allysum desertorum to precipitation. -->
+<!-- ```{r report_models, echo = FALSE} -->
+<!-- summary(lm(mean_density~annual_precip*Treatment, data = filter(dom_annuals, species == "Allyssum desertorum"))) -->
+<!-- ``` -->
 
-\newpage{}
-Table A1-2: Statistical results from linear model relating density of Bromus tectorum to precipitation.
-```{r report_models2, echo = FALSE}
-summary(lm(mean_density~ppt1*Treatment, data = filter(dom_annuals, species == "Bromus tectorum")))
-```
+<!-- \newpage{} -->
+<!-- Table A1-2: Statistical results from linear model relating density of Bromus tectorum to precipitation. -->
+<!-- ```{r report_models2, echo = FALSE} -->
+<!-- summary(lm(mean_density~annual_precip*Treatment, data = filter(dom_annuals, species == "Bromus tectorum"))) -->
+<!-- ``` -->
 
-\newpage{}
+<!-- \newpage{} -->
 
@@
 ```{r rac_plots, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Rank abundance curves for perennial species.
-Basal area of individuals was summed withnin years and plots, and then the total area values were averaged across years and plots for each treatment.", fig.pos="h",fig.align="center"}
+Basal area (grasses) or canopy cover (shrubs and forbs) of individuals was summed withnin years and plots, and then the total area values were averaged across years and plots for each treatment.", fig.pos="h",fig.align="center"}
 ####
@@
   xlab("Species rank")+
-  ylab(expression(paste("Mean basal area (",m^2,")")))+
+  ylab(expression(paste("Mean area (",m^2,")")))+
   facet_wrap(~Treatment, ncol = 3)
@@
 Values of mean area were averaged over plots.
-The four most dominant species are labelled in each panel."}
+The four most dominant species are labelled in each panel.
+Area is measured as basal area for grasses and canopy cover for shrubs and forbs."}
 cover_data <- {}
@@
   xlab("Species rank")+
-  ylab(expression(paste("Mean basal area (",m^2,")")))+
+  ylab(expression(paste("Mean area (",m^2,")")))+
   facet_grid(Treatment~year)
@@
 
-```{r annual_weather, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Relationship between precipitation and abundance of dominant annuals."}
-dom_annuals <- annual_data %>%
-  filter(species %in% c("Allyssum desertorum","Bromus tectorum")) %>%
-  select(-mean_rank) %>%
-  rename(Species = species) %>%
-  left_join(weather_data, by = "year")
+<!-- ```{r annual_weather, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Relationship between precipitation and abundance of dominant annuals."} -->
+<!-- dom_annuals <- annual_data %>% -->
+<!--   filter(species %in% c("Allyssum desertorum","Bromus tectorum")) %>% -->
+<!--   select(-mean_rank) %>% -->
+<!--   rename(Species = species) %>% -->
+<!--   left_join(weather_data, by = "year") -->
 
-ggplot(dom_annuals, aes(x = ppt1, y = mean_density, color = Species))+
-  geom_point()+
-  xlab("Precipitation (mm)")+
-  ylab("Abundance")
-```
+<!-- ggplot(dom_annuals, aes(x = annual_precip, y = mean_density, color = Species))+ -->
+<!--   geom_point()+ -->
+<!--   xlab("Precipitation (mm)")+ -->
+<!--   ylab("Abundance") -->
+<!-- ``` -->
 

### 2018-01-18 manuscript/usses_water_ms.Rmd
trivia

@@
 \new{Water availability is a major driver of annual net primary productivity (ANPP) in grassland ecosystems} [@Huxman2004; @Hsu2012].
-\new{Therefore, projected changes in precipitation regimes, associated with global climate change, will impact grassland ecosystem functioning.
-Increases in precipitation variability will result in prolonged periods of extremely wet or dry years, and predicting the consequences of such chronic resource alterations requires understanding how ecosystems respond to above and below average precipitation levels.}
-
+\new{Therefore, projected changes in precipitation regimes, associated with global climate change, will impact grassland ecosystem functioning.}
 At any given site, the relationship between ANPP and water availability (e.g., soil moisture) can be characterized by regressing historical observations of ANPP on observations of soil moisture. 

### 2018-01-18 manuscript/usses_water_ms.tex
trivia

@@
 \citep{Huxman2004, Hsu2012}.
-\new{Therefore, projected changes in precipitation regimes, associated with global climate change, will impact grassland ecosystem functioning.
-Increases in precipitation variability will result in prolonged periods of extremely wet or dry years, and predicting the consequences of such chronic resource alterations requires understanding how ecosystems respond to above and below average precipitation levels.}
-
+\new{Therefore, projected changes in precipitation regimes, associated with global climate change, will impact grassland ecosystem functioning.}
 At any given site, the relationship between ANPP and water availability
@@
 \new{We used the locally parameterized SOILWAT model to generate daily soil moisture predictions for the duration of the experiment, but only used SOILWAT predictions where there were gaps in our data (Fig.
-2B and Appendix X).
+2B and Appendix 2).
 SOILWAT predicted daily soil moisture under ambient conditions similar to our control plots.
@@
 \(R^2\) values. We then predicted ANPP using the best regression
-equation for each year (Appendix 2).
-\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 4).}
+equation for each year (Appendix 3).
+\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 5).}
 
@@
 A full description of our model and associated R \citep{R2016} code is
-in Appendix 3.
+in Appendix 4.
 
@@
 We also include annotated Stan code in our model description in Appendix
-3.
+4.
 
@@
 historical precipitation regime at our site (Fig. A5-1).
-\new{The 1\% quantile of water-year precipitation at our site is 78 mm, a 26% reduction from the mean, and the 99\% quantile is 545 mm, a 84\% increase from mean growing season precipitation (Appendix 5).
+\new{The 1\% quantile of water-year precipitation at our site is 78 mm, a 26% reduction from the mean, and the 99\% quantile is 545 mm, a 84\% increase from mean growing season precipitation (Appendix 6).
 Thus, our drought treatment represented extreme precipitation amounts, and that is the treatment where we observed a small effect on the slope between soil moisture and ANPP (Fig. 3A).

### 2018-01-19 manuscript/review_response_R1.Rmd
finalish version sent to PBA

@@
 We think this analysis improves the connection of our results to the Hierarchical Response Framework.}
+
+\response{In addition to changes in resposne to reviewer comments, we have added more detail in our description of how we estimated soil moisture based on direct observations.
+While doing this, we discovered an error on our part: we use March-June volumetric water content as our covariate in each year, but we did not actually have soil moisture estimates for March 2012 (sensors were not installed until April 2012).
+Rather than dropping an important month in terms of water availability, we decided to use output from the SOILWAT soil moisture model as parameterized for our site (lines xxx-xxx in the revised manuscript, Figure 2B, and new Appendix 2).
+Importantly, our main results are unchanged: ecosystem functional responses is similar among treatments, though there is now weak evidence that the drought resonse is more steep than the control response.}
 

### 2018-01-19 manuscript/Appendix1.Rmd
finalish version sent to PBA

@@
 ```{r rac_plots, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Rank abundance curves for perennial species.
-Basal area (grasses) or canopy cover (shrubs and forbs) of individuals was summed withnin years and plots, and then the total area values were averaged across years and plots for each treatment.", fig.pos="h",fig.align="center"}
+Area of individuals (either canopy or basal area cover, depending on life form) was summed withnin years and plots, and then the total area values were averaged across years and plots for each treatment.", fig.pos="h",fig.align="center"}
 ####
@@
 The four most dominant species are labelled in each panel.
-Area is measured as basal area for grasses and canopy cover for shrubs and forbs."}
+Area is measured as either basal area or canopy cover area, depending on life form."}
 cover_data <- {}

### 2018-01-19 manuscript/usses_water_ms.Rmd
finalish version sent to PBA

@@
 
-  **Results.** Ecosystem functional response, defined as the relationship between soil moisture and ANPP was similar among drought and control treatments, but the irrigation treatment had a lower slope than the control treatment.
+  **Results.** Ecosystem functional response, defined as the relationship between soil moisture and ANPP was similar among irrigation and control treatments, but the drought treatment had a greater slope than the control treatment.
 However, all estimates for the effect of soil moisture on ANPP overlapped zero, indicating the relationship is weak and uncertain regardless of treatment.
@@
 
-  **Discussion.** Despite some evidence that ecosystem functional response became less sensitive under sustained wet conditions, the response of ANPP to soil moisture was consistently weak and community composition was stable.
+  **Discussion.** Despite some evidence that ecosystem functional response became more sensitive under sustained drought conditions, the response of ANPP to soil moisture was consistently weak and community composition was stable.
 The similarity of ecosystem functional responses across treatments was not related to compensatory shifts at the plant community level, but instead may reflect the insensitivity of the dominant species to soil moisture.
@@
 From these sources, we determined the density of all annuals and perennials forbs, the basal cover of perennial grasses, and the canopy cover of shrubs.
-We made a large plot-treatment-year by species matrix, where columns were filled with either basal cover or density, depending on the measurement made for the particular species.
+We made a large plot-treatment-year by species matrix, where columns were filled with either cover or density, depending on the measurement made for the particular species.
 We standardized the values in each column so we could directly compare species quantified with different metrics (density, basal cover, and canopy cover).
@@
 1.
-\new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots is consistent with a stronger response of productivity to additional water in this ecosystem.
+\new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots is consistent with a stronger response of productivity to additional water in this ecosystem under chronic drought.
 This suggests that if average soil moisture were pushed consistently lower than currently observed ambient conditions, there would be a stronger relationship between precipitation and productivity in this system.}
@@
 A5-1).
-\new{The 1\% quantile of water-year precipitation at our site is 78 mm, a 26% reduction from the mean, and the 99\% quantile is 545 mm, a 84\% increase from mean growing season precipitation (Appendix 6).
+\new{The 1\% quantile of water-year precipitation at our site is 78 mm, a 26\% reduction from the mean, and the 99\% quantile is 545 mm, a 84\% increase from mean growing season precipitation (Appendix 6).
 Thus, our drought treatment represented extreme precipitation amounts, and that is the treatment where we observed a small effect on the slope between soil moisture and ANPP (Fig. 3A).

### 2018-01-19 manuscript/usses_water_ms.tex
finalish version sent to PBA

@@
 \textbf{Results.} Ecosystem functional response, defined as the
-relationship between soil moisture and ANPP was similar among drought
-and control treatments, but the irrigation treatment had a lower slope
+relationship between soil moisture and ANPP was similar among irrigation
+and control treatments, but the drought treatment had a greater slope
 than the control treatment. However, all estimates for the effect of
@@
 \textbf{Discussion.} Despite some evidence that ecosystem functional
-response became less sensitive under sustained wet conditions, the
+response became more sensitive under sustained drought conditions, the
 response of ANPP to soil moisture was consistently weak and community
@@
 plot-treatment-year by species matrix, where columns were filled with
-either basal cover or density, depending on the measurement made for the
+either cover or density, depending on the measurement made for the
 particular species. We standardized the values in each column so we
@@
 4), representing the top right scenario in Fig. 1.
-\new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots is consistent with a stronger response of productivity to additional water in this ecosystem.
+\new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots is consistent with a stronger response of productivity to additional water in this ecosystem under chronic drought.
 This suggests that if average soil moisture were pushed consistently lower than currently observed ambient conditions, there would be a stronger relationship between precipitation and productivity in this system.}
@@
 historical precipitation regime at our site (Fig. A5-1).
-\new{The 1\% quantile of water-year precipitation at our site is 78 mm, a 26% reduction from the mean, and the 99\% quantile is 545 mm, a 84\% increase from mean growing season precipitation (Appendix 6).
+\new{The 1\% quantile of water-year precipitation at our site is 78 mm, a 26\% reduction from the mean, and the 99\% quantile is 545 mm, a 84\% increase from mean growing season precipitation (Appendix 6).
 Thus, our drought treatment represented extreme precipitation amounts, and that is the treatment where we observed a small effect on the slope between soil moisture and ANPP (Fig. 3A).

### 2018-01-19 manuscript/merge_pdfs.sh

@@ -1,3 +1,3 @@
 #!/bin/bash 
 
-"/System/Library/Automator/Combine PDF Pages.action/Contents/Resources/join.py" -o usses_water_ms_appendices.pdf usses_water_ms.pdf Appendix1.pdf Appendix2.pdf Appendix3.pdf
+"/System/Library/Automator/Combine PDF Pages.action/Contents/Resources/join.py" -o usses_water_ms_appendices.pdf usses_water_ms.pdf Appendix1.pdf Appendix2.pdf Appendix3.pdf Appendix4.pdf Appendix5.pdf Appendix6.pdf

### 2018-01-22 manuscript/Appendix1.Rmd
addresses PBA edits and fixes colors to close #16

@@
 
-# Details on the plant community and dynamics
+# Details on plant community structure
 
@@
 
-mycols <- c("#009E73", "#0072B2", "#D55E00")
+mycols <- c("#009E73", "#D55E00", "#0072B2")
 ggplot(nmds_df_annuals, aes(x=MDS1, y=MDS2, color=Treatment))+

### 2018-01-22 manuscript/Appendix3.Rmd
addresses PBA edits and fixes colors to close #16

@@
   - \usepackage{mathtools}
-bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
 csl: components/ecology_letters.csl
@@
 To convert plot NDVI to biomass, we regressed known biomass values from calibration plots against NDVI calculate for those plots.
-Calibration plots were located near our experiment plots, and each year we located a new set of 10 plots in which we clipped all aboveground biomass, dried it to a constant weight at 60$^\circ$ C, and the weighed.
+Calibration plots were located near our experiment plots, and each year we located a new set of 12-16 plots in which we clipped all aboveground biomass, dried it to a constant weight at 60$^\circ$ C, and the weighed.
 We used these biomass values to estimate regression parameters for both AVHRR- and MODIS-based NDVI.

### 2018-01-22 manuscript/Appendix5.Rmd
addresses PBA edits and fixes colors to close #16

@@
   - \usepackage[sc]{mathpazo}
-bibliography: ~/Dropbox/Bibliography/usses_water.bib
+bibliography: /Users/atredenn/Dropbox/Bibliography/usses_water.bib
 csl: components/ecology.csl
@@
 
-```{r plot_betas, fig.cap="Posterior distributions for the effects of drought and irrigation on the intercept and slope of the NDVI-soil moisture relationship.
+```{r plot_betas, fig.cap="Posterior distributions for the effects of drought and irrigation on the intercept and slope of the soil moisture-NDVI relationship.
 Treatment effects show the difference between the coefficients estimated in the treated plots and the control plots.

### 2018-01-22 manuscript/usses_water_ms.Rmd
addresses PBA edits and fixes colors to close #16

@@
 Saturating relationships are actually common [@Hsu2012; @Gherardi2015a], perhaps because other resources, like nitrogen, become more limiting in wet years than dry years.
-Failure to accurately estimate the curvature of the soil moisture-ANPP relationship will lead to over- or underprediction of ANNP under extreme precipitation [@Peters2012].
+Failure to accurately estimate the curvature of the soil moisture-ANPP relationship will lead to over- or underprediction of ANPP under extreme precipitation [@Peters2012].
 
@@
 
-To that end, here we report the results of a five-year precipitation manipulation experiment in a sagebrush steppe grassland.
+To that end, here we report the results of a five-year precipitation manipulation experiment in a sagebrush steppe.
 We imposed drought and irrigation treatments (approximately $\pm50\%$) and measured ecosystem (ANPP) and community (species composition) responses.
@@
 To answer these questions we fit a generalized linear mixed effects model to test whether the regressions differed among treatments.
-We also analyzed community composition and \new{the sensitivity of ANPP to drought and irrigation treatments} over time, allowing us to place our experimental results within the framework our scenarios (Fig.
+We also analyzed community composition and \new{the sensitivity of ANPP to drought and irrigation treatments} over time, allowing us to place our experimental results within the framework of our scenarios (Fig.
 1).
@@
 
-We conducted our precipitation manipulation experiment in a sagebrush steppe community at the USDA, U.S.
-Sheep Experiment Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
-The plant community is dominated by the shrub \emph{Artemesia tripartita}, \new{the perennial forb \emph{Balsamorhiza sagittata}}, and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata} \new{(see Appendix 1 for rank abundance curves)}.
+We conducted our precipitation manipulation experiment in a sagebrush steppe community at the USDA-ARS Sheep Experiment Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
+The plant community is dominated by the shrub \emph{Artimesia tripartita}, \new{the perennial forb \emph{Balsamorhiza sagittata}}, and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata} \new{(see Appendix 1 for rank abundance curves)}.
 During the period of our experiment (2011 -- 2016), mean annual precipitation was 250 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
@@
 \new{We first averaged the observed soil moisture readings for each day ($d$) and plot ($i$), $x_{i,d}$.
-Experimental plots were organized in groups of two, with each group ($g$) containing a drought and irrigated plot.
-In addition a nearby area outside the drought or irrigated plots was monitored for local ambient soil moisture conditions.
+Experimental plots were located in pairs, with each group ($g$) containing a drought and irrigated plot. 
+In addition, a nearby area outside the drought or irrigated plots was monitored for local ambient soil moisture conditions.
 Within each group of plots we standardized the irrigation and drought effects on soil moisture relative to the ambient soil moisture conditions. 
-Specifically we subtracted the ambient daily soil moisture from the soil moisture measured within the drought and irrigation plots within each group and then divided by the standard deviation of daily soil moisture values measured in the ambient conditions $\left( \Delta x_{g,d,\text{trt.}} = (x_{g,d,\text{trt.}} - x_{g,d,\text{ambient}})/\text{s.d.}(x_{g,d,\text{ambient}}) \right)$ where $\Delta x$ is the standardized treatment effect and $x_{g,d,\text{trt.}}$ is the raw soil moisture measure for plot group $g$ on day $d$ and treatment $\text{trt.}$.
+Specifically we subtracted the ambient daily soil moisture from the soil moisture measured within the drought and irrigation plots within each group and then divided by the standard deviation of daily soil moisture values measured in the ambient conditions $\left( \Delta x_{g,d,\text{trt.}} = (x_{g,d,\text{trt.}} - x_{g,d,\text{ambient}})/\text{s.d.}(x_{g,d,\text{ambient}}) \right)$ where $\Delta x$ is the standardized treatment effect and $x_{g,d,\text{trt.}}$ is the raw soil moisture measure for plot group $g$ on day $d$ and treatment $trt.$.
 These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.}
@@
 
-\new{Following the above procedure, we still lacked soil moisture data for March 2012 (observations did not start unitl April) and for a string of days in 2013 during which the soil moisture probes failed take readings.
+\new{Following the above procedure, we still lacked soil moisture data for March 2012 (observations did not start unitl April) and for a string of days in 2013 during which the soil moisture probes failed to take readings.
 To fill in these gaps, we used a version of the SOILWAT soil moisture model} [@Sala1992a] \new{that has been specifically designed for semiarid shrublands and grasslands} [@Bradford2014].
@@
 ## Data Collection
-We estimated aboveground net primary productivity (ANPP) using a radiometer to relate ground reflectance to plant biomass [see @Byrne2011 for a review].
+We estimated aboveground net primary productivity (ANPP) using a radiometer to relate ground reflectance to plant biomass [@Byrne2011].
 We recorded ground reflectance at four wavelengths, two associated with red reflectance (626 nm and 652 nm) and two associated with near-infrared reflectance (875 nm and 859 nm).
@@
 We also took readings in 12 (2015), 15 (2012, 2013, 2014), or 16 (2016) calibration plots adjacent to the experimental site, in which we harvested all aboveground biomass produced in the current year (we excluded litter and standing dead material), dried it to a constant weight at 60$^{\circ}$C, and weighed it to estimate ANPP.
-We harvested at peak biomass each year, typically in late June.
+We harvested at peak green biomass each year, typically in late June.
 
@@
 We then predicted ANPP using the best regression equation for each year (Appendix 3).
-\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 5).}
+\new{Our results do not change when we analyze the soil moisture-NDVI relationship instead of the soil moisture-ANPP relationship (Appendix 5).}
 
@@
 Assuming that rare species will respond to treatments more than common ones, our approach is biased towards detecting compositional changes.
-\new{We limited our community level data to observations from the permanent exclosure with our drought and irrigation treatments for the community composition analyses described below (\textbf{Community composition over time}).
+\new{We limited our analysis of community data to observations from the permanent exclosure containing our drought and irrigation treatments (see \textbf{Community composition over time}).
 This exclosure included six of the control plots and all of the treatment plots, for a total of 22 plots.
@@
 Plot and year of treatment were included as random effects to account for non-independence of observations, as described below.
-We log-transformed ANPP to account for heteroscedasticity.
+We log-transformed ANPP to reduce heteroscedasticity.
 Both log(ANPP) and VWC were standardized to have mean 0 and unit variance before fitting the model [i.e., $(x_i - \bar{x})/\sigma_x$].
@@
 After calculating sensitivity for each drought and irrigation plot in each year, we regressed sensitivity against year of treatment using the \texttt{lm()} function in R} [@R2016].
-\new{This analysis also allows us to link particularly sensitive treatment years to changes in community composition.}
+\new{This analysis also allows us to link particularly sensitive treatment-years to changes in community composition.}
 
@@
 \new{We conducted the above analysis for all species, and then conducted a separate analysis for annual species only (Appendix 1). 
-Annual species have shorter life spans, so conducting a separate analysis allowed us to discover if community responses are restricted to short-lived species that might respond more quickly to altered precipitation regimes.
+Annual species have shorter life spans, so conducting a separate analysis allowed us to test whether we might find stronger evidence for community responses to altered precipitation when we focus on short-lived species.
 Given the dominance of perennial species in our system (Appendix 1), shifts in the annual plant community could be difficult to detect in the analysis of the full community.}
@@
 
-Ambient precipitation was variable over the five years of the study (Fig. 2A), meaning our experiment captured a range of environmental conditions.
+Ambient precipitation and soil moisture were variable over the five years of the study (Fig. 2A-B).
 ANPP varied from a minimum of 77.4 g m$^{-2}$ in 2014 to a maximum of 239.3 g m$^{-2}$ in 2016 when averaged across treatments (Fig. 2C).
@@
 \new{March-June VWC in drought plots was 12\% less than in control plots on average, and March-June VWC in irrigation plots was 19\% higher than in control plots on average across the years of the experiment.}
-Such differences in soil VWC indicate our treatment infrastructure was successful.
+The differences in soil VWC indicate our treatment infrastructure was successful.
 ANPP was highly variable across plots within years, \new{as indicated by the large and overlapping standard deviations} (Fig. 2C).
@@
 Although the parameter estimates for the effect of soil moisture overlap zero, the posterior distributions of the slopes all shrank and shifted to more positive values relative to the prior distributions (Fig.
-A3-2), which indicates the data did influence parameter estimates beyond the information from the uninformative priors.
+A3-2), which indicates the data did influence parameter estimates.
 Ecosystem functional response was similar among treatments (Table 1; Fig. 3B), but there is evidence that the \new{slope for the drought treatment is greater than the slope for the control treatment.}
@@
 A1-4).
-By definition, thsee rare species contribute little to ANPP and thus any compositional changes have little influence on ecosystem functional response.}
+By definition, these rare species contribute little to ANPP and thus any compositional changes have little influence on ecosystem functional response.}
 
@@
 Ecosystem response to a new precipitation regime depends on the physiological responses of constituent species and the rate at which community composition shifts to favor species better able to take advantage of, or cope with, new resource levels [@Smith2009].
-Previous work has shown that community compositional shifts can be both rapid, on the order of years [@Hoover2014], and slow, on order of decades [@Knapp2012; @Wilcox2016].
+Previous work has shown that community compositional shifts can be either rapid, on the order of years [@Hoover2014], or slow, on the order of decades [@Knapp2012; @Wilcox2016].
 A lingering question is how the time scales of ecosystem response and community change vary among ecosystems. 
@@
 
-The results of our five year experiment in a sagebrush steppe \new{conforms to one} of our four predictions.
+The results of our five year experiment in a sagebrush steppe \new{conform to one} of our four predictions.
 Ecosystem functional response under chronic \new{drought} was different from the control treatment, but community composition remained unchanged (Fig.
@@
 1.
-\new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots is consistent with a stronger response of productivity to additional water in this ecosystem under chronic drought.
-This suggests that if average soil moisture were pushed consistently lower than currently observed ambient conditions, there would be a stronger relationship between precipitation and productivity in this system.}
+\new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots indicates increased sensitivity to water availability under chronic drought.
+A strict interpretation of this result implies that if average soil moisture were pushed consistently lower than currently observed ambient conditions, there would be a stronger relationship between precipitation and productivity in this system.}
 
@@
 We can think of three reasons. 
-Two are limitations of our study, and one is the life history traits of the species in our focal communities.
-We first discuss the potential limitations of our study, and then discuss the biological explanation.
+Two are limitations of our study, and one invokes the life history traits of the species at our study site.
 
-First, it could be that our precipitation manipulations were not large enough to induce a response.
-That is, a 50\% decrease in any given year may not be abnormal given our site's historical range of variability [@Knapp2017].
+First, our precipitation manipulations may not have been large enough to induce a response.
+A 50\% decrease in any given year may not be abnormal given our site's historical range of variability [@Knapp2017].
 We cannot definitively rule out this possibility, but we have reason to believe our treatments \emph{should} have been large enough.
@@
 Second, ANPP at our site may be influenced by additional factors, not only the cumulative March-June soil moisture covariate we included in our statistical model.
-\new{For example,} @LaPierre2016 \new{found that site-scale ANPP is better predicted by abiotic factors such as nutrient availability than precipitation.}
+\new{For example,} @LaPierre2016 \new{found that site-scale ANPP is better predicted by nutrient availability than precipitation.}
 Moreover, temperature can impact ANPP directly [@Epstein1997] and by exacerbating the effects of soil moisture [@DeBoeck2011].
@@
 A3-2).
-Unfortunately, we did not collect data on additional factors because our focus was on precipitation.
 
@@
 The dry and variable environment of the sagebrush steppe has likely selected for bet hedging species that can maintain function at low water availability and have weak responses to high water availability.
-In so doing, the dominant species in our ecosystem avoid "boom and bust" cycles, which corresponds to the weak effect of soil moisture on ANPP (i.e., the Bayesian credible intervals for the slopes overlapping zero; Table 1).
+In so doing, the dominant species in our ecosystem avoid "boom and bust" cycles, which may explain the weak effect of soil moisture on ANPP (i.e., the Bayesian credible intervals for the slopes overlapping zero; Table 1).
 
@@
 
-\new{We found that ANPP became more sensitive to the drought over time (Fig.
-3C).
-In other words, relative to the effect of the drought shelters on soil moisture, ANPP was more reduced by the drought conditions later in the experiment than early on in the experiment.
+\new{We found that ANPP became more sensitive to the drought treatment over time (Fig. 3C).
 We interepret this increase in sensitivity as the effect of cumulative impacts of the drought on dominant species.
 Plants may not have shown a large response in terms of ANPP in the first years of the experiment if they could grow from stored carbohydrate reserves or if they abstained from flowering and reproduction.
-As the drought progressed, these same plants may have have started to reduce in size or die. 
-Given the long lived perennial species at this site, this increase in sensitivity may indicate a larger future change in community composition and ecosystem functional response. }
+As the drought progressed, these same plants may have have started to shrink or die. 
+Given the long-lived perennial species at this site, this increase in sensitivity may indicate a larger future change in community composition and ecosystem functional response. }
 
@@
 \new{The length of the perturbation may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}
-Our results suggest compositional shifts would have the largest impact at high rainfall because the current community maintained consistent ecosystem functional response at very low water availability.
-\new{However, the increased sensitivity of ANPP to drought conditions suggests a shift in ecosystem functional response may be near, and simply takes more time than we observed here.}
 
@@
 
-We thank the many summer research technicians who collected the data reported in this paper and the USDA, U.S. Sheep Experiment Station (Dubois, ID) for facilitating work on their property.
+We thank the many summer research technicians who collected the data reported in this paper and the USDA-ARS Sheep Experiment Station (Dubois, ID) for facilitating work on their property.
 We also thank Susan Durham for clarifying our thinking on the statistical analyses and Kevin Wilcox for helpful discussions on analyzing community composition data.
@@
 inyear_caption <- "Results from statistical tests for clustering and dispersion of community composition among precipitation treatments.
-`adonis' tests whether treatments form unique clusters in multidimensial space; `betadisper' tests whether treatments have similar dispersion.
-For both tests, \\emph{P} values greater than 0.05 indicate there is no support that the treatments differ."
+`adonis' tests whether treatments form unique clusters in multidimensional space; `betadisper' tests whether treatments have similar dispersion.
+For both tests, \\emph{P} values greater than 0.05 indicate there is no support for the hypothesis that the treatments differ."
 print(xtable(dplyr::select(sppcomp_stats, -Sums, -Means), caption = inyear_caption, digits = 2),
@@
       \includegraphics[height=7in]{../figures/data_panels.png}
-  \caption{(A) Probability density of historical water-year (Oct.--Sept.) precipitation from 1927-2016, with the years of the experiment shown with arrows on the \emph{x}-axis.
-(B) \new{Satistically estimated (solid lines) and SOILWAT-generated (dashed lines) daily} soil volumetric water content (VWC) in each of the three treatments during the course of the experiment.
+  \caption{(A) Probability density of historical water-year (Oct.--Sept.) precipitation from 1927-2016, with the years of the experiment shown by arrows on the \emph{x}-axis.
+(B) \new{Statistically estimated (solid lines) and SOILWAT-generated (dashed lines) daily} soil volumetric water content (VWC) in each of the three treatments during the course of the experiment.
 \new{We used the estimates from our statistical model, with gaps filled in by SOILWAT predictions, to calculate cumulative March--June VWC as used in our analysis of treatment effects on ecosystem functional response.} (C) Mean (points) ANPP and its standard deviation (error bars) for each year of the experiment.
@@
 Treatment effects show the difference between the coefficients estimated in the treated plots and the control plots.
-Probabilities (``Pr $\lessgtr$ 0 ='') for each coefficient indicate the one-tailed probability that the coefficient is less than or greater than zero, depending on whether the median of the distribution is less than or greater than zero.
+Probabilities (``Pr $</>$ 0 ='') for each coefficient indicate the one-tailed probability that the coefficient is less than or greater than zero, depending on whether the median of the distribution is less than or greater than zero.
 The posterior densities were smoothed for visual clarity by increasing kernel bandwidth by a factor of five.

### 2018-01-22 manuscript/usses_water_ms.tex
addresses PBA edits and fixes colors to close #16

@@
 
-\section{INTRODUCTION}\label{introduction}
+\hypertarget{introduction}{%
+\section{INTRODUCTION}\label{introduction}}
 
 \new{Water availability is a major driver of annual net primary productivity (ANPP) in grassland ecosystems}
-\citep{Huxman2004, Hsu2012}.
+(Huxman et al. 2004; Hsu, Powell, and Adler 2012).
 \new{Therefore, projected changes in precipitation regimes, associated with global climate change, will impact grassland ecosystem functioning.}
@@
 responses can then be used to infer how ANPP may change under future
-precipitation regimes \citep[e.g.,][]{Hsu2012}. A problem with this
-approach is that it requires extrapolation if future precipitation falls
-outside the historical range of variability
-\citep{Smith2011, Peters2012}. For example, the soil moisture-ANPP
+precipitation regimes (e.g., Hsu, Powell, and Adler 2012). A problem
+with this approach is that it requires extrapolation if future
+precipitation falls outside the historical range of variability (Smith
+2011; Peters et al. 2012). For example, the soil moisture-ANPP
 relationship may be linear within the historical range of interannual
 variation, but could saturate at higher levels of soil moisture.
-Saturating relationships are actually common
-\citep{Hsu2012, Gherardi2015a}, perhaps because other resources, like
-nitrogen, become more limiting in wet years than dry years. Failure to
-accurately estimate the curvature of the soil moisture-ANPP relationship
-will lead to over- or underprediction of ANNP under extreme
-precipitation \citep{Peters2012}.
+Saturating relationships are actually common (Hsu, Powell, and Adler
+2012; Laureano A Gherardi and Sala 2015), perhaps because other
+resources, like nitrogen, become more limiting in wet years than dry
+years. Failure to accurately estimate the curvature of the soil
+moisture-ANPP relationship will lead to over- or underprediction of ANPP
+under extreme precipitation (Peters et al. 2012).
 
@@
 All four outcomes are possible in any given ecosystem, but the time
-scales at which the different scenarios play out likely differ
-\citep{Smith2009, Wilcox2016, Knapp2017}. To determine these time
-scales, we need to amass information on how quickly ecosystem functional
-responses change in different ecosystems. We also need to understand
-whether changes at the ecosystem level are driven by community level
-changes or individual level responses.
+scales at which the different scenarios play out likely differ (Smith,
+Knapp, and Collins 2009; Wilcox et al. 2016; Knapp et al. 2017). To
+determine these time scales, we need to amass information on how quickly
+ecosystem functional responses change in different ecosystems. We also
+need to understand whether changes at the ecosystem level are driven by
+community level changes or individual level responses.
 
 To that end, here we report the results of a five-year precipitation
-manipulation experiment in a sagebrush steppe grassland. We imposed
-drought and irrigation treatments (approximately \(\pm50\%\)) and
-measured ecosystem (ANPP) and community (species composition) responses.
-We focus on how the drought and irrigation treatments affect the
-relationship between interannual variation in available soil moisture
-and interannual variation in ANPP, and if community dynamics underlie
-the ecosystem responses. In particular, we are interested in the
-consistency of the soil moisture-ANPP relationship among treatments. Is
-the relationship steeper under the drought treatment at low soil
-moisture? Does the relationship saturate under the irrigation treatment
-at high soil moisture? To answer these questions we fit a generalized
-linear mixed effects model to test whether the regressions differed
-among treatments. We also analyzed community composition and
+manipulation experiment in a sagebrush steppe. We imposed drought and
+irrigation treatments (approximately \(\pm50\%\)) and measured ecosystem
+(ANPP) and community (species composition) responses. We focus on how
+the drought and irrigation treatments affect the relationship between
+interannual variation in available soil moisture and interannual
+variation in ANPP, and if community dynamics underlie the ecosystem
+responses. In particular, we are interested in the consistency of the
+soil moisture-ANPP relationship among treatments. Is the relationship
+steeper under the drought treatment at low soil moisture? Does the
+relationship saturate under the irrigation treatment at high soil
+moisture? To answer these questions we fit a generalized linear mixed
+effects model to test whether the regressions differed among treatments.
+We also analyzed community composition and
 \new{the sensitivity of ANPP to drought and irrigation treatments} over
 time, allowing us to place our experimental results within the framework
-our scenarios (Fig. 1).
+of our scenarios (Fig. 1).
 
-\section{METHODS}\label{methods}
+\hypertarget{methods}{%
+\section{METHODS}\label{methods}}
 
-\subsection{Study Area}\label{study-area}
+\hypertarget{study-area}{%
+\subsection{Study Area}\label{study-area}}
 
 We conducted our precipitation manipulation experiment in a sagebrush
-steppe community at the USDA, U.S. Sheep Experiment Station (USSES) near
+steppe community at the USDA-ARS Sheep Experiment Station (USSES) near
 Dubois, Idaho (44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m above
 sea level. The plant community is dominated by the shrub
-\emph{Artemesia tripartita},
+\emph{Artimesia tripartita},
 \new{the perennial forb \emph{Balsamorhiza sagittata}}, and three
@@
 were missing due to malfunction of the sensors or dataloggers. See
-\citet{Kleinhesselink2017b} for more details on our approach to
-estimating soil moisture.
+Kleinhesselink (2017) for more details on our approach to estimating
+soil moisture.
 
-\new{Following the above procedure, we still lacked soil moisture data for March 2012 (observations did not start unitl April) and for a string of days in 2013 during which the soil moisture probes failed take readings.
+\new{Following the above procedure, we still lacked soil moisture data for March 2012 (observations did not start unitl April) and for a string of days in 2013 during which the soil moisture probes failed to take readings.
 To fill in these gaps, we used a version of the SOILWAT soil moisture model}
-\citep{Sala1992a}
+(Sala, Lauenroth, and Parton 1992)
 \new{that has been specifically designed for semiarid shrublands and grasslands}
-\citep{Bradford2014}.
+(Bradford, Schlaepfer, and Lauenroth 2014).
 \new{The model was parameterized using generic sagebrush steppe parameters and local soil texture, soil bulk density, and precipitation data}
-\citep{Kleinhesselink2017b}.
+(Kleinhesselink 2017).
 \new{We used the locally parameterized SOILWAT model to generate daily soil moisture predictions for the duration of the experiment, but only used SOILWAT predictions where there were gaps in our data (Fig.
@@
 
-\subsection{Data Collection}\label{data-collection}
+\hypertarget{data-collection}{%
+\subsection{Data Collection}\label{data-collection}}
 
 We estimated aboveground net primary productivity (ANPP) using a
-radiometer to relate ground reflectance to plant biomass \citep[see][
-for a review]{Byrne2011}. We recorded ground reflectance at four
-wavelengths, two associated with red reflectance (626 nm and 652 nm) and
-two associated with near-infrared reflectance (875 nm and 859 nm). At
-each plot in each year, we took four readings of ground reflectances at
-the above wavelengths. We also took readings in 12 (2015), 15 (2012,
-2013, 2014), or 16 (2016) calibration plots adjacent to the experimental
-site, in which we harvested all aboveground biomass produced in the
-current year (we excluded litter and standing dead material), dried it
-to a constant weight at 60\(^{\circ}\)C, and weighed it to estimate
-ANPP. We harvested at peak biomass each year, typically in late June.
+radiometer to relate ground reflectance to plant biomass (Byrne et al.
+2011). We recorded ground reflectance at four wavelengths, two
+associated with red reflectance (626 nm and 652 nm) and two associated
+with near-infrared reflectance (875 nm and 859 nm). At each plot in each
+year, we took four readings of ground reflectances at the above
+wavelengths. We also took readings in 12 (2015), 15 (2012, 2013, 2014),
+or 16 (2016) calibration plots adjacent to the experimental site, in
+which we harvested all aboveground biomass produced in the current year
+(we excluded litter and standing dead material), dried it to a constant
+weight at 60\(^{\circ}\)C, and weighed it to estimate ANPP. We harvested
+at peak green biomass each year, typically in late June.
 
@@
 equation for each year (Appendix 3).
-\new{Our analysis, described below, can be done using NDVI directly, but we focus on ANPP to more easily relate our findings to previous work and because our results do not change if we analyze NDVI directly (Appendix 5).}
+\new{Our results do not change when we analyze the soil moisture-NDVI relationship instead of the soil moisture-ANPP relationship (Appendix 5).}
 
 Species composition data came from two sources: yearly census maps for
-each plot made using a pantograph \citep{Hill1920} and yearly counts of
+each plot made using a pantograph (Hill 1920) and yearly counts of
 annual species in each plot. From these sources, we determined the
@@
 changes.
-\new{We limited our community level data to observations from the permanent exclosure with our drought and irrigation treatments for the community composition analyses described below (\textbf{Community composition over time}).
+\new{We limited our analysis of community data to observations from the permanent exclosure containing our drought and irrigation treatments (see \textbf{Community composition over time}).
 This exclosure included six of the control plots and all of the treatment plots, for a total of 22 plots.
@@
 Based on our own observations and previous work at our study site
-\citep{Blaisdell1958, Dalgleish2011, Adler2012}, we chose to use
-cumulative volumetric water content from March through June as our
-metric of soil moisture (hereafter referred to as `VWC'). To achieve
-this goal, we fit a generalized linear mixed effects regression model
-with log(ANPP) as the response variable and VWC and treatment as fixed
-effects. Plot and year of treatment were included as random effects to
-account for non-independence of observations, as described below. We
-log-transformed ANPP to account for heteroscedasticity. Both log(ANPP)
-and VWC were standardized to have mean 0 and unit variance before
-fitting the model {[}i.e., \((x_i - \bar{x})/\sigma_x\){]}.
+(Blaisdell 1958; Dalgleish et al. 2011; Adler, Dalgleish, and Ellner
+2012), we chose to use cumulative volumetric water content from March
+through June as our metric of soil moisture (hereafter referred to as
+`VWC'). To achieve this goal, we fit a generalized linear mixed effects
+regression model with log(ANPP) as the response variable and VWC and
+treatment as fixed effects. Plot and year of treatment were included as
+random effects to account for non-independence of observations, as
+described below. We log-transformed ANPP to reduce heteroscedasticity.
+Both log(ANPP) and VWC were standardized to have mean 0 and unit
+variance before fitting the model {[}i.e.,
+\((x_i - \bar{x})/\sigma_x\){]}.
 
@@
 \new{We used a vague prior distribution for each $\beta$: $\boldsymbol{\beta} \sim \text{Normal}(0,5)$.}
-A full description of our model and associated R \citep{R2016} code is
-in Appendix 4.
+A full description of our model and associated R (R Core Team 2016) code
+is in Appendix 4.
 
@@
 estimates of all unknowns via the No-U-Turn Hamiltonian Monte Carlo
-sampler in Stan \citep{stan2016}. We used the R package `rstan'
-\citep{rstan2016} to link R \citep{R2016} to Stan. We obtained samples
-from the posterior distribution for all model parameters from four
-parallel MCMC chains run for 10,000 iterations, saving every
-\(10^{\text{th}}\) sample. Trace plots of all parameters were visually
-inspected to ensure well-mixed chains and convergence. We also made sure
-all scale reduction factors (\(\hat{R}\) values) were less than 1.1
-\citep{Gelman2009}.
+sampler in Stan (Stan Development Team 2016b). We used the R package
+`rstan' (Stan Development Team 2016a) to link R (R Core Team 2016) to
+Stan. We obtained samples from the posterior distribution for all model
+parameters from four parallel MCMC chains run for 10,000 iterations,
+saving every \(10^{\text{th}}\) sample. Trace plots of all parameters
+were visually inspected to ensure well-mixed chains and convergence. We
+also made sure all scale reduction factors (\(\hat{R}\) values) were
+less than 1.1 (Gelman and Hill 2009).
 
+\hypertarget{sensitivity-of-anpp-over-time}{%
 \subsubsection{Sensitivity of ANPP over
-time}\label{sensitivity-of-anpp-over-time}
+time}\label{sensitivity-of-anpp-over-time}}
 
@@
 After calculating sensitivity for each drought and irrigation plot in each year, we regressed sensitivity against year of treatment using the \texttt{lm()} function in R}
-\citep{R2016}.
-\new{This analysis also allows us to link particularly sensitive treatment years to changes in community composition.}
+(R Core Team 2016).
+\new{This analysis also allows us to link particularly sensitive treatment-years to changes in community composition.}
 
+\hypertarget{community-composition-over-time}{%
 \subsubsection{Community composition over
-time}\label{community-composition-over-time}
+time}\label{community-composition-over-time}}
 
@@
 if community composition overlapped, or not, among treatments in each
-year. We used the \texttt{vegan::metaMDS()} function \citep{Oksanen2016}
-to calculate Bray-Curtis distances and then to run the NMDS analysis. We
-used the \texttt{vegan::adonis()} function \citep{Oksanen2016} to
-perform permutational multivariate analysis of variance to test whether
+year. We used the \texttt{vegan::metaMDS()} function (Oksanen 2016) to
+calculate Bray-Curtis distances and then to run the NMDS analysis. We
+used the \texttt{vegan::adonis()} function (Oksanen 2016) to perform
+permutational multivariate analysis of variance to test whether
 treatment plots formed distinct groupings. To test whether treatment
 plots were equally dispersed, or not, we used the
-\texttt{vegan::betadisper()} function \citep{Oksanen2016}.
+\texttt{vegan::betadisper()} function (Oksanen 2016).
 
 \new{We conducted the above analysis for all species, and then conducted a separate analysis for annual species only (Appendix 1). 
-Annual species have shorter life spans, so conducting a separate analysis allowed us to discover if community responses are restricted to short-lived species that might respond more quickly to altered precipitation regimes.
+Annual species have shorter life spans, so conducting a separate analysis allowed us to test whether we might find stronger evidence for community responses to altered precipitation when we focus on short-lived species.
 Given the dominance of perennial species in our system (Appendix 1), shifts in the annual plant community could be difficult to detect in the analysis of the full community.}
 
-\subsubsection{Reproducibility}\label{reproducibility}
+\hypertarget{reproducibility}{%
+\subsubsection{Reproducibility}\label{reproducibility}}
 
@@
 
-\section{RESULTS}\label{results}
+\hypertarget{results}{%
+\section{RESULTS}\label{results}}
 
-Ambient precipitation was variable over the five years of the study
-(Fig. 2A), meaning our experiment captured a range of environmental
-conditions. ANPP varied from a minimum of 77.4 g m\(^{-2}\) in 2014 to a
-maximum of 239.3 g m\(^{-2}\) in 2016 when averaged across treatments
-(Fig. 2C). ANPP was slightly higher in irrigation plots
-\new{(on average 15\% higher)} and slightly lower in drought plots
-\new{(on average 25\% lower) relative to control plots} (Fig. 2C),
-corresponding to observed and estimated soil volumetric water content
-(VWC) differences among treatments (Fig. 2B).
+Ambient precipitation and soil moisture were variable over the five
+years of the study (Fig. 2A-B). ANPP varied from a minimum of 77.4 g
+m\(^{-2}\) in 2014 to a maximum of 239.3 g m\(^{-2}\) in 2016 when
+averaged across treatments (Fig. 2C). ANPP was slightly higher in
+irrigation plots \new{(on average 15\% higher)} and slightly lower in
+drought plots \new{(on average 25\% lower) relative to control plots}
+(Fig. 2C), corresponding to observed and estimated soil volumetric water
+content (VWC) differences among treatments (Fig. 2B).
 \new{March-June VWC in drought plots was 12\% less than in control plots on average, and March-June VWC in irrigation plots was 19\% higher than in control plots on average across the years of the experiment.}
-Such differences in soil VWC indicate our treatment infrastructure was
+The differences in soil VWC indicate our treatment infrastructure was
 successful. ANPP was highly variable across plots within years,
@@
 positive values relative to the prior distributions (Fig. A3-2), which
-indicates the data did influence parameter estimates beyond the
-information from the uninformative priors. Ecosystem functional response
-was similar among treatments (Table 1; Fig. 3B), but there is evidence
-that the
+indicates the data did influence parameter estimates. Ecosystem
+functional response was similar among treatments (Table 1; Fig. 3B), but
+there is evidence that the
 \new{slope for the drought treatment is greater than the slope for the control treatment.}
@@
 A1-4).
-By definition, thsee rare species contribute little to ANPP and thus any compositional changes have little influence on ecosystem functional response.}
+By definition, these rare species contribute little to ANPP and thus any compositional changes have little influence on ecosystem functional response.}
 
-\section{DISCUSSION}\label{discussion}
+\hypertarget{discussion}{%
+\section{DISCUSSION}\label{discussion}}
 
@@
 community composition shifts to favor species better able to take
-advantage of, or cope with, new resource levels \citep{Smith2009}.
-Previous work has shown that community compositional shifts can be both
-rapid, on the order of years \citep{Hoover2014}, and slow, on order of
-decades \citep{Knapp2012, Wilcox2016}. A lingering question is how the
-time scales of ecosystem response and community change vary among
-ecosystems. Precipitation manipulation experiments can help answer this
-question, especially if they push water availability outside the
-historical range of variability for long periods.
+advantage of, or cope with, new resource levels (Smith, Knapp, and
+Collins 2009). Previous work has shown that community compositional
+shifts can be either rapid, on the order of years (Hoover, Knapp, and
+Smith 2014), or slow, on the order of decades (Knapp, Briggs, and Smith
+2012; Wilcox et al. 2016). A lingering question is how the time scales
+of ecosystem response and community change vary among ecosystems.
+Precipitation manipulation experiments can help answer this question,
+especially if they push water availability outside the historical range
+of variability for long periods.
 
 The results of our five year experiment in a sagebrush steppe
-\new{conforms to one} of our four predictions. Ecosystem functional
+\new{conform to one} of our four predictions. Ecosystem functional
 response under chronic \new{drought} was different from the control
@@
 4), representing the top right scenario in Fig. 1.
-\new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots is consistent with a stronger response of productivity to additional water in this ecosystem under chronic drought.
-This suggests that if average soil moisture were pushed consistently lower than currently observed ambient conditions, there would be a stronger relationship between precipitation and productivity in this system.}
+\new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots indicates increased sensitivity to water availability under chronic drought.
+A strict interpretation of this result implies that if average soil moisture were pushed consistently lower than currently observed ambient conditions, there would be a stronger relationship between precipitation and productivity in this system.}
 
@@
 We can think of three reasons. Two are limitations of our study, and one
-is the life history traits of the species in our focal communities. We
-first discuss the potential limitations of our study, and then discuss
-the biological explanation.
+invokes the life history traits of the species at our study site.
 
-First, it could be that our precipitation manipulations were not large
-enough to induce a response. That is, a 50\% decrease in any given year
-may not be abnormal given our site's historical range of variability
-\citep{Knapp2017}. We cannot definitively rule out this possibility, but
-we have reason to believe our treatments \emph{should} have been large
-enough. Using the methods described by \citet{Lemoine2016}, we
-calculated the percent reduction and increase of \new{water-year}
-precipitation necessary to reach the 1\% and 99\% extremes of the
-historical precipitation regime at our site (Fig. A5-1).
+First, our precipitation manipulations may not have been large enough to
+induce a response. A 50\% decrease in any given year may not be abnormal
+given our site's historical range of variability (Knapp et al. 2017). We
+cannot definitively rule out this possibility, but we have reason to
+believe our treatments \emph{should} have been large enough. Using the
+methods described by Lemoine et al. (2016), we calculated the percent
+reduction and increase of \new{water-year} precipitation necessary to
+reach the 1\% and 99\% extremes of the historical precipitation regime
+at our site (Fig. A5-1).
 \new{The 1\% quantile of water-year precipitation at our site is 78 mm, a 26\% reduction from the mean, and the 99\% quantile is 545 mm, a 84\% increase from mean growing season precipitation (Appendix 6).
@@
 only the cumulative March-June soil moisture covariate we included in
-our statistical model. \new{For example,} \citet{LaPierre2016}
-\new{found that site-scale ANPP is better predicted by abiotic factors such as nutrient availability than precipitation.}
-Moreover, temperature can impact ANPP directly \citep{Epstein1997} and
-by exacerbating the effects of soil moisture \citep{DeBoeck2011}.
-Measurements of soil moisture likely contain a signal of temperature,
-through its impact on evaporation and infiltration, but the measurements
-will not capture the direct effect of temperature on metabolic and
-physiological processes. We also did not redistribute snow across our
-treatments in the winter, and snow melt may spur early spring growth.
-Failure to account for potentially important covariates could explain
-the within-year spread of ANPP (Fig 2C, Fig. 3B) and the resulting
-uncertain relationship we observed between soil moisture and ANPP across
-all treatments (Table 1, Fig. A3-2). Unfortunately, we did not collect
-data on additional factors because our focus was on precipitation.
+our statistical model. \new{For example,} La Pierre et al. (2016)
+\new{found that site-scale ANPP is better predicted by nutrient availability than precipitation.}
+Moreover, temperature can impact ANPP directly (Epstein, Lauenroth, and
+Burke 1997) and by exacerbating the effects of soil moisture (De Boeck
+et al. 2011). Measurements of soil moisture likely contain a signal of
+temperature, through its impact on evaporation and infiltration, but the
+measurements will not capture the direct effect of temperature on
+metabolic and physiological processes. We also did not redistribute snow
+across our treatments in the winter, and snow melt may spur early spring
+growth. Failure to account for potentially important covariates could
+explain the within-year spread of ANPP (Fig 2C, Fig. 3B) and the
+resulting uncertain relationship we observed between soil moisture and
+ANPP across all treatments (Table 1, Fig. A3-2).
 
@@
 high water availability. In so doing, the dominant species in our
-ecosystem avoid ``boom and bust'' cycles, which corresponds to the weak
+ecosystem avoid ``boom and bust'' cycles, which may explain the weak
 effect of soil moisture on ANPP (i.e., the Bayesian credible intervals
@@
 the perennial grasses in our focal ecosystem avoid drought stress by
-growing early in the growing season \citep[A.R. Kleinhesselink, personal
-observation]{Blaisdell1958}. Furthermore, the dominant shrub in our
-focal ecosystem, \emph{Artemisia tripartita}, has access to water deep
-in the soil profile thanks to a deep root system
-\citep{Kulmatiski2017a}.
+growing early in the growing season (Blaisdell 1958, A.R.
+Kleinhesselink, personal observation). Furthermore, the dominant shrub
+in our focal ecosystem, \emph{Artemisia tripartita}, has access to water
+deep in the soil profile thanks to a deep root system (Kulmatiski et al.
+2017).
 \new{The dominance of our site by the shrub \emph{A. tripartita} (Appendix 1) may explain why our results do not conform to broader patterns of grassland sensitivity to precipitation manipulations}
-\citep[e.g.,][]{Huxman2004, Knapp2015, Wilcox2017}.
+(e.g., Huxman et al. 2004; Knapp et al. 2015; Wilcox et al. 2017).
 
-\new{We found that ANPP became more sensitive to the drought over time (Fig.
-3C).
-In other words, relative to the effect of the drought shelters on soil moisture, ANPP was more reduced by the drought conditions later in the experiment than early on in the experiment.
+\new{We found that ANPP became more sensitive to the drought treatment over time (Fig. 3C).
 We interepret this increase in sensitivity as the effect of cumulative impacts of the drought on dominant species.
 Plants may not have shown a large response in terms of ANPP in the first years of the experiment if they could grow from stored carbohydrate reserves or if they abstained from flowering and reproduction.
-As the drought progressed, these same plants may have have started to reduce in size or die. 
-Given the long lived perennial species at this site, this increase in sensitivity may indicate a larger future change in community composition and ecosystem functional response. }
+As the drought progressed, these same plants may have have started to shrink or die. 
+Given the long-lived perennial species at this site, this increase in sensitivity may indicate a larger future change in community composition and ecosystem functional response. }
 
-\section{CONCLUSIONS}\label{conclusions}
+\hypertarget{conclusions}{%
+\section{CONCLUSIONS}\label{conclusions}}
 
@@
 \new{The length of the perturbation may be especially relevant in our focal ecosystem where the perennial species are long-lived, meaning compositional turnover may take many more years than we report on here.}
-Our results suggest compositional shifts would have the largest impact
-at high rainfall because the current community maintained consistent
-ecosystem functional response at very low water availability.
-\new{However, the increased sensitivity of ANPP to drought conditions suggests a shift in ecosystem functional response may be near, and simply takes more time than we observed here.}
 
-\section{ACKNOWLEDGEMENTS}\label{acknowledgements}
+\hypertarget{acknowledgements}{%
+\section{ACKNOWLEDGEMENTS}\label{acknowledgements}}
 
 We thank the many summer research technicians who collected the data
-reported in this paper and the USDA, U.S. Sheep Experiment Station
+reported in this paper and the USDA-ARS Sheep Experiment Station
 (Dubois, ID) for facilitating work on their property. We also thank
@@
 \caption{Results from statistical tests for clustering and dispersion of community composition among precipitation treatments.
-`adonis' tests whether treatments form unique clusters in multidimensial space; `betadisper' tests whether treatments have similar dispersion.
-For both tests, \emph{P} values greater than 0.05 indicate there is no support that the treatments differ.}
+`adonis' tests whether treatments form unique clusters in multidimensional space; `betadisper' tests whether treatments have similar dispersion.
+For both tests, \emph{P} values greater than 0.05 indicate there is no support for the hypothesis that the treatments differ.}
 \begingroup\normalsize
@@
   \hline
-2011 & adonis &  22 &   2 & 1.04 & 0.41 \\ 
+2011 & adonis &  22 &   2 & 1.04 & 0.43 \\ 
   2011 & betadisper &  22 &   2 & 2.22 & 0.14 \\ 
@@
   2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
-  2013 & adonis &  22 &   2 & 1.26 & 0.14 \\ 
+  2013 & adonis &  22 &   2 & 1.26 & 0.12 \\ 
   2013 & betadisper &  22 &   2 & 0.43 & 0.66 \\ 
-  2014 & adonis &  22 &   2 & 0.96 & 0.52 \\ 
+  2014 & adonis &  22 &   2 & 0.96 & 0.58 \\ 
   2014 & betadisper &  22 &   2 & 0.36 & 0.70 \\ 
-  2015 & adonis &  22 &   2 & 1.08 & 0.31 \\ 
+  2015 & adonis &  22 &   2 & 1.08 & 0.33 \\ 
   2015 & betadisper &  22 &   2 & 1.96 & 0.17 \\ 
-  2016 & adonis &  22 &   2 & 1.12 & 0.27 \\ 
+  2016 & adonis &  22 &   2 & 1.12 & 0.26 \\ 
   2016 & betadisper &  22 &   2 & 0.27 & 0.76 \\ 
@@
 
-\bibliography{/Users/atredenn/Dropbox/Bibliography/usses_water.bib}
+\hypertarget{references}{%
+\section*{REFERENCES}\label{references}}
+\addcontentsline{toc}{section}{REFERENCES}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Adler2012}{}%
+Adler, P B, H J Dalgleish, and S P Ellner. 2012. ``Forecasting plant
+community impacts of climate variability and change: when do competitive
+interactions matter?'' \emph{Journal of Ecology} 100:478--87.
+\url{https://doi.org/10.1111/j.1365-2745.2011.01930.x}.
+
+\leavevmode\hypertarget{ref-Avolio2015}{}%
+Avolio, Meghan L, Kimberly J La Pierre, Gregory R Houseman, Sally E
+Koerner, Emily Grman, Forest Isbell, David Samuel Johnson, and Kevin R
+Wilcox. 2015. ``A framework for quantifying the magnitude and
+variability of community responses to global change drivers.''
+\emph{Ecosphere} 6 (12):1--14.
+\url{https://doi.org/10.1890/ES15-00317.1}.
+
+\leavevmode\hypertarget{ref-Bates2015}{}%
+Bates, D., M. Maechler, B. Bolker, and S. Walker. 2015. ``Fitting linear
+mixed-effects models using lme4.''
+\url{https://doi.org/10.18637/jss.v067.i01}.
+
+\leavevmode\hypertarget{ref-Blaisdell1958}{}%
+Blaisdell, James P. 1958. ``Seasonal development and yield of native
+plants on the upper snake river plains and their relation to certain
+climate factors.'' \emph{United States Department of Agriculture
+Technical Bulletin No. 1190}.
+
+\leavevmode\hypertarget{ref-Bradford2014}{}%
+Bradford, John B., Daniel R. Schlaepfer, and William K. Lauenroth. 2014.
+``Ecohydrology of Adjacent Sagebrush and Lodgepole Pine Ecosystems: The
+Consequences of Climate Change and Disturbance.'' \emph{Ecosystems} 17
+(4):590--605. \url{https://doi.org/10.1007/s10021-013-9745-1}.
+
+\leavevmode\hypertarget{ref-Byrne2011}{}%
+Byrne, Kerry M., William K. Lauenroth, Peter B. Adler, and Christine M.
+Byrne. 2011. ``Estimating Aboveground Net Primary Production in
+Grasslands: A Comparison of Nondestructive Methods.'' \emph{Rangeland
+Ecology and Management} 64 (5):498--505.
+\url{https://doi.org/10.2111/REM-D-10-00145.1}.
+
+\leavevmode\hypertarget{ref-Dalgleish2011}{}%
+Dalgleish, Harmony J., David N. Koons, Mevin B. Hooten, Corey A. Moffet,
+and Peter B. Adler. 2011. ``Climate influences the demography of three
+dominant sagebrush steppe plants.'' \emph{Ecology} 92 (1):75--85.
+\url{https://doi.org/10.1890/10-0780.1}.
+
+\leavevmode\hypertarget{ref-DeBoeck2011}{}%
+De Boeck, Hans J., Freja E. Dreesen, Ivan A. Janssens, and Ivan Nijs.
+2011. ``Whole-system responses of experimental plant communities to
+climate extremes imposed in different seasons.'' \emph{New Phytologist}
+189 (3):806--17. \url{https://doi.org/10.1111/j.1469-8137.2010.03515.x}.
+
+\leavevmode\hypertarget{ref-Epstein1997}{}%
+Epstein, H. E., W. K. Lauenroth, and I. C. Burke. 1997. ``Effects of
+temperature and soil texture on ANPP in the U.S. Great plains.''
+\emph{Ecology} 78 (8):2628--31.
+\href{https://doi.org/10.1890/0012-9658(1997)078\%5B2628:EOTAST\%5D2.0.CO;2}{https://doi.org/10.1890/0012-9658(1997)078{[}2628:EOTAST{]}2.0.CO;2}.
+
+\leavevmode\hypertarget{ref-Gelman2009}{}%
+Gelman, Andrew, and Jennifer Hill. 2009. \emph{Data analysis using
+regression and multilevel/hierarchical models}. Cambridge: Cambridge
+University Press.
+
+\leavevmode\hypertarget{ref-Gherardi2015a}{}%
+Gherardi, Laureano A, and Osvaldo E Sala. 2015. ``Enhanced precipitation
+variability decreases grass- and increases shrub-productivity.''
+\emph{Proceedings of the National Academy of Sciences} 112
+(41):12735--40. \url{https://doi.org/10.1073/pnas.1506433112}.
+
+\leavevmode\hypertarget{ref-Gherardi2013}{}%
+Gherardi, Laureano A., and Osvaldo E Sala. 2013. ``Automated rainfall
+manipulation system: a reliable and inexpensive tool for ecologists.''
+\emph{Ecosphere} 4 (2):1--10.
+\url{https://doi.org/10.1890/ES12-00371.1}.
+
+\leavevmode\hypertarget{ref-Gherardi2015}{}%
+Gherardi, Laureano A., and Osvaldo E. Sala. 2015. ``Enhanced interannual
+precipitation variability increases plant functional diversity that in
+turn ameliorates negative impact on productivity.'' \emph{Ecology
+Letters} 18 (12):1293--1300. \url{https://doi.org/10.1111/ele.12523}.
+
+\leavevmode\hypertarget{ref-Hill1920}{}%
+Hill, Robert R. 1920. ``Charting Quadrats with a Pantograph.''
+\emph{Ecology} 1 (4). Ecological Society of America:270--73.
+\url{https://doi.org/10.2307/1929561}.
+
+\leavevmode\hypertarget{ref-Hoover2014}{}%
+Hoover, David L., Alan K. Knapp, and Melinda D. Smith. 2014.
+``Resistance and resilience of a grassland ecosystem to climate
+extremes.'' \emph{Ecology} 95 (9):2646--56.
+\url{https://doi.org/10.1890/13-2186.1}.
+
+\leavevmode\hypertarget{ref-Hsu2012}{}%
+Hsu, Joanna S., James Powell, and Peter B. Adler. 2012. ``Sensitivity of
+mean annual primary production to precipitation.'' \emph{Global Change
+Biology} 18 (7):2246--55.
+\url{https://doi.org/10.1111/j.1365-2486.2012.02687.x}.
+
+\leavevmode\hypertarget{ref-Huxman2004}{}%
+Huxman, Travis E., Melinda D. Smith, Philip A. Fay, Alan K. Knapp, M.
+Rebecca Shaw, Michael E. Loik, Stanley D. Smith, et al. 2004.
+``Convergence across biomes to a common rain-use efficiency.''
+\emph{Nature} 429 (6992):651--54.
+\url{https://doi.org/10.1038/nature02561}.
+
+\leavevmode\hypertarget{ref-Kleinhesselink2017b}{}%
+Kleinhesselink, Andrew R. 2017. ``Direct and indirect effects of climate
+change on plant populations and communities in sagebrush steppe.''
+Dissertation, Utah State University.
+
+\leavevmode\hypertarget{ref-Knapp2017}{}%
+Knapp, Alan K, Meghan L Avolio, Claus Beier, Charles J W Carroll, Scott
+L Collins, Jeffrey S Dukes, Lauchlan H Fraser, et al. 2017. ``Pushing
+precipitation to the extremes in distributed experiments:
+recommendations for simulating wet and dry years.'' \emph{Global Change
+Biology} 23 (5):1774--82. \url{https://doi.org/10.1111/gcb.13504}.
+
+\leavevmode\hypertarget{ref-Knapp2012}{}%
+Knapp, Alan K, John M Briggs, and Melinda D Smith. 2012. ``Community
+stability does not preclude ecosystem sensitivity to chronic resource
+alteration.'' \emph{Functional Ecology} 26 (6):1231--3.
+\url{https://doi.org/10.1111/j.1365-2435.2012.02053.x}.
+
+\leavevmode\hypertarget{ref-Knapp2015}{}%
+Knapp, Alan K., Charles J.W. Carroll, Elsie M. Denton, Kimberly J. La
+Pierre, Scott L. Collins, and Melinda D. Smith. 2015. ``Differential
+sensitivity to regional-scale drought in six central US grasslands.''
+\emph{Oecologia} 177 (4):949--57.
+\url{https://doi.org/10.1007/s00442-015-3233-6}.
+
+\leavevmode\hypertarget{ref-Kulmatiski2017a}{}%
+Kulmatiski, Andrew, Peter B. Adler, John M. Stark, and Andrew T.
+Tredennick. 2017. ``Water and nitrogen uptake are better associated with
+resource availability than root biomass.'' \emph{Ecosphere} 8 (3).
+\url{https://doi.org/10.1002/ecs2.1738}.
+
+\leavevmode\hypertarget{ref-LaPierre2016}{}%
+La Pierre, Kimberly J., Dana M. Blumenthal, Cynthia S. Brown, Julia A.
+Klein, and Melinda D. Smith. 2016. ``Drivers of Variation in Aboveground
+Net Primary Productivity and Plant Community Composition Differ Across a
+Broad Precipitation Gradient.'' \emph{Ecosystems} 19 (3):521--33.
+\url{https://doi.org/10.1007/s10021-015-9949-7}.
+
+\leavevmode\hypertarget{ref-Lemoine2016}{}%
+Lemoine, Nathan P, Justin Sheffield, Jeffrey S Dukes, Alan K Knapp, and
+Melinda D Smith. 2016. ``Terrestrial Precipitation Analysis (TPA): A
+resource for characterizing long-term precipitation regimes and
+extremes.'' \emph{Methods in Ecology and Evolution} 7 (11):1396--1401.
+\url{https://doi.org/10.1111/2041-210X.12582}.
+
+\leavevmode\hypertarget{ref-Oksanen2016}{}%
+Oksanen, Jari. 2016. ``Vegan: ecological diversity.''
+\url{https://doi.org/10.1029/2006JF000545}.
+
+\leavevmode\hypertarget{ref-Peters2012}{}%
+Peters, Debra P C, Jin Yao, Osvaldo E. Sala, and John P. Anderson. 2012.
+``Directional climate change and potential reversal of desertification
+in arid and semiarid ecosystems.'' \emph{Global Change Biology} 18
+(1):151--63. \url{https://doi.org/10.1111/j.1365-2486.2011.02498.x}.
+
+\leavevmode\hypertarget{ref-R2016}{}%
+R Core Team. 2016. \emph{R: A Language and Environment for Statistical
+Computing}. Vienna, Austria.
+
+\leavevmode\hypertarget{ref-Sala1992a}{}%
+Sala, O. E., W. K. Lauenroth, and W. J. Parton. 1992. ``Long-term soil
+water dynamics in the shortgrass steppe.'' \emph{Ecology} 73
+(4):1175--81. \url{https://doi.org/10.2307/1940667}.
+
+\leavevmode\hypertarget{ref-Seger1987}{}%
+Seger, J. 1987. ``What is bet-hedging?'' In \emph{Oxford Surveys in
+Evolutionary Biology}, edited by P.H. Harvey and L. Partridge, 182--211.
+Oxford: Oxford University Press.
+\url{http://ci.nii.ac.jp/naid/10004816581/\%5Cnpapers2://publication/uuid/BE92F8E9-1B87-4E50-85C5-AA134A0EBAF0}.
+
+\leavevmode\hypertarget{ref-Smith2011}{}%
+Smith, M.D. 2011. ``An ecological perspective on extreme climatic
+events: A synthetic definition and framework to guide future research.''
+\emph{Journal of Ecology} 99 (3):656--63.
+\url{https://doi.org/10.1111/j.1365-2745.2011.01798.x}.
+
+\leavevmode\hypertarget{ref-Smith2009}{}%
+Smith, Melinda D., Alan K. Knapp, and Scott L. Collins. 2009. ``A
+framework for assessing ecosystem dynamics in response to chronic
+resource alterations induced by global change.'' \emph{Ecology} 90
+(12):3279--89. \url{https://doi.org/10.1890/08-1815.1}.
+
+\leavevmode\hypertarget{ref-rstan2016}{}%
+Stan Development Team. 2016a. ``Rstan: the R interface to Stan, Version
+2.14.1.'' \url{http://mc-stan.org/rstan.html}.
+
+\leavevmode\hypertarget{ref-stan2016}{}%
+---------. 2016b. ``Stan: A C++ Library for Probability and Sampling,
+Version 2.14.1.''
+
+\leavevmode\hypertarget{ref-Wilcox2016}{}%
+Wilcox, Kevin R, John M Blair, Melinda D Smith, and Alan K Knapp. 2016.
+``Does ecosystem sensitivity to precipitation at the site-level conform
+to regional-scale predictions?'' \emph{Ecology} 97 (3):561--68.
+\url{https://doi.org/10.1890/15-1437}.
+
+\leavevmode\hypertarget{ref-Wilcox2017}{}%
+Wilcox, Kevin R., Zheng Shi, Laureano A. Gherardi, Nathan P. Lemoine,
+Sally E. Koerner, David L. Hoover, Edward Bork, et al. 2017.
+``Asymmetric responses of primary productivity to precipitation
+extremes: A synthesis of grassland precipitation manipulation
+experiments.'' \emph{Global Change Biology} 23 (10):4376--85.
+\url{https://doi.org/10.1111/gcb.13706}.
 

### 2018-01-22 manuscript/Appendix1.Rmd
ready for coauthors

@@
   - \usepackage{float}
-bibliography: ~/Dropbox/Bibliography/usses_water.bib
+bibliography: /Users/atredenn/Dropbox/Bibliography/usses_water.bib
 csl: components/ecology.csl

### 2018-01-22 manuscript/usses_water_ms.Rmd
ready for coauthors

@@
 
-  **Results.** Ecosystem functional response, defined as the relationship between soil moisture and ANPP was similar among irrigation and control treatments, but the drought treatment had a greater slope than the control treatment.
+  **Results.** Ecosystem functional response, defined as the relationship between soil moisture and ANPP, was similar among irrigation and control treatments, but the drought treatment had a greater slope than the control treatment.
 However, all estimates for the effect of soil moisture on ANPP overlapped zero, indicating the relationship is weak and uncertain regardless of treatment.
@@
 Within each group of plots we standardized the irrigation and drought effects on soil moisture relative to the ambient soil moisture conditions. 
-Specifically we subtracted the ambient daily soil moisture from the soil moisture measured within the drought and irrigation plots within each group and then divided by the standard deviation of daily soil moisture values measured in the ambient conditions $\left( \Delta x_{g,d,\text{trt.}} = (x_{g,d,\text{trt.}} - x_{g,d,\text{ambient}})/\text{s.d.}(x_{g,d,\text{ambient}}) \right)$ where $\Delta x$ is the standardized treatment effect and $x_{g,d,\text{trt.}}$ is the raw soil moisture measure for plot group $g$ on day $d$ and treatment $trt.$.
+Specifically, we subtracted the ambient daily soil moisture from the soil moisture measured within the drought and irrigation plots within each group and then divided by the standard deviation of daily soil moisture values measured in the ambient conditions $\left( \Delta x_{g,d,\text{trt.}} = (x_{g,d,\text{trt.}} - x_{g,d,\text{ambient}})/\text{s.d.}(x_{g,d,\text{ambient}}) \right)$ where $\Delta x$ is the standardized treatment effect and $x_{g,d,\text{trt.}}$ is the raw soil moisture measure for plot group $g$ on day $d$ and treatment $trt$.
 These transformations ensured that the treatment effects in each plot were appropriately scaled by the local ambient conditions within each plot group.}
@@
 This evidence comes from interpreting the posterior distributions of the slope offsets for the treatments.
-\new{From these distributions,} we calculate a \new{58}% one-tailed probability that the estimate is greater than zero for the irrigation treatment \new{and a 100\% one-tailed probability that the estimate is greater than zero for the drought treatment} (Fig.
+\new{From these distributions,} we calculate a \new{42}% one-tailed probability that the estimate is less than zero for the irrigation treatment \new{and a 100\% one-tailed probability that the estimate is greater than zero for the drought treatment} (Fig.
 3A, right panel).
@@
 
-The results of our five year experiment in a sagebrush steppe \new{conform to one} of our four predictions.
+The results of our five year experiment in a sagebrush steppe \new{appear to conform to one} of our four predictions.
 Ecosystem functional response under chronic \new{drought} was different from the control treatment, but community composition remained unchanged (Fig.
@@
 We also thank Susan Durham for clarifying our thinking on the statistical analyses and Kevin Wilcox for helpful discussions on analyzing community composition data.
-John Bradford and Caitlin Andrews generously provided SOILWAT model output,
+\new{John Bradford and Caitlin Andrews generously provided SOILWAT model output.}
 \new{Two anonymous reviewers and Elsie Denton provided thoughtful reviews that improved our paper.}
@@
 Treatment effects show the difference between the coefficients estimated in the treated plots and the control plots.
-Probabilities (``Pr $</>$ 0 ='') for each coefficient indicate the one-tailed probability that the coefficient is less than or greater than zero, depending on whether the median of the distribution is less than or greater than zero.
+Probabilities (``Pr $</>$ 0 ='') for each coefficient indicate the one-tailed probability that the coefficient is less than or greater than zero, depending on the median of the intercept offset distributions or our specific hypothesis for the slope offsets.
 The posterior densities were smoothed for visual clarity by increasing kernel bandwidth by a factor of five.

### 2018-01-22 manuscript/usses_water_ms.tex
ready for coauthors

@@
 \textbf{Results.} Ecosystem functional response, defined as the
-relationship between soil moisture and ANPP was similar among irrigation
-and control treatments, but the drought treatment had a greater slope
-than the control treatment. However, all estimates for the effect of
-soil moisture on ANPP overlapped zero, indicating the relationship is
-weak and uncertain regardless of treatment. There was also large spatial
-variation in ANPP within-years, which contributes to the uncertainty of
-the soil moisture effect. Plant community composition was remarkably
-stable over the course of the experiment and did not differ among
-treatments.
+relationship between soil moisture and ANPP, was similar among
+irrigation and control treatments, but the drought treatment had a
+greater slope than the control treatment. However, all estimates for the
+effect of soil moisture on ANPP overlapped zero, indicating the
+relationship is weak and uncertain regardless of treatment. There was
+also large spatial variation in ANPP within-years, which contributes to
+the uncertainty of the soil moisture effect. Plant community composition
+was remarkably stable over the course of the experiment and did not
+differ among treatments.
 
@@
 slope offsets for the treatments. \new{From these distributions,} we
-calculate a \new{58}\% one-tailed probability that the estimate is
-greater than zero for the irrigation treatment
+calculate a \new{42}\% one-tailed probability that the estimate is less
+than zero for the irrigation treatment
 \new{and a 100\% one-tailed probability that the estimate is greater than zero for the drought treatment}
@@
 The results of our five year experiment in a sagebrush steppe
-\new{conform to one} of our four predictions. Ecosystem functional
-response under chronic \new{drought} was different from the control
-treatment, but community composition remained unchanged (Fig. 3A, Fig.
-4), representing the top right scenario in Fig. 1.
+\new{appear to conform to one} of our four predictions. Ecosystem
+functional response under chronic \new{drought} was different from the
+control treatment, but community composition remained unchanged (Fig.
+3A, Fig. 4), representing the top right scenario in Fig. 1.
 \new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots indicates increased sensitivity to water availability under chronic drought.
@@
 Kevin Wilcox for helpful discussions on analyzing community composition
-data. John Bradford and Caitlin Andrews generously provided SOILWAT
-model output,
+data.
+\new{John Bradford and Caitlin Andrews generously provided SOILWAT model output.}
 \new{Two anonymous reviewers and Elsie Denton provided thoughtful reviews that improved our paper.}

### 2018-01-22 manuscript/usses_water_ms.Rmd
trivia

@@
 We conducted our precipitation manipulation experiment in a sagebrush steppe community at the USDA-ARS Sheep Experiment Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
-The plant community is dominated by the shrub \emph{Artimesia tripartita}, \new{the perennial forb \emph{Balsamorhiza sagittata}}, and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata} \new{(see Appendix 1 for rank abundance curves)}.
+The plant community is dominated by the shrub \emph{Artemisia tripartita}, \new{the perennial forb \emph{Balsamorhiza sagittata}}, and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata} \new{(see Appendix 1 for rank abundance curves)}.
 During the period of our experiment (2011 -- 2016), mean annual precipitation was 250 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
@@
 Analyzing annual species on their own produced similar results.
-There is some evidence that the annual community changed in response to treatment in two years (Fig.
-A1-6), but these responses appear to come from very rare species since our analysis weights common and rare species equally (Fig.
+There is some evidence that the annual community changed in response to treatment in two years (Table A1-1; Fig.
+A1-5), but these responses appear to come from very rare species since our analysis weights common and rare species equally (Fig.
 A1-4).
@@
 \begin{description}
-\item [Appendix 1.] \new{Additional information on plant community composition and dynamics, Table A1-1, Table A1-2, Table A1-3, Fig. A1-1, Fig. A1-2, Fig. A1-3, Fig. A1-4, Fig. A1-5, and Fig. A1-6.}
+\item [Appendix 1.] \new{Additional information on plant community composition and dynamics, Table A1-1, Fig. A1-1, Fig. A1-2, Fig. A1-3, Fig. A1-4, and Fig. A1-5.}
 \item [Appendix 2.] \new{Additional details on soil moisture modeling with SOILWAT and Fig. A2-1.}

### 2018-01-22 manuscript/usses_water_ms.tex
trivia

@@
 sea level. The plant community is dominated by the shrub
-\emph{Artimesia tripartita},
+\emph{Artemisia tripartita},
 \new{the perennial forb \emph{Balsamorhiza sagittata}}, and three

### 2018-01-23 manuscript/usses_water_ms.Rmd
adds final PBA edits for R1

@@
 In the near term, ecosystem functioning such as ANPP will reflect the physiological responses of individual species to the manipulated resource level.
-For example, ANPP may decline under simulated drought because the initial community consisted of drought-intolerant species [@Hoover2014].
-Ecosystem functioning may recover over longer time spans as new species colonize or initial species reorder in relative abundance.
-It is also possible that ecosystem functioning shifts to a new mean state, reflecting the suite of species in the new community [@Knapp2012].
+For example, ANPP may initially decline under simulated drought because the community consisted of drought-intolerant species [@Hoover2014], but functioning may recover over longer time spans as new species colonize or resident species reorder in relative abundance.
+It is also possible that ecosystem functioning eventually shifts to a new mean state, reflecting the suite of species in the new community [@Knapp2012].
 
-Experimental manipulations of limiting resources, like precipitation, offer a route to understanding how ecosystems will respond to resource levels that fall outside the historical range of variability [@Avolio2015; @Gherardi2015; @Knapp2017].
+Experimental manipulations of limiting resources, like precipitation, offer a way to test how ecosystems will respond to resource levels that fall outside the historical range of variability [@Avolio2015; @Gherardi2015; @Knapp2017].
 Altering the amount of precipitation over many years should provide insight into the time scales at which water-limited ecosystems respond to chronic resource alteration.
@@
 Third, the ecosystem functional response might be constant but community composition changes (Fig. 1, bottom left).
-In this case, changes in species' identities and/or abundances occur in response to altered precipitation levels and species more suited to the new conditions compensate for reduced function of initial species.
+In this case, changes in species' identities and/or abundances occur in response to altered precipitation levels and species more suited to the new conditions compensate for reduced function of initial residents.
 Fourth, and last, both ecosystem functional response and community composition could change (Fig. 1, bottom right).
@@
 We avoided areas on steep hill slopes, areas with greater than 20\% cover of bare rock, and areas with greater than 10\% cover of the shrubs \emph{Purshia tridentata} and/or \emph{Amelanchier utahensis}.
-We established the new plots in pairs and randomly assigned each plot in a pair to receive a "drought" or "irrigation" treatment.
+We established the new plots in pairs and randomly assigned each plot in a pair to receive the "drought" or "irrigation" treatment.
 \new{Thus, our experiment consisted of $n=14$ control plots, $n=8$ irrigation plots, and $n=8$ irrigation plots, for a total of 30 plots.}
@@
 We could only predict soil moisture in the treated plots on days for which we took at least one ambient soil moisture measurement.
-This procedure allowed us to predict daily soil moisture conditions for all plots, even on those days when some of our direct treatment measurements were missing due to malfunction of the sensors or dataloggers.
+This procedure allowed us to predict daily soil moisture conditions for all plots, even on those days when some of our direct treatment measurements were missing due to malfunction of the sensors or data loggers.
 See @Kleinhesselink2017b for more details on our approach to estimating soil moisture.
@@
 We also took readings in 12 (2015), 15 (2012, 2013, 2014), or 16 (2016) calibration plots adjacent to the experimental site, in which we harvested all aboveground biomass produced in the current year (we excluded litter and standing dead material), dried it to a constant weight at 60$^{\circ}$C, and weighed it to estimate ANPP.
-We harvested at peak green biomass each year, typically in late June.
+We made radiometer measurements and harvested at peak green biomass each year, typically in late June.
 
 For each plot and year, we averaged the four readings for each wavelength and then calculated a greenness index based on the same bands used to calculate NDVI using the MODIS (Moderate Resolution Imaging Spectroradiometer) and AVHRR (Advanced Very High Resolution Radiometer) bands for NDVI.
-We regressed the greenness index against the dry biomass weight from the ten calibration plots to convert the greenness index to ANPP.
+We regressed the greenness index against the dry biomass weight from the calibration plots to convert the greenness index to ANPP.
 We fit regressions to a MODIS-based index and an AVHRR-based index for each year and retained the regression with the better fit based on $R^2$ values.
@@
 This exclosure included six of the control plots and all of the treatment plots, for a total of 22 plots.
-The other eight control plots are in other pastures, which would add a spatial bias to our community analysis focused on temporal dynamics.}
+The other eight control plots are in other pastures. 
+Including these plots would add spatial variation in composition, complicating our goal of describing temporal trends in composition.}
 
@@
 Based on our own observations and previous work at our study site [@Blaisdell1958; @Dalgleish2011; @Adler2012], we chose to use cumulative volumetric water content from March through June as our metric of soil moisture (hereafter referred to as `VWC').
-To achieve this goal, we fit a generalized linear mixed effects regression model with log(ANPP) as the response variable and VWC and treatment as fixed effects.
+We fit a generalized linear mixed effects regression model with log(ANPP) as the response variable and VWC and treatment as fixed effects.
 Plot and year of treatment were included as random effects to account for non-independence of observations, as described below.
@@
 
-The results of our five year experiment in a sagebrush steppe \new{appear to conform to one} of our four predictions.
-Ecosystem functional response under chronic \new{drought} was different from the control treatment, but community composition remained unchanged (Fig.
+We found that ecosystem functional response under chronic \new{drought} was different from the control treatment, but community composition remained unchanged (Fig.
 3A, Fig.
@@
 
-\new{However, we do not want to over-interpret the significance of the slope offsets given that the full distributions of} the slopes of the VWC $\times$ productivity relationship, \new{not just the offsets}, were similar among treatments (Table 1).
+\new{However, we do not want to over-interpret the significance of the slope offsets given that the overall slopes} of the VWC $\times$ productivity relationship, \new{not just the offsets}, were similar among treatments (Table 1).
 We therefore conclude that ecosystem functional response is consistent (\new{similar values}) and weak (\new{all broadly overlapping zero}) across all precipitation treatments.
@@
 
-Second, ANPP at our site may be influenced by additional factors, not only the cumulative March-June soil moisture covariate we included in our statistical model.
+Second, ANPP at our site is likely influenced by additional factors, not only the cumulative March-June soil moisture covariate we included in our statistical model.
 \new{For example,} @LaPierre2016 \new{found that site-scale ANPP is better predicted by nutrient availability than precipitation.}
@@
 One strategy is bet hedging, where species forego short-term gains to reduce the variance of long-term success [@Seger1987].
-In other words, species follow the same conservative strategy every year, designed to minimize response to environmental conditions.
+In other words, species follow the same conservative strategy every year, designed to minimize losses during unfavorable periods.
 The dry and variable environment of the sagebrush steppe has likely selected for bet hedging species that can maintain function at low water availability and have weak responses to high water availability.
@@
 
-Our results suggest the species in our focal plant community are insensitive to to changes in precipitation regime, \new{at least over the five years of our experiment}. 
+Our results suggest that the species in our focal plant community are insensitive to to changes in precipitation regime, \new{at least over the five years of our experiment}. 
 Such insensitivity can buffer species against precipitation variability in this semi-arid ecosystem, making them successful in the long run.

### 2018-01-24 manuscript/Appendix4.Rmd

@@
 
-The first seven rows of the fixed effects design matrix (six control plots and one drought plot) is as follows:
+An example of six rows of the fixed effects design matrix (three control plots and three drought plots) is as follows:
 
@@
 colnames(x) <- c("Int", "VWC", "Drought", "Irrig", "Drought:VWC", "Irrig:VWC")
-head(x, 7)
+x[12:17,]
 ```
@@
 \noindent{}The random year effects ($\boldsymbol{\eta}$) are modeled as intercept offsets centered on zero with a shared variance ($\sigma_{yr}$): $\boldsymbol{\gamma} \sim \text{Normal} \left( 0, \sigma_{yr}  \right)$.
+$\Sigma$ is the variance-covariance matrix for the intercept and slope offsets for each plot, which is defined as the among plot variance ($\sigma^2_{plt}$) times the matrix $\textbf{R}$ that defines the correlation between intercept and slope offsets.
 
@@
 
-# Figures
-
 ```{r scatterplot, fig.cap="Scatterplot, on the arithmetic scale, of the data and model estimates shown as solid lines.
-Model estimates come from treatment level coefficients (colored lines).", echo=FALSE, message=FALSE, warning=FALSE}
+Model estimates come from treatment level coefficients (colored lines).", echo=FALSE, message=FALSE, warning=FALSE, fig.pos="h"}
 library(tidyverse)    # Data munging
@@
 ```
-
-\newpage{}
-
-# References

### 2018-01-24 manuscript/Appendix5.Rmd

@@
 slope_probs$prob_text[1] <- paste("Pr > 0 =", slope_probs$probs[1]) # switch direction of drought intercpet since greater than 0
-slope_probs$xpos <- c(-2.8, -0.6, -2.6, -0.6)
+slope_probs$xpos <- c(-2.8, -0.2, -2.6, -0.2)
 slope_probs$ypos <- c(1.3,1.3,2.55,2.7)

### 2018-01-24 manuscript/usses_water_ms.tex

@@
 ANPP will reflect the physiological responses of individual species to
-the manipulated resource level. For example, ANPP may decline under
-simulated drought because the initial community consisted of
-drought-intolerant species (Hoover, Knapp, and Smith 2014). Ecosystem
+the manipulated resource level. For example, ANPP may initially decline
+under simulated drought because the community consisted of
+drought-intolerant species (Hoover, Knapp, and Smith 2014), but
 functioning may recover over longer time spans as new species colonize
-or initial species reorder in relative abundance. It is also possible
-that ecosystem functioning shifts to a new mean state, reflecting the
-suite of species in the new community (Knapp, Briggs, and Smith 2012).
+or resident species reorder in relative abundance. It is also possible
+that ecosystem functioning eventually shifts to a new mean state,
+reflecting the suite of species in the new community (Knapp, Briggs, and
+Smith 2012).
 
 Experimental manipulations of limiting resources, like precipitation,
-offer a route to understanding how ecosystems will respond to resource
-levels that fall outside the historical range of variability (Avolio et
-al. 2015; Laureano A. Gherardi and Sala 2015; Knapp et al. 2017).
-Altering the amount of precipitation over many years should provide
-insight into the time scales at which water-limited ecosystems respond
-to chronic resource alteration. We propose four alternative scenarios
-for the effect of precipitation manipulation on the ecosystem functional
+offer a way to test how ecosystems will respond to resource levels that
+fall outside the historical range of variability (Avolio et al. 2015;
+Laureano A. Gherardi and Sala 2015; Knapp et al. 2017). Altering the
+amount of precipitation over many years should provide insight into the
+time scales at which water-limited ecosystems respond to chronic
+resource alteration. We propose four alternative scenarios for the
+effect of precipitation manipulation on the ecosystem functional
 response to soil moisture based on the Hierarchical Response Framework
@@
 more suited to the new conditions compensate for reduced function of
-initial species. Fourth, and last, both ecosystem functional response
+initial residents. Fourth, and last, both ecosystem functional response
 and community composition could change (Fig. 1, bottom right). New
@@
 \emph{Amelanchier utahensis}. We established the new plots in pairs and
-randomly assigned each plot in a pair to receive a ``drought'' or
+randomly assigned each plot in a pair to receive the ``drought'' or
 ``irrigation'' treatment.
@@
 plots, even on those days when some of our direct treatment measurements
-were missing due to malfunction of the sensors or dataloggers. See
+were missing due to malfunction of the sensors or data loggers. See
 Kleinhesselink (2017) for more details on our approach to estimating
@@
 (we excluded litter and standing dead material), dried it to a constant
-weight at 60\(^{\circ}\)C, and weighed it to estimate ANPP. We harvested
-at peak green biomass each year, typically in late June.
+weight at 60\(^{\circ}\)C, and weighed it to estimate ANPP. We made
+radiometer measurements and harvested at peak green biomass each year,
+typically in late June.
 
@@
 bands for NDVI. We regressed the greenness index against the dry biomass
-weight from the ten calibration plots to convert the greenness index to
+weight from the calibration plots to convert the greenness index to
 ANPP. We fit regressions to a MODIS-based index and an AVHRR-based index
@@
 through June as our metric of soil moisture (hereafter referred to as
-`VWC'). To achieve this goal, we fit a generalized linear mixed effects
-regression model with log(ANPP) as the response variable and VWC and
-treatment as fixed effects. Plot and year of treatment were included as
-random effects to account for non-independence of observations, as
-described below. We log-transformed ANPP to reduce heteroscedasticity.
-Both log(ANPP) and VWC were standardized to have mean 0 and unit
-variance before fitting the model {[}i.e.,
-\((x_i - \bar{x})/\sigma_x\){]}.
+`VWC'). We fit a generalized linear mixed effects regression model with
+log(ANPP) as the response variable and VWC and treatment as fixed
+effects. Plot and year of treatment were included as random effects to
+account for non-independence of observations, as described below. We
+log-transformed ANPP to reduce heteroscedasticity. Both log(ANPP) and
+VWC were standardized to have mean 0 and unit variance before fitting
+the model {[}i.e., \((x_i - \bar{x})/\sigma_x\){]}.
 
@@
 
-The results of our five year experiment in a sagebrush steppe
-\new{appear to conform to one} of our four predictions. Ecosystem
-functional response under chronic \new{drought} was different from the
-control treatment, but community composition remained unchanged (Fig.
-3A, Fig. 4), representing the top right scenario in Fig. 1.
+We found that ecosystem functional response under chronic \new{drought}
+was different from the control treatment, but community composition
+remained unchanged (Fig. 3A, Fig. 4), representing the top right
+scenario in Fig. 1.
 \new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots indicates increased sensitivity to water availability under chronic drought.
@@
 
-\new{However, we do not want to over-interpret the significance of the slope offsets given that the full distributions of}
-the slopes of the VWC \(\times\) productivity relationship,
+\new{However, we do not want to over-interpret the significance of the slope offsets given that the overall slopes}
+of the VWC \(\times\) productivity relationship,
 \new{not just the offsets}, were similar among treatments (Table 1). We
@@
 
-Second, ANPP at our site may be influenced by additional factors, not
+Second, ANPP at our site is likely influenced by additional factors, not
 only the cumulative March-June soil moisture covariate we included in
@@
 long-term success (Seger 1987). In other words, species follow the same
-conservative strategy every year, designed to minimize response to
-environmental conditions. The dry and variable environment of the
-sagebrush steppe has likely selected for bet hedging species that can
-maintain function at low water availability and have weak responses to
-high water availability. In so doing, the dominant species in our
-ecosystem avoid ``boom and bust'' cycles, which may explain the weak
-effect of soil moisture on ANPP (i.e., the Bayesian credible intervals
-for the slopes overlapping zero; Table 1).
+conservative strategy every year, designed to minimize losses during
+unfavorable periods. The dry and variable environment of the sagebrush
+steppe has likely selected for bet hedging species that can maintain
+function at low water availability and have weak responses to high water
+availability. In so doing, the dominant species in our ecosystem avoid
+``boom and bust'' cycles, which may explain the weak effect of soil
+moisture on ANPP (i.e., the Bayesian credible intervals for the slopes
+overlapping zero; Table 1).
 
@@
 
-Our results suggest the species in our focal plant community are
+Our results suggest that the species in our focal plant community are
 insensitive to to changes in precipitation regime,

### 2018-01-24 manuscript/usses_water_ms.Rmd
tiny edits from ARK

@@
 Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
-In 2007, we relocated 14 of the original quadrats in permenanent livestock exclosures, \new{which we used} as control plots (i.e. ambient precipitation) in the experiment described below.
-\new{We used the relocated plots as our controls because collecting demographic data is time consuming and it was already being collected in these plots for other studies.}
+In 2007, we located 14 of the original quadrats in permenanent livestock exclosures, \new{which we used} as control plots (i.e. ambient precipitation) in the experiment described below.
+\new{We used the original plots as our controls because collecting demographic data is time consuming and it was already being collected in these plots for other studies.}
 
@@
 
-\new{Following the above procedure, we still lacked soil moisture data for March 2012 (observations did not start unitl April) and for a string of days in 2013 during which the soil moisture probes failed to take readings.
+\new{Following the above procedure, we still lacked soil moisture data for March 2012 (observations did not start until April 2012) and for a string of days in 2013 during which the soil moisture probes failed to take readings.
 To fill in these gaps, we used a version of the SOILWAT soil moisture model} [@Sala1992a] \new{that has been specifically designed for semiarid shrublands and grasslands} [@Bradford2014].

### 2018-01-24 manuscript/usses_water_ms.tex
tiny edits from ARK

@@
 % 
-
 \begin{abstract}
@@
 
-\hypertarget{introduction}{%
-\section{INTRODUCTION}\label{introduction}}
+\section{INTRODUCTION}\label{introduction}
 
 \new{Water availability is a major driver of annual net primary productivity (ANPP) in grassland ecosystems}
-(Huxman et al. 2004; Hsu, Powell, and Adler 2012).
+\citep{Huxman2004, Hsu2012}.
 \new{Therefore, projected changes in precipitation regimes, associated with global climate change, will impact grassland ecosystem functioning.}
@@
 responses can then be used to infer how ANPP may change under future
-precipitation regimes (e.g., Hsu, Powell, and Adler 2012). A problem
-with this approach is that it requires extrapolation if future
-precipitation falls outside the historical range of variability (Smith
-2011; Peters et al. 2012). For example, the soil moisture-ANPP
+precipitation regimes \citep[e.g.,][]{Hsu2012}. A problem with this
+approach is that it requires extrapolation if future precipitation falls
+outside the historical range of variability
+\citep{Smith2011, Peters2012}. For example, the soil moisture-ANPP
 relationship may be linear within the historical range of interannual
 variation, but could saturate at higher levels of soil moisture.
-Saturating relationships are actually common (Hsu, Powell, and Adler
-2012; Laureano A Gherardi and Sala 2015), perhaps because other
-resources, like nitrogen, become more limiting in wet years than dry
-years. Failure to accurately estimate the curvature of the soil
-moisture-ANPP relationship will lead to over- or underprediction of ANPP
-under extreme precipitation (Peters et al. 2012).
+Saturating relationships are actually common
+\citep{Hsu2012, Gherardi2015a}, perhaps because other resources, like
+nitrogen, become more limiting in wet years than dry years. Failure to
+accurately estimate the curvature of the soil moisture-ANPP relationship
+will lead to over- or underprediction of ANPP under extreme
+precipitation \citep{Peters2012}.
 
@@
 response to water availability because different species have different
-physiological thresholds. Smith, Knapp, and Collins (2009) introduced
-the `Hierarchical Response Framework' for understanding the interplay of
-community composition and ecosystem functioning in response to long-term
-shifts in resources. In the near term, ecosystem functioning such as
-ANPP will reflect the physiological responses of individual species to
-the manipulated resource level. For example, ANPP may initially decline
+physiological thresholds. \citet{Smith2009} introduced the `Hierarchical
+Response Framework' for understanding the interplay of community
+composition and ecosystem functioning in response to long-term shifts in
+resources. In the near term, ecosystem functioning such as ANPP will
+reflect the physiological responses of individual species to the
+manipulated resource level. For example, ANPP may initially decline
 under simulated drought because the community consisted of
-drought-intolerant species (Hoover, Knapp, and Smith 2014), but
-functioning may recover over longer time spans as new species colonize
-or resident species reorder in relative abundance. It is also possible
-that ecosystem functioning eventually shifts to a new mean state,
-reflecting the suite of species in the new community (Knapp, Briggs, and
-Smith 2012).
+drought-intolerant species \citep{Hoover2014}, but functioning may
+recover over longer time spans as new species colonize or resident
+species reorder in relative abundance. It is also possible that
+ecosystem functioning eventually shifts to a new mean state, reflecting
+the suite of species in the new community \citep{Knapp2012}.
 
@@
 offer a way to test how ecosystems will respond to resource levels that
-fall outside the historical range of variability (Avolio et al. 2015;
-Laureano A. Gherardi and Sala 2015; Knapp et al. 2017). Altering the
-amount of precipitation over many years should provide insight into the
-time scales at which water-limited ecosystems respond to chronic
-resource alteration. We propose four alternative scenarios for the
-effect of precipitation manipulation on the ecosystem functional
-response to soil moisture based on the Hierarchical Response Framework
-(Fig. 1). We define `ecosystem functional response' as the relationship
-between available soil moisture and ANPP. The four scenarios are based
-on possible outcomes at the community (e.g., community composition) and
+fall outside the historical range of variability
+\citep{Avolio2015, Gherardi2015, Knapp2017}. Altering the amount of
+precipitation over many years should provide insight into the time
+scales at which water-limited ecosystems respond to chronic resource
+alteration. We propose four alternative scenarios for the effect of
+precipitation manipulation on the ecosystem functional response to soil
+moisture based on the Hierarchical Response Framework (Fig. 1). We
+define `ecosystem functional response' as the relationship between
+available soil moisture and ANPP. The four scenarios are based on
+possible outcomes at the community (e.g., community composition) and
 ecosystem (e.g., soil moisture-ANPP regression) levels.
@@
 All four outcomes are possible in any given ecosystem, but the time
-scales at which the different scenarios play out likely differ (Smith,
-Knapp, and Collins 2009; Wilcox et al. 2016; Knapp et al. 2017). To
-determine these time scales, we need to amass information on how quickly
-ecosystem functional responses change in different ecosystems. We also
-need to understand whether changes at the ecosystem level are driven by
-community level changes or individual level responses.
+scales at which the different scenarios play out likely differ
+\citep{Smith2009, Wilcox2016, Knapp2017}. To determine these time
+scales, we need to amass information on how quickly ecosystem functional
+responses change in different ecosystems. We also need to understand
+whether changes at the ecosystem level are driven by community level
+changes or individual level responses.
 
@@
 the amount of ambient precipitation by 50\%. To achieve this, we used a
-system of rain-out shelters and automatic irrigation (Gherardi and Sala
-2013). The rain-out shelters consisted of transparent acrylic shingles
-1-1.5 m above the ground that covered an area of \(2.5\times2\) m. The
-shingles intercepted approximately 50\% of incoming rainfall, which was
-channeled into 75 liter containers. Captured rainfall was then pumped
-out of the containers and sprayed on to the adjacent irrigation plot via
-two suspended sprinklers. Pumping was triggered by float switches once
-water levels reached about 20 liters. We disconnected the irrigation
-pumps each October and reconnected them each April. The rain-out
-shelters remained in place throughout the year.
+system of rain-out shelters and automatic irrigation
+\citep{Gherardi2013}. The rain-out shelters consisted of transparent
+acrylic shingles 1-1.5 m above the ground that covered an area of
+\(2.5\times2\) m. The shingles intercepted approximately 50\% of
+incoming rainfall, which was channeled into 75 liter containers.
+Captured rainfall was then pumped out of the containers and sprayed on
+to the adjacent irrigation plot via two suspended sprinklers. Pumping
+was triggered by float switches once water levels reached about 20
+liters. We disconnected the irrigation pumps each October and
+reconnected them each April. The rain-out shelters remained in place
+throughout the year.
 
@@
 snowfall). We fit the model using the \texttt{lme4::lmer()} function
-(Bates et al. 2015) in R (R Core Team 2016), with random effects for
-plot group and date. We weighted observations by the number of unique
-sensors or spot measurements that were taken in each plot on that day.
+\citep{Bates2015} in R \citep{R2016}, with random effects for plot group
+and date. We weighted observations by the number of unique sensors or
+spot measurements that were taken in each plot on that day.
 \new{We then used the model to predict the deviations from ambient conditions produced in the treated plots on each day of the experiment.  
@@
 were missing due to malfunction of the sensors or data loggers. See
-Kleinhesselink (2017) for more details on our approach to estimating
-soil moisture.
+\citet{Kleinhesselink2017b} for more details on our approach to
+estimating soil moisture.
 
@@
 To fill in these gaps, we used a version of the SOILWAT soil moisture model}
-(Sala, Lauenroth, and Parton 1992)
+\citep{Sala1992a}
 \new{that has been specifically designed for semiarid shrublands and grasslands}
-(Bradford, Schlaepfer, and Lauenroth 2014).
+\citep{Bradford2014}.
 \new{The model was parameterized using generic sagebrush steppe parameters and local soil texture, soil bulk density, and precipitation data}
-(Kleinhesselink 2017).
+\citep{Kleinhesselink2017b}.
 \new{We used the locally parameterized SOILWAT model to generate daily soil moisture predictions for the duration of the experiment, but only used SOILWAT predictions where there were gaps in our data (Fig.
@@
 
-\hypertarget{data-collection}{%
-\subsection{Data Collection}\label{data-collection}}
+\subsection{Data Collection}\label{data-collection}
 
 We estimated aboveground net primary productivity (ANPP) using a
-radiometer to relate ground reflectance to plant biomass (Byrne et al.
-2011). We recorded ground reflectance at four wavelengths, two
-associated with red reflectance (626 nm and 652 nm) and two associated
-with near-infrared reflectance (875 nm and 859 nm). At each plot in each
-year, we took four readings of ground reflectances at the above
-wavelengths. We also took readings in 12 (2015), 15 (2012, 2013, 2014),
-or 16 (2016) calibration plots adjacent to the experimental site, in
-which we harvested all aboveground biomass produced in the current year
-(we excluded litter and standing dead material), dried it to a constant
-weight at 60\(^{\circ}\)C, and weighed it to estimate ANPP. We made
-radiometer measurements and harvested at peak green biomass each year,
-typically in late June.
+radiometer to relate ground reflectance to plant biomass
+\citep{Byrne2011}. We recorded ground reflectance at four wavelengths,
+two associated with red reflectance (626 nm and 652 nm) and two
+associated with near-infrared reflectance (875 nm and 859 nm). At each
+plot in each year, we took four readings of ground reflectances at the
+above wavelengths. We also took readings in 12 (2015), 15 (2012, 2013,
+2014), or 16 (2016) calibration plots adjacent to the experimental site,
+in which we harvested all aboveground biomass produced in the current
+year (we excluded litter and standing dead material), dried it to a
+constant weight at 60\(^{\circ}\)C, and weighed it to estimate ANPP. We
+made radiometer measurements and harvested at peak green biomass each
+year, typically in late June.
 
@@
 Species composition data came from two sources: yearly census maps for
-each plot made using a pantograph (Hill 1920) and yearly counts of
+each plot made using a pantograph \citep{Hill1920} and yearly counts of
 annual species in each plot. From these sources, we determined the
@@
 Based on our own observations and previous work at our study site
-(Blaisdell 1958; Dalgleish et al. 2011; Adler, Dalgleish, and Ellner
-2012), we chose to use cumulative volumetric water content from March
-through June as our metric of soil moisture (hereafter referred to as
-`VWC'). We fit a generalized linear mixed effects regression model with
-log(ANPP) as the response variable and VWC and treatment as fixed
-effects. Plot and year of treatment were included as random effects to
-account for non-independence of observations, as described below. We
-log-transformed ANPP to reduce heteroscedasticity. Both log(ANPP) and
-VWC were standardized to have mean 0 and unit variance before fitting
-the model {[}i.e., \((x_i - \bar{x})/\sigma_x\){]}.
+\citep{Blaisdell1958, Dalgleish2011, Adler2012}, we chose to use
+cumulative volumetric water content from March through June as our
+metric of soil moisture (hereafter referred to as `VWC'). We fit a
+generalized linear mixed effects regression model with log(ANPP) as the
+response variable and VWC and treatment as fixed effects. Plot and year
+of treatment were included as random effects to account for
+non-independence of observations, as described below. We log-transformed
+ANPP to reduce heteroscedasticity. Both log(ANPP) and VWC were
+standardized to have mean 0 and unit variance before fitting the model
+{[}i.e., \((x_i - \bar{x})/\sigma_x\){]}.
 
@@
 precipitation-ANPP relationships often saturate with increasing
-precipitation (Hsu, Powell, and Adler 2012; Laureano A Gherardi and Sala
-2015).
+precipitation \citep{Hsu2012, Gherardi2015a}.
 
@@
 \new{We used a vague prior distribution for each $\beta$: $\boldsymbol{\beta} \sim \text{Normal}(0,5)$.}
-A full description of our model and associated R (R Core Team 2016) code
-is in Appendix 4.
+A full description of our model and associated R \citep{R2016} code is
+in Appendix 4.
 
@@
 estimates of all unknowns via the No-U-Turn Hamiltonian Monte Carlo
-sampler in Stan (Stan Development Team 2016b). We used the R package
-`rstan' (Stan Development Team 2016a) to link R (R Core Team 2016) to
-Stan. We obtained samples from the posterior distribution for all model
-parameters from four parallel MCMC chains run for 10,000 iterations,
-saving every \(10^{\text{th}}\) sample. Trace plots of all parameters
-were visually inspected to ensure well-mixed chains and convergence. We
-also made sure all scale reduction factors (\(\hat{R}\) values) were
-less than 1.1 (Gelman and Hill 2009).
+sampler in Stan \citep{stan2016}. We used the R package `rstan'
+\citep{rstan2016} to link R \citep{R2016} to Stan. We obtained samples
+from the posterior distribution for all model parameters from four
+parallel MCMC chains run for 10,000 iterations, saving every
+\(10^{\text{th}}\) sample. Trace plots of all parameters were visually
+inspected to ensure well-mixed chains and convergence. We also made sure
+all scale reduction factors (\(\hat{R}\) values) were less than 1.1
+\citep{Gelman2009}.
 
-\hypertarget{sensitivity-of-anpp-over-time}{%
 \subsubsection{Sensitivity of ANPP over
-time}\label{sensitivity-of-anpp-over-time}}
+time}\label{sensitivity-of-anpp-over-time}
 
@@
 \new{We define `sensitivity' as $\frac{ \text{ANPP}_{\text{control}} - \text{ANPP}_{\text{treatment}} } { \text{VWC}_{\text{ambient}} - \text{VWC}_{\text{treatment}} }$, following}
-Wilcox et al. (2017).
+\citet{Wilcox2017}.
 \new{This metric of sensitivity scales the difference between ANPP in the treated and control plots by the change in soil moisture in the treated plots. 
@@
 After calculating sensitivity for each drought and irrigation plot in each year, we regressed sensitivity against year of treatment using the \texttt{lm()} function in R}
-(R Core Team 2016).
+\citep{R2016}.
 \new{This analysis also allows us to link particularly sensitive treatment-years to changes in community composition.}
 
-\hypertarget{community-composition-over-time}{%
 \subsubsection{Community composition over
-time}\label{community-composition-over-time}}
+time}\label{community-composition-over-time}
 
@@
 if community composition overlapped, or not, among treatments in each
-year. We used the \texttt{vegan::metaMDS()} function (Oksanen 2016) to
-calculate Bray-Curtis distances and then to run the NMDS analysis. We
-used the \texttt{vegan::adonis()} function (Oksanen 2016) to perform
-permutational multivariate analysis of variance to test whether
+year. We used the \texttt{vegan::metaMDS()} function \citep{Oksanen2016}
+to calculate Bray-Curtis distances and then to run the NMDS analysis. We
+used the \texttt{vegan::adonis()} function \citep{Oksanen2016} to
+perform permutational multivariate analysis of variance to test whether
 treatment plots formed distinct groupings. To test whether treatment
 plots were equally dispersed, or not, we used the
-\texttt{vegan::betadisper()} function (Oksanen 2016).
+\texttt{vegan::betadisper()} function \citep{Oksanen2016}.
 
@@
 community composition shifts to favor species better able to take
-advantage of, or cope with, new resource levels (Smith, Knapp, and
-Collins 2009). Previous work has shown that community compositional
-shifts can be either rapid, on the order of years (Hoover, Knapp, and
-Smith 2014), or slow, on the order of decades (Knapp, Briggs, and Smith
-2012; Wilcox et al. 2016). A lingering question is how the time scales
-of ecosystem response and community change vary among ecosystems.
-Precipitation manipulation experiments can help answer this question,
-especially if they push water availability outside the historical range
-of variability for long periods.
+advantage of, or cope with, new resource levels \citep{Smith2009}.
+Previous work has shown that community compositional shifts can be
+either rapid, on the order of years \citep{Hoover2014}, or slow, on the
+order of decades \citep{Knapp2012, Wilcox2016}. A lingering question is
+how the time scales of ecosystem response and community change vary
+among ecosystems. Precipitation manipulation experiments can help answer
+this question, especially if they push water availability outside the
+historical range of variability for long periods.
 
@@
 generally, and sagebrush steppe specifically, are considered
-water-limited systems. \new{For example,} Huxman et al. (2004) and Knapp
-et al. (2015)
+water-limited systems. \new{For example,} \citet{Huxman2004} and
+\citet{Knapp2015}
 \new{showed that semi-arid sites are more sensitive to drought than mesic sites, and}
-Wilcox et al. (2017)
+\citet{Wilcox2017}
 \new{found that semi-arid sites are particularly sensitive to irrigation treatments.}
@@
 induce a response. A 50\% decrease in any given year may not be abnormal
-given our site's historical range of variability (Knapp et al. 2017). We
+given our site's historical range of variability \citep{Knapp2017}. We
 cannot definitively rule out this possibility, but we have reason to
 believe our treatments \emph{should} have been large enough. Using the
-methods described by Lemoine et al. (2016), we calculated the percent
+methods described by \citet{Lemoine2016}, we calculated the percent
 reduction and increase of \new{water-year} precipitation necessary to
@@
 only the cumulative March-June soil moisture covariate we included in
-our statistical model. \new{For example,} La Pierre et al. (2016)
+our statistical model. \new{For example,} \citet{LaPierre2016}
 \new{found that site-scale ANPP is better predicted by nutrient availability than precipitation.}
-Moreover, temperature can impact ANPP directly (Epstein, Lauenroth, and
-Burke 1997) and by exacerbating the effects of soil moisture (De Boeck
-et al. 2011). Measurements of soil moisture likely contain a signal of
-temperature, through its impact on evaporation and infiltration, but the
-measurements will not capture the direct effect of temperature on
-metabolic and physiological processes. We also did not redistribute snow
-across our treatments in the winter, and snow melt may spur early spring
-growth. Failure to account for potentially important covariates could
-explain the within-year spread of ANPP (Fig 2C, Fig. 3B) and the
-resulting uncertain relationship we observed between soil moisture and
-ANPP across all treatments (Table 1, Fig. A3-2).
+Moreover, temperature can impact ANPP directly \citep{Epstein1997} and
+by exacerbating the effects of soil moisture \citep{DeBoeck2011}.
+Measurements of soil moisture likely contain a signal of temperature,
+through its impact on evaporation and infiltration, but the measurements
+will not capture the direct effect of temperature on metabolic and
+physiological processes. We also did not redistribute snow across our
+treatments in the winter, and snow melt may spur early spring growth.
+Failure to account for potentially important covariates could explain
+the within-year spread of ANPP (Fig 2C, Fig. 3B) and the resulting
+uncertain relationship we observed between soil moisture and ANPP across
+all treatments (Table 1, Fig. A3-2).
 
@@
 hedging, where species forego short-term gains to reduce the variance of
-long-term success (Seger 1987). In other words, species follow the same
-conservative strategy every year, designed to minimize losses during
-unfavorable periods. The dry and variable environment of the sagebrush
-steppe has likely selected for bet hedging species that can maintain
-function at low water availability and have weak responses to high water
-availability. In so doing, the dominant species in our ecosystem avoid
-``boom and bust'' cycles, which may explain the weak effect of soil
-moisture on ANPP (i.e., the Bayesian credible intervals for the slopes
-overlapping zero; Table 1).
+long-term success \citep{Seger1987}. In other words, species follow the
+same conservative strategy every year, designed to minimize losses
+during unfavorable periods. The dry and variable environment of the
+sagebrush steppe has likely selected for bet hedging species that can
+maintain function at low water availability and have weak responses to
+high water availability. In so doing, the dominant species in our
+ecosystem avoid ``boom and bust'' cycles, which may explain the weak
+effect of soil moisture on ANPP (i.e., the Bayesian credible intervals
+for the slopes overlapping zero; Table 1).
 
@@
 the perennial grasses in our focal ecosystem avoid drought stress by
-growing early in the growing season (Blaisdell 1958, A.R.
-Kleinhesselink, personal observation). Furthermore, the dominant shrub
-in our focal ecosystem, \emph{Artemisia tripartita}, has access to water
-deep in the soil profile thanks to a deep root system (Kulmatiski et al.
-2017).
+growing early in the growing season \citep[A.R. Kleinhesselink, personal
+observation]{Blaisdell1958}. Furthermore, the dominant shrub in our
+focal ecosystem, \emph{Artemisia tripartita}, has access to water deep
+in the soil profile thanks to a deep root system
+\citep{Kulmatiski2017a}.
 \new{The dominance of our site by the shrub \emph{A. tripartita} (Appendix 1) may explain why our results do not conform to broader patterns of grassland sensitivity to precipitation manipulations}
-(e.g., Huxman et al. 2004; Knapp et al. 2015; Wilcox et al. 2017).
+\citep[e.g.,][]{Huxman2004, Knapp2015, Wilcox2017}.
 
@@
 precipitation alteration might reveal plant community shifts that we did
-not observe (e.g., Wilcox et al. 2016),
+not observe \citep[e.g.,][]{Wilcox2016},
 \new{and the increased sensitivity of ANPP to drought in the final years of our experiment may portend such a shift.}
@@
 
-\hypertarget{references}{%
-\section*{REFERENCES}\label{references}}
-\addcontentsline{toc}{section}{REFERENCES}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Adler2012}{}%
-Adler, P B, H J Dalgleish, and S P Ellner. 2012. ``Forecasting plant
-community impacts of climate variability and change: when do competitive
-interactions matter?'' \emph{Journal of Ecology} 100:478--87.
-\url{https://doi.org/10.1111/j.1365-2745.2011.01930.x}.
-
-\leavevmode\hypertarget{ref-Avolio2015}{}%
-Avolio, Meghan L, Kimberly J La Pierre, Gregory R Houseman, Sally E
-Koerner, Emily Grman, Forest Isbell, David Samuel Johnson, and Kevin R
-Wilcox. 2015. ``A framework for quantifying the magnitude and
-variability of community responses to global change drivers.''
-\emph{Ecosphere} 6 (12):1--14.
-\url{https://doi.org/10.1890/ES15-00317.1}.
-
-\leavevmode\hypertarget{ref-Bates2015}{}%
-Bates, D., M. Maechler, B. Bolker, and S. Walker. 2015. ``Fitting linear
-mixed-effects models using lme4.''
-\url{https://doi.org/10.18637/jss.v067.i01}.
-
-\leavevmode\hypertarget{ref-Blaisdell1958}{}%
-Blaisdell, James P. 1958. ``Seasonal development and yield of native
-plants on the upper snake river plains and their relation to certain
-climate factors.'' \emph{United States Department of Agriculture
-Technical Bulletin No. 1190}.
-
-\leavevmode\hypertarget{ref-Bradford2014}{}%
-Bradford, John B., Daniel R. Schlaepfer, and William K. Lauenroth. 2014.
-``Ecohydrology of Adjacent Sagebrush and Lodgepole Pine Ecosystems: The
-Consequences of Climate Change and Disturbance.'' \emph{Ecosystems} 17
-(4):590--605. \url{https://doi.org/10.1007/s10021-013-9745-1}.
-
-\leavevmode\hypertarget{ref-Byrne2011}{}%
-Byrne, Kerry M., William K. Lauenroth, Peter B. Adler, and Christine M.
-Byrne. 2011. ``Estimating Aboveground Net Primary Production in
-Grasslands: A Comparison of Nondestructive Methods.'' \emph{Rangeland
-Ecology and Management} 64 (5):498--505.
-\url{https://doi.org/10.2111/REM-D-10-00145.1}.
-
-\leavevmode\hypertarget{ref-Dalgleish2011}{}%
-Dalgleish, Harmony J., David N. Koons, Mevin B. Hooten, Corey A. Moffet,
-and Peter B. Adler. 2011. ``Climate influences the demography of three
-dominant sagebrush steppe plants.'' \emph{Ecology} 92 (1):75--85.
-\url{https://doi.org/10.1890/10-0780.1}.
-
-\leavevmode\hypertarget{ref-DeBoeck2011}{}%
-De Boeck, Hans J., Freja E. Dreesen, Ivan A. Janssens, and Ivan Nijs.
-2011. ``Whole-system responses of experimental plant communities to
-climate extremes imposed in different seasons.'' \emph{New Phytologist}
-189 (3):806--17. \url{https://doi.org/10.1111/j.1469-8137.2010.03515.x}.
-
-\leavevmode\hypertarget{ref-Epstein1997}{}%
-Epstein, H. E., W. K. Lauenroth, and I. C. Burke. 1997. ``Effects of
-temperature and soil texture on ANPP in the U.S. Great plains.''
-\emph{Ecology} 78 (8):2628--31.
-\href{https://doi.org/10.1890/0012-9658(1997)078\%5B2628:EOTAST\%5D2.0.CO;2}{https://doi.org/10.1890/0012-9658(1997)078{[}2628:EOTAST{]}2.0.CO;2}.
-
-\leavevmode\hypertarget{ref-Gelman2009}{}%
-Gelman, Andrew, and Jennifer Hill. 2009. \emph{Data analysis using
-regression and multilevel/hierarchical models}. Cambridge: Cambridge
-University Press.
-
-\leavevmode\hypertarget{ref-Gherardi2015a}{}%
-Gherardi, Laureano A, and Osvaldo E Sala. 2015. ``Enhanced precipitation
-variability decreases grass- and increases shrub-productivity.''
-\emph{Proceedings of the National Academy of Sciences} 112
-(41):12735--40. \url{https://doi.org/10.1073/pnas.1506433112}.
-
-\leavevmode\hypertarget{ref-Gherardi2013}{}%
-Gherardi, Laureano A., and Osvaldo E Sala. 2013. ``Automated rainfall
-manipulation system: a reliable and inexpensive tool for ecologists.''
-\emph{Ecosphere} 4 (2):1--10.
-\url{https://doi.org/10.1890/ES12-00371.1}.
-
-\leavevmode\hypertarget{ref-Gherardi2015}{}%
-Gherardi, Laureano A., and Osvaldo E. Sala. 2015. ``Enhanced interannual
-precipitation variability increases plant functional diversity that in
-turn ameliorates negative impact on productivity.'' \emph{Ecology
-Letters} 18 (12):1293--1300. \url{https://doi.org/10.1111/ele.12523}.
-
-\leavevmode\hypertarget{ref-Hill1920}{}%
-Hill, Robert R. 1920. ``Charting Quadrats with a Pantograph.''
-\emph{Ecology} 1 (4). Ecological Society of America:270--73.
-\url{https://doi.org/10.2307/1929561}.
-
-\leavevmode\hypertarget{ref-Hoover2014}{}%
-Hoover, David L., Alan K. Knapp, and Melinda D. Smith. 2014.
-``Resistance and resilience of a grassland ecosystem to climate
-extremes.'' \emph{Ecology} 95 (9):2646--56.
-\url{https://doi.org/10.1890/13-2186.1}.
-
-\leavevmode\hypertarget{ref-Hsu2012}{}%
-Hsu, Joanna S., James Powell, and Peter B. Adler. 2012. ``Sensitivity of
-mean annual primary production to precipitation.'' \emph{Global Change
-Biology} 18 (7):2246--55.
-\url{https://doi.org/10.1111/j.1365-2486.2012.02687.x}.
-
-\leavevmode\hypertarget{ref-Huxman2004}{}%
-Huxman, Travis E., Melinda D. Smith, Philip A. Fay, Alan K. Knapp, M.
-Rebecca Shaw, Michael E. Loik, Stanley D. Smith, et al. 2004.
-``Convergence across biomes to a common rain-use efficiency.''
-\emph{Nature} 429 (6992):651--54.
-\url{https://doi.org/10.1038/nature02561}.
-
-\leavevmode\hypertarget{ref-Kleinhesselink2017b}{}%
-Kleinhesselink, Andrew R. 2017. ``Direct and indirect effects of climate
-change on plant populations and communities in sagebrush steppe.''
-Dissertation, Utah State University.
-
-\leavevmode\hypertarget{ref-Knapp2017}{}%
-Knapp, Alan K, Meghan L Avolio, Claus Beier, Charles J W Carroll, Scott
-L Collins, Jeffrey S Dukes, Lauchlan H Fraser, et al. 2017. ``Pushing
-precipitation to the extremes in distributed experiments:
-recommendations for simulating wet and dry years.'' \emph{Global Change
-Biology} 23 (5):1774--82. \url{https://doi.org/10.1111/gcb.13504}.
-
-\leavevmode\hypertarget{ref-Knapp2012}{}%
-Knapp, Alan K, John M Briggs, and Melinda D Smith. 2012. ``Community
-stability does not preclude ecosystem sensitivity to chronic resource
-alteration.'' \emph{Functional Ecology} 26 (6):1231--3.
-\url{https://doi.org/10.1111/j.1365-2435.2012.02053.x}.
-
-\leavevmode\hypertarget{ref-Knapp2015}{}%
-Knapp, Alan K., Charles J.W. Carroll, Elsie M. Denton, Kimberly J. La
-Pierre, Scott L. Collins, and Melinda D. Smith. 2015. ``Differential
-sensitivity to regional-scale drought in six central US grasslands.''
-\emph{Oecologia} 177 (4):949--57.
-\url{https://doi.org/10.1007/s00442-015-3233-6}.
-
-\leavevmode\hypertarget{ref-Kulmatiski2017a}{}%
-Kulmatiski, Andrew, Peter B. Adler, John M. Stark, and Andrew T.
-Tredennick. 2017. ``Water and nitrogen uptake are better associated with
-resource availability than root biomass.'' \emph{Ecosphere} 8 (3).
-\url{https://doi.org/10.1002/ecs2.1738}.
-
-\leavevmode\hypertarget{ref-LaPierre2016}{}%
-La Pierre, Kimberly J., Dana M. Blumenthal, Cynthia S. Brown, Julia A.
-Klein, and Melinda D. Smith. 2016. ``Drivers of Variation in Aboveground
-Net Primary Productivity and Plant Community Composition Differ Across a
-Broad Precipitation Gradient.'' \emph{Ecosystems} 19 (3):521--33.
-\url{https://doi.org/10.1007/s10021-015-9949-7}.
-
-\leavevmode\hypertarget{ref-Lemoine2016}{}%
-Lemoine, Nathan P, Justin Sheffield, Jeffrey S Dukes, Alan K Knapp, and
-Melinda D Smith. 2016. ``Terrestrial Precipitation Analysis (TPA): A
-resource for characterizing long-term precipitation regimes and
-extremes.'' \emph{Methods in Ecology and Evolution} 7 (11):1396--1401.
-\url{https://doi.org/10.1111/2041-210X.12582}.
-
-\leavevmode\hypertarget{ref-Oksanen2016}{}%
-Oksanen, Jari. 2016. ``Vegan: ecological diversity.''
-\url{https://doi.org/10.1029/2006JF000545}.
-
-\leavevmode\hypertarget{ref-Peters2012}{}%
-Peters, Debra P C, Jin Yao, Osvaldo E. Sala, and John P. Anderson. 2012.
-``Directional climate change and potential reversal of desertification
-in arid and semiarid ecosystems.'' \emph{Global Change Biology} 18
-(1):151--63. \url{https://doi.org/10.1111/j.1365-2486.2011.02498.x}.
-
-\leavevmode\hypertarget{ref-R2016}{}%
-R Core Team. 2016. \emph{R: A Language and Environment for Statistical
-Computing}. Vienna, Austria.
-
-\leavevmode\hypertarget{ref-Sala1992a}{}%
-Sala, O. E., W. K. Lauenroth, and W. J. Parton. 1992. ``Long-term soil
-water dynamics in the shortgrass steppe.'' \emph{Ecology} 73
-(4):1175--81. \url{https://doi.org/10.2307/1940667}.
-
-\leavevmode\hypertarget{ref-Seger1987}{}%
-Seger, J. 1987. ``What is bet-hedging?'' In \emph{Oxford Surveys in
-Evolutionary Biology}, edited by P.H. Harvey and L. Partridge, 182--211.
-Oxford: Oxford University Press.
-\url{http://ci.nii.ac.jp/naid/10004816581/\%5Cnpapers2://publication/uuid/BE92F8E9-1B87-4E50-85C5-AA134A0EBAF0}.
-
-\leavevmode\hypertarget{ref-Smith2011}{}%
-Smith, M.D. 2011. ``An ecological perspective on extreme climatic
-events: A synthetic definition and framework to guide future research.''
-\emph{Journal of Ecology} 99 (3):656--63.
-\url{https://doi.org/10.1111/j.1365-2745.2011.01798.x}.
-
-\leavevmode\hypertarget{ref-Smith2009}{}%
-Smith, Melinda D., Alan K. Knapp, and Scott L. Collins. 2009. ``A
-framework for assessing ecosystem dynamics in response to chronic
-resource alterations induced by global change.'' \emph{Ecology} 90
-(12):3279--89. \url{https://doi.org/10.1890/08-1815.1}.
-
-\leavevmode\hypertarget{ref-rstan2016}{}%
-Stan Development Team. 2016a. ``Rstan: the R interface to Stan, Version
-2.14.1.'' \url{http://mc-stan.org/rstan.html}.
-
-\leavevmode\hypertarget{ref-stan2016}{}%
----------. 2016b. ``Stan: A C++ Library for Probability and Sampling,
-Version 2.14.1.''
-
-\leavevmode\hypertarget{ref-Wilcox2016}{}%
-Wilcox, Kevin R, John M Blair, Melinda D Smith, and Alan K Knapp. 2016.
-``Does ecosystem sensitivity to precipitation at the site-level conform
-to regional-scale predictions?'' \emph{Ecology} 97 (3):561--68.
-\url{https://doi.org/10.1890/15-1437}.
-
-\leavevmode\hypertarget{ref-Wilcox2017}{}%
-Wilcox, Kevin R., Zheng Shi, Laureano A. Gherardi, Nathan P. Lemoine,
-Sally E. Koerner, David L. Hoover, Edward Bork, et al. 2017.
-``Asymmetric responses of primary productivity to precipitation
-extremes: A synthesis of grassland precipitation manipulation
-experiments.'' \emph{Global Change Biology} 23 (10):4376--85.
-\url{https://doi.org/10.1111/gcb.13706}.
+\bibliography{/Users/atredenn/Dropbox/Bibliography/usses_water.bib}
 

### 2018-01-24 code/02_plot_data.R
updates precip axis title

@@ -65,7 +65,7 @@ ppt_histogram <- ggplot(weather, aes(x=annual_precip))+
   scale_x_continuous(expand=c(0,0), limits=c(0,620), breaks=seq(0,600,100))+
   scale_y_continuous(expand=c(0,0), limits=c(0,0.0065))+
   ylab("Density")+
-  xlab(expression(paste("Annual Precipitation (mm ", yr^-1,")")))+
+  xlab(expression(paste("Water-year Precipitation (mm ", yr^-1,")")))+
   theme_bw()+
   theme(panel.grid.minor = element_blank())
 

### 2018-01-24 manuscript/usses_water_ms.tex
adds PDF version of figs for resubmission

@@
 m\(^2\) quadrats to track vegetation change over time. In 2007, we
-relocated 14 of the original quadrats in permenanent livestock
-exclosures, \new{which we used} as control plots (i.e.~ambient
-precipitation) in the experiment described below.
-\new{We used the relocated plots as our controls because collecting demographic data is time consuming and it was already being collected in these plots for other studies.}
+located 14 of the original quadrats in permenanent livestock exclosures,
+\new{which we used} as control plots (i.e.~ambient precipitation) in the
+experiment described below.
+\new{We used the original plots as our controls because collecting demographic data is time consuming and it was already being collected in these plots for other studies.}
 
@@
 
-\new{Following the above procedure, we still lacked soil moisture data for March 2012 (observations did not start unitl April) and for a string of days in 2013 during which the soil moisture probes failed to take readings.
+\new{Following the above procedure, we still lacked soil moisture data for March 2012 (observations did not start until April 2012) and for a string of days in 2013 during which the soil moisture probes failed to take readings.
 To fill in these gaps, we used a version of the SOILWAT soil moisture model}
@@
   \hline
-2011 & adonis &  22 &   2 & 1.04 & 0.43 \\ 
+2011 & adonis &  22 &   2 & 1.04 & 0.41 \\ 
   2011 & betadisper &  22 &   2 & 2.22 & 0.14 \\ 
@@
   2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
-  2013 & adonis &  22 &   2 & 1.26 & 0.12 \\ 
+  2013 & adonis &  22 &   2 & 1.26 & 0.11 \\ 
   2013 & betadisper &  22 &   2 & 0.43 & 0.66 \\ 
-  2014 & adonis &  22 &   2 & 0.96 & 0.58 \\ 
+  2014 & adonis &  22 &   2 & 0.96 & 0.55 \\ 
   2014 & betadisper &  22 &   2 & 0.36 & 0.70 \\ 
-  2015 & adonis &  22 &   2 & 1.08 & 0.33 \\ 
+  2015 & adonis &  22 &   2 & 1.08 & 0.32 \\ 
   2015 & betadisper &  22 &   2 & 1.96 & 0.17 \\ 

### 2018-01-24 code/02_plot_data.R

@@ -162,6 +162,7 @@ suppressWarnings(# ignore warnings about missing values, we know they are empty
                              ncol=1, labels = c("A)","B)","C)"), hjust = -0.4)
 )
 ggsave("../figures/data_panels.png", gout, width=3.3, height=8, units="in", dpi=120)
+ggsave("../figures/data_panels.pdf", gout, width=3.3, height=8, units="in")
 
 
 

### 2018-01-24 code/05_plot_model_results.R

@@ -228,6 +228,7 @@ suppressWarnings( # ignore warnings about NA values
                         rel_heights = c(1, 0.1, 1))
 )
 ggsave("../figures/glmm_main_results.png", plot = gridplot, width = 7, height = 5, units = "in", dpi =120)
+ggsave("../figures/glmm_main_results.pdf", plot = gridplot, width = 7, height = 5, units = "in")
 
 
 

### 2018-01-24 code/06_community_dynamics.R

@@ -178,5 +178,6 @@ ggplot(nmds_df, aes(x=MDS1, y=MDS2, fill=Treatment))+
   theme_few()+
   theme(panel.grid.major = element_line(color="grey90"))
 ggsave(paste0(figure_path,"sppcomp_bray_all.png"), width=6, height = 4, units = "in", dpi = 120)
+ggsave(paste0(figure_path,"sppcomp_bray_all.pdf"), width=6, height = 4, units = "in")
 
 

### 2018-01-24 manuscript/review_response_R1.Rmd
adds pdf figs and line numbers to response letter

@@
 Thank you for the opportunity to revise our manuscript.
+We apologize for the delay of our resubmission.
+The holidays intervened and we caught some errors we wanted to fully correct before resubmitting.
+
 All three reviewers agree our paper is interesting, but point out several areas where we could improve the analysis.
@@
 \response{We provide point-by-point responses to reviewer comments below.
-We have taken care to address Reviewer 1's concerns (also echoed by Reviewer 2) by including the suggested sensitivity analysis (see new section in Methods, lines xxx-xxx).
+We have taken care to address Reviewer 1's concerns (also echoed by Reviewer 2) by including the suggested sensitivity analysis (see new section in Methods, lines 261-273).
 Unfortunately, we could not implement the suggestion of Reviewer 1 to include a VWC$\times$Treatment$\times$Year interaction in our model because our data limits the complexity of our statistical model.
@@
 
-\response{In addition to changes in resposne to reviewer comments, we have added more detail in our description of how we estimated soil moisture based on direct observations.
+\response{In addition to changes in response to reviewer comments, we have added more detail in our description of how we estimated soil moisture based on direct observations (lines 149-186).
 While doing this, we discovered an error on our part: we use March-June volumetric water content as our covariate in each year, but we did not actually have soil moisture estimates for March 2012 (sensors were not installed until April 2012).
-Rather than dropping an important month in terms of water availability, we decided to use output from the SOILWAT soil moisture model as parameterized for our site (lines xxx-xxx in the revised manuscript, Figure 2B, and new Appendix 2).
+Rather than dropping an important month in terms of water availability, we decided to use output from the SOILWAT soil moisture model as parameterized for our site (lines 176-186 in the revised manuscript, Figure 2B, and new Appendix 2).
 Importantly, our main results are unchanged: ecosystem functional responses is similar among treatments, though there is now weak evidence that the drought resonse is more steep than the control response.}
@@
 \response{This is an excellent suggestion, and we have broadened the scope of our Discussion to place our findings in the context suggested by the reviewer.
-In particular, we discuss why our site might be different from the large scale patterns seen in other areas (lines xxx-xxx).}
+In particular, we discuss why our site might be different from the large scale patterns seen in other areas (lines 349-351 and lines 392-394).}
 
@@
 But, in the end, we decided this would muddle the analysis and confuse the reader, which is why we kept it separate.
-In our revision, we have included a new appendix (Appendix X) where we conduct our regression with NDVI rather than ANPP.
-We achieved the same results as when analyzing ANPP, and keep ANPP in the main text since that variable connects more easily to previous research (and what readers might be expecting).}
+In our revision, we have included a new appendix (Appendix 5) where we conduct our regression with NDVI rather than ANPP.
+We achieved the same results as when analyzing ANPP (lines 203-204), and keep ANPP in the main text because that variable connects more easily to previous research (and what readers might be expecting).}
 
 \response{A beneficial consequence of this comment from the reviewer is that we thoroughly checked our code for converting NDVI to ANPP and found an error for one year.
-Now our lowest $R^2$ is 0.39, rather than 0.21 (see new Table A1-2).
+Now our lowest $R^2$ is 0.39, rather than 0.21 (see new Table A3-2).
 This is still low, which is why we conducted the comparison analysis described above.}
@@
 
-\response{Sorry this was not clear. We included the prior specification [$\beta \sim \text{Normal}(0,5)$] in Appendix X on line 59. We have now included this in the main text on line xx.}
+\response{Sorry this was not clear. We included the prior specification [$\beta \sim \text{Normal}(0,5)$] in Appendix 4 on line 58. We have now included this in the main text on lines 251-252.}
 
@@
 \response{That said, the reviewer made an excellent suggestion of conducting a sensitivity analysis \emph{a la} Wilcox et al.
-2017, which we have included (see lines xxx - xxx in the Methods section; new results on lines xxx - xxx; and new discussion items on lines xxx-xxx).
+2017, which we have included (see lines 261-273 in the Methods section; new results on lines 316-317; and new discussion items on lines 395-401).
 The sensitivity analysis allows us to examine how the effect of treatment changes over time, adding another dimension to our analysis that better connects our findings to the Hierachical Response Framework.
@@
 This increase was not large enough to significantly impact ecosystem functional response, and our community composition analysis does not indicate changes in the final years of our experiment.
-However, as we note in the final line of our paper (lines xxx-xxx), the increased sensitivity in the latter years of the experiment may portend a coming shift in ecosystem functional response.}
+However, as we note in the final line of our discussion (lines 399-401), the increased sensitivity in the latter years of the experiment may portend a coming shift in ecosystem functional response.}
 
@@
 
-\response{Thanks for the suggestion. We have added a new introductory paragraph (lines xx-xx).}
+\response{Thanks for the suggestion. We have added new introductory sentences (lines 44-46).}
 
@@
 \response{This comment echoes one of reviewer 1's concerns.
-As we mention above, we have conducted a model comparison where we show that results are consistent whether we use ANPP or NDVI (see new Appendix X).
+As we mention above, we have conducted a model comparison where we show that results are consistent whether we use ANPP or NDVI (see new Appendix 5).
 In the main text we retain the analysis focused on ANPP because we think it connects better to previous research.
-Also, as mentioned above, there was an error in our regression for the year 2015, which has been fixed (see new Table A1-2).}
+Also, as mentioned above, there was an error in our regression for the year 2015, which has been fixed (see new Table A3-2).}
 
@@
 
-\response{This is an interesting point, and one we failed to discuss. We have added a sentence about this in our Conclusion section (lines xxx-xxx).}
+\response{This is an interesting point, and one we failed to discuss. We have added sentences about this in our Discussion (lines 399-401) Conclusion section (lines 409-411).}
 
@@
 \response{Yes, we agree that the rate of community turnover is likely related to the lifespans of constituent species.
-We have added text discussing the impact of lifespan on community turnover in our Conclusion (lines xxx-xxx).}
+We have added text discussing the impact of lifespan on community turnover in our Conclusion (lines 409-411).}
 
@@
 But upon further investigation, we found that these statistical results were detecting changes in rare species because the dominant annuals showed similar dynamics through time, regardless of treatment (Fig.
-Ax-x).
+A1-4).
 Likewise, the stable perennial community contributes most to ANPP is this systems.
-We briefly mention this analysis in our Results section on lines xxx-xxx, and show the full analysis in Appendix X.}
+We briefly mention this analysis in our Results section on lines 322-326, and show the full analysis in Appendix 1.}
 
@@
 \response{These are all good suggestions.
-Reviewer 2 agreed that the paper would benefit from a broader introduction, which we have added (lines xx-xx).
-We have also added a reference to the idea that domoninant species play a large role in determining ecosystem functioning (lines xx-xx).}
+Reviewer 2 agreed that the paper would benefit from a broader introduction, which we have added (lines 44-46).
+We have also added a reference to the idea that dominant species play a large role in determining ecosystem functioning (line 376).}
 
@@
 \response{We have increased the contrast between the colors so that it should be easier to distinguish colors if printed in black and white.
-The colors aren't perfect, but we anticipate most people will read this paper online.}
+The colors aren't perfect, but they are now colorblind friendly and we anticipate most people will read this paper online.}
 
@@
 
-\response{We have revised our description of modeling/extrapolating the discrete soil moisture measurements (lines xxx-xxx). Hopefully it is now easier to understand.}
+\response{We have revised our description of modeling/extrapolating the discrete soil moisture measurements (lines 149-175). Hopefully it is now easier to understand.}
 
@@
 We also use the demographic data to determine species composition.
-Given that this demographic data is time-consuming to collect, we decided to use the historical plots as controls rather than adding even more plots.}
+Given that this demographic data is time-consuming to collect, we decided to use the historical plots as controls rather than adding even more plots.
+We have added our reasoning on lines 119-120}
 
@@
 
-\response{We have revised this sentence to read, "However, all estimates for the effect of soil moisture on ANPP broadly overlapped zero, indicating the relationship is weak and uncertain regardless of treatment." (lines xx-xx).}
+\response{We have revised this sentence to read, "However, all estimates for the effect of soil moisture on ANPP broadly overlapped zero, indicating the relationship is weak and uncertain regardless of treatment." (lines 31-32).}
 
@@
 
-\response{Fair point, and one that echoes Reviewer 2's critique of our introduction. We have added a new opening paragraph that we hope better sets the stage (lines xx-xx).}
+\response{Fair point, and one that echoes Reviewer 2's critique of our introduction. We have added new opening sentences that we hope better set the stage (lines 44-46).}
 
@@
 
-\response{Good point. We have added a sentence describing why we use soil moisture on lines xx-xx.}
+\response{Good point. We have added a sentence describing why we use soil moisture on lines 76-77.}
 
@@
 
-\response{Thanks for the suggestion. We have alluded to this in our new opening paragraph (lines xx-xx)}
+\response{Thanks for the suggestion. We have alluded to this in our new opening paragraph (lines 44-46)}
 
@@
 We also use the demographic data to determine species composition.
-Given that this demographic data is time-consuming to collect, we decided to use the historical plots as controls rather than adding even more plots." We have added this justification in our Methods section on lines xxx-xxx.}
+Given that this demographic data is time-consuming to collect, we decided to use the historical plots as controls rather than adding even more plots." We have added this justification in our Methods section on lines 119-120.}
 
@@
 But, unfortunately, it was not feasible for us to control snow/winter precipitation.
-As the reviewer notes, we do consider this a limitation of our study (lines xxx-xxx).}
+As the reviewer notes, we do consider this a limitation of our study (lines 372-373).}
 
@@
 
-\response{Done.}
+\response{Done (line 141).}
 
@@
 \response{Upon re-reading, we completely agree!
-We have substantially rewritten this section for clarity and have added some mathematical notation to make our description precise (lines xxx-xxx).
+We have substantially rewritten this section for clarity and have added some mathematical notation to make our description precise (lines 149-175).
 Hopefully it is less confusing now.}
@@
 
-\response{We have revised this section to make this clearer (lines xxx-xxx). To answer the reviewer's question, }
+\response{We have revised this section to make this clearer (lines 149-175).}
 
@@
 
-\response{This is a good catch. We actually only use \emph{cumulative} volumetric water content, and we have made this consistent throughout the manuscript.}
+\response{This is a good catch.
+We actually only use \emph{cumulative} volumetric water content, and we have made this consistent throughout the manuscript.
+We also updated the Figure 2 caption to make this more clear.}
 
@@
 
-\response{Good idea. We have added the effect sizes (lines xxx-xxx).}
+\response{Good idea. We have added the effect sizes (lines 298-302).}
 
@@
 
-\response{We have added a new appendix (Appendix X) that provides more details on the dominance and rarity of the communities across and through time.
+\response{We have added a new appendix (Appendix 1) that provides more details on the dominance and rarity of the communities across and through time.
 In brief, rank abundance curves are very similar for all treatments and through time, with the same two perennial species (\emph{Artemisia tripartita} and \emph{Balsamorhiza sagittata}) dominant in all treatments, acounting for about 25\% of total cover.
 Similarly, all treatments had the same dominant annual species, whose dynamics did not vary by treatment.
-These ancillary results add weight to our conclusion that the ecosystem responses is driven by dominant species, which are similar among treatments.}
+These ancillary results add weight to our conclusion that the ecosystem responses is driven by dominant species, which are similar among treatments.
+We mention these results in the manuscript on lines 322-326.}
 
@@
 
-\response{Thanks for bringing this paper to our attention. We have included a new sentence that references this citation on line xxx in the Discussion.}
+\response{Thanks for bringing this paper to our attention. We have included a new sentence that references this citation on lines 367-368 in the Discussion.}
 

### 2018-01-24 manuscript/usses_water_ms.Rmd
adds pdf figs and line numbers to response letter

@@
 We define 'ecosystem functional response' as the relationship between available soil moisture and ANPP.
+\new{We focus on soil moisture rather than precipitation because soil moisture is more directly related to plant resource requirements.}
 The four scenarios are based on possible outcomes at the community (e.g., community composition) and ecosystem (e.g., soil moisture-ANPP regression) levels.
@@
 
-Third, the life history traits of the dominant species in our study ecosystem may explain the consistently positive, but weak and uncertain, effect of soil moisture on ANPP (Table 1, Fig. 3).
+Third, the life history traits of the dominant species, \new{which largely determine ecosystem function} [@Smith2003], in our study ecosystem may explain the weak and uncertain effect of soil moisture on ANPP (Table 1, Fig.
+3).
 Species that live in variable environments, such as cold deserts, must have strategies to ensure long-term success as conditions vary.

### 2018-01-24 manuscript/usses_water_ms.tex
adds pdf figs and line numbers to response letter

@@
 define `ecosystem functional response' as the relationship between
-available soil moisture and ANPP. The four scenarios are based on
-possible outcomes at the community (e.g., community composition) and
-ecosystem (e.g., soil moisture-ANPP regression) levels.
+available soil moisture and ANPP.
+\new{We focus on soil moisture rather than precipitation because soil moisture is more directly related to plant resource requirements.}
+The four scenarios are based on possible outcomes at the community
+(e.g., community composition) and ecosystem (e.g., soil moisture-ANPP
+regression) levels.
 
@@
 
-Third, the life history traits of the dominant species in our study
-ecosystem may explain the consistently positive, but weak and uncertain,
-effect of soil moisture on ANPP (Table 1, Fig. 3). Species that live in
-variable environments, such as cold deserts, must have strategies to
-ensure long-term success as conditions vary. One strategy is bet
-hedging, where species forego short-term gains to reduce the variance of
-long-term success \citep{Seger1987}. In other words, species follow the
-same conservative strategy every year, designed to minimize losses
-during unfavorable periods. The dry and variable environment of the
-sagebrush steppe has likely selected for bet hedging species that can
-maintain function at low water availability and have weak responses to
-high water availability. In so doing, the dominant species in our
-ecosystem avoid ``boom and bust'' cycles, which may explain the weak
-effect of soil moisture on ANPP (i.e., the Bayesian credible intervals
-for the slopes overlapping zero; Table 1).
+Third, the life history traits of the dominant species,
+\new{which largely determine ecosystem function} \citep{Smith2003}, in
+our study ecosystem may explain the weak and uncertain effect of soil
+moisture on ANPP (Table 1, Fig. 3). Species that live in variable
+environments, such as cold deserts, must have strategies to ensure
+long-term success as conditions vary. One strategy is bet hedging, where
+species forego short-term gains to reduce the variance of long-term
+success \citep{Seger1987}. In other words, species follow the same
+conservative strategy every year, designed to minimize losses during
+unfavorable periods. The dry and variable environment of the sagebrush
+steppe has likely selected for bet hedging species that can maintain
+function at low water availability and have weak responses to high water
+availability. In so doing, the dominant species in our ecosystem avoid
+``boom and bust'' cycles, which may explain the weak effect of soil
+moisture on ANPP (i.e., the Bayesian credible intervals for the slopes
+overlapping zero; Table 1).
 

### 2018-01-26 manuscript/usses_water_ms.Rmd
tweaks for PeerJ formatting

@@
 
-\definecolor{blue}{rgb}{0,0,0.7} 
+\definecolor{blue}{rgb}{0,0,0} 
 \newcommand{\new}{\textcolor{blue}}
@@
 
-First, altered precipitation might have no effect on either ecosystem functional response or community composition (Fig. 1, top left).
+First, altered precipitation might have no effect on either ecosystem functional response or community composition (Fig. 1A).
 In this case, changes in ANPP would be well predicted by the current, observed soil moisture-ANPP relationship.
 This corresponds to the early phases of the Hierarchical Response Framework, where ecosystem response follows the physiological responses of individual species.
-Second, the ecosystem functional response might change while community composition remains the same (Fig. 1, top right).
+Second, the ecosystem functional response might change while community composition remains the same (Fig. 1B).
 A saturating soil moisture-ANPP response fits this scenario, where individual species hit physiological thresholds or are limited by some other resource.
-Third, the ecosystem functional response might be constant but community composition changes (Fig. 1, bottom left).
+Third, the ecosystem functional response might be constant but community composition changes (Fig. 1C).
 In this case, changes in species' identities and/or abundances occur in response to altered precipitation levels and species more suited to the new conditions compensate for reduced function of initial residents.
-Fourth, and last, both ecosystem functional response and community composition could change (Fig. 1, bottom right).
+Fourth, and last, both ecosystem functional response and community composition could change (Fig. 1D).
 New species, or newly abundant species, with different physiological responses completely reshape the ecosystem functional response.
@@
 
-Cumulative March-June soil moisture had a weak positive effect on ANPP (Table 1; Fig. 3B).
+Cumulative March-June soil moisture had a weak positive effect on ANPP (Table 1; Fig. 3C).
 The effect of soil moisture for each treatment is associated with high uncertainty, however, with 95\% Bayesian credible intervals that \new{broadly} overlap zero (Table 1).
@@
 A3-2), which indicates the data did influence parameter estimates.
-Ecosystem functional response was similar among treatments (Table 1; Fig. 3B), but there is evidence that the \new{slope for the drought treatment is greater than the slope for the control treatment.}
+Ecosystem functional response was similar among treatments (Table 1; Fig. 3C), but there is evidence that the \new{slope for the drought treatment is greater than the slope for the control treatment.}
 This evidence comes from interpreting the posterior distributions of the slope offsets for the treatments.
 \new{From these distributions,} we calculate a \new{42}% one-tailed probability that the estimate is less than zero for the irrigation treatment \new{and a 100\% one-tailed probability that the estimate is greater than zero for the drought treatment} (Fig.
-3A, right panel).
+3B).
 
-\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3C). 
-Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0007; Fig. 3C)}.
+\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3D). 
+Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0007; Fig. 3D)}.
 
@@
 We found that ecosystem functional response under chronic \new{drought} was different from the control treatment, but community composition remained unchanged (Fig.
-3A, Fig.
-4), representing the top right scenario in Fig.
-1.
+3B, Fig.
+4), representing the scenario in Fig.
+1B.
 \new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots indicates increased sensitivity to water availability under chronic drought.
@@
 
-The similarity of ecosystem functional response (Fig.
-3) and community composition (Fig.
+The similarity of ecosystem functional response (Table 1; Fig.
+3C) and community composition (Fig.
 4) among treatments is surprising because grasslands generally, and sagebrush steppe specifically, are considered water-limited systems.
@@
 \new{The 1\% quantile of water-year precipitation at our site is 78 mm, a 26\% reduction from the mean, and the 99\% quantile is 545 mm, a 84\% increase from mean growing season precipitation (Appendix 6).
-Thus, our drought treatment represented extreme precipitation amounts, and that is the treatment where we observed a small effect on the slope between soil moisture and ANPP (Fig. 3A).
+Thus, our drought treatment represented extreme precipitation amounts, and that is the treatment where we observed a small effect on the slope between soil moisture and ANPP (Fig. 3B).
 The irrigation treatment may not have been extreme enough to induce a response.}
@@
 Failure to account for potentially important covariates could explain the within-year spread of ANPP (Fig 2C, Fig.
-3B) and the resulting uncertain relationship we observed between soil moisture and ANPP across all treatments (Table 1, Fig.
-A3-2).
+3C) and the resulting uncertain relationship we observed between soil moisture and ANPP across all treatments (Table 1).
 
@@
 
-\new{We found that ANPP became more sensitive to the drought treatment over time (Fig. 3C).
+\new{We found that ANPP became more sensitive to the drought treatment over time (Fig. 3D).
 We interepret this increase in sensitivity as the effect of cumulative impacts of the drought on dominant species.
@@
 
-# FUNDING
-
-This research was supported by the Utah Agricultural Experiment Station, Utah State University, and approved as journal paper number 9035.
-The research was also supported by the National Science Foundation, through a Postdoctoral Research Fellowship in Biology and Mathematics to ATT (DBI-1400370), a Graduate Research Fellowship to ARK, and grants DEB-1353078 and DEB-1054040 to PBA.
-
-# AUTHOR CONTRIBUTIONS
-\begin{itemize}
-  \item Andrew T. Tredennick collected data, analyzed the data, wrote the paper, prepared figures and/or tables, reviewed drafts of the paper.
-  \item Andrew R. Kleinhesselink conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed drafts of the paper.
-  \item J. Bret Taylor contributed reagents/materials/analysis tools, reviewed drafts of the paper.
-  \item Peter B. Adler conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed drafts of the paper.
-\end{itemize}
-
 # SUPPLEMENTAL INFORMATION
@@
 \begin{description}
-\item [Appendix 1.] \new{Additional information on plant community composition and dynamics, Table A1-1, Fig. A1-1, Fig. A1-2, Fig. A1-3, Fig. A1-4, and Fig. A1-5.}
-\item [Appendix 2.] \new{Additional details on soil moisture modeling with SOILWAT and Fig. A2-1.}
-\item [Appendix 3.] Additional methods and information on estimating aboveground net primary productivity, Table A3-1, and Table A3-2.
-\item [Appendix 4.] Details of the hierarchical Bayesian model, Fig. A4-1, Fig. A4-2, and Fig. A4-3.
-\item [Appendix 5.] \new{Results from analysis of NDVI without conversion to ANPP, Fig. A5-1.}
-\item [Appendix 6.] Details on analysis of precipitation historical range of variability and Fig. A6-1.
+\item [Appendix 1.] \new{Details on plant community structure.}
+\item [Appendix 2.] \new{Details on SOILWAT predictions.}
+\item [Appendix 3.] Estimating ANPP: additional methods and tables.
+\item [Appendix 4.] Additional details on the statistical model.
+\item [Appendix 5.] \new{Conducting our analysis with NDVI: additional results.}
+\item [Appendix 6.] Characterizing extreme precipitation amounts.
 \end{description}
@@
       \includegraphics[width=5in]{../figures/glmm_main_results.png}
-  \caption{Results from the generalized linear mixed effects model \new{(A-B) and the sensitivity analysis (C)}.
-(A) Posterior distributions for the effects of drought and irrigation on the intercept and slope of the productivity-soil moisture relationship.
+  \caption{Results from the generalized linear mixed effects model \new{(A-C) and the sensitivity analysis (D)}.
+(A-B) Posterior distributions for the effects of drought and irrigation on the intercept (A) and slope (B) of the productivity-soil moisture relationship.
 Treatment effects show the difference between the coefficients estimated in the treated plots and the control plots.
@@
 The posterior densities were smoothed for visual clarity by increasing kernel bandwidth by a factor of five.
-(B) Scatterplot of the data and model estimates shown as solid lines.
+(C) Scatterplot of the data and model estimates shown as solid lines.
 Model estimates come from treatment level coefficients (colored lines).
 Note that we show log(ANPP) on the \emph{y}-axis of panel B; this same plot can be seen on the arithmetic scale in supporting material Fig.
-A2-1.
-\new{(C) Regression of sensitivity against time for each treatment.
+A4-1.
+\new{(D) Regression of sensitivity against time for each treatment.
 Each point represents the sensitivity of ANPP in a plot relative to the mean of the control plots in that year.
@@
       \includegraphics[width=5in]{../figures/sppcomp_bray_all.png}
-  \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot, colored by treatment. \new{Each point represents a plot and the axes represent species composition.}}
+  \caption{Nonmetric multidimensional scaling scores representing plant communities in each plot for each year, colored by treatment.
+\new{Each point represents a plot and the axes represent species composition.}}
 \end{figure}

### 2018-01-26 manuscript/usses_water_ms.tex
tweaks for PeerJ formatting

@@
 
-\definecolor{blue}{rgb}{0,0,0.7} \newcommand{\new}{\textcolor{blue}}
+\definecolor{blue}{rgb}{0,0,0} \newcommand{\new}{\textcolor{blue}}
 
@@
 First, altered precipitation might have no effect on either ecosystem
-functional response or community composition (Fig. 1, top left). In this
-case, changes in ANPP would be well predicted by the current, observed
-soil moisture-ANPP relationship. This corresponds to the early phases of
-the Hierarchical Response Framework, where ecosystem response follows
-the physiological responses of individual species. Second, the ecosystem
+functional response or community composition (Fig. 1A). In this case,
+changes in ANPP would be well predicted by the current, observed soil
+moisture-ANPP relationship. This corresponds to the early phases of the
+Hierarchical Response Framework, where ecosystem response follows the
+physiological responses of individual species. Second, the ecosystem
 functional response might change while community composition remains the
-same (Fig. 1, top right). A saturating soil moisture-ANPP response fits
-this scenario, where individual species hit physiological thresholds or
-are limited by some other resource. Third, the ecosystem functional
-response might be constant but community composition changes (Fig. 1,
-bottom left). In this case, changes in species' identities and/or
-abundances occur in response to altered precipitation levels and species
-more suited to the new conditions compensate for reduced function of
-initial residents. Fourth, and last, both ecosystem functional response
-and community composition could change (Fig. 1, bottom right). New
-species, or newly abundant species, with different physiological
-responses completely reshape the ecosystem functional response.
+same (Fig. 1B). A saturating soil moisture-ANPP response fits this
+scenario, where individual species hit physiological thresholds or are
+limited by some other resource. Third, the ecosystem functional response
+might be constant but community composition changes (Fig. 1C). In this
+case, changes in species' identities and/or abundances occur in response
+to altered precipitation levels and species more suited to the new
+conditions compensate for reduced function of initial residents. Fourth,
+and last, both ecosystem functional response and community composition
+could change (Fig. 1D). New species, or newly abundant species, with
+different physiological responses completely reshape the ecosystem
+functional response.
 
@@
 Cumulative March-June soil moisture had a weak positive effect on ANPP
-(Table 1; Fig. 3B). The effect of soil moisture for each treatment is
+(Table 1; Fig. 3C). The effect of soil moisture for each treatment is
 associated with high uncertainty, however, with 95\% Bayesian credible
@@
 indicates the data did influence parameter estimates. Ecosystem
-functional response was similar among treatments (Table 1; Fig. 3B), but
+functional response was similar among treatments (Table 1; Fig. 3C), but
 there is evidence that the
@@
 \new{and a 100\% one-tailed probability that the estimate is greater than zero for the drought treatment}
-(Fig. 3A, right panel).
+(Fig. 3B).
 
-\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3C). 
-Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0007; Fig. 3C)}.
+\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3D). 
+Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0007; Fig. 3D)}.
 
@@
 was different from the control treatment, but community composition
-remained unchanged (Fig. 3A, Fig. 4), representing the top right
-scenario in Fig. 1.
+remained unchanged (Fig. 3B, Fig. 4), representing the scenario in Fig.
+1B.
 \new{The increase in the slope of the VWC $\times$ productivity relationship in the drought plots indicates increased sensitivity to water availability under chronic drought.
@@
 
-The similarity of ecosystem functional response (Fig. 3) and community
-composition (Fig. 4) among treatments is surprising because grasslands
-generally, and sagebrush steppe specifically, are considered
+The similarity of ecosystem functional response (Table 1; Fig. 3C) and
+community composition (Fig. 4) among treatments is surprising because
+grasslands generally, and sagebrush steppe specifically, are considered
 water-limited systems. \new{For example,} \citet{Huxman2004} and
@@
 Failure to account for potentially important covariates could explain
-the within-year spread of ANPP (Fig 2C, Fig. 3B) and the resulting
+the within-year spread of ANPP (Fig 2C, Fig. 3C) and the resulting
 uncertain relationship we observed between soil moisture and ANPP across
-all treatments (Table 1, Fig. A3-2).
+all treatments (Table 1).
 
@@
 
-\section{FUNDING}\label{funding}
-
-This research was supported by the Utah Agricultural Experiment Station,
-Utah State University, and approved as journal paper number 9035. The
-research was also supported by the National Science Foundation, through
-a Postdoctoral Research Fellowship in Biology and Mathematics to ATT
-(DBI-1400370), a Graduate Research Fellowship to ARK, and grants
-DEB-1353078 and DEB-1054040 to PBA.
-
-\section{AUTHOR CONTRIBUTIONS}\label{author-contributions}
-
-\begin{itemize}
-  \item Andrew T. Tredennick collected data, analyzed the data, wrote the paper, prepared figures and/or tables, reviewed drafts of the paper.
-  \item Andrew R. Kleinhesselink conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed drafts of the paper.
-  \item J. Bret Taylor contributed reagents/materials/analysis tools, reviewed drafts of the paper.
-  \item Peter B. Adler conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed drafts of the paper.
-\end{itemize}
-
 \section{SUPPLEMENTAL INFORMATION}\label{supplemental-information}
@@
   \hline
-2011 & adonis &  22 &   2 & 1.04 & 0.41 \\ 
+2011 & adonis &  22 &   2 & 1.04 & 0.43 \\ 
   2011 & betadisper &  22 &   2 & 2.22 & 0.14 \\ 
-  2012 & adonis &  22 &   2 & 1.09 & 0.32 \\ 
+  2012 & adonis &  22 &   2 & 1.09 & 0.33 \\ 
   2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
-  2013 & adonis &  22 &   2 & 1.26 & 0.11 \\ 
+  2013 & adonis &  22 &   2 & 1.26 & 0.13 \\ 
   2013 & betadisper &  22 &   2 & 0.43 & 0.66 \\ 
-  2014 & adonis &  22 &   2 & 0.96 & 0.55 \\ 
+  2014 & adonis &  22 &   2 & 0.96 & 0.57 \\ 
   2014 & betadisper &  22 &   2 & 0.36 & 0.70 \\ 
-  2015 & adonis &  22 &   2 & 1.08 & 0.32 \\ 
+  2015 & adonis &  22 &   2 & 1.08 & 0.34 \\ 
   2015 & betadisper &  22 &   2 & 1.96 & 0.17 \\ 
-  2016 & adonis &  22 &   2 & 1.12 & 0.26 \\ 
+  2016 & adonis &  22 &   2 & 1.12 & 0.28 \\ 
   2016 & betadisper &  22 &   2 & 0.27 & 0.76 \\ 

### 2018-01-26 code/05_plot_model_results.R

@@ -214,7 +214,7 @@ sens_plot <- ggplot(all_diffs, aes(x = year, y = sensitivity, fill = treatment))
 ##  Version with sensitivity analysis
 suppressWarnings( # ignore warnings about NA values
   bottom_row <- plot_grid(regress_plot, NULL, sens_plot, 
-                          labels = c('B','', 'C'),
+                          labels = c('C','', 'D'),
                           align = 'h', 
                           rel_widths = c(1, 0.1, 1), 
                           nrow = 1)
@@ -228,7 +228,7 @@ suppressWarnings( # ignore warnings about NA values
                         rel_heights = c(1, 0.1, 1))
 )
 ggsave("../figures/glmm_main_results.png", plot = gridplot, width = 7, height = 5, units = "in", dpi =120)
-ggsave("../figures/glmm_main_results.pdf", plot = gridplot, width = 7, height = 5, units = "in")
+ggsave("../figures/Figure3.pdf", plot = gridplot, width = 7, height = 5, units = "in")
 
 
 

### 2018-01-26 code/06_community_dynamics.R

@@ -166,6 +166,7 @@ for(doyr in unique(full_community_matrix_scaled$year)){
 saveRDS(object = out_stats, file = "../results/sppcomp_stats.RDS")
 
 mycols <- c("#009E73", "#D55E00", "#0072B2")
+nmds_df$Year <- paste0(rep(LETTERS[1:6],each=22), ") ", nmds_df$Year)
 ggplot(nmds_df, aes(x=MDS1, y=MDS2, fill=Treatment))+
   geom_point(shape = 21, color = "grey35")+
   scale_fill_manual(values=mycols)+
@@ -178,6 +179,6 @@ ggplot(nmds_df, aes(x=MDS1, y=MDS2, fill=Treatment))+
   theme_few()+
   theme(panel.grid.major = element_line(color="grey90"))
 ggsave(paste0(figure_path,"sppcomp_bray_all.png"), width=6, height = 4, units = "in", dpi = 120)
-ggsave(paste0(figure_path,"sppcomp_bray_all.pdf"), width=6, height = 4, units = "in")
+ggsave(paste0(figure_path,"Figure4.pdf"), width=6, height = 4, units = "in")
 
 
